### 2014-12-02 doc/GMP2015/paper.tex
Added documents.

%
% include the class file for GMP 2015
%
\documentclass[submission]{gmp2015}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%
% put the title of your paper here
%
\title{Optimal Decomposition-Recombination Plans for 2D Well-Constrained Graphs}

%
% put your submission number here
% this number will appear instead of authors and affiliations if the class option "submission" is active
%
\SubNumber{X}

%
% put the author names here
% use the second argument as a reference to the list of affiliations
% no authors and affiliations will appear if the class option "submission" is active
%
\author{Troy Baker}{1}
\author{Meera Sitharam}{1}
\author{Menghan Wang}{1}

%
% put the affiliations of the authors here
%
\affiliation{1}{Department of Computer and Information Science and Engineering, University of Florida, USA}

%
% the rest is as usual
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

\begin{abstract}
Super awesome abstract.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Nulla commodo posuere augue ut aliquet. Morbi at magna sem. Phasellus vitae varius massa. Nam eget quam sed erat euismod malesuada. Aliquam id velit metus, et porta lectus. Donec convallis massa quis eros accumsan consectetur.

Phasellus sit amet aliquam nulla. Vestibulum pretium lectus in ligula tempor convallis. Integer accumsan enim vel ante sollicitudin ultrices. Proin adipiscing, elit quis rhoncus egestas, lectus diam semper purus, sed luctus turpis felis eget magna. Proin eget dolor fringilla justo malesuada imperdiet sed sed erat. Sed sem dolor, egestas eleifend hendrerit dignissim, sagittis id quam. Aenean et nisl eu mi imperdiet mollis sit amet eu nunc. Nunc at turpis eget sapien eleifend lobortis at vitae nunc. Mauris a dui justo. Mauris suscipit quam id elit congue sed volutpat purus dapibus. Etiam eu tellus a enim lobortis posuere. Fusce feugiat sapien eu metus ullamcorper eleifend. Integer vehicula eros vel massa pharetra euismod. Vivamus a lacus ut sem condimentum imperdiet. Maecenas scelerisque vulputate enim, a cursus felis vehicula et. Morbi viverra, ipsum sed commodo lacinia, sapien elit tempus lectus, et consequat leo turpis quis odio.

Nulla eu lorem turpis. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Phasellus tellus elit, lobortis non euismod sed, ultrices vel magna. Vivamus condimentum erat ac erat mollis ut convallis nulla vehicula. Proin sit amet magna metus, quis imperdiet mi. Proin diam libero, laoreet ut auctor et, commodo quis nisi. Curabitur commodo nibh non sapien molestie congue. Etiam et mi non nibh mattis auctor ac in mauris. Donec ipsum urna, accumsan nec fermentum a, eleifend a elit. Aliquam velit justo, hendrerit sed luctus ac, luctus id urna. Aenean viverra blandit augue, a viverra arcu dictum nec. Nunc malesuada, mauris sit amet gravida euismod, magna ipsum porta mi, ut sodales neque tortor vel ligula. Aliquam a ipsum a mi molestie fermentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Fusce porttitor ligula eu mauris auctor sit amet porttitor odio ultrices.

Aliquam sit amet eros vel elit facilisis sodales. Sed pretium lacinia ante venenatis convallis. Etiam iaculis, risus in mattis viverra, sem orci sodales orci, tempus pretium enim est a augue. Suspendisse vitae est a orci lobortis dictum. Pellentesque sed magna ut tellus accumsan vestibulum. Curabitur ultricies mollis ante, interdum mollis metus ultricies a. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean vel magna ipsum.

\begin{figure}\centering
  \includegraphics[scale=.5]{example.eps}
  \caption{Example of a normal figure that spans one column.}
  \label{fig:square}
\end{figure}

Pellentesque commodo adipiscing imperdiet. Proin ultricies, quam vitae semper feugiat, eros diam pulvinar tortor, id venenatis ligula purus non erat. Mauris pulvinar velit sit amet mauris elementum bibendum. Sed mi leo, suscipit nec pharetra a, cursus id libero. Phasellus sit amet sapien diam, vitae scelerisque eros. Etiam vel risus lectus. Etiam molestie euismod turpis, condimentum ultrices dui euismod et.

Donec quis odio a dolor venenatis dictum nec eu dui. Fusce eleifend nibh et massa convallis posuere. Sed justo quam, vulputate sollicitudin pulvinar in, dapibus sit amet urna. Vivamus metus justo, molestie et volutpat a, sagittis ac arcu. Duis in ante ut quam condimentum cursus. Cras purus lacus, dictum vel auctor id, luctus in felis. Praesent hendrerit adipiscing lorem vitae fringilla. Aenean bibendum fermentum tortor, eget commodo ante auctor vel.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}\label{sec:relatedwork}

Vestibulum adipiscing congue orci eu commodo. Aliquam consectetur nibh nisl, ac mattis mauris. Sed vestibulum venenatis tincidunt. Phasellus quis metus id justo porta aliquet. Maecenas et accumsan nisi. Cras malesuada, urna vel venenatis adipiscing, lorem nunc pharetra dolor, ut convallis nisl sapien id eros. Cras commodo, elit a sollicitudin ultricies, erat mi rutrum nisi, in laoreet purus lorem laoreet nibh. Proin tincidunt, purus sit amet faucibus ornare, justo diam posuere dolor, id sollicitudin odio leo eget mauris. Nam mi ante, laoreet a adipiscing vitae, ultricies vitae dolor. Quisque auctor ornare iaculis. Praesent elit urna, porttitor vitae convallis vel, condimentum eu neque; see Figure~\ref{fig:square}.

Duis sollicitudin pulvinar ultrices~\cite{Peters:1997:BBC}. Praesent tincidunt interdum dignissim. Aenean vehicula nisl accumsan purus suscipit volutpat pharetra sapien ornare.
%
Example of an enumerated list:
\begin{enumerate}
  \item First item,
  \item Second item.
\end{enumerate}

Aenean id porttitor mi. Integer nec nisl elit, id dictum diam~\cite{Kazhdan:2008:SMF}. In consequat imperdiet lacus, quis placerat felis dapibus at. In at ligula ante, sollicitudin tristique nisi. Donec commodo justo eget urna luctus convallis. Praesent id diam eros, eu laoreet velit. Vivamus non tellus id nulla cursus sollicitudin.

\begin{figure*}\centering
\includegraphics{example.eps}
\caption{Example of a figure that spans both columns.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The new stuff}

Nam gravida adipiscing imperdiet. Fusce commodo pharetra faucibus. Donec feugiat, metus vel vulputate blandit, dolor nisi ornare libero, eu faucibus dui tortor sed turpis. Nullam porta, orci eu ullamcorper pretium, mauris elit convallis ante, sed gravida urna sapien eu quam. Morbi et quam purus. Nulla aliquam ultricies felis, a viverra eros scelerisque a. Fusce orci orci, dignissim a pellentesque sit amet, elementum eget sapien. Suspendisse potenti. In hac habitasse platea dictumst. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec et nisi dapibus metus mollis tincidunt. Aliquam rutrum sagittis mauris, non pulvinar ipsum sodales a.

\begin{theorem}\label{theorem:pi}
Aliquam nec sapien sit amet diam molestie tristique vitae sit amet libero,
\begin{equation}\label{eq:pi}
\pi = \sqrt{12}\sum^\infty_{k=0} \frac{(-3)^{-k}}{2k+1}.
\end{equation}
\end{theorem}

\begin{proof}
Curabitur metus lorem, rhoncus nec ullamcorper quis, varius eget neque. Phasellus et turpis quis massa porta tincidunt. Donec fringilla luctus libero, at adipiscing justo ullamcorper non. Mauris magna ipsum, semper eget hendrerit nec, ornare sit amet neque. Nullam elementum arcu eget tellus porta mattis.
\end{proof}

\noindent
Morbi ut lorem lectus. Etiam quis dolor justo, according to~\eqref{eq:pi}. Duis suscipit laoreet sollicitudin. Phasellus lacinia adipiscing tortor at posuere. Integer eget neque ac lacus vestibulum sollicitudin. Mauris vitae lacus dui. Aenean ultrices iaculis faucibus. Suspendisse congue, lectus in vulputate convallis, magna eros gravida mi, a molestie metus urna a nibh. Suspendisse adipiscing gravida ullamcorper.
%
Example of an itemized list:
\begin{itemize}
  \item First item,
  \item Second item.
\end{itemize}

\noindent
Nullam eu justo nibh. Vivamus id vehicula elit. Nam pellentesque rhoncus mi, id euismod arcu congue quis. Vestibulum sem urna, hendrerit non semper vel, facilisis vel sem. Ut sit amet massa nisi, vitae tristique metus. Duis venenatis massa quis sapien ullamcorper sagittis~\cite{Li:2010:PSO}. Ut auctor viverra fringilla. Donec at metus ligula. Nulla semper viverra diam, quis lobortis justo rhoncus ut. Aenean eget nisi sed dui iaculis iaculis quis ut libero. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse viverra rutrum ante pulvinar placerat. Nunc gravida laoreet tempus. Quisque non lobortis sem. In convallis fermentum odio, quis tincidunt dui lacinia vel.

\begin{corollary}
Etiam quis mauris orci. Nam blandit elementum diam a molestie. Suspendisse ultricies auctor urna, quis sagittis metus semper vel. Cras vel nisi felis.
\end{corollary}

\begin{proof}
Ut varius interdum pulvinar. Vivamus luctus consequat rutrum. Nulla facilisi. Mauris sit amet eros sit amet justo feugiat facilisis, which follows from Theorem~\ref{theorem:pi}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

Quisque id risus mi. Nunc vel bibendum nulla. Curabitur nec quam est. Quisque commodo leo quis ligula imperdiet vitae gravida orci eleifend. Vestibulum sed mollis turpis. Donec feugiat facilisis lorem, eget hendrerit arcu dignissim sit amet. Suspendisse potenti. Vivamus et felis leo, as shown in Section~\ref{sec:relatedwork}.

Duis consequat neque id libero eleifend a cursus massa sodales. Cras condimentum nulla ac neque dictum varius. Quisque vitae arcu neque, in placerat risus. Sed pellentesque iaculis est, non tincidunt augue sollicitudin id. Aliquam consequat, massa vel faucibus cursus, ipsum quam ultricies lectus, et hendrerit lectus arcu sed enim. Aliquam vitae enim nulla, laoreet fringilla velit. Phasellus eget diam enim.

Fusce iaculis lacus sed eros cursus euismod. Proin nec elit lectus. Mauris tristique accumsan risus, nec adipiscing odio viverra nec. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed a vestibulum eros. Curabitur viverra pretium risus, lacinia feugiat tellus tincidunt ut. Ut sit amet justo nec ipsum ultrices pharetra. Curabitur consectetur nisi nec dui tincidunt varius. Nunc vestibulum sollicitudin sapien, eu aliquam orci porta id. Etiam porttitor felis quis ante sodales euismod. Nam tempor leo vel justo tincidunt vehicula.

Vivamus quis sagittis leo. Nam vel purus risus, ultrices lacinia odio. Sed sagittis consequat luctus. Integer nulla ante, tincidunt sed pellentesque a, blandit vitae velit. Nulla facilisi. Donec id lacus ut justo pharetra consectetur nec nec enim. Praesent ut nisl vel metus facilisis commodo. Curabitur tristique varius ultricies. Praesent vitae mauris odio, vel aliquet metus. Suspendisse tempus erat eget sem lobortis accumsan. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Ut facilisis quam eu tortor fringilla non accumsan mi fringilla. Donec quis leo nec odio elementum tristique ac at enim.

Etiam quis diam ut nibh viverra iaculis. Fusce vehicula, mauris eu convallis consequat, turpis nunc interdum urna, eu malesuada erat justo non velit. In vel lectus id nunc varius cursus. Nullam vitae odio eget urna laoreet aliquam vel semper nisi. Mauris venenatis arcu eget nisi vehicula ac pharetra enim vehicula. Nullam et ipsum tortor, eu volutpat diam. Morbi adipiscing commodo rhoncus. Morbi urna lorem, volutpat non aliquam a, fermentum non diam.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}

### 2014-12-02 doc/GMP2015/sample/sample.tex
Added documents.

--- 2014-12-02 doc/GMP2015/paper.tex
@@
 %
-\title{Optimal Decomposition-Recombination Plans for 2D Well-Constrained Graphs}
+\title{A nice paper that is being submitted to GMP 2015}
 
@@
 %
-\SubNumber{X}
+\SubNumber{142}
 
@@
 %
-\author{Troy Baker}{1}
-\author{Meera Sitharam}{1}
-\author{Menghan Wang}{1}
+\author{Alvin B.\ Ginner}{1}
+\author{Douglas Smith}{1}
+\author{Martha X.\ Pert}{2}
 
@@
 %
-\affiliation{1}{Department of Computer and Information Science and Engineering, University of Florida, USA}
+\affiliation{1}{Department of Computer Science, University of Smalltown, Germany}
+\affiliation{2}{Department of Mathematics, University of Excellence, USA}
 
@@
 \begin{abstract}
-Super awesome abstract.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Suspendisse eleifend diam augue.
+Donec adipiscing mollis nisl vitae porttitor.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Maecenas gravida tincidunt volutpat.
+Morbi sapien ligula, commodo quis dignissim non, aliquam nec urna.
+Vivamus quis massa sit amet mi convallis sagittis ac facilisis velit.
+Duis at eros purus, vel laoreet nibh.
+Suspendisse rutrum est vel massa imperdiet at ultricies massa facilisis.
+Sed vitae pulvinar magna.
 \end{abstract}
@@
 \begin{figure}\centering
-  \includegraphics[scale=.5]{example.eps}
+  \includegraphics[scale=.5]{example}
   \caption{Example of a normal figure that spans one column.}
@@
 \begin{figure*}\centering
-\includegraphics{example.eps}
+\includegraphics{example}
 \caption{Example of a figure that spans both columns.}
@@
 \bibliographystyle{abbrv}
-\bibliography{paper}
+\bibliography{sample}
 

### 2014-12-02 doc/General_Work/DRP_2d.tex
Added documents.

% --------------------------------------------------------------
%                       Initialize
% --------------------------------------------------------------

\documentclass[11pt]{article}

% \usepackage{multicol}
\usepackage{ulem}
\usepackage{comment}
\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}

\usepackage{personalNotes}
% contains special macros for math environment:
%     unit, p, set, abs, floor, ceil, vec, norm, inner, kernel
% contains useful all-purpose macros:
%     n, lessSeparation, append

\usepackage{tikz}
\usetikzlibrary{shapes,snakes}
\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}

% --------------------------------------------------------------
%                    Beginning of Document
% --------------------------------------------------------------

\begin{document}
% \begin{multicols}{2}

% --------------------------------------------------------------
%                           Title
% --------------------------------------------------------------

\pntitle {Decomposition Recombination Plans for 2D Well-Constrained Graphs}
\pnauthor{Troy Baker}
\pndate  {May 13, 2014}

\pnmaketitle
\pnmakefooter

% --------------------------------------------------------------
%                           Body
% --------------------------------------------------------------

\section{Introduction}
\todo{
Other sections...
\begin{itemize}
    \item Intro and motivation (previous work, where this work leads, examples)
    \item Preliminaries
    \item DR-Plan (Canonical) + algo
    \item Recombination + algo (refer to other paper and make 2D)
    \item Solution space navigation
    \item Conclusion
\end{itemize}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminaries}

\subsection{Definitions}
% ISSUE: My definitions don't prevent non-positive densities... This means things aren't guaranteed to be connected...

\begin{definition}[Graph]
Graph $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$. Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
\end{definition}

\begin{definition}[Density]
The density of a graph is defined as:
\[d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}\]
% \[d(G) = \max\set{0,\sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}}\]
\end{definition}

\begin{definition}[Independent or Sparse Graph]
\todo{Needs definition.}
\end{definition}

\begin{definition}[Over-constrained Graph]
Graph $G=(V,E,w_V,w_E)$ is over-constrained if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$. (where S is not trivial... This seems weird because you could have an over-constrained under-constrained graph...)
\end{definition}

\begin{definition}[Well-constrained Graph]
Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
\end{definition}

% \begin{definition}[Well-constrained Graph]
% Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either under-constrained, well-constrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
% \end{definition}

\begin{definition}[Rigid Graph]
Graph $G=(V,E,w_V,w_E)$ is rigid if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
\end{definition}

\begin{definition}[Under-constrained Graph]
Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, $G$ is not rigid.
\end{definition}

% \begin{definition}[Under-constrained Graph]
% Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, there exists some subgraph $G'\subseteq G$ such that $d(G') < k$.
% \end{definition}

\begin{definition}[Trivial Graph]
A trivial graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.

\todo{Maybe leave this out until it's needed later?}
In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
\end{definition}

\begin{definition}[Decomposition Recombination (DR-) Plan]
The DR-plan of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
\begin{enumerate}
    \item The root of the tree `contains' $G$.
    \item For all nodes, $C$, that `contain' nontrivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
    \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
    \item A node with a trivial subgraph is a leaf.
\end{enumerate}
% It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.

An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.

Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}

\begin{definition}[Complete DR-Plan]
The complete DR-plan of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \textit{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$. By this, rule 3 is implicit for a complete DR-plan.
\end{definition}

\begin{definition}[Optimal DR-Plan]
The optimal DR-plan of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Notation}
$G$ is taken to be a graph with values $(V,E,w_V,w_E)$. To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages. It is not useful to solve for two bodies simultaneously if there are no constraints between them. Furthermore, to limit the scope of the paper, $G$ will always be taken to be well-constrained.
% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.

$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.

$C$ is an arbitrary graph in a node in $CompleteDRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$. By definition of a DR-plan, it is implied that $C_i$ is a well-constrained vertex-maximal proper subgraph.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Canonical Optimal DR-Plan}
\begin{figure}
\begin{center}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (1,1) {$R_j$};
    \node[v] (v2) at (3,1) {$R_k$};
    \node[v] (v3) at (2,2) {$D$};
    \node[v] (v4) at (2,3.3) {$R_i$};

    \tedge{v1}{v3}{solid}{}{above};
    \tedge{v2}{v3}{solid}{}{above};
    \tedge{v3}{v4}{solid}{}{right};
\end{tikzpicture}
\ldots with weights \ldots
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (1,1) {y};
    \node[v] (v2) at (3,1) {z};
    \node[v] (v3) at (2,2) {k};
    \node[v] (v4) at (2,3.3) {x};

    \tedge{v1}{v3}{solid}{y}{above};
    \tedge{v2}{v3}{solid}{z}{above};
    \tedge{v3}{v4}{solid}{x}{right};

    % \draw (3.3,2.5) circle (1.9cm);
    % \draw (.7,2.5) circle (1.9cm);
    % \draw (2,.5) circle (1.9cm);
\end{tikzpicture}
\\
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxy};
    \node[v] (v3) at ( 1,-1) {kyz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v5}{solid}{}{};
    \tedge{v3}{v5}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxy};
    \node[v] (v3) at ( 1,-1) {kxz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v5}{solid}{}{};
    \tedge{v3}{v4}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxz};
    \node[v] (v3) at ( 1,-1) {kyz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v6}{solid}{}{};
    \tedge{v3}{v5}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
% \\
% \begin{tikzpicture}
%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
%     \tikzstyle{e}=[]

%     \node[v] (v1) at ( 0, 0) {kxyz};
%     \node[v] (v2) at (-1,-1) {kxy};
%     \node[v] (v3) at (-2,-2) {kx};
%     \node[v] (v4) at (-3,-3) {k};
%     \node[v] (v5) at ( 1,-1) {z plan};
%     \node[v] (v6) at ( 0,-2) {y plan};
%     \node[v] (v7) at (-1,-3) {x plan};
%     \node[v] (v8) at (-2,-4) {k plan};

%     \tedge{v1}{v2}{solid}{}{};
%     \tedge{v1}{v5}{solid}{}{};
%     \tedge{v2}{v3}{solid}{}{};
%     \tedge{v2}{v6}{solid}{}{};
%     \tedge{v3}{v4}{solid}{}{};
%     \tedge{v3}{v7}{solid}{}{};
%     \tedge{v4}{v8}{solid}{}{};
% \end{tikzpicture}
\end{center}

\caption{A well-constrained graph whose nodes are subgraphs labeled with their density. Below are its 3 optimal DR-Plans (partially complete). The center vertex has weight $k$ and is well-constrained itself. The radial nodes have weights $x,y,z \neq k$ with equivalent edge weights, making any union between $k$ and the others well-constrained. \todo{Make this figure more clear.}}
\label{3-plans}
\end{figure}

In the case of a well-constrained graph, it is clear that there always exists a unique vertex-maximum DR-Plan that includes every single well-constrained vertex-maximal proper subgraph at each node, which we call the complete DR-plan. However, in the problem of the optimal DR-Plan there is no such unique plan. For instance, a graph that is formed by the union of 3 well-constrained subgraphs that all intersect on a well-constrained subgraph, demonstrated in Figure \ref{3-plans}. In this case, it results in $3$ unique optimal DR-Plans. Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same. Therefore, all choices of decomposition are in some sense equivalent. The theorem we seek to prove is thus:

\begin{theorem}\label{t1}
Given a well-constrained 2-dimensional bar-joint graph $G$, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
\begin{enumerate}
    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
    \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
\end{enumerate}
\end{theorem}

We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph. By definition, any subgraph of a well-constrained graph will either be (1) trivial, (2) under-constrained, or (3) well-constrained. We will analyze all 3 possibilities in turn. These observations will lead to interesting properties that govern the behavior of optimal DR-plans on well-constrained graphs.

% We show this by carefully considering all of the potential cases of the unions of subgraphs. By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained, or (3) trivial. We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
We wish to prove the following bijections:

\begin{center}
\begin{tabular}{|c|c|}
\hline
\textbf{Union is\ldots} & \textbf{Intersection is\ldots} \\ \hline
\sout{trivial}          & --- \\ \hline
under-constrained       & trivial \\ \hline
well-constrained        & well-constrained \\ \hline
---                     & \sout{under-constrained} \\ \hline
\end{tabular}
\end{center}

\begin{lemma}\label{t-l1}
$F_i\cup F_j$ cannot be trivial.
\end{lemma}

\begin{proof}
% If $F_i,F_j$ are well-constrained, then $F_i\cup F_j \geq k$. Trivial graphs must be over-constrained.

Follows from the definition of a trivial graph. It would imply that the well-constrained graphs were also trivial since they would be subgraphs of a trivial graph.
\end{proof}

\begin{lemma}\label{uc-l1}
$F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
\end{lemma}

\begin{proof}
It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$. In the forward direction, we have that $d(F_i\cup F_j)=l<k$. If it were equal to $k$ it would be well-constrained since, as a subgraph of $F$, it satisfies the other properties of a well-constrained graph; if it were greater than $k$ then it would be over-constrained. Therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=2k-l>k$. The only way to have an over-constrained subgraph in a well-constrained graph is if it is trivial.

In the reverse direction, we know that $d(F_i\cap F_j)>k$ because it is trivial. Thus, by the same math, it is clear that $d(F_i\cup F_j)<k$ and it must be under-constrained.
\end{proof}

\begin{lemma}\label{wc-l1}
$F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
\end{lemma}

\begin{proof}
It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$. In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$. Since $F_i\cap F_j$ is a subgraph of $F$, which is well-constrained, it is also well-constrained.

In the reverse direction the same math proves that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it too is well-constrained.

% \todo{Integrate this}
% Furthermore, all subgraphs of $F_i\cup F_j$ and $F_i\cap F_j$ are also subgraphs of $F_i$ and $F_j$, therefore all properties of well-constrained graphs are maintained.

% In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial. The intersection also cannot be under-constrained. If $F_i \cap F_j$ were under-constrained, then for $F_i$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection. The same is true for $F_j$. If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an over-constrained nontrivial graph. This is a contradiction, by definition of a well-constrained graph. Therefore, the intersection can only be well-constrained.

% reverse!!!!!
\end{proof}

\begin{lemma}\label{iuc-l1}
$F_i\cap F_j$ cannot be under-constrained.
\end{lemma}

\begin{proof}
Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained. So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and well-constrained intersections are possible.
\end{proof}

% WAIT... WHAT ABOUT TWO BANANAS? The intersection is 2 vertices which is under-constrained... oh nevermind... that's well-constrained

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Inferring Graph Structure from Union and Intersection Properties}

\begin{lemma}\label{wc-l2}
% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$,
$C_i\cup C_j$ is well-constrained if and only if $C_i\cup C_j = C$.
\end{lemma}

\begin{proof}
The forwards direction can be shown with a contradiction. If the $C_i\cup C_j$ was well-constrained but $C_i\cup C_j \neq C$, we would have found a larger well-constrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger. The two original subgraphs could not have been vertex-maximal proper.

The reverse direction follows from the fact that $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
\end{proof}

\begin{corollary}\label{wc-c1}
% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$, l
Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
\end{corollary}

\begin{proof}
It follows from the fact that $C_i\cup C_j$ must equal $C$ (proven in Lemma \ref{wc-l2}).
\end{proof}

At this point we limit our discussion to the scope of 2D graphs. All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$. Trivial graphs are a single vertex and empty set.

Also, 2D well-constrained graphs must be connected.
% \todo{Is this actually true?}
The greatest density of a 2D well-constrained graph is $-2$ (the vertex). The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).

\begin{lemma}\label{wc-l3}
If $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
Alternatively, if $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
\end{lemma}

\begin{proof}
Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$). Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.

Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

\begin{itemize}
    \item 3 cases: $C_k$ cannot be $C=Idc(C,R_i\cup R_j\cup D_{i,j})$, $C_i=Idc(C,R_j\cup D_{i,j})$, or $C_j=Idc(C,R_i\cup D_{i,j})$. This is by definition.

    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal. These are the graphs $Idc(C,R'_i\cup D_{i,j})$, $Idc(C,R'_j\cup D_{i,j})$, $Idc(C, D_{i,j})$, $Idc(C,R_i\cup D'_{i,j})$, $Idc(C,R_j\cup D'_{i,j})$, $Idc(C,R'_i\cup D'_{i,j})$, $Idc(C,R'_j\cup D'_{i,j})$, $Idc(C, D'_{i,j})$, $Idc(C,R_i)$, $Idc(C,R_j)$, $Idc(C,R'_i)$, $Idc(C,R'_j)$, and $Idc(C,\emptyset)$.

    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal. These are the graphs $Idc(C,R'_i\cup R_j\cup D_{i,j})$ and $Idc(C,R_i\cup R'_j\cup D_{i,j})$ respectively.

    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary \ref{wc-c1}) and cannot be well-constrained.

    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained. We also know it cannot be trivial because it contains well-constrained subgraphs. This means it must be under-constrained. From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial. This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.

    \item 1 case: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The union with $C_i$ (or $C_j$) is not equal to $C$. Therefore, the union must be under-constrained which means the intersection must be trivial (a single node). However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.

    \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible. The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.

    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that is left!
\end{itemize}

Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.

\todo{Needs a figure.}
\end{proof}

% \pnplaceholder
% WAIT, TRIVIAL GRAPH DEFINITION DOESN'T WORK!!!! in 4D it's the 3 clique, but 3 vertices is a subgraph of that... that is certainly not over-constrained...

% NOT SO! -5C2 = -10. 3 edges and 3 nodes -> 3 - 3*4 = -9... over-constrained!

% \pnplaceholder

\begin{lemma}\label{uc-l2}
% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial (a vertex or empty set).
If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{lemma}

\begin{proof}
If any intersection was not trivial, then it must be well-constrained. If $C_k\cap C_l$ were well-constrained, then by Lemma \ref{wc-l3} all of the intersections are well-constrained. This contradicts the statement that $C_i\cap C_j$ is trivial.
% It follows that if the intersection is trivial then the union is under-constrained (Lemma \ref{uc-l1}). If the union is under-constrained, then the
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Proving the Main Theorem}
Lemmas \ref{wc-l3} and \ref{uc-l2} show that $CompleteDRP(C)$ will result in one of two scenarios. (A) Either all $C_1,\ldots,C_N$ intersect on some common well-constrained subgraph and any two union to make $C$, or (B) all $C_1,\ldots,C_N$ intersect trivially and all of them must be unioned to make $C$.

% \begin{theorem}
% In case (A), any two unique wcvmps of $C$ can be used as the children of $C$ in $OptimalDRP(C)$.
% \end{theorem}

% \begin{proof}

% \end{proof}

% \begin{theorem}
% In case (B), all wcvmps of $C$ must be used as the children of $C$ in $OptimalDRP(C)$.
% \end{theorem}

% \begin{proof}
% If not all of them are used, the union will be under-constrained. $C$ is well-constrained, thus using not all of them is not an option.
% \end{proof}

% \pnplaceholder

% \pagebreak
% See ISSUE at top of paper, we have to assume we have a connected $G$...

% Take $C'_i$ to be $C\setminus C_i$. Then if $C_i\cup C_j$ is well-constrained we have directly from Lemma \ref{wc-l2} for all $C'_i, C'_j$ there can be no edges between them (otherwise $C_i\cup C_j\neq C$).

% The set of edges connecting $C'_i$ to $C_i$ must have a weight that sums to exactly $d(C'_i)$ since we know that $d(C)=d(C_i)=k$.

% Any other $C_k$ must therefore contain $C'_i$ and $C'_j$ because it adds no density to ``tack it on'' to a graph that does not contain them already.

% Now we want to know something about $C_k$. If $C_k \neq C_i$, then $C'_i\not\subseteq C'_k$. It is clear that they cannot be equal otherwise $C_k = C_i$. However, it also cannot be a subset. There are two scenarios. (1) $C'_k$ does not contain vertices that are incident to $C'_i$. If this were so, we know we could ``tack on'' $C'_i$ to the well-constrained $C_k$ without increasing the density and therefore $C_k$ could not have been maximal in the first place. (2) $C'_k$ contains vertices that are incident to $C'_i$. In this case

% Take $D_{i,j}$ to be $C\setminus (C'_i\cup C'_j)$. Clearly $D_{i,j}$ must be well-constrained since ``tacking on'' $C'_i$ and $C'_j$ adds no density and $C$ is well-constrained. Therefore, if there were some other $C_k$, it must contain $D_{i,j,k}\subset D_{i,j}$ such that $D_{i,j,k}=D_{i,j}\setminus C'_k$. It is clear that $D_{i,j,k}$ cannot equal $D_{i,j}$ or else $C_k$ must be $C_i$ or $C_j$ to be vertex-maximal.

% \begin{lemma}
% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial or empty set.
% \end{lemma}

% \begin{proof}
% By Lemmas... the union must be under
% Lemma \ref{uc-l1} that $C_i\cup C_j$ must be under-constrained. Therefore, by the proof
% It follows from Lemma \ref{wc-l1} that the only way to get
%  and its proof that if any intersections
% \end{proof}

\begin{figure}
\begin{center}
\begin{tikzpicture}
    \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
    \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
    \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);

    % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
    % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);

    % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
    % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
    \tikzstyle{e}=[]

    \node[el] (m) at (0,    0) {k};
    \node[v]  (l) at (-1.5, 0) {b};
    \node[v]  (r) at (1.5,  0) {c};

    \node[el] (tl) at (-1,  1) {i};
    \node[el] (tr) at (1,   1) {j};
    \node[v]  (t)  at (0,   2) {a};

    % \node[el] (bl) at (-1,  -1) {l};
    % \node[el] (br) at (1,   -1) {m};
    % \node[v]  (b)  at (0,   -2) {d};

    % \node[el] (v2) at (-1,-1) {b};
    % \node[el] (v3) at (-1,-1) {b};
    % \node[el] (v4) at (-1,-1) {b};
    % \node[el] (v5) at (-1,-1) {b};

    \tedge{m}{l}{solid}{}{};
    \tedge{m}{r}{solid}{}{};

    \tedge{l}{tl}{solid}{}{};
    \tedge{tl}{t}{solid}{}{};
    \tedge{r}{tr}{solid}{}{};
    \tedge{tr}{t}{solid}{}{};

    % \tedge{l}{bl}{solid}{}{};
    % \tedge{bl}{b}{solid}{}{};
    % \tedge{r}{br}{solid}{}{};
    % \tedge{br}{b}{solid}{}{};
\end{tikzpicture}
\end{center}

\caption{Each node contains a subgraph of the entire, well-constrained graph $G$. We take $Idc(G,a\cup i\cup b)$, $Idc(G,a\cup j\cup c)$, and $Idc(G,b\cup k\cup c)$ to be well-constrained. If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c$ must all be trivial subgraphs.
% Each node contains a subgraph of the entire, well-constrained graph. We take $a\cup i\cup b$, $a\cup j\cup c$, $b\cup k\cup c$, $b\cup l\cup d$, and $c\cup m\cup d$ to be well-constrained. If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c,d$ must all be trivial subgraphs.
% THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
}
\label{trivial-intersection}
\end{figure}

\begin{theorem}\label{uc-t1}
% In any solution to $OptimalDRP(G)$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is under-constrained, then all well-constrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.

In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial (case (B)), then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
\end{theorem}

\begin{proof}
From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be under-constrained. If it is under-constrained, it must be part of some larger well-constrained subgraph $D$. If $D$ is not $G$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
 % For $C$ to be well-constrained... WHAT????

If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.

This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid. Such an example is shown in Figure \ref{trivial-intersection}.
\end{proof}

% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$. All of these subgraphs form a complex structure that is well-constrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).

\begin{theorem}\label{wc-t1}
% In $OptimalDRP(G)$ of a well-constrained $G$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is well-constrained, then said node can have only $2$ children.

In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained (case (A)), then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
% Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
\end{theorem}

\begin{proof}
For a valid DR-plan, all children must be proper subgraphs. Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least 2 children. It follows directly from Lemma \ref{wc-l2} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then you found a tree with only 2 children.

Clearly, this is the optimal choice at this level. However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in later on.

Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. Take $R_k=C\setminus C_k$. This means that $D$ is also $\bigcup_{k=1}^N{R_k}$.

After we go down $N$ levels we get to a point where every node is some $Idc(C,R_k\cup D)$. Until this point, the fan-in of every node is exactly 2.

This is demonstrated with $N=3$ in Figure \ref{3-plans}. It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution. This follows from the discussion above. Now, at this level, the fan-in must be equal in all scenarios. In fact, at node $k$ it will be 1 (because of $D$ which is well-constrained) plus whatever fan-in $R_k$ incurs.
\end{proof}

With these two proofs, the original Theorem \ref{t1} has been proven. With this we have formulated the notion of a canonical optimal DR-plan. Any optimal DR-plan of a well-constrained graph must follow this pattern.

% (2) If the union results in a well-constrained subgraph, it must be the entire graph. If this were not the case, .

% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$. It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Further Results}

\begin{corollary}
Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is well-constrained, then $D$ will be a descendant of every $C_k$.
\end{corollary}

\begin{proof}
Explained in Theorem \ref{wc-t1}.
\end{proof}

\begin{corollary}
$\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
\end{corollary}

\begin{proof}
Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
\end{proof}

% \begin{proof}
% It follows from Theorem \ref{uc-t1} that if the intersection of any two well-constrained vertex-maximal proper subgraphs of $C$ are well-constrained then all of them are well-constrained or $\emptyset$.
% \end{proof}

% Figure idea: Large node in the middle with weight $k$. Several other graphs with weight $a, b, c, d\ldots$ attached radially by single edges with weights $a, b, c, d\ldots$ respectively. Circle each vertex-maximal subset.

% $\forall k\in [1,\ldots,N]$ the subgraphs $C\setminus C_k$ will be edge disjoint. Therefore the solving of one is exclusive of the solving of the other.

% Furthermore, the union of all of these subgraphs will be $C\setminus D$. So each subgraph will need to be ``tacked'' onto $D$ for recombination and since they are edge-disjoint the order is unimportant.

% \pnplaceholder
% \begin{lemma}
% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
% \end{lemma}

% \begin{proof}
% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}
% \pnplaceholder

% \begin{corollary}
% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
% \end{corollary}

% \begin{proof}
% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$. That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
% \end{proof}

\begin{corollary}
There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
\end{corollary}

\begin{proof}
Since $C_1\cup C_2$ is the entire graph
\todo{Prove this!!!!}
\end{proof}

% \begin{corollary}
% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
% \end{corollary}

% \begin{proof}
% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Future Work}
While this paper fleshes out the case of a well-constrained input graph, little work has been done on the over-constrained graph. This has to do with the fact that finding a DR-Plan is NP-complete. But is there some algorithm that will find an approximation within a reasonable factor?

Under-constrained is not worth studying in itself, because it will either be that all subgraphs are under-constrained or trivial or that some subgraphs are well-constrained or over-constrained. In this way, it is easily reduced to the other problems.
\\

The current method for finding $DRP(G)$ is the Frontier algorithm. Is there a way to adapt the Pebble Game algorithm? It is used to detect rigidity, but there is no obvious way to use it to find wcvmps (other than cutting out one node at a time).

Or maybe, instead of these top-down algorithms, there is some bottom-up algorithm?

Useful for glass networks.

\begin{comment}
\section{Formal Conjectures}

\subsection{Conjecture 1}

\begin{conjecture}
% Take $G$ to be well-constrained. Then, for all nodes $C$ in $DRP(G)$, either $\forall i,j (C_i \cap C_j \leq 1)$ or $\forall i,j (C_i \cap C_j \geq 2) \wedge (C_i \cup C_j = C)$.

% or...

Take $G$ to be well-constrained. The operator $\cap_V$ returns the count of the vertices in the intersection of two graphs. Then, for all nodes $C$ in $DRP(G)$, exactly one of the following is true:
\begin{align*}
    \forall i,j & (C_i \cap_V C_j = 1) \wedge \p{\bigcup_{k\in\set{1,\ldots,N}\setminus\set{i}}{C_k} \neq C} \text{ or} \\
    \forall i,j & (C_i \cap_V C_j \geq 2) \wedge (C_i \cup C_j = C)
\end{align*}

That is, either any two children share a single vertex and all of them are needed to form the parent node, or there is some significant overlap of 2 or more vertices and because of that only 2 children are needed at node $C$ to satisfy the definition of the DR-plan.

This implies that in the first scenario all proper vertex-maximal rigid subgraphs of $C$ must be in tree $DRP(C)$, and that in the second scenario only 2 such subgraphs need to be in the tree.

$C_i \cap_V C_j$ cannot be $<1$ because 0 is disjoint but $G$ is taken to be well-constrained (negative and non-integral values are meaningless).
\end{conjecture}

\begin{conjecture}
$DRP(G)$ is ``unique'' modulo choice of children. Which means that the choice of children cannot impact the solution of the framework.
\end{conjecture}

\end{comment}

\begin{comment}
\pnplaceholder
Notes:
Edges do not appear as nodes of the DR plan, although vertices do (as leaves). In general, the children of a node are well-constrained proper maximal subgraphs of the parent, whose vertex set union gives the parent's vertex set.

We are trying to formalize here the concept of a canonical DR-plan of a well-constained graph; i.e, whether the non-uniqueness of the decomposition - in the case that some pair of proper vertex maximal rigid subgraphs have a nontrivial intersection - makes any real difference.

Conjecture 1 (should be easy to prove):
Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_2| \ge 2$ (called a nontrivial intersection). Then the subgraph $S_{1,2}$ induced by $V_1\cap V_2$ must be well-constrained.

Here are corollary facts that should be easy to prove:
(1) There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
(2) There is no rigid subgraph induced by a proper, nonempty subset of vertices of   $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.

Corollary of Conjecture 1 and resulting facts:
$C_1\cap C_2$ is a proper vertex maximal rigid subgraph of $C_i$. Either there is a unique second proper vertex maximal rigid subgraph of $C_i$ that intersects  $C_1\cap C_2$  nontrivially (on an edge), or every pair of proper vertex maximal rigid subgraphs of $C_i$ intersect trivially.

This would imply the DR-plans of $C_1$ and $C_2$ will be unique and have a common child in the DR-plan of $C_1\cap C_2$.

Conjecture 2: Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_1| \ge 2$. $C_3 =(V_3,E_3),C_4 = (V_4,E_4)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_3\cup V_4$ and $|V_3\cap V_4| \ge 2$. Then both DR-plans where $C_1$ and $C_2$ are taken as children or where $C_4$ and $C_3$ are taken as children will have the same max fan-in. In particular, either DR-plan will contain $\bigcap\limits_i C_i$ as a node - at the 2nd level below $C$.
\end{comment}

% \pnplaceholder

% Lemma 1: if the graph is well-constrained and any two proper vertex maximal subgraphs have an overlap of only one vertex, then all of the subgraphs will have 1 or 0.

% Overlap of one vertex between any $C_i$ and $C_j$ means that $C\neq C_i\cup C_j$ because we know $C$ is well-constrained and the graph $C_i\cup C_j$ would be able to pivot around the single vertex.

% \pnplaceholder

% Need to define trivial as a graph that, were it only accounting for vertices, would be over-constrained. I.e.\ a single vertex in 2D. Or a vertex and an edge in 3D.

% \pnplaceholder

% \begin{theorem}
% phrasing...
% \end{theorem}

% \begin{proof}
% By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained itself, or (3) an over-constrained graph, referred to as trivial. We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.

% (1) If the union results in an under-constrained subgraph, then the intersection of the two original subgraphs must be an over-constrained trivial subgraph (can be shown with simple arithmetic). As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are over-constrained trivial subgraphs (including $\emptyset$). All of these edge-disjoint subgraphs form a complex structure that is well-constrained and must be the graph itself (otherwise we did not have vertex-maximal subgraphs to begin with).

% (2) If the union results in a well-constrained subgraph, it must be the entire graph. If this were not the case, the two original subgraphs could not have been vertex-maximal proper because we would have found a larger well-constrained subgraph containing them; namely, their union or something containing it would have been larger.

% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$. It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case). The intersection also cannot be under-constrained. If $C_1 \cap C_2$ were under-constrained, then for $C_1$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection. The same is true for $C_2$. If both of these were satisfied, then the union of $C_1$ and $C_2$ would result in an over-constrained nontrivial graph. This is a contradiction, because we have shown the union to be the graph itself, which was defined to be well-constrained. Therefore, the intersection must be well-constrained.

% (3) It is impossible for the union of 2 nontrivial subgraphs to be trivial.
% \end{proof}

% \begin{corollary}
% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
% \end{corollary}

% \begin{proof}
% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$. That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
% \end{proof}

% \begin{corollary}
% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
% \end{corollary}

% \begin{proof}
% Since $C_1\cup C_2$ is the entire graph,

% well... no. if $V_i$ contains $V_1 \setminus (V_1\cap V_2)$ or $V_2 \setminus (V_1\cap V_2)$, then you will induce a well-constrained graph.
% \end{proof}

% \begin{corollary}
% $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every $C_i$.
% \end{corollary}

% \begin{proof}
% First, it must vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}

% \pnplaceholder

% \pnplaceholder

% need to show that that intersection is going to end up being a vertex-maximal subgraph of both of the original subgraphs. As a result, in the second case it doesn't matter which 2 you took.

% --------------------------------------------------------------
%                        End of Body
% --------------------------------------------------------------

% \end{multicols}
\end{document}

% --------------------------------------------------------------
%                       End of Document
% --------------------------------------------------------------

### 2014-12-02 doc/General_Work/DRP_general.tex
Added documents.

% --------------------------------------------------------------
%                       Initialize
% --------------------------------------------------------------

\documentclass[11pt]{article}

\usepackage{personalNotes}
% contains special macros for math environment:
%     unit, p, set, abs, floor, ceil, vec, norm, inner, kernel
% contains useful all-purpose macros:
%     n, lessSeparation, append

\usepackage{tikz}
\usetikzlibrary{shapes,snakes}
\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}

% --------------------------------------------------------------
%                    Beginning of Document
% --------------------------------------------------------------

\begin{document}

% --------------------------------------------------------------
%                           Title
% --------------------------------------------------------------

\pntitle {Graph Theory}
\pnauthor{Troy Baker}
\pndate  {May 13, 2014}

\pnmaketitle
\pnmakefooter

% --------------------------------------------------------------
%                           Body
% --------------------------------------------------------------

\section{Definitions}
ISSUE: My definitions don't prevent non-positive densities... This means things aren't guaranteed to be connected...

\begin{definition}[Graph]
Graph $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$. Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
\end{definition}

\begin{definition}[Density]
The density of a graph is defined as:
\[d(G) = \sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}\]
% \[d(G) = \max\set{0,\sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}}\]
\end{definition}

\begin{definition}[Underconstrained Graph]
Graph $G=(V,E,w_V,w_E)$ is underconstrained if, given constant $k$, $d(G) > k$.
\end{definition}

\begin{definition}[Overconstrained Graph]
Graph $G=(V,E,w_V,w_E)$ is overconstrained if, given constant $k$, $d(G) < k$.
\end{definition}

\begin{definition}[Wellconstrained Graph]
Graph $G=(V,E,w_V,w_E)$ is wellconstrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either underconstrained, wellconstrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
\end{definition}

\begin{definition}[Trivial Graph]
A trivial graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.

In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= {{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
\end{definition}

\begin{definition}[Decomposition Recombination (DR-) Plan]
A DR-plan of graph $G$, $DRP(G)$, is a tree that has the following properties:
\begin{enumerate}
    \item The root of the tree `contains' $G$.
    \item For all nodes, $C$, that `contain' nontrivial graphs, its $N$ children, $C_1, \ldots, C_N$, are either trivial subgraphs or wellconstrained vertex-maximal proper subgraphs of that node $C$.
    \item $\bigcup_{i=1}^N{C_i}=C$.
    \item OR IS IT..... The vertex set of $\bigcup_{i=1}^N{C_i}$ has all the vertices that are in $C$.
\end{enumerate}

% It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.

An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.

Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}

\begin{definition}[Optimal DR-Plan]
The optimal DR-plan of graph $G$, $OPTDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree/DAG.
\end{definition}

\section{Notation}
$G$ is taken to be a graph with values $(V,E,w_V,w_E)$. To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages. It is not useful to solve for two bodies if there are no constraints between them. Furthermore, to limit the scope of the paper, $G$ will always be taken to be wellconstrained. THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.

$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.

$C$ is an arbitrary graph in a node in $DRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$. By definition of a DR-plan, it is implied that $C_i$ is a wellconstrained vertex-maximal proper subgraph.

\section{The Canonical Optimal DR-Plan}
\begin{figure}
\begin{center}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (1,1) {$C'_j$};
    \node[v] (v2) at (3,1) {$C'_k$};
    \node[v] (v3) at (2,2) {$D$};
    \node[v] (v4) at (2,3.3) {$C'_i$};

    \tedge{v1}{v3}{solid}{}{above};
    \tedge{v2}{v3}{solid}{}{above};
    \tedge{v3}{v4}{solid}{}{right};
\end{tikzpicture}
\ldots with weights \ldots
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (1,1) {y};
    \node[v] (v2) at (3,1) {z};
    \node[v] (v3) at (2,2) {k};
    \node[v] (v4) at (2,3.3) {x};

    \tedge{v1}{v3}{solid}{y}{above};
    \tedge{v2}{v3}{solid}{z}{above};
    \tedge{v3}{v4}{solid}{x}{right};

    % \draw (3.3,2.5) circle (1.9cm);
    % \draw (.7,2.5) circle (1.9cm);
    % \draw (2,.5) circle (1.9cm);
\end{tikzpicture}
\\
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxy};
    \node[v] (v3) at ( 1,-1) {kyz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v5}{solid}{}{};
    \tedge{v3}{v5}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxy};
    \node[v] (v3) at ( 1,-1) {kxz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v5}{solid}{}{};
    \tedge{v3}{v4}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{e}=[]

    \node[v] (v1) at (0,  0) {kxyz};
    \node[v] (v2) at (-1,-1) {kxz};
    \node[v] (v3) at ( 1,-1) {kyz};
    \node[v] (v4) at (-2,-2) {kx};
    \node[v] (v5) at ( 0,-2) {ky};
    \node[v] (v6) at ( 2,-2) {kz};

    \tedge{v1}{v2}{solid}{}{};
    \tedge{v1}{v3}{solid}{}{};
    \tedge{v2}{v4}{solid}{}{};
    \tedge{v2}{v6}{solid}{}{};
    \tedge{v3}{v5}{solid}{}{};
    \tedge{v3}{v6}{solid}{}{};
\end{tikzpicture}
% \\
% \begin{tikzpicture}
%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
%     \tikzstyle{e}=[]

%     \node[v] (v1) at ( 0, 0) {kxyz};
%     \node[v] (v2) at (-1,-1) {kxy};
%     \node[v] (v3) at (-2,-2) {kx};
%     \node[v] (v4) at (-3,-3) {k};
%     \node[v] (v5) at ( 1,-1) {z plan};
%     \node[v] (v6) at ( 0,-2) {y plan};
%     \node[v] (v7) at (-1,-3) {x plan};
%     \node[v] (v8) at (-2,-4) {k plan};

%     \tedge{v1}{v2}{solid}{}{};
%     \tedge{v1}{v5}{solid}{}{};
%     \tedge{v2}{v3}{solid}{}{};
%     \tedge{v2}{v6}{solid}{}{};
%     \tedge{v3}{v4}{solid}{}{};
%     \tedge{v3}{v7}{solid}{}{};
%     \tedge{v4}{v8}{solid}{}{};
% \end{tikzpicture}
\end{center}

\caption{A wellconstrained graph whose nodes are subgraphs labeled with their density. Below are its 3 optimal DR-Plans (partially complete). The center vertex has weight $k$ and is wellconstrained itself. The radial nodes have weights $x,y,z \neq k$ with equivalent edge weights, making any union between $k$ and the others wellconstrained.}
\label{3-plans}
\end{figure}

In the case of a wellconstrained graph, it is clear that there always exists a single vertex-maximum DR-Plan that includes every single wellconstrained vertex-maximal proper subgraph at each node. However, in the problem of the optimal DR-Plan there is no such unique plan. For instance, a graph that is formed by the union of 3 wellconstrained subgraphs that all intersect on a wellconstrained subgraph, demonstrated in Figure \ref{3-plans}. In this case, it results in $3$ unique optimal DR-Plans. Indeed, we will prove a union of $N$ wellconstrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same. Therefore, all choices of decomposition are in some sense equivalent.

We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial wellconstrained vertex-maximal proper subgraphs of a wellconstrained graph. By definition, any subgraph of a wellconstrained graph will either be (1) trivial, (2) underconstrained, or (3) wellconstrained. We will analyze all 3 possibilities in turn. These observations will lead to interesting properties that govern the behavior of optimal DR-plans on wellconstrained graphs.

% We show this by carefully considering all of the potential cases of the unions of subgraphs. By definition, any subgraph of a wellconstrained graph will either be (1) underconstrained, (2) wellconstrained, or (3) trivial. We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial wellconstrained vertex-maximal proper subgraphs.

\subsection{Lemmas for Union and Intersection Relationships}
\begin{lemma}\label{t-l1}
Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ cannot be trivial.
\end{lemma}

\begin{proof}
% If $F_i,F_j$ are wellconstrained, then $F_i\cup F_j \geq k$. Trivial graphs must be overconstrained.

Follows from the definition of a trivial graph. It would imply that the wellconstrained graphs were also trivial.
\end{proof}

\begin{lemma}\label{uc-l1}
Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial.
\end{lemma}

\begin{proof}
This can be shown with simple arithmetic. It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$. In the forward direction we have that $d(F_i\cup F_j)=l>k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=2k-l<k$. The only way to have an overconstrained subgraph in a wellconstrained graph is if it is trivial. In the backwards direction the same math proves it.
\end{proof}

\begin{lemma}\label{wc-l1}
Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ is wellconstrained if and only if $F_i\cap F_j$ is wellconstrained.
\end{lemma}

\begin{proof}
This can be shown with simple arithmetic. It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$. In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$. Thus $F_i\cap F_j$ is wellconstrained. In the backwards direction the same math proves it.

Furthermore, all subgraphs of $F_i\cup F_j$ and $F_i\cap F_j$ are also subgraphs of $F_i$ and $F_j$, therefore all properties of wellconstrained graphs are maintained.

% In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial. The intersection also cannot be underconstrained. If $F_i \cap F_j$ were underconstrained, then for $F_i$ to be wellconstrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection. The same is true for $F_j$. If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an overconstrained nontrivial graph. This is a contradiction, by definition of a wellconstrained graph. Therefore, the intersection can only be wellconstrained.

% BACKWARDS!!!!!
\end{proof}

\begin{lemma}\label{iuc-l1}
Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cap F_j$ cannot be underconstrained.
\end{lemma}

\begin{proof}
Subgraphs of wellconstrained graph $G$ can only be trivial, underconstrained, or wellconstrained. So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and wellconstrained intersections are possible.
\end{proof}

\subsection{Lemmas for Intersection and Graph Structure Relationships}

\begin{lemma}\label{wc-l2}
% Given a wellconstrained $G$ and two wellconstrained vertex-maximal proper subgraphs $C_i, C_j$,
$C_i\cup C_j$ is wellconstrained if and only if $C_i\cup C_j = C$.
\end{lemma}

\begin{proof}
The forwards direction can be shown with a contradiction. If the $C_i\cup C_j$ was wellconstrained but $C_i\cup C_j \neq C$, we would have found a larger wellconstrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger. The two original subgraphs could not have been vertex-maximal proper.

The backwards direction follows from the fact that $C$ is either some non-leaf node (wellconstrained by definition of a DR-plan) or $G$ itself (wellconstrained by definition of the problem).
\end{proof}

\begin{corollary}\label{wc-c1}
% Given a wellconstrained $G$ and two wellconstrained vertex-maximal proper subgraphs $C_i, C_j$, l
Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is wellconstrained, then there can be no edges in $G$ between the vertices of $R_i$ and $R_j$.
\end{corollary}

\begin{proof}
It follows from the fact that $C_i\cup C_j$ must equal $C$ (proven in Lemma \ref{wc-l2}).
\end{proof}

% \begin{corollary}
% Given the same conditions: $R_i$ and $R_j$ must be connected to $D_{i,j}=C_i\cap C_j=C\setminus R_i\setminus R_j$.
% \end{corollary}

% \begin{proof}
% From Lemma \ref{wc-l1}, $D_{i,j}$ must be wellconstrained.
% \end{proof}

\begin{lemma}
asdfas
\end{lemma}

\begin{proof}
Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=C\setminus R_i\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$). Furthermore, $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.

Then we assume that there is some third wellconstrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

% \begin{enumerate}
%     \item Assume $C_k\cup C_i \neq C$ and $C_k\cup C_j \neq C$. Then from Lemma \ref{wc-l2} $C_k\cup C_i$ is not wellconstrained and must therefore be underconstrained. Furthermore, $C_k\cup C_i=Idc(C,R'_i\cup C_i)$ because $C_k\neq C_i$. This means the subgraph $R'_i$ and its edges between $C_i$ increase the density overall because $C_i$ is wellconstrained. Similarly, the same is true for $R'_j$ and its edges. Since there can be no edges between $R'_i$ and $R'_j$ (COROLLARY, wouldn't be $C$!!!!) it means that all of the edges connecting them are incident to $D_{i,j}$. Therefore, $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ should be underconstrained. However, by Lemma \ref{uc-l1}, using the two underconstrained graphs we have that $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ so it must be trivial which is a contradiction, so $C_k$ could not be a wellconstrained vertex-maximal proper subgraph.
%     % Similarly, we find the $C_k\cup C_j=Idc(C,R'_j\cup C_j)$ is underconstrained. However, $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ which is wellconstrained. This contradicts Lemma \ref{wc-l1} so $C_k$ could not be wellconstrained.
%     \item Assume $C_k\cup C_i\neq C$ and $C_k\cup C_j= C$. Similar to the first case, we see that

%     If it's just one and the other does equal C, then the other is not vertex-maximal?
%     \item If it's neither... well then you either have C or the solution.
% \end{enumerate}

\begin{itemize}
    \item 3 cases: $C_k$ cannot be $C=Idc(C,R_i\cup R_j\cup D_{i,j})$, $C_i=Idc(C,R_j\cup D_{i,j})$, or $C_j=Idc(C,R_i\cup D_{i,j})$. This is by definition.
    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal. These are the graphs $Idc(C,R'_i\cup D_{i,j})$, $Idc(C,R'_j\cup D_{i,j})$, $Idc(C, D_{i,j})$, $Idc(C,R_i\cup D'_{i,j})$, $Idc(C,R_j\cup D'_{i,j})$, $Idc(C,R'_i\cup D'_{i,j})$, $Idc(C,R'_j\cup D'_{i,j})$, $Idc(C, D'_{i,j})$, $Idc(C,R_i)$, $Idc(C,R_j)$, $Idc(C,R'_i)$, $Idc(C,R'_j)$, and $Idc(C,\emptyset)$.
    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal. These are the graphs $Idc(C,R'_i\cup R_j\cup D_{i,j})$ and $Idc(C,R_i\cup R'_j\cup D_{i,j})$ respectively.
    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$...
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be wellconstrained. We also know it cannot be trivial because it contains wellconstrained subgraphs. This means it must be underconstrained. From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial. This is impossible because $D_{i,j}$ is wellconstrained, thereby contradicting the assumption that $C_k$ is wellconstrained.
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible...
    %
    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible...
    %
    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that is left!
\end{itemize}

The weight of $R_i$ has to be exactly equal to the weight of the edges that connect it to $D_{i,j}$. It must add 0 weight exactly since $D_{i,j}$ is wellconstrained.

There can't be an intersection that is underconstrained between to wellconstrained graphs!

\begin{itemize}
    \item $C_k$ cannot be $Idc(C,R_i\cup R_j)$. From Corollary \ref{wc-c1}, $R_i$ and $R_j$ must not have any edges between them. Without loss of generality, we can say that $d(R_i)=l$ and $d(R_j)=k-l$. Because $C_i$ and $C_j$ are wellconstrained, and so is $D_{i,j}$, $R_i$ with its edges must add 0 density and $R_j$ with its edges must add 0 density. Since edges must have positive weight, $l\geq 0$ and $k-l\geq 0$ so $d(R_i)\in [0,k]$ and $d(R_j)\in [0,k]$ with $d(R_i)+d(R_j)=k$. This means, without loss of generality, that at least $R_i$ is trivial
    %
    % \item 1 case: $C_k=Idc(C,R_i\cup R_j)$ is not possible...
    % \item 1 case: $C_k=Idc(C,R'_i\cup R'_j)$ is not possible...
    % \item 2 cases: $C_k=Idc(C,R'_i\cup R_j)$ and $C_k=Idc(C,R_i\cup R'_j)$ are not possible...
    % \item 4 cases: $C_k$ must contain a subgraph of $D_{i,j}$. Consider if it did not, then $C_k\cap D_{i,j}=\emptyset$ which is necessarily a trivial graph. By Lemma \ref{uc-l1}, this implies that $C_k\cup D_{i,j}$ is underconstrained, so $d(C_k\cup D_{i,j})>k$. Because there is no intersection, $d(C_k\cup D_{i,j})=d(C_k)+d(D_{i,j})$. However, we know that $D_{i,j}$ is wellconstrained, so $d(D_{i,j})=k$. This means that $d(C_k)$ must be greater than $k$ which makes it underconstrained, contradicting the assumption. This means $C_k$ cannot be $C_k=Idc(C,R_i\cup R_j)$, $C_k=Idc(C,R'_i\cup R_j)$, $C_k=Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$.
    %
    \item 4 cases: $C_k$ must contain a subgraph of $D_{i,j}$. Take $C_k=Idc(C,R_i\cup R_j)$. In this case, $C_k\cup C_i = R_i\cup C_i$ which does not equal $C$ because it is missing the edges between $R_i$ and $D_{i,j}$. We have from Lemma \ref{wc-l2} that this union cannot be wellconstrained. It cannot be trivial because it has wellconstrained subgraphs which leaves it to be underconstrained. We know $C_i$ is wellconstrained, therefore it must be $R_i$ that is underconstrained. This can be used to show that $R_j$ is also underconstrained. This means $R_i\cup R_j=C_k$ must be underconstrained, contradicting the assumption. This proof holds for several other cases, meaning $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$.
    %
    %
    %
    % \item $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. If this were wellconstrained the intersection with $C_i$ would be wellconstrained, which is $Idc(C,R'_j\cup D_{i,j})$. Similarly, it would imply
    \item $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since this graph's union with $C_i$ is $Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that this union cannot be wellconstrained (and must therefore be underconstrained). We know $Idc(C,R_j\cup D_{i,j})=C_i$ is wellconstrained which implies $R'_j$ and the edges incident to it are increasing the density. Using the same process, we can show the same for $R'_i$. Since $D_{i,j}$ is wellconstrained, adding two components that increase the density would mean that $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is underconstrained, contradicting the assumption.
    %
    \item $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ and $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ are not possible. In both scenarios we have $C_k\cup C_i \neq C$ and $C_k\cup C_j \neq C$. From Lemma \ref{wc-l2} $C_k\cup C_i$ cannot be wellconstrained and must therefore be underconstrained. Furthermore, $C_k\cup C_i=Idc(C,R'_i\cup C_i)$ because $C_k\neq C_i$. This means the subgraph $R'_i$ and its edges between $C_i$ increase the density overall because $C_i$ is wellconstrained. Similarly, the same is true for $R'_j$ and its edges. Since there can be no edges between $R'_i$ and $R'_j$ (COROLLARY, wouldn't be $C$!!!!) it means that all of the edges connecting them are incident to $D_{i,j}$. Therefore, $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ should be underconstrained. However, by Lemma \ref{uc-l1}, using the two underconstrained graphs we have that $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ so it must be trivial which is a contradiction, so $C_k$ could not be a wellconstrained vertex-maximal proper subgraph.
    % \item $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible...
\end{itemize}
\end{proof}

\pagebreak
See ISSUE at top of paper, we have to assume we have a connected $G$...

Take $C'_i$ to be $C\setminus C_i$. Then if $C_i\cup C_j$ is wellconstrained we have directly from Lemma \ref{wc-l2} for all $C'_i, C'_j$ there can be no edges between them (otherwise $C_i\cup C_j\neq C$).

The set of edges connecting $C'_i$ to $C_i$ must have a weight that sums to exactly $d(C'_i)$ since we know that $d(C)=d(C_i)=k$.

Any other $C_k$ must therefore contain $C'_i$ and $C'_j$ because it adds no density to ``tack it on'' to a graph that does not contain them already.

Now we want to know something about $C_k$. If $C_k \neq C_i$, then $C'_i\not\subseteq C'_k$. It is clear that they cannot be equal otherwise $C_k = C_i$. However, it also cannot be a subset. There are two scenarios. (1) $C'_k$ does not contain vertices that are incident to $C'_i$. If this were so, we know we could ``tack on'' $C'_i$ to the wellconstrained $C_k$ without increasing the density and therefore $C_k$ could not have been maximal in the first place. (2) $C'_k$ contains vertices that are incident to $C'_i$. In this case

Take $D_{i,j}$ to be $C\setminus (C'_i\cup C'_j)$. Clearly $D_{i,j}$ must be wellconstrained since ``tacking on'' $C'_i$ and $C'_j$ adds no density and $C$ is wellconstrained. Therefore, if there were some other $C_k$, it must contain $D_{i,j,k}\subset D_{i,j}$ such that $D_{i,j,k}=D_{i,j}\setminus C'_k$. It is clear that $D_{i,j,k}$ cannot equal $D_{i,j}$ or else $C_k$ must be $C_i$ or $C_j$ to be vertex-maximal.

\begin{lemma}
If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial or empty set.
\end{lemma}

\begin{proof}
By Lemmas... the union must be under
Lemma \ref{uc-l1} that $C_i\cup C_j$ must be underconstrained. Therefore, by the proof
It follows from Lemma \ref{wc-l1} that the only way to get
 and its proof that if any intersections
\end{proof}

\subsection{Theorems for Optimal DR-Plans}

\subsection{Union is Underconstrained}

\begin{figure}
\begin{center}
\begin{tikzpicture}
    \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
    \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
    \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);

    % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
    % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);

    % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
    % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
\end{tikzpicture}
\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
    \tikzstyle{e}=[]

    \node[el] (m) at (0,    0) {k};
    \node[v]  (l) at (-1.5, 0) {b};
    \node[v]  (r) at (1.5,  0) {c};

    \node[el] (tl) at (-1,  1) {i};
    \node[el] (tr) at (1,   1) {j};
    \node[v]  (t)  at (0,   2) {a};

    % \node[el] (bl) at (-1,  -1) {l};
    % \node[el] (br) at (1,   -1) {m};
    % \node[v]  (b)  at (0,   -2) {d};

    % \node[el] (v2) at (-1,-1) {b};
    % \node[el] (v3) at (-1,-1) {b};
    % \node[el] (v4) at (-1,-1) {b};
    % \node[el] (v5) at (-1,-1) {b};

    \tedge{m}{l}{solid}{}{};
    \tedge{m}{r}{solid}{}{};

    \tedge{l}{tl}{solid}{}{};
    \tedge{tl}{t}{solid}{}{};
    \tedge{r}{tr}{solid}{}{};
    \tedge{tr}{t}{solid}{}{};

    % \tedge{l}{bl}{solid}{}{};
    % \tedge{bl}{b}{solid}{}{};
    % \tedge{r}{br}{solid}{}{};
    % \tedge{br}{b}{solid}{}{};
\end{tikzpicture}
\end{center}

\caption{Each node contains a subgraph of the entire, wellconstrained graph. We take $a\cup i\cup b$, $a\cup j\cup c$, and $b\cup k\cup c$ to be wellconstrained. If the union of any two of these listed unions is underconstrained then it immediately follows that all such unions are underconstrained and that nodes $a,b,c$ must all be trivial subgraphs.
THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
% Each node contains a subgraph of the entire, wellconstrained graph. We take $a\cup i\cup b$, $a\cup j\cup c$, $b\cup k\cup c$, $b\cup l\cup d$, and $c\cup m\cup d$ to be wellconstrained. If the union of any two of these listed unions is underconstrained then it immediately follows that all such unions are underconstrained and that nodes $a,b,c,d$ must all be trivial subgraphs.
% THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
}
\label{trivial-intersection}
\end{figure}

\begin{theorem}\label{uc-t1}
% In any solution to $OPTDRP(G)$, if any two wellconstrained vertex-maximal proper subgraphs of a node form a union that is underconstrained, then all wellconstrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.

In $OPTDRP(G)$ of a wellconstrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
\end{theorem}

\begin{proof}
From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be underconstrained. If it is underconstrained, it must be part of some larger wellconstrained subgraph $D$. If $D$ is not $G$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
 For $C$ to be wellconstrained... WHAT????

If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was wellconstrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.

This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid. Such an example is shown in Figure \ref{trivial-intersection}.
\end{proof}

% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$. All of these subgraphs form a complex structure that is wellconstrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).

\subsection{Union is Wellconstrained}

\begin{theorem}\label{wc-t1}
% In $OPTDRP(G)$ of a wellconstrained $G$, if any two wellconstrained vertex-maximal proper subgraphs of a node form a union that is wellconstrained, then said node can have only $2$ children.

In $OPTDRP(G)$ of a wellconstrained $G$, if $C_i \cup C_j$ is wellconstrained, then any 2 wellconstrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
% Namely, these children are any 2 of its wellconstrained vertex-maximal proper subgraphs.
\end{theorem}

\begin{proof}
For a valid DR-plan, all children must be proper subgraphs. Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least 2 children. It follows directly from Lemma \ref{wc-l2} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is wellconstrained, then you found a tree with only 2 children.

Clearly, this is the optimal choice at this level. However, for it to be a solution to $OPTDRP(G)$, it must not affect the maximum fan-in later on.

Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. Take $C\setminus C_k$ to be $F_k$.

After we go down $N$ levels we get to a point where every node is some $F_k$ with $D$ and all the edges connecting them.

This is demonstrated with $N=3$ in Figure \ref{3-plans}. It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution. This follows from the discussion above. Now, at this level, the fan-in must be equal in all scenarios. In fact, at node $k$ it will be 1 (because of $D$ which is wellconstrained) plus whatever fan-in $F_k$ incurs.
\end{proof}

% (2) If the union results in a wellconstrained subgraph, it must be the entire graph. If this were not the case, .

% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$. It cannot be overconstrained, otherwise the union could not be wellconstrained (simple arithmetic, see the first case).

\begin{corollary}
Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is wellconstrained, then $D$ will be a descendant of every $C_k$.
\end{corollary}

\begin{proof}
Explained in Theorem \ref{wc-t1}.
\end{proof}

\begin{corollary}
$\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
\end{corollary}

\begin{proof}
Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
\end{proof}

% \begin{proof}
% It follows from Theorem \ref{uc-t1} that if the intersection of any two wellconstrained vertex-maximal proper subgraphs of $C$ are wellconstrained then all of them are wellconstrained or $\emptyset$.
% \end{proof}

% Figure idea: Large node in the middle with weight $k$. Several other graphs with weight $a, b, c, d\ldots$ attached radially by single edges with weights $a, b, c, d\ldots$ respectively. Circle each vertex-maximal subset.

% $\forall k\in [1,\ldots,N]$ the subgraphs $C\setminus C_k$ will be edge disjoint. Therefore the solving of one is exclusive of the solving of the other.

% Furthermore, the union of all of these subgraphs will be $C\setminus D$. So each subgraph will need to be ``tacked'' onto $D$ for recombination and since they are edge-disjoint the order is unimportant.

% \pnplaceholder
% \begin{lemma}
% If $C_1$ and $C_2$ are wellconstrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is wellconstrained, then $C_1 \cap C_2$ is a wellconstrained vertex-maximal proper subgraph of every child of $C$.
% \end{lemma}

% \begin{proof}
% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained wellconstrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}
% \pnplaceholder

\subsection{Further Results}

% \begin{corollary}
% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
% \end{corollary}

% \begin{proof}
% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$. That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
% \end{proof}

\begin{corollary}
There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
\end{corollary}

\begin{proof}
Since $C_1\cup C_2$ is the entire graph, PROVE THIS!!!!
\end{proof}

% \begin{corollary}
% If $C_1$ and $C_2$ are wellconstrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is wellconstrained, then $C_1 \cap C_2$ is a wellconstrained vertex-maximal proper subgraph of every child of $C$.
% \end{corollary}

% \begin{proof}
% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained wellconstrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}

\section{Future Work}
While this paper fleshes out the case of a wellconstrained input graph, little work has been done on the overconstrained graph. This has to do with the fact that finding a DR-Plan is NP-complete.

Underconstrained is not worth studying in itself, because it will either be that all subgraphs are underconstrained or trivial or that some subgraphs are wellconstrained or overconstrained. In this sense, it is not an independent problem.

% \section{Formal Conjectures}

% \subsection{Conjecture 1}

% \begin{conjecture}
% % Take $G$ to be wellconstrained. Then, for all nodes $C$ in $DRP(G)$, either $\forall i,j (C_i \cap C_j \leq 1)$ or $\forall i,j (C_i \cap C_j \geq 2) \wedge (C_i \cup C_j = C)$.

% % or...

% Take $G$ to be wellconstrained. The operator $\cap_V$ returns the count of the vertices in the intersection of two graphs. Then, for all nodes $C$ in $DRP(G)$, exactly one of the following is true:
% \begin{align*}
%     \forall i,j & (C_i \cap_V C_j = 1) \wedge \p{\bigcup_{k\in\set{1,\ldots,N}\setminus\set{i}}{C_k} \neq C} \text{ or} \\
%     \forall i,j & (C_i \cap_V C_j \geq 2) \wedge (C_i \cup C_j = C)
% \end{align*}

% That is, either any two children share a single vertex and all of them are needed to form the parent node, or there is some significant overlap of 2 or more vertices and because of that only 2 children are needed at node $C$ to satisfy the definition of the DR-plan.

% This implies that in the first scenario all proper vertex-maximal rigid subgraphs of $C$ must be in tree $DRP(C)$, and that in the second scenario only 2 such subgraphs need to be in the tree.

% $C_i \cap_V C_j$ cannot be $<1$ because 0 is disjoint but $G$ is taken to be wellconstrained (negative and non-integral values are meaningless).
% \end{conjecture}

% \begin{conjecture}
% $DRP(G)$ is ``unique'' modulo choice of children. Which means that the choice of children cannot impact the solution of the framework.
% \end{conjecture}

\pnplaceholder
Notes:
Edges do not appear as nodes of the DR plan, although vertices do (as leaves). In general, the children of a node are well-constrained proper maximal subgraphs of the parent, whose vertex set union gives the parent's vertex set.

We are trying to formalize here the concept of a canonical DR-plan of a well-constained graph; i.e, whether the non-uniqueness of the decomposition - in the case that some pair of proper vertex maximal rigid subgraphs have a nontrivial intersection - makes any real difference.

Conjecture 1 (should be easy to prove):
Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_2| \ge 2$ (called a nontrivial intersection). Then the subgraph $S_{1,2}$ induced by $V_1\cap V_2$ must be wellconstrained.

Here are corollary facts that should be easy to prove:
(1) There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
(2) There is no rigid subgraph induced by a proper, nonempty subset of vertices of   $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.

Corollary of Conjecture 1 and resulting facts:
$C_1\cap C_2$ is a proper vertex maximal rigid subgraph of $C_i$. Either there is a unique second proper vertex maximal rigid subgraph of $C_i$ that intersects  $C_1\cap C_2$  nontrivially (on an edge), or every pair of proper vertex maximal rigid subgraphs of $C_i$ intersect trivially.

This would imply the DR-plans of $C_1$ and $C_2$ will be unique and have a common child in the DR-plan of $C_1\cap C_2$.

Conjecture 2: Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_1| \ge 2$. $C_3 =(V_3,E_3),C_4 = (V_4,E_4)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_3\cup V_4$ and $|V_3\cap V_4| \ge 2$. Then both DR-plans where $C_1$ and $C_2$ are taken as children or where $C_4$ and $C_3$ are taken as children will have the same max fan-in. In particular, either DR-plan will contain $\bigcap\limits_i C_i$ as a node - at the 2nd level below $C$.

% \pnplaceholder

% Lemma 1: if the graph is wellconstrained and any two proper vertex maximal subgraphs have an overlap of only one vertex, then all of the subgraphs will have 1 or 0.

% Overlap of one vertex between any $C_i$ and $C_j$ means that $C\neq C_i\cup C_j$ because we know $C$ is wellconstrained and the graph $C_i\cup C_j$ would be able to pivot around the single vertex.

% \pnplaceholder

% Need to define trivial as a graph that, were it only accounting for vertices, would be overconstrained. I.e.\ a single vertex in 2D. Or a vertex and an edge in 3D.

% \pnplaceholder

% \begin{theorem}
% phrasing...
% \end{theorem}

% \begin{proof}
% By definition, any subgraph of a wellconstrained graph will either be (1) underconstrained, (2) wellconstrained itself, or (3) an overconstrained graph, referred to as trivial. We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial wellconstrained vertex-maximal proper subgraphs.

% (1) If the union results in an underconstrained subgraph, then the intersection of the two original subgraphs must be an overconstrained trivial subgraph (can be shown with simple arithmetic). As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are overconstrained trivial subgraphs (including $\emptyset$). All of these edge-disjoint subgraphs form a complex structure that is wellconstrained and must be the graph itself (otherwise we did not have vertex-maximal subgraphs to begin with).

% (2) If the union results in a wellconstrained subgraph, it must be the entire graph. If this were not the case, the two original subgraphs could not have been vertex-maximal proper because we would have found a larger wellconstrained subgraph containing them; namely, their union or something containing it would have been larger.

% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$. It cannot be overconstrained, otherwise the union could not be wellconstrained (simple arithmetic, see the first case). The intersection also cannot be underconstrained. If $C_1 \cap C_2$ were underconstrained, then for $C_1$ to be wellconstrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection. The same is true for $C_2$. If both of these were satisfied, then the union of $C_1$ and $C_2$ would result in an overconstrained nontrivial graph. This is a contradiction, because we have shown the union to be the graph itself, which was defined to be wellconstrained. Therefore, the intersection must be wellconstrained.

% (3) It is impossible for the union of 2 nontrivial subgraphs to be trivial.
% \end{proof}

% \begin{corollary}
% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
% \end{corollary}

% \begin{proof}
% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$. That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
% \end{proof}

% \begin{corollary}
% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
% \end{corollary}

% \begin{proof}
% Since $C_1\cup C_2$ is the entire graph,

% well... no. if $V_i$ contains $V_1 \setminus (V_1\cap V_2)$ or $V_2 \setminus (V_1\cap V_2)$, then you will induce a wellconstrained graph.
% \end{proof}

% \begin{corollary}
% $C_1 \cap C_2$ is a wellconstrained vertex-maximal proper subgraph of every $C_i$.
% \end{corollary}

% \begin{proof}
% First, it must vertex-maximal proper subgraph of $C_1$ and $C_2$. Say that there was some larger proper subgraph in $C_1$ that was well constrained. If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained wellconstrained. The same is true vice versa.

% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
% \end{proof}

% \pnplaceholder

% \pnplaceholder

% need to show that that intersection is going to end up being a vertex-maximal subgraph of both of the original subgraphs. As a result, in the second case it doesn't matter which 2 you took.

% --------------------------------------------------------------
%                        End of Body
% --------------------------------------------------------------

\end{document}

% --------------------------------------------------------------
%                       End of Document
% --------------------------------------------------------------

### 2014-12-02 doc/General_Work/notes-old.tex
Added documents.

--- 2014-12-02 doc/General_Work/DRP_general.tex
@@
 \section{Definitions}
-ISSUE: My definitions don't prevent non-positive densities... This means things aren't guaranteed to be connected...
-
 \begin{definition}[Graph]
-Graph $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$.
-Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
+Graph $G=(V,E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$.
+Additionally, there are two weight functions, $w_V: V \to \mathbb{R}$ and $w_E: E \to \mathbb{R}$.
 \end{definition}
@@
 \[d(G) = \sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}\]
-% \[d(G) = \max\set{0,\sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}}\]
 \end{definition}
@@
 \begin{definition}[Underconstrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is underconstrained if, given constant $k$, $d(G) > k$.
+Graph $G=(V,E)$ is underconstrained if, given constant $k$, $d(G) > k$.
 \end{definition}
@@
 \begin{definition}[Overconstrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is overconstrained if, given constant $k$, $d(G) < k$.
+Graph $G=(V,E)$ is overconstrained if, given constant $k$, $d(G) < k$.
 \end{definition}
@@
 \begin{definition}[Wellconstrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is wellconstrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either underconstrained, wellconstrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
+Graph $G=(V,E)$ is wellconstrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either underconstrained, wellconstrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
 \end{definition}
@@
 \section{Notation}
-$G$ is taken to be a graph with values $(V,E,w_V,w_E)$.
-To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
-It is not useful to solve for two bodies if there are no constraints between them.
-Furthermore, to limit the scope of the paper, $G$ will always be taken to be wellconstrained.
-THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.
+$G$ is taken to be a graph with vertex set $V$ and edge set $E$.
 
@@
 
-$C$ is an arbitrary graph in a node in $DRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$. By definition of a DR-plan, it is implied that $C_i$ is a wellconstrained vertex-maximal proper subgraph.
+$C$ is an arbitrary graph in a node in $DRP(G)$. $C_i$ is the $i^{\text{th}}$ child of $C$ (implied to be a wellconstrained vertex-maximal proper subgraph, by definition of a DR-plan).
 
@@
 \begin{center}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (1,1) {$C'_j$};
-    \node[v] (v2) at (3,1) {$C'_k$};
-    \node[v] (v3) at (2,2) {$D$};
-    \node[v] (v4) at (2,3.3) {$C'_i$};
-
-    \tedge{v1}{v3}{solid}{}{above};
-    \tedge{v2}{v3}{solid}{}{above};
-    \tedge{v3}{v4}{solid}{}{right};
-\end{tikzpicture}
-\ldots with weights \ldots
 \begin{tikzpicture}
@@
 We will analyze all 3 possibilities in turn.
-These observations will lead to interesting properties that govern the behavior of optimal DR-plans on wellconstrained graphs.
 
@@
 
-\subsection{Lemmas for Union and Intersection Relationships}
-\begin{lemma}\label{t-l1}
-Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ cannot be trivial.
+\subsection{Union is Trivial}
+\begin{lemma}
+The union of wellconstrained graphs cannot result in a trivial graph.
 \end{lemma}
@@
 \begin{proof}
-% If $F_i,F_j$ are wellconstrained, then $F_i\cup F_j \geq k$. Trivial graphs must be overconstrained.
+% If $C_i,C_j$ are wellconstrained, then $C_i\cup C_j \geq k$. Trivial graphs must be overconstrained.
 
@@
 
+\subsection{Union is Underconstrained}
 \begin{lemma}\label{uc-l1}
-Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial.
+Given two wellconstrained subgraphs $C_i,C_j$, $C_i\cup C_j$ is underconstrained if and only if $C_i\cap C_j$ is trivial.
 \end{lemma}
@@
 This can be shown with simple arithmetic.
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction we have that $d(F_i\cup F_j)=l>k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=2k-l<k$.
+It follows from definitions that $d(C_i)=k$ and $d(C_j)=k$.
+In the forward direction we have that $d(C_i\cup C_j)=l>k$, therefore $d(C_i\cap C_j)=d(C_i)+d(C_j)-d(C_i\cup C_j)=2k-l<k$.
 The only way to have an overconstrained subgraph in a wellconstrained graph is if it is trivial.
@@
 \end{proof}
-
-\begin{lemma}\label{wc-l1}
-Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cup F_j$ is wellconstrained if and only if $F_i\cap F_j$ is wellconstrained.
-\end{lemma}
-
-\begin{proof}
-This can be shown with simple arithmetic.
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$.
-Thus $F_i\cap F_j$ is wellconstrained.
-In the backwards direction the same math proves it.
-
-Furthermore, all subgraphs of $F_i\cup F_j$ and $F_i\cap F_j$ are also subgraphs of $F_i$ and $F_j$, therefore all properties of wellconstrained graphs are maintained.
-
-% In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial.
-The intersection also cannot be underconstrained.
-If $F_i \cap F_j$ were underconstrained, then for $F_i$ to be wellconstrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
-The same is true for $F_j$.
-If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an overconstrained nontrivial graph.
-This is a contradiction, by definition of a wellconstrained graph.
-Therefore, the intersection can only be wellconstrained.
-
-% BACKWARDS!!!!!
-\end{proof}
-
-\begin{lemma}\label{iuc-l1}
-Given two wellconstrained subgraphs $F_i,F_j$, $F_i\cap F_j$ cannot be underconstrained.
-\end{lemma}
-
-\begin{proof}
-Subgraphs of wellconstrained graph $G$ can only be trivial, underconstrained, or wellconstrained.
-So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and wellconstrained intersections are possible.
-\end{proof}
-
-\subsection{Lemmas for Intersection and Graph Structure Relationships}
-
-\begin{lemma}\label{wc-l2}
-% Given a wellconstrained $G$ and two wellconstrained vertex-maximal proper subgraphs $C_i, C_j$,
-$C_i\cup C_j$ is wellconstrained if and only if $C_i\cup C_j = C$.
-\end{lemma}
-
-\begin{proof}
-The forwards direction can be shown with a contradiction.
-If the $C_i\cup C_j$ was wellconstrained but $C_i\cup C_j \neq C$, we would have found a larger wellconstrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger.
-The two original subgraphs could not have been vertex-maximal proper.
-
-The backwards direction follows from the fact that $C$ is either some non-leaf node (wellconstrained by definition of a DR-plan) or $G$ itself (wellconstrained by definition of the problem).
-\end{proof}
-
-\begin{corollary}\label{wc-c1}
-% Given a wellconstrained $G$ and two wellconstrained vertex-maximal proper subgraphs $C_i, C_j$, l
-Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is wellconstrained, then there can be no edges in $G$ between the vertices of $R_i$ and $R_j$.
-\end{corollary}
-
-\begin{proof}
-It follows from the fact that $C_i\cup C_j$ must equal $C$ (proven in Lemma \ref{wc-l2}).
-\end{proof}
-
-% \begin{corollary}
-% Given the same conditions: $R_i$ and $R_j$ must be connected to $D_{i,j}=C_i\cap C_j=C\setminus R_i\setminus R_j$.
-% \end{corollary}
-
-% \begin{proof}
-% From Lemma \ref{wc-l1}, $D_{i,j}$ must be wellconstrained.
-% \end{proof}
-
-\begin{lemma}
-asdfas
-\end{lemma}
-
-\begin{proof}
-Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=C\setminus R_i\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
-Furthermore, $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.
-
-Then we assume that there is some third wellconstrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$).
-There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
-
-% \begin{enumerate}
-%     \item Assume $C_k\cup C_i \neq C$ and $C_k\cup C_j \neq C$.
-Then from Lemma \ref{wc-l2} $C_k\cup C_i$ is not wellconstrained and must therefore be underconstrained.
-Furthermore, $C_k\cup C_i=Idc(C,R'_i\cup C_i)$ because $C_k\neq C_i$.
-This means the subgraph $R'_i$ and its edges between $C_i$ increase the density overall because $C_i$ is wellconstrained.
-Similarly, the same is true for $R'_j$ and its edges.
-Since there can be no edges between $R'_i$ and $R'_j$ (COROLLARY, wouldn't be $C$!!!!) it means that all of the edges connecting them are incident to $D_{i,j}$.
-Therefore, $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ should be underconstrained.
-However, by Lemma \ref{uc-l1}, using the two underconstrained graphs we have that $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ so it must be trivial which is a contradiction, so $C_k$ could not be a wellconstrained vertex-maximal proper subgraph.
-%     % Similarly, we find the $C_k\cup C_j=Idc(C,R'_j\cup C_j)$ is underconstrained.
-However, $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ which is wellconstrained.
-This contradicts Lemma \ref{wc-l1} so $C_k$ could not be wellconstrained.
-%     \item Assume $C_k\cup C_i\neq C$ and $C_k\cup C_j= C$. Similar to the first case, we see that
-
-%     If it's just one and the other does equal C, then the other is not vertex-maximal?
-%     \item If it's neither... well then you either have C or the solution.
-% \end{enumerate}
-
-\begin{itemize}
-    \item 3 cases: $C_k$ cannot be $C=Idc(C,R_i\cup R_j\cup D_{i,j})$, $C_i=Idc(C,R_j\cup D_{i,j})$, or $C_j=Idc(C,R_i\cup D_{i,j})$. This is by definition.
-    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup D_{i,j})$, $Idc(C,R'_j\cup D_{i,j})$, $Idc(C, D_{i,j})$, $Idc(C,R_i\cup D'_{i,j})$, $Idc(C,R_j\cup D'_{i,j})$, $Idc(C,R'_i\cup D'_{i,j})$, $Idc(C,R'_j\cup D'_{i,j})$, $Idc(C, D'_{i,j})$, $Idc(C,R_i)$, $Idc(C,R_j)$, $Idc(C,R'_i)$, $Idc(C,R'_j)$, and $Idc(C,\emptyset)$.
-    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup R_j\cup D_{i,j})$ and $Idc(C,R_i\cup R'_j\cup D_{i,j})$ respectively.
-    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$...
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be wellconstrained.
-We also know it cannot be trivial because it contains wellconstrained subgraphs.
-This means it must be underconstrained.
-From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial.
-This is impossible because $D_{i,j}$ is wellconstrained, thereby contradicting the assumption that $C_k$ is wellconstrained.
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible...
-    %
-    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible...
-    %
-    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that is left!
-\end{itemize}
-
-The weight of $R_i$ has to be exactly equal to the weight of the edges that connect it to $D_{i,j}$. It must add 0 weight exactly since $D_{i,j}$ is wellconstrained.
-
-There can't be an intersection that is underconstrained between to wellconstrained graphs!
-
-\begin{itemize}
-    \item $C_k$ cannot be $Idc(C,R_i\cup R_j)$.
-From Corollary \ref{wc-c1}, $R_i$ and $R_j$ must not have any edges between them.
-Without loss of generality, we can say that $d(R_i)=l$ and $d(R_j)=k-l$.
-Because $C_i$ and $C_j$ are wellconstrained, and so is $D_{i,j}$, $R_i$ with its edges must add 0 density and $R_j$ with its edges must add 0 density.
-Since edges must have positive weight, $l\geq 0$ and $k-l\geq 0$ so $d(R_i)\in [0,k]$ and $d(R_j)\in [0,k]$ with $d(R_i)+d(R_j)=k$.
-This means, without loss of generality, that at least $R_i$ is trivial
-    %
-    % \item 1 case: $C_k=Idc(C,R_i\cup R_j)$ is not possible...
-    % \item 1 case: $C_k=Idc(C,R'_i\cup R'_j)$ is not possible...
-    % \item 2 cases: $C_k=Idc(C,R'_i\cup R_j)$ and $C_k=Idc(C,R_i\cup R'_j)$ are not possible...
-    % \item 4 cases: $C_k$ must contain a subgraph of $D_{i,j}$.
-Consider if it did not, then $C_k\cap D_{i,j}=\emptyset$ which is necessarily a trivial graph.
-By Lemma \ref{uc-l1}, this implies that $C_k\cup D_{i,j}$ is underconstrained, so $d(C_k\cup D_{i,j})>k$.
-Because there is no intersection, $d(C_k\cup D_{i,j})=d(C_k)+d(D_{i,j})$.
-However, we know that $D_{i,j}$ is wellconstrained, so $d(D_{i,j})=k$.
-This means that $d(C_k)$ must be greater than $k$ which makes it underconstrained, contradicting the assumption.
-This means $C_k$ cannot be $C_k=Idc(C,R_i\cup R_j)$, $C_k=Idc(C,R'_i\cup R_j)$, $C_k=Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$.
-    %
-    \item 4 cases: $C_k$ must contain a subgraph of $D_{i,j}$.
-Take $C_k=Idc(C,R_i\cup R_j)$.
-In this case, $C_k\cup C_i = R_i\cup C_i$ which does not equal $C$ because it is missing the edges between $R_i$ and $D_{i,j}$.
-We have from Lemma \ref{wc-l2} that this union cannot be wellconstrained.
-It cannot be trivial because it has wellconstrained subgraphs which leaves it to be underconstrained.
-We know $C_i$ is wellconstrained, therefore it must be $R_i$ that is underconstrained.
-This can be used to show that $R_j$ is also underconstrained.
-This means $R_i\cup R_j=C_k$ must be underconstrained, contradicting the assumption.
-This proof holds for several other cases, meaning $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$.
-    %
-    %
-    %
-    % \item $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-If this were wellconstrained the intersection with $C_i$ would be wellconstrained, which is $Idc(C,R'_j\cup D_{i,j})$.
-Similarly, it would imply
-    \item $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since this graph's union with $C_i$ is $Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that this union cannot be wellconstrained (and must therefore be underconstrained).
-We know $Idc(C,R_j\cup D_{i,j})=C_i$ is wellconstrained which implies $R'_j$ and the edges incident to it are increasing the density.
-Using the same process, we can show the same for $R'_i$.
-Since $D_{i,j}$ is wellconstrained, adding two components that increase the density would mean that $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is underconstrained, contradicting the assumption.
-    %
-    \item $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ and $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ are not possible.
-In both scenarios we have $C_k\cup C_i \neq C$ and $C_k\cup C_j \neq C$.
-From Lemma \ref{wc-l2} $C_k\cup C_i$ cannot be wellconstrained and must therefore be underconstrained.
-Furthermore, $C_k\cup C_i=Idc(C,R'_i\cup C_i)$ because $C_k\neq C_i$.
-This means the subgraph $R'_i$ and its edges between $C_i$ increase the density overall because $C_i$ is wellconstrained.
-Similarly, the same is true for $R'_j$ and its edges.
-Since there can be no edges between $R'_i$ and $R'_j$ (COROLLARY, wouldn't be $C$!!!!) it means that all of the edges connecting them are incident to $D_{i,j}$.
-Therefore, $Idc(C,R'_i\cup R'_j\cup D_{i,j})$ should be underconstrained.
-However, by Lemma \ref{uc-l1}, using the two underconstrained graphs we have that $(C_k\cup C_i)\cap (C_k\cup C_j)=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ so it must be trivial which is a contradiction, so $C_k$ could not be a wellconstrained vertex-maximal proper subgraph.
-    % \item $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible...
-\end{itemize}
-\end{proof}
-
-\pagebreak
-See ISSUE at top of paper, we have to assume we have a connected $G$...
-
-Take $C'_i$ to be $C\setminus C_i$.
-Then if $C_i\cup C_j$ is wellconstrained we have directly from Lemma \ref{wc-l2} for all $C'_i, C'_j$ there can be no edges between them (otherwise $C_i\cup C_j\neq C$).
-
-The set of edges connecting $C'_i$ to $C_i$ must have a weight that sums to exactly $d(C'_i)$ since we know that $d(C)=d(C_i)=k$.
-
-Any other $C_k$ must therefore contain $C'_i$ and $C'_j$ because it adds no density to ``tack it on'' to a graph that does not contain them already.
-
-Now we want to know something about $C_k$.
-If $C_k \neq C_i$, then $C'_i\not\subseteq C'_k$.
-It is clear that they cannot be equal otherwise $C_k = C_i$.
-However, it also cannot be a subset.
-There are two scenarios.
-(1) $C'_k$ does not contain vertices that are incident to $C'_i$.
-If this were so, we know we could ``tack on'' $C'_i$ to the wellconstrained $C_k$ without increasing the density and therefore $C_k$ could not have been maximal in the first place.
-(2) $C'_k$ contains vertices that are incident to $C'_i$.
-In this case
-
-Take $D_{i,j}$ to be $C\setminus (C'_i\cup C'_j)$.
-Clearly $D_{i,j}$ must be wellconstrained since ``tacking on'' $C'_i$ and $C'_j$ adds no density and $C$ is wellconstrained.
-Therefore, if there were some other $C_k$, it must contain $D_{i,j,k}\subset D_{i,j}$ such that $D_{i,j,k}=D_{i,j}\setminus C'_k$.
-It is clear that $D_{i,j,k}$ cannot equal $D_{i,j}$ or else $C_k$ must be $C_i$ or $C_j$ to be vertex-maximal.
 
@@
 \begin{proof}
-By Lemmas... the union must be under
-Lemma \ref{uc-l1} that $C_i\cup C_j$ must be underconstrained. Therefore, by the proof
-It follows from Lemma \ref{wc-l1} that the only way to get
- and its proof that if any intersections
+Suppose
 \end{proof}
-
-\subsection{Theorems for Optimal DR-Plans}
-
-\subsection{Union is Underconstrained}
 
@@
 \subsection{Union is Wellconstrained}
+
+\begin{lemma}
+Given two wellconstrained subgraphs $C_i,C_j$, $C_i\cup C_j$ is wellconstrained if and only if $C_i\cap C_j$ is wellconstrained.
+\end{lemma}
+
+\begin{proof}
+In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial.
+The intersection also cannot be underconstrained.
+If $C_i \cap C_j$ were underconstrained, then for $C_i$ to be wellconstrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
+The same is true for $C_j$.
+If both of these were satisfied, then the union of $C_i$ and $C_j$ would result in an overconstrained nontrivial graph.
+This is a contradiction, by definition of a wellconstrained graph.
+Therefore, the intersection can only be wellconstrained.
+
+BACKWARDS!!!!!
+\end{proof}
+
+\begin{lemma}\label{wc-l2}
+Given two wellconstrained vertex-maximal proper subgraphs $C_i, C_j$, $C_i\cup C_j$ is wellconstrained if and only if $C_i\cup C_j = C$.
+\end{lemma}
+
+\begin{proof}
+The forwards direction can be shown with a contradiction.
+If the $C_i\cup C_j$ was wellconstrained but $C_i\cup C_j \neq C$, we would have found a larger wellconstrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger.
+The two original subgraphs could not have been vertex-maximal proper.
+
+The backwards direction follows from the definition of a DR-plan.
+\end{proof}
 

### 2014-12-03 doc/GMP2015/paper.tex
Changes to paper.

@@
 \documentclass[submission]{gmp2015}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\usepackage{tikz}
+\usetikzlibrary{shapes,snakes}
+\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
+
+\usepackage{ulem}
+
+\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \maketitle
-
-\begin{abstract}
-Super awesome abstract.
-\end{abstract}
 
@@
 
-\section{Introduction}
+\input{s00_abstract}
 
-Nulla commodo posuere augue ut aliquet.
-Morbi at magna sem.
-Phasellus vitae varius massa.
-Nam eget quam sed erat euismod malesuada.
-Aliquam id velit metus, et porta lectus.
-Donec convallis massa quis eros accumsan consectetur.
+\input{s01_introduction}
 
-Phasellus sit amet aliquam nulla.
-Vestibulum pretium lectus in ligula tempor convallis.
-Integer accumsan enim vel ante sollicitudin ultrices.
-Proin adipiscing, elit quis rhoncus egestas, lectus diam semper purus, sed luctus turpis felis eget magna.
-Proin eget dolor fringilla justo malesuada imperdiet sed sed erat.
-Sed sem dolor, egestas eleifend hendrerit dignissim, sagittis id quam.
-Aenean et nisl eu mi imperdiet mollis sit amet eu nunc.
-Nunc at turpis eget sapien eleifend lobortis at vitae nunc.
-Mauris a dui justo.
-Mauris suscipit quam id elit congue sed volutpat purus dapibus.
-Etiam eu tellus a enim lobortis posuere.
-Fusce feugiat sapien eu metus ullamcorper eleifend.
-Integer vehicula eros vel massa pharetra euismod.
-Vivamus a lacus ut sem condimentum imperdiet.
-Maecenas scelerisque vulputate enim, a cursus felis vehicula et.
-Morbi viverra, ipsum sed commodo lacinia, sapien elit tempus lectus, et consequat leo turpis quis odio.
+\input{s02_preliminaries}
 
-Nulla eu lorem turpis.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Phasellus tellus elit, lobortis non euismod sed, ultrices vel magna.
-Vivamus condimentum erat ac erat mollis ut convallis nulla vehicula.
-Proin sit amet magna metus, quis imperdiet mi.
-Proin diam libero, laoreet ut auctor et, commodo quis nisi.
-Curabitur commodo nibh non sapien molestie congue.
-Etiam et mi non nibh mattis auctor ac in mauris.
-Donec ipsum urna, accumsan nec fermentum a, eleifend a elit.
-Aliquam velit justo, hendrerit sed luctus ac, luctus id urna.
-Aenean viverra blandit augue, a viverra arcu dictum nec.
-Nunc malesuada, mauris sit amet gravida euismod, magna ipsum porta mi, ut sodales neque tortor vel ligula.
-Aliquam a ipsum a mi molestie fermentum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Fusce porttitor ligula eu mauris auctor sit amet porttitor odio ultrices.
+\input{s03_drp}
+\input{s04_recombination}
+\input{s05_solnspacenav}
 
-Aliquam sit amet eros vel elit facilisis sodales.
-Sed pretium lacinia ante venenatis convallis.
-Etiam iaculis, risus in mattis viverra, sem orci sodales orci, tempus pretium enim est a augue.
-Suspendisse vitae est a orci lobortis dictum.
-Pellentesque sed magna ut tellus accumsan vestibulum.
-Curabitur ultricies mollis ante, interdum mollis metus ultricies a.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean vel magna ipsum.
+\input{s06_conclusion}
 
-\begin{figure}\centering
-  \includegraphics[scale=.5]{example.eps}
-  \caption{Example of a normal figure that spans one column.}
-  \label{fig:square}
-\end{figure}
+% \begin{figure}\centering
+%   \includegraphics[scale=.5]{example.eps}
+%   \caption{Example of a normal figure that spans one column.}
+%   \label{fig:square}
+% \end{figure}
 
-Pellentesque commodo adipiscing imperdiet.
-Proin ultricies, quam vitae semper feugiat, eros diam pulvinar tortor, id venenatis ligula purus non erat.
-Mauris pulvinar velit sit amet mauris elementum bibendum.
-Sed mi leo, suscipit nec pharetra a, cursus id libero.
-Phasellus sit amet sapien diam, vitae scelerisque eros.
-Etiam vel risus lectus.
-Etiam molestie euismod turpis, condimentum ultrices dui euismod et.
-
-Donec quis odio a dolor venenatis dictum nec eu dui.
-Fusce eleifend nibh et massa convallis posuere.
-Sed justo quam, vulputate sollicitudin pulvinar in, dapibus sit amet urna.
-Vivamus metus justo, molestie et volutpat a, sagittis ac arcu.
-Duis in ante ut quam condimentum cursus.
-Cras purus lacus, dictum vel auctor id, luctus in felis.
-Praesent hendrerit adipiscing lorem vitae fringilla.
-Aenean bibendum fermentum tortor, eget commodo ante auctor vel.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Related Work}\label{sec:relatedwork}
-
-Vestibulum adipiscing congue orci eu commodo.
-Aliquam consectetur nibh nisl, ac mattis mauris.
-Sed vestibulum venenatis tincidunt.
-Phasellus quis metus id justo porta aliquet.
-Maecenas et accumsan nisi.
-Cras malesuada, urna vel venenatis adipiscing, lorem nunc pharetra dolor, ut convallis nisl sapien id eros.
-Cras commodo, elit a sollicitudin ultricies, erat mi rutrum nisi, in laoreet purus lorem laoreet nibh.
-Proin tincidunt, purus sit amet faucibus ornare, justo diam posuere dolor, id sollicitudin odio leo eget mauris.
-Nam mi ante, laoreet a adipiscing vitae, ultricies vitae dolor.
-Quisque auctor ornare iaculis.
-Praesent elit urna, porttitor vitae convallis vel, condimentum eu neque; see Figure~\ref{fig:square}.
-
-Duis sollicitudin pulvinar ultrices~\cite{Peters:1997:BBC}. Praesent tincidunt interdum dignissim. Aenean vehicula nisl accumsan purus suscipit volutpat pharetra sapien ornare.
-%
-Example of an enumerated list:
-\begin{enumerate}
-  \item First item,
-  \item Second item.
-\end{enumerate}
-
-Aenean id porttitor mi.
-Integer nec nisl elit, id dictum diam~\cite{Kazhdan:2008:SMF}.
-In consequat imperdiet lacus, quis placerat felis dapibus at.
-In at ligula ante, sollicitudin tristique nisi.
-Donec commodo justo eget urna luctus convallis.
-Praesent id diam eros, eu laoreet velit.
-Vivamus non tellus id nulla cursus sollicitudin.
-
-\begin{figure*}\centering
-\includegraphics{example.eps}
-\caption{Example of a figure that spans both columns.}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{The new stuff}
-
-Nam gravida adipiscing imperdiet.
-Fusce commodo pharetra faucibus.
-Donec feugiat, metus vel vulputate blandit, dolor nisi ornare libero, eu faucibus dui tortor sed turpis.
-Nullam porta, orci eu ullamcorper pretium, mauris elit convallis ante, sed gravida urna sapien eu quam.
-Morbi et quam purus.
-Nulla aliquam ultricies felis, a viverra eros scelerisque a.
-Fusce orci orci, dignissim a pellentesque sit amet, elementum eget sapien.
-Suspendisse potenti.
-In hac habitasse platea dictumst.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec et nisi dapibus metus mollis tincidunt.
-Aliquam rutrum sagittis mauris, non pulvinar ipsum sodales a.
-
-\begin{theorem}\label{theorem:pi}
-Aliquam nec sapien sit amet diam molestie tristique vitae sit amet libero,
-\begin{equation}\label{eq:pi}
-\pi = \sqrt{12}\sum^\infty_{k=0} \frac{(-3)^{-k}}{2k+1}.
-\end{equation}
-\end{theorem}
-
-\begin{proof}
-Curabitur metus lorem, rhoncus nec ullamcorper quis, varius eget neque.
-Phasellus et turpis quis massa porta tincidunt.
-Donec fringilla luctus libero, at adipiscing justo ullamcorper non.
-Mauris magna ipsum, semper eget hendrerit nec, ornare sit amet neque.
-Nullam elementum arcu eget tellus porta mattis.
-\end{proof}
-
-\noindent
-Morbi ut lorem lectus.
-Etiam quis dolor justo, according to~\eqref{eq:pi}.
-Duis suscipit laoreet sollicitudin.
-Phasellus lacinia adipiscing tortor at posuere.
-Integer eget neque ac lacus vestibulum sollicitudin.
-Mauris vitae lacus dui.
-Aenean ultrices iaculis faucibus.
-Suspendisse congue, lectus in vulputate convallis, magna eros gravida mi, a molestie metus urna a nibh.
-Suspendisse adipiscing gravida ullamcorper.
-%
-Example of an itemized list:
-\begin{itemize}
-  \item First item,
-  \item Second item.
-\end{itemize}
-
-\noindent
-Nullam eu justo nibh.
-Vivamus id vehicula elit.
-Nam pellentesque rhoncus mi, id euismod arcu congue quis.
-Vestibulum sem urna, hendrerit non semper vel, facilisis vel sem.
-Ut sit amet massa nisi, vitae tristique metus.
-Duis venenatis massa quis sapien ullamcorper sagittis~\cite{Li:2010:PSO}.
-Ut auctor viverra fringilla.
-Donec at metus ligula.
-Nulla semper viverra diam, quis lobortis justo rhoncus ut.
-Aenean eget nisi sed dui iaculis iaculis quis ut libero.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse viverra rutrum ante pulvinar placerat.
-Nunc gravida laoreet tempus.
-Quisque non lobortis sem.
-In convallis fermentum odio, quis tincidunt dui lacinia vel.
-
-\begin{corollary}
-Etiam quis mauris orci. Nam blandit elementum diam a molestie. Suspendisse ultricies auctor urna, quis sagittis metus semper vel. Cras vel nisi felis.
-\end{corollary}
-
-\begin{proof}
-Ut varius interdum pulvinar. Vivamus luctus consequat rutrum. Nulla facilisi. Mauris sit amet eros sit amet justo feugiat facilisis, which follows from Theorem~\ref{theorem:pi}.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
-
-Quisque id risus mi.
-Nunc vel bibendum nulla.
-Curabitur nec quam est.
-Quisque commodo leo quis ligula imperdiet vitae gravida orci eleifend.
-Vestibulum sed mollis turpis.
-Donec feugiat facilisis lorem, eget hendrerit arcu dignissim sit amet.
-Suspendisse potenti.
-Vivamus et felis leo, as shown in Section~\ref{sec:relatedwork}.
-
-Duis consequat neque id libero eleifend a cursus massa sodales.
-Cras condimentum nulla ac neque dictum varius.
-Quisque vitae arcu neque, in placerat risus.
-Sed pellentesque iaculis est, non tincidunt augue sollicitudin id.
-Aliquam consequat, massa vel faucibus cursus, ipsum quam ultricies lectus, et hendrerit lectus arcu sed enim.
-Aliquam vitae enim nulla, laoreet fringilla velit.
-Phasellus eget diam enim.
-
-Fusce iaculis lacus sed eros cursus euismod.
-Proin nec elit lectus.
-Mauris tristique accumsan risus, nec adipiscing odio viverra nec.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed a vestibulum eros.
-Curabitur viverra pretium risus, lacinia feugiat tellus tincidunt ut.
-Ut sit amet justo nec ipsum ultrices pharetra.
-Curabitur consectetur nisi nec dui tincidunt varius.
-Nunc vestibulum sollicitudin sapien, eu aliquam orci porta id.
-Etiam porttitor felis quis ante sodales euismod.
-Nam tempor leo vel justo tincidunt vehicula.
-
-Vivamus quis sagittis leo.
-Nam vel purus risus, ultrices lacinia odio.
-Sed sagittis consequat luctus.
-Integer nulla ante, tincidunt sed pellentesque a, blandit vitae velit.
-Nulla facilisi.
-Donec id lacus ut justo pharetra consectetur nec nec enim.
-Praesent ut nisl vel metus facilisis commodo.
-Curabitur tristique varius ultricies.
-Praesent vitae mauris odio, vel aliquet metus.
-Suspendisse tempus erat eget sem lobortis accumsan.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Ut facilisis quam eu tortor fringilla non accumsan mi fringilla.
-Donec quis leo nec odio elementum tristique ac at enim.
-
-Etiam quis diam ut nibh viverra iaculis.
-Fusce vehicula, mauris eu convallis consequat, turpis nunc interdum urna, eu malesuada erat justo non velit.
-In vel lectus id nunc varius cursus.
-Nullam vitae odio eget urna laoreet aliquam vel semper nisi.
-Mauris venenatis arcu eget nisi vehicula ac pharetra enim vehicula.
-Nullam et ipsum tortor, eu volutpat diam.
-Morbi adipiscing commodo rhoncus.
-Morbi urna lorem, volutpat non aliquam a, fermentum non diam.
+% \begin{figure*}\centering
+%     \includegraphics{example.eps}
+%     \caption{Example of a figure that spans both columns.}
+% \end{figure*}
 

### 2014-12-03 doc/GMP2015/s00_abstract.tex
Changes to paper.

\begin{abstract}
Super awesome abstract.
\end{abstract}

### 2014-12-03 doc/GMP2015/s01_introduction.tex
Changes to paper.

\section{Introduction}
Computational design and analysis of constrained geometric configuration spaces of 2-dimensional, multilayered, aperiodic or self-similar materials and biological structures.

=======================

Many natural and engineered materials are nearly rigid, 2D multilayers, possibly consisting of inter-constraining monolayers each of which is topologically planar (genus 0). They are often aperiodic, recursively self-similar or both \cite{Intro1}, enabling formation of diverse, larger structures with multiple functions.

Examples include:

\begin{itemize}
    \item (cross-sections) of ``necklaces'' of microtubules \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}

    \item interlinked monolayers of cellulose microfibrils e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}

    \item Inorganic examples include silica and glass \cite{SilicaGlass1} \cite{SilicaGlass2}

    \item (bi)layers and other structures that behave like assemblies of 2D particles under non-overlap constraints, i.e, jammed disks on the plane \cite{JammedDisk1}

    \item More recent, engineered examples include disordered graphene layers \cite{Graphene1} \cite{Graphene2}  even reinforced by  microfibrils, and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet bringing other structures \cite{Microfibrils2} within reach.
\end{itemize}

Understanding the structural and mechanical properties of these materials and structures at multiple scales, including rigidity, flexibility and motion is best achieved in the tradition of computer-aided mechanical design:  by modeling the structure as a configuration of appropriate geometric primitives inter-constrained by metric or algebraic constraints; and designing the constraint-system to achieve the desired properties and their robustness/sensitivity to constraint-system variations.

\subsection{Contributions}

The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp. 2DGMCS) belong to a combinatorial class for which algorithms with formally proven efficiency guarantees exist for  decomposing into maximal generically well-constrained components \todo{[Citation Needed]}. However, these systems do not belong in the largest combinatorial class of 2D systems, whose configuration spaces have formally proven  characterizations and algorithmic guarantees (e.g. for algebraic complexity, path and volume measures, and  topology).

Specifically, these results \todo{[Citation Needed]} rely on the fact that rigid systems in this  class, in contrast to GMCS, have a straightforward and immediate {\sl optimal} recursive decomposition, (so-called {\sl optimal} DR-plan) and  recombination or realization (with a formal disambiguation by minimal orientation specification).

It is known \todo{[Citation Needed]} that already the optimal DR-planning problem for general 2DGMCS is NP-hard, intuitively because of a combinatorial explosion of polynomially-sized DR-plans.  In this paper, we  navigate this barrier by defining a so-called {\it canonical} DR-plan and showing that all canonical DR-plans for non-overconstrained 2DGMCS are optimal. We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration. Software implementation can be found at: \todo{Provide Link}.

\subsection{Previous work}

Computer aided mechanical design represents a mature application of geometric constraint systems,  leading to and borrowing from deep and fundamental problems and techniques in distance geometry, convex analysis, combinatorics, algebra, topology and complexity theory.

Broadly, these problems can be classified into three types that draw upon each other, and possess mathematical, algorithmic, and software implementation aspects: (i) combinatorial (generic) minimal rigidity (or isostaticity or well-constrained-ness) characterization, (ii) (optimal) recursive decomposition (DR plans), recombination and realization, (iii) configuration space characterization (topology, complexity, path and volume measures).

In general, proofs of combinatorial (generic) minimal rigidity characterizations for geometric constraint systems rely on underlying matroids \todo{[Citation Needed]} In these cases, more-or-less greedy algorithms such as simplified network-flow, or so-called pebble games can be used.

Nothing nontrivial known until now for optimal DR plans, although efficient algorithms are known for DR-plans satisfying various other formal properties.

Recombination and realization navigation of realizations (disambiguation with orientation, faster with overconstraints, mention also totally different distance geometry method - low rank PSD matrix completions SDP/SVD)

Gelfand/McPherson, Walker's conjecture solution, Convex Cayley, EASAL and Caymos for low Cayley complexity and bijective representation using global rigidity.

### 2014-12-03 doc/GMP2015/s02_preliminaries.tex
Changes to paper.

--- 2014-12-02 doc/General_Work/DRP_2d.tex
@@
-% --------------------------------------------------------------
-%                       Initialize
-% --------------------------------------------------------------
-
-\documentclass[11pt]{article}
-
-% \usepackage{multicol}
-\usepackage{ulem}
-\usepackage{comment}
-\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}
-
-\usepackage{personalNotes}
-% contains special macros for math environment:
-%     unit, p, set, abs, floor, ceil, vec, norm, inner, kernel
-% contains useful all-purpose macros:
-%     n, lessSeparation, append
-
-\usepackage{tikz}
-\usetikzlibrary{shapes,snakes}
-\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
-
-% --------------------------------------------------------------
-%                    Beginning of Document
-% --------------------------------------------------------------
-
-\begin{document}
-% \begin{multicols}{2}
-
-% --------------------------------------------------------------
-%                           Title
-% --------------------------------------------------------------
-
-\pntitle {Decomposition Recombination Plans for 2D Well-Constrained Graphs}
-\pnauthor{Troy Baker}
-\pndate  {May 13, 2014}
-
-\pnmaketitle
-\pnmakefooter
-
-% --------------------------------------------------------------
-%                           Body
-% --------------------------------------------------------------
-
-\section{Introduction}
-\todo{
-Other sections...
-\begin{itemize}
-    \item Intro and motivation (previous work, where this work leads, examples)
-    \item Preliminaries
-    \item DR-Plan (Canonical) + algo
-    \item Recombination + algo (refer to other paper and make 2D)
-    \item Solution space navigation
-    \item Conclusion
-\end{itemize}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Preliminaries}
@@
 \subsection{Definitions}
-% ISSUE: My definitions don't prevent non-positive densities... This means things aren't guaranteed to be connected...
 
-\begin{definition}[Graph]
-Graph $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$.
+In the context of this paper, a \textbf{graph} $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
 Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
-\end{definition}
 
-\begin{definition}[Density]
-The density of a graph is defined as:
-\[d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}\]
-% \[d(G) = \max\set{0,\sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}}\]
-\end{definition}
+The \textbf{density} of graph $G$ is
+\begin{equation}
+    d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}
+\end{equation}
 
-\begin{definition}[Independent or Sparse Graph]
-\todo{Needs definition.}
-\end{definition}
+A graph is \textbf{independent} (or \textbf{sparse}) if \todo{define}.
 
-\begin{definition}[Over-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is over-constrained if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$.
-(where S is not trivial...
-This seems weird because you could have an over-constrained under-constrained graph...)
-\end{definition}
+A graph is \textbf{over-constrained} if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$.
 
-\begin{definition}[Well-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
-\end{definition}
+A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
 
-% \begin{definition}[Well-constrained Graph]
-% Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either under-constrained, well-constrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
-% \end{definition}
+A graph is \textbf{rigid} if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
 
-\begin{definition}[Rigid Graph]
-Graph $G=(V,E,w_V,w_E)$ is rigid if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-\end{definition}
+A graph is \textbf{under-constrained} if, given constant $k$, $G$ is not rigid.
 
-\begin{definition}[Under-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, $G$ is not rigid.
-\end{definition}
-
-% \begin{definition}[Under-constrained Graph]
-% Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, there exists some subgraph $G'\subseteq G$ such that $d(G') < k$.
-% \end{definition}
-
-\begin{definition}[Trivial Graph]
-A trivial graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
-
-\todo{Maybe leave this out until it's needed later?}
+A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
+\todo{Maybe leave the following out until it's needed later?}
 In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
@@
 These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
-\end{definition}
 
-\begin{definition}[Decomposition Recombination (DR-) Plan]
-The DR-plan of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
+The \textbf{decomposition-recombination (DR-) plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
 \begin{enumerate}
@@
 Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
-\end{definition}
 
-\begin{definition}[Complete DR-Plan]
-The complete DR-plan of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \textit{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \textit{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
 By this, rule 3 is implicit for a complete DR-plan.
-\end{definition}
 
-\begin{definition}[Optimal DR-Plan]
-The optimal DR-plan of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
 
 \subsection{Notation}
+
 $G$ is taken to be a graph with values $(V,E,w_V,w_E)$.
@@
 By definition of a DR-plan, it is implied that $C_i$ is a well-constrained vertex-maximal proper subgraph.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{The Canonical Optimal DR-Plan}
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (1,1) {$R_j$};
-    \node[v] (v2) at (3,1) {$R_k$};
-    \node[v] (v3) at (2,2) {$D$};
-    \node[v] (v4) at (2,3.3) {$R_i$};
-
-    \tedge{v1}{v3}{solid}{}{above};
-    \tedge{v2}{v3}{solid}{}{above};
-    \tedge{v3}{v4}{solid}{}{right};
-\end{tikzpicture}
-\ldots with weights \ldots
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (1,1) {y};
-    \node[v] (v2) at (3,1) {z};
-    \node[v] (v3) at (2,2) {k};
-    \node[v] (v4) at (2,3.3) {x};
-
-    \tedge{v1}{v3}{solid}{y}{above};
-    \tedge{v2}{v3}{solid}{z}{above};
-    \tedge{v3}{v4}{solid}{x}{right};
-
-    % \draw (3.3,2.5) circle (1.9cm);
-    % \draw (.7,2.5) circle (1.9cm);
-    % \draw (2,.5) circle (1.9cm);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxy};
-    \node[v] (v3) at ( 1,-1) {kyz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v5}{solid}{}{};
-    \tedge{v3}{v5}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxy};
-    \node[v] (v3) at ( 1,-1) {kxz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v5}{solid}{}{};
-    \tedge{v3}{v4}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxz};
-    \node[v] (v3) at ( 1,-1) {kyz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v6}{solid}{}{};
-    \tedge{v3}{v5}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
-% \\
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at ( 0, 0) {kxyz};
-%     \node[v] (v2) at (-1,-1) {kxy};
-%     \node[v] (v3) at (-2,-2) {kx};
-%     \node[v] (v4) at (-3,-3) {k};
-%     \node[v] (v5) at ( 1,-1) {z plan};
-%     \node[v] (v6) at ( 0,-2) {y plan};
-%     \node[v] (v7) at (-1,-3) {x plan};
-%     \node[v] (v8) at (-2,-4) {k plan};
-
-%     \tedge{v1}{v2}{solid}{}{};
-%     \tedge{v1}{v5}{solid}{}{};
-%     \tedge{v2}{v3}{solid}{}{};
-%     \tedge{v2}{v6}{solid}{}{};
-%     \tedge{v3}{v4}{solid}{}{};
-%     \tedge{v3}{v7}{solid}{}{};
-%     \tedge{v4}{v8}{solid}{}{};
-% \end{tikzpicture}
-\end{center}
-
-\caption{A well-constrained graph whose nodes are subgraphs labeled with their density.
-Below are its 3 optimal DR-Plans (partially complete).
-The center vertex has weight $k$ and is well-constrained itself.
-The radial nodes have weights $x,y,z \neq k$ with equivalent edge weights, making any union between $k$ and the others well-constrained.
-\todo{Make this figure more clear.}}
-\label{3-plans}
-\end{figure}
-
-In the case of a well-constrained graph, it is clear that there always exists a unique vertex-maximum DR-Plan that includes every single well-constrained vertex-maximal proper subgraph at each node, which we call the complete DR-plan.
-However, in the problem of the optimal DR-Plan there is no such unique plan.
-For instance, a graph that is formed by the union of 3 well-constrained subgraphs that all intersect on a well-constrained subgraph, demonstrated in Figure \ref{3-plans}.
-In this case, it results in $3$ unique optimal DR-Plans.
-Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
-Therefore, all choices of decomposition are in some sense equivalent.
-The theorem we seek to prove is thus:
-
-\begin{theorem}\label{t1}
-Given a well-constrained 2-dimensional bar-joint graph $G$, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
-\begin{enumerate}
-    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
-    \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
-\end{enumerate}
-\end{theorem}
-
-We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
-By definition, any subgraph of a well-constrained graph will either be (1) trivial, (2) under-constrained, or (3) well-constrained.
-We will analyze all 3 possibilities in turn.
-These observations will lead to interesting properties that govern the behavior of optimal DR-plans on well-constrained graphs.
-
-% We show this by carefully considering all of the potential cases of the unions of subgraphs.
-By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained, or (3) trivial.
-We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
-For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
-We wish to prove the following bijections:
-
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline
-\textbf{Union is\ldots} & \textbf{Intersection is\ldots} \\ \hline
-\sout{trivial}          & --- \\ \hline
-under-constrained       & trivial \\ \hline
-well-constrained        & well-constrained \\ \hline
----                     & \sout{under-constrained} \\ \hline
-\end{tabular}
-\end{center}
-
-\begin{lemma}\label{t-l1}
-$F_i\cup F_j$ cannot be trivial.
-\end{lemma}
-
-\begin{proof}
-% If $F_i,F_j$ are well-constrained, then $F_i\cup F_j \geq k$. Trivial graphs must be over-constrained.
-
-Follows from the definition of a trivial graph. It would imply that the well-constrained graphs were also trivial since they would be subgraphs of a trivial graph.
-\end{proof}
-
-\begin{lemma}\label{uc-l1}
-$F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
-\end{lemma}
-
-\begin{proof}
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction, we have that $d(F_i\cup F_j)=l<k$.
-If it were equal to $k$ it would be well-constrained since, as a subgraph of $F$, it satisfies the other properties of a well-constrained graph; if it were greater than $k$ then it would be over-constrained.
-Therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=2k-l>k$.
-The only way to have an over-constrained subgraph in a well-constrained graph is if it is trivial.
-
-In the reverse direction, we know that $d(F_i\cap F_j)>k$ because it is trivial. Thus, by the same math, it is clear that $d(F_i\cup F_j)<k$ and it must be under-constrained.
-\end{proof}
-
-\begin{lemma}\label{wc-l1}
-$F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
-\end{lemma}
-
-\begin{proof}
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$.
-Since $F_i\cap F_j$ is a subgraph of $F$, which is well-constrained, it is also well-constrained.
-
-In the reverse direction the same math proves that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it too is well-constrained.
-
-% \todo{Integrate this}
-% Furthermore, all subgraphs of $F_i\cup F_j$ and $F_i\cap F_j$ are also subgraphs of $F_i$ and $F_j$, therefore all properties of well-constrained graphs are maintained.
-
-% In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial.
-The intersection also cannot be under-constrained.
-If $F_i \cap F_j$ were under-constrained, then for $F_i$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
-The same is true for $F_j$.
-If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an over-constrained nontrivial graph.
-This is a contradiction, by definition of a well-constrained graph.
-Therefore, the intersection can only be well-constrained.
-
-% reverse!!!!!
-\end{proof}
-
-\begin{lemma}\label{iuc-l1}
-$F_i\cap F_j$ cannot be under-constrained.
-\end{lemma}
-
-\begin{proof}
-Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained.
-So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and well-constrained intersections are possible.
-\end{proof}
-
-% WAIT... WHAT ABOUT TWO BANANAS? The intersection is 2 vertices which is under-constrained... oh nevermind... that's well-constrained
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Inferring Graph Structure from Union and Intersection Properties}
-
-\begin{lemma}\label{wc-l2}
-% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$,
-$C_i\cup C_j$ is well-constrained if and only if $C_i\cup C_j = C$.
-\end{lemma}
-
-\begin{proof}
-The forwards direction can be shown with a contradiction.
-If the $C_i\cup C_j$ was well-constrained but $C_i\cup C_j \neq C$, we would have found a larger well-constrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger.
-The two original subgraphs could not have been vertex-maximal proper.
-
-The reverse direction follows from the fact that $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
-\end{proof}
-
-\begin{corollary}\label{wc-c1}
-% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$, l
-Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
-\end{corollary}
-
-\begin{proof}
-It follows from the fact that $C_i\cup C_j$ must equal $C$ (proven in Lemma \ref{wc-l2}).
-\end{proof}
-
-At this point we limit our discussion to the scope of 2D graphs.
-All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$.
-Trivial graphs are a single vertex and empty set.
-
-Also, 2D well-constrained graphs must be connected.
-% \todo{Is this actually true?}
-The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
-The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
-
-\begin{lemma}\label{wc-l3}
-If $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
-Alternatively, if $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
-\end{lemma}
-
-\begin{proof}
-Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
-Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.
-
-Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$).
-There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
-
-\begin{itemize}
-    \item 3 cases: $C_k$ cannot be $C=Idc(C,R_i\cup R_j\cup D_{i,j})$, $C_i=Idc(C,R_j\cup D_{i,j})$, or $C_j=Idc(C,R_i\cup D_{i,j})$. This is by definition.
-
-    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup D_{i,j})$, $Idc(C,R'_j\cup D_{i,j})$, $Idc(C, D_{i,j})$, $Idc(C,R_i\cup D'_{i,j})$, $Idc(C,R_j\cup D'_{i,j})$, $Idc(C,R'_i\cup D'_{i,j})$, $Idc(C,R'_j\cup D'_{i,j})$, $Idc(C, D'_{i,j})$, $Idc(C,R_i)$, $Idc(C,R_j)$, $Idc(C,R'_i)$, $Idc(C,R'_j)$, and $Idc(C,\emptyset)$.
-
-    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup R_j\cup D_{i,j})$ and $Idc(C,R_i\cup R'_j\cup D_{i,j})$ respectively.
-
-    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary \ref{wc-c1}) and cannot be well-constrained.
-
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained.
-We also know it cannot be trivial because it contains well-constrained subgraphs.
-This means it must be under-constrained.
-From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial.
-This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
-
-    \item 1 case: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible.
-The union with $C_i$ (or $C_j$) is not equal to $C$.
-Therefore, the union must be under-constrained which means the intersection must be trivial (a single node).
-However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.
-
-    \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
-
-    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that is left!
-\end{itemize}
-
-Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.
-
-\todo{Needs a figure.}
-\end{proof}
-
-% \pnplaceholder
-% WAIT, TRIVIAL GRAPH DEFINITION DOESN'T WORK!!!! in 4D it's the 3 clique, but 3 vertices is a subgraph of that... that is certainly not over-constrained...
-
-% NOT SO! -5C2 = -10. 3 edges and 3 nodes -> 3 - 3*4 = -9... over-constrained!
-
-% \pnplaceholder
-
-\begin{lemma}\label{uc-l2}
-% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial (a vertex or empty set).
-If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
-\end{lemma}
-
-\begin{proof}
-If any intersection was not trivial, then it must be well-constrained.
-If $C_k\cap C_l$ were well-constrained, then by Lemma \ref{wc-l3} all of the intersections are well-constrained.
-This contradicts the statement that $C_i\cap C_j$ is trivial.
-% It follows that if the intersection is trivial then the union is under-constrained (Lemma \ref{uc-l1}). If the union is under-constrained, then the
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Proving the Main Theorem}
-Lemmas \ref{wc-l3} and \ref{uc-l2} show that $CompleteDRP(C)$ will result in one of two scenarios.
-(A) Either all $C_1,\ldots,C_N$ intersect on some common well-constrained subgraph and any two union to make $C$, or (B) all $C_1,\ldots,C_N$ intersect trivially and all of them must be unioned to make $C$.
-
-% \begin{theorem}
-% In case (A), any two unique wcvmps of $C$ can be used as the children of $C$ in $OptimalDRP(C)$.
-% \end{theorem}
-
-% \begin{proof}
-
-% \end{proof}
-
-% \begin{theorem}
-% In case (B), all wcvmps of $C$ must be used as the children of $C$ in $OptimalDRP(C)$.
-% \end{theorem}
-
-% \begin{proof}
-% If not all of them are used, the union will be under-constrained. $C$ is well-constrained, thus using not all of them is not an option.
-% \end{proof}
-
-% \pnplaceholder
-
-% \pagebreak
-% See ISSUE at top of paper, we have to assume we have a connected $G$...
-
-% Take $C'_i$ to be $C\setminus C_i$.
-Then if $C_i\cup C_j$ is well-constrained we have directly from Lemma \ref{wc-l2} for all $C'_i, C'_j$ there can be no edges between them (otherwise $C_i\cup C_j\neq C$).
-
-% The set of edges connecting $C'_i$ to $C_i$ must have a weight that sums to exactly $d(C'_i)$ since we know that $d(C)=d(C_i)=k$.
-
-% Any other $C_k$ must therefore contain $C'_i$ and $C'_j$ because it adds no density to ``tack it on'' to a graph that does not contain them already.
-
-% Now we want to know something about $C_k$.
-If $C_k \neq C_i$, then $C'_i\not\subseteq C'_k$.
-It is clear that they cannot be equal otherwise $C_k = C_i$.
-However, it also cannot be a subset.
-There are two scenarios.
-(1) $C'_k$ does not contain vertices that are incident to $C'_i$.
-If this were so, we know we could ``tack on'' $C'_i$ to the well-constrained $C_k$ without increasing the density and therefore $C_k$ could not have been maximal in the first place.
-(2) $C'_k$ contains vertices that are incident to $C'_i$.
-In this case
-
-% Take $D_{i,j}$ to be $C\setminus (C'_i\cup C'_j)$.
-Clearly $D_{i,j}$ must be well-constrained since ``tacking on'' $C'_i$ and $C'_j$ adds no density and $C$ is well-constrained.
-Therefore, if there were some other $C_k$, it must contain $D_{i,j,k}\subset D_{i,j}$ such that $D_{i,j,k}=D_{i,j}\setminus C'_k$.
-It is clear that $D_{i,j,k}$ cannot equal $D_{i,j}$ or else $C_k$ must be $C_i$ or $C_j$ to be vertex-maximal.
-
-% \begin{lemma}
-% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial or empty set.
-% \end{lemma}
-
-% \begin{proof}
-% By Lemmas... the union must be under
-% Lemma \ref{uc-l1} that $C_i\cup C_j$ must be under-constrained. Therefore, by the proof
-% It follows from Lemma \ref{wc-l1} that the only way to get
-%  and its proof that if any intersections
-% \end{proof}
-
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-    \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
-    \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
-    \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);
-
-    % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
-    % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);
-
-    % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
-    % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
-    \tikzstyle{e}=[]
-
-    \node[el] (m) at (0,    0) {k};
-    \node[v]  (l) at (-1.5, 0) {b};
-    \node[v]  (r) at (1.5,  0) {c};
-
-    \node[el] (tl) at (-1,  1) {i};
-    \node[el] (tr) at (1,   1) {j};
-    \node[v]  (t)  at (0,   2) {a};
-
-    % \node[el] (bl) at (-1,  -1) {l};
-    % \node[el] (br) at (1,   -1) {m};
-    % \node[v]  (b)  at (0,   -2) {d};
-
-    % \node[el] (v2) at (-1,-1) {b};
-    % \node[el] (v3) at (-1,-1) {b};
-    % \node[el] (v4) at (-1,-1) {b};
-    % \node[el] (v5) at (-1,-1) {b};
-
-    \tedge{m}{l}{solid}{}{};
-    \tedge{m}{r}{solid}{}{};
-
-    \tedge{l}{tl}{solid}{}{};
-    \tedge{tl}{t}{solid}{}{};
-    \tedge{r}{tr}{solid}{}{};
-    \tedge{tr}{t}{solid}{}{};
-
-    % \tedge{l}{bl}{solid}{}{};
-    % \tedge{bl}{b}{solid}{}{};
-    % \tedge{r}{br}{solid}{}{};
-    % \tedge{br}{b}{solid}{}{};
-\end{tikzpicture}
-\end{center}
-
-\caption{Each node contains a subgraph of the entire, well-constrained graph $G$.
-We take $Idc(G,a\cup i\cup b)$, $Idc(G,a\cup j\cup c)$, and $Idc(G,b\cup k\cup c)$ to be well-constrained.
-If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c$ must all be trivial subgraphs.
-% Each node contains a subgraph of the entire, well-constrained graph.
-We take $a\cup i\cup b$, $a\cup j\cup c$, $b\cup k\cup c$, $b\cup l\cup d$, and $c\cup m\cup d$ to be well-constrained.
-If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c,d$ must all be trivial subgraphs.
-% THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
-}
-\label{trivial-intersection}
-\end{figure}
-
-\begin{theorem}\label{uc-t1}
-% In any solution to $OptimalDRP(G)$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is under-constrained, then all well-constrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.
-
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial (case (B)), then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
-\end{theorem}
-
-\begin{proof}
-From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be under-constrained.
-If it is under-constrained, it must be part of some larger well-constrained subgraph $D$.
-If $D$ is not $G$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
- % For $C$ to be well-constrained... WHAT????
-
-If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
-
-This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid.
-Such an example is shown in Figure \ref{trivial-intersection}.
-\end{proof}
-
-% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$.
-All of these subgraphs form a complex structure that is well-constrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).
-
-\begin{theorem}\label{wc-t1}
-% In $OptimalDRP(G)$ of a well-constrained $G$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is well-constrained, then said node can have only $2$ children.
-
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained (case (A)), then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
-% Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
-\end{theorem}
-
-\begin{proof}
-For a valid DR-plan, all children must be proper subgraphs.
-Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least 2 children.
-It follows directly from Lemma \ref{wc-l2} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then you found a tree with only 2 children.
-
-Clearly, this is the optimal choice at this level. However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in later on.
-
-Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. Take $R_k=C\setminus C_k$. This means that $D$ is also $\bigcup_{k=1}^N{R_k}$.
-
-After we go down $N$ levels we get to a point where every node is some $Idc(C,R_k\cup D)$. Until this point, the fan-in of every node is exactly 2.
-
-This is demonstrated with $N=3$ in Figure \ref{3-plans}.
-It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution.
-This follows from the discussion above.
-Now, at this level, the fan-in must be equal in all scenarios.
-In fact, at node $k$ it will be 1 (because of $D$ which is well-constrained) plus whatever fan-in $R_k$ incurs.
-\end{proof}
-
-With these two proofs, the original Theorem \ref{t1} has been proven.
-With this we have formulated the notion of a canonical optimal DR-plan.
-Any optimal DR-plan of a well-constrained graph must follow this pattern.
-
-% (2) If the union results in a well-constrained subgraph, it must be the entire graph. If this were not the case, .
-
-% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$.
-It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Further Results}
-
-\begin{corollary}
-Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is well-constrained, then $D$ will be a descendant of every $C_k$.
-\end{corollary}
-
-\begin{proof}
-Explained in Theorem \ref{wc-t1}.
-\end{proof}
-
-\begin{corollary}
-$\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
-\end{corollary}
-
-\begin{proof}
-Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
-\end{proof}
-
-% \begin{proof}
-% It follows from Theorem \ref{uc-t1} that if the intersection of any two well-constrained vertex-maximal proper subgraphs of $C$ are well-constrained then all of them are well-constrained or $\emptyset$.
-% \end{proof}
-
-% Figure idea: Large node in the middle with weight $k$.
-Several other graphs with weight $a, b, c, d\ldots$ attached radially by single edges with weights $a, b, c, d\ldots$ respectively.
-Circle each vertex-maximal subset.
-
-% $\forall k\in [1,\ldots,N]$ the subgraphs $C\setminus C_k$ will be edge disjoint. Therefore the solving of one is exclusive of the solving of the other.
-
-% Furthermore, the union of all of these subgraphs will be $C\setminus D$.
-So each subgraph will need to be ``tacked'' onto $D$ for recombination and since they are edge-disjoint the order is unimportant.
-
-% \pnplaceholder
-% \begin{lemma}
-% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
-% \end{lemma}
-
-% \begin{proof}
-% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-% \pnplaceholder
-
-% \begin{corollary}
-% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-% \end{corollary}
-
-% \begin{proof}
-% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
-% \end{proof}
-
-\begin{corollary}
-There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-\end{corollary}
-
-\begin{proof}
-Since $C_1\cup C_2$ is the entire graph
-\todo{Prove this!!!!}
-\end{proof}
-
-% \begin{corollary}
-% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
-% \end{corollary}
-
-% \begin{proof}
-% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Future Work}
-While this paper fleshes out the case of a well-constrained input graph, little work has been done on the over-constrained graph.
-This has to do with the fact that finding a DR-Plan is NP-complete.
-But is there some algorithm that will find an approximation within a reasonable factor?
-
-Under-constrained is not worth studying in itself, because it will either be that all subgraphs are under-constrained or trivial or that some subgraphs are well-constrained or over-constrained.
-In this way, it is easily reduced to the other problems.
-\\
-
-The current method for finding $DRP(G)$ is the Frontier algorithm.
-Is there a way to adapt the Pebble Game algorithm?
-It is used to detect rigidity, but there is no obvious way to use it to find wcvmps (other than cutting out one node at a time).
-
-Or maybe, instead of these top-down algorithms, there is some bottom-up algorithm?
-
-Useful for glass networks.
-
-\begin{comment}
-\section{Formal Conjectures}
-
-\subsection{Conjecture 1}
-
-\begin{conjecture}
-% Take $G$ to be well-constrained. Then, for all nodes $C$ in $DRP(G)$, either $\forall i,j (C_i \cap C_j \leq 1)$ or $\forall i,j (C_i \cap C_j \geq 2) \wedge (C_i \cup C_j = C)$.
-
-% or...
-
-Take $G$ to be well-constrained. The operator $\cap_V$ returns the count of the vertices in the intersection of two graphs. Then, for all nodes $C$ in $DRP(G)$, exactly one of the following is true:
-\begin{align*}
-    \forall i,j & (C_i \cap_V C_j = 1) \wedge \p{\bigcup_{k\in\set{1,\ldots,N}\setminus\set{i}}{C_k} \neq C} \text{ or} \\
-    \forall i,j & (C_i \cap_V C_j \geq 2) \wedge (C_i \cup C_j = C)
-\end{align*}
-
-That is, either any two children share a single vertex and all of them are needed to form the parent node, or there is some significant overlap of 2 or more vertices and because of that only 2 children are needed at node $C$ to satisfy the definition of the DR-plan.
-
-This implies that in the first scenario all proper vertex-maximal rigid subgraphs of $C$ must be in tree $DRP(C)$, and that in the second scenario only 2 such subgraphs need to be in the tree.
-
-$C_i \cap_V C_j$ cannot be $<1$ because 0 is disjoint but $G$ is taken to be well-constrained (negative and non-integral values are meaningless).
-\end{conjecture}
-
-\begin{conjecture}
-$DRP(G)$ is ``unique'' modulo choice of children. Which means that the choice of children cannot impact the solution of the framework.
-\end{conjecture}
-
-\end{comment}
-
-\begin{comment}
-\pnplaceholder
-Notes:
-Edges do not appear as nodes of the DR plan, although vertices do (as leaves).
-In general, the children of a node are well-constrained proper maximal subgraphs of the parent, whose vertex set union gives the parent's vertex set.
-
-We are trying to formalize here the concept of a canonical DR-plan of a well-constained graph; i.e, whether the non-uniqueness of the decomposition - in the case that some pair of proper vertex maximal rigid subgraphs have a nontrivial intersection - makes any real difference.
-
-Conjecture 1 (should be easy to prove):
-Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_2| \ge 2$ (called a nontrivial intersection).
-Then the subgraph $S_{1,2}$ induced by $V_1\cap V_2$ must be well-constrained.
-
-Here are corollary facts that should be easy to prove:
-(1) There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-(2) There is no rigid subgraph induced by a proper, nonempty subset of vertices of   $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-
-Corollary of Conjecture 1 and resulting facts:
-$C_1\cap C_2$ is a proper vertex maximal rigid subgraph of $C_i$.
-Either there is a unique second proper vertex maximal rigid subgraph of $C_i$ that intersects  $C_1\cap C_2$  nontrivially (on an edge), or every pair of proper vertex maximal rigid subgraphs of $C_i$ intersect trivially.
-
-This would imply the DR-plans of $C_1$ and $C_2$ will be unique and have a common child in the DR-plan of $C_1\cap C_2$.
-
-Conjecture 2: Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_1| \ge 2$.
-$C_3 =(V_3,E_3),C_4 = (V_4,E_4)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_3\cup V_4$ and $|V_3\cap V_4| \ge 2$.
-Then both DR-plans where $C_1$ and $C_2$ are taken as children or where $C_4$ and $C_3$ are taken as children will have the same max fan-in.
-In particular, either DR-plan will contain $\bigcap\limits_i C_i$ as a node - at the 2nd level below $C$.
-\end{comment}
-
-% \pnplaceholder
-
-% Lemma 1: if the graph is well-constrained and any two proper vertex maximal subgraphs have an overlap of only one vertex, then all of the subgraphs will have 1 or 0.
-
-% Overlap of one vertex between any $C_i$ and $C_j$ means that $C\neq C_i\cup C_j$ because we know $C$ is well-constrained and the graph $C_i\cup C_j$ would be able to pivot around the single vertex.
-
-% \pnplaceholder
-
-% Need to define trivial as a graph that, were it only accounting for vertices, would be over-constrained. I.e.\ a single vertex in 2D. Or a vertex and an edge in 3D.
-
-% \pnplaceholder
-
-% \begin{theorem}
-% phrasing...
-% \end{theorem}
-
-% \begin{proof}
-% By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained itself, or (3) an over-constrained graph, referred to as trivial.
-We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.
-
-% (1) If the union results in an under-constrained subgraph, then the intersection of the two original subgraphs must be an over-constrained trivial subgraph (can be shown with simple arithmetic).
-As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are over-constrained trivial subgraphs (including $\emptyset$).
-All of these edge-disjoint subgraphs form a complex structure that is well-constrained and must be the graph itself (otherwise we did not have vertex-maximal subgraphs to begin with).
-
-% (2) If the union results in a well-constrained subgraph, it must be the entire graph.
-If this were not the case, the two original subgraphs could not have been vertex-maximal proper because we would have found a larger well-constrained subgraph containing them; namely, their union or something containing it would have been larger.
-
-% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$.
-It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case).
-The intersection also cannot be under-constrained.
-If $C_1 \cap C_2$ were under-constrained, then for $C_1$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
-The same is true for $C_2$.
-If both of these were satisfied, then the union of $C_1$ and $C_2$ would result in an over-constrained nontrivial graph.
-This is a contradiction, because we have shown the union to be the graph itself, which was defined to be well-constrained.
-Therefore, the intersection must be well-constrained.
-
-% (3) It is impossible for the union of 2 nontrivial subgraphs to be trivial.
-% \end{proof}
-
-% \begin{corollary}
-% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-% \end{corollary}
-
-% \begin{proof}
-% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
-% \end{proof}
-
-% \begin{corollary}
-% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-% \end{corollary}
-
-% \begin{proof}
-% Since $C_1\cup C_2$ is the entire graph,
-
-% well... no. if $V_i$ contains $V_1 \setminus (V_1\cap V_2)$ or $V_2 \setminus (V_1\cap V_2)$, then you will induce a well-constrained graph.
-% \end{proof}
-
-% \begin{corollary}
-% $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every $C_i$.
-% \end{corollary}
-
-% \begin{proof}
-% First, it must vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-
-% \pnplaceholder
-
-% \pnplaceholder
-
-% need to show that that intersection is going to end up being a vertex-maximal subgraph of both of the original subgraphs. As a result, in the second case it doesn't matter which 2 you took.
-
-% --------------------------------------------------------------
-%                        End of Body
-% --------------------------------------------------------------
-
-% \end{multicols}
-\end{document}
-
-% --------------------------------------------------------------
-%                       End of Document
-% --------------------------------------------------------------

### 2014-12-03 doc/GMP2015/s03_drp.tex
Changes to paper.

--- 2014-12-02 doc/General_Work/DRP_2d.tex
@@
-% --------------------------------------------------------------
-%                       Initialize
-% --------------------------------------------------------------
+\section{Decomposition-Recombination Plan}
 
-\documentclass[11pt]{article}
-
-% \usepackage{multicol}
-\usepackage{ulem}
-\usepackage{comment}
-\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}
-
-\usepackage{personalNotes}
-% contains special macros for math environment:
-%     unit, p, set, abs, floor, ceil, vec, norm, inner, kernel
-% contains useful all-purpose macros:
-%     n, lessSeparation, append
-
-\usepackage{tikz}
-\usetikzlibrary{shapes,snakes}
-\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
-
-% --------------------------------------------------------------
-%                    Beginning of Document
-% --------------------------------------------------------------
-
-\begin{document}
-% \begin{multicols}{2}
-
-% --------------------------------------------------------------
-%                           Title
-% --------------------------------------------------------------
-
-\pntitle {Decomposition Recombination Plans for 2D Well-Constrained Graphs}
-\pnauthor{Troy Baker}
-\pndate  {May 13, 2014}
-
-\pnmaketitle
-\pnmakefooter
-
-% --------------------------------------------------------------
-%                           Body
-% --------------------------------------------------------------
-
-\section{Introduction}
-\todo{
-Other sections...
-\begin{itemize}
-    \item Intro and motivation (previous work, where this work leads, examples)
-    \item Preliminaries
-    \item DR-Plan (Canonical) + algo
-    \item Recombination + algo (refer to other paper and make 2D)
-    \item Solution space navigation
-    \item Conclusion
-\end{itemize}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Preliminaries}
-
-\subsection{Definitions}
-% ISSUE: My definitions don't prevent non-positive densities... This means things aren't guaranteed to be connected...
-
-\begin{definition}[Graph]
-Graph $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$.
-Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
-\end{definition}
-
-\begin{definition}[Density]
-The density of a graph is defined as:
-\[d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}\]
-% \[d(G) = \max\set{0,\sum_{v\in V}{w_V(v)} - \sum_{e\in E}{w_E(e)}}\]
-\end{definition}
-
-\begin{definition}[Independent or Sparse Graph]
-\todo{Needs definition.}
-\end{definition}
-
-\begin{definition}[Over-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is over-constrained if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$.
-(where S is not trivial...
-This seems weird because you could have an over-constrained under-constrained graph...)
-\end{definition}
-
-\begin{definition}[Well-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
-\end{definition}
-
-% \begin{definition}[Well-constrained Graph]
-% Graph $G=(V,E,w_V,w_E)$ is well-constrained if, given constant $k$ and a set of trivial graphs $T$, the graph satisfies (1) $d(G)=k$ and (2) all subgraphs are either under-constrained, well-constrained, or trivial (i.e.\ are isomorphic to one of the graphs in $T$).
-% \end{definition}
-
-\begin{definition}[Rigid Graph]
-Graph $G=(V,E,w_V,w_E)$ is rigid if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-\end{definition}
-
-\begin{definition}[Under-constrained Graph]
-Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, $G$ is not rigid.
-\end{definition}
-
-% \begin{definition}[Under-constrained Graph]
-% Graph $G=(V,E,w_V,w_E)$ is under-constrained if, given constant $k$, there exists some subgraph $G'\subseteq G$ such that $d(G') < k$.
-% \end{definition}
-
-\begin{definition}[Trivial Graph]
-A trivial graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
-
-\todo{Maybe leave this out until it's needed later?}
-In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
-Trivial graphs for 2D would be a single vertex.
-Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between).
-Etc.
-These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
-\end{definition}
-
-\begin{definition}[Decomposition Recombination (DR-) Plan]
-The DR-plan of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
-\begin{enumerate}
-    \item The root of the tree `contains' $G$.
-    \item For all nodes, $C$, that `contain' nontrivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
-    \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
-    \item A node with a trivial subgraph is a leaf.
-\end{enumerate}
-% It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
-
-An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.
-
-Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
-\end{definition}
-
-\begin{definition}[Complete DR-Plan]
-The complete DR-plan of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \textit{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
-By this, rule 3 is implicit for a complete DR-plan.
-\end{definition}
-
-\begin{definition}[Optimal DR-Plan]
-The optimal DR-plan of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Notation}
-$G$ is taken to be a graph with values $(V,E,w_V,w_E)$.
-To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
-It is not useful to solve for two bodies simultaneously if there are no constraints between them.
-Furthermore, to limit the scope of the paper, $G$ will always be taken to be well-constrained.
-% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.
-
-$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.
-
-$C$ is an arbitrary graph in a node in $CompleteDRP(G)$.
-$C_i$ is the $i^{\text{th}}$ child of $C$.
-By definition of a DR-plan, it is implied that $C_i$ is a well-constrained vertex-maximal proper subgraph.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{The Canonical Optimal DR-Plan}
-\begin{figure}
-\begin{center}
+\begin{figure*}\centering
 \begin{tikzpicture}
@@
 % \end{tikzpicture}
-\end{center}
 
@@
 \label{3-plans}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
-\begin{center}
+\begin{figure}\centering
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\end{center}
 
@@
 % \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Future Work}
-While this paper fleshes out the case of a well-constrained input graph, little work has been done on the over-constrained graph.
-This has to do with the fact that finding a DR-Plan is NP-complete.
-But is there some algorithm that will find an approximation within a reasonable factor?
-
-Under-constrained is not worth studying in itself, because it will either be that all subgraphs are under-constrained or trivial or that some subgraphs are well-constrained or over-constrained.
-In this way, it is easily reduced to the other problems.
-\\
-
-The current method for finding $DRP(G)$ is the Frontier algorithm.
-Is there a way to adapt the Pebble Game algorithm?
-It is used to detect rigidity, but there is no obvious way to use it to find wcvmps (other than cutting out one node at a time).
-
-Or maybe, instead of these top-down algorithms, there is some bottom-up algorithm?
-
-Useful for glass networks.
-
-\begin{comment}
-\section{Formal Conjectures}
-
-\subsection{Conjecture 1}
-
-\begin{conjecture}
-% Take $G$ to be well-constrained. Then, for all nodes $C$ in $DRP(G)$, either $\forall i,j (C_i \cap C_j \leq 1)$ or $\forall i,j (C_i \cap C_j \geq 2) \wedge (C_i \cup C_j = C)$.
-
-% or...
-
-Take $G$ to be well-constrained. The operator $\cap_V$ returns the count of the vertices in the intersection of two graphs. Then, for all nodes $C$ in $DRP(G)$, exactly one of the following is true:
-\begin{align*}
-    \forall i,j & (C_i \cap_V C_j = 1) \wedge \p{\bigcup_{k\in\set{1,\ldots,N}\setminus\set{i}}{C_k} \neq C} \text{ or} \\
-    \forall i,j & (C_i \cap_V C_j \geq 2) \wedge (C_i \cup C_j = C)
-\end{align*}
-
-That is, either any two children share a single vertex and all of them are needed to form the parent node, or there is some significant overlap of 2 or more vertices and because of that only 2 children are needed at node $C$ to satisfy the definition of the DR-plan.
-
-This implies that in the first scenario all proper vertex-maximal rigid subgraphs of $C$ must be in tree $DRP(C)$, and that in the second scenario only 2 such subgraphs need to be in the tree.
-
-$C_i \cap_V C_j$ cannot be $<1$ because 0 is disjoint but $G$ is taken to be well-constrained (negative and non-integral values are meaningless).
-\end{conjecture}
-
-\begin{conjecture}
-$DRP(G)$ is ``unique'' modulo choice of children. Which means that the choice of children cannot impact the solution of the framework.
-\end{conjecture}
-
-\end{comment}
-
-\begin{comment}
-\pnplaceholder
-Notes:
-Edges do not appear as nodes of the DR plan, although vertices do (as leaves).
-In general, the children of a node are well-constrained proper maximal subgraphs of the parent, whose vertex set union gives the parent's vertex set.
-
-We are trying to formalize here the concept of a canonical DR-plan of a well-constained graph; i.e, whether the non-uniqueness of the decomposition - in the case that some pair of proper vertex maximal rigid subgraphs have a nontrivial intersection - makes any real difference.
-
-Conjecture 1 (should be easy to prove):
-Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_2| \ge 2$ (called a nontrivial intersection).
-Then the subgraph $S_{1,2}$ induced by $V_1\cap V_2$ must be well-constrained.
-
-Here are corollary facts that should be easy to prove:
-(1) There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-(2) There is no rigid subgraph induced by a proper, nonempty subset of vertices of   $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-
-Corollary of Conjecture 1 and resulting facts:
-$C_1\cap C_2$ is a proper vertex maximal rigid subgraph of $C_i$.
-Either there is a unique second proper vertex maximal rigid subgraph of $C_i$ that intersects  $C_1\cap C_2$  nontrivially (on an edge), or every pair of proper vertex maximal rigid subgraphs of $C_i$ intersect trivially.
-
-This would imply the DR-plans of $C_1$ and $C_2$ will be unique and have a common child in the DR-plan of $C_1\cap C_2$.
-
-Conjecture 2: Let $C = (V,E)$ be well-constrained and $C_1 =(V_1,E_1),C_2 = (V_2,E_2)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_1\cup V_2$ and $|V_1\cap V_1| \ge 2$.
-$C_3 =(V_3,E_3),C_4 = (V_4,E_4)$ any two proper vertex-maximal well-constrained subgraphs with $V = V_3\cup V_4$ and $|V_3\cap V_4| \ge 2$.
-Then both DR-plans where $C_1$ and $C_2$ are taken as children or where $C_4$ and $C_3$ are taken as children will have the same max fan-in.
-In particular, either DR-plan will contain $\bigcap\limits_i C_i$ as a node - at the 2nd level below $C$.
-\end{comment}
-
-% \pnplaceholder
-
-% Lemma 1: if the graph is well-constrained and any two proper vertex maximal subgraphs have an overlap of only one vertex, then all of the subgraphs will have 1 or 0.
-
-% Overlap of one vertex between any $C_i$ and $C_j$ means that $C\neq C_i\cup C_j$ because we know $C$ is well-constrained and the graph $C_i\cup C_j$ would be able to pivot around the single vertex.
-
-% \pnplaceholder
-
-% Need to define trivial as a graph that, were it only accounting for vertices, would be over-constrained. I.e.\ a single vertex in 2D. Or a vertex and an edge in 3D.
-
-% \pnplaceholder
-
-% \begin{theorem}
-% phrasing...
-% \end{theorem}
-
-% \begin{proof}
-% By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained itself, or (3) an over-constrained graph, referred to as trivial.
-We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.
-
-% (1) If the union results in an under-constrained subgraph, then the intersection of the two original subgraphs must be an over-constrained trivial subgraph (can be shown with simple arithmetic).
-As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are over-constrained trivial subgraphs (including $\emptyset$).
-All of these edge-disjoint subgraphs form a complex structure that is well-constrained and must be the graph itself (otherwise we did not have vertex-maximal subgraphs to begin with).
-
-% (2) If the union results in a well-constrained subgraph, it must be the entire graph.
-If this were not the case, the two original subgraphs could not have been vertex-maximal proper because we would have found a larger well-constrained subgraph containing them; namely, their union or something containing it would have been larger.
-
-% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$.
-It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case).
-The intersection also cannot be under-constrained.
-If $C_1 \cap C_2$ were under-constrained, then for $C_1$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
-The same is true for $C_2$.
-If both of these were satisfied, then the union of $C_1$ and $C_2$ would result in an over-constrained nontrivial graph.
-This is a contradiction, because we have shown the union to be the graph itself, which was defined to be well-constrained.
-Therefore, the intersection must be well-constrained.
-
-% (3) It is impossible for the union of 2 nontrivial subgraphs to be trivial.
-% \end{proof}
-
-% \begin{corollary}
-% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-% \end{corollary}
-
-% \begin{proof}
-% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
-% \end{proof}
-
-% \begin{corollary}
-% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-% \end{corollary}
-
-% \begin{proof}
-% Since $C_1\cup C_2$ is the entire graph,
-
-% well... no. if $V_i$ contains $V_1 \setminus (V_1\cap V_2)$ or $V_2 \setminus (V_1\cap V_2)$, then you will induce a well-constrained graph.
-% \end{proof}
-
-% \begin{corollary}
-% $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every $C_i$.
-% \end{corollary}
-
-% \begin{proof}
-% First, it must vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-
-% \pnplaceholder
-
-% \pnplaceholder
-
-% need to show that that intersection is going to end up being a vertex-maximal subgraph of both of the original subgraphs. As a result, in the second case it doesn't matter which 2 you took.
-
-% --------------------------------------------------------------
-%                        End of Body
-% --------------------------------------------------------------
-
-% \end{multicols}
-\end{document}
-
-% --------------------------------------------------------------
-%                       End of Document
-% --------------------------------------------------------------

### 2014-12-03 doc/GMP2015/s04_recombination.tex
Changes to paper.

\section{Recombination}

Discussion about algorithm.

Drop an appropriate edge. Add an appropriate non-edge to get tree-decomposable graph. Change length of edge until dropped edge is correct length (as supplied as input).

### 2014-12-03 src/GUI/old_font_stuff.cpp
Several changes. Trying to improve pebble game to get clusters.

@@ -1,0 +1,184 @@
+#include "fontstash.h"
+#include <fstream>
+
+void do_font_stuff() {
+    char *ttf_file = "src/OpenSans-Regular.ttf";
+    // std::ifstream
+    std::ifstream f(ttf_file);
+    if (f.is_open()) {
+        cout << "Open file." << endl;
+    } else {
+        cout << "Did not open file." << endl;
+    }
+
+
+    // fp = fopen("DroidSerif-Regular.ttf", "rb");
+    // if (!fp) goto error_add_font;
+    // fseek(fp, 0, SEEK_END);
+    // datasize = (int)ftell(fp);
+    // fseek(fp, 0, SEEK_SET);
+    // data = (unsigned char*)malloc(datasize);
+    // if (data == NULL) goto error_add_font;
+    // fread(data, 1, datasize, fp);
+    // fclose(fp);
+
+
+     // create a font stash with a maximum texture size of 512 x 512
+    struct sth_stash* stash = sth_create(512, 512);
+     // load truetype font
+    int font_id = sth_add_font(stash, ttf_file);
+     // position of the text
+
+    // float x = 10, y = 10;
+    float x = 0, y = 0;
+
+
+    // draw text during your OpenGL render loop
+    sth_begin_draw(stash);
+    // position: (x, y); font size: 24
+    sth_draw_text(stash, font_id, 24.0f, x, y, "Hello world!", &x);
+    // now, the float x contains the x position of the next char
+    sth_end_draw(stash);
+}
+
+// void draw_font() {
+// }
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+// #define STB_TRUETYPE_IMPLEMENTATION
+// #include "../ext/stb/stb_truetype.h"
+
+
+
+
+
+// stbtt_bakedchar cdata[96]; // ASCII 32..126 is 95 glyphs
+// GLuint ftex;
+
+// void my_stbtt_initfont(void)
+// {
+//     char *ttf_file = "src/OpenSans-Regular.ttf";
+
+//     unsigned char ttf_buffer[1<<20];
+//     unsigned char temp_bitmap[512*512];
+
+//     fread(ttf_buffer, 1, 1<<20, fopen(ttf_file, "rb"));
+//     // cout << ttf_buffer << endl;
+
+//     stbtt_BakeFontBitmap(ttf_buffer, 0,
+//         32.0,
+//         temp_bitmap, 512, 512,
[…]

### 2014-12-03 src/Graph/Graph.cpp

@@ -30,7 +30,7 @@
 
 //   for (int y = (int)min_point[1]; y <= (int)max_point[1]; ++y) {
 //     for (int x = (int)min_point[0]; x <= (int)max_point[0]; ++x) {
-//       typename graph_traits<Graph>::vertex_iterator vi, vi_end;
+//       typename graph_traits<Graph>::vertex_Iteratorator vi, vi_end;
 //       // Find vertex at this position
 //       typename graph_traits<Graph>::vertices_size_type index = 0;
 //       for (tie(vi, vi_end) = vertices(g); vi != vi_end; ++vi, ++index) {
@@ -51,14 +51,14 @@
 //   std::ofstream out((name + ".dot").c_str());
 //   out << "graph " << name << " {" << std::endl;
 
-//   typename graph_traits<Graph>::vertex_iterator vi, vi_end;
+//   typename graph_traits<Graph>::vertex_Iteratorator vi, vi_end;
 //   for (tie(vi, vi_end) = vertices(g); vi != vi_end; ++vi) {
 //     out << "  n" << get(vertex_index, g, *vi) << "[ pos=\""
 //         << (int)position[*vi][0] + 25 << ", " << (int)position[*vi][1] + 25
 //         << "\" ];\n";
 //   }
 
-//   typename graph_traits<Graph>::edge_iterator ei, ei_end;
+//   typename graph_traits<Graph>::edge_Iteratorator ei, ei_end;
 //   for (tie(ei, ei_end) = edges(g); ei != ei_end; ++ei) {
 //     out << "  n" << get(vertex_index, g, source(*ei, g)) << " -- n"
 //         << get(vertex_index, g, target(*ei, g)) << ";\n";
@@ -71,14 +71,14 @@
 // test_circle_layout(Graph*, typename graph_traits<Graph>::vertices_size_type n)
 // {
 //   typedef typename graph_traits<Graph>::vertex_descriptor vertex;
-//   typedef typename graph_traits<Graph>::vertex_iterator vertex_iterator;
+//   typedef typename graph_traits<Graph>::vertex_Iteratorator vertex_Iteratorator;
 //   typedef typename graph_traits<Graph>::vertices_size_type vertices_size_type;
 //   typedef typename graph_traits<Graph>::edges_size_type edges_size_type;
 
 //   Graph g(n);
 
 //   // Initialize vertex indices
-//   vertex_iterator vi = vertices(g).first;
+//   vertex_Iteratorator vi = vertices(g).first;
[…]

### 2014-12-03 src/Graph/Graph.hpp

@@ -60,29 +60,35 @@ struct Edge_Properties {
 
 // undirected is simplified adjacency list graph
 // don't want adjacency matrix as these graphs have |E|=O(|V|)
-typedef boost::undirected_graph<Vertex_Properties, Edge_Properties> gtype;
-typedef boost::graph_traits<gtype>::vertex_descriptor Vertex;
-typedef boost::graph_traits<gtype>::edge_descriptor Edge;
+typedef boost::undirected_graph<Vertex_Properties, Edge_Properties> Graph_Type;
+typedef boost::graph_traits<Graph_Type>::vertex_descriptor Vertex_ID;
+typedef boost::graph_traits<Graph_Type>::edge_descriptor Edge_ID;
 
-typedef boost::graph_traits<gtype>::vertex_iterator Vertex_iter;
-typedef boost::graph_traits<gtype>::edge_iterator Edge_iter;
-
-
+typedef boost::graph_traits<Graph_Type>::vertex_iterator Vertex_Iterator;
+typedef boost::graph_traits<Graph_Type>::edge_iterator Edge_Iterator;
 
 
 
-class Graph : public gtype
+
+
+class Graph : public Graph_Type
 {
 public:
-    Vertex add_vertex();
-    Edge add_edge(Vertex v0, Vertex v1, double distance);
+    Graph() : Graph_Type() {}
+    // Graph(const Graph& g);
 
-    std::pair<Vertex_iter, Vertex_iter> vertices();
-    std::pair<Edge_iter, Edge_iter> edges();
+    Vertex_ID add_vertex(Vertex_Properties vp);
+    Vertex_ID add_vertex();
+
+    Edge_ID add_edge(Vertex_ID v0, Vertex_ID v1, Edge_Properties ep);
+    Edge_ID add_edge(Vertex_ID v0, Vertex_ID v1, double distance);
+
+    std::pair<Vertex_Iterator, Vertex_Iterator> vertices() const;
+    std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
     unsigned int num_vertices();
 
-    std::pair<Vertex, Vertex> verts_on_edge(Edge_iter e);
+    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_Iterator e);
 
     void set_layout();
 

### 2014-12-03 src/Graph/Pebbled_Graph.old.cpp

@@ -4,13 +4,6 @@
 
 
 
-
-
-
-
-
-
-
 Pebbled_Graph::Pebbled_Graph(Graph *g) {
     this->g = g;
 }
@@ -78,137 +71,14 @@ Clusters Pebbled_Graph::pebble_game_2D_exclude(Vertex_ID excluded_vert) {
         if (v1_cid != cid_end && v2_cid != cid_end && v1_cid == v2_cid)
             continue;
 
-        // gather l+1 pebbles around the edge
-        // bool found_l_plus_1 = false;
-        bool done_with_v1 = false, done_with_v2 = false;
-        while (true) {
-            unsigned int
-                pebs_on_v1 = num_pebbles_on_vert(v1),
-                pebs_on_v2 = num_pebbles_on_vert(v2);
-
-            if (pebs_on_v1 + pebs_on_v2 > this->l) {
-                // found_l_plus_1 = true;
-                break;
-            }
-
-            // initialize
-            std::pair<Vertex_Iterator, Vertex_Iterator> vs = g->vertices();
-            for (Vertex_Iterator it = vs.first; it != vs.second; it++)
-            {
-                seen[*it] = false;
-            }
-
-            if (!done_with_v1) {
-                if (!make_pebble_available(v1)) {
-                    done_with_v1 = true;
-                }
-            }
-            else if (!done_with_v2) {
-                if (!make_pebble_available(v2)) {
-                    done_with_v2 = true;
-                }
-            }
-            else {
-                // found_l stays false
-                break;
-            }
-        }
-
-        // 2 pebbles should be on each for (2,3)
-        // but just to be safe and have some generality...
-        unsigned int
-            pebs_on_v1 = num_pebbles_on_vert(v1),
-            pebs_on_v2 = num_pebbles_on_vert(v2);
-        if (pebs_on_v1 + pebs_on_v2 > this->l) {
-            if (pebs_on_v1 > 0)
-                place_available_pebble(v1, v2);
-            else
-                place_available_pebble(v2, v1);
-        }
-
-        // just for generality...
-        if (pebs_on_v1 + pebs_on_v2 - 1 > this->l)
-            continue;
-
-
-
         // if everything is fine, enlarge cover
         this->enlarge_cover(it);
 
     }
 
     return cs;
-
-
-
-
-
-    //     // initialize
[…]

### 2014-12-03 src/Graph/Pebbled_Graph.old.hpp

@@ -25,7 +25,6 @@ private:
     // pebbles per node = k = 2
     // pebbles for sparsity/tightness = l = 3
     // solving for (k, l)-sparse graphs
-    // ONLY GUARANTEED TO WORK ON (2,3)
     static const unsigned int k = 2;
     static const unsigned int l = 3;
 
@@ -62,9 +61,9 @@ private:
         vert_peb() {}
         vert_peb(Vertex_ID v, unsigned int p) {vertex = v; pebble = p;}
     };
-    // // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
-    // // used to reconstruct the path for rearranging.
-    // std::map<Vertex_ID, vert_peb> path;
+    // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
+    // used to reconstruct the path for rearranging.
+    std::map<Vertex_ID, vert_peb> path;
 
 
     // enum pebble {
@@ -91,14 +90,8 @@ private:
     // returns the index of a free pebble or k if there is none
     unsigned int identify_free_pebble(Vertex_ID v);
 
-    // returns the number of pebbles free on given vertex
-    unsigned int num_pebbles_on_vert(Vertex_ID v);
-
-    // gives the inverted directed pebble game graph
-    std::map<Vertex_ID, pebbles> get_reversed_pebbles();
-
     // Finds a free pebble if there is one and sets up the path to it.
-    bool find_pebble(Vertex_ID v, std::map<Vertex_ID, vert_peb> *path);
+    bool find_pebble(Vertex_ID v);
 
     // Frees a pebble at vertex_ID v, if there is one. Returns true if successful
     bool make_pebble_available(Vertex_ID v);

### 2014-12-03 src/outside_code/fontstash.cpp

@@ -1,0 +1,815 @@
+//
+// Copyright (c) 2011-2013 Andreas Krinke andreas.krinke@gmx.de
+// Copyright (c) 2009 Mikko Mononen memon@inside.org
+//
+// This software is provided 'as-is', without any express or implied
+// warranty.  In no event will the authors be held liable for any damages
+// arising from the use of this software.
+// Permission is granted to anyone to use this software for any purpose,
+// including commercial applications, and to alter it and redistribute it
+// freely, subject to the following restrictions:
+// 1. The origin of this software must not be misrepresented; you must not
+//    claim that you wrote the original software. If you use this software
+//    in a product, an acknowledgment in the product documentation would be
+//    appreciated but is not required.
+// 2. Altered source versions must be plainly marked as such, and must not be
+//    misrepresented as being the original software.
+// 3. This notice may not be removed or altered from any source distribution.
+//
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <math.h> /* @rlyeh: floorf() */
+
+#include <GL/glew.h>  /* @rlyeh: before including GL. doesnt hurt and makes life better */
+
+#ifdef __MACOSX__
+#include <OpenGL/gl.h>
+#else
+#include <GL/gl.h>
+#endif
+
+/* @rlyeh: removed STB_TRUETYPE_IMPLENTATION. We link it externally */
+#include "stb_truetype.h"
+
+#include "fontstash.h"
+
+#define HASH_LUT_SIZE 256
+#define MAX_ROWS 128
+#define VERT_COUNT (6*128)
+#define VERT_STRIDE (sizeof(float)*4)
+
+#define TTFONT_FILE 1
+#define TTFONT_MEM  2
+#define BMFONT      3
+
+static int idx = 1;
+
+static unsigned int hashint(unsigned int a)
+{
+	a += ~(a<<15);
+	a ^=  (a>>10);
[…]

### 2014-12-03 src/outside_code/main.cpp

@@ -115,7 +115,7 @@ void do_font_stuff() {
 Graph set_up_graph() {
     Graph g;
 
-    vector<Vertex> v;
+    vector<Vertex_ID> v;
     for (int i = 0; i < 10; i++) {
         v.push_back(g.add_vertex());
     }
@@ -157,7 +157,7 @@ Graph set_up_graph() {
 vector<gl_obj::pos_vec> verts_from_graph(Graph &g) {
     vector<gl_obj::pos_vec> v;
 
-    for (std::pair<Vertex_iter, Vertex_iter> verts = g.vertices();
+    for (std::pair<Vertex_Iterator, Vertex_Iterator> verts = g.vertices();
         verts.first != verts.second;
         verts.first++)
     {
@@ -171,11 +171,11 @@ vector<gl_obj::pos_vec> verts_from_graph(Graph &g) {
 vector<gl_obj::pos_vec> edges_from_graph(Graph &g) {
     vector<gl_obj::pos_vec> v;
 
-    for (std::pair<Edge_iter, Edge_iter> edges = g.edges();
+    for (std::pair<Edge_Iterator, Edge_Iterator> edges = g.edges();
         edges.first != edges.second;
         edges.first++)
     {
-        std::pair<Vertex, Vertex> uv = g.verts_on_edge(edges.first);
+        std::pair<Vertex_ID, Vertex_ID> uv = g.verts_on_edge(edges.first);
         v.push_back(gl_obj::pos_vec(g[uv.first].x, g[uv.first].y));
         v.push_back(gl_obj::pos_vec(g[uv.second].x, g[uv.second].y));
     }
@@ -287,6 +287,9 @@ int main(int argc, char **argv) {
     Graph g;
     g.read_from_file(dot_file.c_str());
 
+    Graph copy(g);
+    g.write_to_file("out.dot");
+
     g.set_layout();
 
 
@@ -307,7 +310,7 @@ int main(int argc, char **argv) {
 
 
     Pebbled_Graph pg(&g);
-    cout << "Pebbles remaining: " << pg.DRP_2D() << endl;
+    cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
 
 
 

### 2014-12-03 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 \section{Introduction}
-Computational design and analysis of constrained geometric configuration spaces of 2-dimensional, multilayered, aperiodic or self-similar materials and biological structures.
+Constrained geometric configurations in layered materials and biological structures.
 
@@
 
-Many natural and engineered materials are nearly rigid, 2D multilayers, possibly consisting of inter-constraining monolayers each of which is topologically planar (genus 0).
-They are often aperiodic, recursively self-similar or both \cite{Intro1}, enabling formation of diverse, larger structures with multiple functions.
+Many natural and engineered materials are essentially two dimensional multilayers 
+that not planar (genus 0). i.e., they consist of interconstraining planar monolayers. In addition,
+they are often  
+aperiodic and recursively self-similar \cite{Intro1}, enabling minimization of mass, 
+optimal distribution of stresses, and formation  
+of diverse, larger structures with multiple functions.
 
@@
 
-    \item interlinked monolayers of cellulose microfibrils e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
+    \item interlinked or interwoven cellulose or collagen microfibrils e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
 
@@
 
-    \item (bi)layers and other structures that behave like assemblies of 2D particles under non-overlap constraints, i.e, jammed disks on the plane \cite{JammedDisk1}
+    \item (bi)layers and other structures that behave like assemblies of 2D particles under non-overlap constraints, i.e, jammed 
+    disks on the plane \cite{JammedDisk1}
 
-    \item More recent, engineered examples include disordered graphene layers \cite{Graphene1} \cite{Graphene2}  even reinforced by  microfibrils, and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet bringing other structures \cite{Microfibrils2} within reach.
+    \item More recent, engineered examples include disordered graphene layers \cite{Graphene1} \cite{Graphene2}  even reinforced 
+    by  microfibrils, and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet bringing other structures 
+    \cite{Microfibrils2} within reach.
 \end{itemize}
 
-Understanding the structural and mechanical properties of these materials and structures at multiple scales, including rigidity, flexibility and motion is best achieved in the tradition of computer-aided mechanical design:  by modeling the structure as a configuration of appropriate geometric primitives inter-constrained by metric or algebraic constraints; and designing the constraint-system to achieve the desired properties and their robustness/sensitivity to constraint-system variations.
+Understanding the structural and mechanical properties of these materials and structures at multiple scales, including rigidity, 
+flexibility and motion is best achieved in the tradition of computer-aided mechanical design:  by modeling the structure as a 
+configuration of appropriate geometric primitives inter-constrained by metric or algebraic constraints; and designing the 
+constraint-system to achieve the desired properties and their robustness/sensitivity to constraint-system variations.
 
@@
 
-The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp.
-2DGMCS) belong to a combinatorial class for which algorithms with formally proven efficiency guarantees exist for  decomposing into maximal generically well-constrained components \todo{[Citation Needed]}.
-However, these systems do not belong in the largest combinatorial class of 2D systems, whose configuration spaces have formally proven  characterizations and algorithmic guarantees (e.g.
-for algebraic complexity, path and volume measures, and  topology).
+The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp. 2DGMCS) 
+belong to a combinatorial class for which algorithms with formally proven efficiency guarantees exist for  
+decomposing into maximal generically well-constrained components \todo{[Citation Needed]}. However, these systems do not 
+belong in the largest combinatorial class of 2D systems, whose configuration spaces have formally proven  characterizations and 
+algorithmic guarantees (e.g. for algebraic complexity, path and volume measures, and  topology).
 
-Specifically, these results \todo{[Citation Needed]} rely on the fact that rigid systems in this  class, in contrast to GMCS, have a straightforward and immediate {\sl optimal} recursive decomposition, (so-called {\sl optimal} DR-plan) and  recombination or realization (with a formal disambiguation by minimal orientation specification).
+Specifically, these results \todo{[Citation Needed]} rely on the fact that rigid systems in this  class, in contrast to GMCS, 
+have a straightforward and immediate {\sl optimal} recursive decomposition, (so-called {\sl optimal} DR-plan) and  recombination 
+or realization (with a formal disambiguation by minimal orientation specification).
 
-It is known \todo{[Citation Needed]} that already the optimal DR-planning problem for general 2DGMCS is NP-hard, intuitively because of a combinatorial explosion of polynomially-sized DR-plans.
-In this paper, we  navigate this barrier by defining a so-called {\it canonical} DR-plan and showing that all canonical DR-plans for non-overconstrained 2DGMCS are optimal.
-We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS.
-The notion of a canonical DR-plan illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan.
-We then show how to judiciously combine existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration.
+It is known \todo{[Citation Needed]} that already the optimal DR-planning problem for general 2DGMCS is NP-hard, intuitively 
+because of a combinatorial explosion of polynomially-sized DR-plans.  In this paper, we  navigate this barrier by defining a 
+so-called {\it canonical} DR-plan and showing that all canonical DR-plans for non-overconstrained 2DGMCS are optimal. 
+We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan 
+illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of 
+overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine 
+existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, 
+to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration. 
 Software implementation can be found at: \todo{Provide Link}.
@@
 
-Computer aided mechanical design represents a mature application of geometric constraint systems,  leading to and borrowing from deep and fundamental problems and techniques in distance geometry, convex analysis, combinatorics, algebra, topology and complexity theory.
+Computer aided mechanical design represents a mature application of geometric constraint systems,  leading to and borrowing 
+from deep and fundamental problems and techniques in distance geometry, convex analysis, combinatorics, algebra, topology 
+and complexity theory.
 
-Broadly, these problems can be classified into three types that draw upon each other, and possess mathematical, algorithmic, and software implementation aspects: (i) combinatorial (generic) minimal rigidity (or isostaticity or well-constrained-ness) characterization, (ii) (optimal) recursive decomposition (DR plans), recombination and realization, (iii) configuration space characterization (topology, complexity, path and volume measures).
+Broadly, these problems can be classified into three types that draw upon each other, and possess mathematical, algorithmic, 
+and software implementation aspects: (i) combinatorial (generic) minimal rigidity (or isostaticity or well-constrained-ness) 
+characterization, (ii) (optimal) recursive decomposition (DR plans), recombination and realization, (iii) configuration space 
+characterization (topology, complexity, path and volume measures).
 
-In general, proofs of combinatorial (generic) minimal rigidity characterizations for geometric constraint systems rely on underlying matroids \todo{[Citation Needed]} In these cases, more-or-less greedy algorithms such as simplified network-flow, or so-called pebble games can be used.
+In general, proofs of combinatorial (generic) minimal rigidity characterizations for geometric constraint systems rely on 
+underlying matroids \todo{[Citation Needed]} In these cases, more-or-less greedy algorithms such as simplified network-flow, 
+or so-called pebble games can be used.
 
-Nothing nontrivial known until now for optimal DR plans, although efficient algorithms are known for DR-plans satisfying various other formal properties.
+Nothing nontrivial known until now for optimal DR plans, although efficient algorithms are known for DR-plans satisfying 
+various other formal properties.
 
-Recombination and realization navigation of realizations (disambiguation with orientation, faster with overconstraints, mention also totally different distance geometry method - low rank PSD matrix completions SDP/SVD)
+Recombination and realization navigation of realizations (disambiguation with orientation, faster with overconstraints, mention 
+also totally different distance geometry method - low rank PSD matrix completions SDP/SVD)
 
-Gelfand/McPherson, Walker's conjecture solution, Convex Cayley, EASAL and Caymos for low Cayley complexity and bijective representation using global rigidity.
+Gelfand/McPherson, Walker's conjecture solution, Convex Cayley, EASAL and Caymos for low Cayley complexity and bijective 
+representation using global rigidity.

### 2014-12-04 src/main.cpp
Makefile works for both Ubuntu and Mac OS now

@@ -23,7 +23,7 @@ using namespace std;
 
 
 
-#include "GUI/font.hpp"
+#include "GUI/Font.hpp"
 
 
 

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 \section{Introduction}
-Constrained geometric configurations in layered materials and biological structures.
+Modeling, analysis and design of self-similar, aperiodic layered materials as geometric constraint systems
 
-=======================
-
-Many natural and engineered materials are essentially two dimensional multilayers 
-that not planar (genus 0). i.e., they consist of interconstraining planar monolayers. In addition,
-they are often  
-aperiodic and recursively self-similar \cite{Intro1}, enabling minimization of mass, 
-optimal distribution of stresses, and formation  
-of diverse, larger structures with multiple functions.
+Many natural and engineered materials are recursively self-similar \cite{Intro1}, 
+spanning multiple scales. Within any one scale, the materials are quasiuniform, but aperiodic, 
+consisting of a few different motifs repeating in random combinations.
+In addition, many materials are stacked two dimensional (2D) layers, where each layer is independently self-similar and aperiodic.
+Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
+Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. This is 
+consistent with a self-assembled structure that minimizes mass, optimally distributes external stresses and itself 
+participates in the assembly of diverse, 
+larger structures with multiple functions, all within the same layer.
 
@@
 \begin{itemize}
-    \item (cross-sections) of ``necklaces'' of microtubules \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
 
-    \item interlinked or interwoven cellulose or collagen microfibrils e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
+    \item (cross-section) of microtubules \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
 
-    \item Inorganic examples include silica and glass \cite{SilicaGlass1} \cite{SilicaGlass2}
+    \item (cross-section) of hierarchical bone structure \cite{XX}
 
-    \item (bi)layers and other structures that behave like assemblies of 2D particles under non-overlap constraints, i.e, jammed 
+    \item interlinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
+
+    \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced 
+    by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other structures 
+    \cite{Microfibrils2} within reach.
+
+    \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 
+    2D particles under non-overlap constraints, e.g, jammed 
     disks on the plane \cite{JammedDisk1}
 
-    \item More recent, engineered examples include disordered graphene layers \cite{Graphene1} \cite{Graphene2}  even reinforced 
-    by  microfibrils, and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet bringing other structures 
-    \cite{Microfibrils2} within reach.
 \end{itemize}
 
-Understanding the structural and mechanical properties of these materials and structures at multiple scales, including rigidity, 
-flexibility and motion is best achieved in the tradition of computer-aided mechanical design:  by modeling the structure as a 
-configuration of appropriate geometric primitives inter-constrained by metric or algebraic constraints; and designing the 
-constraint-system to achieve the desired properties and their robustness/sensitivity to constraint-system variations.
+It is natural to model such a self-similar, aperiodic material using appropriate geometric primitives under metric or algebraic constraints; design the 
+constraint-system to achieve the desired properties; and study robustness/sensitivity to constraint variations.
+
+Two key observations serve as the motivations for this paper:
+As constraint systems, the above 2D {\em sesamacs} (Self-similar, aperiodic material constraint systems) involve multiple types of 
+well-studied geometric primitives and constraints: they can be modeled as so-called
+{\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3).  
+All of these systems have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}) 
+for locating maximal, generically rigid components \cite{XX}.
+
+Recursive decompositions of {\em slamacs} into rigid subsystems is For , is important to understand or design the structural 
+and mechanical properties - such as rigidity, flexes and distribution of external stresses within the layer - at all scales.
 
@@
 
-The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp. 2DGMCS) 
-belong to a combinatorial class for which algorithms with formally proven efficiency guarantees exist for  
-decomposing into maximal generically well-constrained components \todo{[Citation Needed]}. However, these systems do not 
+The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp. 2D-GEMS) 
+belong to a
+
+However, these systems do not 
 belong in the largest combinatorial class of 2D systems, whose configuration spaces have formally proven  characterizations and 

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 \section{Introduction}
-Modeling, analysis and design of self-similar, aperiodic layered materials as geometric constraint systems
+Modeling, analysis and design of self-similar, aperiodic layered materials using geometric constraint systems
 
-Many natural and engineered materials are recursively self-similar \cite{Intro1}, 
-spanning multiple scales. Within any one scale, the materials are quasiuniform, but aperiodic, 
-consisting of a few different motifs repeating in random combinations.
-In addition, many materials are stacked two dimensional (2D) layers, where each layer is independently self-similar and aperiodic.
+Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer is recursively 
+self-similar \cite{Intro1}, 
+spanning multiple scales. Within any one scale, the material layer is aperiodic, and quasi-uniform,
+consisting of a few repeated motifs appearing in disordered arrangments.
 Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
-Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. This is 
-consistent with a self-assembled structure that minimizes mass, optimally distributes external stresses and itself 
-participates in the assembly of diverse, 
-larger structures with multiple functions, all within the same layer.
+Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. These properties are
+consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself 
+participates in the assembly of diverse and multifunctional, 
+larger 2D structures.
 
@@
 
-It is natural to model such a self-similar, aperiodic material using appropriate geometric primitives under metric or algebraic constraints; design the 
-constraint-system to achieve the desired properties; and study robustness/sensitivity to constraint variations.
+In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
+within the material layer, 
+it is natural to model them using appropriate geometric primitives under metric or algebraic constraints, 
+Such  2D {\em sesamacs} (Self-similar, aperiodic material constraint systems) can be used to 
+study robustness/sensitivity to constraint variations and to design material layers 
+to achieve the desired properties.
 
 Two key observations serve as the motivations for this paper:
-As constraint systems, the above 2D {\em sesamacs} (Self-similar, aperiodic material constraint systems) involve multiple types of 
-well-studied geometric primitives and constraints: they can be modeled as so-called
+As constraint systems, 2D sesamacs involve multiple types of 
+well-studied geometric constraint systems: so-called
 {\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3).  
@@
 
-Recursive decompositions of {\em slamacs} into rigid subsystems is For , is important to understand or design the structural 
-and mechanical properties - such as rigidity, flexes and distribution of external stresses within the layer - at all scales.
+Recursive decompositions of 2D sesamacs into rigid subsystems is important in order to understand or design rigidity, flexes and distribution 
+of external stresses  -- at all scales.
 

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 Two key observations serve as the motivations for this paper:
-As constraint systems, 2D sesamacs involve multiple types of 
-well-studied geometric constraint systems: so-called
+As constraint systems, 2D sesamacs  are  well-studied: so-called
 {\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3).  
@@
 
-Recursive decompositions of 2D sesamacs into rigid subsystems is important in order to understand or design rigidity, flexes and distribution 
-of external stresses  -- at all scales.
+Especially due to their recursive structure, recursive decompositions of 2D sesamacs 
+is imperative
+in order to understand or design rigidity, flexes and distribution 
+of external stresses  -- at all scales. In fact, this has been studied combinatorially as the Decomposition-Recombination (DR)
+-planning problem \cite{XX}.  
+In particular, since many different such decompositions typically exist, is to ensure that any system is decomposed into a minimal set 
+of maximal rigid proper subsystems  
+such that they can be recombined efficiently
 

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

--- 2014-12-04 doc/GMP2015/s01_introduction.tex
@@
 
-Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer is recursively 
+Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer may be recursively 
 self-similar \cite{Intro1}, 
@@
 In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
-within the material layer, 
-it is natural to model them using appropriate geometric primitives under metric or algebraic constraints, 
-Such  2D {\em sesamacs} (Self-similar, aperiodic material constraint systems) can be used to 
-study robustness/sensitivity to constraint variations and to design material layers 
-to achieve the desired properties.
+within the material layer, it is natural to model a material layer as a solution of geometric constraint systems 
+arising from appropriate geometric primitives under metric or algebraic constraints. 
+Such  2D {\em sesamacs} (Self-similar or aperiodic material constraint systems) can be used to 
+to understand or design material layers (their solutions) with desired properties.
 
-Two key observations serve as the motivations for this paper:
-As constraint systems, 2D sesamacs involve multiple types of 
-well-studied geometric constraint systems: so-called
+As constraint systems, 2D sesamacs  are  well-studied: so-called
 {\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3).  
 All of these systems have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}) 
-for locating maximal, generically rigid components \cite{XX}.
+for locating maximal, generically rigid subsystems \cite{XX}.
 
-Recursive decompositions of 2D sesamacs into rigid subsystems is important in order to understand or design rigidity, flexes and distribution 
-of external stresses  -- at all scales.
+Whether self-similar, or just aperiodic, recursive decomposition of 2D sesamacs 
+is imperative for understanding or designing rigidity, flexes and distribution 
+of external stresses as well as robustness/sensitivity to constraint variations -- at all scales. 
+In fact, the Decomposition-Recombination (DR)
+-planning problem  for a geometric constraint system has been formalized \cite{XX} and 
+studied by many \cite{XX}. The problem is purely combinatorial  - since decomposition does not require
+knowledge of the geometry. In addition, there are efficient  algorithms 
+for optimal recombination of solutions to child systems, in order to obtain a solution to the parent system at any
+given level of the DR-plan.
+
+In particular, since many different such decompositions typically exist, is to ensure 
+that any system is decomposed into a minimal set 
+of maximal rigid proper subsystems  
+such that they can be recombined efficiently 
+ Two key observations serve as the motivations for this paper:
 

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 \section{Introduction}
-Modeling, analysis and design of self-similar, aperiodic layered materials using geometric constraint systems
+Modeling, analysis and design of quasi-uniform or self-similar material layers using geometric constraint systems
 
@@
 self-similar \cite{Intro1}, 
-spanning multiple scales. Within any one scale, the material layer is aperiodic, and quasi-uniform,
+spanning multiple scales. Within any one scale, the material layer is generally aperiodic, and quasi-uniform,
 consisting of a few repeated motifs appearing in disordered arrangments.
@@
 arising from appropriate geometric primitives under metric or algebraic constraints. 
-Such  2D {\em sesamacs} (Self-similar or aperiodic material constraint systems) can be used to 
+Such  2D {\em qusecs} (quasi-uniform or self-similar constraint systems) can be used to 
 to understand or design material layers (their solutions) with desired properties.
 
-As constraint systems, 2D sesamacs  are  well-studied: so-called
+We now briefly survey existing techniques and their limitations, for 2D {\em qusecs,} which are most naturally treated as
 {\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3). 
-All of these systems have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}) 
-for locating maximal, generically rigid subsystems \cite{XX}.
+The limitations directly motivate the contributions of this paper.
 
-Whether self-similar, or just aperiodic, recursive decomposition of 2D sesamacs 
-is imperative for understanding or designing rigidity, flexes and distribution 
+First, all of these have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}), 
+for locating maximal, generically rigid subsystems \cite{XX}. When the input is rigid, these algorithms do nothing, i.e, the output is the same as the input.
+However, both for self-similar or just aperiodic 2D qusecs, recursive decomposition into rigid subsystems
+is imperative for understanding or designing rigidity, flexes, distribution 
 of external stresses as well as robustness/sensitivity to constraint variations -- at all scales. 
-In fact, the Decomposition-Recombination (DR)
--planning problem  for a geometric constraint system has been formalized \cite{XX} and 
-studied by many \cite{XX}. The problem is purely combinatorial  - since decomposition does not require
-knowledge of the geometry. In addition, there are efficient  algorithms 
-for optimal recombination of solutions to child systems, in order to obtain a solution to the parent system at any
+In particular, recursive decomposition of rigid systems down to the geometric primitives is necessary.
+
+Second,  decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
+as the Decomposition-Recombination (DR)
+-planning problem.
+For the types of 2D qusecs listed above, generic rigidity is 
+a combinatorial property and hence decomposition does not require
+knowledge of the geometric information in the constraint system. 
+In particular, since many different such decompositions typically exist, 
+some criteria  to ensure good or optimal DR-plans were given in \cite{XX}. An optimal DR-plan is one
+that minimizes the maximum number of child subsystems, over all its parent systems, a property that is 
+crucial for obtaining a solution to the parent constraint system from the solutions of the child systems.
+For special 2D qusecs such as so-called {\em tree-decomposable} systems, optimal DR-planning is immediate.
+
+the optimal DR-planning problem was shown to be NP-hard for general 2D qusecs.
+
+Other criteria where any parent system is composed of  
+a minimal set of maximal rigid proper subsystems
+
+Third, In addition, there are general algorithms for
+recombination of  child system solutions, in order to obtain a solution to the parent system at any
 given level of the DR-plan.
 
-In particular, since many different such decompositions typically exist, is to ensure 
-that any system is decomposed into a minimal set 
-of maximal rigid proper subsystems  
-such that they can be recombined efficiently 
  Two key observations serve as the motivations for this paper:

### 2014-12-05 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

--- 2014-12-04 doc/GMP2015/s01_introduction.tex
@@
 \section{Introduction}
-Modeling, analysis and design of self-similar, aperiodic layered materials using geometric constraint systems
+Modeling, analysis and design of quasi-uniform or self-similar layered materials using geometric constraint systems
 
-Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer may be recursively 
+Some properties of natural and engineered materials can be studied by treating them as stacked two dimensional (2D) layers, 
+where the structure within each layer may be recursively 
 self-similar \cite{Intro1}, 
-spanning multiple scales. Within any one scale, the material layer is aperiodic, and quasi-uniform,
+spanning multiple scales. Within any one scale, the material layer is generally aperiodic, and quasi-uniform,
 consisting of a few repeated motifs appearing in disordered arrangments.
@@
 arising from appropriate geometric primitives under metric or algebraic constraints. 
-Such  2D {\em sesamacs} (Self-similar or aperiodic material constraint systems) can be used to 
+Such  2D {\em qusecs} (quasi-uniform or self-similar constraint systems) can be used to 
 to understand or design material layers (their solutions) with desired properties.
 
-As constraint systems, 2D sesamacs  are  well-studied: so-called
+\subsection{Previous Work}
+We now briefly survey existing techniques for studying 2D {\em qusecs,}  many of which are
 {\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3). 
-All of these systems have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}) 
-for locating maximal, generically rigid subsystems \cite{XX}.
+The limitations of these techniques directly motivate the contributions of this paper.
 
-Whether self-similar, or just aperiodic, recursive decomposition of 2D sesamacs 
-is imperative for understanding or designing rigidity, flexes and distribution 
-of external stresses as well as robustness/sensitivity to constraint variations -- at all scales. 
-In fact, the Decomposition-Recombination (DR)
--planning problem  for a geometric constraint system has been formalized \cite{XX} and 
-studied by many \cite{XX}. The problem is purely combinatorial  - since decomposition does not require
-knowledge of the geometry. In addition, there are efficient  algorithms 
-for optimal recombination of solutions to child systems, in order to obtain a solution to the parent system at any
-given level of the DR-plan.
+Fast, graph-based algorithms exist (pebble-game \cite{XX}),   
+for locating all maximal, generically rigid subsystems \cite{XX}. 
+When the input is rigid, these algorithms do nothing, i.e, the output is the same as the input.
+However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems 
+into their rigid subsystems, down to the level of geometric primitives, in order to 
+understand or design properties at all scales, such as: rigidity, flexes, distribution 
+of external stresses as well as robustness/sensitivity to constraint variations.
 
-In particular, since many different such decompositions typically exist, is to ensure 
-that any system is decomposed into a minimal set 
-of maximal rigid proper subsystems  
-such that they can be recombined efficiently 
- Two key observations serve as the motivations for this paper:
+Recursive decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
+as the Decomposition-Recombination (DR)-planning problem.
+For the above-mentioned classes of 2D qusecs, generic rigidity is 
+a combinatorial property and hence decomposition can, in principle, be done by a graph-based algorithm, without
+involving geometric information in the constraint system.  
+However, since many  such decompositions can exist for a given constraint system, 
+criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{XX}. An {\em optimal DR-plan} is one
+that minimizes the maximum number of child subsystems of any parent system, a property that is 
+crucial for obtaining a solution to the parent constraint system from the solutions of the child systems.
+
+For a special class of 2D qusecs, namely  {\em tree-decomposable} systems \cite{XX} common in computer aided mechanical design, 
+(which includes ruler-and-compass and Henneberg-I constuctible systems), all DR-plans turn out to 
+be optimal. This satisfies the so-called {\em Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
+However, for general 2D qusecs, even when restricted to bar-joint systems, 
+the optimal DR-planning problem was shown to be NP-hard \cite{XX}.
+Subsequently, alternate criteria were suggested   
+such as {\em cluster minimality} 
+requiring parent systems to be composed of  
+a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); 
+and {\em proper maximality}, requiring child subsystems
+to be maximal rigid proper subsystems of the parent system.
+While polynomial time algorithms were given \cite{XX} to generate DR-plans meeting the cluster minimality criterion,
+no such algorithm is known for the latter criterion.
+
+For fairly general constraint systems, there are algorithms that use a constraint system's DR-plan \cite{XX} for solving the system.
+They optimize the algebraic complexity of (re)combining child system solutions, 
+and algebraically solve the optimized recombination system to give a solution to the parent system at any
+given level of the DR-plan. However, the generality of these algorithms trades-off with efficiency. 
+Faster algorithms should exist for 2D cusecs that leverage recent developments on so-called Cayley configuration spaces
+\cite{XX,XX}.
+
+For underconstrained 2D cusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed \cite{XX} that 
+leverage Cayley configuration space theory to provide descriptions of connected components and find paths within the connected components 
+of the standard cartesian configuration spaces. However, no such algorithms are known for non-tree-decomposable systems.
 
 \subsection{Contributions}
+In this paper, we  navigate the NP-hardness barrier for finding optimal DR-plans by defining a 
+so-called {\it canonical} DR-plan and showing that all canonical DR-plans for isostatic 2DGMCS are optimal. 
+We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan 
+illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of 
+overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine 
+existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, 
+to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration. 
+Software implementation can be found at: \todo{Provide Link}.
 
@@
 It is known \todo{[Citation Needed]} that already the optimal DR-planning problem for general 2DGMCS is NP-hard, intuitively 
-because of a combinatorial explosion of polynomially-sized DR-plans.  In this paper, we  navigate this barrier by defining a 
-so-called {\it canonical} DR-plan and showing that all canonical DR-plans for non-overconstrained 2DGMCS are optimal. 
-We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan 
-illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of 
-overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine 
-existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, 
-to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration. 
-Software implementation can be found at: \todo{Provide Link}.
+because of a combinatorial explosion of polynomially-sized DR-plans.
 

### 2014-12-05 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 \section{Introduction}
+\label{sec:prev}
 Modeling, analysis and design of quasi-uniform or self-similar layered materials using geometric constraint systems
@@
 Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
-Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. These properties are
+Furthermore, a layer is often  either {\sl isostatic or flexible, not self-stressed}, at any scale. These properties are
 consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself 
@@
 arising from appropriate geometric primitives under metric or algebraic constraints. 
-Such  2D {\em qusecs} (quasi-uniform or self-similar constraint systems) can be used to 
+Such  2D {\bf{\em qusecs}} (quasi-uniform or self-similar constraint systems) can be used to 
 to understand or design material layers (their solutions) with desired properties.
@@
 
+\medskip\noindent\underbar{{\sl (1) Finding (vertex)-maximal generically rigid subsystems}}
 Fast, graph-based algorithms exist (pebble-game \cite{XX}),   
@@
 
+\medskip\noindent\underbar{{\sl (2) Optimal Recursive Decomposition (DR-planning)}}
 Recursive decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
-as the Decomposition-Recombination (DR)-planning problem.
+as the {\sl Decomposition-Recombination (DR)-planning} problem.
 For the above-mentioned classes of 2D qusecs, generic rigidity is 
-a combinatorial property and hence decomposition can, in principle, be done by a graph-based algorithm, without
+a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without
 involving geometric information in the constraint system.  
-However, since many  such decompositions can exist for a given constraint system, 
+Since many  such decompositions can exist for a given constraint system, 
 criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{XX}. An {\em optimal DR-plan} is one
@@
 crucial for obtaining a solution to the parent constraint system from the solutions of the child systems.
+However, for general 2D qusecs, even when restricted to bar-joint systems, 
+the optimal DR-planning problem was shown to be NP-hard \cite{XX}.
 
+\medskip\noindent\underbar{{\sl (3) DR-plans for special classes and with other criteria}}
 For a special class of 2D qusecs, namely  {\em tree-decomposable} systems \cite{XX} common in computer aided mechanical design, 
@@
 be optimal. This satisfies the so-called {\em Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
-However, for general 2D qusecs, even when restricted to bar-joint systems, 
-the optimal DR-planning problem was shown to be NP-hard \cite{XX}.
 Subsequently, alternate criteria were suggested   
@@
 
-For fairly general constraint systems, there are algorithms that use a constraint system's DR-plan \cite{XX} for solving the system.
-They optimize the algebraic complexity of (re)combining child system solutions, 
+\medskip\noindent\underbar{{\sl (4) Optimal Recombination and Solution Space Navigation}}
+For fairly general 3D constraint systems, there are algorithms with formal guarantees that 
+solve the parent (sub)systems that occur in the DR-plan \cite{XX, XX,XX}, 
+provided the DR-plan meets some of the above criteria.
+These algorithms  uncover underlying matroids to combinatorially optimize the algebraic complexity of (re)combining child system solutions, 
 and algebraically solve the optimized recombination system to give a solution to the parent system at any
-given level of the DR-plan. However, the generality of these algorithms trades-off with efficiency. 
-Faster algorithms should exist for 2D cusecs that leverage recent developments on so-called Cayley configuration spaces
-\cite{XX,XX}.
+given level of the DR-plan. However, the generality of these algorithms trades-off against efficiency, and despite the optimization,
+the best algorithms can still take
+exponential time in the number of child subsystems in order to guarantee all solutions even for a single (sub)system in a DR-plan, 
+and are prohibitively slow in practice. For the entire DR-plan, finding all solutions is barely tractable even with fast recombination algorithms
+for the subsystems in the DR-plan. This is because even for the simplest systems, the problem is NP-hard \cite{XX} and there is a 
+combinatorial explosion of solutions \cite{XX}.  Using the DR-plan and optimal recombination as a principled basis, high performance software exists \cite{XX,XX} to 
+tame combinatorial explosion via user intervention.
 
-For underconstrained 2D cusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed \cite{XX} that 
-leverage Cayley configuration space theory to provide descriptions of connected components and find paths within the connected components 
-of the standard cartesian configuration spaces. However, no such algorithms are known for non-tree-decomposable systems.
+\medskip\noindent\underbar{{\sl (5) Configuration Spaces of Underconstrained Systems}}
+For underconstrained 2D bar-joint and multibody-pin cusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed \cite{XX, XX} that 
+to complete them into isostatic systems \cite{XX, XX} and to find paths within the connected components \cite{XX,XX} of
+standard cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{XX,XX}
+to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation 
+of connected components and paths.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, starting with
+underlying partial 2-tree graphs (alternately, tree-width 2, series-parallel, $K_4$ minor avoiding),  moving on   
+to 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), 
+and further to general 1-dof tree-decomposable graphs. 
+While software suites exist \cite{XX,XX}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
 
 \subsection{Contributions}
-In this paper, we  navigate the NP-hardness barrier for finding optimal DR-plans by defining a 
-so-called {\it canonical} DR-plan and showing that all canonical DR-plans for isostatic 2DGMCS are optimal. 
+\label{sec:cont}
+In this paper, we  navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a 
+so-called {\it canonical} DR-plan and showing that all canonical DR-plans for isostatic 2D qusecs are optimal. 
 We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan 

### 2014-12-05 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

@@
 
-\subsection{Previous Work}
+\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
 We now briefly survey existing techniques for studying 2D {\em qusecs,}  many of which are
@@
 understand or design properties at all scales, such as: rigidity, flexes, distribution 
-of external stresses as well as robustness/sensitivity to constraint variations.
+of external stresses,  well as robustness/sensitivity to constraint variations.
 
@@
 In this paper, we  navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a 
-so-called {\it canonical} DR-plan and showing that all canonical DR-plans for isostatic 2D qusecs are optimal. 
-We then give an efficient ($O(n^2)$) algorithm to find a canonical DR-plan for all 2DGMCS. The notion of a canonical DR-plan 
-illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained 2DGMCS and gives classes of 
+so-called {\it canonical} DR-plan and showing a strong Church-Rosser property: {\it all canonical DR-plans for isostatic or underconstrained (independent) 
+2D qusecs are optimal}. 
+We then give an efficient ($O(n^2)$) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
+canonical DR-plan 
+illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained  and gives classes of 
 overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine 

### 2014-12-05 src/Graph/Graph.cpp
Trying to get component detection. It's close.

@@ -287,6 +287,10 @@ unsigned int Graph::num_vertices() {
     return boost::num_vertices(*this);
 }
 
+unsigned int Graph::num_edges() {
+    return boost::num_edges(*this);
+}
+
 std::pair<Vertex_ID, Vertex_ID> Graph::verts_on_edge(Edge_Iterator e) {
     return std::make_pair(boost::source(*e, *this), boost::target(*e, *this));
 }

### 2014-12-05 src/Graph/Graph.hpp

@@ -3,6 +3,7 @@
 
 #include <boost/graph/graph_traits.hpp>
 #include <boost/graph/undirected_graph.hpp>
+#include <boost/graph/subgraph.hpp>
 // #include <boost/graph/adjacency_list.hpp>
 
 // for layout
@@ -87,6 +88,7 @@ public:
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
     unsigned int num_vertices();
+    unsigned int num_edges();
 
     std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_Iterator e);
 
@@ -95,6 +97,11 @@ public:
     void write_to_file(const char* filename);
     void read_from_file(const char* filename);
 };
+
+
+
+typedef boost::subgraph<Graph> Subgraph_Type;
+
 
 
 #endif

### 2014-12-05 src/Graph/Pebbled_Graph copy.cpp

@@ -54,14 +54,6 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
     // init pebbles
     reset_pebbles();
 
-    // for differencing
-    std::set<Vertex_ID> all_vertices;
-    std::pair<Vertex_Iterator, Vertex_Iterator> vs = this->g->vertices();
-    for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
-        all_vertices.insert(*v_it);
-    }
-
-
     // the clusters found
     std::set<Cluster*> cs;
 
@@ -85,7 +77,7 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
         std::cout << "    Pebbles before: "; print_all_verts_with_pebbles();
 
         // if edge can't be added, move on
-        if (!enlarge_cover(it)) {
+        if (!component_enlarge_cover(it)) {
             std::cout << "    Pebbles after : "; print_all_verts_with_pebbles();
             continue;
         }
@@ -93,54 +85,11 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
         std::cout << "  Successfully added." << std::endl;;
 
 
-
-
-        // work on v1
-        while (num_pebbles_on_vert(v1) != this->k &&
-            num_pebbles_on_vert(v1) + num_pebbles_on_vert(v2) < this->l)
-        {
-            std::map<Vertex_ID, bool> seen;
-
-            std::pair<Vertex_Iterator, Vertex_Iterator> vs = g->vertices();
-            for (Vertex_Iterator it = vs.first; it != vs.second; it++)
-            {
-                seen[*it] = false;
-            }
-
-            // leave if you can't get any more pebbles
-            if (!make_pebble_available(v1, &seen, &v2))
-                break;
-        }
-
-
-        // work on v2
-        while (num_pebbles_on_vert(v2) != this->k &&
-            num_pebbles_on_vert(v1) + num_pebbles_on_vert(v2) < this->l)
-        {
-            std::map<Vertex_ID, bool> seen;
-
-            std::pair<Vertex_Iterator, Vertex_Iterator> vs = g->vertices();
-            for (Vertex_Iterator it = vs.first; it != vs.second; it++)
-            {
-                seen[*it] = false;
-            }
-            // seen[v1] = true;
-
-            // leave if you can't get any more pebbles
[…]

### 2014-12-05 src/Graph/Pebbled_Graph copy.hpp

@@ -6,9 +6,18 @@
 #include <map>
 #include <vector>
 #include <array>
+#include <set>
 
 
-typedef std::vector<std::vector<Vertex_ID> > Clusters;
+struct Cluster {
+    std::set<Vertex_ID> vertices;
+    std::set<Edge_ID> edges;
+
+    void add_vertex(Vertex_ID v) {vertices.insert(v);}
+    void add_edge(Edge_ID e) {edges.insert(e);}
+
+    unsigned int num_edges() {return edges.size();}
+};
 
 
 class Pebbled_Graph {
@@ -17,14 +26,17 @@ public:
 
     // if there are exactly 2 pebbles free in the end, the graph is rigid
     unsigned int pebble_game_2D();
+    // if there is exactly 1 cluster, the graph is rigid
+    std::set<Cluster*> component_pebble_game_2D();
 
-    Clusters DRP_2D();
+    std::set<Cluster*> DRP_2D();
 
 
 private:
     // pebbles per node = k = 2
     // pebbles for sparsity/tightness = l = 3
     // solving for (k, l)-sparse graphs
+    // ONLY GUARANTEED TO WORK ON (2,3)
     static const unsigned int k = 2;
     static const unsigned int l = 3;
 
@@ -51,8 +63,8 @@ private:
 
 
 
-    // a flag to prevent repeat work and infinite recursion
-    std::map<Vertex_ID, bool> seen;
+    // // a flag to prevent repeat work and infinite recursion
+    // std::map<Vertex_ID, bool> seen;
 
     // A vertex and an important pebble (index)
     struct vert_peb {
@@ -61,9 +73,9 @@ private:
         vert_peb() {}
         vert_peb(Vertex_ID v, unsigned int p) {vertex = v; pebble = p;}
     };
-    // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
-    // used to reconstruct the path for rearranging.
-    std::map<Vertex_ID, vert_peb> path;
+    // // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
+    // // used to reconstruct the path for rearranging.
+    // std::map<Vertex_ID, vert_peb> path;
 
 
     // enum pebble {
@@ -90,11 +102,28 @@ private:
     // returns the index of a free pebble or k if there is none
     unsigned int identify_free_pebble(Vertex_ID v);
 
+    // returns the number of pebbles free on given vertex
[…]

### 2014-12-05 src/Graph/Pebbled_Graph.hpp

@@ -6,9 +6,18 @@
 #include <map>
 #include <vector>
 #include <array>
+#include <set>
 
 
-typedef std::vector<std::vector<Vertex_ID> > Clusters;
+struct Cluster {
+    std::set<Vertex_ID> vertices;
+    std::set<Edge_ID> edges;
+
+    void add_vertex(Vertex_ID v) {vertices.insert(v);}
+    void add_edge(Edge_ID e) {edges.insert(e);}
+
+    unsigned int num_edges() {return edges.size();}
+};
 
 
 class Pebbled_Graph {
@@ -17,8 +26,10 @@ public:
 
     // if there are exactly 2 pebbles free in the end, the graph is rigid
     unsigned int pebble_game_2D();
+    // if there is exactly 1 cluster, the graph is rigid
+    std::set<Cluster*> component_pebble_game_2D();
 
-    Clusters DRP_2D();
+    std::set<Cluster*> DRP_2D();
 
 
 private:
@@ -52,8 +63,8 @@ private:
 
 
 
-    // a flag to prevent repeat work and infinite recursion
-    std::map<Vertex_ID, bool> seen;
+    // // a flag to prevent repeat work and infinite recursion
+    // std::map<Vertex_ID, bool> seen;
 
     // A vertex and an important pebble (index)
     struct vert_peb {
@@ -94,14 +105,25 @@ private:
     // returns the number of pebbles free on given vertex
     unsigned int num_pebbles_on_vert(Vertex_ID v);
 
-    // gives the inverted directed pebble game graph
-    std::map<Vertex_ID, pebbles> get_reversed_pebbles();
+
+    void print_verts(std::set<Vertex_ID> &v);
+    void print_all_verts_with_pebbles();
+
+
 
     // Finds a free pebble if there is one and sets up the path to it.
-    bool find_pebble(Vertex_ID v, std::map<Vertex_ID, vert_peb> *path);
+    bool find_pebble(
+        Vertex_ID v,
+        std::map<Vertex_ID, bool> *seen,
+        std::map<Vertex_ID, vert_peb> *path,
+        Vertex_ID *dont_take_pebble_from = NULL);
 
     // Frees a pebble at vertex_ID v, if there is one. Returns true if successful
-    bool make_pebble_available(Vertex_ID v);
+    bool make_pebble_available(
+        Vertex_ID v,
+        std::map<Vertex_ID, bool> *seen,
[…]

### 2014-12-05 src/Graph/Pebbled_Graph.old.cpp

@@ -1,227 +1,0 @@
-#include "Pebbled_Graph.hpp"
-
-#include <vector>
-
-
-
-Pebbled_Graph::Pebbled_Graph(Graph *g) {
-    this->g = g;
-}
-
-// if there are exactly 2 pebbles free in the end, the graph is rigid
-unsigned int Pebbled_Graph::pebble_game_2D() {
-    reset_pebbles();
-
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
-    {
-        this->enlarge_cover(it);
-    }
-
-    return _total_pebbles_available;
-}
-
-Clusters Pebbled_Graph::DRP_2D() {
-    std::map<Vertex_ID, Clusters> clusters_for_excluded_vert;
-
-    std::pair<Vertex_Iterator, Vertex_Iterator> vs = this->g->vertices();
-    for (Vertex_Iterator it = vs.first; it != vs.second; it++)
-    {
-        clusters_for_excluded_vert[*it] = this->pebble_game_2D_exclude(*it);
-    }
-
-    // this->pebble_game_2D();
-
-    return clusters_for_excluded_vert[0];
-}
-
-
-Clusters Pebbled_Graph::pebble_game_2D_exclude(Vertex_ID excluded_vert) {
-    // init pebbles
-    reset_pebbles();
-    _total_pebbles_available -= this->k;
-
-    // init cluster map
-    std::map<Vertex_ID, unsigned int> cluster_id_for_vert;
-    unsigned int current_cluster_id = 0;
-    Clusters cs;
-
-
-    //
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
-    {
-        std::pair<Vertex_ID, Vertex_ID> vs_on_e = g->verts_on_edge(it);
-        Vertex_ID v1 = vs_on_e.first, v2 = vs_on_e.second;
-
-        // check if this edge lies on the excluded vertex
-        if (v1 == excluded_vert || v2 == excluded_vert)
-            continue;
-
-        // get the cluster_ids
-        std::map<Vertex_ID, unsigned int>::iterator v1_cid
-            = cluster_id_for_vert.find(v1);
-        std::map<Vertex_ID, unsigned int>::iterator v2_cid
-            = cluster_id_for_vert.find(v2);
-
[…]

### 2014-12-05 src/main.cpp

@@ -288,9 +288,12 @@ int main(int argc, char **argv) {
     g.read_from_file(dot_file.c_str());
 
     Graph copy(g);
-    g.write_to_file("out.dot");
+    g.write_to_file("test_files/out.dot");
 
     g.set_layout();
+
+    // Subgraph_Type sg;
+    // sg.read_from_file(dot_file.c_str());
 
 
 
@@ -310,8 +313,8 @@ int main(int argc, char **argv) {
 
 
     Pebbled_Graph pg(&g);
-    cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
-
+    // cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
+    pg.component_pebble_game_2D();
 
 
 

### 2014-12-06 src/Graph/Graph.cpp
Component detection works (although inefficiently) and it can do the first level of the DR-plan.

@@ -1,3 +1,290 @@
+
+
+#include "Graph.hpp"
+
+
+
+// #include <boost/graph/topological_sort.hpp>
+
+
+#include <boost/graph/copy.hpp>
+
+
+
+// for layout
+#include <boost/config.hpp>
+// #include <boost/graph/adjacency_list.hpp>
+#include <boost/graph/graph_utility.hpp>
+#include <boost/graph/simple_point.hpp>
+#include <boost/property_map/property_map.hpp>
+
+
+#include <boost/graph/circle_layout.hpp>
+#include <boost/graph/kamada_kawai_spring_layout.hpp>
+
+#include <boost/graph/random_layout.hpp>
+#include <boost/graph/fruchterman_reingold.hpp>
+
+
+#include <boost/graph/graphviz.hpp>
+
+#include <boost/graph/iteration_macros.hpp>
+
+
+
+#include <iostream>
+#include <fstream>
+#include <map>
+#include <vector>
+
+
+
+// Graph::Graph(const Graph& other) {
+//     // // boost::copy_graph<Graph, Graph>(other, *this);
+
+
+
+//     // typedef std::map<Vertex_ID, size_t> IndexMap;
+//     // IndexMap mapIndex;
+//     // boost::associative_property_map<IndexMap> propmapIndex(mapIndex);
+
+//     // int i=0;
+//     // std::pair<Vertex_Iterator, Vertex_Iterator> vs = other.vertices();
+//     // for (Vertex_Iterator it = vs.first; it != vs.second; it++)
+//     // {
+//     //     put(propmapIndex, *it, i++);
+//     // }
+
+//     // this->clear();
+//     // boost::copy_graph(other, *this, vertex_index_map( propmapIndex ) );
+
+
+
+
+//     this->clear();
+
+//     std::vector<Vertex_ID> verts;
+
+//     std::pair<Vertex_Iterator, Vertex_Iterator> vs = other.vertices();
+//     for (Vertex_Iterator it = vs.first; it != vs.second; it++)
+//     {
+//         verts.push_back(this->add_vertex(other[*it]));
+//     }
+
+//     std::pair<Edge_Iterator, Edge_Iterator> es = other.edges();
+//     for (Edge_Iterator it = es.first; it != es.second; it++)
+//     {
+//         std::pair<Vertex_ID, Vertex_ID> vs_on_e = verts_on_edge(it);
+//         this->add_edge(vs_on_e.first, vs_on_e.second, other[*it]);
[…]

### 2014-12-06 src/Graph/Graph.hpp

@@ -36,16 +36,16 @@ struct Vertex_Properties {
 
 
 struct Edge_Properties {
-    double distance;
+    double length;
     // unsigned int pebbles;
 
     Edge_Properties() {
-        distance = 1;
+        length = 1;
         // pebbles = 0;
     }
 
     Edge_Properties(double d) {
-        distance = d;
+        length = d;
         // pebbles = 0;
     }
 };

### 2014-12-06 src/Graph/Pebbled_Graph copy.cpp

@@ -1,711 +1,0 @@
-#include "Pebbled_Graph.hpp"
-
-#include <vector>
-#include <queue>
-#include <algorithm>
-
-#include <iostream>
-
-
-
-
-
-
-
-
-
-Pebbled_Graph::Pebbled_Graph(Graph *g) {
-    this->g = g;
-}
-
-// if there are exactly 2 pebbles free in the end, the graph is rigid
-unsigned int Pebbled_Graph::pebble_game_2D() {
-    reset_pebbles();
-
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
-    {
-        this->enlarge_cover(it);
-    }
-
-    return _total_pebbles_available;
-}
-
-
-void Pebbled_Graph::print_verts(std::set<Vertex_ID> &v) {
-    for (std::set<Vertex_ID>::iterator v_it = v.begin(); v_it != v.end(); v_it++) {
-        std::cout << (*g)[*v_it].name << " ";
-    }
-    std::cout << std::endl;
-}
-
-void Pebbled_Graph::print_all_verts_with_pebbles() {
-    std::pair<Vertex_Iterator, Vertex_Iterator> vs = this->g->vertices();
-    for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
-        std::cout << (*g)[*v_it].name << "("
-            << (*g)[pebbles_at_vertex[*v_it][0]].name
-            << (*g)[pebbles_at_vertex[*v_it][1]].name << ") ";
-    }
-    std::cout << std::endl;
-}
-
-
-std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
-    // init pebbles
-    reset_pebbles();
-
-    // the clusters found
-    std::set<Cluster*> cs;
-
-    // true if they're in the same cluster, not in map if not
-    std::map<std::pair<Vertex_ID, Vertex_ID>, bool> in_same_clust;
-
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
-    {
-        std::pair<Vertex_ID, Vertex_ID> vs_on_e = g->verts_on_edge(it);
-        Vertex_ID v1 = vs_on_e.first, v2 = vs_on_e.second;
-
-        std::cout << "Edge ("<< (*g)[v1].name << "--" << (*g)[v2].name << "): " << std::endl;;
-
[…]

### 2014-12-06 src/Graph/Pebbled_Graph copy.hpp

@@ -1,161 +1,0 @@
-#ifndef PEBBLED_GRAPH_HPP
-#define PEBBLED_GRAPH_HPP
-
-#include "Graph.hpp"
-
-#include <map>
-#include <vector>
-#include <array>
-#include <set>
-
-
-struct Cluster {
-    std::set<Vertex_ID> vertices;
-    std::set<Edge_ID> edges;
-
-    void add_vertex(Vertex_ID v) {vertices.insert(v);}
-    void add_edge(Edge_ID e) {edges.insert(e);}
-
-    unsigned int num_edges() {return edges.size();}
-};
-
-
-class Pebbled_Graph {
-public:
-    Pebbled_Graph(Graph *g);
-
-    // if there are exactly 2 pebbles free in the end, the graph is rigid
-    unsigned int pebble_game_2D();
-    // if there is exactly 1 cluster, the graph is rigid
-    std::set<Cluster*> component_pebble_game_2D();
-
-    std::set<Cluster*> DRP_2D();
-
-
-private:
-    // pebbles per node = k = 2
-    // pebbles for sparsity/tightness = l = 3
-    // solving for (k, l)-sparse graphs
-    // ONLY GUARANTEED TO WORK ON (2,3)
-    static const unsigned int k = 2;
-    static const unsigned int l = 3;
-
-
-
-
-
-    // the graph we're playing the game on
-    Graph *g;
-
-
-
-    // An array to store the Vertex_ID that a pebble is associated with
-    typedef std::array<Vertex_ID, k> pebbles;
-
-    // Used to keep track of each vertex's pebbles. Say pebbles_at_vertex[v][i] = n
-    // if v == n, then the pebble is available (on the vertex v). Else if v != n
-    // then it's on the edge between v and n
-    std::map<Vertex_ID, pebbles> pebbles_at_vertex;
-
-    // starts at 2 * the number of vertices
-    unsigned int _total_pebbles_available;
-
-
-
-
-    // // a flag to prevent repeat work and infinite recursion
-    // std::map<Vertex_ID, bool> seen;
-
-    // A vertex and an important pebble (index)
-    struct vert_peb {
-        Vertex_ID vertex;
-        unsigned int pebble;
-        vert_peb() {}
-        vert_peb(Vertex_ID v, unsigned int p) {vertex = v; pebble = p;}
-    };
-    // // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
-    // // used to reconstruct the path for rearranging.
[…]

### 2014-12-06 src/Graph/Pebbled_Graph.cpp

@@ -32,25 +32,8 @@ unsigned int Pebbled_Graph::pebble_game_2D() {
 }
 
 
-void Pebbled_Graph::print_verts(std::set<Vertex_ID> &v) {
-    for (std::set<Vertex_ID>::iterator v_it = v.begin(); v_it != v.end(); v_it++) {
-        std::cout << (*g)[*v_it].name << " ";
-    }
-    std::cout << std::endl;
-}
 
-void Pebbled_Graph::print_all_verts_with_pebbles() {
-    std::pair<Vertex_Iterator, Vertex_Iterator> vs = this->g->vertices();
-    for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
-        std::cout << (*g)[*v_it].name << "("
-            << (*g)[pebbles_at_vertex[*v_it][0]].name
-            << (*g)[pebbles_at_vertex[*v_it][1]].name << ") ";
-    }
-    std::cout << std::endl;
-}
-
-
-std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
+std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D(Vertex_ID *exclude) {
     // init pebbles
     reset_pebbles();
 
@@ -74,23 +57,26 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D() {
         std::pair<Vertex_ID, Vertex_ID> vs_on_e = g->verts_on_edge(it);
         Vertex_ID v1 = vs_on_e.first, v2 = vs_on_e.second;
 
-        std::cout << "Edge ("<< (*g)[v1].name << "--" << (*g)[v2].name << "): " << std::endl;;
+        // std::cout << "Edge ("<< (*g)[v1].name << "--" << (*g)[v2].name << "): " << std::endl;;
+
+        if (exclude != NULL && (v1 == *exclude || v2 == *exclude))
+            continue;
 
         // if both verts are already in the cluter, no need to add the edge
         // wait, is this even possible?! in wellconstrained? I don't think so
         if (in_same_clust.find(vs_on_e) != in_same_clust.end())
             continue;
 
-        std::cout << "  Total pebbles remaining: " << _total_pebbles_available << std::endl;
-        std::cout << "    Pebbles before: "; print_all_verts_with_pebbles();
+        // std::cout << "  Total pebbles remaining: " << _total_pebbles_available << std::endl;
[…]

### 2014-12-06 src/Graph/Pebbled_Graph.hpp

@@ -27,7 +27,7 @@ public:
     // if there are exactly 2 pebbles free in the end, the graph is rigid
     unsigned int pebble_game_2D();
     // if there is exactly 1 cluster, the graph is rigid
-    std::set<Cluster*> component_pebble_game_2D();
+    std::set<Cluster*> component_pebble_game_2D(Vertex_ID *exclude = NULL);
 
     std::set<Cluster*> DRP_2D();
 
@@ -112,26 +112,34 @@ private:
 
 
     // Finds a free pebble if there is one and sets up the path to it.
-    bool find_pebble(
+    bool _find_pebble_aux(
         Vertex_ID v,
         std::map<Vertex_ID, bool> *seen,
         std::map<Vertex_ID, vert_peb> *path,
-        Vertex_ID *dont_take_pebble_from = NULL);
+        Vertex_ID *dont_take_pebble_from = NULL,
+        Vertex_ID *exclude = NULL);
+    bool find_pebble(
+        Vertex_ID v,
+        std::map<Vertex_ID, vert_peb> *path,
+        Vertex_ID *dont_take_pebble_from = NULL,
+        Vertex_ID *exclude = NULL);
 
     // Frees a pebble at vertex_ID v, if there is one. Returns true if successful
     bool make_pebble_available(
         Vertex_ID v,
-        std::map<Vertex_ID, bool> *seen,
-        Vertex_ID *dont_take_pebble_from = NULL);
+        Vertex_ID *dont_take_pebble_from = NULL,
+        Vertex_ID *exclude = NULL);
 
+    // true means the edge was added to the cover
+    bool enlarge_cover(
+        Edge_Iterator e,
+        Vertex_ID *exclude = NULL);
+    // bool component_enlarge_cover(Edge_Iterator e);
 
     // takes one of origin's available pebbles and places it on the edge to
     // the destination vertex_ID. Returns false if there was no available pebble
     bool place_available_pebble(Vertex_ID origin, Vertex_ID destination);
 
-    // true means the edge was added to the cover
-    bool enlarge_cover(Edge_Iterator e);
-    bool component_enlarge_cover(Edge_Iterator e);
 
 
 

### 2014-12-06 src/main.cpp

@@ -233,6 +233,8 @@ void display_graph(Window &wind, Graph &g) {
 int main(int argc, char **argv) {
     string dot_file = "test_files/test.dot";
 
+    int runtime_option = 0;
+
     // handle command-line arguments
     for (int i = 1; i < argc; i++) {
         string inp = argv[i];
@@ -259,6 +261,19 @@ int main(int argc, char **argv) {
                     return 0;
                 }
                 dot_file = argv[i];
+            } else {
+                cerr << "Invalid value for flag " << flag << "." << endl;
+                return 0;
+            }
+        }
+
+        else if (!flag.compare("-opt")) {
+            if (!val.compare("0")) {
+                runtime_option = 0;
+            } else if (!val.compare("1")) {
+                runtime_option = 1;
+            } else if (!val.compare("2")) {
+                runtime_option = 2;
             } else {
                 cerr << "Invalid value for flag " << flag << "." << endl;
                 return 0;
@@ -313,9 +328,15 @@ int main(int argc, char **argv) {
 
 
     Pebbled_Graph pg(&g);
-    // cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
-    pg.component_pebble_game_2D();
 
+    if (runtime_option == 0) {
+        pg.component_pebble_game_2D();
+    } else if (runtime_option == 1) {
+        pg.DRP_2D();
+    } else if (runtime_option == 2) {
+        int pebbles_remaining = pg.pebble_game_2D();
+        cout << "Pebbles remaining: " << pebbles_remaining << endl;
+    }
 
 
     while (!myWindow.should_close()) {

### 2014-12-10 src/Graph/Graph.cpp
Now uses subgraphs. Ready for the next step: get the whole DR plan.

@@ -37,6 +37,8 @@
 #include <map>
 #include <vector>
 
+// #include <cstring>
+
 
 
 // Graph::Graph(const Graph& other) {
@@ -115,6 +117,18 @@ std::pair<Edge_Iterator, Edge_Iterator> Graph::edges() const {
     return boost::edges(*this);
 }
 
+Vertex_Iterator Graph::find_vertex(const char *name) {
+    Vertex_Iterator v, v_end;
+    for (boost::tie(v, v_end) = this->vertices(); v != v_end; v++) {
+        if ((*this)[*v].name.compare(name) == 0) {
+            return v;
+        }
+    }
+
+    return v_end;
+}
+
+
 unsigned int Graph::num_vertices() {
     return boost::num_vertices(*this);
 }
@@ -123,9 +137,10 @@ unsigned int Graph::num_edges() {
     return boost::num_edges(*this);
 }
 
-std::pair<Vertex_ID, Vertex_ID> Graph::verts_on_edge(Edge_Iterator e) {
-    return std::make_pair(boost::source(*e, *this), boost::target(*e, *this));
+std::pair<Vertex_ID, Vertex_ID> Graph::verts_on_edge(Edge_ID e) {
+    return std::make_pair(boost::source(e, *this), boost::target(e, *this));
 }
+
 
 
 
@@ -433,3 +448,134 @@ void Graph::read_from_file(const char* filename) {
 //   std::cout << "Triangle layout (Kamada-Kawai).\n";
 //   print_graph_layout(g, get(vertex_position, g));
 // }
+
+
+
+
+
+
+
+Subgraph::Subgraph(Graph *g) {
+    _graph = g;
+}
+
+const Graph* Subgraph::graph() {
+    return _graph;
+}
+
+// void Subgraph::induce(Vertex_Iterator begin, Vertex_Iterator end) {
+//     for (Vertex_Iterator v_it = begin; v_it != end; v_it++) {
+//         _vertices.insert(*v_it);
+//     }
+
+
+//     std::pair<Edge_Iterator, Edge_Iterator> es = _graph->edges();
+//     for (Edge_Iterator e_it = es.first; e_it != es.second; e_it++) {
+//         std::pair<Vertex_ID, Vertex_ID> vs = _graph->verts_on_edge(e_it);
+//         if (_vertices.find(vs.first) != _vertices.end()
+//             && _vertices.find(vs.second) != _vertices.end())
+//         {
[…]

### 2014-12-10 src/Graph/Graph.hpp

@@ -12,6 +12,10 @@ typedef boost::square_topology<>::point_type Point;
 
 
 
+#include <set>
+#include <string>
+
+
 
 struct Graph_Properties { };
 
@@ -87,10 +91,12 @@ public:
     std::pair<Vertex_Iterator, Vertex_Iterator> vertices() const;
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
+    Vertex_Iterator find_vertex(const char *name);
+
     unsigned int num_vertices();
     unsigned int num_edges();
 
-    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_Iterator e);
+    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_ID e);
 
     void set_layout();
 
@@ -100,7 +106,40 @@ public:
 
 
 
-typedef boost::subgraph<Graph> Subgraph_Type;
+// typedef boost::subgraph<Graph> Subgraph_Type;
+
+
+
+typedef std::set<Vertex_ID>::iterator Sg_Vertex_Iterator;
+typedef std::set<Edge_ID>::iterator Sg_Edge_Iterator;
+
+// An induced subgraph of the input graph. Initially empty.
+class Subgraph {
+public:
+    Subgraph(Graph *g);
+
+    void induce(Vertex_Iterator begin, Vertex_Iterator end);
+    void induce(std::set<Vertex_ID> *vertices);
+
+    void add_vertex(Vertex_ID vertex);
+    void remove_vertex(Vertex_ID vertex);
+
+    const Graph* graph();
+    std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> vertices();
+    std::pair<Sg_Edge_Iterator, Sg_Edge_Iterator> edges();
+
+    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_ID e);
+
+    unsigned int num_vertices();
+    unsigned int num_edges();
+
+    Vertex_Properties& operator[](Vertex_ID vertex);
+private:
+    Graph *_graph;
+
+    std::set<Vertex_ID> _vertices;
+    std::set<Edge_ID> _edges;
+};
 
 
 

### 2014-12-10 src/Graph/Pebbled_Graph.cpp

@@ -15,17 +15,23 @@
 
 
 Pebbled_Graph::Pebbled_Graph(Graph *g) {
-    this->g = g;
+    this->in_subgraph = new Subgraph(g);
+    std::pair<Vertex_Iterator, Vertex_Iterator> vs = g->vertices();
+    in_subgraph->induce(vs.first, vs.second);
+}
+
+Pebbled_Graph::Pebbled_Graph(Subgraph *g) {
+    this->in_subgraph = g;
 }
 
 // if there are exactly 2 pebbles free in the end, the graph is rigid
 unsigned int Pebbled_Graph::pebble_game_2D() {
     reset_pebbles();
 
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
+    std::pair<Sg_Edge_Iterator, Sg_Edge_Iterator> es = this->in_subgraph->edges();
+    for (Sg_Edge_Iterator it = es.first; it != es.second; it++)
     {
-        this->enlarge_cover(it);
+        this->enlarge_cover(*it);
     }
 
     return _total_pebbles_available;
@@ -37,10 +43,11 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D(Vertex_ID *exclude) {
     // init pebbles
     reset_pebbles();
 
+    // std::cout << "Pebbled_Graph::component_pebble_game_2D: here0" << std::endl;
     // for differencing
     std::set<Vertex_ID> all_vertices;
-    std::pair<Vertex_Iterator, Vertex_Iterator> vs = this->g->vertices();
-    for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
+    std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> vs = this->in_subgraph->vertices();
+    for (Sg_Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
         all_vertices.insert(*v_it);
     }
 
@@ -51,10 +58,10 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D(Vertex_ID *exclude) {
     // true if they're in the same cluster, not in map if not
     std::map<std::pair<Vertex_ID, Vertex_ID>, bool> in_same_clust;
 
-    std::pair<Edge_Iterator, Edge_Iterator> es = this->g->edges();
-    for (Edge_Iterator it = es.first; it != es.second; it++)
[…]

### 2014-12-10 src/Graph/Pebbled_Graph.hpp

@@ -23,6 +23,7 @@ struct Cluster {
 class Pebbled_Graph {
 public:
     Pebbled_Graph(Graph *g);
+    Pebbled_Graph(Subgraph *g);
 
     // if there are exactly 2 pebbles free in the end, the graph is rigid
     unsigned int pebble_game_2D();
@@ -45,7 +46,7 @@ private:
 
 
     // the graph we're playing the game on
-    Graph *g;
+    Subgraph *in_subgraph;
 
 
 
@@ -58,13 +59,10 @@ private:
     std::map<Vertex_ID, pebbles> pebbles_at_vertex;
 
     // starts at 2 * the number of vertices
+    // set in reset_pebbles, decremented in place_available_pebble
     unsigned int _total_pebbles_available;
 
 
-
-
-    // // a flag to prevent repeat work and infinite recursion
-    // std::map<Vertex_ID, bool> seen;
 
     // A vertex and an important pebble (index)
     struct vert_peb {
@@ -73,28 +71,6 @@ private:
         vert_peb() {}
         vert_peb(Vertex_ID v, unsigned int p) {vertex = v; pebble = p;}
     };
-    // // typedef std::pair<Vertex_ID, unsigned int> vert_peb;
-    // // used to reconstruct the path for rearranging.
-    // std::map<Vertex_ID, vert_peb> path;
-
-
-    // enum pebble {
-    //     p1, p2
-    // };
-
-    // // the edge a pebble from a vertex_ID is on, represented as the vertex_ID it's
-    // // pointing to. If it maps to nothing, then the pebble is available
-    // typedef std::pair<Vertex_ID, pebble> vert_peb;
-    // std::map<vert_peb, Vertex_ID> pebbles_on_verts;
-
-    // // set in the constructor, decremented in place_available_pebble
-    // unsigned int _total_pebbles_available;
-
-
-    // // used to reconstruct the path for rearranging. the key is the vertex_ID
-    // // the value is the pair of the next vertex_ID in the path and the pebble
-    // // used from the original to cover the edge.
-    // std::map<Vertex_ID, vert_peb> path;
 
 
 
@@ -132,7 +108,7 @@ private:
 
     // true means the edge was added to the cover
     bool enlarge_cover(
-        Edge_Iterator e,
+        Edge_ID e,
         Vertex_ID *exclude = NULL);
[…]

### 2014-12-10 src/main.cpp

@@ -175,7 +175,7 @@ vector<gl_obj::pos_vec> edges_from_graph(Graph &g) {
         edges.first != edges.second;
         edges.first++)
     {
-        std::pair<Vertex_ID, Vertex_ID> uv = g.verts_on_edge(edges.first);
+        std::pair<Vertex_ID, Vertex_ID> uv = g.verts_on_edge(*edges.first);
         v.push_back(gl_obj::pos_vec(g[uv.first].x, g[uv.first].y));
         v.push_back(gl_obj::pos_vec(g[uv.second].x, g[uv.second].y));
     }
@@ -303,7 +303,8 @@ int main(int argc, char **argv) {
     g.read_from_file(dot_file.c_str());
 
     Graph copy(g);
-    g.write_to_file("test_files/out.dot");
+    copy.write_to_file("test_files/out.dot");
+
 
     g.set_layout();
 
@@ -327,13 +328,27 @@ int main(int argc, char **argv) {
 
 
 
-    Pebbled_Graph pg(&g);
 
     if (runtime_option == 0) {
-        pg.component_pebble_game_2D();
-    } else if (runtime_option == 1) {
+
+        Subgraph sg(&g);
+        std::pair<Vertex_Iterator, Vertex_Iterator> vs = g.vertices();
+        sg.induce(vs.first, vs.second);
+
+        // sg.remove_vertex(*(g.find_vertex("7")));
+        // sg.remove_vertex(*(g.find_vertex("8")));
+        // sg.remove_vertex(*(g.find_vertex("9")));
+        // sg.remove_vertex(*(g.find_vertex("10")));
+
+
+        Pebbled_Graph pg(&sg);
+
         pg.DRP_2D();
+    } else if (runtime_option == 1) {
+        Pebbled_Graph pg(&g);
+        pg.component_pebble_game_2D();
     } else if (runtime_option == 2) {
+        Pebbled_Graph pg(&g);
         int pebbles_remaining = pg.pebble_game_2D();
         cout << "Pebbles remaining: " << pebbles_remaining << endl;
     }

### 2014-12-10 src/Graph/Graph.cpp
Now finds the entire DR-plan, optimal or complete.

@@ -117,7 +117,7 @@ std::pair<Edge_Iterator, Edge_Iterator> Graph::edges() const {
     return boost::edges(*this);
 }
 
-Vertex_Iterator Graph::find_vertex(const char *name) {
+Vertex_Iterator Graph::find_vertex(const char *name) const {
     Vertex_Iterator v, v_end;
     for (boost::tie(v, v_end) = this->vertices(); v != v_end; v++) {
         if ((*this)[*v].name.compare(name) == 0) {
@@ -129,15 +129,15 @@ Vertex_Iterator Graph::find_vertex(const char *name) {
 }
 
 
-unsigned int Graph::num_vertices() {
+unsigned int Graph::num_vertices() const {
     return boost::num_vertices(*this);
 }
 
-unsigned int Graph::num_edges() {
+unsigned int Graph::num_edges() const {
     return boost::num_edges(*this);
 }
 
-std::pair<Vertex_ID, Vertex_ID> Graph::verts_on_edge(Edge_ID e) {
+std::pair<Vertex_ID, Vertex_ID> Graph::verts_on_edge(Edge_ID e) const {
     return std::make_pair(boost::source(e, *this), boost::target(e, *this));
 }
 
@@ -455,7 +455,7 @@ void Graph::read_from_file(const char* filename) {
 
 
 
-Subgraph::Subgraph(Graph *g) {
+Subgraph::Subgraph(const Graph *g) {
     _graph = g;
 }
 
@@ -517,8 +517,10 @@ void Subgraph::add_vertex(Vertex_ID vertex) {
     for (boost::tie(v, v_end) = boost::adjacent_vertices(vertex, *_graph);
         v != v_end; v++)
     {
-        Edge_ID e = boost::edge(vertex, *v, *_graph).first;
-        _edges.insert(e);
+        if (_vertices.find(*v) != _vertices.end()) {
+            Edge_ID e = boost::edge(vertex, *v, *_graph).first;
+            _edges.insert(e);
+        }
     }
 }
 
@@ -576,6 +578,6 @@ unsigned int Subgraph::num_edges() {
 }
 
 
-Vertex_Properties& Subgraph::operator[](Vertex_ID vertex) {
+const Vertex_Properties& Subgraph::operator[](Vertex_ID vertex) {
     return (*_graph)[vertex];
 }

### 2014-12-10 src/Graph/Graph.hpp

@@ -91,12 +91,12 @@ public:
     std::pair<Vertex_Iterator, Vertex_Iterator> vertices() const;
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
-    Vertex_Iterator find_vertex(const char *name);
+    Vertex_Iterator find_vertex(const char *name) const;
 
-    unsigned int num_vertices();
-    unsigned int num_edges();
+    unsigned int num_vertices() const;
+    unsigned int num_edges() const;
 
-    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_ID e);
+    std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_ID e) const;
 
     void set_layout();
 
@@ -116,7 +116,7 @@ typedef std::set<Edge_ID>::iterator Sg_Edge_Iterator;
 // An induced subgraph of the input graph. Initially empty.
 class Subgraph {
 public:
-    Subgraph(Graph *g);
+    Subgraph(const Graph *g);
 
     void induce(Vertex_Iterator begin, Vertex_Iterator end);
     void induce(std::set<Vertex_ID> *vertices);
@@ -133,9 +133,9 @@ public:
     unsigned int num_vertices();
     unsigned int num_edges();
 
-    Vertex_Properties& operator[](Vertex_ID vertex);
+    const Vertex_Properties& operator[](Vertex_ID vertex);
 private:
-    Graph *_graph;
+    const Graph *_graph;
 
     std::set<Vertex_ID> _vertices;
     std::set<Edge_ID> _edges;

### 2014-12-10 src/Graph/Pebbled_Graph.cpp

@@ -10,6 +10,23 @@
 
 
 
+void Cluster::print_tree(Graph &g, unsigned int tabs) {
+    // std::string t = "";
+    // for (int i = 0; i < tabs; i++) t.append("\t");
+
+
+    for (int i = 0; i < tabs; i++) std::cout << "\t";
+    for (std::set<Vertex_ID>::iterator v_it = vertices.begin(); v_it != vertices.end(); v_it++) {
+        std::cout << g[*v_it].name << " ";
+    }
+    std::cout << std::endl;
+
+    for (std::set<Cluster*>::iterator c_it = children.begin(); c_it != children.end(); c_it++) {
+        (*c_it)->print_tree(g, tabs+1);
+    }
+}
+
+
 
 
 
@@ -236,12 +253,14 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D(Vertex_ID *exclude) {
 
 
 
-std::set<Cluster*> Pebbled_Graph::DRP_2D() {
+std::set<Cluster*> Pebbled_Graph::DRP_2D(bool optimalDRP) {
     std::set<Cluster*> all_clusts;
 
     // the system is not wellconstrained
-    if (pebble_game_2D() != this->l)
+    if (pebble_game_2D() != this->l) {
+        std::cout << "Not wellconstrained, " << _total_pebbles_available << ": "; print_all_verts_with_pebbles();
         return all_clusts;
+    }
 
     std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> vs = this->in_subgraph->vertices();
     for (Sg_Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++)
@@ -300,15 +319,43 @@ std::set<Cluster*> Pebbled_Graph::DRP_2D() {
 
     }
 
-    // this->pebble_game_2D();
+    // If looking for optimalDRP, check if intersection is wellconstrained
+    // if it is, only save two clusters as children.
+    if (optimalDRP) {
+        if (all_clusts.size() != 0) {
+            std::set<Cluster*>::iterator c_it1, c_it2;
+            c_it1 = all_clusts.begin();
+            c_it2 = all_clusts.begin();
+            c_it2++;
+                    std::cout << *c_it2 << " " << *c_it1 << std::endl;
 
-    std::cout << "\n" << std::endl;
-    for (std::set<Cluster*>::iterator c_it = all_clusts.begin(); c_it != all_clusts.end(); c_it++) {
-        std::cout << "Cluster: ";
[…]

### 2014-12-10 src/Graph/Pebbled_Graph.hpp

@@ -11,12 +11,16 @@
 
 struct Cluster {
     std::set<Vertex_ID> vertices;
-    std::set<Edge_ID> edges;
+    std::set<Cluster*> children;
+
+    // std::set<Edge_ID> edges;
 
     void add_vertex(Vertex_ID v) {vertices.insert(v);}
-    void add_edge(Edge_ID e) {edges.insert(e);}
+    // void add_edge(Edge_ID e) {edges.insert(e);}
 
-    unsigned int num_edges() {return edges.size();}
+    // unsigned int num_edges() {return edges.size();}
+
+    void print_tree(Graph &g, unsigned int tabs = 0);
 };
 
 
@@ -30,7 +34,8 @@ public:
     // if there is exactly 1 cluster, the graph is rigid
     std::set<Cluster*> component_pebble_game_2D(Vertex_ID *exclude = NULL);
 
-    std::set<Cluster*> DRP_2D();
+    // If optimal is false, you find completeDRP
+    std::set<Cluster*> DRP_2D(bool optimalDRP = true);
 
 
 private:

### 2014-12-10 src/main.cpp

@@ -343,7 +343,13 @@ int main(int argc, char **argv) {
 
         Pebbled_Graph pg(&sg);
 
-        pg.DRP_2D();
+        Cluster c;
+        std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> sgvs = sg.vertices();
+        for (Sg_Vertex_Iterator v_it = sgvs.first; v_it != sgvs.second; v_it++)
+            c.add_vertex(*v_it);
+
+        c.children = pg.DRP_2D();
+        c.print_tree(g);
     } else if (runtime_option == 1) {
         Pebbled_Graph pg(&g);
         pg.component_pebble_game_2D();

### 2014-12-12 src/Graph/Graph.cpp
Changed DRP finding a little. Still doesn't use pointers correctly. Encased in auxilliary function now

@@ -561,6 +561,21 @@ std::pair<Sg_Edge_Iterator, Sg_Edge_Iterator> Subgraph::edges() {
     return std::pair<Sg_Edge_Iterator, Sg_Edge_Iterator>(_edges.begin(), _edges.end());
 }
 
+std::set<Vertex_ID> Subgraph::vertices_adjacent(std::set<Vertex_ID> &v_set) {
+    std::set<Vertex_ID> ret;
+    for (std::set<Vertex_ID>::iterator set_v = v_set.begin(); set_v != v_set.end(); set_v++) {
+        boost::graph_traits<Graph>::adjacency_iterator v, v_end;
+        for (boost::tie(v, v_end) = boost::adjacent_vertices(*set_v, *_graph);
+            v != v_end; v++)
+        {
+            ret.insert(*v);
+        }
+    }
+
+    return ret;
+}
+
+
 // const std::set<Vertex_ID>* Subgraph::vertices() {return &_vertices; }
 // const std::set<Edge_ID>* Subgraph::edges() {return &_edges; }
 

### 2014-12-12 src/Graph/Pebbled_Graph.cpp

@@ -11,11 +11,8 @@
 
 
 void Cluster::print_tree(Graph &g, unsigned int tabs) {
-    // std::string t = "";
-    // for (int i = 0; i < tabs; i++) t.append("\t");
-
-
     for (int i = 0; i < tabs; i++) std::cout << "\t";
+    std::cout << this << " " << finished << ": ";
     for (std::set<Vertex_ID>::iterator v_it = vertices.begin(); v_it != vertices.end(); v_it++) {
         std::cout << g[*v_it].name << " ";
     }
@@ -251,116 +248,350 @@ std::set<Cluster*> Pebbled_Graph::component_pebble_game_2D(Vertex_ID *exclude) {
 }
 
 
+Cluster Pebbled_Graph::DRP_2D(bool optimalDRP) {
+    Cluster ret;
+    std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> sgvs = in_subgraph->vertices();
+    for (Sg_Vertex_Iterator v_it = sgvs.first; v_it != sgvs.second; v_it++)
+            ret.add_vertex(*v_it);
+
+    ret.children = _DRP_2D_aux(NULL, optimalDRP);
+    ret.finished = true;
+
+    return ret;
+}
 
 
-std::set<Cluster*> Pebbled_Graph::DRP_2D(bool optimalDRP) {
-    std::set<Cluster*> all_clusts;
+
+
+
+
+// TODO: this is quite inefficient, the paper says there is a better way
+void erase_and_delete_all_subseteqs(Cluster *super, std::set<Cluster*> &sub) {
+    std::vector<std::set<Cluster*>::iterator> to_erase;
+
+    for (std::set<Cluster*>::iterator sub_it = sub.begin(); sub_it != sub.end(); sub_it++) {
+        bool is_subset = std::includes(
+            super->vertices.begin(), super->vertices.end(),
+            (*sub_it)->vertices.begin(), (*sub_it)->vertices.end());
+        if (is_subset) {
+            to_erase.push_back(sub_it);
+        }
+    }
+    for (int i = to_erase.size()-1; i >= 0; i--) {
+        // std::cout << "Delete" <<std::endl;
+        delete *(to_erase[i]);
+        // std::cout << "Delete - done" <<std::endl;
+
+        sub.erase(to_erase[i]);
+    }
+}
+
+// TODO: this is quite inefficient, the paper says there is a better way
[…]

### 2014-12-12 src/Graph/Pebbled_Graph.hpp

@@ -12,13 +12,19 @@
 struct Cluster {
     std::set<Vertex_ID> vertices;
     std::set<Cluster*> children;
-
     // std::set<Edge_ID> edges;
 
-    void add_vertex(Vertex_ID v) {vertices.insert(v);}
-    // void add_edge(Edge_ID e) {edges.insert(e);}
+    bool finished;
 
-    // unsigned int num_edges() {return edges.size();}
+
+    Cluster() {finished = false;}
+
+
+    void add_vertex(Vertex_ID v) {vertices.insert(v);}
+
+    // void finished(bool f) {_finished = f}
+    // bool finished() {return finished;}
+    // void add_edge(Edge_ID e) {edges.insert(e);}
 
     void print_tree(Graph &g, unsigned int tabs = 0);
 };
@@ -35,7 +41,7 @@ public:
     std::set<Cluster*> component_pebble_game_2D(Vertex_ID *exclude = NULL);
 
     // If optimal is false, you find completeDRP
-    std::set<Cluster*> DRP_2D(bool optimalDRP = true);
+    Cluster DRP_2D(bool optimalDRP = true);
 
 
 private:
@@ -76,6 +82,16 @@ private:
         vert_peb() {}
         vert_peb(Vertex_ID v, unsigned int p) {vertex = v; pebble = p;}
     };
+
+
+    std::set<Cluster*> _DRP_2D_aux(
+        Cluster* known_cluster,
+        bool optimalDRP);
+
+    std::set<Cluster*> _DRP_2D_linear_aux(
+        Cluster* known_cluster,
+        bool optimalDRP);
+
 
 
 

### 2014-12-12 src/main.cpp

@@ -112,45 +112,6 @@ void do_font_stuff() {
 
 
 
-Graph set_up_graph() {
-    Graph g;
-
-    vector<Vertex_ID> v;
-    for (int i = 0; i < 10; i++) {
-        v.push_back(g.add_vertex());
-    }
-
-    g.add_edge(v[0], v[4], 1);
-    g.add_edge(v[1], v[4], 1);
-    g.add_edge(v[2], v[4], 1);
-    g.add_edge(v[3], v[4], 1);
-    g.add_edge(v[0], v[1], 1);
-    g.add_edge(v[1], v[3], 1);
-    g.add_edge(v[2], v[3], 1);
-    g.add_edge(v[4], v[5], 1);
-    g.add_edge(v[5], v[6], 1);
-    g.add_edge(v[6], v[7], 1);
-    g.add_edge(v[7], v[8], 1);
-    g.add_edge(v[8], v[9], 1);
-    g.add_edge(v[9], v[3], 1);
-    g.add_edge(v[6], v[2], 1);
-    g.add_edge(v[5], v[8], 1);
-
-
-    // // Perform a topological sort.
-    // std::deque<int> topo_order;
-    // boost::topological_sort(g, std::front_inserter(topo_order));
-
-    // // Print the results.
-    // for(std::deque<int>::const_iterator i = topo_order.begin();
-    //     i != topo_order.end();
-    //     ++i)
-    // {
-    //     cout << tasks[v] << endl;
-    // }
-
-    return g;
-}
 
 
 
@@ -343,12 +304,7 @@ int main(int argc, char **argv) {
 
         Pebbled_Graph pg(&sg);
 
-        Cluster c;
-        std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> sgvs = sg.vertices();
-        for (Sg_Vertex_Iterator v_it = sgvs.first; v_it != sgvs.second; v_it++)
-            c.add_vertex(*v_it);
-
-        c.children = pg.DRP_2D();
+        Cluster c = pg.DRP_2D();
         c.print_tree(g);
     } else if (runtime_option == 1) {
         Pebbled_Graph pg(&g);

### 2014-12-12 doc/GMP2015/s01_introduction.tex

@@
 
-\begin{itemize}
-
+\begin{enumerate}
+  %  
     \item (cross-section) of microtubules \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
-
+   % 
     \item (cross-section) of hierarchical bone structure \cite{XX}
-
+    %
     \item interlinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
-
+  %  
     \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced 
@@
     \cite{Microfibrils2} within reach.
-
+    %
     \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 
@@
     disks on the plane \cite{JammedDisk1}
-
-\end{itemize}
-
+%
+\end{enumerate}
+%
 In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
@@
 to understand or design material layers (their solutions) with desired properties.
-
+%
 \subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
@@
 
-\medskip\noindent\underbar{{\sl (1) Finding (vertex)-maximal generically rigid subsystems}}
+\medskip\noindent\underbar{{\sl (i) Finding (vertex)-maximal generically rigid subsystems}}
 Fast, graph-based algorithms exist (pebble-game \cite{XX}),   
@@
 understand or design properties at all scales, such as: rigidity, flexes, distribution 
-of external stresses,  well as robustness/sensitivity to constraint variations.
+of external stresses, boundary conditions for isostaticity, as well as behavior under constraint variations.
 
-\medskip\noindent\underbar{{\sl (2) Optimal Recursive Decomposition (DR-planning)}}
+\medskip\noindent\underbar{{\sl (ii) Optimal Recursive Decomposition (DR-planning)}}
 Recursive decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
@@
 
-\medskip\noindent\underbar{{\sl (3) DR-plans for special classes and with other criteria}}
+\medskip\noindent\underbar{{\sl (iii) DR-plans for special classes and with other criteria}}
 For a special class of 2D qusecs, namely  {\em tree-decomposable} systems \cite{XX} common in computer aided mechanical design, 
@@
 
-\medskip\noindent\underbar{{\sl (4) Optimal Recombination and Solution Space Navigation}}
+\medskip\noindent\underbar{{\sl (iv) Optimal Recombination and Solution Space Navigation}}
 For fairly general 3D constraint systems, there are algorithms with formal guarantees that 
@@
 
-\medskip\noindent\underbar{{\sl (5) Configuration Spaces of Underconstrained Systems}}
+\medskip\noindent\underbar{{\sl (v) Configuration Spaces of Underconstrained Systems}}
 For underconstrained 2D bar-joint and multibody-pin cusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed \cite{XX, XX} that 
@@
 While software suites exist \cite{XX,XX}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
-
+%
 \subsection{Contributions}
 \label{sec:cont}
-In this paper, we  navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a 
+The contributions of this paper are the following.
+\begin{itemize}
+\item
+We  navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a 
 so-called {\it canonical} DR-plan and showing a strong Church-Rosser property: {\it all canonical DR-plans for isostatic or underconstrained (independent) 
 2D qusecs are optimal}. 
-We then give an efficient ($O(n^2)$) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
-canonical DR-plan 
-illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained  and gives classes of 
-overconstrained 2DGMCS for which the above  algorithm finds an optimal DR-plan. We then show how to judiciously combine 
-existing methods for recombination, search of so-called, convex Cayley configuration spaces, and navigation of realizations, 
-to explore a connected component of the configuration space of a flexible 2DGMCS, given a starting configuration. 
-Software implementation can be found at: \todo{Provide Link}.
+\item
+We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
+The canonical DR-plan  
+illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained systems, and characterizes those 
+overconstrained 2D qusecs for which the above algorithm does find an optimal DR-plan. 
+\item
+We show how to judiciously combine 
+existing methods for recombination with searching  convex or low complexity Cayley configuration spaces to  navigate the solution space of isostatic 2D qusecs, and
+explore connected components of the configuration space of underconstrained 2D qusecs. 
+\item
+We briefly describe applications of the above techniques to modeling and analyzing specific 2D material layers mentioned above.
+\item
+Software implementation and videos can be found at: \todo{Provide Link}.
+\end{itemize}
 
-The starting point of this paper is the observation that the above 2D Geometric Materials Constraint Systems (resp. 2D-GEMS) 
-belong to a
-
-However, these systems do not 
-belong in the largest combinatorial class of 2D systems, whose configuration spaces have formally proven  characterizations and 
-algorithmic guarantees (e.g. for algebraic complexity, path and volume measures, and  topology).
-
-Specifically, these results \todo{[Citation Needed]} rely on the fact that rigid systems in this  class, in contrast to GMCS, 
-have a straightforward and immediate {\sl optimal} recursive decomposition, (so-called {\sl optimal} DR-plan) and  recombination 
-or realization (with a formal disambiguation by minimal orientation specification).
-
-It is known \todo{[Citation Needed]} that already the optimal DR-planning problem for general 2DGMCS is NP-hard, intuitively 
-because of a combinatorial explosion of polynomially-sized DR-plans.
-
-\subsection{Previous work}
+%%%%%%%
+XXXXXXXXXXXXXXXXXXXXXXx
 

### 2014-12-12 script/p2dot.py
Uploaded the script generating dot file from subdivision tiling file.

@@ -1,0 +1,72 @@
+#!/usr/bin/python
+
+#Convert a .p packing file to a .dot graph file 
+
+#currently start vertex index given by command line arguments
+
+import sys
+import math
+
+narg = len(sys.argv)
+if narg < 3:
+    print 'Usage: p2dot .p_filename start_vertex_idx'
+    sys.exit(1)
+p_filename = sys.argv[1]
+start_idx = int(sys.argv[2])
+
+f = open(p_filename, 'r')
+line = f.readline()
+end_idx = int(line.split()[1])
+
+# return whether the given idx is actually a vertex in tiling (not a barycenter)
+def is_vertex(idx): return idx >= start_idx and idx <= end_idx
+
+coordinate_dict = dict() # coordinates of vertices
+edge_dict = dict() # edge -- length
+
+# read combinatorics
+while not line.startswith('FLOWERS:'):  line = f.readline()
+line = f.readline()
+while not line.strip()=='':
+    a = line.split() # a[0]: i, a[1]: degree, a[2]--a[d+2]: neighbors
+    i=int(a[0])
+    if is_vertex(i): 
+        for j in range(2, len(a)):
+            nidx = int(a[j])
+            if nidx > i and  is_vertex(nidx): 
+                edge_dict[(i, nidx)] = 0
+    line = f.readline()
+
+
+# read coordinates of vertices
+while not line.startswith('CENTERS:'):  line = f.readline()
+s = " ".join(line.strip() for line in f) # assume center is the last piece of info ... 
+coords = s.split()
+for i in range(start_idx, end_idx+1):
+    xi = 2*(i-1); yi = xi + 1
+    x = float(coords[xi]); y = float(coords[yi])
+    coordinate_dict[i] = (x,y)
+
+# compute edge lengths
+for (i,j) in edge_dict.keys():
+    pi = coordinate_dict[i]; pj = coordinate_dict[j]
+    d = math.sqrt( (pi[0] - pj[0])**2 + (pi[1] - pj[1])**2 )
+    edge_dict[(i,j)] = d
+
+print '# vertices: %d, # edges: %d' % (len(coordinate_dict), len(edge_dict))
+
+# write .dot file
+fout = open(p_filename+".dot", 'w')
+fout.write('graph G {\n');
+# write vertices
+for i in range(start_idx, end_idx+1): 
+    fout.write(str(i) + ' ')
[…]

### 2014-12-12 script/subdivision programs by William Floyd/subdivide.c

@@ -1,0 +1,1123 @@
+/* subdivide.c */
+/* Description:
+ * Reads a subdivision rule from a file (e.g., filename.r).
+ * Reads a tiling of a quadrilateral from a file (e.g., filename.l0).
+ * Subdivides the tiling according to the rules file and
+ * writes the new tiling to a file (e.g., filename.l1).
+ *
+ * Copyright (c) 1997, 2000, 2008 James W. Cannon, William J. Floyd, and
+ * Walter R. Parry
+ *
+ * This work was supported in part by National Science Foundation
+ * grants DMS-9803868 and DMS-9971783.
+ * 
+ * This is free software and may be freely copied, modified, and
+ * redistributed as noted below.
+ *	The copyright notice must remain intact.
+ *	Modifications must include a notice giving the reason
+ * 	for the modification and including name and date.
+ *	This is provided "as is" and comes with no warranty or
+ * 	guarantee of fitness. Comments, bugs, and suggestions may be sent
+ * 	to floyd@math.vt.edu.
+ *
+ */
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#define MAXVALENCE 50
+        /* The assumed maximal number of adjacencies to a tile
+         * that is not an edge tile. */
+        /* For the edge tiles we assume no more than nooftiles
+         * as the maximum number of adjacencies.*/
+#define MAXCYCLE 50
+        /* The assumed maximal length in a vertex cycle. */
+
+/* Variables for the old tiling.*/
+int noofedges; /* Acts as size of *edges. */
+int nooftiles; /* Acts as size of each of next three arrays.*/
+int *facetype; /* Array, dynamically allocated: gives type of each tile.*/
+int *facesize; /* Array: gives no of edges of each tile.*/
+int *facebegin;/* Array: gives first edgeid associated with each tile.*/
+int *edges;    /* Array: gives tileid adjacent across each edge of each tile.*/
+               /* Since each tile has several edges, there are more edges than
+                * tiles.*/
+
+/* Variables for the new tiling. */ /* See comments for previous section.*/
+int newnooftiles;
+int *newfacetype;
+int *newfacesize;
+int *newfacebegin;
+int *newedges;
+
+/* Variables for the relationship between old and new tiling. */
[…]

### 2014-12-12 script/examples/dodec.r
Added several examples (with pictures showing the realization of subdivision)

@@ -1,0 +1,162 @@
+   Number_of_tile-types_(not_counting_the_four_standard_end_types):
+   3
+   Size_(i.e.,_number_of_edges)_of_each_of_these_types:
+   3 4 5
+   Subdivision-tiling:
+   Type_number_(the_first_one_should_be_4_since_the_ends_are_typed_0_1_2_3):
+   4
+   Number_of_tiles_into_which_that_tile_type_is_subdivided:
+   37
+   Type_of_each_of_these_tiles_in_the_subdivision:
+   4 6 5 6 5 6 5 4 5 4
+   5 4 5 4 5 4 5 6 5 6
+   5 6 5 6 5 4 5 4 5 6
+   5 4 5 4 5 5 6
+   Tile-ids_--_Adjacent_tiles_(in_correct_order_according_to_edge_number):
+   0 20 28 30
+   1 -1 -1 6 8 2
+   2 1 9 3 -1
+   3 -1 2 10 12 4
+   4 3 13 5 -1
+   5 -1 4 14 16 -1
+   6 1 -1 17 7
+   7 6 18 8
+   8 1 7 19 9
+   9 2 8 10
+   10 3 9 19 11
+   11 10 20 12
+   12 3 11 21 13
+   13 4 12 14
+   14 5 13 21 15
+   15 16 14 22
+   16 5 15 23 -1
+   17 -1 24 26 18 6
+   18 17 27 19 7
+   19 8 18 28 20 10
+   20 19 0 21 11
+   21 12 20 30 22 14
+   22 21 31 23 15
+   23 -1 16 22 32 34
+   24 17 -1 36 25
+   25 26 24 35
+   26 17 25 29 27
+   27 18 26 28
+   28 19 27 29 0
+   29 28 26 35 32 30
+   30 21 0 29 31
+   31 30 32 22
+   32 23 31 29 33
+   33 32 35 34
+   34 23 33 36 -1
+   35 29 25 36 33
+   36 -1 -1 34 35 24
+   Corresponding_boundary-tiling:
+   Boundary_of_type_number_(the_first_should_be_4):
+   4
+   Number_of_edges_before_subdivision:
+   3
+   Number_of_edges_into_which_each_of_these_original_edges_is_subdivided:
+   5 5 5
+   Original_edgeid_-_numbers_in_pairs_for_subdivision_edges_(tileno_edgeno):
[…]

### 2014-12-12 script/examples/pent.r

@@ -1,0 +1,31 @@
+   Number_of_tile-types_(not_counting_the_four_standard_end_types):
+   1
+   Size_(i.e.,_number_of_edges)_of_each_of_these_types:
+   5
+   Subdivision-tiling:
+   Type_number_(the_first_one_should_be_4_since_the_ends_are_typed_0_1_2_3):
+   4
+   Number_of_tiles_into_which_that_tile_type_is_subdivided:
+   6
+   Type_of_each_of_these_tiles_in_the_subdivision:
+   4 4 4 4 4 4
+   Tile-ids_--_Adjacent_tiles_(in_correct_order_according_to_edge_number):
+   0 -1 1 5 4 -1
+   1  -1 -1 2 5 0
+   2 1 -1 -1 3 5
+   3 5 2 -1 -1 4
+   4 0 5 3 -1 -1
+   5 0 1 2 3 4
+   Corresponding_boundary-tiling:
+   Boundary_of_type_number_(the_first_should_be_4):
+   4
+   Number_of_edges_before_subdivision:
+   5
+   Number_of_edges_into_which_each_of_these_original_edges_is_subdivided:
+   2 2 2 2 2
+   Original_edgeid_-_numbers_in_pairs_for_subdivision_edges_(tileno_edgeno):
+   0  1 0 0 0
+   1  2 1 1 1
+   2  3 2 2 2
+   3  4 3 3 3
+   4  0 4 4 4 

### 2014-12-14 doc/GMP2015/s00_abstract.tex
Fixed subdivision script and used it as -opt=2 for main code. Added Meera's updated intro.

@@
 \begin{abstract}
-Super awesome abstract.
+A variety of quasi-uniform (aperiodic) and possibly self-similar, layered material structures are natural consequences of the design objectives of rigidity and flexibility, minimizing mass, optimally distributing external stresses, and participating in assembly of diverse and multifunctional, larger structures.
+
+Crucial for design and analysis of these properties is the recursive decomposition of diverse types of underlying isostatic or underconstrained 2D geometric constraint systems (which we collectively call \textit{qusecs}).
+While the optimal recursive decomposition (\textit{DR-planning}) problem is NP-hard even for 2D bar-joint constraint systems, we describe an $O(n^2)$ algorithm  for qusecs.
+Besides elucidating the source of NP-hardness, the algorithm outputs \textit{canonical} DR-plans that meet previously studied desirable criteria.
+When a DR-plan is input, we show reductions between problems related to realizing qusecs,  leading to efficient algorithms that leverage recent results on Cayley configuration spaces.
+We briefly discuss two specific 2D layered materials as applications of the above algorithms, whose software implementation is available.
+
+Keywords: rigidity, geometric constraint solving, configuration spaces, self-similar structures, layered materials
+
+% =====================CUT HERE ==============
+
+% Examples include:
+% cellulose and collagen microfibrils in cell walls;
+% actin cytoskeleton matrix; hierarchical bone cross-section;
+% ciliary membranes and transitions;
+% self-similar DNA self-assemblies; silica bilayers;
+% disordered graphene; etc.
+
+% By  defining a canonical recursive decomposition
+% ({\it DR plan}),  we show that the optimal DR-planning
+% problem for
+
+% We  navigate the NP-hardness of finding
+% optimal DR-plans by defining a
+% so-called {\it canonical} DR-plan and showing a strong Church-Rosser
+% property: {\it all canonical DR-plans for isostatic or underconstrained
+% 2D qusecs are optimal}.
+% \item
+% We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence
+% optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
+% The canonical DR-plan
+% illuminates the essence of the NP-hardness of finding optimal DR-plans for
+% overconstrained systems, and characterizes those
+% overconstrained 2D qusecs for which the above algorithm does find an
+% optimal DR-plan.
+% \item
+% We show how to judiciously combine
+% existing methods for recombination with searching  convex or low
+% complexity Cayley configuration spaces to  navigate the solution space of
+% isostatic 2D qusecs, and
+% explore connected components of the configuration space of
+% underconstrained 2D qusecs.
+% \item
+
+% \item
+% Software implementation and videos can be found at: \todo{Provide Link}.
 \end{abstract}

### 2014-12-14 doc/GMP2015/s01_introduction.tex
Fixed subdivision script and used it as -opt=2 for main code. Added Meera's updated intro.

@@
 \label{sec:prev}
-Modeling, analysis and design of quasi-uniform or self-similar layered materials using geometric constraint systems
 
-Some properties of natural and engineered materials can be studied by treating them as stacked two dimensional (2D) layers, 
-where the structure within each layer may be recursively 
-self-similar \cite{Intro1}, 
-spanning multiple scales. Within any one scale, the material layer is generally aperiodic, and quasi-uniform,
-consisting of a few repeated motifs appearing in disordered arrangments.
-Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
-Furthermore, a layer is often  either {\sl isostatic or flexible, not self-stressed}, at any scale. These properties are
-consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself 
+Some properties of natural and engineered materials can be analyzed  by
+treating them as stacked two dimensional (2D) layers.
+The structure within each layer could be
+self-similar \cite{Intro1}
+spanning multiple scales,
+generally aperiodic and quasi-uniform within any one scale,
+and consisting of a few repeated motifs appearing in disordered
+arrangements. Each layer is not necessarily planar, i.e.,
+it consists of multiple,
+inter-constraining planar (genus 0) monolayers.
+Furthermore, a layer is often  either
+{\sl isostatic or underconstrained, not
+self-stressed} (See Section \ref{prelim} for definitions).
+These properties are
+consistent with a self-assembled 2D structure that minimizes mass,
+optimally distributes external stresses and itself
 participates in the assembly of diverse and multifunctional,
@@
 
-Examples include:
+Examples of such materials (See Figure \ref{examples}) include:
 
@@
   %
-    \item (cross-section) of microtubules \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
+    \item (cross-section) of microtubules \cite{Necklace1} e.g., in
+ciliary membranes and transitions \cite{Necklace2}
    %
-    \item (cross-section) of hierarchical bone structure \cite{XX}
+    \item (cross-section) of hierarchical bone structure \cite{XX} %
+    \item crosslinked cellulose or collagen microfibril monolayers e.g.,
+in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked
+actin filaments in the cytoskeleton matrix.
   %
-    \item interlinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}
-  %  
-    \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced 
-    by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other structures 
+    \item more recent, engineered examples including  disordered graphene
+layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced
+    by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a
+recent Sierpinski carpet, bringing other self-similar structures
     \cite{Microfibrils2} within reach.
     %
-    \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 
+    \item  silica bi-layers, glass \cite{SilicaGlass1}
+\cite{SilicaGlass2}, and materials that behave like assemblies of
     2D particles under non-overlap constraints, e.g, jammed
@@
 %
-In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
-within the material layer, it is natural to model a material layer as a solution of geometric constraint systems 
-arising from appropriate geometric primitives under metric or algebraic constraints. 
-Such  2D {\bf{\em qusecs}} (quasi-uniform or self-similar constraint systems) can be used to 
-to understand or design material layers (their solutions) with desired properties.
+In order to study structural and mechanical properties
+such as rigidity, flexes and distribution of external stresses
+within the material layer, it is natural to model a material
+layer as a solution of a geometric constraint system
+of appropriate types of geometric primitives, under
+metric or algebraic constraints.
+Such  2D {\bf{\em qusecs}} (quasi-uniform or self-similar
+constraint systems) can be used to
+understand or design material layers (their solutions)
+with desired properties.
 %
-\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
-We now briefly survey existing techniques for studying 2D {\em qusecs,}  many of which are
-{\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3). 
-The limitations of these techniques directly motivate the contributions of this paper.
+\subsection{Previous Work on Relevant 2D Geometric Constraint Systems} We
+now briefly survey existing techniques for studying 2D {\em qusecs,}  many
+of which are {\it bar-joint} systems (Examples 1,2 above, see Sections
+\ref{dr}, \ref{recomb}),
+{\it multibody-pin} systems (Example 4,5, see Section \ref{app}) or  {\it
+pinned-line incidence} systems (Example 3, see Section \ref{app}). The
+limitations of these techniques directly motivate
+the contributions of this paper.
 
-\medskip\noindent\underbar{{\sl (i) Finding (vertex)-maximal generically rigid subsystems}}
+\medskip\noindent\underbar{{\sl (i) Finding (vertex)-maximal
+generically rigid subsystems}}
 Fast, graph-based algorithms exist (pebble-game \cite{XX}),
 for locating all maximal, generically rigid subsystems \cite{XX}.
-When the input is rigid, these algorithms do nothing, i.e, the output is the same as the input.
-However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems 
-into their rigid subsystems, down to the level of geometric primitives, in order to 
-understand or design properties at all scales, such as: rigidity, flexes, distribution 
-of external stresses, boundary conditions for isostaticity, as well as behavior under constraint variations.
+When the input is rigid, these algorithms compute the identity and  output
+the input.
 
-\medskip\noindent\underbar{{\sl (ii) Optimal Recursive Decomposition (DR-planning)}}
-Recursive decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
+However, both for self-similar or just aperiodic 2D qusecs, it is
+imperative to recursively decompose rigid systems
+into their rigid subsystems, down to the level of
+geometric primitives, in order to
+understand or design properties at all scales, such as:
+rigidity, flexes, distribution
+of external stresses, boundary conditions for isostaticity,
+as well as behavior under constraint variations.
+
+\medskip\noindent\underbar{{\sl (ii) Optimal Recursive Decomposition
+(DR-planning)}}
+Recursive decomposition of geometric constraint systems has been
+formalized \cite{XX} and well-studied \cite{XX}
 as the {\sl Decomposition-Recombination (DR)-planning} problem.
 For the above-mentioned classes of 2D qusecs, generic rigidity is
-a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without
+a combinatorial property and hence each level of the decomposition should,
+in principle, be achievable by a graph-based algorithm as in (1), without
 involving geometric information in the constraint system.
 Since many  such decompositions can exist for a given constraint system,
-criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{XX}. An {\em optimal DR-plan} is one
-that minimizes the maximum number of child subsystems of any parent system, a property that is 
-crucial for obtaining a solution to the parent constraint system from the solutions of the child systems.
+criteria defining desirable or optimal DR-plans and DR-planning algorithms
+were given in \cite{XX}. An {\em optimal DR-plan} is one
+that minimizes the maximum number of child subsystems of any parent
+system, a property that is
+crucial for obtaining a solution to the parent constraint system from the
+solutions of the child systems.
+
 However, for general 2D qusecs, even when restricted to bar-joint systems,
@@
 
-\medskip\noindent\underbar{{\sl (iii) DR-plans for special classes and with other criteria}}
-For a special class of 2D qusecs, namely  {\em tree-decomposable} systems \cite{XX} common in computer aided mechanical design, 
-(which includes ruler-and-compass and Henneberg-I constuctible systems), all DR-plans turn out to 
-be optimal. This satisfies the so-called {\em Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
-Subsequently, alternate criteria were suggested   
+\medskip\noindent\underbar{{\sl (iii) DR-plans for special classes and
+with other criteria}}
+For a special class of 2D qusecs, namely  {\em tree-decomposable} systems
+\cite{XX} common in computer aided mechanical design,
+(which includes ruler-and-compass and Henneberg-I constructible systems),
+all DR-plans turn out to
+be optimal. This satisfies the so-called {\em Church-Rosser} criterion,
+leading to highly efficient DR-planning algorithms.
+For general 2D qusecs, alternate criteria were suggested
 such as {\em cluster minimality}
 requiring parent systems to be composed of
-a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); 
+a minimal set of at least 2 rigid proper subsystems (i.e., no proper
+subset forms a rigid system);
 and {\em proper maximality}, requiring child subsystems
 to be maximal rigid proper subsystems of the parent system.
-While polynomial time algorithms were given \cite{XX} to generate DR-plans meeting the cluster minimality criterion,
+
+While polynomial time algorithms were given \cite{XX} to generate DR-plans
+meeting the cluster minimality criterion,
 no such algorithm is known for the latter criterion.
 
-\medskip\noindent\underbar{{\sl (iv) Optimal Recombination and Solution Space Navigation}}
-For fairly general 3D constraint systems, there are algorithms with formal guarantees that 
+\medskip\noindent\underbar{{\sl (iv) Optimal Recombination and Solution
+Space Navigation}}
+For fairly general 3D constraint systems, there are algorithms with formal
+guarantees that
 solve the parent (sub)systems that occur in the DR-plan \cite{XX, XX,XX},
 provided the DR-plan meets some of the above criteria.
-These algorithms  uncover underlying matroids to combinatorially optimize the algebraic complexity of (re)combining child system solutions, 
-and algebraically solve the optimized recombination system to give a solution to the parent system at any
-given level of the DR-plan. However, the generality of these algorithms trades-off against efficiency, and despite the optimization,
+These algorithms  uncover underlying matroids to combinatorially optimize
+the algebraic complexity of (re)combining child system solutions,
+and algebraically solve the optimized recombination system to give a
+solution to the parent system at any
+given level of the DR-plan.
+
+However, the generality of these algorithms
+trades-off against efficiency, and despite the optimization,
 the best algorithms can still take
-exponential time in the number of child subsystems in order to guarantee all solutions even for a single (sub)system in a DR-plan, 
-and are prohibitively slow in practice. For the entire DR-plan, finding all solutions is barely tractable even with fast recombination algorithms
-for the subsystems in the DR-plan. This is because even for the simplest systems, the problem is NP-hard \cite{XX} and there is a 
-combinatorial explosion of solutions \cite{XX}.  Using the DR-plan and optimal recombination as a principled basis, high performance software exists \cite{XX,XX} to 
+exponential time in the number of child subsystems in order to guarantee
+all solutions even for a single (sub)system in a DR-plan. They
+are prohibitively slow in practice. For the entire DR-plan, finding all
+solutions is barely tractable even with fast recombination algorithms for
+the subsystems in the DR-plan. This is because even for the simplest
+systems, the problem is NP-hard \cite{XX} and there is a
+combinatorial explosion of solutions \cite{XX}.  Using the DR-plan and
+optimal recombination as a principled basis, high performance software
+exists \cite{XX,XX} to
 tame combinatorial explosion via user intervention.
 
-\medskip\noindent\underbar{{\sl (v) Configuration Spaces of Underconstrained Systems}}
-For underconstrained 2D bar-joint and multibody-pin cusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed \cite{XX, XX} that 
-to complete them into isostatic systems \cite{XX, XX} and to find paths within the connected components \cite{XX,XX} of
-standard cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{XX,XX}
-to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation 
+\medskip\noindent\underbar{{\sl (v) Configuration Spaces of
+Underconstrained Systems}}
+For underconstrained 2D bar-joint and multibody-pin cusecs obtained from
+various subclasses of tree-decomposable systems, algorithms have been
+developed \cite{XX, XX} that
+to complete them into isostatic systems \cite{XX, XX} and to find paths
+within the connected components \cite{XX,XX} of
+standard cartesian configuration spaces. Most of the algorithms with
+formal guarantees leverage Cayley configuration space theory \cite{XX,XX}
+to characterize subclasses of graphs and additional constraints that
+control combinatorial explosion, and provide faithful bijective
+representation
 of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems gets bigger, starting with
-underlying partial 2-tree graphs (alternately, tree-width 2, series-parallel, $K_4$ minor avoiding),  moving on   
-to 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), 
-and further to general 1-dof tree-decomposable graphs. 
-While software suites exist \cite{XX,XX}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
+These algorithms have decreasing efficiency as the subclass of systems
+gets bigger,  with highest efficiency for
+underlying partial 2-tree graphs (alternately, tree-width 2,
+series-parallel, $K_4$ minor avoiding),   moderate efficiency for
+1 degree-of-freedom (dof) graphs with low Cayley complexity (which include
+common linkages such as the Strandbeest, Limacon and Cardioid), and
+decreased efficiency for general 1-dof tree-decomposable graphs. While
+software suites exist \cite{XX,XX}, no such formal algorithms and
+guarantees are known for non-tree-decomposable systems.
 %
-\subsection{Contributions}
+\subsection{Contributions and Organization}
 \label{sec:cont}
@@
 \item
-We  navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a 
-so-called {\it canonical} DR-plan and showing a strong Church-Rosser property: {\it all canonical DR-plans for isostatic or underconstrained (independent) 
-2D qusecs are optimal}. 
+We  navigate the NP-hardness barrier mentioned in (2) above, for finding
+optimal DR-plans by defining a
+so-called {\it canonical} DR-plan and showing a strong Church-Rosser
+property: {\it all canonical DR-plans for isostatic or
+underconstrained
+2D qusecs are optimal} (Section \ref{dr}).
 \item
-We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
+We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence
+optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections
+\ref{dr} and \ref{app}.
 The canonical DR-plan
-illuminates the essence of the NP-hardness of finding optimal DR-plans for overconstrained systems, and characterizes those 
-overconstrained 2D qusecs for which the above algorithm does find an optimal DR-plan. 
+illuminates the essence of the NP-hardness of finding optimal DR-plans for
+overconstrained systems.
 \item
-We show how to judiciously combine 
-existing methods for recombination with searching  convex or low complexity Cayley configuration spaces to  navigate the solution space of isostatic 2D qusecs, and
-explore connected components of the configuration space of underconstrained 2D qusecs. 
+We show reductions between different well-known problems related to
+realization or solution by
+recombining the realizations or solutions of subsystems (at any given level
+of the DR-plan). Example problems include optimal completion of
+underconstrained  systems  \cite{XX}, and optimal parametrization in
+isostatic systems \cite{XX}). When completions and parametrizations are
+given from certain well-known classes, we obtain new,
+efficient algorithms for realizing both isostatic and underconstrained
+qusecs by leveraging recent results about Cayley parameters in
+\cite{XX,XX}  (Section \ref{recomb}).
 \item
-We briefly describe applications of the above techniques to modeling and analyzing specific 2D material layers mentioned above.
+We briefly describe applications of the above techniques to modeling,
+analyzing and designing specific properties in 2D material layers.
+For Examples 1,2 (achieving isostaticity, distribution of stresses in
+self-similar, bar-joint systems); Example 3 (canonical and optimal
+DR-plans for pinned line incidence systems \cite{XX}) and Examples 4,5
+(boundary-conditions for achieving various desired properties
+of multi-body  pin systems).
+
 \item
@@
 \end{itemize}
-
-%%%%%%%
-XXXXXXXXXXXXXXXXXXXXXXx
-
-Computer aided mechanical design represents a mature application of geometric constraint systems,  leading to and borrowing 
-from deep and fundamental problems and techniques in distance geometry, convex analysis, combinatorics, algebra, topology 
-and complexity theory.
-
-Broadly, these problems can be classified into three types that draw upon each other, and possess mathematical, algorithmic, 
-and software implementation aspects: (i) combinatorial (generic) minimal rigidity (or isostaticity or well-constrained-ness) 
-characterization, (ii) (optimal) recursive decomposition (DR plans), recombination and realization, (iii) configuration space 
-characterization (topology, complexity, path and volume measures).
-
-In general, proofs of combinatorial (generic) minimal rigidity characterizations for geometric constraint systems rely on 
-underlying matroids \todo{[Citation Needed]} In these cases, more-or-less greedy algorithms such as simplified network-flow, 
-or so-called pebble games can be used.
-
-Nothing nontrivial known until now for optimal DR plans, although efficient algorithms are known for DR-plans satisfying 
-various other formal properties.
-
-Recombination and realization navigation of realizations (disambiguation with orientation, faster with overconstraints, mention 
-also totally different distance geometry method - low rank PSD matrix completions SDP/SVD)
-
-Gelfand/McPherson, Walker's conjecture solution, Convex Cayley, EASAL and Caymos for low Cayley complexity and bijective 
-representation using global rigidity.

### 2014-12-14 script/p2dot.py

@@ -1,6 +1,6 @@
 #!/usr/bin/python
 
-#Convert a .p packing file to a .dot graph file 
+#Convert a .p packing file to a .dot graph file
 
 #currently start vertex index given by command line arguments
 
@@ -9,7 +9,7 @@ import math
 
 narg = len(sys.argv)
 if narg < 3:
-    print 'Usage: p2dot .p_filename start_vertex_idx'
+    print('Usage: p2dot .p_filename start_vertex_idx')
     sys.exit(1)
 p_filename = sys.argv[1]
 start_idx = int(sys.argv[2])
@@ -30,17 +30,17 @@ line = f.readline()
 while not line.strip()=='':
     a = line.split() # a[0]: i, a[1]: degree, a[2]--a[d+2]: neighbors
     i=int(a[0])
-    if is_vertex(i): 
+    if is_vertex(i):
         for j in range(2, len(a)):
             nidx = int(a[j])
-            if nidx > i and  is_vertex(nidx): 
+            if nidx > i and  is_vertex(nidx):
                 edge_dict[(i, nidx)] = 0
     line = f.readline()
 
 
 # read coordinates of vertices
 while not line.startswith('CENTERS:'):  line = f.readline()
-s = " ".join(line.strip() for line in f) # assume center is the last piece of info ... 
+s = " ".join(line.strip() for line in f) # assume center is the last piece of info ...
 coords = s.split()
 for i in range(start_idx, end_idx+1):
     xi = 2*(i-1); yi = xi + 1
@@ -53,19 +53,18 @@ for (i,j) in edge_dict.keys():
     d = math.sqrt( (pi[0] - pj[0])**2 + (pi[1] - pj[1])**2 )
     edge_dict[(i,j)] = d
 
-print '# vertices: %d, # edges: %d' % (len(coordinate_dict), len(edge_dict))
+print('# vertices: %d, # edges: %d' % (len(coordinate_dict), len(edge_dict)))
 
 # write .dot file
 fout = open(p_filename+".dot", 'w')
 fout.write('graph G {\n');
 # write vertices
-for i in range(start_idx, end_idx+1): 
-    fout.write(str(i) + ' ')
-    fout.write('[x='+str(coordinate_dict[i][0])+' y='+str(coordinate_dict[i][1])+'];\n')
+for i in range(start_idx, end_idx+1):
[…]

### 2014-12-14 src/Graph/Graph.cpp

@@ -172,17 +172,50 @@ void Graph::set_layout() {
 
 
     Vertex_Iterator i, end;
-    boost::tie(i, end) = this->vertices();
+    for (boost::tie(i, end) = this->vertices(); i != end; i++) {
+        (*this)[*i].x = vp_map[*i][0];
+        (*this)[*i].y = vp_map[*i][1];
+    }
+}
+
+
+// TODO: Potentially scale edge length
+void Graph::get_graph_in_range(float x_min, float x_max, float y_min, float y_max) {
+    // Vertex_Iterator v, v_end;
+    // boost::tie(v, v_end) = vertices();
+
+    // float
+    //     min_x_seen = (*this)[*v].x,
+    //     max_x_seen = (*this)[*v].x,
+    //     min_y_seen = (*this)[*v].y,
+    //     max_y_seen = (*this)[*v].y;
+
+    // v++;
+    // for (; v != v_end; v++) {
+    //     float v_x = (*this)[*v].x, v_y = (*this)[*v].y;
+    //     if ()
+    //     add_vertex(*v_it);
+    // }
+
+    float
+        del_x = x_max - x_min,
+        del_y = y_max - y_min;
+
+
+
+
+    Vertex_Iterator v, end;
+    boost::tie(v, end) = this->vertices();
     double
-        maxx = vp_map[*i][0],
-        minx = vp_map[*i][0],
-        maxy = vp_map[*i][1],
-        miny = vp_map[*i][1];
+        maxx = (*this)[*v].x,
+        minx = (*this)[*v].x,
+        maxy = (*this)[*v].y,
+        miny = (*this)[*v].y;
     double centerx=0, centery=0;
 
-    for (; i != end; i++) {
-        double x = vp_map[*i][0];
-        double y = vp_map[*i][1];
+    for (; v != end; v++) {
+        double x = (*this)[*v].x;
+        double y = (*this)[*v].y;
 
         centerx += x; centery += y;
 
@@ -203,10 +236,18 @@ void Graph::set_layout() {
     // std::cout << maxx << minx << std::endl;
     // std::cout << maxy << miny << std::endl;
 
-    double diffx, diffy;
-    diffx = (maxx > -minx)? maxx: -minx;
-    diffy = (maxy > -miny)? maxy: -miny;
-    double diff = (diffx > diffy)? diffx: diffy;
+    double
+        diff_x = (maxx > -minx)? maxx: -minx,
+        diff_y = (maxy > -miny)? maxy: -miny,
[…]

### 2014-12-14 src/Graph/Graph.hpp

@@ -99,6 +99,7 @@ public:
     std::pair<Vertex_ID, Vertex_ID> verts_on_edge(Edge_ID e) const;
 
     void set_layout();
+    void get_graph_in_range(float x_min, float x_max, float y_min, float y_max);
 
     void write_to_file(const char* filename);
     void read_from_file(const char* filename);

### 2014-12-14 src/main.cpp

@@ -187,14 +187,18 @@ void display_graph(Window &wind, Graph &g) {
 
 
 
-
+enum runtime_options {
+    ro_drp_2d,
+    ro_display_realized_graph,
+    ro_test
+};
 
 
 
 int main(int argc, char **argv) {
     string dot_file = "test_files/test.dot";
 
-    int runtime_option = 0;
+    runtime_options runtime_option = ro_drp_2d;
 
     // handle command-line arguments
     for (int i = 1; i < argc; i++) {
@@ -230,11 +234,11 @@ int main(int argc, char **argv) {
 
         else if (!flag.compare("-opt")) {
             if (!val.compare("0")) {
-                runtime_option = 0;
+                runtime_option = ro_drp_2d;
             } else if (!val.compare("1")) {
-                runtime_option = 1;
+                runtime_option = ro_display_realized_graph;
             } else if (!val.compare("2")) {
-                runtime_option = 2;
+                runtime_option = ro_test;
             } else {
                 cerr << "Invalid value for flag " << flag << "." << endl;
                 return 0;
@@ -257,21 +261,21 @@ int main(int argc, char **argv) {
     }
 
 
-    // Graph g = set_up_graph();
-    // g.write_to_file("test.dot");
-
+    // read in
     Graph g;
     g.read_from_file(dot_file.c_str());
 
-    Graph copy(g);
-    copy.write_to_file("test_files/out.dot");
+    // // test output
+    // Graph copy(g);
+    // copy.write_to_file("test_files/out.dot");
 
-
-    g.set_layout();
-
-    // Subgraph_Type sg;
-    // sg.read_from_file(dot_file.c_str());
-
+    // get layout... x/y coords
+    if (runtime_option != ro_display_realized_graph) {
+        g.set_layout();
+        g.get_graph_in_range(-0.87, 0.87, -0.87, 0.87);
+    } else {
+        g.get_graph_in_range(-0.87, 0.87, -0.87, 0.87);
+    }
 
 
     MainGuiManager mgm;
@@ -290,29 +294,31 @@ int main(int argc, char **argv) {
 
 
 
-    if (runtime_option == 0) {
+    if (runtime_option == ro_drp_2d) {
+        cout << "Running DRP_2D on \"" << dot_file << "\"." << endl;
 
         Subgraph sg(&g);
[…]

### 2014-12-14 doc/GMP2015/s07_problempinning.tex
Added a section on problem pinning. A bit in its infancy

\section{Problem Pinning}

We can use DR plans to help in solving questions in designs for graphene and silica bilayers (\textcolor{red}{Add Citation}). Specifically, we will look into the problem when you have a monolayer and it is underconstrained. To make this thing isostatic, there are a few methods we can apply.

\begin{itemize}
    \item Pin together 2 underconstrained monolayers in such a way that the resulting bilayer becomes isostatic.
    \item Pin the boundary in such a way that the resulting system is isostatic.
    \item Apply a self-similarity constraint and essentially wrap the monolayer and pin it to itself (pin one edge to another in certain places)
\end{itemize}

In each case, we are specifically interested in how to pin these things so that the resultiing structure also has a small DR plan.

\subsection{Body-Pin}

To answer these questions, we first restrict ourselves to a specific case of Body-Pin networks such that:

\begin{itemize}
    \item Each pin is shared by at most 2 bodies
    \item No 2 bodies share more than one pin 
\end{itemize}

Such a body-pin system can also be seen as a {\em body-bar system} where the bodies are the original bodies and each pin represents 2 edges from one body to another. Then, such a system is sparse iff its body-bar system is $(3,3)$-tight (\textcolor{red}{cite needed?}). The goal is to take a general underconstrained graph and add in some number of bodies and pins to make it well-constrained with as small of a DR plan as possible.

\begin{theorem}
\label{thm:34tight}
    $(3,4)$-tight is equivalent to 1 dof for these body bar systems. 
\end{theorem}

\begin{proof}
    The follows because $(3,3)$-tightness is isostatic.
\end{proof}

\begin{theorem}
    We can get a canonical DR plan using $(3,4)$-tightness and $(3,5)$ tighness as a condition exactly like we have been doing for wellconstrained.
\end{theorem}

\begin{proof}
    It was shown in \cite{XX} that $(3,4)$ and $(3,5)$ tight graphs are matroidal and have a pebble game analog. Thus, we can use similar techniques to find such maximal subgraphs and obtain a similar DR plan as the well-constrained case.
\end{proof}

Suppose we are given a body pin system $G$ and have obtained the DR plan guaranteed by Theorem \ref{thm:34tight} for $G$, denoted $T$. Each node of $T$ will then be a vertex maximal proper 1 dof subgraph of $G$. To make the entire graph wellconstrained, we can add a body and pin it to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$. We can do this easily by considering this theorem.

\begin{theorem}
    Each node of $T$ will fall into one of the following
    \begin{enumerate}
        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on a nontrivial 1 dof subgraph
        \item Its children will be a number of proper maximal 1 dof subgraphs, joined pairwise by at most one pin  
    \end{enumerate}
\end{theorem}

\begin{proof}
    \textcolor{red}{Item 1 Follows from paper?}

    For item 2, consider the case where we have more than 2 proper vertex maximal 1 dof subgraphs $s_1, ..., s_k, k > 2$. Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$ tight and hence $k_i$ and $k_j$ are not vertex maximal
\end{proof}

Thus, the body we need to add to $v$ must be pinned to 2 separate children of $v$. Doing so will ensure the following

\begin{itemize}
    \item All children of $b$ will remain 1 dof
    \item All ancestors of $b$ (including $b$) will now be wellconstrained
\end{itemize}

Then, we can form a valid wellconstrained DR plan $T_b$ from $T$. In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic. Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$ excluding $b'$. Then, for any node $b$ that we choose, $T_b$ is a valid DR plan. Thus, if we want to minimize the size of our DR plan, we simply need to take the $b$ that has the $T_b$ of smallest size. Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.

\begin{theorem}
\label{thm:35tight}

\end{theorem}

### 2014-12-14 doc/GMP2015/s03_drp.tex
Brushed up my sections. Added new figures.

@@
-\section{Decomposition-Recombination Plan}
+\section{The Canonical Decomposition-Recombination Plan}
 
-\begin{figure*}\centering
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
+\subsection{Theory}
 
-    \node[v] (v1) at (1,1) {$R_j$};
-    \node[v] (v2) at (3,1) {$R_k$};
-    \node[v] (v3) at (2,2) {$D$};
-    \node[v] (v4) at (2,3.3) {$R_i$};
+% \begin{figure*}\centering
+% \begin{tikzpicture}
+%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+%     \tikzstyle{e}=[]
 
-    \tedge{v1}{v3}{solid}{}{above};
-    \tedge{v2}{v3}{solid}{}{above};
-    \tedge{v3}{v4}{solid}{}{right};
-\end{tikzpicture}
-\ldots with weights \ldots
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
+%     \node[v] (v1) at (1,1) {$R_j$};
+%     \node[v] (v2) at (3,1) {$R_k$};
+%     \node[v] (v3) at (2,2) {$D$};
+%     \node[v] (v4) at (2,3.3) {$R_i$};
 
-    \node[v] (v1) at (1,1) {y};
-    \node[v] (v2) at (3,1) {z};
-    \node[v] (v3) at (2,2) {k};
-    \node[v] (v4) at (2,3.3) {x};
+%     \tedge{v1}{v3}{solid}{}{above};
+%     \tedge{v2}{v3}{solid}{}{above};
+%     \tedge{v3}{v4}{solid}{}{right};
+% \end{tikzpicture}
+% \ldots with weights \ldots
+% \begin{tikzpicture}
+%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+%     \tikzstyle{e}=[]
 
-    \tedge{v1}{v3}{solid}{y}{above};
-    \tedge{v2}{v3}{solid}{z}{above};
-    \tedge{v3}{v4}{solid}{x}{right};
+%     \node[v] (v1) at (1,1) {y};
+%     \node[v] (v2) at (3,1) {z};
+%     \node[v] (v3) at (2,2) {k};
+%     \node[v] (v4) at (2,3.3) {x};
 
-    % \draw (3.3,2.5) circle (1.9cm);
-    % \draw (.7,2.5) circle (1.9cm);
-    % \draw (2,.5) circle (1.9cm);
-\end{tikzpicture}
-\\
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
+%     \tedge{v1}{v3}{solid}{y}{above};
+%     \tedge{v2}{v3}{solid}{z}{above};
+%     \tedge{v3}{v4}{solid}{x}{right};
 
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxy};
-    \node[v] (v3) at ( 1,-1) {kyz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v5}{solid}{}{};
-    \tedge{v3}{v5}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxy};
-    \node[v] (v3) at ( 1,-1) {kxz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v5}{solid}{}{};
-    \tedge{v3}{v4}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{e}=[]
-
-    \node[v] (v1) at (0,  0) {kxyz};
-    \node[v] (v2) at (-1,-1) {kxz};
-    \node[v] (v3) at ( 1,-1) {kyz};
-    \node[v] (v4) at (-2,-2) {kx};
-    \node[v] (v5) at ( 0,-2) {ky};
-    \node[v] (v6) at ( 2,-2) {kz};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v4}{solid}{}{};
-    \tedge{v2}{v6}{solid}{}{};
-    \tedge{v3}{v5}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-\end{tikzpicture}
+%     % \draw (3.3,2.5) circle (1.9cm);
+%     % \draw (.7,2.5) circle (1.9cm);
+%     % \draw (2,.5) circle (1.9cm);
+% \end{tikzpicture}
 % \\
@@
 
-%     \node[v] (v1) at ( 0, 0) {kxyz};
+%     \node[v] (v1) at (0,  0) {kxyz};
 %     \node[v] (v2) at (-1,-1) {kxy};
-%     \node[v] (v3) at (-2,-2) {kx};
-%     \node[v] (v4) at (-3,-3) {k};
-%     \node[v] (v5) at ( 1,-1) {z plan};
-%     \node[v] (v6) at ( 0,-2) {y plan};
-%     \node[v] (v7) at (-1,-3) {x plan};
-%     \node[v] (v8) at (-2,-4) {k plan};
+%     \node[v] (v3) at ( 1,-1) {kyz};
+%     \node[v] (v4) at (-2,-2) {kx};
+%     \node[v] (v5) at ( 0,-2) {ky};
+%     \node[v] (v6) at ( 2,-2) {kz};
 
 %     \tedge{v1}{v2}{solid}{}{};
-%     \tedge{v1}{v5}{solid}{}{};
-%     \tedge{v2}{v3}{solid}{}{};
+%     \tedge{v1}{v3}{solid}{}{};
+%     \tedge{v2}{v4}{solid}{}{};
+%     \tedge{v2}{v5}{solid}{}{};
+%     \tedge{v3}{v5}{solid}{}{};
+%     \tedge{v3}{v6}{solid}{}{};
+% \end{tikzpicture}
+% \begin{tikzpicture}
+%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+%     \tikzstyle{e}=[]
+
+%     \node[v] (v1) at (0,  0) {kxyz};
+%     \node[v] (v2) at (-1,-1) {kxy};
+%     \node[v] (v3) at ( 1,-1) {kxz};
+%     \node[v] (v4) at (-2,-2) {kx};
+%     \node[v] (v5) at ( 0,-2) {ky};
+%     \node[v] (v6) at ( 2,-2) {kz};
+
+%     \tedge{v1}{v2}{solid}{}{};
+%     \tedge{v1}{v3}{solid}{}{};
+%     \tedge{v2}{v4}{solid}{}{};
+%     \tedge{v2}{v5}{solid}{}{};
+%     \tedge{v3}{v4}{solid}{}{};
+%     \tedge{v3}{v6}{solid}{}{};
+% \end{tikzpicture}
+% \begin{tikzpicture}
+%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+%     \tikzstyle{e}=[]
+
+%     \node[v] (v1) at (0,  0) {kxyz};
+%     \node[v] (v2) at (-1,-1) {kxz};
+%     \node[v] (v3) at ( 1,-1) {kyz};
+%     \node[v] (v4) at (-2,-2) {kx};
+%     \node[v] (v5) at ( 0,-2) {ky};
+%     \node[v] (v6) at ( 2,-2) {kz};
+
+%     \tedge{v1}{v2}{solid}{}{};
+%     \tedge{v1}{v3}{solid}{}{};
+%     \tedge{v2}{v4}{solid}{}{};
 %     \tedge{v2}{v6}{solid}{}{};
-%     \tedge{v3}{v4}{solid}{}{};
-%     \tedge{v3}{v7}{solid}{}{};
-%     \tedge{v4}{v8}{solid}{}{};
+%     \tedge{v3}{v5}{solid}{}{};
+%     \tedge{v3}{v6}{solid}{}{};
 % \end{tikzpicture}
+% % \\
+% % \begin{tikzpicture}
+% %     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+% %     \tikzstyle{e}=[]
 
-\caption{A well-constrained graph whose nodes are subgraphs labeled with their density.
+% %     \node[v] (v1) at ( 0, 0) {kxyz};
+% %     \node[v] (v2) at (-1,-1) {kxy};
+% %     \node[v] (v3) at (-2,-2) {kx};
+% %     \node[v] (v4) at (-3,-3) {k};
+% %     \node[v] (v5) at ( 1,-1) {z plan};
+% %     \node[v] (v6) at ( 0,-2) {y plan};
+% %     \node[v] (v7) at (-1,-3) {x plan};
+% %     \node[v] (v8) at (-2,-4) {k plan};
+
+% %     \tedge{v1}{v2}{solid}{}{};
+% %     \tedge{v1}{v5}{solid}{}{};
+% %     \tedge{v2}{v3}{solid}{}{};
+% %     \tedge{v2}{v6}{solid}{}{};
+% %     \tedge{v3}{v4}{solid}{}{};
+% %     \tedge{v3}{v7}{solid}{}{};
+% %     \tedge{v4}{v8}{solid}{}{};
+% % \end{tikzpicture}
+
+% \caption{A well-constrained graph whose nodes are subgraphs labeled with their density.
 Below are its 3 optimal DR-Plans (partially complete).
@@
 \todo{Make this figure more clear.}}
-\label{3-plans}
-\end{figure*}
+% \label{3-plans}
+% \end{figure*}
 
-In the case of a well-constrained graph, it is clear that there always exists a unique vertex-maximum DR-Plan that includes every single well-constrained vertex-maximal proper subgraph at each node, which we call the complete DR-plan.
-However, in the problem of the optimal DR-Plan there is no such unique plan.
-For instance, a graph that is formed by the union of 3 well-constrained subgraphs that all intersect on a well-constrained subgraph, demonstrated in Figure \ref{3-plans}.
-In this case, it results in $3$ unique optimal DR-Plans.
-Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
+In the problem of the optimal DR-Plan there is generally not a unique plan.
+For example, a graph that is formed by the union of 3 well-constrained subgraphs that all intersect on a well-constrained subgraph has ${3\choose 2} = 3$ unique optimal DR-Plans; any two of these subgraphs will suffice as children.
+% Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
 Therefore, all choices of decomposition are in some sense equivalent.
 The theorem we seek to prove is thus:
+However, we will show that regardless of which children are chosen for the plan, the recombination will require solving of the same systems.
+Thus, this definition of an optimal DR-plan could be considered the canonical DR-plan.
+To assist in showing this, we prove this core theorem throughout this section:
 
@@
 
-We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
+% We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
 By definition, any subgraph of a well-constrained graph will either be (1) trivial, (2) under-constrained, or (3) well-constrained.
@@
 
-\subsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
+\subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
 For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
@@
 
-\subsection{Inferring Graph Structure from Union and Intersection Properties}
+\subsubsection{Inferring Parent Structure from Union and Intersection of Children}
 
@@
     \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
+The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.
 
-    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that is left!
+    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that remains.
 \end{itemize}
@@
 \begin{proof}
-If any intersection was not trivial, then it must be well-constrained.
-If $C_k\cap C_l$ were well-constrained, then by Lemma \ref{wc-l3} all of the intersections are well-constrained.
+If some $C_i\cap C_k$ is not trivial, then it must be well-constrained.
+Then, by Lemma \ref{wc-l3}, all of the intersections are well-constrained.
 This contradicts the statement that $C_i\cap C_j$ is trivial.
@@
 
-\subsection{Proving the Main Theorem}
+\subsubsection{Proving the Main Theorem \ref{t1}}\label{sec:maintheorem}
 Lemmas \ref{wc-l3} and \ref{uc-l2} show that $CompleteDRP(C)$ will result in one of two scenarios.
@@
 
-\begin{figure}\centering
-\begin{tikzpicture}
-    \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
-    \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
-    \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);
+% \begin{figure}\centering
+% \begin{tikzpicture}
+%     \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
+%     \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
+%     \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);
 
-    % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
-    % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);
+%     % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
+%     % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);
 
-    % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
-    % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
-\end{tikzpicture}
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
-    \tikzstyle{e}=[]
+%     % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
+%     % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
+% \end{tikzpicture}
+% \begin{tikzpicture}
+%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+%     \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
+%     \tikzstyle{e}=[]
 
-    \node[el] (m) at (0,    0) {k};
-    \node[v]  (l) at (-1.5, 0) {b};
-    \node[v]  (r) at (1.5,  0) {c};
+%     \node[el] (m) at (0,    0) {k};
+%     \node[v]  (l) at (-1.5, 0) {b};
+%     \node[v]  (r) at (1.5,  0) {c};
 
-    \node[el] (tl) at (-1,  1) {i};
-    \node[el] (tr) at (1,   1) {j};
-    \node[v]  (t)  at (0,   2) {a};
+%     \node[el] (tl) at (-1,  1) {i};
+%     \node[el] (tr) at (1,   1) {j};
+%     \node[v]  (t)  at (0,   2) {a};
 
-    % \node[el] (bl) at (-1,  -1) {l};
-    % \node[el] (br) at (1,   -1) {m};
-    % \node[v]  (b)  at (0,   -2) {d};
+%     % \node[el] (bl) at (-1,  -1) {l};
+%     % \node[el] (br) at (1,   -1) {m};
+%     % \node[v]  (b)  at (0,   -2) {d};
 
-    % \node[el] (v2) at (-1,-1) {b};
-    % \node[el] (v3) at (-1,-1) {b};
-    % \node[el] (v4) at (-1,-1) {b};
-    % \node[el] (v5) at (-1,-1) {b};
+%     % \node[el] (v2) at (-1,-1) {b};
+%     % \node[el] (v3) at (-1,-1) {b};
+%     % \node[el] (v4) at (-1,-1) {b};
+%     % \node[el] (v5) at (-1,-1) {b};
 
-    \tedge{m}{l}{solid}{}{};
-    \tedge{m}{r}{solid}{}{};
+%     \tedge{m}{l}{solid}{}{};
+%     \tedge{m}{r}{solid}{}{};
 
-    \tedge{l}{tl}{solid}{}{};
-    \tedge{tl}{t}{solid}{}{};
-    \tedge{r}{tr}{solid}{}{};
-    \tedge{tr}{t}{solid}{}{};
+%     \tedge{l}{tl}{solid}{}{};
+%     \tedge{tl}{t}{solid}{}{};
+%     \tedge{r}{tr}{solid}{}{};
+%     \tedge{tr}{t}{solid}{}{};
 
-    % \tedge{l}{bl}{solid}{}{};
-    % \tedge{bl}{b}{solid}{}{};
-    % \tedge{r}{br}{solid}{}{};
-    % \tedge{br}{b}{solid}{}{};
-\end{tikzpicture}
+%     % \tedge{l}{bl}{solid}{}{};
+%     % \tedge{bl}{b}{solid}{}{};
+%     % \tedge{r}{br}{solid}{}{};
+%     % \tedge{br}{b}{solid}{}{};
+% \end{tikzpicture}
 
-\caption{Each node contains a subgraph of the entire, well-constrained graph $G$.
+% \caption{Each node contains a subgraph of the entire, well-constrained graph $G$.
 We take $Idc(G,a\cup i\cup b)$, $Idc(G,a\cup j\cup c)$, and $Idc(G,b\cup k\cup c)$ to be well-constrained.
 If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c$ must all be trivial subgraphs.
-% Each node contains a subgraph of the entire, well-constrained graph.
+% % Each node contains a subgraph of the entire, well-constrained graph.
 We take $a\cup i\cup b$, $a\cup j\cup c$, $b\cup k\cup c$, $b\cup l\cup d$, and $c\cup m\cup d$ to be well-constrained.
 If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c,d$ must all be trivial subgraphs.
-% THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
-}
-\label{trivial-intersection}
-\end{figure}
-
-\begin{theorem}\label{uc-t1}
-% In any solution to $OptimalDRP(G)$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is under-constrained, then all well-constrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.
-
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial (case (B)), then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
-\end{theorem}
-
-\begin{proof}
-From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be under-constrained.
-If it is under-constrained, it must be part of some larger well-constrained subgraph $D$.
-If $D$ is not $G$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
- % For $C$ to be well-constrained... WHAT????
-
-If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
-
-This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid.
-Such an example is shown in Figure \ref{trivial-intersection}.
-\end{proof}
-
-% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$.
-All of these subgraphs form a complex structure that is well-constrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).
+% % THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
+% }
+% \label{trivial-intersection}
+% \end{figure}
 
@@
 
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained (case (A)), then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
+In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
 % Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
@@
 \begin{proof}
-For a valid DR-plan, all children must be proper subgraphs.
-Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least 2 children.
-It follows directly from Lemma \ref{wc-l2} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then you found a tree with only 2 children.
+In a valid DR-plan, all children must be proper subgraphs.
+Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
+It follows directly from Lemma \ref{wc-l3} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
+% you found a tree with only 2 children.
+% Clearly, this is the optimal choice at this level.
+However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
 
-Clearly, this is the optimal choice at this level. However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in later on.
+Take $D=\bigcap_{k=1}^N{C_k}$ and $R_k=C\setminus C_k$.
+% This means that $D$ is also $\bigcup_{k=1}^N{R_k}$.
+Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
+% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
+% the children of this node will be
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
+% and
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
+$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
+the children of this node will be
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
+and
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
+for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two, at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
 
-Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. Take $R_k=C\setminus C_k$. This means that $D$ is also $\bigcup_{k=1}^N{R_k}$.
-
-After we go down $N$ levels we get to a point where every node is some $Idc(C,R_k\cup D)$. Until this point, the fan-in of every node is exactly 2.
-
-This is demonstrated with $N=3$ in Figure \ref{3-plans}.
+% This is demonstrated with $N=3$ in Figure \ref{3-plans}.
 It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution.
@@
 
-With these two proofs, the original Theorem \ref{t1} has been proven.
-With this we have formulated the notion of a canonical optimal DR-plan.
-Any optimal DR-plan of a well-constrained graph must follow this pattern.
+\begin{theorem}\label{uc-t1}
+% In any solution to $OptimalDRP(G)$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is under-constrained, then all well-constrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.
+
+In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
+\end{theorem}
+
+% \begin{proof}
+% From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be under-constrained.
+If it is under-constrained, it must be part of some larger well-constrained subgraph $D$.
+If $D$ is not $C$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
+%  % For $C$ to be well-constrained... WHAT????
+
+% If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
+
+% This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid.
+Such an example is shown in Figure \ref{trivial-intersection}.
+% \end{proof}
+
+\begin{proof}
+Take $X\subset \{1,\ldots,N\}$ and $D=\bigcup_{i\in X}{C_i}$.
+If $D\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
+
+\todo{Uses 2D requirement:}
+If $D= C$, then we contradict the assumption that $C_i \cap C_j$ is trivial.
+By Lemma \ref{uc-l2}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
+Furthermore, by definition of a DR-plan, $C_k=C\cap C_k=D\cap C_k$.
+Together, this shows that $C_k$ cannot be well-constrained, violating the properties of a DR-plan.
+\end{proof}
+
+% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$.
+All of these subgraphs form a complex structure that is well-constrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).
+
+With these two proofs, the original Theorem \ref{t1} has been proven. With this we have formulated the notion of a canonical DR-plan.
+% Any optimal DR-plan of a well-constrained graph must follow this pattern.
+No matter which two children you choose in case (A) the recombination steps will always involve the same systems; in case (B), all children must always be included.
 
@@
 
-\subsection{Further Results}
+% \subsection{Further Results}
 
-\begin{corollary}
-Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is well-constrained, then $D$ will be a descendant of every $C_k$.
-\end{corollary}
+% \begin{corollary}
+% Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is well-constrained, then $D$ will be a descendant of every $C_k$.
+% \end{corollary}
 
-\begin{proof}
-Explained in Theorem \ref{wc-t1}.
-\end{proof}
+% \begin{proof}
+% Explained in Theorem \ref{wc-t1}.
+% \end{proof}
 
-\begin{corollary}
-$\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
-\end{corollary}
+% \begin{corollary}
+% $\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
+% \end{corollary}
 
-\begin{proof}
-Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
-\end{proof}
+% \begin{proof}
+% Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
+% \end{proof}
 
@@
 
-\begin{corollary}
-There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-\end{corollary}
+% \begin{corollary}
+% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
+% \end{corollary}
 
-\begin{proof}
-Since $C_1\cup C_2$ is the entire graph
-\todo{Prove this!!!!}
-\end{proof}
+% \begin{proof}
+% Since $C_1\cup C_2$ is the entire graph
+% \todo{Prove this!!!!}
+% \end{proof}
 
@@
 % \end{proof}
+
+\subsection{Examples}
+
+We now present two examples of non-tree-decomposable graphs that illustrate the two cases found in decomposition -- intersection on a well-constrained subgraph and intersection on a trivial subgraph.
+
+\begin{figure*}\centering
+\begin{tikzpicture}[scale=3]
+    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+    \tikzstyle{e}=[]
+
+    \node[v] (v1) at (0,0.866) {$a_1$};
+    \node[v] (v2) at (-1,-0.866) {$a_2$};
+    \node[v] (v3) at (1,-0.866) {$a_3$};
+
+    \node[v] (v4) at (0,0.577-.1) {$b_1$};
+    \node[v] (v5) at (-0.667,-0.577-.1) {$b_2$};
+    \node[v] (v6) at (0.667,-0.577-.1) {$b_3$};
+
+    \node[v] (v7) at (0,0.289-.2) {$c_1$};
+    \node[v] (v8) at (-0.333,-0.289-.2) {$c_2$};
+    \node[v] (v9) at (0.333,-0.289-.2) {$c_3$};
+
+    \tedge{v1}{v2}{solid}{}{};
+    \tedge{v1}{v3}{solid}{}{};
+    \tedge{v2}{v3}{solid}{}{};
+
+    \tedge{v4}{v5}{solid}{}{};
+    \tedge{v4}{v6}{solid}{}{};
+    \tedge{v5}{v6}{solid}{}{};
+
+    \tedge{v7}{v8}{solid}{}{};
+    \tedge{v7}{v9}{solid}{}{};
+    \tedge{v8}{v9}{solid}{}{};
+
+    \tedge{v1}{v4}{solid}{}{};
+    \tedge{v2}{v5}{solid}{}{};
+    \tedge{v3}{v6}{solid}{}{};
+
+    \tedge{v4}{v7}{solid}{}{};
+    \tedge{v5}{v8}{solid}{}{};
+    \tedge{v6}{v9}{solid}{}{};
+\end{tikzpicture}
+
+\begin{tikzpicture}
+    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+    \tikzstyle{e}=[]
+
+    \node (c0) at (0,0) {$\{a_1,a_2,a_3,b_1,b_2,b_3,c_1,c_2,c_3\}$};
+    \node (c1a) at (-2,-1) {$\{a_1,a_2,a_3,b_1,b_2,b_3\}$};
+    \node (c1b) at (2,-1) {$\{b_1,b_2,b_3,c_1,c_2,c_3\}$};
+    \node (c2a) at (-2,-2) {$\{a_1,a_2,a_3\}$};
+    \node (c2b) at (0,-2) {$\{b_1,b_2,b_3\}$};
+    \node (c2c) at (2,-2) {$\{c_1,c_2,c_3\}$};
+    \node (c2ab1) at (-5.6,-2) {$\{a_1,b_1\}$};
+    \node (c2ab2) at (-4.5,-2) {$\{a_2,b_2\}$};
+    \node (c2ab3) at (-3.4,-2) {$\{a_3,b_3\}$};
+    \node (c2bc1) at (5.6,-2) {$\{b_1,c_1\}$};
+    \node (c2bc2) at (4.5,-2) {$\{b_2,c_2\}$};
+    \node (c2bc3) at (3.4,-2) {$\{b_3,c_3\}$};
+
+    \tedge{c0}{c1a}{solid}{}{};
+    \tedge{c0}{c1b}{solid}{}{};
+
+    \tedge{c1a}{c2ab1}{solid}{}{};
+    \tedge{c1a}{c2ab2}{solid}{}{};
+    \tedge{c1a}{c2ab3}{solid}{}{};
+    \tedge{c1a}{c2a}{solid}{}{};
+    \tedge{c1a}{c2b}{solid}{}{};
+
+    \tedge{c1b}{c2bc1}{solid}{}{};
+    \tedge{c1b}{c2bc2}{solid}{}{};
+    \tedge{c1b}{c2bc3}{solid}{}{};
+    \tedge{c1b}{c2b}{solid}{}{};
+    \tedge{c1b}{c2c}{solid}{}{};
+\end{tikzpicture}
+
+\caption{Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
+Also showing the optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
+\label{doublets}
+\end{figure*}
+
+In figure \ref{doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
+These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$.
+This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{wc-t1}.
+Also shown is the decomposition of the individual doublets.
+These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them.
+As explained in Theorem \ref{uc-t1}, all of them must be included in the optimal DR-plan.
+Were any left out, the union of the children would not be the entire parent.
+
+\begin{figure*}\centering
+\begin{tikzpicture}[scale=3]
+    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+    \tikzstyle{e}=[]
+
+    \node[v] (v1) at (0,0.866) {$a_1$};
+    \node[v] (v2) at (-1,-0.866) {$a_2$};
+    \node[v] (v3) at (1,-0.866) {$a_3$};
+
+    \node[v] (v4) at (0,0.433-.125) {$b_1$};
+    \node[v] (v5) at (-0.5,-0.433-.125) {$b_2$};
+    \node[v] (v6) at (0.5,-0.433-.125) {$b_3$};
+
+    \tedge{v1}{v2}{solid}{}{};
+    \tedge{v1}{v3}{solid}{}{};
+    \tedge{v2}{v3}{solid}{}{};
+
+    \tedge{v4}{v5}{solid}{}{};
+    \tedge{v4}{v6}{solid}{}{};
+    \tedge{v5}{v6}{solid}{}{};
+
+    \tedge{v1}{v4}{solid}{}{};
+    \tedge{v2}{v5}{solid}{}{};
+    \tedge{v3}{v6}{solid}{}{};
+
+    % sin(30deg) = 0.5
+    % cos(30deg) = 0.866
+
+    % o/i -> outside/inside triangle
+    % b/l/r -> bottom/left/right edge of triangle
+
+    \coordinate (ob0) at (-0.333,-0.866-0.1);
+    \coordinate (ob1) at (0.333,-0.866-0.1);
+    \coordinate (ob2) at (-0.333,-0.866+0.1);
+    \coordinate (ob3) at (0.333,-0.866+0.1);
+    \draw (v2) -- (ob0) -- (ob1) -- (v3);
+    \draw (v2) -- (ob2) -- (ob3) -- (v3);
+    \draw (ob0) -- (ob3);
+    \draw (ob2) -- (ob1);
+
+    \coordinate (ib0) at (-0.167,-0.433-.125-0.1);
+    \coordinate (ib1) at (0.167,-0.433-.125-0.1);
+    \coordinate (ib2) at (-0.167,-0.433-.125+0.1);
+    \coordinate (ib3) at (0.167,-0.433-.125+0.1);
+    \draw (v5) -- (ib0) -- (ib1) -- (v6);
+    \draw (v5) -- (ib2) -- (ib3) -- (v6);
+    \draw (ib0) -- (ib3);
+    \draw (ib2) -- (ib1);
+
+    % \draw[rotate around={60:(-1,-0.866)}] (v2) -- (-0.333,-0.766) -- (0.333,-0.766) -- (v1);
+    % \draw (v2) -- (ob2) -- (ob3) -- (v1);
+    % \draw (ob0) -- (ob3);
+    % \draw (ob2) -- (ob1);
+% \newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
+
+    % \draw (-1,-0.866) -- (-0.333,-0.966);
+
+\end{tikzpicture}
+
+\caption{A doublet with each edge replaced by a $K3,3$. \todo{Drawing incomplete.}}
+\label{c2c3ofk33s}
+\end{figure*}
+
+In figure \ref{c2c3ofk33s}... \todo{Is it worth working through this? The first example has trivial and non-trivial.}
+
+\subsection{Extensions}
+This framework immediately pushes through for body-pin systems via a simple reduction.
+If there are $N$ pins on a body, it can be represented as a 2-tree with $N$ vertices, each corresponding to a pin, making sure to select edge distances such that the distance between pins is preserved.
+E.g.\ a body with two pins is an edge, three pins is a triangle, etc.
+Any bodies that share a pin now intersect on their vertex that corresponds to that pin.
+Now we have a bar-joint representation of the body-pin system in 2D and all proofs follow.
+
+With more effort, it can be shown that pinned line-incidence systems can also use this framework. This is done in section \ref{XXX}.
+
+\subsection{Algorithm}
+The algorithm to find the optimal DR-plan is a straightforward application of the theory.
+It begins by finding the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
+If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
+Otherwise, all the subgraphs are the children.
+Apply recursively to the children.
+This naturally terminates when you reach trivial subgraphs.

### 2014-12-14 doc/GMP2015/s07_problempinning.tex
Added more to Problem pinning

@@
 
-To answer these questions, we first restrict ourselves to a specific case of Body-Pin networks such that:
+To answer these questions, we first inroduce the structures we model the problems with:
 
+\begin{definition}
+    A multi-body-pin qusec is a qusec where the obejct are rigid bodies that are pinned together by some number of pins.
+\end{definition}
+
+\begin{remark}
+    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying. As such, the DR planning discussed in Sections \ref{dr} and \ref{recomb} will still go through.
+\end{remark}
+
+\begin{proof}
+    We can replace each body that has only one pin by a single vertex.
+A body with 2 pins can be replaced by an edge.
+A body with 3 pins can be replaced by a triangle.
+In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
+When looking for a DR plan, we treat each body as trivial, so they become the leaves of our plan.
+\end{proof}
+
+In this section we will go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{intro}.
+
+\begin{defintion}
+    A body-pin qusec is a multi-body-pin qusec with the following conditions:
     \begin{itemize}
@@
     \end{itemize}
+    Such a body-pin qusec can also be seen as a {\em body-bar qusec} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
+For such body-bar qusecs, graphs with 1 and 2 dofs can be characterized by being $(3,4)$ and $(3,5)$ tight respectively.
+\end{defintion}
 
-Such a body-pin system can also be seen as a {\em body-bar system} where the bodies are the original bodies and each pin represents 2 edges from one body to another.
-Then, such a system is sparse iff its body-bar system is $(3,3)$-tight (\textcolor{red}{cite needed?}).
-The goal is to take a general underconstrained graph and add in some number of bodies and pins to make it well-constrained with as small of a DR plan as possible.
+\begin{definition}
+    A multi-triangle-pin qusec is a multi-body-pin qusec where each body is a triangle, meaning it is pinned in at most 3 places.
+This is also represented as a hyper qusec where each pin is a vertex and each triangle represents a tri-hyperedge.
+For such hyper qusecs, 1 and 2 dofs can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
+\end{definition}
 
-\begin{theorem}
-\label{thm:34tight}
-    $(3,4)$-tight is equivalent to 1 dof for these body bar systems. 
-\end{theorem}
+Body-pin qusecs are of particular interest to us in the context of Example 4 in \ref{intro}.
+Multi-triangle-pin qusec can be used to reprsent the silica bi-layers described in Example 5 of \ref{intro}, where each triangle is the junction of ``disks''.
+Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR plan:
 
-\begin{proof}
-    The follows because $(3,3)$-tightness is isostatic.
-\end{proof}
+\begin{defnition}
+    A $(k,l)$ tight DR plan is one in which each child node is either a vertex maximal proper $(k,l)$ tight subgraph of the parent node or it is trivial.
+In our case, the trivial nodes are just the bodies.
+\end{defnition}
 
-\begin{theorem}
-    We can get a canonical DR plan using $(3,4)$-tightness and $(3,5)$ tighness as a condition exactly like we have been doing for wellconstrained.
-\end{theorem}
+We can develop the notion of a canonical DR plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{dr}.
 
-\begin{proof}
-    It was shown in \cite{XX} that $(3,4)$ and $(3,5)$ tight graphs are matroidal and have a pebble game analog.
-Thus, we can use similar techniques to find such maximal subgraphs and obtain a similar DR plan as the well-constrained case.
-\end{proof}
+\begin{remark}
+\label{rem:1dofcanon}
+    For the 1 dof multi-body-pin qusecs described above, a DR plan node will follow one of the following:
 
-Suppose we are given a body pin system $G$ and have obtained the DR plan guaranteed by Theorem \ref{thm:34tight} for $G$, denoted $T$.
-Each node of $T$ will then be a vertex maximal proper 1 dof subgraph of $G$.
-To make the entire graph wellconstrained, we can add a body and pin it to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$.
-We can do this easily by considering this theorem.
-
-\begin{theorem}
-    Each node of $T$ will fall into one of the following
-    \begin{enumerate}
-        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on a nontrivial 1 dof subgraph
+    \begin{itemize}
+        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on a single body
         \item Its children will be a number of proper maximal 1 dof subgraphs, joined pairwise by at most one pin  
-    \end{enumerate}
-\end{theorem}
+    \end{itemize}
+\end{remark}
 
@@
     For item 2, consider the case where we have more than 2 proper vertex maximal 1 dof subgraphs $s_1, ..., s_k, k > 2$.
-Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$ tight and hence $k_i$ and $k_j$ are not vertex maximal
+Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$ tight and hence $k_i$ and $k_j$ are not vertex maximal.
 \end{proof}
 
-Thus, the body we need to add to $v$ must be pinned to 2 separate children of $v$. Doing so will ensure the following
+\begin{remark}
+    All $(k,l)$ tight canonical DR plans are optimal. We can find such a DR plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{dr}.
+\end{remark}
+
+\begin{theorem}
+    There is a quadratic algorithm for the 1 dof and 2 dof optimal completion problem of Seciton \ref{recomb} for body-pin and multi-triangle-pin qusecs.
+\end{theorem}
+
+\begin{proof}
+    Suppose we are given a body-pin qusec $G$ and have obtained the DR plan denoted $T$.
+Each node of $T$ will then be a vertex maximal proper 1 dof subgraph of $G$.
+To make the entire graph wellconstrained, we can add a body and pin it to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$.
+From Remark \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a single body, so we simply have to pin the new body to one part on each side of the pin or body.
+Doing so will ensure the following
 
@@
 
-Then, we can form a valid wellconstrained DR plan $T_b$ from $T$.
+    Then, we can form a valid isostatic DR plan $T_b$ from $T$.
 In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
@@
 
-\begin{theorem}
-\label{thm:35tight}
+    For the 2 dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
+These can be the same node, and if it is the same node, we can find a wellconstrained DR plan in quadratic time, we would just be doing the same thing as the 1 dof case.
+\end{proof}
 
-\end{theorem}
+%TODO write out the stuff for 2 dof

### 2014-12-14 doc/GMP2015/s02_preliminaries.tex

@@
 
+% TODO
+
+% A paragraph of preliminaries such as rigidity matrix R, stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex)); flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R) should precede the definition of independent, overconstrained (not independent), rigid (contains an independent set with sufficiently many eedges/constraints) isostatic (wellconstrained/minimally rigid), , flexible (not rigid), underconstrained (independent and not rigid).
+
+% After that, define DR plans etc.
+
+\begin{definition}
    In the context of this paper, a \textbf{graph} $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
 Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
+\end{definition}
 
+\begin{definition}
     The \textbf{density} of graph $G$ is
@@
     \end{equation}
+\end{definition}
 
+\begin{definition}
     A graph is \textbf{independent} (or \textbf{sparse}) if \todo{define}.
+\end{definition}
 
+\begin{definition}
     A graph is \textbf{over-constrained} if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$.
+\end{definition}
 
+\begin{definition}
     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
+\end{definition}
 
+\begin{definition}
     A graph is \textbf{rigid} if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+\end{definition}
 
+\begin{definition}
     A graph is \textbf{under-constrained} if, given constant $k$, $G$ is not rigid.
+\end{definition}
 
+\begin{definition}
     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
@@
 These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
+\end{definition}
 
+\begin{definition}
     The \textbf{decomposition-recombination (DR-) plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
@@
     \end{enumerate}
+
     % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
@@
     Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
+\end{definition}
 
-The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \textit{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+\begin{definition}
+    The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
 By this, rule 3 is implicit for a complete DR-plan.
+\end{definition}
 
+\begin{definition}
     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+\end{definition}
 
@@
 
-$G$ is taken to be a graph with values $(V,E,w_V,w_E)$.
-To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
+$G$ is taken to be a graph with values $(V,E,w_V,w_E)$. The most important limitation of this paper is that we take $G$ to be well-constrained.
+\todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
 It is not useful to solve for two bodies simultaneously if there are no constraints between them.
-Furthermore, to limit the scope of the paper, $G$ will always be taken to be well-constrained.
 % THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.

### 2014-12-14 doc/GMP2015/s08_pinnedline.tex
Added section on pinned line incidence

\section{Application: cross-linking microfibrils}

The optimal qusecs DR-plan developed in this paper can be applied to study 
the structure of  cross-linking microfibrils in collagen and cellulose in biology.

Collagen is an important protein material in biological tissues with highly elastic mechanical property. 
Cellulose is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}). 
Both of these substances consist of a large number microfibrils,
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally 
each of which is cross-linked at 2 places with some two other fibrils, 
where the cross-linking is like an incidence constraint that 
the two crosslinked fibrils can slide with each other (see Figure~\ref{fig:cross-linking}(a). ).

%Collagen
%
%layers of microfibril with cross-linking??? 
%
%
%- different cross-linking densities
%larger cross-link densities lead to larger yield strains, larger yield stresses as well as larger fracture stresses.
%the maximum fracture stress of the collagen fibril does not increase with increasing cross-link densities (2 links per molecule)
%
%
%Plant cell wall, cellulose
%
%Non-charged cellulose is considered to form a random mesh, the rigidity of which would depend on the number of interfibriller cross links, i.e. H bonds.
%new microfibrils were laid down on top of the existing ones, forming a new layer.

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.5\linewidth]{img/AFM_Innventia_nanocellulose}
\caption{Fibrils of carboxymethylated nanocellulose adsorbed on a silica surface. \cite{xxx}}
\label{fig:cellulose}
\end{figure}

\subsection{Modeling the fibrils as a pinned line incidence system}

The cross-linking of microfibrils can be modeled using 2-dimensional pinned line incidence constraint system.

%TODO system? problem?
\begin{definition} [2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
A simple graph $G=(V,E)$ is a 2-dimensional pinned line incidence graph, 
if there is  a  set $X$ of $|E|$ pins with fixed positions in $\mathbb{R}^2$.
such that 
 every edge $e \in E$ corresponds to a pin $x \in X$ where $e$ is constrained to lie on a line passing through $x$.
%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$

%Let $X$ be a  set of $m$ pins that are fixed in $\mathbb{R}^2$, and $V$ be the index set of a set of vertices. 
%For every pin $x \in X$, we are also given the indices of two vertices $v_1,v_2$ in $V$, 
%such that the line (edge) spanned by $v_1$ and $v_2$ must pass through $x$. 
% % % % % % % % % % % % % % % % % % %
%Find any such set of points $P$ satisfying the given incidences. 
%For every pin $x\in X$, we are also given a line $\supp{x}$, %of %a set of $s$-dimensional subspaces passing through 
%i.e, a size-two index subset of an unknown set of points $D = \{v_1,\ldots,v_n\}$ in $\mathbb{R}^2$, %$P = \{p_1,\ldots, p_n \}$, 
%such that  $x_i$ lies on the line spanned by $\supp{x}$. 
%%assume additionally that  
%%no $s$ pins lie on any single subspace defined by one of the above hyperedges.
%Find any such set D %$P[X]$ 
%that satisfies the given subspace incidences. % specified by $\supp[P]{x}$.%$\{S_i\}$. % on $P[X]$ and $X$.
\end{definition}

In the case of microfibril cross-linking, each fibril is 
attached to some fixed larger organelle/membrane at one spot, 
so it can be modeled as an edge/line segment incident at a pin in $X$. 
%For cross-linking density 2, 
%i.e.\ each fibril has two cross-linkings with other fibrils, 
The cross-linkings can be modeled as vertices in $V$.

\begin{figure}
\centering
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=\linewidth]{img/crosslink}
  \caption{}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.8\linewidth]{img/teq1}
  \caption{}
  \label{fig:sub2}
\end{subfigure}
\caption{(a) Cross-linking of microfibrils~\cite{xxx}. (b) A pinned line incidence graph.}
\label{fig:cross-linking}
\end{figure}

As an example, Figure~\ref{fig:cross-linking}, 
where the grey ovals stand for pins/attachments of fibrils and the black dots stand for vertices/cross-linkings.
There are 5 edges/fibrils,  and 10 vertices/cross-linkings in the graph.

\subsection{Optimal DR-plan for pinned line incidence systems}

%Recall the definitions in Section~\ref{priliminary}.

In this section, we will adapt the results in Section~\ref{xxx}
to give the Optimal DR-plan for 2-dimensional pinned line incidence graphs.
%The 2-dimensional pinned line incidence graphs and their DR-plans have 
Note the following differences from the bar-joint graphs:

\begin{itemize}

\item For pinned line incidence graphs, all vertex weights are $2$, all edge weights are $1$, 
and the constant $k = 0$ \cite{xxxx}. 
%In other words, a pinned line incidence graph is well-constrained if $2|V| = |E|$ and $2|V'| \le |E'|$ for all induced non-trivial subgraphs $S=(V',E') \subseteq G$.

\item 
\todo{Trivial graphs (must be overconstrained in the original definition) does not exist??? }\
\todo{Both a single vertex and a single edge are underconstrained}
\begin{itemize}

\item For  pinned line incidence graphs:
no trivial motion exists since the pins have fixed locations.

\item 
\todo{We just define the trivial graph/DR-plan leaf for 2-dimensional pinned line incidence graphs to be a single vertex.  }

\item \todo{A single edge can be a node of the DR-plan}

\end{itemize}

\item 
Notice that for pinned line incidence graphs, 
since the pins in $X$ have fixed positions on the plane, 
a well-constrained vertex-maximal proper subgraphs can be disconnected.\\
Without loss of generality, 
we slightly modify the requirement of DR-plan for pinned line incidence graphs
such that only connected subgraphs can be considered as a node of the DR-plan.

\end{itemize}

With the above modification, 
it is not hard to adapt Theorem~\ref{t1} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
to show the structure of optimal DR-plan for 2-dimensional pinned line incidence graphs.

\begin{corollary}
Given a well-constrained 2-dimensional pinned line incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
\begin{enumerate}
    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
    \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
\end{enumerate}
\label{cor:pinned}
\end{corollary}

\todo{All Lemmas before Lemma 8 are proved with respect to general constraint systems,
and they go through for 2-dimensional pinned line incidence as well. }
Now we  show that Lemma 8 also holds for 2-dimensional pinned line incidence graphs.

\begin{proof}[Proof of Lemma~\ref{wc-l3} for 2D pinned line incidence graphs]
We use the same notations as in the proof of Lemma~\ref{wc-c1} in Section~\ref{xxx}.
\begin{itemize}
    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary~\ref{wc-c1}) and by our modified requirement of DR-plan nodes.

%    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained. We also know it cannot be trivial because it contains well-constrained subgraphs. This means it must be under-constrained. From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial. This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
%
%    \item 1 case: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The union with $C_i$ (or $C_j$) is not equal to $C$. Therefore, the union must be under-constrained which means the intersection must be trivial (a single node). However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.
%
%    \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible. The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
\end{itemize}

\todo{All other cases are similar to the original proof.}

\end{proof}

%How theorems go through when only considering connected rigidity components

Given Lemma~\ref{wc-l3}, Lemma~\ref{uc-l2}, Theorem~\ref{uc-t1} and \ref{wc-t1} also apply to  2-dimensional pinned line incidence graphs,
proving Corollary~\ref{cor:pinned}.

Therefore, we can analyze microfibril cross-linkings by finding the optimal DR-plan for  2-dimensional pinned line incidence graphs in the same way as for 2-dimensional bar-joint graphs.

### 2014-12-15 doc/GMP2015/s08_pinnedline.tex
Modified definition of trivial graph

@@
 where the cross-linking is like an incidence constraint that 
-the two crosslinked fibrils can slide with each other (see Figure~\ref{fig:cross-linking}(a). ).
+the two crosslinked fibrils can slide with each other (see Figure~\ref{fig:cross-linking}(a)).
 
@@
 A simple graph $G=(V,E)$ is a 2-dimensional pinned line incidence graph, 
-if there is  a  set $X$ of $|E|$ pins with fixed positions in $\mathbb{R}^2$.
-such that 
- every edge $e \in E$ corresponds to a pin $x \in X$ where $e$ is constrained to lie on a line passing through $x$.
+if there is  a  set $X$ of $|E|$ pins with fixed positions in $\mathbb{R}^2$,
+such that  every edge $e \in E$ corresponds to a pin $x \in X$ where $e$ is constrained to lie on a line passing through $x$.
 %There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
@@
 \item 
-\todo{We just define the trivial graph/DR-plan leaf for 2-dimensional pinned line incidence graphs to be a single vertex.  }
+\todo{We just define the %trivial graph/
+DR-plan leaves for 2-dimensional pinned line incidence graphs to be single vertices...  }
 
@@
 \todo{All Lemmas before Lemma 8 are proved with respect to general constraint systems,
-and they go through for 2-dimensional pinned line incidence as well. }
+and they go through for 2-dimensional pinned line incidence as well (without trivial graph??). }
+
 Now we  show that Lemma 8 also holds for 2-dimensional pinned line incidence graphs.

### 2014-12-15 doc/GMP2015/s02_preliminaries.tex
Clean up + changes to preliminaries.

@@
-\section{Preliminaries}
+\section{Preliminaries}\label{prelim}
 
@@
 
-% A paragraph of preliminaries such as rigidity matrix R, stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex)); flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R) should precede the definition of independent, overconstrained (not independent), rigid (contains an independent set with sufficiently many eedges/constraints) isostatic (wellconstrained/minimally rigid), , flexible (not rigid), underconstrained (independent and not rigid).
-
+% A paragraph of preliminaries such as:
+%     - rigidity matrix R,
+%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
+%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
+% Should precede the definition of
+%     + independent,
+%     + overconstrained (not independent),
+%     - rigid (contains an independent set with sufficiently many eedges/constraints)
+%     - isostatic (wellconstrained/minimally rigid), ,
+%     - flexible (not rigid),
+%     + underconstrained (independent and not rigid).
 % After that, define DR plans etc.
@@
 \begin{definition}
-    A graph is \textbf{independent} (or \textbf{sparse}) if \todo{define}.
+    A graph is \textbf{independent} (or \textbf{sparse}) if, given constant $k$, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
 \end{definition}
 
+% \begin{definition}
+%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
+%     % induced
+%     subgraph $S\subseteq G$ such that $d(S) > k$.
+% \end{definition}
+
 \begin{definition}
-    A graph is \textbf{over-constrained} if, given constant $k$, there exists some induced subgraph $S\subseteq G$ such that $d(S) > k$.
+    A graph is \textbf{over-constrained} if it is not independent.
 \end{definition}
 
+% \begin{definition}
+%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
+%     % induced
+%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
+% \end{definition}
+
 \begin{definition}
-    A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all induced subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
+    A graph is \textbf{well-constrained} (or \textbf{tight}) if, given constant $k$, $G$ is independent and $d(G)=k$.
 \end{definition}
@@
 \begin{definition}
-    A graph is \textbf{under-constrained} if, given constant $k$, $G$ is not rigid.
+    A graph is \textbf{under-constrained} if $G$ is independent and not rigid.
 \end{definition}
@@
         \item The root of the tree `contains' $G$.
-        \item For all nodes, $C$, that `contain' nontrivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
         \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.

### 2014-12-15 doc/GMP2015/s03_drp.tex
Clean up + changes to preliminaries.

@@
 
-% We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
+% We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two non-trivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
 By definition, any subgraph of a well-constrained graph will either be (1) trivial, (2) under-constrained, or (3) well-constrained.
@@
 By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained, or (3) trivial.
-We will analyze all 3 possibilities in turn, in the case of a union of any two nontrivial well-constrained vertex-maximal proper subgraphs.
+We will analyze all 3 possibilities in turn, in the case of a union of any two non-trivial well-constrained vertex-maximal proper subgraphs.
 
@@
 The same is true for $F_j$.
-If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an over-constrained nontrivial graph.
+If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an over-constrained non-trivial graph.
 This is a contradiction, by definition of a well-constrained graph.
@@
 
+\begin{example}
     In figure \ref{doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
@@
 Were any left out, the union of the children would not be the entire parent.
+\end{example}
 
@@
 
+\begin{example}
     In figure \ref{c2c3ofk33s}... \todo{Is it worth working through this? The first example has trivial and non-trivial.}
+\end{example}
 

### 2014-12-15 doc/GMP2015/s07_problempinning.tex
Added more to Problem pinning
Clean up + changes to preliminaries.

@@
 
-We can use DR plans to help in solving questions in designs for graphene and silica bilayers (\textcolor{red}{Add Citation}).
+We can use DR plans to help in solving questions in designs for graphene and silica bilayers (\todo{Add Citation}).
 Specifically, we will look into the problem when you have a monolayer and it is underconstrained.
@@
 
-\begin{defintion}
+\begin{definition}
     A body-pin qusec is a multi-body-pin qusec with the following conditions:
@@
 For such body-bar qusecs, graphs with 1 and 2 dofs can be characterized by being $(3,4)$ and $(3,5)$ tight respectively.
-\end{defintion}
+\end{definition}
 
@@
 
-\begin{defnition}
+\begin{definition}
     A $(k,l)$ tight DR plan is one in which each child node is either a vertex maximal proper $(k,l)$ tight subgraph of the parent node or it is trivial.
 In our case, the trivial nodes are just the bodies.
-\end{defnition}
+\end{definition}
 
@@
 \begin{proof}
-    \textcolor{red}{Item 1 Follows from paper?}
+    \todo{Item 1 Follows from paper?}
 
@@
 \end{proof}
-
-%TODO write out the stuff for 2 dof

### 2014-12-15 doc/GMP2015/s08_pinnedline.tex
Changes in lemmas and proof

@@
 \item 
-\todo{Trivial graphs (must be overconstrained in the original definition) does not exist??? }\
+\todo{Over-constrained trivial subgraphs does not exist??? }\
 \todo{Both a single vertex and a single edge are underconstrained}
@@
 
-\todo{All Lemmas before Lemma 8 are proved with respect to general constraint systems,
-and they go through for 2-dimensional pinned line incidence as well (without trivial graph??). }
+%\todo{All Lemmas before Lemma 8 are proved with respect to general constraint systems,
+%and they go through for 2-dimensional pinned line incidence as well (without trivial graph??). }
+Lemmas similar to Lemma 2--5 hold for 2-dimensional pinned line incidence graphs.
 
-Now we  show that Lemma 8 also holds for 2-dimensional pinned line incidence graphs.
+\begin{lemma}
+Let $F_j$ and $F_j$ be well-constrained subgraphs of the same well-constrained graph $F$. 
+ $F_i \cup F_j$ can only be well-constrained.
+\end{lemma}
+
+\begin{proof}
+\begin{enumerate}
+\item Obviously $F_i \cup F_j$ cannot be over-constrained since it is a subgraph of well-constrained graph $F$.
+\item $F_i \cup F_j$ cannot be under-constrained, otherwise $F_i \cap F_j$ would be over-constrained while there are no trivial over-constrained graphs. 
+\item The only remaining case is that $F_i \cup F_j$ is well-constrained, which is true if and only if $F_i \cap F_j$ is well-constrained.
+\end{enumerate}
+\end{proof}
+
+Lemma 6 and 7 straightforwardly holds for 2-dimensional pinned line incidence graphs. 
+Now we  show that Lemma 8 also holds.
 

### 2014-12-15 doc/GMP2015/s03_drp.tex
Change to figure.

@@
 \begin{figure*}\centering
-\begin{tikzpicture}[scale=3]
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+\begin{subfigure}{.3\linewidth}
+  \centering
+  \begin{tikzpicture}[scale=2]
+        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
         \tikzstyle{e}=[]
@@
 
-\begin{tikzpicture}
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+    \caption{}
+\end{subfigure}%
+\begin{subfigure}{.7\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.9, transform shape]
+        \tikzstyle{v}=[draw, circle, minimum size=0.75cm, font=\footnotesize]
         \tikzstyle{e}=[]
@@
 
-\caption{Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
-Also showing the optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
+    \caption{}
+\end{subfigure}
+
+\caption{(a) Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
+(b) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
 \label{doublets}

### 2014-12-15 doc/GMP2015/s08_pinnedline.tex
Changed section title. Restated the lemmas and several proofs. Changed figure. Mentioned recombination.

--- 2014-12-14 doc/GMP2015/s08_pinnedline.tex
@@
-\section{Application: cross-linking microfibrils}
+\section{Application: finding optimal DR-plans and realizations for cross-linking microfibrils }
 
-The optimal qusecs DR-plan developed in this paper can be applied to study 
-the structure of  cross-linking microfibrils in collagen and cellulose in biology.
+The optimal qusecs DR-plan can also be applied to analyze and solve
+the structure of  cross-linking microfibrils in collagen and cellulose in biology,
+by modeling these structures as {\em pinned line incidence systems}.
 
-Collagen is an important protein material in biological tissues with highly elastic mechanical property. 
-Cellulose is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}). 
-Both of these substances consist of a large number microfibrils,
+{\em Collagen} is an important protein material in biological tissues with highly elastic mechanical property. 
+{\em Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}). 
+Both of these substances consist of a large number {\em microfibrils},
 %fibril itself straight?
@@
 each of which is cross-linked at 2 places with some two other fibrils, 
-where the cross-linking is like an incidence constraint that 
-the two crosslinked fibrils can slide with each other (see Figure~\ref{fig:cross-linking}(a). ).
+where the {\em  cross-linking} is like an incidence constraint that 
+the two crosslinked fibrils can slide with each other  \cite{xxx} (see Figure~\ref{fig:crosslink}).
 
@@
 
-\begin{figure}[hbtp]
+\begin{figure}
 \centering
-\includegraphics[width=0.5\linewidth]{img/AFM_Innventia_nanocellulose}
-\caption{Fibrils of carboxymethylated nanocellulose adsorbed on a silica surface. \cite{xxx}}
+
+\begin{subfigure}{.49\linewidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{img/AFM_Innventia_nanocellulose}
+  \caption{}
   \label{fig:cellulose}
+\end{subfigure}
+\begin{subfigure}{.49\linewidth}
+  \centering
+  \includegraphics[width=\linewidth]{img/crosslink}
+  \caption{}
+  \label{fig:crosslink}
+\end{subfigure}%
+\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface~\cite{xxx}. (b) Cross-linking of cellulose microfibrils~\cite{xxx}.} 
 \end{figure}
@@
 
-The cross-linking of microfibrils can be modeled using 2-dimensional pinned line incidence constraint system.
+The cross-linking  microfibrils can be modeled as a 2-dimensional pinned line incidence constraint system, 
+where incidence constraints are used instead of distance constraints.
 
-%TODO system? problem?
-\begin{definition} [2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
-A simple graph $G=(V,E)$ is a 2-dimensional pinned line incidence graph, 
-if there is  a  set $X$ of $|E|$ pins with fixed positions in $\mathbb{R}^2$.
-such that 
- every edge $e \in E$ corresponds to a pin $x \in X$ where $e$ is constrained to lie on a line passing through $x$.
+\begin{definition} %[2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
+A  graph $G=(V,E, w_V, w_E)$ is a \textbf{2-dimensional pinned line incidence graph}, 
+if  $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$, and
+there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
+such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where 
+for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
+such that $e$ is constrained to lie on a line passing through $x$.
 %There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
@@
 
+\begin{remark}
+For 2-dimensional pinned line incidence graphs, 
+the constant $k$ used in the definition of independence and rigidity is $0$  \cite{xxx} (see Definition~\ref{xxx}--\ref{xxx}).
+\end{remark}
+
 In the case of microfibril cross-linking, each fibril is 
-attached to some fixed larger organelle/membrane at one spot, 
-so it can be modeled as an edge/line segment incident at a pin in $X$. 
+attached to some fixed larger organelle/membrane at one spot.
+Consequently, each fibril can be modeled as an edge of the graph, 
+with the attachment being the corresponding pin in $X$. 
+The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge. 
+%so it can be modeled as an edge/line segment incident at a pin in $X$. 
 %For cross-linking density 2, 
 %i.e.\ each fibril has two cross-linkings with other fibrils, 
-The cross-linkings can be modeled as vertices in $V$.
+%The cross-linkings can be modeled as vertices in $V$.
+
+Figure~\ref{fig:pinned_line} shows an example of a 2-dimensional pinned line incidence graph, 
+where the grey ovals stand for pins representing attachments of fibrils, 
+and the vertices $a_1,a_2,\ldots, d_3$ represent cross-linkings.
+% and the black dots stand for vertices/cross-linkings.
+%There are  and 10 vertices/cross-linkings in the graph.
+The graph is well-constrained, with 12 vertices and 24 edges/pins.
 
@@
   \centering
-\begin{subfigure}{.5\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{img/crosslink}
-  \caption{}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\linewidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{img/teq1}
-  \caption{}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{(a) Cross-linking of microfibrils~\cite{xxx}. (b) A pinned line incidence graph.}
-\label{fig:cross-linking}
+  \includegraphics[width=.9\linewidth]{img/pinned}
+\caption{ A well-constrained pinned line incidence graph.}
+\label{fig:pinned_line}
 \end{figure}
-
-As an example, Figure~\ref{fig:cross-linking}, 
-where the grey ovals stand for pins/attachments of fibrils and the black dots stand for vertices/cross-linkings.
-There are 5 edges/fibrils,  and 10 vertices/cross-linkings in the graph.
 
@@
 
-In this section, we will adapt the results in Section~\ref{xxx}
+In this section, we will adapt the results in Section~\ref{dr}
 to give the Optimal DR-plan for 2-dimensional pinned line incidence graphs.
-%The 2-dimensional pinned line incidence graphs and their DR-plans have 
-Note the following differences from the bar-joint graphs:
+First, we note the following differences of pinned line incidence graphs from the bar-joint graphs.
 
@@
 
-\item For pinned line incidence graphs, all vertex weights are $2$, all edge weights are $1$, 
-and the constant $k = 0$ \cite{xxxx}. 
-%In other words, a pinned line incidence graph is well-constrained if $2|V| = |E|$ and $2|V'| \le |E'|$ for all induced non-trivial subgraphs $S=(V',E') \subseteq G$.
+\item For pinned line incidence graphs, 
+since the pins in $X$ have fixed positions on the plane, 
+o trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
+In particular, note that both a single vertex and a single edge are under-constrained graphs.
+
+%\todo{Over-constrained trivial subgraphs does not exist??? }\
+%\todo{Both a single vertex and a single edge are underconstrained}
+%\begin{itemize}
+%%\item For  pinned line incidence graphs:
+%%no trivial motion exists since the pins have fixed locations. 
+%\item 
+%\todo{We just define the %trivial graph/
+%DR-plan leaves for 2-dimensional pinned line incidence graphs to be single vertices...  }
+%\item \todo{A single edge can be a node of the DR-plan}
+%\end{itemize}
 
 \item 
-\todo{Trivial graphs (must be overconstrained in the original definition) does not exist??? }\
-\todo{Both a single vertex and a single edge are underconstrained}
-\begin{itemize}
-
-\item For  pinned line incidence graphs:
-no trivial motion exists since the pins have fixed locations.
-
-\item 
-\todo{We just define the trivial graph/DR-plan leaf for 2-dimensional pinned line incidence graphs to be a single vertex.  }
-
-\item \todo{A single edge can be a node of the DR-plan}
+For pinned line incidence graphs, 
+since the pins in $X$ have fixed positions on the plane, 
+{\em a well-constrained vertex-maximal proper subgraph can be disconnected}, 
+being the disjoint union of two or more well-constrained subgraphs. \\
+%Without loss of generality, 
+%we slightly modify the requirement of DR-plan for pinned line incidence graphs
+%such that only connected subgraphs can be considered as a node of the DR-plan.
 
@@
 
-\item 
-Notice that for pinned line incidence graphs, 
-since the pins in $X$ have fixed positions on the plane, 
-a well-constrained vertex-maximal proper subgraphs can be disconnected.\\
-Without loss of generality, 
-we slightly modify the requirement of DR-plan for pinned line incidence graphs
-such that only connected subgraphs can be considered as a node of the DR-plan.
+We make  a few corresponding modifications to the definition of the DR-plan:
 
-\end{itemize}
+%\begin{itemize}
+%\item For any $C$, its $N$ children, $C_1, \ldots, C_N$, are {\em edges} or {\em connected} well-constrained vertex-maximal proper subgraphs of that node $C$.
+%\item A node containing a single edge is a leaf.
+%%\item 
+%%\todo{We just define the %trivial graph/
+%%DR-plan leaves for 2-dimensional pinned line incidence graphs to be single vertices...  }
+%%\item \todo{A single edge can be a node of the DR-plan}
+%%\item only connected subgraphs can be considered as a node of the DR-plan.
+%\end{itemize}
+
+\begin{definition}
+
+ The \textbf{DR-plan} of  a 2-dimensional pinned line incidence graph $G$ is Definition~\ref{def:drp} with  modified rules number 2 and 4:
+%    The \textbf{decomposition-recombination (DR-) plan} of  a 2-dimensional pinned line incidence graph $G$, $DRP(G)$, is  a tree that has the following properties:
+    \begin{enumerate}
+        %\item The root of the tree `contains' $G$.
+        \item[2] For any node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
+        \item[4] A node containing a single edge is a leaf.
+    \end{enumerate}
+\end{definition}
 
 With the above modification, 
-it is not hard to adapt Theorem~\ref{t1} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
-to show the structure of optimal DR-plan for 2-dimensional pinned line incidence graphs.
+we now adapt Theorem~\ref{t1} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
+for the optimal DR-plan for 2-dimensional pinned line incidence graphs.
 
@@
 \begin{enumerate}
-    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
+    \item if $C_i \cap C_j$ is a single vertex then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
     \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
@@
 
-\todo{All Lemmas before Lemma 8 are proved with respect to general constraint systems,
-and they go through for 2-dimensional pinned line incidence as well. }
-Now we  show that Lemma 8 also holds for 2-dimensional pinned line incidence graphs.
+The following lemma serves a similar function to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection} 
+%Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection}, 
+and can be proved using a similar counting based argument.
+
+\begin{lemma}
+\label{lem:pinned_union_intersection}
+Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$, 
+where each of them can be either  a single edge or a connected well-constrained subgraph. 
+There are only two possible cases: 
+\begin{itemize}
+\item At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is a vertex. 
+\item Both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained. 
+\end{itemize}
+
+%The following table summarizes all bijections: 
+%\setlength{\tabcolsep}{3pt}
+%{\small
+%\begin{center}
+%\begin{tabular}{|p{2.8cm}|c|p{2.3cm}|}
+%\hline
+%\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
+% \centering--- &  \sout{over-constrained}          &  \centering --- \tabularnewline \hline
+%at least one of $F_i,F_j$ is an edge & under-constrained       &  \centering a vertex (under-constrained)  \tabularnewline \hline
+%both $F_i$ and $F_j$ are well-constrained & well-constrained        &  \centering well-constrained  \tabularnewline \hline
+% \centering --- & ---                     &  \centering \sout{under-constrained} \tabularnewline \hline
+%\end{tabular}
+%\end{center}
+%}
+%\setlength{\tabcolsep}{6pt}
+
+\end{lemma}
+
+%The proof of Lemma~\ref{lem:pinned_union_intersection} is similar to the proofs of Lemma~\ref{t-l1}--\ref{iuc-l1}.
+
+%\begin{lemma}
+%Let $F_j$ and $F_j$ be well-constrained subgraphs of the same well-constrained graph $F$. 
+%Both $F_i \cup F_j$ and $F_i \cap F_j$ can only be well-constrained.
+%\end{lemma}
+%
+%\begin{proof}
+%\begin{enumerate}
+%\item Obviously $F_i \cup F_j$ cannot be over-constrained since it is a subgraph of well-constrained graph $F$.
+%\item $F_i \cup F_j$ cannot be under-constrained, otherwise $F_i \cap F_j$ would be over-constrained while there are no trivial over-constrained graphs. 
+%\item The only remaining case is that $F_i \cup F_j$ is well-constrained, which is true if and only if $F_i \cap F_j$ is well-constrained.
+%\end{enumerate}
+%\end{proof}
+
+Proofs of Lemma~\ref{wc-l2} and \ref{wc-c1} straightforwardly extend to 2-dimensional pinned line incidence graphs 
+given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan.
+Now we  show that Lemma~\ref{wc-l3} also holds.
 
 \begin{proof}[Proof of Lemma~\ref{wc-l3} for 2D pinned line incidence graphs]
-We use the same notations as in the proof of Lemma~\ref{wc-c1} in Section~\ref{xxx}.
+We use the same notations as in the original proof of Lemma~\ref{wc-l3} in Section~\ref{sec:infer_parent}.
+
+First notice that since $C_i \cup C_j$ is well-constrained, by Lemma~\ref{lem:pinned_union_intersection}, 
+both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.
+Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a well-constrained subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected well-constrained vertex-maximal proper subgraph of  $C$.
+
+Then we analyze all the possible cases for $C_k$.
 \begin{itemize}
-    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary~\ref{wc-c1}) and by our modified requirement of DR-plan nodes.
-
-%    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
+%    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary~\ref{wc-c1}) and by our modified requirement of DR-plan nodes.
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
 Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained.
-We also know it cannot be trivial because it contains well-constrained subgraphs.
+    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained, 
+    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are well-constrained.
+%    We also know it cannot be trivial because it contains well-constrained subgraphs.
 This means it must be under-constrained.
@@
 This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
-%
-%    \item 1 case: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible.
-The union with $C_i$ (or $C_j$) is not equal to $C$.
+
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.  
+%    union with $C_i$ (or $C_j$) is not equal to $C$.
 Therefore, the union must be under-constrained which means the intersection must be trivial (a single node).
 However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.
-%
-%    \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
+
+    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible. 
+    The proof is similar to the previous case.  
+    %The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
 \end{itemize}
 
-\todo{All other cases are similar to the original proof.}
+All the remaining cases are similar to the original proof.
 
@@
 
-Given Lemma~\ref{wc-l3}, Lemma~\ref{uc-l2}, Theorem~\ref{uc-t1} and \ref{wc-t1} also apply to  2-dimensional pinned line incidence graphs,
-proving Corollary~\ref{cor:pinned}.
+The following lemma serves as Lemma~\ref{uc-l2} for pinned line incidence and can be proved similarly.
+\begin{lemma}
+\label{lem:vertex_intersection}
+If $C_i\cap C_j$ is a single vertex, then $\forall k: C_i\cap C_k$ is a single vertex.
+\end{lemma}
 
-Therefore, we can analyze microfibril cross-linkings by finding the optimal DR-plan for  2-dimensional pinned line incidence graphs in the same way as for 2-dimensional bar-joint graphs.
+Given Lemma~\ref{wc-l3} and Lemma~\ref{lem:vertex_intersection}, 
+it is straightforward to adapt the proof of Theorem~\ref{uc-t1} and \ref{wc-t1} 
+to  2-dimensional pinned line incidence graphs,
+thus proving Corollary~\ref{cor:pinned}. 
+Consequently, we can efficiently find the optimal DR-plan for 2-dimensional pinned line incidence graphs using basically the same algorithm as for 2-dimensional bar-joint graphs.
+%we can analyze microfibril cross-linkings by finding the optimal DR-plan for  2-dimensional pinned line incidence graphs in the same way as for 2-dimensional bar-joint graphs.
+
+%TODO ???
+The recombination problem for 2-dimensional pinned line incidence systems is trivial. 
+Since the pins are given fixed positions in the plane, 
+once we get a solution a well-constrained sub-system, 
+the solution will automatically be consistent with the remaining of the system.

### 2014-12-15 doc/GMP2015/appendix_a.tex
Major changes to my section. Meera's major changes to her sections. Updated labels and their references.

\section{Appendix}
\label{sec:appendix}

\subsection{Section \ref{sec:DRP} -- {\sffamily \textbf{Lemma \ref{lemma:wc_intersection_makes_all_wc}}}}

% \textbf{Section \ref{sec:DRP}}

% {\sffamily \textbf{Lemma \ref{lemma:main_trivial}}

\begin{proof}
The alternative phrasing is due to Lemma \ref{lemma:wc_intersection_is_C}.

Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$). Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.

Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

For convenience, it is implied that all graphs are the induced graph on $C$.
\newcommand{\inducedOnC}[1]{#1}

% \newcommand{\inducedOnC}[1]{Idc\left(C,#1\right)}

\begin{itemize}
    \item 3 cases: $C_k$ cannot be $C=\inducedOnC{R_i\cup R_j\cup D_{i,j}}$, $C_i=\inducedOnC{R_j\cup D_{i,j}}$, or $C_j=\inducedOnC{R_i\cup D_{i,j}}$. This is by definition.

    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal. These are the graphs $\inducedOnC{R'_i\cup D_{i,j}}$, $\inducedOnC{R'_j\cup D_{i,j}}$, $\inducedOnC{ D_{i,j}}$, $\inducedOnC{R_i\cup D'_{i,j}}$, $\inducedOnC{R_j\cup D'_{i,j}}$, $\inducedOnC{R'_i\cup D'_{i,j}}$, $\inducedOnC{R'_j\cup D'_{i,j}}$, $\inducedOnC{ D'_{i,j}}$, $\inducedOnC{R_i}$, $\inducedOnC{R_j}$, $\inducedOnC{R'_i}$, $\inducedOnC{R'_j}$, and $\inducedOnC{\emptyset}$.

    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal. These are the graphs $\inducedOnC{R'_i\cup R_j\cup D_{i,j}}$ and $\inducedOnC{R_i\cup R'_j\cup D_{i,j}}$ respectively.

    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.

    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be well-constrained. We also know it cannot be trivial because it contains well-constrained subgraphs. This means it must be under-constrained. From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial. This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.

    \item 1 case: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible. Since $C_i\cup C_k\neq C$ (and $C_j\cup C_k\neq C$), we know by the same logic as the previous case that the $C_i\cap C_k$ must be trivial (a single node). However, $C_i\cap C_k=\inducedOnC{R'_j\cup D'_{i,j}}$. This causes a contradiction, the intersection cannot be trivial because $R'_j$ and $D'_{i,j}$ are not empty sets and are disjoint.

    \item 2 cases: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible. The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.

    \item 1 case: $C_k=\inducedOnC{R_i\cup R_j\cup D'_{i,j}}$ is all that remains.
\end{itemize}

Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.

\todo{Need a figure?}
\end{proof}

### 2014-12-15 doc/GMP2015/s01_introduction.tex
Major changes to my section. Meera's major changes to her sections. Updated labels and their references.

@@
 \section{Introduction}
-\label{sec:prev}
+\label{sec:intro}
 
@@
 {\sl isostatic or underconstrained, not
-self-stressed} (See Section \ref{prelim} for definitions).
+self-stressed} (See Section \ref{sec:prelim} for definitions).
 These properties are
@@
 of which are {\it bar-joint} systems (Examples 1,2 above, see Sections
-\ref{dr}, \ref{recomb}),
-{\it multibody-pin} systems (Example 4,5, see Section \ref{app}) or  {\it
-pinned-line incidence} systems (Example 3, see Section \ref{app}). The
+\ref{sec:DRP}, \ref{sec:recomb}),
+{\it multibody-pin} systems (Example 4,5, see Section \ref{sec:bodypin}) or  {\it
+pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The
 limitations of these techniques directly motivate
@@
 Space Navigation}}
-For fairly general 3D constraint systems, there are algorithms with formal
-guarantees that
-solve the parent (sub)systems that occur in the DR-plan \cite{XX, XX,XX},
-provided the DR-plan meets some of the above criteria.
-These algorithms  uncover underlying matroids to combinatorially optimize
-the algebraic complexity of (re)combining child system solutions,
-and algebraically solve the optimized recombination system to give a
+For the entire DR-plan, finding all
+desired solutions is barely tractable even if recombination
+of solved subsystems is easy for each indecomposable parent system in the
+DR-plan.
+This is because even for the simplest, highly decomposable systems
+with small DR-plans, the problem of finding even a single
+solution to the input system at the root of the DR-plan
+is NP-hard \cite{XX} and there is a
+combinatorial explosion of solutions \cite{XX}. Typically, however,
+the desired solution has a given orientation type, in which case,
+the crux of the difficulty is concentrated in the algebraic complexity of
+(re)combining child system solutions, to give a
 solution to the parent system at any
 given level of the DR-plan.
+For fairly general 3D constraint systems,
+there are algorithms with formal
+guarantees that uncover underlying matroids to combinatorially
+obtain an optimal parameterization to minimize the algebraic
+complexity of the indecomposable parent (sub)systems that occur in the
+DR-plan \cite{XX, XX,XX},
+provided the DR-plan meets some of the abovementioned criteria.
 
@@
 the best algorithms can still take
-exponential time in the number of child subsystems in order to guarantee
-all solutions even for a single (sub)system in a DR-plan. They
-are prohibitively slow in practice. For the entire DR-plan, finding all
-solutions is barely tractable even with fast recombination algorithms for
-the subsystems in the DR-plan. This is because even for the simplest
-systems, the problem is NP-hard \cite{XX} and there is a
-combinatorial explosion of solutions \cite{XX}.  Using the DR-plan and
-optimal recombination as a principled basis, high performance software
+exponential time in the number of child subsystems (which can be arbitrarily
+large even for optimal DR-plans) in order to guarantee
+all solutions of a given orientation type,
+even for a single (sub)system in a DR-plan. They
+are prohibitively slow in practice. We note that utilizing the DR-plan and
+optimal recombination as a principled basis, high performance heuristics
+and software
 exists \cite{XX,XX} to
@@
 underconstrained
-2D qusecs are optimal} (Section \ref{dr}).
+2D qusecs are optimal} (Section \ref{sec:DRP}).
 \item
@@
 optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections
-\ref{dr} and \ref{app}.
+\ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}.
 The canonical DR-plan
@@
 \item
-We show reductions between different well-known problems related to
-realization or solution by
-recombining the realizations or solutions of subsystems (at any given level
-of the DR-plan). Example problems include optimal completion of
-underconstrained  systems  \cite{XX}, and optimal parametrization in
-isostatic systems \cite{XX}). When completions and parametrizations are
-given from certain well-known classes, we obtain new,
-efficient algorithms for realizing both isostatic and underconstrained
+We give a method to deal with the
+algebraic complexity of recombining the realizations or
+solutions of  child subsystems into a solution of the parent system.
+Specifically, we define the problem of  minimally modifying the
+indecomposable recombination system so that it becomes decomposable via a
+ small DR-plan and yet preserves the original
+solutions in an efficiently searchable manner. We show formal connection
+to well known problems such as optimal completion of
+underconstrained  systems  \cite{XX}.  When the required modifications
+are bounded, we obtain new, efficient algorithms for
+realizing both isostatic and underconstrained
 qusecs by leveraging recent results about Cayley parameters in
-\cite{XX,XX}  (Section \ref{recomb}).
+\cite{XX,XX}  (Section \ref{sec:recomb}).
 \item

### 2014-12-15 doc/GMP2015/s02_preliminaries.tex
Major changes to my section. Meera's major changes to her sections. Updated labels and their references.

@@
-\section{Preliminaries}\label{prelim}
+\section{Preliminaries}
+\label{sec:prelim}
 
@@
 
-\begin{definition}
-   In the context of this paper, a \textbf{graph} $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
+In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
+$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
 Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
-\end{definition}
+The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
 
-\begin{definition}
-    The \textbf{density} of graph $G$ is
-    \begin{equation}
-        d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}
-    \end{equation}
-\end{definition}
+Given a constant $k$, a graph $G$ is:
+\begin{itemize}
+    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
 
-\begin{definition}
-    A graph is \textbf{independent} (or \textbf{sparse}) if, given constant $k$, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
-\end{definition}
+    \item \textbf{over-constrained} if it is not independent.
+
+    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
+
+    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+
+    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+
+    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
+\end{itemize}
+
+\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.
 
@@
 
-\begin{definition}
-    A graph is \textbf{over-constrained} if it is not independent.
-\end{definition}
-
 % \begin{definition}
@@
 
-\begin{definition}
-    A graph is \textbf{well-constrained} (or \textbf{tight}) if, given constant $k$, $G$ is independent and $d(G)=k$.
-\end{definition}
-
-\begin{definition}
-    A graph is \textbf{rigid} if, given constant $k$, there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-\end{definition}
-
-\begin{definition}
-    A graph is \textbf{under-constrained} if $G$ is independent and not rigid.
-\end{definition}
-
-\begin{definition}
-    A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
-    \todo{Maybe leave the following out until it's needed later?}
-    In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
+% \begin{definition}
+%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
+%     \todo{Maybe leave the following out until it's needed later?}
+%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
 Trivial graphs for 2D would be a single vertex.
@@
 These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
-\end{definition}
+% \end{definition}
 
-\begin{definition}
-    The \textbf{decomposition-recombination (DR-) plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
+\begin{definition}\label{def:drp}
+    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
     \begin{enumerate}
@@
     Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
+
+    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+This makes rule 3 implicit.
+The complete DR-plan is unique.
+
+    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
 \end{definition}
 
-\begin{definition}
-    The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+% \begin{definition}
+%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
 By this, rule 3 is implicit for a complete DR-plan.
-\end{definition}
+% \end{definition}
 
-\begin{definition}
-    The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-\end{definition}
+% \begin{definition}
+%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+% \end{definition}
 
@@
 
-$G$ is taken to be a graph with values $(V,E,w_V,w_E)$. The most important limitation of this paper is that we take $G$ to be well-constrained.
-\todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
+In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
+% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
 It is not useful to solve for two bodies simultaneously if there are no constraints between them.
@@
 
-$C$ is an arbitrary graph in a node in $CompleteDRP(G)$.
+$C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
 $C_i$ is the $i^{\text{th}}$ child of $C$.

### 2014-12-15 doc/GMP2015/s07_problempinning.tex
Major changes to my section. Meera's major changes to her sections. Updated labels and their references.

@@
 \section{Problem Pinning}
+\label{sec:bodypin}
 
@@
 \begin{remark}
-    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying. As such, the DR planning discussed in Sections \ref{dr} and \ref{recomb} will still go through.
+    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying. As such, the DR planning discussed in Sections \ref{sec:DRP} and \ref{sec:recomb} will still go through.
 \end{remark}
@@
 
-In this section we will go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{intro}.
+In this section we will go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
 
@@
 
-Body-pin qusecs are of particular interest to us in the context of Example 4 in \ref{intro}.
-Multi-triangle-pin qusec can be used to reprsent the silica bi-layers described in Example 5 of \ref{intro}, where each triangle is the junction of ``disks''.
+Body-pin qusecs are of particular interest to us in the context of Example 4 in \ref{sec:intro}.
+Multi-triangle-pin qusec can be used to reprsent the silica bi-layers described in Example 5 of \ref{sec:intro}, where each triangle is the junction of ``disks''.
 Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR plan:
@@
 
-We can develop the notion of a canonical DR plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{dr}.
+We can develop the notion of a canonical DR plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{sec:DRP}.
 
@@
 \begin{remark}
-    All $(k,l)$ tight canonical DR plans are optimal. We can find such a DR plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{dr}.
+    All $(k,l)$ tight canonical DR plans are optimal. We can find such a DR plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{sec:DRP}.
 \end{remark}
@@
 \begin{theorem}
-    There is a quadratic algorithm for the 1 dof and 2 dof optimal completion problem of Seciton \ref{recomb} for body-pin and multi-triangle-pin qusecs.
+    There is a quadratic algorithm for the 1 dof and 2 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
 \end{theorem}

### 2014-12-15 doc/GMP2015/s08_pinnedline.tex
Major changes to my section. Meera's major changes to her sections. Updated labels and their references.

@@
 \section{Application: finding optimal DR-plans and realizations for cross-linking microfibrils }
+\label{sec:pinnedline}
 
@@
   \centering
-  \includegraphics[width=.9\linewidth]{img/pinned}
+  % \includegraphics[width=.9\linewidth]{img/pinned}
 \caption{ A well-constrained pinned line incidence graph.}
@@
 
-In this section, we will adapt the results in Section~\ref{dr}
+In this section, we will adapt the results in Section~\ref{sec:DRP}
 to give the Optimal DR-plan for 2-dimensional pinned line incidence graphs.
@@
 With the above modification,
-we now adapt Theorem~\ref{t1} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
+we now adapt Theorem~\ref{theorem:main} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
 for the optimal DR-plan for 2-dimensional pinned line incidence graphs.
@@
 
-The following lemma serves a similar function to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection} 
+The following lemma serves a similar function to Lemma~\ref{lemma:union_intersection} in Section~\ref{sec:union_intersection}
 %Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
@@
 
-Proofs of Lemma~\ref{wc-l2} and \ref{wc-c1} straightforwardly extend to 2-dimensional pinned line incidence graphs 
+Proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to 2-dimensional pinned line incidence graphs
 given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan.
-Now we  show that Lemma~\ref{wc-l3} also holds.
+Now we  show that Lemma~\ref{lemma:wc_intersection_makes_all_wc} also holds.
 
-\begin{proof}[Proof of Lemma~\ref{wc-l3} for 2D pinned line incidence graphs]
-We use the same notations as in the original proof of Lemma~\ref{wc-l3} in Section~\ref{sec:infer_parent}.
+\begin{proof}[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
+We use the same notations as in the original proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} in Section~\ref{sec:infer_parent}.
 
@@
     \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained.
+Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be well-constrained.
     By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
@@
 
-The following lemma serves as Lemma~\ref{uc-l2} for pinned line incidence and can be proved similarly.
+The following lemma serves as Lemma~\ref{lemma:uc_intersection_makes_all_uc} for pinned line incidence and can be proved similarly.
 \begin{lemma}
@@
 
-Given Lemma~\ref{wc-l3} and Lemma~\ref{lem:vertex_intersection}, 
-it is straightforward to adapt the proof of Theorem~\ref{uc-t1} and \ref{wc-t1} 
+Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lem:vertex_intersection},
+it is straightforward to adapt the proof of Theorem~\ref{theorem:main_trivial} and \ref{theorem:main_wellconstrained}
 to  2-dimensional pinned line incidence graphs,

### 2014-12-15 doc/GMP2015/s08_pinnedline.tex
Fixed image and moved proof of lemma to appendix.

@@
 
-\begin{definition} %[2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
-A  graph $G=(V,E, w_V, w_E)$ is a \textbf{2-dimensional pinned line incidence graph},
-if  $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$, and
-there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
+%[2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
+A \textbf{2-dimensional pinned line incidence graph}  $G=(V,E, w_V, w_E)$ 
+%is a \textbf{2-dimensional pinned line incidence graph},
+ has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
+Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
 such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
@@
 %that satisfies the given subspace incidences. % specified by $\supp[P]{x}$.%$\{S_i\}$. % on $P[X]$ and $X$.
-\end{definition}
 
@@
 For 2-dimensional pinned line incidence graphs,
-the constant $k$ used in the definition of independence and rigidity is $0$  \cite{xxx} (see Definition~\ref{xxx}--\ref{xxx}).
+the constant $k$ used in the definition of independence and rigidity is $0$  \cite{xxx} (see Section~\ref{sec:prelim_defs}).
 \end{remark}
@@
 The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge.
-%so it can be modeled as an edge/line segment incident at a pin in $X$.
-%For cross-linking density 2,
-%i.e.\ each fibril has two cross-linkings with other fibrils,
-%The cross-linkings can be modeled as vertices in $V$.
 
@@
   \centering
-  % \includegraphics[width=.9\linewidth]{img/pinned}
-\caption{ A well-constrained pinned line incidence graph.}
+   \includegraphics[width=.9\linewidth]{img/pinned}
+\caption{ A well-constrained 2-dimensional pinned line incidence graph.}
 \label{fig:pinned_line}
@@
         %\item The root of the tree `contains' $G$.
-        \item[2] For any node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
         %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
@@
 \end{itemize}
-
 %The following table summarizes all bijections:
@@
 %\setlength{\tabcolsep}{6pt}
-
 \end{lemma}
 
-%The proof of Lemma~\ref{lem:pinned_union_intersection} is similar to the proofs of Lemma~\ref{t-l1}--\ref{iuc-l1}.
-
-%\begin{lemma}
-%Let $F_j$ and $F_j$ be well-constrained subgraphs of the same well-constrained graph $F$.
-%Both $F_i \cup F_j$ and $F_i \cap F_j$ can only be well-constrained.
-%\end{lemma}
-%
-%\begin{proof}
-%\begin{enumerate}
-%\item Obviously $F_i \cup F_j$ cannot be over-constrained since it is a subgraph of well-constrained graph $F$.
-%\item $F_i \cup F_j$ cannot be under-constrained, otherwise $F_i \cap F_j$ would be over-constrained while there are no trivial over-constrained graphs.
-%\item The only remaining case is that $F_i \cup F_j$ is well-constrained, which is true if and only if $F_i \cap F_j$ is well-constrained.
-%\end{enumerate}
-%\end{proof}
-
-Proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to 2-dimensional pinned line incidence graphs
-given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan.
-Now we  show that Lemma~\ref{lemma:wc_intersection_makes_all_wc} also holds.
-
-\begin{proof}[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
-We use the same notations as in the original proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} in Section~\ref{sec:infer_parent}.
-
-First notice that since $C_i \cup C_j$ is well-constrained, by Lemma~\ref{lem:pinned_union_intersection},
-both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.
-Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a well-constrained subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected well-constrained vertex-maximal proper subgraph of  $C$.
-
-Then we analyze all the possible cases for $C_k$.
-\begin{itemize}
-%    \item 4 cases: $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary~\ref{wc-c1}) and by our modified requirement of DR-plan nodes.
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be well-constrained.
-    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
-    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are well-constrained.
-%    We also know it cannot be trivial because it contains well-constrained subgraphs.
-This means it must be under-constrained.
-From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial.
-This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
-
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.
-%    union with $C_i$ (or $C_j$) is not equal to $C$.
-Therefore, the union must be under-constrained which means the intersection must be trivial (a single node).
-However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.
-
-    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-    The proof is similar to the previous case.
-    %The proof mirrors the previous case, except you must choose $C_i$ and $C_j$, respectively, here.
-\end{itemize}
-
-All the remaining cases are similar to the original proof.
-
-\end{proof}
-
-%How theorems go through when only considering connected rigidity components
-
+Given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan,
+proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to 2-dimensional pinned line incidence graphs.
+The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2-dimensional pinned line incidence graphs 
+is given in Appendix~\ref{sec:appendix_pinned}.
 The following lemma serves as Lemma~\ref{lemma:uc_intersection_makes_all_uc} for pinned line incidence and can be proved similarly.

### 2014-12-16 doc/GMP2015/s07_problempinning.tex
Added figures into the problem pinning section

--- 2014-12-15 doc/GMP2015/s07_problempinning.tex
@@
 To make this thing isostatic, there are a few methods we can apply.
+
+\begin{figure}\centering
+    \includegraphics[width=0.5\linewidth]{img/Silica}
+    \caption{Example of 1 layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.}
+\end{figure}
 
@@
 \begin{remark}
-    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying. As such, the DR planning discussed in Sections \ref{dr} and \ref{recomb} will still go through.
+    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying.
+As such, the DR planning discussed in Section \ref{dr} is unchanged and the work in Section \ref{recomb} will still go through with slight modifications.
 \end{remark}
@@
 A body with 2 pins can be replaced by an edge.
-A body with 3 pins can be replaced by a triangle.
 In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
 When looking for a DR plan, we treat each body as trivial, so they become the leaves of our plan.
+The optimal completion problem and approach of Section \ref{dr} are unchanged.
+The optimal paramaterization problem in Section \ref{dr} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
@@
 \begin{definition}
+\label{def:body-pin}
     A body-pin qusec is a multi-body-pin qusec with the following conditions:
@@
 
-Body-pin qusecs are of particular interest to us in the context of Example 4 in \ref{intro}.
-Multi-triangle-pin qusec can be used to reprsent the silica bi-layers described in Example 5 of \ref{intro}, where each triangle is the junction of ``disks''.
+Body-pin qusecs are of particular interest to us in the context of Example 4 in Section \ref{intro}.
+Multi-triangle-pin qusec can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{intro}, where each triangle is the junction of ``disks'' in the plane.
 Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR plan:
@@
 \end{remark}
+\begin{figure*}\centering
+\begin{subfigure}{0.2\linewidth}\centering
+    \includegraphics[width=\linewidth]{img/bodypin}
+    \caption{}
+\end{subfigure}
+\begin{subfigure}{0.2\linewidth}\centering
+    \begin{tikzpicture}[scale=0.8]
+        % \tikzstyle{v}=[draw, circle, minimum size=0.1cm, scale=0.7, font=\footnotesize]
+        \tikzstyle{v}=[font=\footnotesize]
+
+        \node[v] (a) at (0,0) {A};
+        \node[v] (b) at (0,1) {B};
+        \node[v] (c) at (1,0) {C};
+        \node[v] (d) at (1,1) {D};
+        \node[v] (e) at (0,2) {E};
+        \node[v] (f) at (1,2) {F};
+        \node[v] (g) at (-1,0) {G};
+        \node[v] (h) at (-1,1) {H};
+
+        \draw (a) to [bend right=10](b);
+        \draw (a) to [bend left=10](b);
+        \draw (a) to [bend right=10](c);
+        \draw (a) to [bend left=10](c);
+        \draw (d) to [bend right=10](c);
+        \draw (d) to [bend left=10](c);
+        \draw (b) to [bend right=10](d);
+        \draw (b) to [bend left=10](d);
+        \draw (b) to [bend right=10](e);
+        \draw (b) to [bend left=10](e);
+        \draw (b) to [bend right=10](h);
+        \draw (b) to [bend left=10](h);
+        \draw (g) to [bend right=10](h);
+        \draw (g) to [bend left=10](h);
+        \draw (g) to [bend right=10](a);
+        \draw (g) to [bend left=10](a);
+        \draw (f) to [bend right=10](e);
+        \draw (f) to [bend left=10](e);
+        \draw (f) to [bend right=10](d);
+        \draw (f) to [bend left=10](d);
+    \end{tikzpicture}
+    \caption{}
+\end{subfigure}
+\begin{subfigure}{0.3\linewidth}\centering\scriptsize
+    \begin{tikzpicture}
+    [level 1/.style={sibling distance=8em},
+     level 3/.style={sibling distance=1em}, level distance=0.75cm]
+        \tikzstyle{n}=[draw, fill, circle]
+
+        \node (a) at (0,0) {A,B,C,D,E,F,G,H}
+            child { node {A,B,C,D,E,F} 
+                child { node {C,D,E,F}
+                    child {node { }}
+                    child {node {...}}
+                    child {node { }}
+                    child {node { }}
+                    } 
+                child { node (shared) {A,B,C,D}
+                    child {node {A}}
+                    child {node {B}}
+                    child {node {C}}
+                    child {node {D}}
+                    } 
+                }
+            child { node {A,B,C,D,G,H} 
+                child { node { } }
+                child { node {{\bf A,B,G,H}} 
+                    child {node { }}
+                    child {node {...}}
+                    child {node { }}
+                    child {node { }}
+                }
+            };
+
+    \end{tikzpicture}
+    \caption{}
+\end{subfigure}
+\begin{subfigure}{0.2\linewidth}\centering
+    \includegraphics[width=\linewidth]{img/bodypin2}
+    \caption{}
+\end{subfigure}
+
+% \begin{subfigure}
+
+% \end{subfigure}
+\caption{(a) is an orignal body-pin qusec.
+(b) is the corresponding body-bar qusec explained in Definition \ref{def:body-pin}.
+(c) is the 1 dof DR plan for the qusec.
+In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the node in the second level.
+(d) is the result of adding one such body tot he bold-faced node.}
+
+\end{figure*}
 

### 2014-12-16 doc/GMP2015/appendix_a.tex

@@
 \end{proof}
+
+\subsection{Section \ref{sec:pinnedline} -- proof of {\sffamily \textbf{Lemma \ref{lemma:wc_intersection_makes_all_wc}}} for  2-dimensional pinned line incidence graphs }
+\label{sec:appendix_pinned}
+
+\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
+We use the same notations as in the original proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} in the last section.
+
+First notice that since $C_i \cup C_j$ is well-constrained, by Lemma~\ref{lem:pinned_union_intersection},
+both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.
+Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a well-constrained subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected well-constrained vertex-maximal proper subgraph of  $C$.
+
+Then we analyze all the possible cases for $C_k$.
+\begin{itemize}
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
+Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be well-constrained.
+    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
+    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are well-constrained.
+
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.
+
+    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
+    The proof is similar to the previous case.
+\end{itemize}
+
+All remaining cases are similar to the original proof.
+\end{proof}

### 2014-12-16 doc/GMP2015/s00_abstract.tex
Brushed up my sections. Added new figures.

@@
 \begin{abstract}
-A variety of quasi-uniform (aperiodic) and possibly self-similar, layered material structures are natural consequences of the design objectives of rigidity and flexibility, minimizing mass, optimally distributing external stresses, and participating in assembly of diverse and multifunctional, larger structures.
+A variety of quasi-uniform (aperiodic) and possibly self-similar, layered material structures are a natural consequence of the design objectives of rigidity and flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
 
-Crucial for design and analysis of these properties is the recursive decomposition of diverse types of underlying isostatic or underconstrained 2D geometric constraint systems (which we collectively call \textit{qusecs}).
-While the optimal recursive decomposition (\textit{DR-planning}) problem is NP-hard even for 2D bar-joint constraint systems, we describe an $O(n^2)$ algorithm  for qusecs.
-Besides elucidating the source of NP-hardness, the algorithm outputs \textit{canonical} DR-plans that meet previously studied desirable criteria.
-When a DR-plan is input, we show reductions between problems related to realizing qusecs,  leading to efficient algorithms that leverage recent results on Cayley configuration spaces.
-We briefly discuss two specific 2D layered materials as applications of the above algorithms, whose software implementation is available.
-
-Keywords: rigidity, geometric constraint solving, configuration spaces, self-similar structures, layered materials
+Crucial for the design and analysis of the abovementioned properties is the recursive decomposition of diverse types of underlying 2D geometric constraint systems that are isostatic or underconstrained (which we call collectively as {\it qusecs}).
+While the optimal recursive decomposition ({\it DR-planning}) problem is NP-hard even for 2D bar-joint constraint systems, we describe an $O(n^2)$ algorithm  for qusecs.
+Besides elucidating the source of NP-hardness, the algorithm outputs {\it canonical} DR-plans that meet various desirable, previously studied criteria.
+By leveraging recent results on Cayley configuration spaces, we show how to decompose the indecomposable algebraic systems that recombine the solutions to child subsystems at a node of an optimal DR-plan: we minimally modify the system to have a small DR-plan, leading to efficient realization algorithms.
+We show formal connections to well-known problems such as completion of underconstrained systems.
+We briefly discuss the modeling and analysis of two specific 2D layered materials as applications of the above algorithms, whose software implementation is available.
 
@@
 % Software implementation and videos can be found at: \todo{Provide Link}.
+
+% A variety of quasi-uniform (aperiodic) and possibly self-similar, layered material structures are natural consequences of the design objectives of rigidity and flexibility, minimizing mass, optimally distributing external stresses, and participating in assembly of diverse and multifunctional, larger structures.
+
+% Crucial for design and analysis of these properties is the recursive decomposition of diverse types of underlying isostatic or underconstrained 2D geometric constraint systems (which we collectively call \textit{qusecs}).
+While the optimal recursive decomposition (\textit{DR-planning}) problem is NP-hard even for 2D bar-joint constraint systems, we describe an $O(n^2)$ algorithm  for qusecs.
+Besides elucidating the source of NP-hardness, the algorithm outputs \textit{canonical} DR-plans that meet previously studied desirable criteria.
+When a DR-plan is input, we show reductions between problems related to realizing qusecs,  leading to efficient algorithms that leverage recent results on Cayley configuration spaces.
+We briefly discuss two specific 2D layered materials as applications of the above algorithms, whose software implementation is available.
+
+% \textit{Keywords: rigidity, geometric constraint solving, configuration spaces, self-similar structures, layered materials}
+
+% % =====================CUT HERE ==============
+
+% % Examples include:
+% % cellulose and collagen microfibrils in cell walls;
+% % actin cytoskeleton matrix; hierarchical bone cross-section;
+% % ciliary membranes and transitions;
+% % self-similar DNA self-assemblies; silica bilayers;
+% % disordered graphene; etc.
+
+% % By  defining a canonical recursive decomposition
+% % ({\it DR plan}),  we show that the optimal DR-planning
+% % problem for
+
+% % We  navigate the NP-hardness of finding
+% % optimal DR-plans by defining a
+% % so-called {\it canonical} DR-plan and showing a strong Church-Rosser
+% % property: {\it all canonical DR-plans for isostatic or underconstrained
+% % 2D qusecs are optimal}.
+% % \item
+% % We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence
+% % optimal) DR-plan for all 3 types of 2D qusecs mentioned above.
+% % The canonical DR-plan
+% % illuminates the essence of the NP-hardness of finding optimal DR-plans for
+% % overconstrained systems, and characterizes those
+% % overconstrained 2D qusecs for which the above algorithm does find an
+% % optimal DR-plan.
+% % \item
+% % We show how to judiciously combine
+% % existing methods for recombination with searching  convex or low
+% % complexity Cayley configuration spaces to  navigate the solution space of
+% % isostatic 2D qusecs, and
+% % explore connected components of the configuration space of
+% % underconstrained 2D qusecs.
+% % \item
+
+% % \item
+% % Software implementation and videos can be found at: \todo{Provide Link}.
 \end{abstract}

### 2014-12-16 doc/GMP2015/s03_drp.tex

@@
 \section{The Canonical Decomposition-Recombination Plan}
+\label{sec:DRP}
 
@@
 
-\begin{theorem}\label{t1}
+\begin{theorem}\label{theorem:main}
 Given a well-constrained 2-dimensional bar-joint graph $G$, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
@@
 
-% We develop this notion of the canonical optimal DR-plan by carefully considering the case of a union of any two non-trivial well-constrained vertex-maximal proper subgraphs of a well-constrained graph.
-By definition, any subgraph of a well-constrained graph will either be (1) trivial, (2) under-constrained, or (3) well-constrained.
-We will analyze all 3 possibilities in turn.
-These observations will lead to interesting properties that govern the behavior of optimal DR-plans on well-constrained graphs.
+% \begin{proof}
+% Immediate from Lemma \ref{lemma:main_trivial} and \ref{lemma:main_wellconstrained}.
+% \end{proof}
 
-% We show this by carefully considering all of the potential cases of the unions of subgraphs.
-By definition, any subgraph of a well-constrained graph will either be (1) under-constrained, (2) well-constrained, or (3) trivial.
-We will analyze all 3 possibilities in turn, in the case of a union of any two non-trivial well-constrained vertex-maximal proper subgraphs.
+% \begin{lemma}\label{lemma:main_trivial}
+% In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
+% \end{lemma}
+
+% \begin{lemma}\label{lemma:main_wellconstrained}
+% In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
+% \end{lemma}
 
@@
 \subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
-For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
-We wish to prove the following bijections:
+\label{sec:union_intersection}
 
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline
-\textbf{Union is\ldots} & \textbf{Intersection is\ldots} \\ \hline
-\sout{trivial}          & --- \\ \hline
-under-constrained       & trivial \\ \hline
-well-constrained        & well-constrained \\ \hline
----                     & \sout{under-constrained} \\ \hline
-\end{tabular}
-\end{center}
-
-\begin{lemma}\label{t-l1}
-$F_i\cup F_j$ cannot be trivial.
+\begin{lemma}\label{lemma:union_intersection}
+If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that:
+\begin{enumerate}
+    \item $F_i\cup F_j$ is not trivial.
+    \item $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
+    \item $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
+    \item $F_i\cap F_j$ is not under-constrained.
+\end{enumerate}
 \end{lemma}
@@
 \begin{proof}
-% If $F_i,F_j$ are well-constrained, then $F_i\cup F_j \geq k$. Trivial graphs must be over-constrained.
+Note that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
+\begin{enumerate}
+    \item By definition of trivial graph; if $F_i\cup F_j$ were trivial, then $F_i$ and $F_j$ must be trivial.
 
-Follows from the definition of a trivial graph. It would imply that the well-constrained graphs were also trivial since they would be subgraphs of a trivial graph.
+    \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
+Therefore $d(F_i\cap F_j)=2k-l>k$.
+This means $F_i\cap F_j$ is trivial.
+\textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial.
+By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.
+
+    \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
+Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
+\textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.
+
+    \item Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained. So it follows from the above that an under-constrained intersection is not possible.
+\end{enumerate}
 \end{proof}
 
-\begin{lemma}\label{uc-l1}
-$F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
-\end{lemma}
+% For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
+% We wish to prove the following bijections:
 
-\begin{proof}
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
+% \begin{center}
+% \begin{tabular}{|c|c|}
+% \hline
+% \textbf{Union is\ldots} & \textbf{Intersection is\ldots} \\ \hline
+% \sout{trivial}          & --- \\ \hline
+% under-constrained       & trivial \\ \hline
+% well-constrained        & well-constrained \\ \hline
+% ---                     & \sout{under-constrained} \\ \hline
+% \end{tabular}
+% \end{center}
+
+% \begin{lemma}\label{t-l1}
+% $F_i\cup F_j$ cannot be trivial.
+% \end{lemma}
+
+% \begin{proof}
+% Follows from the definition of a trivial graph. It would imply that the well-constrained graphs were also trivial since they would be subgraphs of a trivial graph.
+% \end{proof}
+
+% \begin{lemma}\label{uc-l1}
+% $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
+% \end{lemma}
+
+% \begin{proof}
+% It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
 In the forward direction, we have that $d(F_i\cup F_j)=l<k$.
@@
 
-In the reverse direction, we know that $d(F_i\cap F_j)>k$ because it is trivial. Thus, by the same math, it is clear that $d(F_i\cup F_j)<k$ and it must be under-constrained.
-\end{proof}
+% In the reverse direction, we know that $d(F_i\cap F_j)>k$ because it is trivial. Thus, by the same math, it is clear that $d(F_i\cup F_j)<k$ and it must be under-constrained.
+% \end{proof}
 
-\begin{lemma}\label{wc-l1}
-$F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
-\end{lemma}
+% \begin{lemma}\label{wc-l1}
+% $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
+% \end{lemma}
 
-\begin{proof}
-It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
+% \begin{proof}
+% It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
 In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$.
@@
 
-In the reverse direction the same math proves that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it too is well-constrained.
+% In the reverse direction the same math proves that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it too is well-constrained.
+% \end{proof}
 
-% \todo{Integrate this}
-% Furthermore, all subgraphs of $F_i\cup F_j$ and $F_i\cap F_j$ are also subgraphs of $F_i$ and $F_j$, therefore all properties of well-constrained graphs are maintained.
+% \begin{lemma}\label{iuc-l1}
+% $F_i\cap F_j$ cannot be under-constrained.
+% \end{lemma}
 
-% In the forward direction: From Lemma \ref{uc-l1}, the intersection cannot be trivial.
-The intersection also cannot be under-constrained.
-If $F_i \cap F_j$ were under-constrained, then for $F_i$ to be well-constrained you would need to add more constraints (edge weight) than degrees of freedom (vertex weight) to the intersection.
-The same is true for $F_j$.
-If both of these were satisfied, then the union of $F_i$ and $F_j$ would result in an over-constrained non-trivial graph.
-This is a contradiction, by definition of a well-constrained graph.
-Therefore, the intersection can only be well-constrained.
-
-% reverse!!!!!
-\end{proof}
-
-\begin{lemma}\label{iuc-l1}
-$F_i\cap F_j$ cannot be under-constrained.
-\end{lemma}
-
-\begin{proof}
-Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained.
+% \begin{proof}
+% Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained.
 So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and well-constrained intersections are possible.
-\end{proof}
-
-% WAIT... WHAT ABOUT TWO BANANAS? The intersection is 2 vertices which is under-constrained... oh nevermind... that's well-constrained
+% \end{proof}
 
@@
 \subsubsection{Inferring Parent Structure from Union and Intersection of Children}
+\label{sec:infer_parent}
 
-\begin{lemma}\label{wc-l2}
-% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$,
+\begin{lemma}\label{lemma:wc_intersection_is_C}
 $C_i\cup C_j$ is well-constrained if and only if $C_i\cup C_j = C$.
@@
 \begin{proof}
-The forwards direction can be shown with a contradiction.
-If the $C_i\cup C_j$ was well-constrained but $C_i\cup C_j \neq C$, we would have found a larger well-constrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it would have been larger.
+\textit{Forward direction:} Proof by contradiction.
+If the $C_i\cup C_j$ was well-constrained but $C_i\cup C_j \neq C$, we would have found a larger well-constrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it.
 The two original subgraphs could not have been vertex-maximal proper.
 
-The reverse direction follows from the fact that $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
+\textit{Reverse direction:} $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
+Thus, $C_i\cup C_j$ is well-constrained.
 \end{proof}
 
-\begin{corollary}\label{wc-c1}
-% Given a well-constrained $G$ and two well-constrained vertex-maximal proper subgraphs $C_i, C_j$, l
+\begin{corollary}\label{corollary:no_edges_between_diff}
 Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
@@
 \begin{proof}
-It follows from the fact that $C_i\cup C_j$ must equal $C$ (proven in Lemma \ref{wc-l2}).
+Lemma \ref{lemma:wc_intersection_is_C} shows that $C_i\cup C_j$ must equal the entire parent $C$.
 \end{proof}
@@
 Trivial graphs are a single vertex and empty set.
-
-Also, 2D well-constrained graphs must be connected.
-% \todo{Is this actually true?}
-The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
+Furthermore, 2D well-constrained graphs must be connected.
+% The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
 The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
 
-\begin{lemma}\label{wc-l3}
-If $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
-Alternatively, if $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
+\begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
+If $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
+Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
 \end{lemma}
@@
 \begin{proof}
-Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
-Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.
-
-Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$).
-There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
-
-\begin{itemize}
-    \item 3 cases: $C_k$ cannot be $C=Idc(C,R_i\cup R_j\cup D_{i,j})$, $C_i=Idc(C,R_j\cup D_{i,j})$, or $C_j=Idc(C,R_i\cup D_{i,j})$. This is by definition.
-
-    \item 13 cases: $C_k$ cannot be a proper subgraph of $C_i$ and $C_j$ or else $C_k$ would not be vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup D_{i,j})$, $Idc(C,R'_j\cup D_{i,j})$, $Idc(C, D_{i,j})$, $Idc(C,R_i\cup D'_{i,j})$, $Idc(C,R_j\cup D'_{i,j})$, $Idc(C,R'_i\cup D'_{i,j})$, $Idc(C,R'_j\cup D'_{i,j})$, $Idc(C, D'_{i,j})$, $Idc(C,R_i)$, $Idc(C,R_j)$, $Idc(C,R'_i)$, $Idc(C,R'_j)$, and $Idc(C,\emptyset)$.
-
-    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal.
-These are the graphs $Idc(C,R'_i\cup R_j\cup D_{i,j})$ and $Idc(C,R_i\cup R'_j\cup D_{i,j})$ respectively.
-
-    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $Idc(C,R_i\cup R_j)$, $Idc(C,R'_i\cup R_j)$, $Idc(C,R_i\cup R'_j)$, or $Idc(C,R'_i\cup R'_j)$ because these are all disconnected (Corollary \ref{wc-c1}) and cannot be well-constrained.
-
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{wc-l2} that $C_k\cup C_i$ cannot be well-constrained.
-We also know it cannot be trivial because it contains well-constrained subgraphs.
-This means it must be under-constrained.
-From Lemma \ref{uc-l1}, we know that $C_k\cap C_i=Idc(C,R'_j\cup D_{i,j})$ must then be trivial.
-This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
-
-    \item 1 case: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible.
-The union with $C_i$ (or $C_j$) is not equal to $C$.
-Therefore, the union must be under-constrained which means the intersection must be trivial (a single node).
-However, $C_k\cap C_i=Idc(C,R'_j\cup D'_{i,j})$ and since $R'_j$ and $D'_{i,j}$ are not empty sets, it cannot be trivial making a contradiction.
-
-    \item 2 cases: \todo{Uses 2D requirement:} $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.
-
-    \item 1 case: $C_k=Idc(C,R_i\cup R_j\cup D'_{i,j})$ is all that remains.
-\end{itemize}
-
-Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.
-
-\todo{Needs a figure.}
+In appendix (\ref{sec:appendix}).
 \end{proof}
 
-% \pnplaceholder
-% WAIT, TRIVIAL GRAPH DEFINITION DOESN'T WORK!!!! in 4D it's the 3 clique, but 3 vertices is a subgraph of that... that is certainly not over-constrained...
-
-% NOT SO! -5C2 = -10. 3 edges and 3 nodes -> 3 - 3*4 = -9... over-constrained!
-
-% \pnplaceholder
-
-\begin{lemma}\label{uc-l2}
-% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial (a vertex or empty set).
+\begin{lemma}\label{lemma:uc_intersection_makes_all_uc}
 If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
@@
 \begin{proof}
-If some $C_i\cap C_k$ is not trivial, then it must be well-constrained.
-Then, by Lemma \ref{wc-l3}, all of the intersections are well-constrained.
+By Lemma \ref{lemma:union_intersection}, if some $C_i\cap C_k$ is not trivial, then it must be well-constrained.
+Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, all of the intersections are well-constrained.
 This contradicts the statement that $C_i\cap C_j$ is trivial.
-% It follows that if the intersection is trivial then the union is under-constrained (Lemma \ref{uc-l1}). If the union is under-constrained, then the
 \end{proof}
@@
 
-\subsubsection{Proving the Main Theorem \ref{t1}}\label{sec:maintheorem}
-Lemmas \ref{wc-l3} and \ref{uc-l2} show that $CompleteDRP(C)$ will result in one of two scenarios.
-(A) Either all $C_1,\ldots,C_N$ intersect on some common well-constrained subgraph and any two union to make $C$, or (B) all $C_1,\ldots,C_N$ intersect trivially and all of them must be unioned to make $C$.
-
-% \begin{theorem}
-% In case (A), any two unique wcvmps of $C$ can be used as the children of $C$ in $OptimalDRP(C)$.
-% \end{theorem}
-
-% \begin{proof}
-
-% \end{proof}
-
-% \begin{theorem}
-% In case (B), all wcvmps of $C$ must be used as the children of $C$ in $OptimalDRP(C)$.
-% \end{theorem}
-
-% \begin{proof}
-% If not all of them are used, the union will be under-constrained. $C$ is well-constrained, thus using not all of them is not an option.
-% \end{proof}
-
-% \pnplaceholder
-
-% \pagebreak
-% See ISSUE at top of paper, we have to assume we have a connected $G$...
-
-% Take $C'_i$ to be $C\setminus C_i$.
-Then if $C_i\cup C_j$ is well-constrained we have directly from Lemma \ref{wc-l2} for all $C'_i, C'_j$ there can be no edges between them (otherwise $C_i\cup C_j\neq C$).
-
-% The set of edges connecting $C'_i$ to $C_i$ must have a weight that sums to exactly $d(C'_i)$ since we know that $d(C)=d(C_i)=k$.
-
-% Any other $C_k$ must therefore contain $C'_i$ and $C'_j$ because it adds no density to ``tack it on'' to a graph that does not contain them already.
-
-% Now we want to know something about $C_k$.
-If $C_k \neq C_i$, then $C'_i\not\subseteq C'_k$.
-It is clear that they cannot be equal otherwise $C_k = C_i$.
-However, it also cannot be a subset.
-There are two scenarios.
-(1) $C'_k$ does not contain vertices that are incident to $C'_i$.
-If this were so, we know we could ``tack on'' $C'_i$ to the well-constrained $C_k$ without increasing the density and therefore $C_k$ could not have been maximal in the first place.
-(2) $C'_k$ contains vertices that are incident to $C'_i$.
-In this case
-
-% Take $D_{i,j}$ to be $C\setminus (C'_i\cup C'_j)$.
-Clearly $D_{i,j}$ must be well-constrained since ``tacking on'' $C'_i$ and $C'_j$ adds no density and $C$ is well-constrained.
-Therefore, if there were some other $C_k$, it must contain $D_{i,j,k}\subset D_{i,j}$ such that $D_{i,j,k}=D_{i,j}\setminus C'_k$.
-It is clear that $D_{i,j,k}$ cannot equal $D_{i,j}$ or else $C_k$ must be $C_i$ or $C_j$ to be vertex-maximal.
-
-% \begin{lemma}
-% If there exist $C_i, C_j$ such that $C_i\cap C_j$ is trivial, then for all $C_k,C_l$, $C_k\cap C_l$ is trivial or empty set.
-% \end{lemma}
-
-% \begin{proof}
-% By Lemmas... the union must be under
-% Lemma \ref{uc-l1} that $C_i\cup C_j$ must be under-constrained. Therefore, by the proof
-% It follows from Lemma \ref{wc-l1} that the only way to get
-%  and its proof that if any intersections
-% \end{proof}
+\subsubsection{Proving the Main Theorem \ref{theorem:main}}
+\label{sec:maintheorem}
+Lemmas \ref{lemma:wc_intersection_makes_all_wc} and \ref{lemma:uc_intersection_makes_all_uc} show that $CompleteDRP(C)$ will result in one of two scenarios.
+(A) Either all $C_1,\ldots,C_N$ intersect trivially and all of them must be unioned to make $C$, or (B) all $C_1,\ldots,C_N$ intersect on some common well-constrained subgraph and any two union to make $C$.
 
@@
 
-\begin{theorem}\label{wc-t1}
-% In $OptimalDRP(G)$ of a well-constrained $G$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is well-constrained, then said node can have only $2$ children.
-
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
-% Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
-\end{theorem}
-
-\begin{proof}
-In a valid DR-plan, all children must be proper subgraphs.
-Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
-It follows directly from Lemma \ref{wc-l3} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
-% you found a tree with only 2 children.
-% Clearly, this is the optimal choice at this level.
-However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
-
-Take $D=\bigcap_{k=1}^N{C_k}$ and $R_k=C\setminus C_k$.
-% This means that $D$ is also $\bigcup_{k=1}^N{R_k}$.
-Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
-% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
-% the children of this node will be
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
-% and
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
-$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
-the children of this node will be
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
-and
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two, at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
-
-% This is demonstrated with $N=3$ in Figure \ref{3-plans}.
-It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution.
-This follows from the discussion above.
-Now, at this level, the fan-in must be equal in all scenarios.
-In fact, at node $k$ it will be 1 (because of $D$ which is well-constrained) plus whatever fan-in $R_k$ incurs.
-\end{proof}
-
-\begin{theorem}\label{uc-t1}
-% In any solution to $OptimalDRP(G)$, if any two well-constrained vertex-maximal proper subgraphs of a node form a union that is under-constrained, then all well-constrained vertex-maximal proper subgraphs of said node must be included in the optimal plan.
-
+\begin{theorem}\label{theorem:main_wellconstrained}
 In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
@@
 If $D= C$, then we contradict the assumption that $C_i \cap C_j$ is trivial.
-By Lemma \ref{uc-l2}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
+By Lemma \ref{lemma:uc_intersection_makes_all_uc}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
 Furthermore, by definition of a DR-plan, $C_k=C\cap C_k=D\cap C_k$.
@@
 
-With these two proofs, the original Theorem \ref{t1} has been proven. With this we have formulated the notion of a canonical DR-plan.
+\begin{theorem}\label{theorem:main_trivial}
+In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
+% Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
+\end{theorem}
+
+\begin{proof}
+In a valid DR-plan, all children must be proper subgraphs.
+Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
+It follows directly from Lemma \ref{lemma:wc_intersection_makes_all_wc} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
+% you found a tree with only 2 children.
+% Clearly, this is the optimal choice at this level.
+However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
+
+Take $D=\bigcap_{k=1}^N{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
+% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
+% the children of this node will be
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
+% and
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
+$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
+the children of this node will be
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
+and
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
+for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two, at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
+\end{proof}
+
+% This is demonstrated with $N=3$ in Figure \ref{3-plans}.
+It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution.
+This follows from the discussion above.
+Now, at this level, the fan-in must be equal in all scenarios.
+In fact, at node $k$ it will be 1 (because of $D$ which is well-constrained) plus whatever fan-in $R_k$ incurs.
+
+With these two proofs, the original Theorem \ref{theorem:main} has been proven. With this we have formulated the notion of a canonical DR-plan.
 % Any optimal DR-plan of a well-constrained graph must follow this pattern.
-No matter which two children you choose in case (A) the recombination steps will always involve the same systems; in case (B), all children must always be included.
-
-% (2) If the union results in a well-constrained subgraph, it must be the entire graph. If this were not the case, .
-
-% More can be said about the intersection of these two subgraphs, $C_1=(V_1, E_1)$ and $C_2=(V_2, E_2)$.
-It cannot be over-constrained, otherwise the union could not be well-constrained (simple arithmetic, see the first case).
+No matter which two children you choose in case (A) all children must always be included; in case (B), the recombination steps will always involve the same systems.
 
@@
 
-% \subsection{Further Results}
+% \subsubsection{Further Results}
 
@@
 % \begin{proof}
-% Explained in Theorem \ref{wc-t1}.
+% Explained in Theorem \ref{theorem:main_wellconstrained}.
 % \end{proof}
@@
 % \begin{proof}
-% Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{wc-l2} proves is the entire graph.
+% Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{lemma:wc_intersection_is_C} proves is the entire graph.
 % \end{proof}
@@
 % \begin{proof}
-% It follows from Theorem \ref{uc-t1} that if the intersection of any two well-constrained vertex-maximal proper subgraphs of $C$ are well-constrained then all of them are well-constrained or $\emptyset$.
+% It follows from Theorem \ref{theorem:main_trivial} that if the intersection of any two well-constrained vertex-maximal proper subgraphs of $C$ are well-constrained then all of them are well-constrained or $\emptyset$.
 % \end{proof}
@@
 These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$.
-This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{wc-t1}.
+This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main_wellconstrained}.
 Also shown is the decomposition of the individual doublets.
 These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them.
-As explained in Theorem \ref{uc-t1}, all of them must be included in the optimal DR-plan.
+As explained in Theorem \ref{theorem:main_trivial}, all of them must be included in the optimal DR-plan.
 Were any left out, the union of the children would not be the entire parent.

### 2014-12-16 doc/GMP2015/s04_recombination.tex

@@
-\section{Recombination}
+\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
+\label{sec:recomb}
+In this section, we consider the {\em optimal recombination}
+problem of combining specific solutions of
+subsystems in a DR-plan into a solution of their parent system $I$, (without
+loss of generality, at the top level of the DR-plan). In the case of isostatic
+qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
+seek  {\em solution(s) (among a finite large number of solutions)
+with a specific orientation or chirality}.
+In the case of underconstrained qusecs the subsystems are the multiple roots
+of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
+an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
+chirality}.
 
-Discussion about algorithm.
+The main barrier in recombination when given an optimal DR-plan (of smallest possible
+size or max fan-in),  is that the number of children (resp. number of
+roots) - and correspondingly the  size and complexity of the (indecomposable) algebraic
+system $I$ to be solved -
+could be arbitrarily large as a function of the size of the input system.
+This difficulty can persist even after optimal parametrization of the
+indecomposable system $I$ as in \cite{XX}
+to minimize its algebraic complexity.
 
-Drop an appropriate edge. Add an appropriate non-edge to get tree-decomposable graph. Change length of edge until dropped edge is correct length (as supplied as input).
+In the following, we formulate the problem of {\em optimal
+modification},
+of the indecomposable
+algebraic system $I$ into a decomposable system with a small DR-plan (low
+algebraic complexity).
+Leveraging recent results on {\em Cayley
+configuration spaces}, our approach to the optimal modification problem
+achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
+$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$ ,  all of which have
+small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
+potentially be solved or realized easily once the orientation type of the solution
+is known  (when the DR-plan size is small enough).
+{\em (b) Solution preservation.} Moreover, the union of solution spaces of the systems in the family
+$I_{\lambda_F}$
+is guaranteed to contain
+all of $I$'s solutions. {\em (c) Efficient search.} Finally, this union can be projected into the
+so-called {\em Cayley} or distance parameter space $\lambda_F$ that is convex
+or otherwise easy to traverse in order to search for $I$'s solution
+(or connected component)
+of the desired orientation type.
+For the case when the modification is bounded, this approach provides an efficient
+algorithm for recombination. We first define the decision version of the
+problem of optimal modification
+for decomposition. The standard optimization versions are straightforward.
+
+\medskip\noindent{\bf Optimal Modification for Decomposition (OMD) Problem.} Given a generically
+independent graph $G = (V,E)$
+with no non-trivial proper isostatic subgraph (indecomposable), and 2 constants $k$ and $s$,
+does there exist a set of at most $k$  edges $E_1$ and a set of nonedges $F$
+such that the modified graph $H = (V, E\setminus E_1 \cup F)$ has a DR-plan
+of size at most $s$?  The OMD$_k$ problem is  OMD where $k$ is a fixed bound (not
+part of the input).   We say that such a tuple $(G,s)$ is a member of the
+set OMD$_k$. We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
+with appropriately small $s$.
+
+It is immediately clear that indecomposable graphs $G$ that belong in OMD$_k$ for small $k$ and
+$s$  lend themselves to modification  into decomposable graphs
+satisfying Criteria (a) and (b) above.
+However, it is not clear how Criterion (c) is met by OMD graphs.
+%
+\subsection{Using Convex Cayley Configuration Spaces}
+%
+\ref{2-tree-reduction}
+Next we provide the necessary background to describe a specific approach for achieving the requirements
+(a)-(c) mentioned above, by restricting the class of reduced graphs $G' =
+G\setminus E_1$ and their isostatic completions $H$ in the above definition of the
+OMD problem, and using a key theorem of Convex Cayley configuration
+spaces \cite{XX}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley
+parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of
+the nonedges $F$
+is always convex for any
+given lengths $\delta$ for the edges of $H$ (i.e, over all the realization
+s of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
+This set is called the (2-dimensional) {\em Cayley configuration space}
+of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted
+$\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian
+realization space of $(H,\delta)$ on the Cayley parameters $F$.
+Such graphs $H$ are said to have {\em convexifiable Cayley configuration spaces
+with parameters $F$} (short: {\em convexifiable}). To state the theorem, we first have to define the
+notion of {\em 2-sums} and {\em 2-trees}.
+Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and
+$V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
+respectively.  A {\em 2-sum} of
+$H_1$ and $H_2$ is a
+graph $H$  obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$.
+In this case, $H_1$ and $H_2$ are called {\em 2-sum components} of $H$.
+A {\em minimal 2-sum component} of $H$ is  one that cannot be further split
+into 2-sum components.
+A {\em 2-tree} is recursively obtained by taking
+a 2-sum of two 2-trees, with the base case of 2-tree being a triangle.
+A {\em partial 2-tree} is a 2-tree minus some edges.
+Partial 2-trees are characterized as having $K_4$ as forbidden minor are also called series parallel graphs.
+
+\begin{theorem}
+    \label{convexcayley}
+    \cite{XX}
+    $H$ has a convexifiable Cayley configuration space  with parameters $F$
+    if and only if for each $f\in F$  all the minimal 2-sum components
+    of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees.
+    The Cayley configuration space $\Phi_F(H,\delta)$
+    of a bar-joint system or linkage $(H,\delta)$ is a convex polytope.
+    When $H\cup F$ is a 2-tree, the
+    bounding hyperplanes of this polytope are triangle inequalities
+    relating the lengths of edges of the triangles in $H\cup F$.
+\end{theorem}
+
+The idea of our approach to achieve the criteria
+(a)-(c) begins with the following simple but useful theorem.
+
+\begin{theorem}
+    \label{omdk}
+Given an indecomposable graph $G$,
+let $G'$ be a spanning partial 2-tree subgraph in $G$ with
+$k$ fewer edges than $G$.
+Then  $(G,2)$ belongs in the set OMD$_k$.
+\end{theorem}
+
+\begin{proof}
+The proof follows from the fact that 2-trees are well decomposable and have
+simple DR-plans of size 2.
+We know that $G$ can be reduced
+by removing $k$ edges to create a partial-2-tree $G'$
+which can then be completed to an (isostatic) 2-tree by adding some set of non-edges
+$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2, proving the theorem.
+\end{proof}.
+
+We refer to such graphs $G$ in short as {\em $k$-approximately
+convexifiable,} where the reduced graphs $G'$ and isostatic completions $H$
+are convexifiable.
+As observed earlier, since graphs such as $G$ are in OMD$_k$, Criteria (a) and (b)
+are automatically met for small enough $k$.
+Criterion $c$ is addressed
+as described in the following efficient search procedure
+which clarifies the dependence of the complexity on the number and ranges  of Cayley
+parameters $F$.
+
+\begin{theorem}
+    \label{criterionc}
+    [Efficient search]
+    For an indecomposable, $k$-approximately convexifiable graph $G =
+    (V,E)$,
+   let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph
+    where $|D| \le  k$. Let  $F$ be a set of nonedges of $G$ such that
+    $H = (V, E'\cup F)$ is a 2-tree.
+    Each solution $p$ (or connected component of a solution space)
+    of $(G,\delta)$
+    of an orientation type $\sigma_p$ can be found in time $O(\log(W))$ where
+    $W$ is the number of cells of desired accuracy (discrete volume) of
+    the convex
+    polytope $\Phi_F(G',\delta_E')$.  The (discrete) volume $W$ is exponential in
+    $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
+\end{theorem}
+
+\begin{proof}
+    The cartesian realization
+    space of $(H,<\delta_E', \lambda_F>)$ is
+    computed easily with a DR-plan of size 2, and is the union of $2^t$
+    solutions (modulo orientation preserving isometries) each with a distinct orientation type, where $t$ is the number of triangles in the 2-tree
+    $H$;
+    here $\delta_{S}$ is the restriction of the length vector $\delta$ to
+    the edges in $S$.
+    A desired solution $p$ (or connected component of a solution space)
+    of $(G,\delta)$
+    of an orientation type $\sigma_p$ can be found
+    by a subdivided binary search   of the cartesian realization space of
+    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ -
+    as $\lambda_F$ ranges over
+    the discretized convex
+    polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
+    Theorem \ref{convexcayley}.
+    A solution $p$  is found  when the lengths for nonedges in $D$ match
+    $\delta_D$.
+\end{proof}
+
+\medskip\noindent{\bf Example}.
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
+graphs, i.e, its DR-plan has a root and 9
+children corresponding to the 9 edges. Solving or recombining the system
+$(G,\delta)$
+corresponding to the root of
+this DR-plan involves solving a quadratic system with 8 equations and variables,
+Instead of simultaneously solving this system, we could instead use the fact
+that $G=K_{3,3}$ is in OMD$_1$: remove the
+edge $e$ in Figure \ref{omdexample} to give a partial 2-tree $G'$.
+Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
+The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
+is a single interval of attainable length values $\lambda_F$ for
+the edge $f$.
+When $\delta$ is generic, i.e, does not admit collinearities or coincidences
+in the realizations of $(G,\delta)$,
+the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
+solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
+with distinct orientation types $\sigma_p$,
+(two orientation choices for each of the 4 triangles)
+that can be obtained by solving a sequence of 4 single quadratics in 1
+variable (DR-plan of size 2).
+By subdivided
+binary search
+in the interval $\lambda_f \in
+\Phi_f(G', \delta_{E\setminus e})$, the desired
+solution $p$  of $(G,\delta)$ is found when the length of the nonedge
+$e$ in  the realization
+$q^p_{\lambda_f}$
+is $\delta_e$.
+%
+\subsection{Optimized Modification by Enlarging the class of Reduced
+Graphs}
+\label{tdecomp}
+It is possible in principle to decrease $k$ for a OMD$_k$ graph
+(i.e, the number of edges to be removed
+to ensure an isostatic completion that is decomposable with a small DR-plan) by
+considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
+partial 2-trees but nevertheless have convex Cayley configuration spaces at
+least when the realization space is restricted to a sufficiently
+comprehensive orientation type.
+In particular, the so-called {\em tree-decomposable graphs of low Cayley
+complexity} \cite{XX} include the partial 2-trees and many others that are
+not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
+OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
+- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
+linkages based on such graphs $G'$
+are known to be single intervals
+when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
+In addition, the bounds of these intervals are of low algebraic
+complexity.
+More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
+i.e, the computation of these bounds is tree-decomposable. Alternatively,
+the bounds are in a simple quadratic or radically solvable extension field over the rationals,
+or can be computed by solving a
+triangularized system of quadratics.

### 2014-12-16 doc/GMP2015/s08_pinnedline.tex

@@
 Since the pins are given fixed positions in the plane,
-once we get a solution a well-constrained sub-system,
-the solution will automatically be consistent with the remaining of the system.
+ the solutions of a well-constrained sub-system will automatically be consistent with solutions of the remaining of the system.

### 2014-12-16 doc/GMP2015/s09_problems.tex

\section{Problem Relationships and Unification}
\label{sec:table}
In this section we provide a unified view of  the various problems
studied in the previous 2 sections, along with formal reductions between
them,  and relationships to other known problems and results as well as open
questions.
%
\subsection{Special classes of small DR-plans}
As seen in the previous section, 2-trees and tree-decomposable graphs
have not only small, but also special DR-plans that permit easy
solving - essentially a single quadratic at a time.

The {\em restricted optimal DR-planning problem} requires
DR-plans of one of these types, which reduces to recognizing if
the input graph is a 2-tree or a tree-decomposable graph for which simple
near-linear time algorithms are available \cite{XX} and the DR-plan is a
by-product output of the recognition algorithm.

In the recombination setting, the corresponding {\em restricted
OMD$_k$  problem}
requires the reduced graph $G'$  and its isostatic completion $H$ to be
2-trees as in Section \ref{2-tree-completion}
or to be a low Cayley complexity tree-decomposable graph as in Section
\ref{tdecomp}. Clearly these problems have deterministic polynomial time
algorithms, the algorithms run in time exponential in $k$.
As a result, when $k$ is part of the input, the {\em restricted OMD problem}
has the potential to be difficult.
For example,  when the isostatic completion $H$ is required to be a 2-tree
the restricted OMD problem is reducible to the maximum spanning series-parallel
subgraph problem shown by \cite{XX,XX} to be NP-complete
even if the input graph is planar of maximum degree at most 6.
However, since the OMD problem has other input restrictions
such as not having any proper isostatic subgraphs,
it is not clear if the reverse reduction exists and hence it is unclear
whether the OMD problem is NP-complete.
The same holds for the restricted OMD problem where the isostatic completion
$H$ is required to be a tree-decomposable graph of low Cayley complexity (i.e, have
special, small DR-plans).
One potential obstacle to an indecomposable graph $G$'s membership in the
restricted OMD$_k$ for small $k$ is
if $G$ is triconnected and has large girth. In fact,  6-connected (hence
rigid) graphs with
arbitrarily large girth have been constructed in \cite{XX}.
%
\subsection{Optimal Modification, Completion and Recombination: formal
connections}
%
The OMD problem is closely related to a well-studied problem of
completion of an underconstrained system to an isostatic one with a small
DR-plan.
\begin{observation}
The (decision version of) the {\bf optimal completion problem (OC)} from \cite{bib:survey, j-a,
fang} is OMD$_0$.
\end{observation}.
In fact, a {\em restricted OC} problem was studied by \cite{j-a} requiring the
completion to be tree-decomposable.
It is not clear if the reverse reduction holds, i.e, is OMD reducible to OC?

We now connect the OMD problem to the informal {\em optimal recombination
(OR)} problem
mentioned as motivation at the beginning of
Section \ref{sec:recomb}.

In order to connect the OR problem to OMD,
when the input graph for is the isostatic graph at the DR-plan root, we do not consider the case
where the two child {\em solved subgraphs}
(corresponding to already solved subsystems)
have a nontrivial intersection (in this case
the recombination is trivial). We only consider the case where no two child
solved subgraphs (resp. two root subgraphs when the input graph is
underconstrained) share more than 1 vertex.
We replace such solved subgraphs  by isostatic graphs as
follows. If a solved subgraph shares at most one vertex with the remainder of
the graph, simply replace it by an edge one of whose endpoints  is
the shared vertex. Otherwise, replace it by  a 2-tree graph of the shared vertices.
Finally, we add the additional
restriction to the OM Problem that
when any edge in a solved subgraph is chosen among the $k$ edges to be
removed, in fact the entire solved subgraph must be removed  and all of its
edges must be counted in $k$.

This reduction is used also for transferrring  algorithms for
optimal DR-planning, recombination, completion, OMD, and other problems and
to so-called {\em
multi-body pin} from so-called{\em bar-joint-systems}, by showing that the corresponding
prblems for the latter are reduced to the correspnding problems on
bar-joint systems.

{\bf Note.} This section deals only with bar-joint systems, however,
extensions of several of these problems, approaches and solutions
to other qusecs beyond  bar-joint - specifically pinned line incidence
systems and
multi-body-pin systems
will be expanded in Sections \ref{sec:pinnedline} and \ref{sec:bodypin}.

### 2014-12-16 doc/GMP2015/s04_recombination.tex

@@
 \label{sec:recomb}
+
+\begin{figure*}\centering
+\begin{subfigure}{.3\linewidth}\centering
+  \begin{tikzpicture}[scale=2]
+        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
+        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
+        \tikzstyle{e}=[]
+
+        \node[c] (v0) at (  -1,  0)     [label=left:$v_0$]{};
+        \node[c] (v1) at (-0.5,  0.866) [label=above left:$v_1$]{};
+        \node[c] (v2) at ( 0.5,  0.866) [label=above right:$v_2$]{};
+        \node[c] (v3) at (   1,  0)     [label=right:$v_3$]{};
+        \node[c] (v4) at ( 0.5, -0.866) [label=below right:$v_4$]{};
+        \node[c] (v5) at (-0.5, -0.866) [label=below left:$v_5$]{};
+
+        \tedge{v3}{v2}{solid}{}{};
+        \tedge{v2}{v1}{solid}{}{};
+        \tedge{v1}{v0}{solid}{}{};
+        \tedge{v0}{v5}{solid}{}{};
+        \tedge{v5}{v4}{solid}{}{};
+        \tedge{v4}{v3}{solid}{}{};
+
+        \tedge{v3}{v0}{solid}{}{};
+        \tedge{v2}{v5}{solid}{$e$}{left, pos=0.15};
+        \tedge{v1}{v4}{solid}{}{};
+    \end{tikzpicture}
+
+    \caption{}
+\end{subfigure}
+%
+\begin{subfigure}{.3\linewidth}\centering
+  \begin{tikzpicture}[scale=2]
+        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
+        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
+        \tikzstyle{e}=[]
+
+        \node[c] (v0) at (0,  1) [label=above:$v_0$]{};
+        \node[c] (v1) at (0,  0)     [label=below:$v_1$]{};
+        \node[c] (v2) at (-0.5, 1.5) [label=above:$v_2$]{};
+        \node[c] (v3) at (1, 1)     [label=above:$v_3$]{};
+        \node[c] (v4) at (1, 0) [label=below:$v_4$]{};
+        \node[c] (v5) at (1.5,  1.5) [label=above:$v_5$]{};
+
+        \tedge{v3}{v2}{solid}{}{};
+        \tedge{v2}{v1}{solid}{}{};
+        \tedge{v1}{v0}{solid}{}{};
+        \tedge{v0}{v5}{solid}{}{};
+        \tedge{v5}{v4}{solid}{}{};
+        \tedge{v4}{v3}{solid}{}{};
+
+        \tedge{v3}{v0}{solid}{}{};
+        \tedge{v2}{v5}{dashed}{$e$}{above};
+        \tedge{v1}{v4}{solid}{}{};
+    \end{tikzpicture}
+
+    \caption{}
+\end{subfigure}
+%
+\begin{subfigure}{.3\linewidth}\centering
+  \begin{tikzpicture}[scale=2]
+        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
+        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
+        \tikzstyle{e}=[]
+
+        \node[c] (v0) at (0,  1) [label=above:$v_0$]{};
+        \node[c] (v1) at (0,  0)     [label=below:$v_1$]{};
+        \node[c] (v2) at (-0.5, 1.5) [label=above:$v_2$]{};
+        \node[c] (v3) at (1, 1)     [label=above:$v_3$]{};
+        \node[c] (v4) at (1, 0) [label=below:$v_4$]{};
+        \node[c] (v5) at (1.5,  1.5) [label=above:$v_5$]{};
+
+        \tedge{v3}{v2}{solid}{}{};
+        \tedge{v2}{v1}{solid}{}{};
+        \tedge{v1}{v0}{solid}{}{};
+        \tedge{v0}{v5}{solid}{}{};
+        \tedge{v5}{v4}{solid}{}{};
+        \tedge{v4}{v3}{solid}{}{};
+
+        \tedge{v3}{v0}{solid}{}{};
+        \tedge{v2}{v5}{dashed}{$e$}{above};
+        \tedge{v1}{v4}{solid}{}{};
+
+        \tedge{v1}{v3}{solid, ultra thick}{$f$}{left, pos=0.4};
+    \end{tikzpicture}
+
+    \caption{}
+\end{subfigure}
+
+\caption{(a) The $K3,3$ with one labeled edge $e$.
+(b) The $K3,3$ with $e$ removed (dashed line) and rearranged to illustrate that it is now a partial 2-tree.
+(c) The $K3,3$ with $e$ removed and $f$ (bold line) added to make a 2-tree.
+Note that the non-edge $(v_0,v_4)$ would also be a valid choice.}
+\label{fig:omd_k33_example}
+\end{figure*}
+
 In this section, we consider the {\em optimal recombination}
@@
 that $G=K_{3,3}$ is in OMD$_1$: remove the
-edge $e$ in Figure \ref{omdexample} to give a partial 2-tree $G'$.
+edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
 Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.

### 2014-12-16 doc/GMP2015/s03_drp.tex
Major restructuring of DRP section.

@@
 
+\newcommand{\usestwod}{\todo{Uses 2D requirement:}}
+\renewcommand{\usestwod}{}
+
 \subsection{Theory}
 
-% \begin{figure*}\centering
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at (1,1) {$R_j$};
-%     \node[v] (v2) at (3,1) {$R_k$};
-%     \node[v] (v3) at (2,2) {$D$};
-%     \node[v] (v4) at (2,3.3) {$R_i$};
-
-%     \tedge{v1}{v3}{solid}{}{above};
-%     \tedge{v2}{v3}{solid}{}{above};
-%     \tedge{v3}{v4}{solid}{}{right};
-% \end{tikzpicture}
-% \ldots with weights \ldots
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at (1,1) {y};
-%     \node[v] (v2) at (3,1) {z};
-%     \node[v] (v3) at (2,2) {k};
-%     \node[v] (v4) at (2,3.3) {x};
-
-%     \tedge{v1}{v3}{solid}{y}{above};
-%     \tedge{v2}{v3}{solid}{z}{above};
-%     \tedge{v3}{v4}{solid}{x}{right};
-
-%     % \draw (3.3,2.5) circle (1.9cm);
-%     % \draw (.7,2.5) circle (1.9cm);
-%     % \draw (2,.5) circle (1.9cm);
-% \end{tikzpicture}
-% \\
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at (0,  0) {kxyz};
-%     \node[v] (v2) at (-1,-1) {kxy};
-%     \node[v] (v3) at ( 1,-1) {kyz};
-%     \node[v] (v4) at (-2,-2) {kx};
-%     \node[v] (v5) at ( 0,-2) {ky};
-%     \node[v] (v6) at ( 2,-2) {kz};
-
-%     \tedge{v1}{v2}{solid}{}{};
-%     \tedge{v1}{v3}{solid}{}{};
-%     \tedge{v2}{v4}{solid}{}{};
-%     \tedge{v2}{v5}{solid}{}{};
-%     \tedge{v3}{v5}{solid}{}{};
-%     \tedge{v3}{v6}{solid}{}{};
-% \end{tikzpicture}
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at (0,  0) {kxyz};
-%     \node[v] (v2) at (-1,-1) {kxy};
-%     \node[v] (v3) at ( 1,-1) {kxz};
-%     \node[v] (v4) at (-2,-2) {kx};
-%     \node[v] (v5) at ( 0,-2) {ky};
-%     \node[v] (v6) at ( 2,-2) {kz};
-
-%     \tedge{v1}{v2}{solid}{}{};
-%     \tedge{v1}{v3}{solid}{}{};
-%     \tedge{v2}{v4}{solid}{}{};
-%     \tedge{v2}{v5}{solid}{}{};
-%     \tedge{v3}{v4}{solid}{}{};
-%     \tedge{v3}{v6}{solid}{}{};
-% \end{tikzpicture}
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{e}=[]
-
-%     \node[v] (v1) at (0,  0) {kxyz};
-%     \node[v] (v2) at (-1,-1) {kxz};
-%     \node[v] (v3) at ( 1,-1) {kyz};
-%     \node[v] (v4) at (-2,-2) {kx};
-%     \node[v] (v5) at ( 0,-2) {ky};
-%     \node[v] (v6) at ( 2,-2) {kz};
-
-%     \tedge{v1}{v2}{solid}{}{};
-%     \tedge{v1}{v3}{solid}{}{};
-%     \tedge{v2}{v4}{solid}{}{};
-%     \tedge{v2}{v6}{solid}{}{};
-%     \tedge{v3}{v5}{solid}{}{};
-%     \tedge{v3}{v6}{solid}{}{};
-% \end{tikzpicture}
-% % \\
-% % \begin{tikzpicture}
-% %     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-% %     \tikzstyle{e}=[]
-
-% %     \node[v] (v1) at ( 0, 0) {kxyz};
-% %     \node[v] (v2) at (-1,-1) {kxy};
-% %     \node[v] (v3) at (-2,-2) {kx};
-% %     \node[v] (v4) at (-3,-3) {k};
-% %     \node[v] (v5) at ( 1,-1) {z plan};
-% %     \node[v] (v6) at ( 0,-2) {y plan};
-% %     \node[v] (v7) at (-1,-3) {x plan};
-% %     \node[v] (v8) at (-2,-4) {k plan};
-
-% %     \tedge{v1}{v2}{solid}{}{};
-% %     \tedge{v1}{v5}{solid}{}{};
-% %     \tedge{v2}{v3}{solid}{}{};
-% %     \tedge{v2}{v6}{solid}{}{};
-% %     \tedge{v3}{v4}{solid}{}{};
-% %     \tedge{v3}{v7}{solid}{}{};
-% %     \tedge{v4}{v8}{solid}{}{};
-% % \end{tikzpicture}
-
-% \caption{A well-constrained graph whose nodes are subgraphs labeled with their density.
-Below are its 3 optimal DR-Plans (partially complete).
-The center vertex has weight $k$ and is well-constrained itself.
-The radial nodes have weights $x,y,z \neq k$ with equivalent edge weights, making any union between $k$ and the others well-constrained.
-\todo{Make this figure more clear.}}
-% \label{3-plans}
-% \end{figure*}
-
 In the problem of the optimal DR-Plan there is generally not a unique plan.
-For example, a graph that is formed by the union of 3 well-constrained subgraphs that all intersect on a well-constrained subgraph has ${3\choose 2} = 3$ unique optimal DR-Plans; any two of these subgraphs will suffice as children.
 % Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
@@
 The theorem we seek to prove is thus:
-However, we will show that regardless of which children are chosen for the plan, the recombination will require solving of the same systems.
-Thus, this definition of an optimal DR-plan could be considered the canonical DR-plan.
-To assist in showing this, we prove this core theorem throughout this section:
+However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems.
+Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
+% To assist in showing this, we prove this core theorem throughout this section:
+
+In this section, we discuss 2D bar-joint graphs.
+All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$.
+Trivial graphs are a single vertex and empty set.
+Furthermore, 2D well-constrained graphs must be connected.
+% The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
+The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
 
@@
 \end{enumerate}
+This covers all potential cases.
 \end{theorem}
 
-% \begin{proof}
-% Immediate from Lemma \ref{lemma:main_trivial} and \ref{lemma:main_wellconstrained}.
-% \end{proof}
+\begin{proof}
+Remark \ref{lemma:union_intersection} shows that the intersection of any two well-constrained subgraphs (such as $C_i$ and $C_j$) can only result in trivial or well-constrained subgraphs.
+Therefore, these are the only possibilities to consider.
 
-% \begin{lemma}\label{lemma:main_trivial}
-% In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
-% \end{lemma}
+\begin{enumerate}
+\item
+% In a valid DR-plan, all children must be proper subgraphs. Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
+It follows directly from Lemma \ref{lemma:wc_intersection_makes_all_wc} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
+% you found a tree with only 2 children.
+% Clearly, this is the optimal choice at this level.
+However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
 
-% \begin{lemma}\label{lemma:main_wellconstrained}
-% In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
-% \end{lemma}
+Take $D=\bigcap_{k\in N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
+% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
+% the children of this node will be
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
+% and
+% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
+$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
+the children of this node will be
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
+and
+$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
+for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
+
+%%%%%%%%%%%%%%%
+
+\item
+Take $X\subset \{1,\ldots,N\}$ and $D=\bigcup_{i\in X}{C_i}$.
+If $D\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
+
+\usestwod
+By Lemma \ref{lemma:uc_intersection_makes_all_uc}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
+Since $C_k=C\cap C_k=D\cap C_k$, by definition of a DR-plan, the assumption that $C_k$ is well-constrained has been contradicted.
+Thus, $X$ must be all of $\{1,\ldots,N\}$.
+\end{enumerate}
+\end{proof}
+
+The following subsections provide the supporting proofs.
 
@@
 
-\begin{lemma}\label{lemma:union_intersection}
+% This is general, not for 2d
+\begin{remark}\label{lemma:union_intersection}
 If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that:
@@
 \end{enumerate}
-\end{lemma}
+\end{remark}
 
@@
 \end{proof}
-
-% For the following lemmas, we define $F_i$ and $F_j$ as well-constrained subgraphs of the same well-constrained graph $F$.
-% We wish to prove the following bijections:
-
-% \begin{center}
-% \begin{tabular}{|c|c|}
-% \hline
-% \textbf{Union is\ldots} & \textbf{Intersection is\ldots} \\ \hline
-% \sout{trivial}          & --- \\ \hline
-% under-constrained       & trivial \\ \hline
-% well-constrained        & well-constrained \\ \hline
-% ---                     & \sout{under-constrained} \\ \hline
-% \end{tabular}
-% \end{center}
-
-% \begin{lemma}\label{t-l1}
-% $F_i\cup F_j$ cannot be trivial.
-% \end{lemma}
-
-% \begin{proof}
-% Follows from the definition of a trivial graph. It would imply that the well-constrained graphs were also trivial since they would be subgraphs of a trivial graph.
-% \end{proof}
-
-% \begin{lemma}\label{uc-l1}
-% $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
-% \end{lemma}
-
-% \begin{proof}
-% It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction, we have that $d(F_i\cup F_j)=l<k$.
-If it were equal to $k$ it would be well-constrained since, as a subgraph of $F$, it satisfies the other properties of a well-constrained graph; if it were greater than $k$ then it would be over-constrained.
-Therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=2k-l>k$.
-The only way to have an over-constrained subgraph in a well-constrained graph is if it is trivial.
-
-% In the reverse direction, we know that $d(F_i\cap F_j)>k$ because it is trivial. Thus, by the same math, it is clear that $d(F_i\cup F_j)<k$ and it must be under-constrained.
-% \end{proof}
-
-% \begin{lemma}\label{wc-l1}
-% $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
-% \end{lemma}
-
-% \begin{proof}
-% It follows from definitions that $d(F_i)=k$ and $d(F_j)=k$.
-In the forward direction we have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)=k$.
-Since $F_i\cap F_j$ is a subgraph of $F$, which is well-constrained, it is also well-constrained.
-
-% In the reverse direction the same math proves that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it too is well-constrained.
-% \end{proof}
-
-% \begin{lemma}\label{iuc-l1}
-% $F_i\cap F_j$ cannot be under-constrained.
-% \end{lemma}
-
-% \begin{proof}
-% Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained.
-So it follows from Lemmas \ref{t-l1}, \ref{uc-l1}, and \ref{wc-l1} that only trivial and well-constrained intersections are possible.
-% \end{proof}
 
@@
 
-At this point we limit our discussion to the scope of 2D graphs.
-All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$.
-Trivial graphs are a single vertex and empty set.
-Furthermore, 2D well-constrained graphs must be connected.
-% The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
-The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
-
 \begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
@@
 \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Proving the Main Theorem \ref{theorem:main}}
-\label{sec:maintheorem}
-Lemmas \ref{lemma:wc_intersection_makes_all_wc} and \ref{lemma:uc_intersection_makes_all_uc} show that $CompleteDRP(C)$ will result in one of two scenarios.
-(A) Either all $C_1,\ldots,C_N$ intersect trivially and all of them must be unioned to make $C$, or (B) all $C_1,\ldots,C_N$ intersect on some common well-constrained subgraph and any two union to make $C$.
-
-% \begin{figure}\centering
-% \begin{tikzpicture}
-%     \draw[rotate=0]  (0,0) ellipse (1.5cm and .3cm);
-%     \draw[rotate around={-59:(.7,1.2)}] (.8,1.1) ellipse (1.5cm and .3cm);
-%     \draw[rotate around={59:(-.7,1.2)}] (-.8,1.1) ellipse (1.5cm and .3cm);
-
-%     % \draw[rotate around={59:(.7,-1.2)}] (.8,-1.1) ellipse (1.5cm and .3cm);
-%     % \draw[rotate around={-59:(-.7,-1.2)}] (-.8,-1.1) ellipse (1.5cm and .3cm);
-
-%     % \draw[rotate=0] (1,1) ellipse (.1cm and .1cm);
-%     % \draw[rotate=66] (0,0) ellipse (.1cm and .1cm);
-% \end{tikzpicture}
-% \begin{tikzpicture}
-%     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-%     \tikzstyle{el}=[draw, ellipse, minimum height=.75cm, minimum width=1.5cm]
-%     \tikzstyle{e}=[]
-
-%     \node[el] (m) at (0,    0) {k};
-%     \node[v]  (l) at (-1.5, 0) {b};
-%     \node[v]  (r) at (1.5,  0) {c};
-
-%     \node[el] (tl) at (-1,  1) {i};
-%     \node[el] (tr) at (1,   1) {j};
-%     \node[v]  (t)  at (0,   2) {a};
-
-%     % \node[el] (bl) at (-1,  -1) {l};
-%     % \node[el] (br) at (1,   -1) {m};
-%     % \node[v]  (b)  at (0,   -2) {d};
-
-%     % \node[el] (v2) at (-1,-1) {b};
-%     % \node[el] (v3) at (-1,-1) {b};
-%     % \node[el] (v4) at (-1,-1) {b};
-%     % \node[el] (v5) at (-1,-1) {b};
-
-%     \tedge{m}{l}{solid}{}{};
-%     \tedge{m}{r}{solid}{}{};
-
-%     \tedge{l}{tl}{solid}{}{};
-%     \tedge{tl}{t}{solid}{}{};
-%     \tedge{r}{tr}{solid}{}{};
-%     \tedge{tr}{t}{solid}{}{};
-
-%     % \tedge{l}{bl}{solid}{}{};
-%     % \tedge{bl}{b}{solid}{}{};
-%     % \tedge{r}{br}{solid}{}{};
-%     % \tedge{br}{b}{solid}{}{};
-% \end{tikzpicture}
-
-% \caption{Each node contains a subgraph of the entire, well-constrained graph $G$.
-We take $Idc(G,a\cup i\cup b)$, $Idc(G,a\cup j\cup c)$, and $Idc(G,b\cup k\cup c)$ to be well-constrained.
-If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c$ must all be trivial subgraphs.
-% % Each node contains a subgraph of the entire, well-constrained graph.
-We take $a\cup i\cup b$, $a\cup j\cup c$, $b\cup k\cup c$, $b\cup l\cup d$, and $c\cup m\cup d$ to be well-constrained.
-If the union of any two of these listed unions is under-constrained then it immediately follows that all such unions are under-constrained and that nodes $a,b,c,d$ must all be trivial subgraphs.
-% % THIS IS NOT TECHNICALLY CORRECT... $a\cup i\cup b$ doesn't include the edge between them... Is there a better way to phrase this?
-% }
-% \label{trivial-intersection}
-% \end{figure}
-
-\begin{theorem}\label{theorem:main_wellconstrained}
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cap C_j$ is trivial, then all $C_1,\ldots, C_N$ must be included in the optimal plan as children of $C$.
-\end{theorem}
-
-% \begin{proof}
-% From Lemma \ref{uc-l1}, if it is trivial then $C_i \cup C_j$ must be under-constrained.
-If it is under-constrained, it must be part of some larger well-constrained subgraph $D$.
-If $D$ is not $C$ itself, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
-%  % For $C$ to be well-constrained... WHAT????
-
-% If there was some $X\subset [1,\ldots,N]$ such that $\bigcup_{i\in X}{C_i}\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
-
-% This means that in the case of a trivial intersection, all of the children come together to form some complex structure that is itself rigid.
-Such an example is shown in Figure \ref{trivial-intersection}.
-% \end{proof}
-
-\begin{proof}
-Take $X\subset \{1,\ldots,N\}$ and $D=\bigcup_{i\in X}{C_i}$.
-If $D\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
-
-\todo{Uses 2D requirement:}
-If $D= C$, then we contradict the assumption that $C_i \cap C_j$ is trivial.
-By Lemma \ref{lemma:uc_intersection_makes_all_uc}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
-Furthermore, by definition of a DR-plan, $C_k=C\cap C_k=D\cap C_k$.
-Together, this shows that $C_k$ cannot be well-constrained, violating the properties of a DR-plan.
-\end{proof}
-
-% As such, there must be additional vertex-maximal proper subgraphs, all of whose intersections with one another are trivial subgraphs or $\emptyset$.
-All of these subgraphs form a complex structure that is well-constrained and must be the parent itself (otherwise we did not have vertex-maximal subgraphs to begin with).
-
-\begin{theorem}\label{theorem:main_trivial}
-In $OptimalDRP(G)$ of a well-constrained $G$, if $C_i \cup C_j$ is well-constrained, then any 2 well-constrained vertex-maximal proper subgraphs of $C$ will be sufficient as its children.
-% Namely, these children are any 2 of its well-constrained vertex-maximal proper subgraphs.
-\end{theorem}
-
-\begin{proof}
-In a valid DR-plan, all children must be proper subgraphs.
-Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
-It follows directly from Lemma \ref{lemma:wc_intersection_makes_all_wc} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
-% you found a tree with only 2 children.
-% Clearly, this is the optimal choice at this level.
-However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
-
-Take $D=\bigcap_{k=1}^N{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
-% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
-% the children of this node will be
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
-% and
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
-$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
-the children of this node will be
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
-and
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two, at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
-\end{proof}
-
-% This is demonstrated with $N=3$ in Figure \ref{3-plans}.
-It is clear that until this point, the fan-in at every node is exactly 2 in the optimal solution.
-This follows from the discussion above.
-Now, at this level, the fan-in must be equal in all scenarios.
-In fact, at node $k$ it will be 1 (because of $D$ which is well-constrained) plus whatever fan-in $R_k$ incurs.
-
-With these two proofs, the original Theorem \ref{theorem:main} has been proven. With this we have formulated the notion of a canonical DR-plan.
-% Any optimal DR-plan of a well-constrained graph must follow this pattern.
-No matter which two children you choose in case (A) all children must always be included; in case (B), the recombination steps will always involve the same systems.
 
@@
 \begin{figure*}\centering
-\begin{subfigure}{.3\linewidth}
-  \centering
+\begin{subfigure}{.3\linewidth}\centering
   \begin{tikzpicture}[scale=2]
         \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
+        \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
         \tikzstyle{e}=[]
 
-        \node[v] (v1) at (0,0.866) {$a_1$};
-        \node[v] (v2) at (-1,-0.866) {$a_2$};
-        \node[v] (v3) at (1,-0.866) {$a_3$};
+        \node[c] (v1) at (0,0.866) [label={left,inner sep=.555}:$a_1$]{};
+        \node[c] (v2) at (-1,-0.866) [label={below,inner sep=.555}:$a_2$]{};
+        \node[c] (v3) at (1,-0.866) [label={above,inner sep=.555}:$a_3$]{};
 
-        \node[v] (v4) at (0,0.577-.1) {$b_1$};
-        \node[v] (v5) at (-0.667,-0.577-.1) {$b_2$};
-        \node[v] (v6) at (0.667,-0.577-.1) {$b_3$};
+        \node[c] (v4) at (0,0.577-.1) [label={left,inner sep=.555}:$b_1$]{};
+        \node[c] (v5) at (-0.667,-0.577-.1) [label={below,inner sep=.555}:$b_2$]{};
+        \node[c] (v6) at (0.667,-0.577-.1) [label={above,inner sep=.555}:$b_3$]{};
 
-        \node[v] (v7) at (0,0.289-.2) {$c_1$};
-        \node[v] (v8) at (-0.333,-0.289-.2) {$c_2$};
-        \node[v] (v9) at (0.333,-0.289-.2) {$c_3$};
+        \node[c] (v7) at (0,0.289-.2) [label={left,inner sep=.555}:$c_1$]{};
+        \node[c] (v8) at (-0.333,-0.289-.2) [label={below,inner sep=.555}:$c_2$]{};
+        \node[c] (v9) at (0.333,-0.289-.2) [label={above,inner sep=.555}:$c_3$]{};
 
@@
 \end{subfigure}%
-\begin{subfigure}{.7\linewidth}
-    \centering
+\begin{subfigure}{.7\linewidth}\centering
     \begin{tikzpicture}[scale=.9, transform shape]
         \tikzstyle{v}=[draw, circle, minimum size=0.75cm, font=\footnotesize]
+        % \tikzstyle{b}=[draw, font=\footnotesize]
+        \tikzstyle{b}=[]
         \tikzstyle{e}=[]
 
-        \node (c0) at (0,0) {$\{a_1,a_2,a_3,b_1,b_2,b_3,c_1,c_2,c_3\}$};
-        \node (c1a) at (-2,-1) {$\{a_1,a_2,a_3,b_1,b_2,b_3\}$};
-        \node (c1b) at (2,-1) {$\{b_1,b_2,b_3,c_1,c_2,c_3\}$};
-        \node (c2a) at (-2,-2) {$\{a_1,a_2,a_3\}$};
-        \node (c2b) at (0,-2) {$\{b_1,b_2,b_3\}$};
-        \node (c2c) at (2,-2) {$\{c_1,c_2,c_3\}$};
-        \node (c2ab1) at (-5.6,-2) {$\{a_1,b_1\}$};
-        \node (c2ab2) at (-4.5,-2) {$\{a_2,b_2\}$};
-        \node (c2ab3) at (-3.4,-2) {$\{a_3,b_3\}$};
-        \node (c2bc1) at (5.6,-2) {$\{b_1,c_1\}$};
-        \node (c2bc2) at (4.5,-2) {$\{b_2,c_2\}$};
-        \node (c2bc3) at (3.4,-2) {$\{b_3,c_3\}$};
+        \node[b] (c0) at (0,0) {$\{a_1,a_2,a_3,b_1,b_2,b_3,c_1,c_2,c_3\}$};
+        \node[b] (c1a) at (-2,-1) {$\{a_1,a_2,a_3,b_1,b_2,b_3\}$};
+        \node[b] (c1b) at (2,-1) {$\{b_1,b_2,b_3,c_1,c_2,c_3\}$};
+        \node[b] (c2a) at (-2,-2) {$\{a_1,a_2,a_3\}$};
+        \node[b] (c2b) at (0,-2) {$\{b_1,b_2,b_3\}$};
+        \node[b] (c2c) at (2,-2) {$\{c_1,c_2,c_3\}$};
+        \node[b] (c2ab1) at (-5.6,-2) {$\{a_1,b_1\}$};
+        \node[b] (c2ab2) at (-4.5,-2) {$\{a_2,b_2\}$};
+        \node[b] (c2ab3) at (-3.4,-2) {$\{a_3,b_3\}$};
+        \node[b] (c2bc1) at (5.6,-2) {$\{b_1,c_1\}$};
+        \node[b] (c2bc2) at (4.5,-2) {$\{b_2,c_2\}$};
+        \node[b] (c2bc3) at (3.4,-2) {$\{b_3,c_3\}$};
 
@@
     \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
+    \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
     \tikzstyle{e}=[]
 
-    \node[v] (v1) at (0,0.866) {$a_1$};
-    \node[v] (v2) at (-1,-0.866) {$a_2$};
-    \node[v] (v3) at (1,-0.866) {$a_3$};
+    \node[circle,fill=white,inner sep=7] (center) at (0,0-.125-.1) {};
 
-    \node[v] (v4) at (0,0.433-.125) {$b_1$};
-    \node[v] (v5) at (-0.5,-0.433-.125) {$b_2$};
-    \node[v] (v6) at (0.5,-0.433-.125) {$b_3$};
+    \node[c] (v1) at (0,0.866) [label={left,inner sep=.555}:$a_1$]{};
+    \node[c] (v2) at (-1,-0.866) [label={below,inner sep=.555}:$a_2$]{};
+    \node[c] (v3) at (1,-0.866) [label={above,inner sep=.555}:$a_3$]{};
+
+    \node[c] (v4) at (0,0.433-.125) [label={left,inner sep=.555}:$b_1$]{};
+    \node[c] (v5) at (-0.5,-0.433-.125) [label={below,inner sep=.555}:$b_2$]{};
+    \node[c] (v6) at (0.5,-0.433-.125) [label={above,inner sep=.555}:$b_3$]{};
 
@@
     \tedge{v3}{v6}{solid}{}{};
+
+    \tedge{v4}{center}{dashed}{}{};
+    \tedge{v5}{center}{dashed}{}{};
+    \tedge{v6}{center}{dashed}{}{};
 
@@
 
-    \coordinate (ib0) at (-0.167,-0.433-.125-0.1);
-    \coordinate (ib1) at (0.167,-0.433-.125-0.1);
-    \coordinate (ib2) at (-0.167,-0.433-.125+0.1);
-    \coordinate (ib3) at (0.167,-0.433-.125+0.1);
+    \draw[rotate around={60:(-1,-0.866)}] (v2) -- (-0.333,-0.766) -- (0.333,-0.766) -- (v1);
+    \draw[rotate around={60:(-1,-0.866)}]  (v2) -- (-0.333,-0.966) -- (0.333,-0.966) -- (v1);
+    \draw[rotate around={60:(-1,-0.866)}]  (-0.333,-0.766) -- (0.333,-0.966);
+    \draw[rotate around={60:(-1,-0.866)}]  (-0.333,-0.966) -- (0.333,-0.766);
+
+    \draw[rotate around={-60:(1,-0.866)}] (v3) -- (0.333,-0.766) -- (-0.333,-0.766) -- (v1);
+    \draw[rotate around={-60:(1,-0.866)}]  (v3) -- (0.333,-0.966) -- (-0.333,-0.966) -- (v1);
+    \draw[rotate around={-60:(1,-0.866)}]  (-0.333,-0.766) -- (0.333,-0.966);
+    \draw[rotate around={-60:(1,-0.866)}]  (-0.333,-0.966) -- (0.333,-0.766);
+
+    \coordinate (ib0) at (-0.167,-0.433-.125-0.1); %(-.167,-.658)
+    \coordinate (ib1) at (0.167,-0.433-.125-0.1); %(.167,-.658)
+    \coordinate (ib2) at (-0.167,-0.433-.125+0.1);%(-.167,-.458)
+    \coordinate (ib3) at (0.167,-0.433-.125+0.1);%(.167,-.458)
     \draw (v5) -- (ib0) -- (ib1) -- (v6);
@@
 
-    % \draw[rotate around={60:(-1,-0.866)}] (v2) -- (-0.333,-0.766) -- (0.333,-0.766) -- (v1);
-    % \draw (v2) -- (ob2) -- (ob3) -- (v1);
-    % \draw (ob0) -- (ob3);
-    % \draw (ob2) -- (ob1);
+    \draw[rotate around={60:(-0.5,-0.558)}] (v5) -- (-.167,-.658) -- (.167,-.658) -- (v4);
+    \draw[rotate around={60:(-0.5,-0.558)}]  (v5) -- (-.167,-.458) -- (.167,-.458) -- (v4);
+    \draw[rotate around={60:(-0.5,-0.558)}]  (-.167,-.658) -- (.167,-.458);
+    \draw[rotate around={60:(-0.5,-0.558)}]  (-.167,-.458) -- (.167,-.658);
+
+    \draw[rotate around={-60:(0.5,-0.558)}] (v6) -- (.167,-.658) -- (-.167,-.658) -- (v4);
+    \draw[rotate around={-60:(0.5,-0.558)}]  (v6) -- (.167,-.458) -- (-.167,-.458) -- (v4);
+    \draw[rotate around={-60:(0.5,-0.558)}]  (-.167,-.658) -- (.167,-.458);
+    \draw[rotate around={-60:(0.5,-0.558)}]  (-.167,-.458) -- (.167,-.658);
 % \newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
@@
 
-\subsection{Extensions}
-This framework immediately pushes through for body-pin systems via a simple reduction.
+% \subsection{Extensions}
+% This framework immediately pushes through for body-pin systems via a simple reduction.
 If there are $N$ pins on a body, it can be represented as a 2-tree with $N$ vertices, each corresponding to a pin, making sure to select edge distances such that the distance between pins is preserved.
@@
 
-With more effort, it can be shown that pinned line-incidence systems can also use this framework. This is done in section \ref{XXX}.
+% With more effort, it can be shown that pinned line-incidence systems can also use this framework. This is done in section \ref{XXX}.
 

### 2014-12-16 doc/GMP2015/s07_problempinning.tex
Fixed a few typos

@@
 % \end{subfigure}
-\caption{(a) is an orignal body-pin qusec.
+\caption{(a) is a 1 dof body-pin qusec.
 (b) is the corresponding body-bar qusec explained in Definition \ref{def:body-pin}.
@@
 \begin{theorem}
-    There is a quadratic algorithm for the 1 dof and 2 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
+    There is a quadratic algorithm for the 1 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
 \end{theorem}
@@
 
-    For the 2 dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
+    For the multi-triangle pin qusecs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
+
+    % For the 2 dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
 These can be the same node, and if it is the same node, we can find a wellconstrained DR plan in quadratic time, we would just be doing the same thing as the 1 dof case.
 \end{proof}
+
+\begin{theorem}
+    There is a \textcolor{red}{$O(?)$} algorithm for the 1 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
+\end{theorem}
+
+\begin{proof}
+    We can do a similar algorithm to the 1 dof case, except we constrain a given node 2 times (by adding 2 triangles or 2 bodies and 4 pins).
+This causes the nodes along the path to the given node to become well-constrained while all others will be underconstrained.
+Thus, our DR plan will have the same size as the 1 dof case.
+This is still a quadratic algorithm
+\end{proof}
+
+An open question remains as to the complexity of the optimal completion problem when there are more than 2 dofs. For higher number of dofs, the $(l,k)$ characterization is no longer matroidal.

### 2014-12-16 doc/GMP2015/s07_problempinning.tex
Added some figures to Problem Pinning. Updated Prose

@@
-\section{Problem Pinning}
+\section{Application: Finding extensions of underconstrained glassy structures to isostatic}
 \label{sec:bodypin}
 
-We can use DR plans to help in solving questions in designs for graphene and silica bilayers (\todo{Add Citation}).
+We can use DR-plans to help in solving questions in designs for graphene and silica bilayers (\todo{Add Citation}).
 Specifically, we will look into the problem when you have a monolayer and it is underconstrained.
-To make this thing isostatic, there are a few methods we can apply.
+Our goal is to make such a layer isostatic.
+This can be done in a number of ways.
 
@@
     \caption{Example of 1 layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.}
+    \label{fig:silica_glass}
 \end{figure}
@@
     \item Pin the boundary in such a way that the resulting system is isostatic.
-    \item Apply a self-similarity constraint and essentially wrap the monolayer and pin it to itself (pin one edge to another in certain places)
+    \item Build the structures by repeated subdivision to ensure they are isostatic at each layer (see Figure \ref{fig:subdivision})
 \end{itemize}
 
-In each case, we are specifically interested in how to pin these things so that the resultiing structure also has a small DR plan.
+We will discuss item 2 in detail in this section.
+In each case, we are specifically interested in how to they can be applied to get a resulting isostatic structure with a small DR-plan, so a realization can be found quickly.
+To answer these questions, we first inroduce the structures we model the problems with:
+
+\begin{figure}\centering
+    \begin{subfigure}{0.25\linewidth}
+        \includegraphics[width=\linewidth]{img/pentl1}
+    \end{subfigure}
+    \begin{subfigure}{0.25\linewidth}
+        \includegraphics[width=\linewidth]{img/pentl2}
+    \end{subfigure}
+    \begin{subfigure}{0.4\linewidth}
+        \includegraphics[width=\linewidth]{img/pentl3}
+    \end{subfigure}
+    \caption{An example of self-similarity in the form of repeated subdivisions.
+Each figure follows fromt he previous due to a set subdivision rule.
+If each figure was guaranteed to be isostatic, then isostatic systems of arbitrary detail could be made.}
+    \label{fig:subdivision}
+\end{figure}
 
 \subsection{Body-Pin}
-
-To answer these questions, we first inroduce the structures we model the problems with:
 
@@
     A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying.
-As such, the DR planning discussed in Section \ref{dr} is unchanged and the work in Section \ref{recomb} will still go through with slight modifications.
+As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
 \end{remark}
@@
 In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
-When looking for a DR plan, we treat each body as trivial, so they become the leaves of our plan.
-The optimal completion problem and approach of Section \ref{dr} are unchanged.
-The optimal paramaterization problem in Section \ref{dr} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
+The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged.
+The optimal paramaterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
 
-In this section we will go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
+We now go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
 
@@
     \end{itemize}
-    Such a body-pin qusec can also be seen as a {\em body-bar qusec} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
-For such body-bar qusecs, graphs with 1 and 2 dofs can be characterized by being $(3,4)$ and $(3,5)$ tight respectively.
+    Such a body-pin qusec can also be seen as a {\em body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
+For such body-bar qusecs, graphs with 1 and 2 dofs can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
 \end{definition}
@@
     A multi-triangle-pin qusec is a multi-body-pin qusec where each body is a triangle, meaning it is pinned in at most 3 places.
-This is also represented as a hyper qusec where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hyper qusecs, 1 and 2 dofs can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
+This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
+For such hyper-graphs, 1 and 2 dofs can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
 \end{definition}
 
-Body-pin qusecs are of particular interest to us in the context of Example 4 in Section \ref{intro}.
-Multi-triangle-pin qusec can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{intro}, where each triangle is the junction of ``disks'' in the plane.
-Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR plan:
+Body-pin qusecs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
+Multi-triangle-pin qusecs can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (se Figure \ref{fig:silica_glass}).
+Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan:
 
 \begin{definition}
-    A $(k,l)$ tight DR plan is one in which each child node is either a vertex maximal proper $(k,l)$ tight subgraph of the parent node or it is trivial.
+    A $(k,l)$-tight DR-plan is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
 In our case, the trivial nodes are just the bodies.
@@
 
-We can develop the notion of a canonical DR plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{sec:DRP}.
+We can develop the notion of a canonical DR-plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{sec:DRP}.
 
@@
 \label{rem:1dofcanon}
-    For the 1 dof multi-body-pin qusecs described above, a DR plan node will follow one of the following:
+    For the 1 dof multi-body-pin qusecs described above, a DR-plan node will follow one of the following:
 
     \begin{itemize}
-        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on a single body
+        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on another 1 dof graph
         \item Its children will be a number of proper maximal 1 dof subgraphs, joined pairwise by at most one pin
@@
     For item 2, consider the case where we have more than 2 proper vertex maximal 1 dof subgraphs $s_1, ..., s_k, k > 2$.
-Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$ tight and hence $k_i$ and $k_j$ are not vertex maximal.
+Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
 \end{proof}
@@
 \begin{remark}
-    All $(k,l)$ tight canonical DR plans are optimal. We can find such a DR plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{sec:DRP}.
+    All $(k,l)$-tight canonical DR-plans are optimal. We can find such a DR-plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{sec:DRP}.
 \end{remark}
+
 \begin{figure*}\centering
@@
 \caption{(a) is a 1 dof body-pin qusec.
-(b) is the corresponding body-bar qusec explained in Definition \ref{def:body-pin}.
-(c) is the 1 dof DR plan for the qusec.
-In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the node in the second level.
-(d) is the result of adding one such body tot he bold-faced node.}
+(b) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}.
+(c) is the 1 dof DR-plan for the graph.
+In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level.
+(d) is the result of adding one such body to the bold-faced node.}
 
@@
 \begin{proof}
-    Suppose we are given a body-pin qusec $G$ and have obtained the DR plan denoted $T$.
+    Suppose we are given a body-pin qusec and its corresponding body-bar graph $G$ and have obtained the DR-plan $T$.
 Each node of $T$ will then be a vertex maximal proper 1 dof subgraph of $G$.
-To make the entire graph wellconstrained, we can add a body and pin it to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$.
-From Remark \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a single body, so we simply have to pin the new body to one part on each side of the pin or body.
-Doing so will ensure the following
+To make the entire graph wellconstrained, we can add a body (vertex) and pin it (add a 2-multiedge) to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$.
+From Remark \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a subgraph, so we simply have to pin the new body to the 2 bodies joined by the pin or that are not in the intersecting graph.
+Doing so will ensure the following:
 
@@
 
-    Then, we can form a valid isostatic DR plan $T_b$ from $T$.
+    Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
 In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
 Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$ excluding $b'$.
-Then, for any node $b$ that we choose, $T_b$ is a valid DR plan.
-Thus, if we want to minimize the size of our DR plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
+Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
+Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
 Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
@@
     % For the 2 dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
-These can be the same node, and if it is the same node, we can find a wellconstrained DR plan in quadratic time, we would just be doing the same thing as the 1 dof case.
+These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1 dof case.
 \end{proof}
+
+% THIS NEEDS SOME MORE THINKING. THERE ARE DIFFERENT TYPES OF CHILDREN NODES FOR THIS PART. NOT JUST THE TWO LISTED ABOVE
 
@@
 This causes the nodes along the path to the given node to become well-constrained while all others will be underconstrained.
-Thus, our DR plan will have the same size as the 1 dof case.
-This is still a quadratic algorithm
+Thus, our DR-plan will have the same size as the 1 dof case.
+This is still a quadratic algorithm.
+
 \end{proof}

### 2014-12-16 script/subdivision programs by William Floyd/subdivide.c

@@ -13,12 +13,12 @@
  * 
  * This is free software and may be freely copied, modified, and
  * redistributed as noted below.
- *	The copyright notice must remain intact.
- *	Modifications must include a notice giving the reason
- * 	for the modification and including name and date.
- *	This is provided "as is" and comes with no warranty or
- * 	guarantee of fitness. Comments, bugs, and suggestions may be sent
- * 	to floyd@math.vt.edu.
+ *  The copyright notice must remain intact.
+ *  Modifications must include a notice giving the reason
+ *  for the modification and including name and date.
+ *  This is provided "as is" and comes with no warranty or
+ *  guarantee of fitness. Comments, bugs, and suggestions may be sent
+ *  to floyd@math.vt.edu.
  *
  */
 #include <stdio.h>

### 2014-12-16 doc/GMP2015/img/tikz/omd_k33_example-1.tikz.tex
Added figures to intro. Updated tikz figs, put them in separate files.

--- 2014-12-03 doc/GMP2015/s04_recombination.tex
@@
-\section{Recombination}
+% \newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
 
-Discussion about algorithm.
+\begin{tikzpicture}
+    \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
+    \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
+    \tikzstyle{e}=[]
 
-Drop an appropriate edge. Add an appropriate non-edge to get tree-decomposable graph. Change length of edge until dropped edge is correct length (as supplied as input).
+    \node[c] (v0) at (  0,   0) [label=below:$v_0$]{};
+    \node[c] (v1) at (  0,   1) [label={above left,inner sep=0pt,outer sep=0pt}:$v_1$]{};
+    \node[c] (v2) at (  1,   1) [label={above right,inner sep=0pt,outer sep=0pt}:$v_2$]{};
+    \node[c] (v3) at (  1,   0) [label=below:$v_3$]{};
+    \node[c] (v4) at (1.5, 1.5) [label=above:$v_4$]{};
+    \node[c] (v5) at (-.5, 1.5) [label=above:$v_5$]{};
+
+    \tedge{v3}{v2}{solid}{}{};
+    \tedge{v2}{v1}{solid}{}{};
+    \tedge{v1}{v0}{solid}{}{};
+    \tedge{v0}{v5}{solid}{}{};
+    \tedge{v5}{v4}{solid}{}{};
+    \tedge{v4}{v3}{solid}{}{};
+
+    \tedge{v3}{v0}{solid}{}{};
+    \tedge{v2}{v5}{dashed}{$e$}{above, pos=.35};
+    \tedge{v1}{v4}{solid}{}{};
+\end{tikzpicture}

### 2014-12-16 doc/GMP2015/s01_introduction.tex

@@
 Examples of such materials (See Figure \ref{examples}) include:
+
+\begin{figure*}\centering
+\begin{subfigure}{.31\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/Axopodium_Mikrotubuli}
+  \caption{Cross section of a Heliozoa's psuedopod, formed by a spiral structure of microtubules.}
+  % https://de.wikipedia.org/wiki/Mikrotubulus#mediaviewer/File:Axopodium_Mikrotubuli.jpg
+  \label{fig:microtubule}
+\end{subfigure}\hfill
+\begin{subfigure}{.31\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/Transverse_Section_Of_Bone}
+  \caption{Cross section of an osteon (the fundamental unit of compact bone) exhibiting its hierarchical structure.}
+  % https://en.wikipedia.org/wiki/Osteon#mediaviewer/File:Transverse_Section_Of_Bone.png
+  \label{fig:osteon}
+\end{subfigure}\hfill
+\begin{subfigure}{.31\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/Rothemund-DNA-SierpinskiGasket}
+  \caption{A DNA array exhibiting the Sierpinski triangle.}
+  % https://en.wikipedia.org/wiki/DNA_nanotechnology#mediaviewer/File:Rothemund-DNA-SierpinskiGasket.jpg
+  \label{fig:sierpinski}
+\end{subfigure}
+\caption{}
+\end{figure*}
 

### 2014-12-16 doc/GMP2015/s07_problempinning.tex
Prettified/Separated the tikz code to match Troy's format

@@
 \begin{subfigure}{0.2\linewidth}\centering
-    \begin{tikzpicture}[scale=0.8]
-        % \tikzstyle{v}=[draw, circle, minimum size=0.1cm, scale=0.7, font=\footnotesize]
-        \tikzstyle{v}=[font=\footnotesize]
-
-        \node[v] (a) at (0,0) {A};
-        \node[v] (b) at (0,1) {B};
-        \node[v] (c) at (1,0) {C};
-        \node[v] (d) at (1,1) {D};
-        \node[v] (e) at (0,2) {E};
-        \node[v] (f) at (1,2) {F};
-        \node[v] (g) at (-1,0) {G};
-        \node[v] (h) at (-1,1) {H};
-
-        \draw (a) to [bend right=10](b);
-        \draw (a) to [bend left=10](b);
-        \draw (a) to [bend right=10](c);
-        \draw (a) to [bend left=10](c);
-        \draw (d) to [bend right=10](c);
-        \draw (d) to [bend left=10](c);
-        \draw (b) to [bend right=10](d);
-        \draw (b) to [bend left=10](d);
-        \draw (b) to [bend right=10](e);
-        \draw (b) to [bend left=10](e);
-        \draw (b) to [bend right=10](h);
-        \draw (b) to [bend left=10](h);
-        \draw (g) to [bend right=10](h);
-        \draw (g) to [bend left=10](h);
-        \draw (g) to [bend right=10](a);
-        \draw (g) to [bend left=10](a);
-        \draw (f) to [bend right=10](e);
-        \draw (f) to [bend left=10](e);
-        \draw (f) to [bend right=10](d);
-        \draw (f) to [bend left=10](d);
-    \end{tikzpicture}
+    \input{img/tikz/bodypin_graph.tikz}
     \caption{}
@@
 \begin{subfigure}{0.3\linewidth}\centering\scriptsize
-    \begin{tikzpicture}
-    [level 1/.style={sibling distance=8em},
-     level 3/.style={sibling distance=1em}, level distance=0.75cm]
-        \tikzstyle{n}=[draw, fill, circle]
-
-        \node (a) at (0,0) {A,B,C,D,E,F,G,H}
-            child { node {A,B,C,D,E,F} 
-                child { node {C,D,E,F}
-                    child {node { }}
-                    child {node {...}}
-                    child {node { }}
-                    child {node { }}
-                    } 
-                child { node (shared) {A,B,C,D}
-                    child {node {A}}
-                    child {node {B}}
-                    child {node {C}}
-                    child {node {D}}
-                    } 
-                }
-            child { node {A,B,C,D,G,H} 
-                child { node { } }
-                child { node {{\bf A,B,G,H}} 
-                    child {node { }}
-                    child {node {...}}
-                    child {node { }}
-                    child {node { }}
-                }
-            };
-
-    \end{tikzpicture}
+    \input{img/tikz/bodypin_drp.tikz}
     \caption{}

### 2014-12-16 doc/GMP2015/s07_problempinning.tex
Adjusted a few more figures, glass in particular

@@
 \begin{figure}\centering
-    \includegraphics[width=0.5\linewidth]{img/Silica}
-    \caption{Example of 1 layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.}
+    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
+    \raisebox{0.5\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
+    \caption{Example of a signle layer of a Silica Oxygen glassy structure.
+This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
+Not shown here are other layers stacked on this one.
+Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in another layer.}
     \label{fig:silica_glass}
@@
 \begin{itemize}
-    \item Pin together 2 underconstrained monolayers in such a way that the resulting bilayer becomes isostatic.
+    \item Pin together 2 underconstrained monolayers in such a way that the resulting bilayer becomes isostatic (see Figure \ref{fig:silica_glass}).
     \item Pin the boundary in such a way that the resulting system is isostatic.
@@
     \end{subfigure}
+    % 
     \begin{subfigure}{0.25\linewidth}
@@
     \end{subfigure}
+    % 
     \begin{subfigure}{0.4\linewidth}
@@
 \end{subfigure}
+%
 \begin{subfigure}{0.2\linewidth}\centering
@@
 \end{subfigure}
+%
 \begin{subfigure}{0.3\linewidth}\centering\scriptsize
@@
 \end{subfigure}
-
-% \begin{subfigure}
-
-% \end{subfigure}
+%
 \caption{(a) is a 1 dof body-pin qusec.

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Added a Basics section (incomplete)

@@
 
-\subsection{Definitions}
+\subsection{Basics}
+%TODO!
 
-% TODO
+%Standard except for generality and ordering and the fact that I want to
+%avoid the use of the word framework if possible. Do everything
+%with constraint systems, linearization and genericity using indeterminates.
 
-% A paragraph of preliminaries such as:
-%     - rigidity matrix R,
-%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
-%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
-% Should precede the definition of
-%     + independent,
-%     + overconstrained (not independent),
-%     - rigid (contains an independent set with sufficiently many eedges/constraints)
-%     - isostatic (wellconstrained/minimally rigid), ,
-%     - flexible (not rigid),
-%     + underconstrained (independent and not rigid).
-% After that, define DR plans etc.
+%--general  geometric constraint system --
 
-In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
+A $d$-dimensional {\it geometric constraint system} in general consists of
+
+- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.
+
+- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint values
+
+In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).
+
+A familiar example of geometric constraint system is {\it linkage (barjoint system)},
+where the underlying constraint graph is a simple graph and
+the constraints are Euclidean distances on the edges of the graph.
+
+%--Give example as
+%linkage(barjoint system) 
+\todo{-- refer to others (citations
++ pinned line incidence and multi-body-pin (in application
+section)}
+
+%--underlying constraint graph or hypergraph (common treatment)
+\todo{???The (hyper)graph of a is sometimes refered to???}
+
+%--linearization: Jacobian of constraint system,
+%rigidity matrix of indeterminates
+
+The algebraic system can usually be linearized at {\it generic} points
+by taking the Jacobian of the system,
+giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
+whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.
+
+%-- give example for the case of
+%linkages
+
+In the case of linkage, 
+let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
+The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
+The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.
+$v$) and zero in the other entries.
+
+%%Let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
+%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
+The row corresponding to the hyperedge $e = \{v_1, v_2, \ldots, v_k\} \in E$ represents the constraint between the vertices $ \{v_1, v_2, \ldots, v_k\}$,
+%has $d$ non-zero indeterminate entries %$\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) 
+%in the $d$ columns corresponding to each vertex $v_i \in e$ and zero in the other entries.
+
+%generic independence, infinitesimal rigidity
+%(based on indeterminates)
+
+A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
+This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.
+
+The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
+if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
+In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
+A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
+
+%--stress vectors and flex vectors
+%(also based on indeterminates)
+
+The left nullspace of the rigidity matrix is called the {\it space of stress}, consisting of the {\it flex vectors}.
+
+%--A generalized version of Asimow Roth (implicit fn thm)
+%generic rigidity equivalent to infinitesimal rigidity (easy
+%when the latter def is based on indeterminates)
+
+ A generalized version of Asimow Roth (implicit fn thm) states that
+ generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
+  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.
+
+%--generic isostaticity (minimal rigidity, wellconstrainedness),
+%flexibility (not rigid), underconstrained (independent
+%and not rigid), overconstrained (dependent -- only briefly
+%mentioned
+%in this paper; all results are about independent - underconstrained or
+%isostatic)
+
+A graph $G = (V,E)$ is:
+\begin{itemize}
+	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
+	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+	\item {\it flexible} if $G$ is not rigid
+    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+    \item \textbf{over-constrained} if it is not independent.
+\end{itemize}
+In this paper all results are about independent - underconstrained or
+isostatic
+
+\todo{--rigidity matroid\\
+--Combinatorial characterization of rigidity matroid
+Example: Laman's theorem for bar-joint (refer to others)
+Lovasz Yemini
+Tay white and whiteley
+(citations)
++ pinned line incidence and multi-body-pin (in application
+section)}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+%The problem is equivalent to combinatorially determining the generic rank of the 3-dimensional bar-joint rigidity matrix of a graph.
+The $d$-dimensional bar-joint rigidity matrix of a graph $G = (V,E)$, denoted $R_d(G)$, is a matrix of indeterminates.
+Let $p_1(v), p_2(v), \ldots, p_d(v)$  represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$of the joint corresponding to a vertex $v \in V$.
+.
+The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
+The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.
+$v$) and zero in the other entries.
+%
+%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be independent (we drop ``bar-joint'' from now on) in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
+This yields the $d$-dimensional generic rigidity matroid associated with a graph $G$.
+The graph is rigid if the number of generically independent rows or the rank of $R_d(G)$ is maximal, i.e., $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
+
+\subsection{DR-plan}
+%\subsection{Definitions}
+%\label{sec:prelim_defs}
+%
+%% TODO
+%
+%% A paragraph of preliminaries such as:
+%%     - rigidity matrix R,
+%%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
+%%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
+%% Should precede the definition of
+%%     + independent,
+%%     + overconstrained (not independent),
+%%     - rigid (contains an independent set with sufficiently many eedges/constraints)
+%%     - isostatic (wellconstrained/minimally rigid), ,
+%%     - flexible (not rigid),
+%%     + underconstrained (independent and not rigid).
+%% After that, define DR plans etc.
+%
+%In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
 $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
@@
 The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
-
-Given a constant $k$, a graph $G$ is:
-\begin{itemize}
-    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
-
-    \item \textbf{over-constrained} if it is not independent.
-
-    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
-
-    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-
-    \item \textbf{under-constrained} if $G$ is independent and not rigid.
-
-    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
-\end{itemize}
-
+%
+%Given a constant $k$, a graph $G$ is:
+%\begin{itemize}
+%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
+%
+%    \item \textbf{over-constrained} if it is not independent.
+%
+%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
+%
+%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+%
+%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+%
+%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
+%\end{itemize}
+%
 \textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.

### 2014-12-16 doc/GMP2015/s08_pinnedline.tex
Added a Basics section (incomplete)

@@
 the structure of  cross-linking microfibrils in collagen and cellulose in biology,
-by modeling these structures as {\em pinned line incidence systems}.
+by modeling these structures as {\it pinned line incidence systems}.
 
-{\em Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
-{\em Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
-Both of these substances consist of a large number {\em microfibrils},
+{\it Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
+{\it Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
+Both of these substances consist of a large number {\it microfibrils},
 %fibril itself straight?
@@
 each of which is cross-linked at 2 places with some two other fibrils,
-where the {\em  cross-linking} is like an incidence constraint that
+where the {\it  cross-linking} is like an incidence constraint that
 the two crosslinked fibrils can slide with each other  \cite{xxx} (see Figure~\ref{fig:crosslink}).
@@
 
-%TODO ???
+\begin{remark}
 The recombination problem for 2-dimensional pinned line incidence systems is trivial.
@@
  the solutions of a well-constrained sub-system will automatically be consistent with solutions of the remaining of the system.
+ \end{remark}

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Cut premilinary basics.

@@
 %TODO!
+
+In this paper, a {\em geometric constraint system} is a multivariate 
+polynomial (usually bilinear or quadratic) system
+$G(x,\delta)=0$, representing constraints with parameters $\delta$ 
+between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
+%
+When the type of constraint (system) is fixed, 
+the system is simply represented as $(G,\delta)$, where 
+$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
+with the vertices $V$ representing the geometric primitives in 
+$\mathbb{R}^2$ and (hyper)edges
+$E$ representing the constraints, each with an associated 
+parameter $\delta$.
+
+For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
+$G=(V,E)$  
+with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
+this represents the distance constraint system $\| x_u -x_v \|_2 =
+\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
+$u\in V$.
+
+In this paper, we will consider 3 types of constraint systems: bar-joint
+as well as {\em pinned line-incidence}, and {\em body-hyperpin}. 
+In all of these cases, a cartesian \emph{realization} or 
+\emph{solution} $G(p)$ of $(G,\delta)$ is 
+an assignment of coordinates or Euclidean transformations (poses), 
+$p: V \rightarrow
+\mathbb{R}^2$ or $\mathbb{R}^3$ to the vertices of $G$ satisfying the 
+constraints with parameters $\delta$, 
+modulo orientation preserving isometries 
+(Euclidean rigid body motions except for pinned line-incidence).
+
+Although the realization space itself depends on the constraint parameters 
+$\delta$, many  relevant {\em generic} properties of 
+the constraint system $G(x,\delta)$
+are properties of the constraint (hyper)graph $G$ 
+and do not depend on $\delta$ (or they hold for all but a measure 
+zero set of $\delta$
+values). Many of these are properties of the jacobian 
+$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of
+$G$} 
+(a matrix of indeterminates).
+For example, the {\em bar-joint rigidity matrix 
+of the graph $G = (V,E)$}
+is a matrix of indeterminates representing the 
+jacobian of the distance map 
+$\| x_u -x_v \|_2$  for $(u,v) \in E$.
+The matrix has one row per edge $(u,v)$ in $E$,  
+with 2 columns for $u$ (resp.
+$v$) containing the 2 coordinate indeterminates 
+for $x_u -x_v$ (resp. $x_v-x_u$), 
+that is 4 non-zero entries per row.
+
+Some of these generic constraint system or 
+(hyper)graph properties are the following ({\bf Note:} we refer to these as
+properties of the constraint system or as properties of the underlying
+(hyper)graph interchangeably).
+%
+A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
+has independent rows
+(i.e, the determinant of some square submatrix is not
+identically zero). 
+It is {\em isostatic (minimally rigid, well-constrained)} 
+if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
+For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
+where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
+The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
+It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.
+
+In this paper, 
+a {\em qusec} is any {\em independent} geometric 
+constraint system of one of the 3
+types mentioned above.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
+%This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.
+
+%The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
+%if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
+%In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
+%A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
+
+%A graph $G = (V,E)$ is:
+%\begin{itemize}
+%	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
+%	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+%	\item {\it flexible} if $G$ is not rigid
+%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+%    \item \textbf{over-constrained} if it is not independent.
+%\end{itemize}
+%In this paper all results are about independent - underconstrained or
+%isostatic
+
+\todo{Where to put these}
+A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
+The left nullspace of the rigidity matrix is called the {\it space of stress}. A vector in the stress space is called a {\it flex vector}.
+
+%--A generalized version of Asimow Roth (implicit fn thm)
+%generic rigidity equivalent to infinitesimal rigidity (easy
+%when the latter def is based on indeterminates)
+
+% A generalized version of Asimow Roth (implicit fn thm) states that
+% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
+%  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.
+
+%\todo{--rigidity matroid\\
+%--Combinatorial characterization of rigidity matroid
+%Example: Laman's theorem for bar-joint (refer to others)
+%Lovasz Yemini
+%Tay white and whiteley
+%(citations)
+%+ pinned line incidence and multi-body-pin (in application
+%section)}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 
-A $d$-dimensional {\it geometric constraint system} in general consists of
-
-- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.
-
-- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint values
-
-In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).
-
-A familiar example of geometric constraint system is {\it linkage (barjoint system)},
-where the underlying constraint graph is a simple graph and
-the constraints are Euclidean distances on the edges of the graph.
-
-%--Give example as
-%linkage(barjoint system) 
-\todo{-- refer to others (citations
-+ pinned line incidence and multi-body-pin (in application
-section)}
+%A $d$-dimensional {\it geometric constraint system} in general consists of
+%
+%- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.
+%- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint value
+%In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).
 
 %--underlying constraint graph or hypergraph (common treatment)
-\todo{???The (hyper)graph of a is sometimes refered to???}
 
@@
 %rigidity matrix of indeterminates
-
-The algebraic system can usually be linearized at {\it generic} points
-by taking the Jacobian of the system,
-giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
-whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.
+%
+%The algebraic system can usually be linearized at {\it generic} points
+%by taking the Jacobian of the system,
+%giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
+%whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.
 
@@
 %linkages
-
-In the case of linkage, 
-let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
-The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
+%
+%In the case of linkage, 
+%let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
+%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
 The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.
@@
 
-A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
-This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.
-
-The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
-if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
-In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
-A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
-
 %--stress vectors and flex vectors
 %(also based on indeterminates)
-
-The left nullspace of the rigidity matrix is called the {\it space of stress}, consisting of the {\it flex vectors}.
-
-%--A generalized version of Asimow Roth (implicit fn thm)
-%generic rigidity equivalent to infinitesimal rigidity (easy
-%when the latter def is based on indeterminates)
-
- A generalized version of Asimow Roth (implicit fn thm) states that
- generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
-  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.
 
@@
 %isostatic)
-
-A graph $G = (V,E)$ is:
-\begin{itemize}
-	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
-	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-	\item {\it flexible} if $G$ is not rigid
-    \item \textbf{under-constrained} if $G$ is independent and not rigid.
-    \item \textbf{over-constrained} if it is not independent.
-\end{itemize}
-In this paper all results are about independent - underconstrained or
-isostatic
-
-\todo{--rigidity matroid\\
---Combinatorial characterization of rigidity matroid
-Example: Laman's theorem for bar-joint (refer to others)
-Lovasz Yemini
-Tay white and whiteley
-(citations)
-+ pinned line incidence and multi-body-pin (in application
-section)}
 

### 2014-12-16 doc/GMP2015/s01_introduction.tex
Completed DRP example 2. Assembled recomb image. Got different images for intro and cleaned up tex file.

--- 2014-12-15 doc/GMP2015/s01_introduction.tex
@@
 
+\newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}
+
 Some properties of natural and engineered materials can be analyzed by
-treating them as stacked two dimensional (2D) layers.
-The structure within each layer could be
-self-similar \cite{Intro1}
-spanning multiple scales,
-generally aperiodic and quasi-uniform within any one scale,
-and consisting of a few repeated motifs appearing in disordered
-arrangements. Each layer is not necessarily planar, i.e.,
-it consists of multiple,
-inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either
-{\sl isostatic or underconstrained, not
-self-stressed} (See Section \ref{sec:prelim} for definitions).
-These properties are
-consistent with a self-assembled 2D structure that minimizes mass,
-optimally distributes external stresses and itself
-participates in the assembly of diverse and multifunctional,
-larger 2D structures.
+treating them as stacked two dimensional (2D) layers. The structure
+within each layer could be self-similar \cite{Intro1} spanning
+multiple scales, generally aperiodic and quasi-uniform within any one
+scale, and consisting of a few repeated motifs appearing in disordered
+arrangements. Each layer is not necessarily planar, i.e., it consists
+of multiple, inter-constraining planar (genus 0) monolayers.
+Furthermore, a layer is often  either {\sl isostatic or
+underconstrained, not self-stressed} (See Section \ref{sec:prelim} for
+definitions). These properties are consistent with a self-assembled 2D
+structure that minimizes mass, optimally distributes external stresses
+and itself participates in the assembly of diverse and
+multifunctional, larger 2D structures.
 
-Examples of such materials (See Figure \ref{examples}) include:
+Examples of such materials (See Figure \ref{fig:material_examples})
+include:
+
+\begin{figure*}\centering
+% \begin{subfigure}{.31\linewidth}\centering
+%   \includegraphics[width=\linewidth]{img/Axopodium_Mikrotubuli}
+%   \caption{Cross section of a Heliozoa's psuedopod, formed by a spiral structure of microtubules.}
+%   % https://de.wikipedia.org/wiki/Mikrotubulus#mediaviewer/File:Axopodium_Mikrotubuli.jpg
+%   \label{fig:material_examples:microtubule}
+% \end{subfigure}\hfill
+% \begin{subfigure}{.31\linewidth}\centering
+%   \includegraphics[width=\linewidth]{img/Transverse_Section_Of_Bone}
+%   \caption{Cross section of an osteon (the fundamental unit of compact bone) exhibiting its hierarchical structure.}
+%   % https://en.wikipedia.org/wiki/Osteon#mediaviewer/File:Transverse_Section_Of_Bone.png
+%   \label{fig:material_examples:osteon}
+% \end{subfigure}\hfill
+\begin{subfigure}{.28\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/Chlamydomonas_TEM_17}
+  \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules.}
+  % https://en.wikipedia.org/wiki/Axoneme#mediaviewer/File:Chlamydomonas_TEM_17.jpg
+  \label{fig:material_examples:microtubule}
+\end{subfigure}\hfill
+\begin{subfigure}{.40\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/ligten2}
+  \caption{Cross section of a tendon displaying the hierarchical structure.}
+  % http://www.umich.edu/~bme332/ch10ligten/ligten2.jpg
+  \label{fig:material_examples:tendon}
+\end{subfigure}\hfill
+\begin{subfigure}{.30\linewidth}\centering
+  \includegraphics[width=\linewidth]{img/Rothemund-DNA-SierpinskiGasket}
+  \caption{A DNA array exhibiting the Sierpinski triangle.}
+  % https://en.wikipedia.org/wiki/DNA_nanotechnology#mediaviewer/File:Rothemund-DNA-SierpinskiGasket.jpg
+  \label{fig:material_examples:sierpinski}
+\end{subfigure}
+\caption{}\label{fig:material_examples}
+\end{figure*}
 
 \begin{enumerate}
-  %
-    \item (cross-section) of microtubules \cite{Necklace1} e.g., in
-ciliary membranes and transitions \cite{Necklace2}
-   %
-    \item (cross-section) of hierarchical bone structure \cite{XX} %
-    \item crosslinked cellulose or collagen microfibril monolayers e.g.,
-in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked
-actin filaments in the cytoskeleton matrix.
-  %
-    \item more recent, engineered examples including  disordered graphene
-layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced
-    by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a
-recent Sierpinski carpet, bringing other self-similar structures
-    \cite{Microfibrils2} within reach.
-    %
-    \item  silica bi-layers, glass \cite{SilicaGlass1}
-\cite{SilicaGlass2}, and materials that behave like assemblies of
-    2D particles under non-overlap constraints, e.g, jammed
-    disks on the plane \cite{JammedDisk1}
-%
+    \item Cross-sections of microtubule structures (Figure \ref{fig:material_examples:microtubule}) \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
+
+    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \cite{XX}
+
+    \item crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+
+    \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.
+
+    \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
 \end{enumerate}
 %
-In order to study structural and mechanical properties
-such as rigidity, flexes and distribution of external stresses
-within the material layer, it is natural to model a material
-layer as a solution of a geometric constraint system
-of appropriate types of geometric primitives, under
-metric or algebraic constraints.
-Such  2D {\bf{\em qusecs}} (quasi-uniform or self-similar
-constraint systems) can be used to
-understand or design material layers (their solutions)
-with desired properties.
+In order to study structural and mechanical properties such as
+rigidity, flexes and distribution of external stresses within the
+material layer, it is natural to model a material layer as a solution
+of a geometric constraint system of appropriate types of geometric
+primitives, under metric or algebraic constraints. Such  2D {\bf{\em
+qusecs}} (quasi-uniform or self-similar constraint systems) can be
+used to understand or design material layers (their solutions) with
+desired properties.
 %
-\subsection{Previous Work on Relevant 2D Geometric Constraint Systems} We
-now briefly survey existing techniques for studying 2D {\em qusecs,}  many
-of which are {\it bar-joint} systems (Examples 1,2 above, see Sections
-\ref{sec:DRP}, \ref{sec:recomb}),
-{\it multibody-pin} systems (Example 4,5, see Section \ref{sec:bodypin}) or  {\it
-pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The
-limitations of these techniques directly motivate
-the contributions of this paper.
+\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
+We now briefly survey existing techniques for studying 2D {\em
+qusecs,}  many of which are {\it bar-joint} systems (Examples 1,2
+above, see Sections \ref{sec:DRP}, \ref{sec:recomb}), {\it multibody-
+pin} systems (Example 4,5, see Section \ref{sec:bodypin}) or  {\it
+pinned-line incidence} systems (Example 3, see Section
+\ref{sec:pinnedline}). The limitations of these techniques directly
+motivate the contributions of this paper.
 
-\medskip\noindent\underbar{{\sl (i) Finding (vertex)-maximal
-generically rigid subsystems}}
-Fast, graph-based algorithms exist (pebble-game \cite{XX}),
-for locating all maximal, generically rigid subsystems \cite{XX}.
-When the input is rigid, these algorithms compute the identity and  output
-the input.
+\meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
+Fast, graph-based algorithms exist (pebble-game \cite{XX}), for
+locating all maximal, generically rigid subsystems \cite{XX}. When the
+input is rigid, these algorithms compute the identity and  output the
+input.
 
 However, both for self-similar or just aperiodic 2D qusecs, it is
-imperative to recursively decompose rigid systems
-into their rigid subsystems, down to the level of
-geometric primitives, in order to
-understand or design properties at all scales, such as:
-rigidity, flexes, distribution
-of external stresses, boundary conditions for isostaticity,
-as well as behavior under constraint variations.
+imperative to recursively decompose rigid systems into their rigid
+subsystems, down to the level of geometric primitives, in order to
+understand or design properties at all scales, such as: rigidity,
+flexes, distribution of external stresses, boundary conditions for
+isostaticity, as well as behavior under constraint variations.
 
-\medskip\noindent\underbar{{\sl (ii) Optimal Recursive Decomposition
-(DR-planning)}}
+\meeraheader{ii}{Optimal Recursive Decomposition (DR-planning)}
 Recursive decomposition of geometric constraint systems has been
-formalized \cite{XX} and well-studied \cite{XX}
-as the {\sl Decomposition-Recombination (DR)-planning} problem.
-For the above-mentioned classes of 2D qusecs, generic rigidity is
-a combinatorial property and hence each level of the decomposition should,
-in principle, be achievable by a graph-based algorithm as in (1), without
-involving geometric information in the constraint system.
-Since many  such decompositions can exist for a given constraint system,
-criteria defining desirable or optimal DR-plans and DR-planning algorithms
-were given in \cite{XX}. An {\em optimal DR-plan} is one
-that minimizes the maximum number of child subsystems of any parent
-system, a property that is
-crucial for obtaining a solution to the parent constraint system from the
-solutions of the child systems.
+formalized \cite{XX} and well-studied \cite{XX} as the {\sl
+Decomposition-Recombination (DR)-planning} problem. For the above-
+mentioned classes of 2D qusecs, generic rigidity is a combinatorial
+property and hence each level of the decomposition should, in
+principle, be achievable by a graph-based algorithm as in (1), without
+involving geometric information in the constraint system. Since many
+such decompositions can exist for a given constraint system, criteria
+defining desirable or optimal DR-plans and DR-planning algorithms were
+given in \cite{XX}. An {\em optimal DR-plan} is one that minimizes the
+maximum number of child subsystems of any parent system, a property
+that is crucial for obtaining a solution to the parent constraint
+system from the solutions of the child systems.
 
-However, for general 2D qusecs, even when restricted to bar-joint systems,
-the optimal DR-planning problem was shown to be NP-hard \cite{XX}.
+However, for general 2D qusecs, even when restricted to bar-joint
+systems, the optimal DR-planning problem was shown to be NP-hard
+\cite{XX}.
 
-\medskip\noindent\underbar{{\sl (iii) DR-plans for special classes and
-with other criteria}}
-For a special class of 2D qusecs, namely  {\em tree-decomposable} systems
-\cite{XX} common in computer aided mechanical design,
-(which includes ruler-and-compass and Henneberg-I constructible systems),
-all DR-plans turn out to
-be optimal. This satisfies the so-called {\em Church-Rosser} criterion,
-leading to highly efficient DR-planning algorithms.
-For general 2D qusecs, alternate criteria were suggested
-such as {\em cluster minimality}
-requiring parent systems to be composed of
-a minimal set of at least 2 rigid proper subsystems (i.e., no proper
-subset forms a rigid system);
-and {\em proper maximality}, requiring child subsystems
-to be maximal rigid proper subsystems of the parent system.
+\meeraheader{iii}{DR-plans for special classes and with other
+criteria}
+For a special class of 2D qusecs, namely  {\em tree-decomposable}
+systems \cite{XX} common in computer aided mechanical design, (which
+includes ruler-and-compass and Henneberg-I constructible systems), all
+DR-plans turn out to be optimal. This satisfies the so-called {\em
+Church-Rosser} criterion, leading to highly efficient DR-planning
+algorithms. For general 2D qusecs, alternate criteria were suggested
+such as {\em cluster minimality} requiring parent systems to be
+composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
+no proper subset forms a rigid system); and {\em proper maximality},
+requiring child subsystems to be maximal rigid proper subsystems of
+the parent system.
 
-While polynomial time algorithms were given \cite{XX} to generate DR-plans
-meeting the cluster minimality criterion,
-no such algorithm is known for the latter criterion.
+While polynomial time algorithms were given \cite{XX} to generate DR-
+plans meeting the cluster minimality criterion, no such algorithm is
+known for the latter criterion.
 
-\medskip\noindent\underbar{{\sl (iv) Optimal Recombination and Solution
-Space Navigation}}
-For the entire DR-plan, finding all
-desired solutions is barely tractable even if recombination
-of solved subsystems is easy for each indecomposable parent system in the
-DR-plan.
-This is because even for the simplest, highly decomposable systems
-with small DR-plans, the problem of finding even a single
-solution to the input system at the root of the DR-plan
-is NP-hard \cite{XX} and there is a
-combinatorial explosion of solutions \cite{XX}. Typically, however,
-the desired solution has a given orientation type, in which case,
-the crux of the difficulty is concentrated in the algebraic complexity of
-(re)combining child system solutions, to give a
-solution to the parent system at any
-given level of the DR-plan.
-For fairly general 3D constraint systems,
-there are algorithms with formal
-guarantees that uncover underlying matroids to combinatorially
-obtain an optimal parameterization to minimize the algebraic
-complexity of the indecomposable parent (sub)systems that occur in the
-DR-plan \cite{XX, XX,XX},
-provided the DR-plan meets some of the abovementioned criteria.
+\meeraheader{iv}{Optimal Recombination and Solution Space Navigation}
+For the entire DR-plan, finding all desired solutions is barely
+tractable even if recombination of solved subsystems is easy for each
+indecomposable parent system in the DR-plan. This is because even for
+the simplest, highly decomposable systems with small DR-plans, the
+problem of finding even a single solution to the input system at the
+root of the DR-plan is NP-hard \cite{XX} and there is a combinatorial
+explosion of solutions \cite{XX}. Typically, however, the desired
+solution has a given orientation type, in which case, the crux of the
+difficulty is concentrated in the algebraic complexity of
+(re)combining child system solutions, to give a solution to the parent
+system at any given level of the DR-plan. For fairly general 3D
+constraint systems, there are algorithms with formal guarantees that
+uncover underlying matroids to combinatorially obtain an optimal
+parameterization to minimize the algebraic complexity of the
+indecomposable parent (sub)systems that occur in the DR-plan \cite{XX,
+XX,XX}, provided the DR-plan meets some of the abovementioned
+criteria.
 
-However, the generality of these algorithms
-trades-off against efficiency, and despite the optimization,
-the best algorithms can still take
-exponential time in the number of child subsystems (which can be arbitrarily
-large even for optimal DR-plans) in order to guarantee
-all solutions of a given orientation type,
-even for a single (sub)system in a DR-plan. They
-are prohibitively slow in practice. We note that utilizing the DR-plan and
-optimal recombination as a principled basis, high performance heuristics
-and software
-exists \cite{XX,XX} to
-tame combinatorial explosion via user intervention.
+However, the generality of these algorithms trades-off against
+efficiency, and despite the optimization, the best algorithms can
+still take exponential time in the number of child subsystems (which
+can be arbitrarily large even for optimal DR-plans) in order to
+guarantee all solutions of a given orientation type, even for a single
+(sub)system in a DR-plan. They are prohibitively slow in practice. We
+note that utilizing the DR-plan and optimal recombination as a
+principled basis, high performance heuristics and software exists
+\cite{XX,XX} to tame combinatorial explosion via user intervention.
 
-\medskip\noindent\underbar{{\sl (v) Configuration Spaces of
-Underconstrained Systems}}
-For underconstrained 2D bar-joint and multibody-pin cusecs obtained from
-various subclasses of tree-decomposable systems, algorithms have been
-developed \cite{XX, XX} that
-to complete them into isostatic systems \cite{XX, XX} and to find paths
-within the connected components \cite{XX,XX} of
-standard cartesian configuration spaces. Most of the algorithms with
-formal guarantees leverage Cayley configuration space theory \cite{XX,XX}
-to characterize subclasses of graphs and additional constraints that
-control combinatorial explosion, and provide faithful bijective
-representation
-of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems
-gets bigger,  with highest efficiency for
-underlying partial 2-tree graphs (alternately, tree-width 2,
-series-parallel, $K_4$ minor avoiding),   moderate efficiency for
-1 degree-of-freedom (dof) graphs with low Cayley complexity (which include
+\meeraheader{v}{Configuration Spaces of Underconstrained Systems}
+For underconstrained 2D bar-joint and multibody-pin cusecs obtained
+from various subclasses of tree-decomposable systems, algorithms have
+been developed \cite{XX, XX} that to complete them into isostatic
+systems \cite{XX, XX} and to find paths within the connected
+components \cite{XX,XX} of standard cartesian configuration spaces.
+Most of the algorithms with formal guarantees leverage Cayley
+configuration space theory \cite{XX,XX} to characterize subclasses of
+graphs and additional constraints that control combinatorial
+explosion, and provide faithful bijective representation of connected
+components and paths. These algorithms have decreasing efficiency as
+the subclass of systems gets bigger,  with highest efficiency for
+underlying partial 2-tree graphs (alternately, tree-width 2, series-
+parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-
+of-freedom (dof) graphs with low Cayley complexity (which include
 common linkages such as the Strandbeest, Limacon and Cardioid), and
@@
 \label{sec:cont}
+
 The contributions of this paper are the following.
@@
 \item
-We  navigate the NP-hardness barrier mentioned in (2) above, for finding
-optimal DR-plans by defining a
-so-called {\it canonical} DR-plan and showing a strong Church-Rosser
-property: {\it all canonical DR-plans for isostatic or
-underconstrained
-2D qusecs are optimal} (Section \ref{sec:DRP}).
+We navigate the NP-hardness barrier mentioned in (2) above, for
+finding optimal DR-plans by defining a so-called {\it canonical} DR-
+plan and showing a strong Church-Rosser property: {\it all canonical
+DR-plans for isostatic or underconstrained 2D qusecs are optimal}
+(Section \ref{sec:DRP}).
+
 \item
-We give an efficient ($O(n^2)$) algorithm to find a canonical (and hence
-optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections
-\ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}.
-The canonical DR-plan
-illuminates the essence of the NP-hardness of finding optimal DR-plans for
-overconstrained systems.
+We give an efficient ($O(n^2)$) algorithm to find a canonical (and
+hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above
+(Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}.
+The canonical DR-plan illuminates the essence of the NP-hardness of
+finding optimal DR-plans for overconstrained systems.
+
 \item
-We give a method to deal with the
-algebraic complexity of recombining the realizations or
-solutions of  child subsystems into a solution of the parent system.
-Specifically, we define the problem of  minimally modifying the
-indecomposable recombination system so that it becomes decomposable via a
- small DR-plan and yet preserves the original
-solutions in an efficiently searchable manner. We show formal connection
-to well known problems such as optimal completion of
+We give a method to deal with the algebraic complexity of recombining
+the realizations or solutions of  child subsystems into a solution of
+the parent system. Specifically, we define the problem of  minimally
+modifying the indecomposable recombination system so that it becomes
+decomposable via a small DR-plan and yet preserves the original
+solutions in an efficiently searchable manner. We show formal
+connection to well known problems such as optimal completion of
 underconstrained  systems  \cite{XX}.  When the required modifications
-are bounded, we obtain new, efficient algorithms for
-realizing both isostatic and underconstrained
-qusecs by leveraging recent results about Cayley parameters in
-\cite{XX,XX}  (Section \ref{sec:recomb}).
+are bounded, we obtain new, efficient algorithms for realizing both
+isostatic and underconstrained qusecs by leveraging recent results
+about Cayley parameters in \cite{XX,XX}  (Section \ref{sec:recomb}).
+
 \item
 We briefly describe applications of the above techniques to modeling,
-analyzing and designing specific properties in 2D material layers.
-For Examples 1,2 (achieving isostaticity, distribution of stresses in
-self-similar, bar-joint systems); Example 3 (canonical and optimal
-DR-plans for pinned line incidence systems \cite{XX}) and Examples 4,5
-(boundary-conditions for achieving various desired properties
-of multi-body  pin systems).
+analyzing and designing specific properties in 2D material layers. For
+Examples 1,2 (achieving isostaticity, distribution of stresses in
+self-similar, bar-joint systems); Example 3 (canonical and optimal DR-
+plans for pinned line incidence systems \cite{XX}) and Examples 4,5
+(boundary-conditions for achieving various desired properties of
+multi-body  pin systems).
 
 \item
-Software implementation and videos can be found at: \todo{Provide Link}.
+Software implementation and videos can be found at: \todo{Provide
+Link}.
 \end{itemize}

### 2014-12-16 doc/GMP2015/s03_drp.tex
Added figures to intro. Updated tikz figs, put them in separate files.
Really added figures this time
Completed DRP example 2. Assembled recomb image. Got different images for intro and cleaned up tex file.

@@
 \begin{proof}
+%I wouldn't do the forward and backward direction italics, or say proof by contradiction.
+Just start the first paragraph with Suppose C_i union C_j is well constained maximal proper and assume it is not C.
 \textit{Forward direction:} Proof by contradiction.
@@
 
-We now present two examples of non-tree-decomposable graphs that illustrate the two cases found in decomposition -- intersection on a well-constrained subgraph and intersection on a trivial subgraph.
+We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.
 
@@
 \begin{subfigure}{.3\linewidth}\centering
-  \begin{tikzpicture}[scale=2]
-        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
-        \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
-        \tikzstyle{e}=[]
-
-        \node[c] (v1) at (0,0.866) [label={left,inner sep=.555}:$a_1$]{};
-        \node[c] (v2) at (-1,-0.866) [label={below,inner sep=.555}:$a_2$]{};
-        \node[c] (v3) at (1,-0.866) [label={above,inner sep=.555}:$a_3$]{};
-
-        \node[c] (v4) at (0,0.577-.1) [label={left,inner sep=.555}:$b_1$]{};
-        \node[c] (v5) at (-0.667,-0.577-.1) [label={below,inner sep=.555}:$b_2$]{};
-        \node[c] (v6) at (0.667,-0.577-.1) [label={above,inner sep=.555}:$b_3$]{};
-
-        \node[c] (v7) at (0,0.289-.2) [label={left,inner sep=.555}:$c_1$]{};
-        \node[c] (v8) at (-0.333,-0.289-.2) [label={below,inner sep=.555}:$c_2$]{};
-        \node[c] (v9) at (0.333,-0.289-.2) [label={above,inner sep=.555}:$c_3$]{};
-
-        \tedge{v1}{v2}{solid}{}{};
-        \tedge{v1}{v3}{solid}{}{};
-        \tedge{v2}{v3}{solid}{}{};
-
-        \tedge{v4}{v5}{solid}{}{};
-        \tedge{v4}{v6}{solid}{}{};
-        \tedge{v5}{v6}{solid}{}{};
-
-        \tedge{v7}{v8}{solid}{}{};
-        \tedge{v7}{v9}{solid}{}{};
-        \tedge{v8}{v9}{solid}{}{};
-
-        \tedge{v1}{v4}{solid}{}{};
-        \tedge{v2}{v5}{solid}{}{};
-        \tedge{v3}{v6}{solid}{}{};
-
-        \tedge{v4}{v7}{solid}{}{};
-        \tedge{v5}{v8}{solid}{}{};
-        \tedge{v6}{v9}{solid}{}{};
-    \end{tikzpicture}
-
-    \caption{}
+    \input{img/tikz/two_doublets-0.tikz}
+    \caption{}\label{fig:doublets:a}
 \end{subfigure}%
 \begin{subfigure}{.7\linewidth}\centering
-    \begin{tikzpicture}[scale=.9, transform shape]
-        \tikzstyle{v}=[draw, circle, minimum size=0.75cm, font=\footnotesize]
-        % \tikzstyle{b}=[draw, font=\footnotesize]
-        \tikzstyle{b}=[]
-        \tikzstyle{e}=[]
-
-        \node[b] (c0) at (0,0) {$\{a_1,a_2,a_3,b_1,b_2,b_3,c_1,c_2,c_3\}$};
-        \node[b] (c1a) at (-2,-1) {$\{a_1,a_2,a_3,b_1,b_2,b_3\}$};
-        \node[b] (c1b) at (2,-1) {$\{b_1,b_2,b_3,c_1,c_2,c_3\}$};
-        \node[b] (c2a) at (-2,-2) {$\{a_1,a_2,a_3\}$};
-        \node[b] (c2b) at (0,-2) {$\{b_1,b_2,b_3\}$};
-        \node[b] (c2c) at (2,-2) {$\{c_1,c_2,c_3\}$};
-        \node[b] (c2ab1) at (-5.6,-2) {$\{a_1,b_1\}$};
-        \node[b] (c2ab2) at (-4.5,-2) {$\{a_2,b_2\}$};
-        \node[b] (c2ab3) at (-3.4,-2) {$\{a_3,b_3\}$};
-        \node[b] (c2bc1) at (5.6,-2) {$\{b_1,c_1\}$};
-        \node[b] (c2bc2) at (4.5,-2) {$\{b_2,c_2\}$};
-        \node[b] (c2bc3) at (3.4,-2) {$\{b_3,c_3\}$};
-
-        \tedge{c0}{c1a}{solid}{}{};
-        \tedge{c0}{c1b}{solid}{}{};
-
-        \tedge{c1a}{c2ab1}{solid}{}{};
-        \tedge{c1a}{c2ab2}{solid}{}{};
-        \tedge{c1a}{c2ab3}{solid}{}{};
-        \tedge{c1a}{c2a}{solid}{}{};
-        \tedge{c1a}{c2b}{solid}{}{};
-
-        \tedge{c1b}{c2bc1}{solid}{}{};
-        \tedge{c1b}{c2bc2}{solid}{}{};
-        \tedge{c1b}{c2bc3}{solid}{}{};
-        \tedge{c1b}{c2b}{solid}{}{};
-        \tedge{c1b}{c2c}{solid}{}{};
-    \end{tikzpicture}
-
-    \caption{}
+    \input{img/tikz/two_doublets-1.tikz}
+    \caption{}\label{fig:doublets:b}
 \end{subfigure}
 
-\caption{(a) Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
-(b) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
-\label{doublets}
+\caption{(\ref{fig:doublets:a}) Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
+(\ref{fig:doublets:b}) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
+\label{fig:doublets}
 \end{figure*}
@@
 \begin{example}
-    In figure \ref{doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
+    In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
 These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$.
-This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main_wellconstrained}.
+This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 2.
 Also shown is the decomposition of the individual doublets.
 These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them.
-As explained in Theorem \ref{theorem:main_trivial}, all of them must be included in the optimal DR-plan.
+As explained in Theorem \ref{theorem:main}, point 1, all of them must be included in the optimal DR-plan.
 Were any left out, the union of the children would not be the entire parent.
@@
 \begin{figure*}\centering
-\begin{tikzpicture}[scale=3]
-    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
-    \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
-    \tikzstyle{e}=[]
+\begin{subfigure}{.3\linewidth}\centering
+    \input{img/tikz/c2c3_of_k33s-0.tikz}
+    \caption{}\label{fig:c2c3ofk33s:a}
+\end{subfigure}%
+\begin{subfigure}{.7\linewidth}\centering
+    \input{img/tikz/c2c3_of_k33s-1.tikz}
+    \caption{}\label{fig:c2c3ofk33s:b}
+\end{subfigure}
 
-    \node[circle,fill=white,inner sep=7] (center) at (0,0-.125-.1) {};
-
-    \node[c] (v1) at (0,0.866) [label={left,inner sep=.555}:$a_1$]{};
-    \node[c] (v2) at (-1,-0.866) [label={below,inner sep=.555}:$a_2$]{};
-    \node[c] (v3) at (1,-0.866) [label={above,inner sep=.555}:$a_3$]{};
-
-    \node[c] (v4) at (0,0.433-.125) [label={left,inner sep=.555}:$b_1$]{};
-    \node[c] (v5) at (-0.5,-0.433-.125) [label={below,inner sep=.555}:$b_2$]{};
-    \node[c] (v6) at (0.5,-0.433-.125) [label={above,inner sep=.555}:$b_3$]{};
-
-    \tedge{v1}{v2}{solid}{}{};
-    \tedge{v1}{v3}{solid}{}{};
-    \tedge{v2}{v3}{solid}{}{};
-
-    \tedge{v4}{v5}{solid}{}{};
-    \tedge{v4}{v6}{solid}{}{};
-    \tedge{v5}{v6}{solid}{}{};
-
-    \tedge{v1}{v4}{solid}{}{};
-    \tedge{v2}{v5}{solid}{}{};
-    \tedge{v3}{v6}{solid}{}{};
-
-    \tedge{v4}{center}{dashed}{}{};
-    \tedge{v5}{center}{dashed}{}{};
-    \tedge{v6}{center}{dashed}{}{};
-
-    % sin(30deg) = 0.5
-    % cos(30deg) = 0.866
-
-    % o/i -> outside/inside triangle
-    % b/l/r -> bottom/left/right edge of triangle
-
-    \coordinate (ob0) at (-0.333,-0.866-0.1);
-    \coordinate (ob1) at (0.333,-0.866-0.1);
-    \coordinate (ob2) at (-0.333,-0.866+0.1);
-    \coordinate (ob3) at (0.333,-0.866+0.1);
-    \draw (v2) -- (ob0) -- (ob1) -- (v3);
-    \draw (v2) -- (ob2) -- (ob3) -- (v3);
-    \draw (ob0) -- (ob3);
-    \draw (ob2) -- (ob1);
-
-    \draw[rotate around={60:(-1,-0.866)}] (v2) -- (-0.333,-0.766) -- (0.333,-0.766) -- (v1);
-    \draw[rotate around={60:(-1,-0.866)}]  (v2) -- (-0.333,-0.966) -- (0.333,-0.966) -- (v1);
-    \draw[rotate around={60:(-1,-0.866)}]  (-0.333,-0.766) -- (0.333,-0.966);
-    \draw[rotate around={60:(-1,-0.866)}]  (-0.333,-0.966) -- (0.333,-0.766);
-
-    \draw[rotate around={-60:(1,-0.866)}] (v3) -- (0.333,-0.766) -- (-0.333,-0.766) -- (v1);
-    \draw[rotate around={-60:(1,-0.866)}]  (v3) -- (0.333,-0.966) -- (-0.333,-0.966) -- (v1);
-    \draw[rotate around={-60:(1,-0.866)}]  (-0.333,-0.766) -- (0.333,-0.966);
-    \draw[rotate around={-60:(1,-0.866)}]  (-0.333,-0.966) -- (0.333,-0.766);
-
-    \coordinate (ib0) at (-0.167,-0.433-.125-0.1); %(-.167,-.658)
-    \coordinate (ib1) at (0.167,-0.433-.125-0.1); %(.167,-.658)
-    \coordinate (ib2) at (-0.167,-0.433-.125+0.1);%(-.167,-.458)
-    \coordinate (ib3) at (0.167,-0.433-.125+0.1);%(.167,-.458)
-    \draw (v5) -- (ib0) -- (ib1) -- (v6);
-    \draw (v5) -- (ib2) -- (ib3) -- (v6);
-    \draw (ib0) -- (ib3);
-    \draw (ib2) -- (ib1);
-
-    \draw[rotate around={60:(-0.5,-0.558)}] (v5) -- (-.167,-.658) -- (.167,-.658) -- (v4);
-    \draw[rotate around={60:(-0.5,-0.558)}]  (v5) -- (-.167,-.458) -- (.167,-.458) -- (v4);
-    \draw[rotate around={60:(-0.5,-0.558)}]  (-.167,-.658) -- (.167,-.458);
-    \draw[rotate around={60:(-0.5,-0.558)}]  (-.167,-.458) -- (.167,-.658);
-
-    \draw[rotate around={-60:(0.5,-0.558)}] (v6) -- (.167,-.658) -- (-.167,-.658) -- (v4);
-    \draw[rotate around={-60:(0.5,-0.558)}]  (v6) -- (.167,-.458) -- (-.167,-.458) -- (v4);
-    \draw[rotate around={-60:(0.5,-0.558)}]  (-.167,-.658) -- (.167,-.458);
-    \draw[rotate around={-60:(0.5,-0.558)}]  (-.167,-.458) -- (.167,-.658);
-% \newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}
-
-    % \draw (-1,-0.866) -- (-0.333,-0.966);
-
-\end{tikzpicture}
-
-\caption{A doublet with each edge replaced by a $K3,3$. \todo{Drawing incomplete.}}
-\label{c2c3ofk33s}
+\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K3,3$.
+This pattern continues for a total of $N$ triangles.
+(\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K3,3$ subgraphs into the separate 9 edges and of edges into the component nodes.
+$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K3,3$ subgraphs.
+The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
+\label{fig:c2c3ofk33s}
 \end{figure*}
 
-\begin{example}
-    In figure \ref{c2c3ofk33s}... \todo{Is it worth working through this? The first example has trivial and non-trivial.}
-\end{example}
+% \begin{example}
+%     In
+% \end{example}
 

### 2014-12-16 doc/GMP2015/s04_recombination.tex
Added figures to intro. Updated tikz figs, put them in separate files.
Completed DRP example 2. Assembled recomb image. Got different images for intro and cleaned up tex file.

--- 2014-12-16 doc/GMP2015/s04_recombination.tex
@@
 \label{sec:recomb}
+
 In this section, we consider the {\em optimal recombination}
@@
 
+\begin{figure*}\centering
+\begin{subfigure}{.2\linewidth}\centering
+    \input{img/tikz/omd_k33_example_new-0.tikz}
+    \caption{}\label{fig:omd_k33_example:a}
+\end{subfigure}\hfill
+\begin{subfigure}{.15\linewidth}\centering
+    \input{img/tikz/omd_k33_example_new-1.tikz}
+    \caption{}\label{fig:omd_k33_example:b}
+\end{subfigure}
+\begin{subfigure}{.15\linewidth}\centering
+    \input{img/tikz/omd_k33_example_new-2.tikz}
+    \caption{}\label{fig:omd_k33_example:c}
+\end{subfigure}\hfill
+\begin{subfigure}{.15\linewidth}\centering
+    \input{img/tikz/omd_k33_example-1.tikz}
+    \caption{}\label{fig:omd_k33_example:d}
+\end{subfigure}
+\begin{subfigure}{.15\linewidth}\centering
+    \input{img/tikz/omd_k33_example-2.tikz}
+    \caption{}\label{fig:omd_k33_example:e}
+\end{subfigure}
+
+\caption{
+(\ref{fig:omd_k33_example:a}) The $K3,3$ with two labeled edges, $e_1$ and $e_2$.
+(\ref{fig:omd_k33_example:b}) The $K3,3$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree.
+(\ref{fig:omd_k33_example:c}) The $K3,3$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K3,3$ is at least $OMD_2$.
+(\ref{fig:omd_k33_example:d}) The $K3,3$ with only $e_2$ removed (dashed line).
+(\ref{fig:omd_k33_example:e}) The $K3,3$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K3,3$ is $OMD_1$.
+}
+\label{fig:omd_k33_example}
+\end{figure*}
+
 \medskip\noindent{\bf Example}.

### 2014-12-16 doc/GMP2015/s01_introduction.tex
Slight changes to wording. Added some references.

@@
 Furthermore, a layer is often  either {\sl isostatic or
-underconstrained, not self-stressed} (See Section \ref{sec:prelim} for
+under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
 definitions). These properties are consistent with a self-assembled 2D
@@
 multifunctional, larger 2D structures.
-
-Examples of such materials (See Figure \ref{fig:material_examples})
-include:
 
@@
 
+Examples of such materials (See Figure \ref{fig:material_examples})
+include:
+
 \begin{enumerate}
@@
 
-    \item crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
 
-    \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.
+    \item More recent, engineered examples, including disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.
 
-    \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
+    \item Silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
 \end{enumerate}
@@
 desired properties.
-%
+
 \subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
-We now briefly survey existing techniques for studying 2D {\em
-qusecs,}  many of which are {\it bar-joint} systems (Examples 1,2
-above, see Sections \ref{sec:DRP}, \ref{sec:recomb}), {\it multibody-
-pin} systems (Example 4,5, see Section \ref{sec:bodypin}) or  {\it
-pinned-line incidence} systems (Example 3, see Section
-\ref{sec:pinnedline}). The limitations of these techniques directly
-motivate the contributions of this paper.
+We now briefly survey existing techniques for studying 2D
+\textit{qusecs},  many of which are \textit{bar-joint} systems
+(Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
+\textit{body-hyperpin} systems (Example 4,5, see Section
+\ref{sec:bodypin}) or \textit{pinned-line incidence} systems (Example
+3, see Section \ref{sec:pinnedline}). The limitations of these
+techniques directly motivate the contributions of this paper.
 
 \meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-game \cite{XX}), for
+Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
 locating all maximal, generically rigid subsystems \cite{XX}. When the
@@
 defining desirable or optimal DR-plans and DR-planning algorithms were
-given in \cite{XX}. An {\em optimal DR-plan} is one that minimizes the
-maximum number of child subsystems of any parent system, a property
-that is crucial for obtaining a solution to the parent constraint
-system from the solutions of the child systems.
+given in \cite{XX}. An \emph{optimal DR-plan} is one that minimizes
+the maximum number of child subsystems of any parent system, a
+property that is crucial for obtaining a solution to the parent
+constraint system from the solutions of the child systems.
 
@@
 criteria}
-For a special class of 2D qusecs, namely  {\em tree-decomposable}
+For a special class of 2D qusecs, namely \emph{tree- decomposable}
 systems \cite{XX} common in computer aided mechanical design, (which
 includes ruler-and-compass and Henneberg-I constructible systems), all
-DR-plans turn out to be optimal. This satisfies the so-called {\em
-Church-Rosser} criterion, leading to highly efficient DR-planning
+DR-plans turn out to be optimal. This satisfies the so-called \emph
+{Church-Rosser} criterion, leading to highly efficient DR-planning
 algorithms. For general 2D qusecs, alternate criteria were suggested
-such as {\em cluster minimality} requiring parent systems to be
+such as \emph{cluster minimality} requiring parent systems to be
 composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
-no proper subset forms a rigid system); and {\em proper maximality},
+no proper subset forms a rigid system); and \emph{proper maximality},
 requiring child subsystems to be maximal rigid proper subsystems of
@@
 (sub)system in a DR-plan. They are prohibitively slow in practice. We
-note that utilizing the DR-plan and optimal recombination as a
+note that, utilizing the DR-plan and optimal recombination as a
 principled basis, high performance heuristics and software exists
@@
 
-\meeraheader{v}{Configuration Spaces of Underconstrained Systems}
-For underconstrained 2D bar-joint and multibody-pin cusecs obtained
+\meeraheader{v}{Configuration Spaces of Under-constrained Systems}
+For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
 from various subclasses of tree-decomposable systems, algorithms have
-been developed \cite{XX, XX} that to complete them into isostatic
-systems \cite{XX, XX} and to find paths within the connected
-components \cite{XX,XX} of standard cartesian configuration spaces.
-Most of the algorithms with formal guarantees leverage Cayley
-configuration space theory \cite{XX,XX} to characterize subclasses of
-graphs and additional constraints that control combinatorial
-explosion, and provide faithful bijective representation of connected
-components and paths. These algorithms have decreasing efficiency as
-the subclass of systems gets bigger,  with highest efficiency for
-underlying partial 2-tree graphs (alternately, tree-width 2, series-
-parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-
-of-freedom (dof) graphs with low Cayley complexity (which include
-common linkages such as the Strandbeest, Limacon and Cardioid), and
-decreased efficiency for general 1-dof tree-decomposable graphs. While
-software suites exist \cite{XX,XX}, no such formal algorithms and
-guarantees are known for non-tree-decomposable systems.
-%
+been developed \cite{XX, XX} to complete them into isostatic systems
+\cite{XX, XX} and to find paths within the connected components
+\cite{XX,XX} of standard Cartesian configuration spaces. Most of the
+algorithms with formal guarantees leverage Cayley configuration space
+theory \cite{XX,XX} to characterize subclasses of graphs and
+additional constraints that control combinatorial explosion, and
+provide faithful bijective representation of connected components and
+paths. These algorithms have decreasing efficiency as the subclass of
+systems gets bigger, with highest efficiency for underlying partial
+2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
+minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
+graphs with low Cayley complexity (which include common linkages such
+as the Strandbeest, Limacon and Cardioid), and decreased efficiency
+for general 1-dof tree-decomposable graphs. While software suites
+exist \cite{XX,XX}, no such formal algorithms and guarantees are known
+for non-tree-decomposable systems.
+
 \subsection{Contributions and Organization}
@@
 \begin{itemize}
-\item
-We navigate the NP-hardness barrier mentioned in (2) above, for
-finding optimal DR-plans by defining a so-called {\it canonical} DR-
-plan and showing a strong Church-Rosser property: {\it all canonical
-DR-plans for isostatic or underconstrained 2D qusecs are optimal}
-(Section \ref{sec:DRP}).
+  \item We navigate the NP-hardness barrier mentioned in (2) above,
+  for finding optimal DR-plans by defining a so-called \textit{canonical}
+  DR-plan and showing a strong Church-Rosser property: \textit{all
+  canonical DR-plans for isostatic or under-constrained 2D qusecs are
+  optimal} (Section \ref{sec:DRP}).
 
-\item
-We give an efficient ($O(n^2)$) algorithm to find a canonical (and
-hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above
-(Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}.
-The canonical DR-plan illuminates the essence of the NP-hardness of
-finding optimal DR-plans for overconstrained systems.
+  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
+  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
+  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
+  \ref{sec:pinnedline}). The canonical DR-plan illuminates the essence
+  of the NP-hardness of finding optimal DR-plans for over-constrained
+  systems.
 
-\item
-We give a method to deal with the algebraic complexity of recombining
-the realizations or solutions of  child subsystems into a solution of
-the parent system. Specifically, we define the problem of  minimally
-modifying the indecomposable recombination system so that it becomes
-decomposable via a small DR-plan and yet preserves the original
-solutions in an efficiently searchable manner. We show formal
-connection to well known problems such as optimal completion of
-underconstrained  systems  \cite{XX}.  When the required modifications
-are bounded, we obtain new, efficient algorithms for realizing both
-isostatic and underconstrained qusecs by leveraging recent results
-about Cayley parameters in \cite{XX,XX}  (Section \ref{sec:recomb}).
+  \item We give a method to deal with the algebraic complexity of
+  recombining the realizations or solutions of child subsystems into a
+  solution of the parent system. Specifically, we define the problem
+  of minimally modifying the indecomposable recombination system so
+  that it becomes decomposable via a small DR-plan and yet preserves
+  the original solutions in an efficiently searchable manner. We show
+  formal connection to well known problems such as optimal completion
+  of under-constrained systems \cite{XX}. When the required
+  modifications are bounded, we obtain new, efficient algorithms for
+  realizing both isostatic and under-constrained qusecs by leveraging
+  recent results about Cayley parameters in \cite{XX,XX} (Section
+  \ref{sec:recomb}).
 
-\item
-We briefly describe applications of the above techniques to modeling,
-analyzing and designing specific properties in 2D material layers. For
-Examples 1,2 (achieving isostaticity, distribution of stresses in
-self-similar, bar-joint systems); Example 3 (canonical and optimal DR-
-plans for pinned line incidence systems \cite{XX}) and Examples 4,5
-(boundary-conditions for achieving various desired properties of
-multi-body  pin systems).
+  \item We briefly describe applications of the above techniques to
+  modeling, analyzing, and designing specific properties in 2D material
+  layers. For Examples 1,2 (achieving isostaticity, distribution of
+  stresses in self-similar, bar-joint systems); Example 3 (canonical
+  and optimal DR-plans for pinned line incidence systems \cite{XX})
+  and Examples 4,5 (boundary-conditions for achieving various desired
+  properties of body-hyperpin systems).
 
-\item
-Software implementation and videos can be found at: \todo{Provide
-Link}.
+  \item Software implementation and videos can be found at:
+  \todo{Provide Link}.
 \end{itemize}

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Updated Basics.

@@
 
-In this paper, a {\em geometric constraint system} is a multivariate 
-polynomial (usually bilinear or quadratic) system
-$G(x,\delta)=0$, representing constraints with parameters $\delta$ 
+In this paper, a {\em geometric constraint system} is a multivariate polynomial 
+(usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
 between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
 %
-When the type of constraint (system) is fixed, 
-the system is simply represented as $(G,\delta)$, where 
+When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where 
 $G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
@@
 parameter $\delta$.
-
+%
 For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
@@
 an assignment of coordinates or Euclidean transformations (poses), 
-$p: V \rightarrow
-\mathbb{R}^2$ or $\mathbb{R}^3$ to the vertices of $G$ satisfying the 
+$p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the 
 constraints with parameters $\delta$, 
 modulo orientation preserving isometries 
-(Euclidean rigid body motions except for pinned line-incidence).
+(Euclidean rigid body motions for distance constraint).
 
@@
 $\| x_u -x_v \|_2$  for $(u,v) \in E$.
-The matrix has one row per edge $(u,v)$ in $E$,  
-with 2 columns for $u$ (resp.
-$v$) containing the 2 coordinate indeterminates 
-for $x_u -x_v$ (resp. $x_v-x_u$), 
+The matrix has $2$ columns per vertex in $V$ and one row per edge  in $E$, 
+where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates 
+for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), 
 that is 4 non-zero entries per row.
@@
 identically zero). 
+It is {\em rigid} if the number of independent rows of the rigidity matrix is
+at least the number of columns less the number of rigid body motions, 
+which is 3 for distance constraint systems. 
 It is {\em isostatic (minimally rigid, well-constrained)} 
-if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
-For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
-where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
-The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
+if it is both rigid and independent. 
+%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
+%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
+%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
+%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
 It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.
 
-In this paper, 
-a {\em qusec} is any {\em independent} geometric 
+In this paper, a {\em qusec} is any {\em independent} geometric 
 constraint system of one of the 3
 types mentioned above.
+
+%Define rigidity matroid and combinatorial characterization of independence/rigidity next
+%Laman's theorem (for bar-joint rigidity matroid), 
+%cite also Lovasz Yemini
+%Cite other combinatorial characterizations for other types of constraint systems/hyper graphs:
+%Tay and Whiteey, White and Whitelely,  
+%Streinu Theran (k,l)-sparsity matroid
+
+The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
+where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
+if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
+
+There exist various results on combinatorial characterization of independence and rigidity
+for different types of  rigidity matroid. 
+For bar-joint rigidity matroid, 
+the famous Laman's theorem~\cite{???} states that the underlying graph
+is minimally rigid if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. 
+The result by Lovasz and Yemini~\cite{???} shows that all 6-vertex-connected graphs are rigid in the plane. 
+For bar-body rigidity matroid,
+Tay~\cite{???} proved that
+the underlying multigraph is minimally rigid if and only if it can be decomposed as $3$ edge disjoint spanning trees. 
+White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
+More generally, 
+Lee, Streinu and Theran~\cite{???} defined the $(k,l)$-sparsity matroid,
+where a hypergraph is called $(k,l)$-sparse if $|E'| \le k|V| - l$ for any induced subgraph with at lease 2 vertices.
 
@@
 
-\todo{Where to put these}
-A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
-The left nullspace of the rigidity matrix is called the {\it space of stress}. A vector in the stress space is called a {\it flex vector}.
+For bar-joint graphs,  
+a vector  in the right null space of its rigidity matrix
+is called a {\em flex vector}. 
+It has 2 entries of indeterminates per vertex and represents the internal motions of the system
+(i.e.\ modulo rigid body motions).
+For rigid graphs, all flex vectors are  identically zero. 
+A vector in the left null space of the rigidity matrix
+is called a {\em stress vector}, with one entry of indeterminate per edge.
+For independent graphs, all stress vectors are identically zero.
 

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Add (k,l)-sparsity matroidal condition in Basics. Slight modification in pinned line incidence to match with Basics.

@@
 
-There exist various results on combinatorial characterization of independence and rigidity
+There are various results on combinatorial characterization of independence and rigidity
 for different types of  rigidity matroid. 
@@
 More generally, 
-Lee, Streinu and Theran~\cite{???} defined the $(k,l)$-sparsity matroid,
-where a hypergraph is called $(k,l)$-sparse if $|E'| \le k|V| - l$ for any induced subgraph with at lease 2 vertices.
+Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid}.
+A hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
+and {\em $(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. 
+In general, if you have a $d$-uniform hypergraph, a $(k,l)$-sparsity
+condition is matroidal as long as $l \le dk-1$.
 

### 2014-12-16 doc/GMP2015/s08_pinnedline.tex
Add (k,l)-sparsity matroidal condition in Basics. Slight modification in pinned line incidence to match with Basics.

--- 2014-12-16 doc/GMP2015/s08_pinnedline.tex
@@
 
-The cross-linking  microfibrils can be modeled as a 2-dimensional pinned line incidence constraint system,
+The cross-linking  microfibrils can be modeled as a pinned line incidence constraint system in $\mathbb{R}^2$,
 where incidence constraints are used instead of distance constraints.
@@
 %[2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
-A \textbf{2-dimensional pinned line incidence graph}  $G=(V,E, w_V, w_E)$ 
-%is a \textbf{2-dimensional pinned line incidence graph},
- has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
-Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
-such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
-for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
-such that $e$ is constrained to lie on a line passing through $x$.
-%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
-
-%Let $X$ be a  set of $m$ pins that are fixed in $\mathbb{R}^2$, and $V$ be the index set of a set of vertices.
-%For every pin $x \in X$, we are also given the indices of two vertices $v_1,v_2$ in $V$,
-%such that the line (edge) spanned by $v_1$ and $v_2$ must pass through $x$.
-% % % % % % % % % % % % % % % % % % %
-%Find any such set of points $P$ satisfying the given incidences.
-%For every pin $x\in X$, we are also given a line $\supp{x}$, %of %a set of $s$-dimensional subspaces passing through
-%i.e, a size-two index subset of an unknown set of points $D = \{v_1,\ldots,v_n\}$ in $\mathbb{R}^2$, %$P = \{p_1,\ldots, p_n \}$,
-%such that  $x_i$ lies on the line spanned by $\supp{x}$.
-%%assume additionally that
-%%no $s$ pins lie on any single subspace defined by one of the above hyperedges.
-%Find any such set D %$P[X]$
-%that satisfies the given subspace incidences. % specified by $\supp[P]{x}$.%$\{S_i\}$. % on $P[X]$ and $X$.
-
-\begin{remark}
-For 2-dimensional pinned line incidence graphs,
-the constant $k$ used in the definition of independence and rigidity is $0$  \cite{xxx} (see Section~\ref{sec:prelim_defs}).
-\end{remark}
+A {\em pinned line incidence system} $(G,\delta)$  
+is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$, 
+such that each edge is constrained to lie on a line passing through 
+the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
+%
+% $G=(V,E, w_V, w_E)$  
+%%is a \textbf{2-dimensional pinned line incidence graph},
+% has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
+%Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
+%such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
+%for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
+%such that $e$ is constrained to lie on a line passing through $x$.
+%%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
+%
+A pinned line incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$  \cite{xxx}. 
+Note that no trivial motion exists since the pins have fixed positions on the plane.
 
@@
 Consequently, each fibril can be modeled as an edge of the graph,
-with the attachment being the corresponding pin in $X$.
+with the attachment being the corresponding pin.
 The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge.
 
-Figure~\ref{fig:pinned_line} shows an example of a 2-dimensional pinned line incidence graph,
+Figure~\ref{fig:pinned_line} shows an example of a  pinned line incidence graph,
 where the grey ovals stand for pins representing attachments of fibrils,
-and the vertices $a_1,a_2,\ldots, d_3$ represent cross-linkings.
+and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
 % and the black dots stand for vertices/cross-linkings.
@@
   \centering
-   \includegraphics[width=.9\linewidth]{img/pinned}
-\caption{ A well-constrained 2-dimensional pinned line incidence graph.}
+   \includegraphics[width=.8\linewidth]{img/pinned}
+\caption{ A well-constrained pinned line incidence graph.}
 \label{fig:pinned_line}
@@
 In this section, we will adapt the results in Section~\ref{sec:DRP}
-to give the Optimal DR-plan for 2-dimensional pinned line incidence graphs.
+to give the Optimal DR-plan for pinned line incidence graphs.
 First, we note the following differences of pinned line incidence graphs from the bar-joint graphs.
@@
 \item For pinned line incidence graphs,
-since the pins in $X$ have fixed positions on the plane,
-o trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
+since the pins have fixed positions on the plane,
+no trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
 In particular, note that both a single vertex and a single edge are under-constrained graphs.
-
-%\todo{Over-constrained trivial subgraphs does not exist??? }\
-%\todo{Both a single vertex and a single edge are underconstrained}
-%\begin{itemize}
-%%\item For  pinned line incidence graphs:
-%%no trivial motion exists since the pins have fixed locations.
-%\item
-%\todo{We just define the %trivial graph/
-%DR-plan leaves for 2-dimensional pinned line incidence graphs to be single vertices...  }
-%\item \todo{A single edge can be a node of the DR-plan}
-%\end{itemize}
 
@@
 For pinned line incidence graphs,
-since the pins in $X$ have fixed positions on the plane,
+since the pins have fixed positions on the plane,
 {\em a well-constrained vertex-maximal proper subgraph can be disconnected},
@@
 With the above modification,
-we now adapt Theorem~\ref{theorem:main} %, the optimal DR-plan for 2-dimensional bar-joint graphs,
-for the optimal DR-plan for 2-dimensional pinned line incidence graphs.
+we now adapt Theorem~\ref{theorem:main} 
+for the optimal DR-plan for  pinned line incidence graphs.
 
 \begin{corollary}
-Given a well-constrained 2-dimensional pinned line incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
+Given a well-constrained  pinned line incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
 \begin{enumerate}
@@
 Given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan,
-proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to 2-dimensional pinned line incidence graphs.
-The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2-dimensional pinned line incidence graphs 
+proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to  pinned line incidence graphs.
+The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line incidence graphs 
 is given in Appendix~\ref{sec:appendix_pinned}.
@@
 Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lem:vertex_intersection},
-it is straightforward to adapt the proof of Theorem~\ref{theorem:main_trivial} and \ref{theorem:main_wellconstrained}
-to  2-dimensional pinned line incidence graphs,
+it is straightforward to adapt the proof of Theorem~\ref{theorem:main}
+to  pinned line incidence graphs,
 thus proving Corollary~\ref{cor:pinned}.
-Consequently, we can efficiently find the optimal DR-plan for 2-dimensional pinned line incidence graphs using basically the same algorithm as for 2-dimensional bar-joint graphs.
-%we can analyze microfibril cross-linkings by finding the optimal DR-plan for  2-dimensional pinned line incidence graphs in the same way as for 2-dimensional bar-joint graphs.
+Consequently, we can efficiently find the optimal DR-plan for  pinned line incidence graphs using basically the same algorithm as for  bar-joint graphs.
 
-%TODO ???
-The recombination problem for 2-dimensional pinned line incidence systems is trivial.
+{\bf Note. }The recombination problem for  pinned line incidence systems is trivial.
 Since the pins are given fixed positions in the plane,

### 2014-12-17 doc/GMP2015/appendix_a.tex
Implemented Joel's suggestions for my section.

@@
 % {\sffamily \textbf{Lemma \ref{lemma:main_trivial}}
+\begin{corollary}\label{corollary:no_edges_between_diff}
+Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
+\end{corollary}
 
 \begin{proof}
+Lemma \ref{lemma:wc_intersection_is_C} shows that $C_i\cup C_j$ must equal the entire parent $C$.
+\end{proof}
+
+\begin{proof} (Of Lemma \ref{lemma:wc_intersection_makes_all_wc})
+
 The alternative phrasing is due to Lemma \ref{lemma:wc_intersection_is_C}.

### 2014-12-17 doc/GMP2015/s03_drp.tex
Implemented Joel's suggestions for my section.

@@
 \item
-% In a valid DR-plan, all children must be proper subgraphs. Therefore, for the union to be the parent (and satisfy the conditions of a DR-plan) there must be at least two children.
-It follows directly from Lemma \ref{lemma:wc_intersection_makes_all_wc} that if there exists $C_i, C_j$ such that $C_i \cup C_j$ is well-constrained, then the smallest fan-in at this node can be made by selecting any two $C_i, C_j$.
-% you found a tree with only 2 children.
-% Clearly, this is the optimal choice at this level.
-However, for it to be a solution to $OptimalDRP(G)$, it must not affect the maximum fan-in for any descendant.
+We show that no subset of the children can union to form $C$, thereby requiring all of them be included in an optimal plan.
 
-Take $D=\bigcap_{k\in N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
-% \[C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)\]
-% the children of this node will be
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)\]
-% and
-% \[Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)\]
-$C_i=Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i\}}{R_k}\right)$
-the children of this node will be
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,m\}}{R_k}\right)$
-and
-$Idc\left(C,D\cup\bigcup_{k\in N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $m\neq n$. This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,D\cup R_k)$.
+Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \emph{assume} is well-constrained.
+If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
+So, it must be that $U=C$.
+\usestwod
+However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
+By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
+Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$.
+Thus, the assumption is wrong and $U$ cannot be well-constrained.
+As $C$ is well-constrained, this means no proper subset of the children in $CompleteDRP(C)$ can union to form $C$.
 
@@
 \item
-Take $X\subset \{1,\ldots,N\}$ and $D=\bigcup_{i\in X}{C_i}$.
-If $D\neq C$ but it was well-constrained, then we just found a larger proper subgraph and the graphs in the set were not vertex-maximal to begin with.
+If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
+This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
+Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
 
-\usestwod
-By Lemma \ref{lemma:uc_intersection_makes_all_uc}, if $C_i \cap C_j$ is trivial then for $k\notin X$, $D\cap C_k$ must be one or more trivial subgraphs.
-Since $C_k=C\cap C_k=D\cap C_k$, by definition of a DR-plan, the assumption that $C_k$ is well-constrained has been contradicted.
-Thus, $X$ must be all of $\{1,\ldots,N\}$.
+However, to guarantee that any two are \emph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant.
+Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
+Suppose we select $i$ and $j$, where $i\neq j$, as the children.
+Since
+% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
+% the children of this node will be
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
+% and
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
+$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
+the children of this node will be
+$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
+and
+$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
+for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other.
+\todo{Prove these are valid children?
+Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once.
+Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
 \end{enumerate}
+
+This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
 \end{proof}
@@
 Just start the first paragraph with Suppose C_i union C_j is well constained maximal proper and assume it is not C.
-\textit{Forward direction:} Proof by contradiction.
-If the $C_i\cup C_j$ was well-constrained but $C_i\cup C_j \neq C$, we would have found a larger well-constrained proper subgraph containing them; namely, $C_i\cup C_j$ or something containing it.
-The two original subgraphs could not have been vertex-maximal proper.
+Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
 
-\textit{Reverse direction:} $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
+In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
 Thus, $C_i\cup C_j$ is well-constrained.
-\end{proof}
-
-\begin{corollary}\label{corollary:no_edges_between_diff}
-Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
-\end{corollary}
-
-\begin{proof}
-Lemma \ref{lemma:wc_intersection_is_C} shows that $C_i\cup C_j$ must equal the entire parent $C$.
 \end{proof}
@@
 \begin{proof}
-By Lemma \ref{lemma:union_intersection}, if some $C_i\cap C_k$ is not trivial, then it must be well-constrained.
-Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, all of the intersections are well-constrained.
-This contradicts the statement that $C_i\cap C_j$ is trivial.
+Assume there is some $k$ such that $C_i\cap C_k$ is not trivial.
+By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be well-constrained.
+Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be well-constrained.
+This means that $C_i\cap C_j$ is well-constrained.
+Therefore, such a $k$ cannot exist and all intersections are trivial.
 \end{proof}

### 2014-12-17 doc/GMP2015/s03_drp.tex
Minor changes to sec:DRP.

@@
 \begin{enumerate}
-    \item By definition of trivial graph; if $F_i\cup F_j$ were trivial, then $F_i$ and $F_j$ must be trivial.
+    \item If $F_i\cup F_j$ were trivial, then $F_i$ and $F_j$ must be trivial, by definition.
 
-    \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
+    \item Observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$. Then it is straightforward application of the equation above.
+
+    % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
+
+    \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.
+
+    % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
 Therefore $d(F_i\cap F_j)=2k-l>k$.
@@
 
-    \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
+    % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
 Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
@@
 
-    \item Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained. So it follows from the above that an under-constrained intersection is not possible.
+    \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
 \end{enumerate}
@@
 Just start the first paragraph with Suppose C_i union C_j is well constained maximal proper and assume it is not C.
-Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
+Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
 
 In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
-Thus, $C_i\cup C_j$ is well-constrained.
+Thus, $C_i\cup C_j=C$ is well-constrained.
 \end{proof}
@@
 \subsection{Algorithm}
+\todo{Cut this section out?}
 The algorithm to find the optimal DR-plan is a straightforward application of the theory.

### 2014-12-17 doc/GMP2015/s02_preliminaries.tex
Updated preliminary

@@
 
-%Define rigidity matroid and combinatorial characterization of independence/rigidity next
-%Laman's theorem (for bar-joint rigidity matroid), 
-%cite also Lovasz Yemini
-%Cite other combinatorial characterizations for other types of constraint systems/hyper graphs:
-%Tay and Whiteey, White and Whitelely,  
-%Streinu Theran (k,l)-sparsity matroid
-
 The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
@@
 
-% % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
-%This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.
-
-%The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
-%if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
-%In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
-%A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
-
-%A graph $G = (V,E)$ is:
-%\begin{itemize}
-%	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
-%	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-%	\item {\it flexible} if $G$ is not rigid
-%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
-%    \item \textbf{over-constrained} if it is not independent.
-%\end{itemize}
-%In this paper all results are about independent - underconstrained or
-%isostatic
-
 For bar-joint graphs,  
@@
 
-% A generalized version of Asimow Roth (implicit fn thm) states that
-% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
-%  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
-%\todo{--rigidity matroid\\
-%--Combinatorial characterization of rigidity matroid
-%Example: Laman's theorem for bar-joint (refer to others)
-%Lovasz Yemini
-%Tay white and whiteley
-%(citations)
-%+ pinned line incidence and multi-body-pin (in application
-%section)}
+%XXXXXXXXXXXXXXX BASICS until combinatorial characterization of 
+%independence and rigidity XXXXXXXXXXX
 
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+NOTE: geometric constraint systems can have inequalities (the parameters
+in $\delta$ being inexact, small intervals of values).
 
-%Standard except for generality and ordering and the fact that I want to
-%avoid the use of the word framework if possible. Do everything
-%with constraint systems, linearization and genericity using indeterminates.
+XXXXXXXXXsomething has to be said 
+along with asimow roth (infinitesimal rigidity 
+implies rigidity)
+when the rigidity matrix has appropriate rank
+the realizations or solutions of the 
+of the corresponding constraint system 
+are generically isolated and zero-dimensional 
+(if the constraint system has equalities and exact 
+values for the parameters $\delta$.
+If the constraint system has inequalities, i.e, 
+if the $\delta$ lie in an interval,
+this claim is approximate: 
+the solutions are isolated small, full-dimensional  connected components.
 
-%--general  geometric constraint system --
+%XXXXXXXX Also in the recombination section, it should be noted
+%that the HUGE advantage of the convex Cayley method
+%is that it is completely unaffected by having the $\delta$ be intervals
+%cite my paper with Heping!!!
 
-%A $d$-dimensional {\it geometric constraint system} in general consists of
-%
-%- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.
-%- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint value
-%In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).
+%XXXXXXXXXX CUrrently there are open problems strewn throughout the paper. 
+%Should they be collected to the conclusion?
+%Or at least clearly labeled as open problems?
+%XXXXXXXXX
 
-%--underlying constraint graph or hypergraph (common treatment)
+\noindent{\bf Note:}
+In the remainder of this section and the next two sections 
+\ref{sec:DRP} and \ref{sec:recomb} 
+we only consider bar-joint qusecs and graphs
+Relevant formal analogies  for the other 2 types of qusecs and (hyper)graphs 
+are given in the subsequent 2 sections on Applications.
 
-%--linearization: Jacobian of constraint system,
-%rigidity matrix of indeterminates
-%
-%The algebraic system can usually be linearized at {\it generic} points
-%by taking the Jacobian of the system,
-%giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
-%whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.
+XXXNow comes the definition of stress vector and flex vector. 
+NOTE: something has to be said about what flexes and stresses mean
 
-%-- give example for the case of
-%linkages
-%
-%In the case of linkage, 
-%let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
-%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
-The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.
-$v$) and zero in the other entries.
+An {\em external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates)
+which specifies a 2D vector acting at each vertex
+such that the stress balance equation $sR = t$ holds, where $s$ is a vector
+of internal stresses, one for each bar, and $R$ is the rigidity matrix.
 
-%%Let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
-%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
-The row corresponding to the hyperedge $e = \{v_1, v_2, \ldots, v_k\} \in E$ represents the constraint between the vertices $ \{v_1, v_2, \ldots, v_k\}$,
-%has $d$ non-zero indeterminate entries %$\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) 
-%in the $d$ columns corresponding to each vertex $v_i \in e$ and zero in the other entries.
+\subsection*{Decomposition-Recombination (DR) Plans}
 
-%generic independence, infinitesimal rigidity
-%(based on indeterminates)
-
-%--stress vectors and flex vectors
-%(also based on indeterminates)
-
-%--generic isostaticity (minimal rigidity, wellconstrainedness),
-%flexibility (not rigid), underconstrained (independent
-%and not rigid), overconstrained (dependent -- only briefly
-%mentioned
-%in this paper; all results are about independent - underconstrained or
-%isostatic)
-
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-%The problem is equivalent to combinatorially determining the generic rank of the 3-dimensional bar-joint rigidity matrix of a graph.
-The $d$-dimensional bar-joint rigidity matrix of a graph $G = (V,E)$, denoted $R_d(G)$, is a matrix of indeterminates.
-Let $p_1(v), p_2(v), \ldots, p_d(v)$  represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$of the joint corresponding to a vertex $v \in V$.
-.
-The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.
-The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.
-$v$) and zero in the other entries.
-%
-%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be independent (we drop ``bar-joint'' from now on) in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries.
-This yields the $d$-dimensional generic rigidity matroid associated with a graph $G$.
-The graph is rigid if the number of generically independent rows or the rank of $R_d(G)$ is maximal, i.e., $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
-
-\subsection{DR-plan}
-%\subsection{Definitions}
-%\label{sec:prelim_defs}
-%
-%% TODO
-%
-%% A paragraph of preliminaries such as:
-%%     - rigidity matrix R,
-%%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
-%%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
-%% Should precede the definition of
-%%     + independent,
-%%     + overconstrained (not independent),
-%%     - rigid (contains an independent set with sufficiently many eedges/constraints)
-%%     - isostatic (wellconstrained/minimally rigid), ,
-%%     - flexible (not rigid),
-%%     + underconstrained (independent and not rigid).
-%% After that, define DR plans etc.
-%
-%In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
-$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
+%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
 Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
 The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
-%
+
 %Given a constant $k$, a graph $G$ is:
@@
 %
-\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.
-
 % \begin{definition}
@@
 
+A \textbf{trivial} graph is   a single vertex.
 \begin{definition}\label{def:drp}
-    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
+    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
+    \cite{XX}
+    $DRP(G)$, is defined as a forest that has the following properties:
     \begin{enumerate}
-        \item The root of the tree `contains' $G$.
-        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
-        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
-        \item A node with a trivial subgraph is a leaf.
+        \item Each  node represents/contains/is  a rigid vertex-induced subgraph of $G$
+        %\item For a node, $C$, that is  a non-trivial graph, its $N$
+        %    children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper subgraphs of $C$.
+        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the 
+            vertex set of $C$.
+        \item A leaf node is a trivial subgraph.
+       \item A root node is a maximal rigid subgraph of $G$
     \end{enumerate}
@@
 
-    An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.
+    An equivalent DAG can be constructed from the above forest such that all nodes
+    containing the same subgraphs are identified into a single node.
+    In this manuscript, however, we retain a forest as defined above.
+    When a graph is isostatic, its DR plan is a tree.
 
-    Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
+    %Note that nodes will be referred to interchangeably as ``the node that
+    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
 
-    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
-This makes rule 3 implicit.
-The complete DR-plan is unique.
+    A DR-plan is \textbf{complete} if it satisfies an additional rule: 
+        \item For a node, $C$, that is  a non-trivial graph, its $N$
+            children, $C_1, \ldots, C_N$, are all of the 
+            rigid vertex-maximal proper subgraphs of $C$.
+    This makes Rule 2 implicit, and in fact, $C$ contains all the edges in
+    the union of its children as well. The complete DR-plan is unique.
 
@@
 
-\subsection{Notation}
-
-In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
+%\emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
 % \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
@@
 
-$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.
+\noindent{\bf Note:}
+In Section \ref{sec:DRP} $G$ is assumed to be isostatic.
+$Idc(G,X)$ is the subgraph of $G$ induced by the vertex set $X\subseteq V$. 
+For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
+In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$. $C_i$
+is the $i^{\text{th}}$ child of $C$. It is implied that 
+$C_i$ is an isostatic vertex-maximal proper subgraph.
 
-$C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
+\subsection*{Importance of an Optimal DR-plan}
+The use of a DR-plan in designing especially self-similar qusecs is
+immediately clear, as shown in Figures \ref{XX} and \ref{XX}.
+In fact, knowing an optimal DR-plan is crucial for design and analysis of 
+even non-self-similar qusecs 
+and their realizations with various desirable properties.
+
+First, note that the basic task of \underbar{{\sl realizing given qusecs}} would require finding real
+ solutions 
+ a large multivariate polynomial system (of inequalities and equalities
+ representing the constraints). Without DR-planning, this requires double
+ exponential time in the number of variables
+ (even if orientation type is specified).
+With DR-planning, the complexity is dominated by the size of the largest 
+subsystem that is
+solved, or recombined from the solutions of its child subsystems i.e.,the maximum fan-in occuring in a DR-plan.
+Thus finding an optimal DR-plan is important for realization.
+
+Next, using the DR-plan we obtain a \underbar{{\sl recursive block
+decomposition of the rigidity matrix}}   
+with each block being the rigidity matrix of 
+an isostatic subgraph.
+The blocks intersect on isostatic or trival subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ 
+correspondingly \underbar{{\sl recursively decomposes the stress vectors}}
+$s$ that balance a given
+external stress $t$ since $sR = t$, and moreover, the external stress vector
+$t$ now \underbar{{\sl recursively "distributes" as external stresses}} 
+for each isostatic subgraph of the DR-plan.
+When the rigidity matrix and stress vectors are indeterminates, it can be
+used to design  qusecs underlying graph and parameters $\delta$ to obtain an  
+optimal distribution of stresses on the geometric primitives. 
+When the realization of the qusec is given, this can be used to 
+analyze the distribution of a  given external load.
+
+%\subsection{DR-plan}
+%%\subsection{Definitions}
+%%\label{sec:prelim_defs}
+%%
+%%% TODO
+%%
+%%% A paragraph of preliminaries such as:
+%%%     - rigidity matrix R,
+%%%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
+%%%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
+%%% Should precede the definition of
+%%%     + independent,
+%%%     + overconstrained (not independent),
+%%%     - rigid (contains an independent set with sufficiently many eedges/constraints)
+%%%     - isostatic (wellconstrained/minimally rigid), ,
+%%%     - flexible (not rigid),
+%%%     + underconstrained (independent and not rigid).
+%%% After that, define DR plans etc.
+%%
+%%In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
+$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
+Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
+The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
+%%
+%%Given a constant $k$, a graph $G$ is:
+%%\begin{itemize}
+%%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
+%%
+%%    \item \textbf{over-constrained} if it is not independent.
+%%
+%%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
+%%
+%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+%%
+%%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+%%
+%%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
+%%\end{itemize}
+%%
+%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.
+%
+%% \begin{definition}
+%%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
+%%     % induced
+%%     subgraph $S\subseteq G$ such that $d(S) > k$.
+%% \end{definition}
+%
+%% \begin{definition}
+%%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
+%%     % induced
+%%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
+%% \end{definition}
+%
+%
+%
+%% \begin{definition}
+%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
+%%     \todo{Maybe leave the following out until it's needed later?}
+%%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
+Trivial graphs for 2D would be a single vertex.
+Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between).
+Etc.
+These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
+%% \end{definition}
+%
+%
+%\begin{definition}\label{def:drp}
+%    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
+%    \begin{enumerate}
+%        \item The root of the tree `contains' $G$.
+%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+%        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
+%        \item A node with a trivial subgraph is a leaf.
+%    \end{enumerate}
+%
+%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
+%
+%    An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.
+%
+%    Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
+%
+%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+This makes rule 3 implicit.
+The complete DR-plan is unique.
+%
+%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+%\end{definition}
+%
+%% \begin{definition}
+%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+By this, rule 3 is implicit for a complete DR-plan.
+%% \end{definition}
+%
+%% \begin{definition}
+%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+%% \end{definition}
+%
+%
+%\subsection{Notation}
+%
+%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
+%% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
+It is not useful to solve for two bodies simultaneously if there are no constraints between them.
+%% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.
+%
+%$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.
+%
+%$C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
 $C_i$ is the $i^{\text{th}}$ child of $C$.

### 2014-12-17 doc/GMP2015/s07_problempinning.tex
Cleaned up Body-pin. Added more to proof...

@@
-\section{Application: Finding extensions of underconstrained glassy structures to isostatic}
+\section{Application: Finding completions of underconstrained glassy structures to isostatic}
 \label{sec:bodypin}
@@
     \raisebox{0.5\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
-    \caption{Example of a signle layer of a Silica Oxygen glassy structure.
+    \caption{Example of a single mono-layer of a Silica Oxygen glassy structure.
 This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
 Not shown here are other layers stacked on this one.
-Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in another layer.}
+Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.}
     \label{fig:silica_glass}
@@
 
-\begin{itemize}
+\begin{enumerate}
     \item Pin together 2 underconstrained monolayers in such a way that the resulting bilayer becomes isostatic (see Figure \ref{fig:silica_glass}).
@@
     \item Build the structures by repeated subdivision to ensure they are isostatic at each layer (see Figure \ref{fig:subdivision})
-\end{itemize}
+\end{enumerate}
 
@@
 \begin{definition}
-    A multi-body-pin qusec is a qusec where the obejct are rigid bodies that are pinned together by some number of pins.
+    A \emph{body-hyper-pin qusec} is a qusec where the obejct are rigid bodies that are pinned together by some number of pins.
 \end{definition}
@@
 \begin{remark}
-    A multi-body-pin qusec is a special case of bar-joint qusecs that we have been studying.
+    A body-hyper-pin qusec is a special case of bar-joint qusecs that we have been studying.
 As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
@@
 
-We now go over 2 subclasses of multi-body-pin qusecs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
+For the rest of this section, we will only be dealing with the DR-pla n of such qusecs.
+As such, we will refer only to the underlying graph of the qusecs.
+We now go over 2 subclasses of body-hyper-pin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
 
@@
 \label{def:body-pin}
-    A body-pin qusec is a multi-body-pin qusec with the following conditions:
+    A body-pin graph is a body-hyper-pin graph with the following conditions:
     \begin{itemize}
@@
     \end{itemize}
-    Such a body-pin qusec can also be seen as a {\em body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
-For such body-bar qusecs, graphs with 1 and 2 dofs can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
+    Such a body-pin graph can also be seen as a \emph{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
+Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
 \end{definition}
@@
 \begin{definition}
-    A multi-triangle-pin qusec is a multi-body-pin qusec where each body is a triangle, meaning it is pinned in at most 3 places.
+    A \emph{multi-triangle-pin graph} is a body-hyper-pin graph where each body is a triangle, meaning it is pinned in at most 3 places.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hyper-graphs, 1 and 2 dofs can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
+For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
 \end{definition}
 
-Body-pin qusecs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
-Multi-triangle-pin qusecs can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (se Figure \ref{fig:silica_glass}).
+Body-pin graphs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
+Multi-triangle-pin graphs can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (se Figure \ref{fig:silica_glass}).
 Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan:
@@
 \begin{definition}
-    A $(k,l)$-tight DR-plan is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
+    A \emph{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
 In our case, the trivial nodes are just the bodies.
@@
 
-\begin{remark}
-\label{rem:1dofcanon}
-    For the 1 dof multi-body-pin qusecs described above, a DR-plan node will follow one of the following:
+\begin{observation}
+    For the 1-dof body-hyper-pin graphs described above, a DR-plan node will follow one of the following:
 
     \begin{itemize}
-        \item Its children will be 2 proper vertex maximal 1 dof graphs that intersect on another 1 dof graph
-        \item Its children will be a number of proper maximal 1 dof subgraphs, joined pairwise by at most one pin
+        \item Its children will be 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph
+        \item Its children will be a number of proper maximal 1-dof subgraphs, joined pairwise by at most one pin
     \end{itemize}
-\end{remark}
+\end{observation}
 
@@
 
-    For item 2, consider the case where we have more than 2 proper vertex maximal 1 dof subgraphs $s_1, ..., s_k, k > 2$.
-Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \Cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
+    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof subgraphs $s_1, ..., s_k, k > 2$.
+Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
 \end{proof}
@@
 \begin{remark}
-    All $(k,l)$-tight canonical DR-plans are optimal. We can find such a DR-plan in the same time complexity as the $(2,3)$ tight case for bar and joint qusecs discussed in Section \ref{sec:DRP}.
+\label{rem:1dofcanon}
+
+    All $(k,l)$-tight canonical DR-plans are optimal.
+We can find such a DR-plan in the same time complexity as the $(2,3)$ tight case for bar and joint graphs discussed in Section \ref{sec:DRP}.
+This is due tot he fact that such $(k,l)$-tight graphs are matroidal and have a pebble game.
 \end{remark}
@@
 %
-\caption{(a) is a 1 dof body-pin qusec.
+\caption{(a) is a 1-dof body-pin graph.
 (b) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}.
-(c) is the 1 dof DR-plan for the graph.
+(c) is the 1-dof DR-plan for the graph.
 In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level.
@@
 \begin{theorem}
-    There is a quadratic algorithm for the 1 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
+\label{thm:1dofcase}
+    There is a quadratic algorithm for the 1-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
 \end{theorem}
@@
 \begin{proof}
-    Suppose we are given a body-pin qusec and its corresponding body-bar graph $G$ and have obtained the DR-plan $T$.
-Each node of $T$ will then be a vertex maximal proper 1 dof subgraph of $G$.
-To make the entire graph wellconstrained, we can add a body (vertex) and pin it (add a 2-multiedge) to a node $b$ in $T$ such that the body is pinned to 2 separate children of $b$.
-From Remark \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a subgraph, so we simply have to pin the new body to the 2 bodies joined by the pin or that are not in the intersecting graph.
-Doing so will ensure the following:
+    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof subgraph of $G$.
 
-    \begin{itemize}
-        \item All children of $b$ will remain 1 dof
-        \item All ancestors of $b$ (including $b$) will now be wellconstrained
-    \end{itemize}
+    To make the graph isostatic, we need only add one body and pin it to 2 other bodies.
+Doing so will cause $G$ to become $(3,3)$-tight.
+We choose the 2 boodies to pin to by choosing a node $b$ in $T$ and looking at its children.
+From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a subgraph.
+We pin the new body to bodies in two separate children.
+Doing so will ensure that all children of $b$ will have 1-dof and all ancestors of $b$ (including $b$) will now be isostatic.
+
+    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+In $T_b$, $fanin(b)$is the number of leaves in the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
+Similarly, for any other node $w$ that is an ancestor of $b$, $fanin(w)$ is the the number leaf nodes in the tree rooted at $w$, excluding the subtree constaining $b$, plus 1 (for the node leading to $b$).
+Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
+The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
+Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
 
@@
 In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
-Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$ excluding $b'$.
+Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$, excluding $b'$.
 Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
+The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
 Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
+
+    We claim that pinning a new body to a node in the manner described covers all possible ways of adding a new body.
+Assume we add a new body $b_n$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
+Then, there will be some lowest 1-dof node $v$ in our tree such that $b_i$ and $b_j$ appear in $v$.
+This could indeed be the root node.
+Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning a new body to $b_i$ and $b_j$.
+
+    It is the case that the node we need to pin can wlog be the deepest nontrivial node of some path in $T$.
+Suppose we chose to pin a node $b$ that has a nontrivial child $v$.
+Then, $fanin(b)$ is the number of leaves in the subtree rooted at $b$, including the subtree rooted at $v$.
+If we had instead chosen to pin $v$, then $fanin(b')$ will just be the number of nodes in all its subtrees except for the one rooted at $v$.
+Then, $\max{fanin(b'),fanin(v)} < fanin(b)$, and for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $T_v$ has size at most $T_b$.
+Thus we only have to check the deepest non-trivial nodes.
+
     Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
 
-    For the multi-triangle pin qusecs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
+    For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
 
-    % For the 2 dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
-These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1 dof case.
+    % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
+These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1-dof case.
 \end{proof}
@@
 
-\begin{theorem}
-    There is a \textcolor{red}{$O(?)$} algorithm for the 1 dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin qusecs.
-\end{theorem}
+\begin{observation}
+    For the 2-dof case, if a similar statement to Remark \ref{rem:1dofcanon} can be made, then Theorem \ref{thm:1dofcase} is true for 2-dof as well.
+\end{observation}
 
 \begin{proof}
-    We can do a similar algorithm to the 1 dof case, except we constrain a given node 2 times (by adding 2 triangles or 2 bodies and 4 pins).
+    The only difference from the 1-dof case is that now we need to remove 2-dof from our graph.
+We build a 2-dof DR-plan $T$.
+Like above, we need to add a body and 2 pins to 2 nodes now to get to isostatic.
+
+    Suppose we pin 2 distinct nodes $v_i$ and $v_j$.
+Then, there will be some common ancestor $a$ of $v_i$ and $v_j$.
+Then, in $T_{v_i,v_j}$, $fanin(a)$ will be the number of leaves in the subtree rooted at $a$.
+However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin(a)'$ will be the number of leaves in its subtree excluding the subtree containing the node we pinned.
+Thus $fanin(a)' < fanin(a)$ and the size of $T_{v_i}$ is at most size of $T_{v_i,v_j}$.
+
+    So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
+\end{proof}
+
+% \begin{theorem}
+%     There is a \todo{$O(?)$} algorithm for the 2-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
+% \end{theorem}
+
+% \begin{proof}
+%     We can do a similar algorithm to the 1-dof case, except we constrain a given node 2 times (by adding 2 triangles or 2 bodies and 4 pins).
 This causes the nodes along the path to the given node to become well-constrained while all others will be underconstrained.
-Thus, our DR-plan will have the same size as the 1 dof case.
+Thus, our DR-plan will have the same size as the 1-dof case.
 This is still a quadratic algorithm.
@@
 
-An open question remains as to the complexity of the optimal completion problem when there are more than 2 dofs. For higher number of dofs, the $(l,k)$ characterization is no longer matroidal.
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX QUESTION XXXXXXXXXXXXXXXXXXXXXXXXX
+An open question remains as to the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(l,k)$ characterization is no longer matroidal.

### 2014-12-17 doc/GMP2015/s02_preliminaries.tex
Updated preliminary. Added not to recombination. Modified pinned line-incidence according to prelimary and updated DRP section.

@@
 \subsection{Basics}
-%TODO!
 
@@
 (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
-between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
+between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.   
 %
@@
 $u\in V$.
+
+\noindent {\bf Note:} Geometric constraint systems can also have inequalities, 
+where the parameters in $\delta$ are small intervals of values rather than exact values.
 
@@
 and do not depend on $\delta$ (or they hold for all but a measure 
-zero set of $\delta$
-values). Many of these are properties of the jacobian 
-$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of
-$G$} 
+zero set of $\delta$ values). Many of these are properties of the jacobian 
+$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of $G$} 
 (a matrix of indeterminates).
@@
 
-Some of these generic constraint system or 
-(hyper)graph properties are the following ({\bf Note:} we refer to these as
+%XXXXXXXXXsomething has to be said 
+%along with asimow roth (infinitesimal rigidity 
+%implies rigidity)
+%When the rigidity matrix has appropriate rank, the realizations or solutions 
+%of the corresponding constraint system are generically isolated and zero-dimensional 
+%if the constraint system has equalities and exact values for the parameters $\delta$.
+%If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
+%this claim is approximate: the solutions are isolated small, full-dimensional  connected components.
+
+One important property of generic constraint system or 
+(hyper)graph %that depends on the appropriate rigidity matrix of indeterminates
+is \emph{rigidity} ({\bf Note:} we refer to these as
 properties of the constraint system or as properties of the underlying
-(hyper)graph interchangeably).
+(hyper)graph interchangeably),
+i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.
+%(if the constraint system has equalities and exact values for the parameters $\delta$).
+The result by Asimow and Roth~\cite{???} shows a constraint (hyper)graph is rigid
+if it is generically {\em infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
+%This condition is called generic {\em infinitesimal rigidity}.
+
+\noindent {\bf Note:} If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
+the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.
+% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
+
+Other generic constraint system or (hyper)graph properties are the following: %({\bf Note:} we refer to these as
+%properties of the constraint system or as properties of the underlying
+%(hyper)graph interchangeably).
 %
@@
 identically zero). 
-It is {\em rigid} if the number of independent rows of the rigidity matrix is
-at least the number of columns less the number of rigid body motions, 
-which is 3 for distance constraint systems. 
+%It is {\em rigid} if the number of independent rows of the rigidity matrix is
+%at least the number of columns less the number of rigid body motions, 
+%which is 3 for distance constraint systems. 
 It is {\em isostatic (minimally rigid, well-constrained)} 
@@
 
-In this paper, a {\em qusec} is any {\em independent} geometric 
-constraint system of one of the 3
-types mentioned above.
-
-The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
-where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
-if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-
+Defining the independence of a subset of edges $E' \subseteq E$ 
+to be the independence of corresponding rows in the appropriate rigidity matrix of indeterminates,
+we obtain the {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$. 
+%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
+%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
+%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent. 
 There are various results on combinatorial characterization of independence and rigidity
@@
 White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
-More generally, 
-Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid}.
-A hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
+Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid},
+where a hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
 and {\em $(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. 
-In general, if you have a $d$-uniform hypergraph, a $(k,l)$-sparsity
+In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity
 condition is matroidal as long as $l \le dk-1$.
 
-For bar-joint graphs,  
-a vector  in the right null space of its rigidity matrix
-is called a {\em flex vector}. 
-It has 2 entries of indeterminates per vertex and represents the internal motions of the system
-(i.e.\ modulo rigid body motions).
-For rigid graphs, all flex vectors are  identically zero. 
-A vector in the left null space of the rigidity matrix
-is called a {\em stress vector}, with one entry of indeterminate per edge.
-For independent graphs, all stress vectors are identically zero.
-
-%--A generalized version of Asimow Roth (implicit fn thm)
-%generic rigidity equivalent to infinitesimal rigidity (easy
-%when the latter def is based on indeterminates)
-
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-%XXXXXXXXXXXXXXX BASICS until combinatorial characterization of 
-%independence and rigidity XXXXXXXXXXX
-
-NOTE: geometric constraint systems can have inequalities (the parameters
-in $\delta$ being inexact, small intervals of values).
-
-XXXXXXXXXsomething has to be said 
-along with asimow roth (infinitesimal rigidity 
-implies rigidity)
-when the rigidity matrix has appropriate rank
-the realizations or solutions of the 
-of the corresponding constraint system 
-are generically isolated and zero-dimensional 
-(if the constraint system has equalities and exact 
-values for the parameters $\delta$.
-If the constraint system has inequalities, i.e, 
-if the $\delta$ lie in an interval,
-this claim is approximate: 
-the solutions are isolated small, full-dimensional  connected components.
-
-%XXXXXXXX Also in the recombination section, it should be noted
-%that the HUGE advantage of the convex Cayley method
-%is that it is completely unaffected by having the $\delta$ be intervals
-%cite my paper with Heping!!!
-
-%XXXXXXXXXX CUrrently there are open problems strewn throughout the paper. 
-%Should they be collected to the conclusion?
-%Or at least clearly labeled as open problems?
-%XXXXXXXXX
-
-\noindent{\bf Note:}
-In the remainder of this section and the next two sections 
-\ref{sec:DRP} and \ref{sec:recomb} 
-we only consider bar-joint qusecs and graphs
+In this paper, a {\em qusec} is any {\em independent} geometric 
+constraint system of one of the 3 types mentioned above.
+In the remainder of this section and the next two sections~\ref{sec:DRP} and \ref{sec:recomb} 
+we only consider bar-joint qusecs and graphs. 
 Relevant formal analogies  for the other 2 types of qusecs and (hyper)graphs 
@@
 
-XXXNow comes the definition of stress vector and flex vector. 
-NOTE: something has to be said about what flexes and stresses mean
+%XXXNow comes the definition of stress vector and flex vector. 
+%NOTE: something has to be said about what flexes and stresses mean
 
+%TODO
+Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix
+is called a {\em flex vector}. 
+It has 2 entries per vertex and represents the internal motions of the system
+(i.e.\ modulo rigid body motions).
+For rigid graphs, all flex vectors are identically zero. 
+%
+A vector (of indeterminates) in the left null space of the rigidity matrix
+is called a {\em stress vector}.
+It has one entry per edge and represents a {\em self stress} of the system.
+For independent graphs, all stress vectors are identically zero. 
+%
 An {\em external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates)
@@
 
-\subsection*{Decomposition-Recombination (DR) Plans}
+\subsection{Decomposition-Recombination (DR) Plans}
 
@@
 
-A \textbf{trivial} graph is   a single vertex.
+A {\em trivial} graph is   a single vertex.
 \begin{definition}\label{def:drp}
-    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
+    The  {\em decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
     \cite{XX}
@@
 
-    A DR-plan is \textbf{complete} if it satisfies an additional rule: 
+    A DR-plan is {\em complete} if it satisfies an additional rule: 
         \item For a node, $C$, that is  a non-trivial graph, its $N$
@@
 
-    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+    A DR-plan is {\em optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
 \end{definition}
@@
 
-\subsection*{Importance of an Optimal DR-plan}
+\subsection{Importance of an Optimal DR-plan}
 The use of a DR-plan in designing especially self-similar qusecs is

### 2014-12-17 doc/GMP2015/s04_recombination.tex
Updated preliminary. Added not to recombination. Modified pinned line-incidence according to prelimary and updated DRP section.

@@
 achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
-$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$ ,  all of which have
+$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$,  all of which have
 small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
@@
 \caption{
-(\ref{fig:omd_k33_example:a}) The $K3,3$ with two labeled edges, $e_1$ and $e_2$.
-(\ref{fig:omd_k33_example:b}) The $K3,3$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree.
-(\ref{fig:omd_k33_example:c}) The $K3,3$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K3,3$ is at least $OMD_2$.
-(\ref{fig:omd_k33_example:d}) The $K3,3$ with only $e_2$ removed (dashed line).
-(\ref{fig:omd_k33_example:e}) The $K3,3$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K3,3$ is $OMD_1$.
+(\ref{fig:omd_k33_example:a}) The $K_{3,3}$ with two labeled edges, $e_1$ and $e_2$.
+(\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree.
+(\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$.
+(\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line).
+(\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
 }
@@
 \end{figure*}
+
+\noindent{\bf Note:} A huge advantage of the convex Cayley method
+is that it is completely unaffected when $\delta$ are intervals rather than exact values~\cite{GaoSitharam}.
+%cite my paper with Heping!!!
 
@@
 corresponding to the root of
-this DR-plan involves solving a quadratic system with 8 equations and variables,
+this DR-plan involves solving a quadratic system with 8 equations and variables.
 Instead of simultaneously solving this system, we could instead use the fact
@@
 the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
-solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
-with distinct orientation types $\sigma_p$,
+solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries),
+with distinct orientation types $\sigma_p$ 
 (two orientation choices for each of the 4 triangles)

### 2014-12-17 doc/GMP2015/s08_pinnedline.tex
Updated preliminary. Added not to recombination. Modified pinned line-incidence according to prelimary and updated DRP section.

@@
 the structure of  cross-linking microfibrils in collagen and cellulose in biology,
-by modeling these structures as {\em pinned line incidence systems}.
+by modeling these structures as {\em pinned line-incidence systems}.
 
@@
 
-\subsection{Modeling the fibrils as a pinned line incidence system}
+\subsection{Modeling the fibrils as a pinned line-incidence system}
 
-The cross-linking  microfibrils can be modeled as a pinned line incidence constraint system in $\mathbb{R}^2$,
+The cross-linking  microfibrils can be modeled as a pinned line-incidence constraint system in $\mathbb{R}^2$,
 where incidence constraints are used instead of distance constraints.
 
-%[2-Dimensional Pinned Line Incidence Graph] \label{prob:pinned_subspace}
-A {\em pinned line incidence system} $(G,\delta)$  
+A {\em pinned line-incidence system} $(G,\delta)$  
 is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$, 
@@
 % $G=(V,E, w_V, w_E)$  
-%%is a \textbf{2-dimensional pinned line incidence graph},
+%%is a \textbf{2-dimensional pinned line-incidence graph},
 % has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
@@
 %
-A pinned line incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$  \cite{xxx}. 
+A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$  \cite{xxx}.
 Note that no trivial motion exists since the pins have fixed positions on the plane.
+In particular, both a single vertex and a single edge are under-constrained graphs, 
+and the smallest over-constrained graph has $6$ vertices.
 
@@
 
-Figure~\ref{fig:pinned_line} shows an example of a  pinned line incidence graph,
+Figure~\ref{fig:pinned_line} shows an example of a  pinned line-incidence graph,
 where the grey ovals stand for pins representing attachments of fibrils,
@@
    \includegraphics[width=.8\linewidth]{img/pinned}
-\caption{ A well-constrained pinned line incidence graph.}
+\caption{ A well-constrained pinned line-incidence graph.}
 \label{fig:pinned_line}
@@
 
-\subsection{Optimal DR-plan for pinned line incidence systems}
+\subsection{Optimal DR-plan for pinned line-incidence systems}
 
@@
 In this section, we will adapt the results in Section~\ref{sec:DRP}
-to give the Optimal DR-plan for pinned line incidence graphs.
-First, we note the following differences of pinned line incidence graphs from the bar-joint graphs.
-
-\begin{itemize}
-
-\item For pinned line incidence graphs,
-since the pins have fixed positions on the plane,
-no trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
-In particular, note that both a single vertex and a single edge are under-constrained graphs.
-
-\item
-For pinned line incidence graphs,
-since the pins have fixed positions on the plane,
-{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
-being the disjoint union of two or more well-constrained subgraphs. \\
-%Without loss of generality,
-%we slightly modify the requirement of DR-plan for pinned line incidence graphs
-%such that only connected subgraphs can be considered as a node of the DR-plan.
-
-\end{itemize}
-
-We make  a few corresponding modifications to the definition of the DR-plan:
+to give the Optimal DR-plan for pinned line-incidence graphs.
+First, we note that %unlike bar-joint graphs, 
+{\em a well-constrained pinned line-incidence graph can be disconnected},
+being the disjoint union of two or more well-constrained subgraphs. 
+This is because the pins have fixed positions on the plane.
+%
+%the following difference of pinned line-incidence graphs from the bar-joint graphs.
+%
+%For pinned line-incidence graphs,
+%since the pins have fixed positions on the plane,
+%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
+%being the disjoint union of two or more well-constrained subgraphs. 
+%
+%\begin{itemize}
+%
+%\item For pinned line-incidence graphs,
+%since the pins have fixed positions on the plane,
+%no trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
+%In particular, note that both a single vertex and a single edge are under-constrained graphs.
+%
+%\item
+%For pinned line-incidence graphs,
+%since the pins have fixed positions on the plane,
+%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
+%being the disjoint union of two or more well-constrained subgraphs. \\
+%%Without loss of generality,
+%%we slightly modify the requirement of DR-plan for pinned line-incidence graphs
+%%such that only connected subgraphs can be considered as a node of the DR-plan.
+%
+%\end{itemize}
+%
+We make the following corresponding modification to the definition of the DR-plan:
 
@@
 %%\todo{We just define the %trivial graph/
-%%DR-plan leaves for 2-dimensional pinned line incidence graphs to be single vertices...  }
+%%DR-plan leaves for 2-dimensional pinned line-incidence graphs to be single vertices...  }
 %%\item \todo{A single edge can be a node of the DR-plan}
@@
 
- The \textbf{DR-plan} of  a 2-dimensional pinned line incidence graph $G$ is Definition~\ref{def:drp} with  modified rules number 2 and 4:
-%    The \textbf{decomposition-recombination (DR-) plan} of  a 2-dimensional pinned line incidence graph $G$, $DRP(G)$, is  a tree that has the following properties:
-    \begin{enumerate}
-        %\item The root of the tree `contains' $G$.
-        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
-        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
-        \item[4] A node containing a single edge is a leaf.
-    \end{enumerate}
+ The {\em DR-plan} of  a well-constrained pinned line-incidence graph $G$ is one in which 
+ each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+ % Definition~\ref{def:drp} with  modified rules number 2 and 4:
+%    The \textbf{decomposition-recombination (DR-) plan} of  a 2-dimensional pinned line-incidence graph $G$, $DRP(G)$, is  a tree that has the following properties:
+%    \begin{enumerate}
+%        %\item The root of the tree `contains' $G$.
+%        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+%        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
+%        \item[4] A node containing a single edge is a leaf.
+%    \end{enumerate}
 \end{definition}
@@
 we now adapt Theorem~\ref{theorem:main} 
-for the optimal DR-plan for  pinned line incidence graphs.
+for the optimal DR-plan for  pinned line-incidence graphs.
 
 \begin{corollary}
-Given a well-constrained  pinned line incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
+Given a well-constrained  pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
 \begin{enumerate}
-    \item if $C_i \cap C_j$ is a single vertex then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
+    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$.
     \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
@@
 \begin{itemize}
-\item At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is a vertex.
+\item At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial.
 \item Both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
@@
 
-Given Lemma~\ref{lem:pinned_union_intersection} and the modified definition of DR-plan,
-proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{corollary:no_edges_between_diff} straightforwardly extend to  pinned line incidence graphs.
-The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line incidence graphs 
+Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
+proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
+The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs 
 is given in Appendix~\ref{sec:appendix_pinned}.
-The following lemma serves as Lemma~\ref{lemma:uc_intersection_makes_all_uc} for pinned line incidence and can be proved similarly.
-\begin{lemma}
-\label{lem:vertex_intersection}
-If $C_i\cap C_j$ is a single vertex, then $\forall k: C_i\cap C_k$ is a single vertex.
-\end{lemma}
+%The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.
+%\begin{lemma}
+%\label{lem:vertex_intersection}
+%If $C_i\cap C_j$ is a single vertex, then $\forall k: C_i\cap C_k$ is a single vertex.
+%\end{lemma}
 
-Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lem:vertex_intersection},
+Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lemma:uc_intersection_makes_all_uc},
 it is straightforward to adapt the proof of Theorem~\ref{theorem:main}
-to  pinned line incidence graphs,
+to  pinned line-incidence graphs,
 thus proving Corollary~\ref{cor:pinned}.
-Consequently, we can efficiently find the optimal DR-plan for  pinned line incidence graphs using basically the same algorithm as for  bar-joint graphs.
+Consequently, we can efficiently find the optimal DR-plan for  pinned line-incidence graphs using basically the same algorithm as for  bar-joint graphs.
 
-{\bf Note. }The recombination problem for  pinned line incidence systems is trivial.
+\noindent {\bf Note:} The recombination problem for  pinned line-incidence systems is trivial.
 Since the pins are given fixed positions in the plane,

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Added Meera's subdivision figure and worked out some of the proof

@@
 \begin{figure}\centering
-    \begin{subfigure}{0.25\linewidth}
+    (a)
+    \begin{subfigure}{0.2\linewidth}
         \includegraphics[width=\linewidth]{img/pentl1}
@@
     % 
-    \begin{subfigure}{0.25\linewidth}
+    $\rightarrow$
+    \begin{subfigure}{0.2\linewidth}
         \includegraphics[width=\linewidth]{img/pentl2}
@@
     %
-    \begin{subfigure}{0.4\linewidth}
-        \includegraphics[width=\linewidth]{img/pentl3}
+
+    (b)
+    \begin{subfigure}{0.6\linewidth}
+        \includegraphics[width=\linewidth]{img/pentawesome}
     \end{subfigure}
-    \caption{An example of self-similarity in the form of repeated subdivisions.
-Each figure follows fromt he previous due to a set subdivision rule.
-If each figure was guaranteed to be isostatic, then isostatic systems of arbitrary detail could be made.}
+    \caption{Examples of self-similarity in the form of repeated subdivision.
+In (a), a simple subdivision is done that does not guarantee isostatic.
+In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.}
     \label{fig:subdivision}
@@
 \label{thm:1dofcase}
-    There is a quadratic algorithm for the 1-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
+    Given a 1-dof body-pin or multi-triangle graph $G$ and correpsonding 1-dof DR plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
 \end{theorem}
@@
 \begin{proof}
-    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof subgraph of $G$.
+    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof subgraph of $G$.
 
@@
 
-    Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+    Such a pinning covers all possible ways of adding a new body.
+Assume we add a new body $b$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
+Then, there will be some lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$.
+Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning $b$ to $b_i$ and $b_j$.
+
+    For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$.
+$|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$.
+We are looking for $b$ that minimizes $|T_b|$.
+Denote the subtree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$.
+Note that $fanin(b)= nl(T^b)$ because no descendent of $b$ is isostatic.
+Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$.
+All other nodes are not isostatic and so have no fanin.
+
+    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
 In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
@@
 
-    We claim that pinning a new body to a node in the manner described covers all possible ways of adding a new body.
-Assume we add a new body $b_n$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
-Then, there will be some lowest 1-dof node $v$ in our tree such that $b_i$ and $b_j$ appear in $v$.
-This could indeed be the root node.
-Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning a new body to $b_i$ and $b_j$.
-
-    It is the case that the node we need to pin can wlog be the deepest nontrivial node of some path in $T$.
+    The node we need to pin will always be the deepest nontrivial node of some path in $T$.
 Suppose we chose to pin a node $b$ that has a nontrivial child $v$.
-Then, $fanin(b)$ is the number of leaves in the subtree rooted at $b$, including the subtree rooted at $v$.
-If we had instead chosen to pin $v$, then $fanin(b')$ will just be the number of nodes in all its subtrees except for the one rooted at $v$.
-Then, $\max{fanin(b'),fanin(v)} < fanin(b)$, and for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $T_v$ has size at most $T_b$.
+Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$.
+If we had instead chosen to pin $v$, then $fanin_v(b) = nl(T^b) - nl(T^{b'}) + 1 \leq fanin_b(v)$.
+And for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $|T_v| \leq |T_b|$.
 Thus we only have to check the deepest non-trivial nodes.
@@
 Then, there will be some common ancestor $a$ of $v_i$ and $v_j$.
-Then, in $T_{v_i,v_j}$, $fanin(a)$ will be the number of leaves in the subtree rooted at $a$.
-However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin(a)'$ will be the number of leaves in its subtree excluding the subtree containing the node we pinned.
-Thus $fanin(a)' < fanin(a)$ and the size of $T_{v_i}$ is at most size of $T_{v_i,v_j}$.
+Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$.
+However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin_v(a) = nl(T^a) - nl(T^{a'}) +1$ .
+Thus $fanin_v(a)' \leq fanin_{v_i,v_j}(a)$.
+All ancestors of $a$ are unchanged.
+So $|T_v| \leq |T_{v_i,v_j}|$.
 
@@
 This is still a quadratic algorithm.
-
-\end{proof}
+% \end{proof}
 

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Typos

@@
-\section{Application: Finding completions of underconstrained glassy structures to isostatic}
+\section{Application: Finding completions of under-constrained glassy structures to isostatic}
 \label{sec:bodypin}
 
-We can use DR-plans to help in solving questions in designs for graphene and silica bilayers (\todo{Add Citation}).
-Specifically, we will look into the problem when you have a monolayer and it is underconstrained.
+We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers (\todo{Add Citation}).
+Specifically, we will look into the problem when you have a mono-layer and it is under-constrained.
 Our goal is to make such a layer isostatic.
@@
 \begin{enumerate}
-    \item Pin together 2 underconstrained monolayers in such a way that the resulting bilayer becomes isostatic (see Figure \ref{fig:silica_glass}).
+    \item Pin together 2 under-constrained mono-layers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
     \item Pin the boundary in such a way that the resulting system is isostatic.
@@
 In each case, we are specifically interested in how to they can be applied to get a resulting isostatic structure with a small DR-plan, so a realization can be found quickly.
-To answer these questions, we first inroduce the structures we model the problems with:
+To answer these questions, we first introduce the structures we model the problems with:
 
@@
 \begin{definition}
-    A \emph{body-hyper-pin qusec} is a qusec where the obejct are rigid bodies that are pinned together by some number of pins.
+    A \emph{body-hyper-pin qusec} is a qusec where the object are rigid bodies that are pinned together by some number of pins.
 \end{definition}
@@
 The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged.
-The optimal paramaterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
 
-For the rest of this section, we will only be dealing with the DR-pla n of such qusecs.
+For the rest of this section, we will only be dealing with the DR-plan of such qusecs.
 As such, we will refer only to the underlying graph of the qusecs.
-We now go over 2 subclasses of body-hyper-pin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
+We now go over 2 sub-classes of body-hyper-pin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
 
@@
 Body-pin graphs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
-Multi-triangle-pin graphs can be used to reprsent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (se Figure \ref{fig:silica_glass}).
+Multi-triangle-pin graphs can be used to represent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (see Figure \ref{fig:silica_glass}).
 Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan:
@@
 \begin{definition}
-    A \emph{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
+    A \emph{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
 In our case, the trivial nodes are just the bodies.
@@
         \item Its children will be 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph
-        \item Its children will be a number of proper maximal 1-dof subgraphs, joined pairwise by at most one pin
+        \item Its children will be a number of proper maximal 1-dof sub-graphs, joined pairwise by at most one pin
     \end{itemize}
@@
 
-    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof subgraphs $s_1, ..., s_k, k > 2$.
+    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
 Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
@@
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or multi-triangle graph $G$ and correpsonding 1-dof DR plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
+    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
 \end{theorem}
@@
 \begin{proof}
-    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof subgraph of $G$.
+    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
+Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.
 
@@
 Doing so will cause $G$ to become $(3,3)$-tight.
-We choose the 2 boodies to pin to by choosing a node $b$ in $T$ and looking at its children.
-From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a subgraph.
+We choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
+From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a sub-graph.
 We pin the new body to bodies in two separate children.
@@
 We are looking for $b$ that minimizes $|T_b|$.
-Denote the subtree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$.
-Note that $fanin(b)= nl(T^b)$ because no descendent of $b$ is isostatic.
+Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$.
+Note that $fanin(b)= nl(T^b)$ because no descendant of $b$ is isostatic.
 Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$.

### 2014-12-18 doc/GMP2015/s08_pinnedline.tex
Minor modification in pinned line-incidence

@@
 Note that no trivial motion exists since the pins have fixed positions on the plane.
-In particular, both a single vertex and a single edge are under-constrained graphs, 
-and the smallest over-constrained graph has $6$ vertices.
+In particular, both a single vertex and a single edge are under-constrained graphs.
+%and the smallest over-constrained graph has $6$ vertices.
 
@@
 %
-We make the following corresponding modification to the definition of the DR-plan:
+We define a {\em trivial} graph to be a single vertex and make the following modification to the definition of the DR-plan:
 
@@
  The {\em DR-plan} of  a well-constrained pinned line-incidence graph $G$ is one in which 
- each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+(1) each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$, and
+(2) a leaf node is a single edge.
  % Definition~\ref{def:drp} with  modified rules number 2 and 4:

### 2014-12-18 doc/GMP2015/s02_preliminaries.tex
Attempt to unify text decoration/emphasis. Make sure to pull, affects all tex files.

@@
 
-In this paper, a {\em geometric constraint system} is a multivariate polynomial 
-(usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
-between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.   
+In this paper, a \dfn{geometric constraint system} is a multivariate
+polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$,
+representing constraints with parameters $\delta$ between geometric
+primitives  in $\mathbb{R}^2$ represented collectively as $x\in
+\mathbb{R}^n$.
 %
-When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where 
-$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
-with the vertices $V$ representing the geometric primitives in 
-$\mathbb{R}^2$ and (hyper)edges
-$E$ representing the constraints, each with an associated 
-parameter $\delta$.
+When the type of constraint (system) is fixed, the system is simply
+represented as $(G,\delta)$, where $G$ is the underlying constraint
+(hyper)graph $G = (V,E)$ with the vertices $V$ representing the
+geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$
+representing the constraints, each with an associated parameter
+$\delta$.
 %
-For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
-$G=(V,E)$  
-with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
-this represents the distance constraint system $\| x_u -x_v \|_2 =
-\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
-$u\in V$.
+For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a
+graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E
+\rightarrow \mathbb{R}$; this represents the distance constraint
+system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where
+$x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
 
-\noindent {\bf Note:} Geometric constraint systems can also have inequalities, 
-where the parameters in $\delta$ are small intervals of values rather than exact values.
+\noindent
+\note Geometric constraint systems can also have
+inequalities, where the parameters in $\delta$ are small intervals of
+values rather than exact values.
 
-In this paper, we will consider 3 types of constraint systems: bar-joint
-as well as {\em pinned line-incidence}, and {\em body-hyperpin}. 
-In all of these cases, a cartesian \emph{realization} or 
-\emph{solution} $G(p)$ of $(G,\delta)$ is 
-an assignment of coordinates or Euclidean transformations (poses), 
-$p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the 
-constraints with parameters $\delta$, 
-modulo orientation preserving isometries 
+In this paper, we will consider 3 types of constraint systems: bar-
+joint as well as \dfn{pinned line-incidence}, and \dfn{body-hyperpin}.
+In all of these cases, a cartesian \dfn{realization} or \dfn{solution}
+$G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean
+transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or
+$\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with
+parameters $\delta$, modulo orientation preserving isometries
 (Euclidean rigid body motions for distance constraint).
 
-Although the realization space itself depends on the constraint parameters 
-$\delta$, many  relevant {\em generic} properties of 
-the constraint system $G(x,\delta)$
-are properties of the constraint (hyper)graph $G$ 
-and do not depend on $\delta$ (or they hold for all but a measure 
-zero set of $\delta$ values). Many of these are properties of the jacobian 
-$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of $G$} 
-(a matrix of indeterminates).
-For example, the {\em bar-joint rigidity matrix 
-of the graph $G = (V,E)$}
-is a matrix of indeterminates representing the 
-jacobian of the distance map 
-$\| x_u -x_v \|_2$  for $(u,v) \in E$.
-The matrix has $2$ columns per vertex in $V$ and one row per edge  in $E$, 
-where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates 
-for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), 
-that is 4 non-zero entries per row.
+Although the realization space itself depends on the constraint
+parameters $\delta$, many relevant \dfn{generic} properties of the
+constraint system $G(x,\delta)$ are properties of the constraint
+(hyper)graph $G$ and do not depend on $\delta$ (or they hold for all
+but a measure zero set of $\delta$ values). Many of these are
+properties of the jacobian $\Delta_x G(x,\delta)$, often called the
+appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
+For example, the \dfn{bar-joint rigidity matrix of the graph $G =
+(V,E)$} is a matrix of indeterminates representing the jacobian of the
+distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$
+columns per vertex in $V$ and one row per edge in $E$, where the row
+corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates
+for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\
+$v$), that is 4 non-zero entries per row.
 
-%XXXXXXXXXsomething has to be said 
-%along with asimow roth (infinitesimal rigidity 
-%implies rigidity)
-%When the rigidity matrix has appropriate rank, the realizations or solutions 
-%of the corresponding constraint system are generically isolated and zero-dimensional 
-%if the constraint system has equalities and exact values for the parameters $\delta$.
-%If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
-%this claim is approximate: the solutions are isolated small, full-dimensional  connected components.
+% XXXXXXXXX
+% something has to be said along with asimow roth
+% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
+% appropriate rank, the realizations or solutions of the corresponding
+% constraint system are generically isolated and zero-dimensional if the
+% constraint system has equalities and exact values for the parameters
+% $\delta$. If the constraint system has inequalities, i.e, if the
+% $\delta$ lie in an interval, this claim is approximate: the solutions
+% are isolated small, full-dimensional  connected components.
 
-One important property of generic constraint system or 
-(hyper)graph %that depends on the appropriate rigidity matrix of indeterminates
-is \emph{rigidity} ({\bf Note:} we refer to these as
-properties of the constraint system or as properties of the underlying
-(hyper)graph interchangeably),
-i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.
-%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth~\cite{???} shows a constraint (hyper)graph is rigid
-if it is generically {\em infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
+One important property of generic constraint system or (hyper)graph
+%that depends on the appropriate rigidity matrix of indeterminates
+is \dfn{rigidity} (\note we refer to these as properties of
+the constraint system or as properties of the underlying (hyper)graph
+interchangeably), i.e., the realizations or solutions of the
+corresponding constraint system being generically isolated and zero-
+dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
+The result by Asimow and Roth \uncited shows a constraint
+(hyper)graph is rigid if it is generically \dfn{infinitesimally
+rigid}, i.e.\ the number of independent rows of its appropriate
+rigidity matrix is at least the number of columns less the number of
+rigid body motions, which is 3 for bar-joint systems.
 %This condition is called generic {\em infinitesimal rigidity}.
 
-\noindent {\bf Note:} If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
-the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.
+\noindent
+\note If the constraint system has inequalities, i.e, if the $\delta$
+lie in an interval, the definition of rigidity is approximate: the
+solutions are isolated small, full-dimensional  connected components.
 % generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
 
-Other generic constraint system or (hyper)graph properties are the following: %({\bf Note:} we refer to these as
-%properties of the constraint system or as properties of the underlying
-%(hyper)graph interchangeably).
-%
-A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
-has independent rows
-(i.e, the determinant of some square submatrix is not
-identically zero). 
+Other generic constraint system or (hyper)graph properties are the
+following:
+% (\note we refer to these as properties of the constraint
+% system or as properties of the underlying (hyper)graph
+% interchangeably).
+A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate
+rigidity matrix of indeterminates has independent rows (i.e, the
+determinant of some square submatrix is not identically zero).
 %It is {\em rigid} if the number of independent rows of the rigidity matrix is
@@
 %which is 3 for distance constraint systems.
-It is {\em isostatic (minimally rigid, well-constrained)} 
-if it is both rigid and independent. 
+It is \dfn{isostatic (minimally rigid, well-constrained)} if it is
+both rigid and independent.
 %if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
@@
 %The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.
+It is \dfn{flexible} if it is not rigid, \dfn{under-constrained} if it is independent and not rigid, or \dfn{over-constrained} if it is not independent.
 
-Defining the independence of a subset of edges $E' \subseteq E$ 
-to be the independence of corresponding rows in the appropriate rigidity matrix of indeterminates,
-we obtain the {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$. 
+Defining the independence of a subset of edges $E' \subseteq E$ to be
+the independence of corresponding rows in the appropriate rigidity
+matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a
+constraint (hyper)graph $G = (V,E)$.
 %The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
@@
 %if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-There are various results on combinatorial characterization of independence and rigidity
-for different types of  rigidity matroid. 
-For bar-joint rigidity matroid, 
-the famous Laman's theorem~\cite{???} states that the underlying graph
-is minimally rigid if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. 
-The result by Lovasz and Yemini~\cite{???} shows that all 6-vertex-connected graphs are rigid in the plane. 
-For bar-body rigidity matroid,
-Tay~\cite{???} proved that
-the underlying multigraph is minimally rigid if and only if it can be decomposed as $3$ edge disjoint spanning trees. 
-White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
-Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid},
-where a hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
-and {\em $(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. 
-In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity
+There are various results on combinatorial characterization of
+independence and rigidity for different types of  rigidity matroid.
+For bar-joint rigidity matroid, the famous Laman's theorem \uncited
+states that the underlying graph is minimally rigid if and only if
+$|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
+at least 2 vertices. The result by Lovasz and Yemini \uncited shows
+that all 6-vertex-connected graphs are rigid in the plane. For bar-
+body rigidity matroid, Tay \uncited proved that the underlying
+multigraph is minimally rigid if and only if it can be decomposed as
+$3$ edge disjoint spanning trees. White and Whiteley \uncited gave
+the same characterization using a different technique to study the
+algebraic - geomtric conditions called pure condition. Lee, Streinu
+and Theran \uncited defined the \dfn{$(k,l)$-sparsity matroid},
+where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le
+k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2
+vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V|
+- l$. In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity
 condition is matroidal as long as $l \le dk-1$.
 
-In this paper, a {\em qusec} is any {\em independent} geometric 
-constraint system of one of the 3 types mentioned above.
-In the remainder of this section and the next two sections~\ref{sec:DRP} and \ref{sec:recomb} 
-we only consider bar-joint qusecs and graphs. 
-Relevant formal analogies  for the other 2 types of qusecs and (hyper)graphs 
-are given in the subsequent 2 sections on Applications.
+In this paper, a \dfn{qusec} is any \dfn{independent} geometric
+constraint system of one of the 3 types mentioned above. In the
+remainder of this section and the next two sections~\ref{sec:DRP} and
+\ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and
+(hyper)graphs are given in the subsequent 2 sections on Applications.
 
@@
 %TODO
-Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix
-is called a {\em flex vector}. 
-It has 2 entries per vertex and represents the internal motions of the system
-(i.e.\ modulo rigid body motions).
-For rigid graphs, all flex vectors are identically zero. 
+Given a bar-joint graph, a vector (of indeterminates) in the right
+null space of its rigidity matrix is called a \dfn{flex vector}. It
+has 2 entries per vertex and represents the internal motions of the
+system (i.e.\ modulo rigid body motions). For rigid graphs, all flex
+vectors are identically zero.
 %
-A vector (of indeterminates) in the left null space of the rigidity matrix
-is called a {\em stress vector}.
-It has one entry per edge and represents a {\em self stress} of the system.
-For independent graphs, all stress vectors are identically zero. 
+A vector (of indeterminates) in the left null space of the rigidity
+matrix is called a \dfn{stress vector}. It has one entry per edge and
+represents a \dfn{self stress} of the system. For independent graphs,
+all stress vectors are identically zero.
 %
-An {\em external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates)
-which specifies a 2D vector acting at each vertex
-such that the stress balance equation $sR = t$ holds, where $s$ is a vector
-of internal stresses, one for each bar, and $R$ is the rigidity matrix.
+An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of
+indeterminates) which specifies a 2D vector acting at each vertex such
+that the stress balance equation $sR = t$ holds, where $s$ is a vector
+of internal stresses, one for each bar, and $R$ is the rigidity
+matrix.
 
@@
 
-A {\em trivial} graph is   a single vertex.
+A \dfn{trivial} graph is a single vertex.
+%
 \begin{definition}\label{def:drp}
-    The  {\em decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
-    \cite{XX}
-    $DRP(G)$, is defined as a forest that has the following properties:
+    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
+    $G$, \uncited $DRP(G)$, is defined as a forest that has the
+    following properties:
     \begin{enumerate}
-        \item Each  node represents/contains/is  a rigid vertex-induced subgraph of $G$
-        %\item For a node, $C$, that is  a non-trivial graph, its $N$
-        %    children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper subgraphs of $C$.
-        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the 
-            vertex set of $C$.
+        \item Each node represents/contains/is a rigid vertex-induced
+        subgraph of $G$.
+        % \item For a node, $C$, that is  a non-trivial graph, its $N$
+        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
+        % subgraphs of $C$.
+        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
+        set of $C$.
         \item A leaf node is a trivial subgraph.
-       \item A root node is a maximal rigid subgraph of $G$
+        \item A root node is a maximal rigid subgraph of $G$.
     \end{enumerate}
@@
 
-    An equivalent DAG can be constructed from the above forest such that all nodes
-    containing the same subgraphs are identified into a single node.
-    In this manuscript, however, we retain a forest as defined above.
-    When a graph is isostatic, its DR plan is a tree.
+    An equivalent DAG can be constructed from the above forest such
+    that all nodes containing the same subgraphs are identified into a
+    single node. In this manuscript, however, we retain a forest as
+    defined above. When a graph is isostatic, its DR-plan is a tree.
 
@@
 
-    A DR-plan is {\em complete} if it satisfies an additional rule: 
+    A DR-plan is \dfn{complete} if it satisfies an additional rule:
         \item For a node, $C$, that is  a non-trivial graph, its $N$
-            children, $C_1, \ldots, C_N$, are all of the 
-            rigid vertex-maximal proper subgraphs of $C$.
-    This makes Rule 2 implicit, and in fact, $C$ contains all the edges in
-    the union of its children as well. The complete DR-plan is unique.
-
-    A DR-plan is {\em optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
+        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
+        and in fact, $C$ contains all the edges in the union of its
+        children as well. The complete DR-plan is unique.
+    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
 \end{definition}
@@
 
-\noindent{\bf Note:}
-In Section \ref{sec:DRP} $G$ is assumed to be isostatic.
-$Idc(G,X)$ is the subgraph of $G$ induced by the vertex set $X\subseteq V$. 
-For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
-In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$. $C_i$
-is the $i^{\text{th}}$ child of $C$. It is implied that 
-$C_i$ is an isostatic vertex-maximal proper subgraph.
+\noindent
+\note In Section \ref{sec:DRP} $G$ is assumed to be isostatic.
+$Idc(G,X)$ is the subgraph of $G$ induced by the vertex set
+$X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
+In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
+$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
+an isostatic vertex-maximal proper subgraph.
 
@@
 The use of a DR-plan in designing especially self-similar qusecs is
-immediately clear, as shown in Figures \ref{XX} and \ref{XX}.
-In fact, knowing an optimal DR-plan is crucial for design and analysis of 
-even non-self-similar qusecs 
-and their realizations with various desirable properties.
+immediately clear, as shown in Figures \ref{XX} and \ref{XX}
+\todo{Reference?}. In fact, knowing an optimal DR-plan is crucial for
+design and analysis of even non-self-similar qusecs and their
+realizations with various desirable properties.
 
-First, note that the basic task of \underbar{{\sl realizing given qusecs}} would require finding real
- solutions 
- a large multivariate polynomial system (of inequalities and equalities
- representing the constraints). Without DR-planning, this requires double
- exponential time in the number of variables
- (even if orientation type is specified).
-With DR-planning, the complexity is dominated by the size of the largest 
-subsystem that is
-solved, or recombined from the solutions of its child subsystems i.e.,the maximum fan-in occuring in a DR-plan.
-Thus finding an optimal DR-plan is important for realization.
+First, note that the basic task of \header{realizing given qusecs}
+would require finding real solutions a large multivariate polynomial
+system (of inequalities and equalities representing the constraints).
+Without DR-planning, this requires double exponential time in the
+number of variables (even if orientation type is specified). With DR-
+planning, the complexity is dominated by the size of the largest
+subsystem that is solved, or recombined from the solutions of its
+child subsystems i.e.,the maximum fan-in occuring in a DR-plan. Thus
+finding an optimal DR-plan is important for realization.
 
-Next, using the DR-plan we obtain a \underbar{{\sl recursive block
-decomposition of the rigidity matrix}}   
-with each block being the rigidity matrix of 
-an isostatic subgraph.
-The blocks intersect on isostatic or trival subgraphs.
-This recursive block decomposition of the rigidity matrix $R$ 
-correspondingly \underbar{{\sl recursively decomposes the stress vectors}}
-$s$ that balance a given
-external stress $t$ since $sR = t$, and moreover, the external stress vector
-$t$ now \underbar{{\sl recursively "distributes" as external stresses}} 
-for each isostatic subgraph of the DR-plan.
-When the rigidity matrix and stress vectors are indeterminates, it can be
-used to design  qusecs underlying graph and parameters $\delta$ to obtain an  
-optimal distribution of stresses on the geometric primitives. 
-When the realization of the qusec is given, this can be used to 
-analyze the distribution of a  given external load.
+Next, using the DR-plan we obtain a \header{recursive block
+decomposition of the rigidity matrix} with each block being the
+rigidity matrix of an isostatic subgraph. The blocks intersect on
+isostatic or trival subgraphs. This recursive block decomposition of
+the rigidity matrix $R$ correspondingly \header{recursively decomposes
+the stress vectors} $s$ that balance a given external stress $t$ since
+$sR = t$, and moreover, the external stress vector $t$ now
+\header{recursively ``distributes'' as external stresses} for each
+isostatic subgraph of the DR-plan. When the rigidity matrix and stress
+vectors are indeterminates, it can be used to design  qusecs
+underlying graph and parameters $\delta$ to obtain an optimal
+distribution of stresses on the geometric primitives. When the
+realization of the qusec is given, this can be used to analyze the
+distribution of a given external load.
 

### 2014-12-18 doc/GMP2015/s04_recombination.tex
Attempt to unify text decoration/emphasis. Make sure to pull, affects all tex files.

@@
 
-In this section, we consider the {\em optimal recombination}
-problem of combining specific solutions of
-subsystems in a DR-plan into a solution of their parent system $I$, (without
-loss of generality, at the top level of the DR-plan). In the case of isostatic
-qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
-seek  {\em solution(s) (among a finite large number of solutions)
-with a specific orientation or chirality}.
-In the case of underconstrained qusecs the subsystems are the multiple roots
-of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
-an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
-chirality}.
+In this section, we consider the \dfn{optimal recombination} problem
+of combining specific solutions of subsystems in a DR-plan into a
+solution of their parent system $I$, (without loss of generality, at
+the top level of the DR-plan). In the case of isostatic qusecs, the
+parent system $I$ is isostatic (the root of the DR-plan), and we seek
+\dfn{solution(s) (among a finite large number of solutions) with a
+specific orientation or chirality}. In the case of underconstrained
+qusecs the subsystems are the multiple roots of the DR-plan, the
+parent system $I$ is underconstrained, and we typically seek an
+efficient algorithmic description of \dfn{connected component(s) of
+solutions with a specific orientation or chirality}.
 
-The main barrier in recombination when given an optimal DR-plan (of smallest possible
-size or max fan-in),  is that the number of children (resp. number of
-roots) - and correspondingly the  size and complexity of the (indecomposable) algebraic
-system $I$ to be solved -
-could be arbitrarily large as a function of the size of the input system.
-This difficulty can persist even after optimal parametrization of the
-indecomposable system $I$ as in \cite{XX}
-to minimize its algebraic complexity.
+The main barrier in recombination when given an optimal DR-plan (of
+smallest possible size or max fan-in),  is that the number of children
+(resp. number of roots) - and correspondingly the  size and complexity
+of the (indecomposable) algebraic system $I$ to be solved - could be
+arbitrarily large as a function of the size of the input system. This
+difficulty can persist even after optimal parametrization of the
+indecomposable system $I$ as in \uncited to minimize its algebraic
+complexity.
 
-In the following, we formulate the problem of {\em optimal
-modification},
-of the indecomposable
-algebraic system $I$ into a decomposable system with a small DR-plan (low
-algebraic complexity).
-Leveraging recent results on {\em Cayley
-configuration spaces}, our approach to the optimal modification problem
-achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
-$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$,  all of which have
-small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
-potentially be solved or realized easily once the orientation type of the solution
-is known  (when the DR-plan size is small enough).
-{\em (b) Solution preservation.} Moreover, the union of solution spaces of the systems in the family
-$I_{\lambda_F}$
-is guaranteed to contain
-all of $I$'s solutions. {\em (c) Efficient search.} Finally, this union can be projected into the
-so-called {\em Cayley} or distance parameter space $\lambda_F$ that is convex
-or otherwise easy to traverse in order to search for $I$'s solution
-(or connected component)
-of the desired orientation type.
-For the case when the modification is bounded, this approach provides an efficient
-algorithm for recombination. We first define the decision version of the
-problem of optimal modification
-for decomposition. The standard optimization versions are straightforward.
+In the following, we formulate the problem of \dfn{optimal
+modification}, of the indecomposable algebraic system $I$ into a
+decomposable system with a small DR-plan (low algebraic complexity).
+Leveraging recent results on \dfn{Cayley configuration spaces}, our
+approach to the optimal modification problem achieves the following:
+\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
+$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters
+$F$,  all of which have small DR-plans. Thus, given a value $v$ for
+$\lambda_F$, the system $I_v$ can potentially be solved or realized
+easily once the orientation type of the solution is known  (when the
+DR-plan size is small enough). \dfn{(b) Solution preservation.}
+Moreover, the union of solution spaces of the systems in the family
+$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
+so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
+convex or otherwise easy to traverse in order to search for $I$'s
+solution (or connected component) of the desired orientation type. For
+the case when the modification is bounded, this approach provides an
+efficient algorithm for recombination. We first define the decision
+version of the problem of optimal modification for decomposition. The
+standard optimization versions are straightforward.
 
-\medskip\noindent{\bf Optimal Modification for Decomposition (OMD) Problem.} Given a generically
-independent graph $G = (V,E)$
-with no non-trivial proper isostatic subgraph (indecomposable), and 2 constants $k$ and $s$,
-does there exist a set of at most $k$  edges $E_1$ and a set of nonedges $F$
-such that the modified graph $H = (V, E\setminus E_1 \cup F)$ has a DR-plan
-of size at most $s$?  The OMD$_k$ problem is  OMD where $k$ is a fixed bound (not
-part of the input).   We say that such a tuple $(G,s)$ is a member of the
-set OMD$_k$. We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
+\problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
+Given a generically independent graph $G = (V,E)$ with no non-trivial
+proper isostatic subgraph (indecomposable), and 2 constants $k$ and
+$s$, does there exist a set of at most $k$  edges $E_1$ and a set of
+nonedges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
+F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
+where $k$ is a fixed bound (not part of the input).   We say that such
+a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to
+graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
 with appropriately small $s$.
 
-It is immediately clear that indecomposable graphs $G$ that belong in OMD$_k$ for small $k$ and
-$s$  lend themselves to modification  into decomposable graphs
-satisfying Criteria (a) and (b) above.
-However, it is not clear how Criterion (c) is met by OMD graphs.
+It is immediately clear that indecomposable graphs $G$ that belong in
+OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
+decomposable graphs satisfying Criteria (a) and (b) above. However, it
+is not clear how Criterion (c) is met by OMD graphs.
 %
 \subsection{Using Convex Cayley Configuration Spaces}
+\label{sec:2-tree-reduction}
 %
-\ref{2-tree-reduction}
-Next we provide the necessary background to describe a specific approach for achieving the requirements
-(a)-(c) mentioned above, by restricting the class of reduced graphs $G' =
-G\setminus E_1$ and their isostatic completions $H$ in the above definition of the
-OMD problem, and using a key theorem of Convex Cayley configuration
-spaces \cite{XX}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley
-parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of
-the nonedges $F$
-is always convex for any
-given lengths $\delta$ for the edges of $H$ (i.e, over all the realization
-s of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
-This set is called the (2-dimensional) {\em Cayley configuration space}
-of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted
-$\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian
-realization space of $(H,\delta)$ on the Cayley parameters $F$.
-Such graphs $H$ are said to have {\em convexifiable Cayley configuration spaces
-with parameters $F$} (short: {\em convexifiable}). To state the theorem, we first have to define the
-notion of {\em 2-sums} and {\em 2-trees}.
-Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and
-$V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
-respectively.  A {\em 2-sum} of
-$H_1$ and $H_2$ is a
-graph $H$  obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$.
-In this case, $H_1$ and $H_2$ are called {\em 2-sum components} of $H$.
-A {\em minimal 2-sum component} of $H$ is  one that cannot be further split
-into 2-sum components.
-A {\em 2-tree} is recursively obtained by taking
-a 2-sum of two 2-trees, with the base case of 2-tree being a triangle.
-A {\em partial 2-tree} is a 2-tree minus some edges.
-Partial 2-trees are characterized as having $K_4$ as forbidden minor are also called series parallel graphs.
+Next we provide the necessary background to describe a specific
+approach for achieving the requirements (a)-(c) mentioned above, by
+restricting the class of reduced graphs $G' = G\setminus E_1$ and
+their isostatic completions $H$ in the above definition of the OMD
+problem, and using a key theorem of Convex Cayley configuration spaces
+\uncited. This theorem characterizes the class of graphs $H$ and non-
+edges $F$ (Cayley parameters), such that the set of vectors
+$\lambda_F$ of  attainable lengths of the nonedges $F$ is always
+convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
+all the realization s of the bar-joint constraint system or linkage
+$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
+\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
+Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
+a ``projection'' of the cartesian realization space of $(H,\delta)$ on
+the Cayley parameters $F$. Such graphs $H$ are said to have
+\dfn{convexifiable Cayley configuration spaces with parameters $F$}
+(short: \dfn{convexifiable}). To state the theorem, we first have to
+define the notion of \dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and
+$H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with
+edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
+respectively.  A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
+obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
+and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
+components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
+that cannot be further split into 2-sum components. A \dfn{2-tree} is
+recursively obtained by taking a 2-sum of two 2-trees, with the base
+case of 2-tree being a triangle. A \dfn{partial 2-tree} is a 2-tree
+minus some edges. Partial 2-trees are characterized as having $K_4$ as
+forbidden minor are also called series parallel graphs.
 
-\begin{theorem}
-    \label{convexcayley}
-    \cite{XX}
-    $H$ has a convexifiable Cayley configuration space  with parameters $F$
-    if and only if for each $f\in F$  all the minimal 2-sum components
-    of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees.
-    The Cayley configuration space $\Phi_F(H,\delta)$
-    of a bar-joint system or linkage $(H,\delta)$ is a convex polytope.
-    When $H\cup F$ is a 2-tree, the
-    bounding hyperplanes of this polytope are triangle inequalities
-    relating the lengths of edges of the triangles in $H\cup F$.
+\begin{theorem}\label{theorem:convexcayley}
+    \uncited $H$ has a convexifiable Cayley configuration space  with
+    parameters $F$ if and only if for each $f\in F$  all the minimal
+    2-sum components of $H\cup F$ that contain both endpoints of $f$
+    are partial 2-trees. The Cayley configuration space
+    $\Phi_F(H,\delta)$ of a bar-joint system or linkage $(H,\delta)$
+    is a convex polytope. When $H\cup F$ is a 2-tree, the bounding
+    hyperplanes of this polytope are triangle inequalities relating
+    the lengths of edges of the triangles in $H\cup F$.
 \end{theorem}
 
-The idea of our approach to achieve the criteria
-(a)-(c) begins with the following simple but useful theorem.
+The idea of our approach to achieve the criteria (a)-(c) begins with
+the following simple but useful theorem.
 
-\begin{theorem}
-    \label{omdk}
-Given an indecomposable graph $G$,
-let $G'$ be a spanning partial 2-tree subgraph in $G$ with
-$k$ fewer edges than $G$.
-Then  $(G,2)$ belongs in the set OMD$_k$.
+\begin{theorem}\label{theorem:omdk}
+    Given an indecomposable graph $G$, let $G'$ be a spanning partial
+    2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then
+    $(G,2)$ belongs in the set OMD$_k$.
 \end{theorem}
@@
 \begin{proof}
-The proof follows from the fact that 2-trees are well decomposable and have
-simple DR-plans of size 2.
-We know that $G$ can be reduced
-by removing $k$ edges to create a partial-2-tree $G'$
-which can then be completed to an (isostatic) 2-tree by adding some set of non-edges
-$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2, proving the theorem.
+The proof follows from the fact that 2-trees are well decomposable and
+have simple DR-plans of size 2. We know that $G$ can be reduced by
+removing $k$ edges to create a partial-2-tree $G'$ which can then be
+completed to an (isostatic) 2-tree by adding some set of non-edges
+$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
+proving the theorem.
 \end{proof}.
 
-We refer to such graphs $G$ in short as {\em $k$-approximately
-convexifiable,} where the reduced graphs $G'$ and isostatic completions $H$
-are convexifiable.
-As observed earlier, since graphs such as $G$ are in OMD$_k$, Criteria (a) and (b)
-are automatically met for small enough $k$.
-Criterion $c$ is addressed
-as described in the following efficient search procedure
-which clarifies the dependence of the complexity on the number and ranges  of Cayley
-parameters $F$.
+We refer to such graphs $G$ in short as \dfn{$k$-approximately
+convexifiable}, where the reduced graphs $G'$ and isostatic
+completions $H$ are convexifiable. As observed earlier, since graphs
+such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
+for small enough $k$. Criterion $c$ is addressed as described in the
+following efficient search procedure which clarifies the dependence of
+the complexity on the number and ranges  of Cayley parameters $F$.
 
-\begin{theorem}
-    \label{criterionc}
+\begin{theorem}\label{theorem:criterionc}
     [Efficient search]
     For an indecomposable, $k$-approximately convexifiable graph $G =
-    (V,E)$,
-   let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph
-    where $|D| \le  k$. Let  $F$ be a set of nonedges of $G$ such that
-    $H = (V, E'\cup F)$ is a 2-tree.
-    Each solution $p$ (or connected component of a solution space)
-    of $(G,\delta)$
-    of an orientation type $\sigma_p$ can be found in time $O(\log(W))$ where
-    $W$ is the number of cells of desired accuracy (discrete volume) of
-    the convex
-    polytope $\Phi_F(G',\delta_E')$.  The (discrete) volume $W$ is exponential in
-    $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
+    (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
+    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of nonedges
+    of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
+    $p$ (or connected component of a solution space) of $(G,\delta)$
+    of an orientation type $\sigma_p$ can be found in time
+    $O(\log(W))$ where $W$ is the number of cells of desired accuracy
+    (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
+    The (discrete) volume $W$ is exponential in $|F|$ and polynomial
+    in the (discrete range) of the parameters in $F$.
 \end{theorem}
@@
 \begin{proof}
-    The cartesian realization
-    space of $(H,<\delta_E', \lambda_F>)$ is
-    computed easily with a DR-plan of size 2, and is the union of $2^t$
-    solutions (modulo orientation preserving isometries) each with a distinct orientation type, where $t$ is the number of triangles in the 2-tree
-    $H$;
-    here $\delta_{S}$ is the restriction of the length vector $\delta$ to
-    the edges in $S$.
-    A desired solution $p$ (or connected component of a solution space)
-    of $(G,\delta)$
-    of an orientation type $\sigma_p$ can be found
-    by a subdivided binary search   of the cartesian realization space of
-    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ -
-    as $\lambda_F$ ranges over
-    the discretized convex
-    polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
-    Theorem \ref{convexcayley}.
-    A solution $p$  is found  when the lengths for nonedges in $D$ match
-    $\delta_D$.
+    The Cartesian realization space of $(H,<\delta_E', \lambda_F>)$ is
+    computed easily with a DR-plan of size 2, and is the union of
+    $2^t$ solutions (modulo orientation preserving isometries) each
+    with a distinct orientation type, where $t$ is the number of
+    triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction
+    of the length vector $\delta$ to the edges in $S$. A desired
+    solution $p$ (or connected component of a solution space) of
+    $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a
+    subdivided binary search   of the cartesian realization space of
+    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ - as
+    $\lambda_F$ ranges over the discretized convex polytope
+    $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
+    Theorem \ref{theorem:convexcayley}. A solution $p$  is found  when
+    the lengths for nonedges in $D$ match $\delta_D$.
 \end{proof}
@@
 
-\noindent{\bf Note:} A huge advantage of the convex Cayley method
-is that it is completely unaffected when $\delta$ are intervals rather than exact values~\cite{GaoSitharam}.
+\noindent
+\note A huge advantage of the convex Cayley method is that it is
+completely unaffected when $\delta$ are intervals rather than exact
+values~\cite{GaoSitharam}.
 %cite my paper with Heping!!!
 
-\medskip\noindent{\bf Example}.
-A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
-graphs, i.e, its DR-plan has a root and 9
-children corresponding to the 9 edges. Solving or recombining the system
-$(G,\delta)$
-corresponding to the root of
-this DR-plan involves solving a quadratic system with 8 equations and variables.
-Instead of simultaneously solving this system, we could instead use the fact
-that $G=K_{3,3}$ is in OMD$_1$: remove the
-edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
-Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
-The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
-is a single interval of attainable length values $\lambda_F$ for
-the edge $f$.
-When $\delta$ is generic, i.e, does not admit collinearities or coincidences
-in the realizations of $(G,\delta)$,
-the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
-solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries),
-with distinct orientation types $\sigma_p$ 
-(two orientation choices for each of the 4 triangles)
-that can be obtained by solving a sequence of 4 single quadratics in 1
-variable (DR-plan of size 2).
-By subdivided
-binary search
-in the interval $\lambda_f \in
-\Phi_f(G', \delta_{E\setminus e})$, the desired
-solution $p$  of $(G,\delta)$ is found when the length of the nonedge
-$e$ in  the realization
-$q^p_{\lambda_f}$
-is $\delta_e$.
+\medskip\noindent
+\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
+isostatic graphs, i.e, its DR-plan has a root and 9 children
+corresponding to the 9 edges. Solving or recombining the system
+$(G,\delta)$ corresponding to the root of this DR-plan involves
+solving a quadratic system with 8 equations and variables. Instead of
+simultaneously solving this system, we could instead use the fact that
+$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
+\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
+nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
+configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
+interval of attainable length values $\lambda_F$ for the edge $f$.
+When $\delta$ is generic, i.e, does not admit collinearities or
+coincidences in the realizations of $(G,\delta)$, the realization
+space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16 solutions
+$q^p_{\lambda_f}$ (modulo orientation preserving isometries), with
+distinct orientation types $\sigma_p$ (two orientation choices for
+each of the 4 triangles) that can be obtained by solving a sequence of
+4 single quadratics in 1 variable (DR-plan of size 2). By subdivided
+binary search in the interval $\lambda_f \in \Phi_f(G',
+\delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is
+found when the length of the nonedge $e$ in  the realization
+$q^p_{\lambda_f}$ is $\delta_e$.
 %
@@
 Graphs}
-\label{tdecomp}
-It is possible in principle to decrease $k$ for a OMD$_k$ graph
-(i.e, the number of edges to be removed
-to ensure an isostatic completion that is decomposable with a small DR-plan) by
-considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
-partial 2-trees but nevertheless have convex Cayley configuration spaces at
-least when the realization space is restricted to a sufficiently
-comprehensive orientation type.
-In particular, the so-called {\em tree-decomposable graphs of low Cayley
-complexity} \cite{XX} include the partial 2-trees and many others that are
-not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
-OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
-- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
-linkages based on such graphs $G'$
-are known to be single intervals
-when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
-In addition, the bounds of these intervals are of low algebraic
-complexity.
-More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
-i.e, the computation of these bounds is tree-decomposable. Alternatively,
-the bounds are in a simple quadratic or radically solvable extension field over the rationals,
-or can be computed by solving a
-triangularized system of quadratics.
+\label{sec:tdecomp}
+%
+It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
+the number of edges to be removed to ensure an isostatic completion
+that is decomposable with a small DR-plan) by considering reduced
+graphs $G'$ (and modified graphs $H$) that come from a larger class
+than partial 2-trees but nevertheless have convex Cayley configuration
+spaces at least when the realization space is restricted to a
+sufficiently comprehensive orientation type. In particular, the so-
+called \dfn{tree-decomposable graphs of low Cayley complexity}
+\uncited include the partial 2-trees and many others that are not
+partial 2-trees. These too result in DR-plans of size 2 or 3, putting
+$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
+Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
+configuration spaces of linkages based on such graphs $G'$ are known
+to be single intervals when a comprehensive orientation type
+$\sigma_p$ of the sought solution $p$ is given. In addition, the
+bounds of these intervals are of low algebraic complexity. More
+precisely, the bounds  can themselves be computed using a DR-plan of
+size 2 or 3, i.e, the computation of these bounds is tree-
+decomposable. Alternatively, the bounds are in a simple quadratic or
+radically solvable extension field over the rationals, or can be
+computed by solving a triangularized system of quadratics.

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Updated the subdivision figure
Attempt to unify text decoration/emphasis. Make sure to pull, affects all tex files.

@@
 
-We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers (\todo{Add Citation}).
+We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \uncited.
 Specifically, we will look into the problem when you have a mono-layer and it is under-constrained.
@@
     \caption{Example of a single mono-layer of a Silica Oxygen glassy structure.
-This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
+This can be thought of as a multi-triangle-pin qusecs where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
 Not shown here are other layers stacked on this one.
@@
     (b)
-    \begin{subfigure}{0.6\linewidth}
+    \begin{subfigure}{0.8\linewidth}
         \includegraphics[width=\linewidth]{img/pentawesome}
@@
 \begin{definition}
-    A \emph{body-hyper-pin qusec} is a qusec where the object are rigid bodies that are pinned together by some number of pins.
+    A \dfn{body-hyperpin qusecs} is a qusecs where the object are rigid bodies that are pinned together by some number of pins.
 \end{definition}
@@
 \begin{remark}
-    A body-hyper-pin qusec is a special case of bar-joint qusecs that we have been studying.
+    A body-hyperpin qusecs is a special case of bar-joint qusecs that we have been studying.
 As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
@@
 As such, we will refer only to the underlying graph of the qusecs.
-We now go over 2 sub-classes of body-hyper-pin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
+We now go over 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
 
@@
 \label{def:body-pin}
-    A body-pin graph is a body-hyper-pin graph with the following conditions:
+    A body-pin graph is a body-hyperpin graph with the following conditions:
     \begin{itemize}
@@
     \end{itemize}
-    Such a body-pin graph can also be seen as a \emph{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
+    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
 Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
@@
 \begin{definition}
-    A \emph{multi-triangle-pin graph} is a body-hyper-pin graph where each body is a triangle, meaning it is pinned in at most 3 places.
+    A \dfn{multi-triangle-pin graph} is a body-hyperpin graph where each body is a triangle, meaning it is pinned in at most 3 places.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
@@
 \begin{definition}
-    A \emph{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
+    A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
 In our case, the trivial nodes are just the bodies.
@@
 \begin{observation}
-    For the 1-dof body-hyper-pin graphs described above, a DR-plan node will follow one of the following:
+    For the 1-dof body-hyperpin graphs described above, a DR-plan node will follow one of the following:
 
@@
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
+    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
 \end{theorem}
@@
 
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX QUESTION XXXXXXXXXXXXXXXXXXXXXXXXX
-An open question remains as to the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(l,k)$ characterization is no longer matroidal.
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(k,l)$ characterization is no longer matroidal.
+\end{openproblem}

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Added another observation about algebraic complexity

--- 2014-12-18 doc/GMP2015/s07_problempinning.tex
@@
     \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
-    \raisebox{0.5\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
+    \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
     \caption{Example of a single mono-layer of a Silica Oxygen glassy structure.
@@
 
-% THIS NEEDS SOME MORE THINKING. THERE ARE DIFFERENT TYPES OF CHILDREN NODES FOR THIS PART. NOT JUST THE TWO LISTED ABOVE
-
 \begin{observation}
@@
 
-% \begin{theorem}
-%     There is a \todo{$O(?)$} algorithm for the 2-dof optimal completion problem of Seciton \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
-% \end{theorem}
+\begin{observation}
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan with minimum fanin, we did not consider the algebraic complexity of realizing the DR-plan.
+If that is preferred, we can no longer only consider the deepest nodes.
+\end{observation}
 
-% \begin{proof}
-%     We can do a similar algorithm to the 1-dof case, except we constrain a given node 2 times (by adding 2 triangles or 2 bodies and 4 pins).
-This causes the nodes along the path to the given node to become well-constrained while all others will be underconstrained.
-Thus, our DR-plan will have the same size as the 1-dof case.
-This is still a quadratic algorithm.
-% \end{proof}
+\begin{proof}
+    An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton).
+These are the Euclidean motions.
+A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has.
+A 2-dof has 5 parameters, and so on.
+These paramters denote the algebraic complexity of giving a realization for the graphs.
+
+    Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}.
+When we pin a node $b$, we will have the same structure as before.
+Suppose we are looking at an isostatic node $v$ after pinning $b$.
+Then, the children of $v$ (except one if $v \neq b$) will be 1-dof.
+The complexity for that node is then just that of solving each of its children.
+In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.
+
+    Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$.
+In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children.
+SO we will have to try pinning all nodes to see which gives the lowest algebraic complexity.
+This is still quadratic for the 1-dof case.
+
+    For the 2-dof case, there are more cases to consider.
+If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$.
+If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof.
+Note that solving $v'$ will result in solving $v$.
+Then, we need to consider nodes above and includiing $v'$ in our complexity.
+$np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
+The only remaining case will be if we pin two nodes that are not descendent/ancestor.
+the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.
+For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.
+
+    Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin.
+However, we cannot also assume pinning one node twice will give us the best algebraic complexity.
+Hence, we will need to check each pair of node to pin.
+This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
+\end{proof}
 

### 2014-12-18 doc/GMP2015/s01_introduction.tex

@@
 
-\newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}
+% \newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}
 
@@
 of multiple, inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either {\sl isostatic or
+Furthermore, a layer is often  either \vemph{isostatic or
 under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
@@
 
-    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \cite{XX}
+    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited
 
@@
 of a geometric constraint system of appropriate types of geometric
-primitives, under metric or algebraic constraints. Such 2D {\bf{\em
-qusecs}} (quasi-uniform or self-similar constraint systems) can be
+primitives, under metric or algebraic constraints. Such 2D
+\dfn{qusecs} (quasi-uniform or self-similar constraint systems) can be
 used to understand or design material layers (their solutions) with
@@
 We now briefly survey existing techniques for studying 2D
-\textit{qusecs},  many of which are \textit{bar-joint} systems
+\dfn{qusecs},  many of which are \dfn{bar-joint} systems
 (Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
-\textit{body-hyperpin} systems (Example 4,5, see Section
-\ref{sec:bodypin}) or \textit{pinned-line incidence} systems (Example
+\dfn{body-hyperpin} systems (Example 4,5, see Section
+\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
 3, see Section \ref{sec:pinnedline}). The limitations of these
@@
 
-\meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
+\medskip\noindent
+\header{(i) Finding (vertex)-maximal generically rigid subsystems}
 Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
-locating all maximal, generically rigid subsystems \cite{XX}. When the
+locating all maximal, generically rigid subsystems \uncited. When the
 input is rigid, these algorithms compute the identity and output the
@@
 
-\meeraheader{ii}{Optimal Recursive Decomposition (DR-planning)}
+\medskip\noindent
+\header{(ii) Optimal Recursive Decomposition (DR-planning)}
 Recursive decomposition of geometric constraint systems has been
-formalized \cite{XX} and well-studied \cite{XX} as the {\sl
-Decomposition-Recombination (DR)-planning} problem. For the above-
+formalized \uncited and well-studied \uncited as the \dfn
+{Decomposition-Recombination (DR)-planning} problem. For the above-
 mentioned classes of 2D qusecs, generic rigidity is a combinatorial
@@
 defining desirable or optimal DR-plans and DR-planning algorithms were
-given in \cite{XX}. An \emph{optimal DR-plan} is one that minimizes
+given in \uncited. An \dfn{optimal DR-plan} is one that minimizes
 the maximum number of child subsystems of any parent system, a
@@
 systems, the optimal DR-planning problem was shown to be NP-hard
-\cite{XX}.
+\uncited.
 
-\meeraheader{iii}{DR-plans for special classes and with other
+\medskip\noindent
+\header{(iii) DR-plans for special classes and with other
 criteria}
-For a special class of 2D qusecs, namely \emph{tree- decomposable}
-systems \cite{XX} common in computer aided mechanical design, (which
+For a special class of 2D qusecs, namely \dfn{tree-decomposable}
+systems \uncited common in computer aided mechanical design, (which
 includes ruler-and-compass and Henneberg-I constructible systems), all
-DR-plans turn out to be optimal. This satisfies the so-called \emph
+DR-plans turn out to be optimal. This satisfies the so-called \dfn
 {Church-Rosser} criterion, leading to highly efficient DR-planning
 algorithms. For general 2D qusecs, alternate criteria were suggested
-such as \emph{cluster minimality} requiring parent systems to be
+such as \dfn{cluster minimality} requiring parent systems to be
 composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
-no proper subset forms a rigid system); and \emph{proper maximality},
+no proper subset forms a rigid system); and \dfn{proper maximality},
 requiring child subsystems to be maximal rigid proper subsystems of
@@
 
-While polynomial time algorithms were given \cite{XX} to generate DR-
+While polynomial time algorithms were given \uncited to generate DR-
 plans meeting the cluster minimality criterion, no such algorithm is
@@
 
-\meeraheader{iv}{Optimal Recombination and Solution Space Navigation}
+\medskip\noindent
+\header{(iv) Optimal Recombination and Solution Space Navigation}
 For the entire DR-plan, finding all desired solutions is barely
@@
 problem of finding even a single solution to the input system at the
-root of the DR-plan is NP-hard \cite{XX} and there is a combinatorial
-explosion of solutions \cite{XX}. Typically, however, the desired
+root of the DR-plan is NP-hard \uncited and there is a combinatorial
+explosion of solutions \uncited. Typically, however, the desired
 solution has a given orientation type, in which case, the crux of the
@@
 parameterization to minimize the algebraic complexity of the
-indecomposable parent (sub)systems that occur in the DR-plan \cite{XX,
-XX,XX}, provided the DR-plan meets some of the abovementioned
+indecomposable parent (sub)systems that occur in the DR-plan \uncited\uncited\uncited, provided the DR-plan meets some of the abovementioned
 criteria.
@@
 principled basis, high performance heuristics and software exists
-\cite{XX,XX} to tame combinatorial explosion via user intervention.
+\uncited\uncited to tame combinatorial explosion via user intervention.
 
-\meeraheader{v}{Configuration Spaces of Under-constrained Systems}
+\medskip\noindent
+\header{(v) Configuration Spaces of Under-constrained Systems}
 For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
 from various subclasses of tree-decomposable systems, algorithms have
-been developed \cite{XX, XX} to complete them into isostatic systems
-\cite{XX, XX} and to find paths within the connected components
-\cite{XX,XX} of standard Cartesian configuration spaces. Most of the
+been developed \uncited\uncited to complete them into isostatic systems
+\uncited\uncited and to find paths within the connected components
+\uncited\uncited of standard Cartesian configuration spaces. Most of the
 algorithms with formal guarantees leverage Cayley configuration space
-theory \cite{XX,XX} to characterize subclasses of graphs and
+theory \uncited\uncited to characterize subclasses of graphs and
 additional constraints that control combinatorial explosion, and
@@
 for general 1-dof tree-decomposable graphs. While software suites
-exist \cite{XX,XX}, no such formal algorithms and guarantees are known
+exist \uncited\uncited, no such formal algorithms and guarantees are known
 for non-tree-decomposable systems.
@@
   \item We navigate the NP-hardness barrier mentioned in (2) above,
-  for finding optimal DR-plans by defining a so-called \textit{canonical}
-  DR-plan and showing a strong Church-Rosser property: \textit{all
+  for finding optimal DR-plans by defining a so-called \dfn{canonical}
+  DR-plan and showing a strong Church-Rosser property: \vemph{all
   canonical DR-plans for isostatic or under-constrained 2D qusecs are
@@
   formal connection to well known problems such as optimal completion
-  of under-constrained systems \cite{XX}. When the required
+  of under-constrained systems \uncited. When the required
   modifications are bounded, we obtain new, efficient algorithms for
   realizing both isostatic and under-constrained qusecs by leveraging
-  recent results about Cayley parameters in \cite{XX,XX} (Section
+  recent results about Cayley parameters in \uncited\uncited (Section
   \ref{sec:recomb}).
@@
   stresses in self-similar, bar-joint systems); Example 3 (canonical
-  and optimal DR-plans for pinned line incidence systems \cite{XX})
+  and optimal DR-plans for pinned line incidence systems \uncited)
   and Examples 4,5 (boundary-conditions for achieving various desired

### 2014-12-18 doc/GMP2015/s03_drp.tex

--- 2014-12-17 doc/GMP2015/s03_drp.tex
@@
 
-Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \emph{assume} is well-constrained.
+Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained.
 If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
@@
 
-However, to guarantee that any two are \emph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant.
+However, to guarantee that any two are \vemph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant.
 Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
@@
 \begin{remark}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that:
-\begin{enumerate}
-    \item $F_i\cup F_j$ is not trivial.
-    \item $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
-    \item $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
-    \item $F_i\cap F_j$ is not under-constrained.
-\end{enumerate}
+If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
 \end{remark}
@@
 \begin{proof}
-Note that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
-\begin{enumerate}
-    \item By definition of trivial graph; if $F_i\cup F_j$ were trivial, then $F_i$ and $F_j$ must be trivial.
+For (1), simply note that if $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
 
-    \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
+For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$.
+For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained.
+Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$.
+Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
+
+For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
+% \begin{enumerate}
+%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
+
+%     \item Observe that  Then it is straightforward application of the equation above.
+
+%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
+
+%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.
+
+%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
 Therefore $d(F_i\cap F_j)=2k-l>k$.
@@
 
-    \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
+%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
 Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
@@
 
-    \item Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained. So it follows from the above that an under-constrained intersection is not possible.
-\end{enumerate}
+%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
+% \end{enumerate}
 \end{proof}
@@
 
-We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.
+\todo{Explain just second figure now.}
+% We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.
 
-\begin{figure*}\centering
-\begin{subfigure}{.3\linewidth}\centering
-    \input{img/tikz/two_doublets-0.tikz}
-    \caption{}\label{fig:doublets:a}
-\end{subfigure}%
-\begin{subfigure}{.7\linewidth}\centering
-    \input{img/tikz/two_doublets-1.tikz}
-    \caption{}\label{fig:doublets:b}
-\end{subfigure}
+% \begin{figure*}\centering
+% \begin{subfigure}{.3\linewidth}\centering
+%     \input{img/tikz/two_doublets-0.tikz}
+%     \caption{}\label{fig:doublets:a}
+% \end{subfigure}%
+% \begin{subfigure}{.7\linewidth}\centering
+%     \input{img/tikz/two_doublets-1.tikz}
+%     \caption{}\label{fig:doublets:b}
+% \end{subfigure}
 
-\caption{(\ref{fig:doublets:a}) Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
+% \caption{(\ref{fig:doublets:a}) Two doublets ($C_2\times C_3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
 (\ref{fig:doublets:b}) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
-\label{fig:doublets}
-\end{figure*}
+% \label{fig:doublets}
+% \end{figure*}
 
-\begin{example}
-    In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
+% \begin{example}
+%     In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
 These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$.
-This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 2.
+This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 1.
 Also shown is the decomposition of the individual doublets.
 These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them.
-As explained in Theorem \ref{theorem:main}, point 1, all of them must be included in the optimal DR-plan.
+As explained in Theorem \ref{theorem:main}, point 2, all of them must be included in the optimal DR-plan.
 Were any left out, the union of the children would not be the entire parent.
-\end{example}
+% \end{example}
 
@@
 
-\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K3,3$.
+\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K_{3,3}$.
 This pattern continues for a total of $N$ triangles.
-(\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K3,3$ subgraphs into the separate 9 edges and of edges into the component nodes.
-$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K3,3$ subgraphs.
+(\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
 The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
@@
 \subsection{Algorithm}
+
+\todo{Update, explain pebble game and frontier, complexity}
+
 The algorithm to find the optimal DR-plan is a straightforward application of the theory.

### 2014-12-18 doc/GMP2015/s08_pinnedline.tex

@@
 The optimal qusecs DR-plan can also be applied to analyze and solve
-the structure of  cross-linking microfibrils in collagen and cellulose in biology,
-by modeling these structures as {\em pinned line-incidence systems}.
+the structure of  cross-linking microfibrils in collagen and cellulose
+in biology, by modeling these structures as \dfn{pinned line-incidence
+systems}.
 
-{\em Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
-{\em Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
-Both of these substances consist of a large number {\em microfibrils},
+\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
+\dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
+Both of these substances consist of a large number \dfn{microfibrils},
 %fibril itself straight?
@@
 each of which is cross-linked at 2 places with some two other fibrils,
-where the {\em  cross-linking} is like an incidence constraint that
-the two crosslinked fibrils can slide with each other  \cite{xxx} (see Figure~\ref{fig:crosslink}).
+where the \dfn{cross-linking} is like an incidence constraint that the
+two crosslinked fibrils can slide with each other  \uncited (see
+Figure~\ref{fig:crosslink}).
 
@@
 \end{subfigure}%
-\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface~\cite{xxx}. (b) Cross-linking of cellulose microfibrils~\cite{xxx}.}
+\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \uncited. (b) Cross-linking of cellulose microfibrils \uncited.}
 \end{figure}
@@
 
-The cross-linking  microfibrils can be modeled as a pinned line-incidence constraint system in $\mathbb{R}^2$,
-where incidence constraints are used instead of distance constraints.
+The cross-linking microfibrils can be modeled as a pinned line-
+incidence constraint system in $\mathbb{R}^2$, where incidence
+constraints are used instead of distance constraints.
 
-A {\em pinned line-incidence system} $(G,\delta)$  
-is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$, 
-such that each edge is constrained to lie on a line passing through 
-the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
+A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
+together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
+fixed positions in $\mathbb{R}^2$, such that each edge is constrained
+to lie on a line passing through the corresponding pin, i.e.\ $\delta:
+E \rightarrow \mathbb{R}^2$.
 %
@@
 %
-A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$  \cite{xxx}.
+A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \uncited.
 Note that no trivial motion exists since the pins have fixed positions on the plane.
@@
 First, we note that %unlike bar-joint graphs,
-{\em a well-constrained pinned line-incidence graph can be disconnected},
+\vemph{a well-constrained pinned line-incidence graph can be disconnected},
 being the disjoint union of two or more well-constrained subgraphs.
@@
 %
-We define a {\em trivial} graph to be a single vertex and make the following modification to the definition of the DR-plan:
+We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of the DR-plan:
 
@@
 \begin{definition}
-
- The {\em DR-plan} of  a well-constrained pinned line-incidence graph $G$ is one in which 
-(1) each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$, and
+The \dfn{DR-plan} of a well-constrained pinned line-incidence graph $G$ is one in which
+(1) each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a \dfn{connected} well-constrained vertex-maximal proper subgraph of $C$, and
 (2) a leaf node is a single edge.
@@
 
-With the above modification,
-we now adapt Theorem~\ref{theorem:main} 
+With the above modification, we now adapt Theorem~\ref{theorem:main}
 for the optimal DR-plan for pinned line-incidence graphs.
 
-\begin{corollary}
+\begin{corollary}\label{cor:pinned}
 Given a well-constrained pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
@@
 \end{enumerate}
-\label{cor:pinned}
 \end{corollary}
 
-The following lemma serves a similar function to Lemma~\ref{lemma:union_intersection} in Section~\ref{sec:union_intersection}
+The following lemma serves a similar function to Remark~\ref{lemma:union_intersection} in Section~\ref{sec:union_intersection}
 %Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
@@
 
-\begin{lemma}
-\label{lem:pinned_union_intersection}
+\begin{remark}\label{lem:pinned_union_intersection}
 Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$,
@@
 There are only two possible cases:
-\begin{itemize}
-\item At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial.
-\item Both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
-\end{itemize}
+(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
 %The following table summarizes all bijections:
@@
 %\setlength{\tabcolsep}{6pt}
-\end{lemma}
+\end{remark}
 
-Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
+Given Remark~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
 proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
-The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs 
-is given in Appendix~\ref{sec:appendix_pinned}.
+The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned
+line-incidence graphs is given in Appendix~\ref{sec:appendix_pinned}.
 %The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.
@@
 
-\noindent {\bf Note:} The recombination problem for  pinned line-incidence systems is trivial.
+\noindent
+\note The recombination problem for  pinned line-incidence systems is trivial.
 Since the pins are given fixed positions in the plane,

### 2014-12-18 doc/GMP2015/s09_problems.tex

@@
 In this section we provide a unified view of  the various problems
-studied in the previous 2 sections, along with formal reductions between
-them,  and relationships to other known problems and results as well as open
-questions.
+studied in the previous 2 sections, along with formal reductions
+between them, and relationships to other known problems and results as
+well as open questions.
 %
@@
 
-The {\em restricted optimal DR-planning problem} requires
-DR-plans of one of these types, which reduces to recognizing if
-the input graph is a 2-tree or a tree-decomposable graph for which simple
-near-linear time algorithms are available \cite{XX} and the DR-plan is a
+The \dfn{restricted optimal DR-planning problem} requires DR-plans of
+one of these types, which reduces to recognizing if the input graph is
+a 2-tree or a tree-decomposable graph for which simple near-linear
+time algorithms are available \uncited and the DR-plan is a
 by-product output of the recognition algorithm.
 
-In the recombination setting, the corresponding {\em restricted
-OMD$_k$  problem}
-requires the reduced graph $G'$  and its isostatic completion $H$ to be
-2-trees as in Section \ref{2-tree-completion}
+In the recombination setting, the corresponding \dfn{restricted
+OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
+completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
 or to be a low Cayley complexity tree-decomposable graph as in Section
-\ref{tdecomp}. Clearly these problems have deterministic polynomial time
-algorithms, the algorithms run in time exponential in $k$.
-As a result, when $k$ is part of the input, the {\em restricted OMD problem}
-has the potential to be difficult.
-For example,  when the isostatic completion $H$ is required to be a 2-tree
-the restricted OMD problem is reducible to the maximum spanning series-parallel
-subgraph problem shown by \cite{XX,XX} to be NP-complete
-even if the input graph is planar of maximum degree at most 6.
-However, since the OMD problem has other input restrictions
-such as not having any proper isostatic subgraphs,
-it is not clear if the reverse reduction exists and hence it is unclear
-whether the OMD problem is NP-complete.
-The same holds for the restricted OMD problem where the isostatic completion
-$H$ is required to be a tree-decomposable graph of low Cayley complexity (i.e, have
-special, small DR-plans).
-One potential obstacle to an indecomposable graph $G$'s membership in the
-restricted OMD$_k$ for small $k$ is
-if $G$ is triconnected and has large girth. In fact,  6-connected (hence
-rigid) graphs with
-arbitrarily large girth have been constructed in \cite{XX}.
+\ref{sec:tdecomp}. Clearly these problems have deterministic
+polynomial time algorithms, the algorithms run in time exponential in
+$k$. As a result, when $k$ is part of the input, the \dfn{restricted
+OMD problem} has the potential to be difficult. For example, when the
+isostatic completion $H$ is required to be a 2-tree the restricted OMD
+problem is reducible to the maximum spanning series-parallel subgraph
+problem shown by \uncited\uncited to be NP-complete even if the input
+graph is planar of maximum degree at most 6. However, since the OMD
+problem has other input restrictions such as not having any proper
+isostatic subgraphs, it is not clear if the reverse reduction exists
+and hence it is unclear whether the OMD problem is NP-complete. The
+same holds for the restricted OMD problem where the isostatic
+completion $H$ is required to be a tree-decomposable graph of low
+Cayley complexity (i.e, have special, small DR-plans). One potential
+obstacle to an indecomposable graph $G$'s membership in the restricted
+OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
+have been constructed in \uncited.
 %
@@
 The OMD problem is closely related to a well-studied problem of
-completion of an underconstrained system to an isostatic one with a small
-DR-plan.
+completion of an under-constrained system to an isostatic one with a
+small DR-plan.
 \begin{observation}
-The (decision version of) the {\bf optimal completion problem (OC)} from \cite{bib:survey, j-a,
-fang} is OMD$_0$.
+    The (decision version of) the {\bf optimal completion problem
+    (OC)} from \cite{bib:survey, j-a, fang} is OMD$_0$.
 \end{observation}.
-In fact, a {\em restricted OC} problem was studied by \cite{j-a} requiring the
-completion to be tree-decomposable.
-It is not clear if the reverse reduction holds, i.e, is OMD reducible to OC?
+In fact, a \dfn{restricted OC} problem was studied by \cite{j-a}
+requiring the completion to be tree-decomposable. It is not clear if
+the reverse reduction holds, i.e, is OMD reducible to OC?
 
-We now connect the OMD problem to the informal {\em optimal recombination
-(OR)} problem
-mentioned as motivation at the beginning of
-Section \ref{sec:recomb}.
+We now connect the OMD problem to the informal \dfn{optimal
+recombination (OR)} problem mentioned as motivation at the beginning
+of Section \ref{sec:recomb}.
 
-In order to connect the OR problem to OMD,
-when the input graph for is the isostatic graph at the DR-plan root, we do not consider the case
-where the two child {\em solved subgraphs}
-(corresponding to already solved subsystems)
-have a nontrivial intersection (in this case
-the recombination is trivial). We only consider the case where no two child
-solved subgraphs (resp. two root subgraphs when the input graph is
-underconstrained) share more than 1 vertex.
-We replace such solved subgraphs  by isostatic graphs as
-follows. If a solved subgraph shares at most one vertex with the remainder of
-the graph, simply replace it by an edge one of whose endpoints  is
-the shared vertex. Otherwise, replace it by  a 2-tree graph of the shared vertices.
-Finally, we add the additional
-restriction to the OM Problem that
-when any edge in a solved subgraph is chosen among the $k$ edges to be
-removed, in fact the entire solved subgraph must be removed  and all of its
-edges must be counted in $k$.
+In order to connect the OR problem to OMD, when the input graph for is
+the isostatic graph at the DR-plan root, we do not consider the case
+where the two child \dfn{solved subgraphs} (corresponding to already
+solved subsystems) have a nontrivial intersection (in this case the
+recombination is trivial). We only consider the case where no two
+child solved subgraphs (resp. two root subgraphs when the input graph
+is under-constrained) share more than 1 vertex. We replace such solved
+subgraphs  by isostatic graphs as follows. If a solved subgraph shares
+at most one vertex with the remainder of the graph, simply replace it
+by an edge one of whose endpoints  is the shared vertex. Otherwise,
+replace it by  a 2-tree graph of the shared vertices. Finally, we add
+the additional restriction to the OM Problem that when any edge in a
+solved subgraph is chosen among the $k$ edges to be removed, in fact
+the entire solved subgraph must be removed  and all of its edges must
+be counted in $k$.
 
-This reduction is used also for transferrring  algorithms for
-optimal DR-planning, recombination, completion, OMD, and other problems and
-to so-called {\em
-multi-body pin} from so-called{\em bar-joint-systems}, by showing that the corresponding
-prblems for the latter are reduced to the correspnding problems on
-bar-joint systems.
+This reduction is used also for transferring  algorithms for optimal
+DR-planning, recombination, completion, OMD, and other problems and to
+so-called \dfn{multi-body pin} from so-called \dfn{bar-joint-systems},
+by showing that the corresponding problems for the latter are reduced
+to the corresponding problems on bar-joint systems.
 
-{\bf Note.} This section deals only with bar-joint systems, however,
-extensions of several of these problems, approaches and solutions
-to other qusecs beyond  bar-joint - specifically pinned line incidence
-systems and
-multi-body-pin systems
-will be expanded in Sections \ref{sec:pinnedline} and \ref{sec:bodypin}.
+\note This section deals only with bar-joint systems, however,
+extensions of several of these problems, approaches and solutions to
+other qusecs beyond  bar-joint - specifically pinned line incidence
+systems and multi-body-pin systems will be expanded in Sections
+\ref{sec:pinnedline} and \ref{sec:bodypin}.

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Further proof reading. Fixed \emph problem!

--- 2014-12-18 doc/GMP2015/s07_problempinning.tex
@@
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
+For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively.
 \end{definition}
@@
 
-We can develop the notion of a canonical DR-plan for certain $(k,l)$ tightness conditions similar to the way we did in Section \ref{sec:DRP}.
+We can develop the notion of a canonical DR-plan for certain $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}.
 
@@
     All $(k,l)$-tight canonical DR-plans are optimal.
-We can find such a DR-plan in the same time complexity as the $(2,3)$ tight case for bar and joint graphs discussed in Section \ref{sec:DRP}.
+We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar and joint graphs discussed in Section \ref{sec:DRP}.
 This is due tot he fact that such $(k,l)$-tight graphs are matroidal and have a pebble game.
@@
     For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
-
+    % Don't leave a blank line between last paragraph and end!
     % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.

### 2014-12-18 doc/GMP2015/s03_drp.tex
Further proof reading. Fixed \emph problem!
Expanded on the DR-planning example and algorithm. Minor change to recomb notation.

@@
 \end{enumerate}
-
+%
 This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
@@
 For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
+\end{proof}
+
+% \begin{proof}
+% Use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
 % \begin{enumerate}
@@
 % \end{enumerate}
-\end{proof}
+% \end{proof}
 
@@
 \begin{proof}
-%I wouldn't do the forward and backward direction italics, or say proof by contradiction.
-Just start the first paragraph with Suppose C_i union C_j is well constained maximal proper and assume it is not C.
 Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
-
+%
 In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
@@
 
-\subsection{Examples}
+\subsection{Example}
 
@@
 
-\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K_{3,3}$.
-This pattern continues for a total of $N$ triangles.
-(\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+\caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
+This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
+(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
 $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
+The dashed lines indicated that this exact structure is repeated.}
 \label{fig:c2c3ofk33s}
 \end{figure*}
+
+This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
+The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$.
+These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained.
+As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must occur 2 levels below the parent.
+Just as expected, it is on the third level, as it is a child of both of $G$'s children.
+Furthermore, it intersects trivially with the edges that connect triangle $1$ and $N$ to $2$ and $N-1$, respectively; these necessarily intersect with triangle $1$ and $N$ themselves trivially (by Theorem \ref{theorem:main}, part 1) making all of these children.
+The triangles then decompose into their constituent $K_{3,3}$ graphs, which then decompose into 9 edges.
+
+Now, the self-similar nature of the graph is clear; $G$ without triangle $1$ and $N$ has the same structure, and thus the same pattern is repeated until only triangle $(N+1)/2$ remains in the case that $N$ is odd, or until only triangles $N/2$ and $N/2+1$ remain if $N$ is even.
 
@@
 
-\todo{Update, explain pebble game and frontier, complexity}
-
-The algorithm to find the optimal DR-plan is a straightforward application of the theory.
-It begins by finding the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
+The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
+There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \uncited to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \uncited on each node and taking the largest components that are not subgraphs of any larger ones.
+After the subgraphs have been found, it is a straightforward application of the theory.
 If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
@@
 This naturally terminates when you reach trivial subgraphs.
+The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
+\todo{Think more about complexity.}

### 2014-12-18 doc/GMP2015/s04_recombination.tex
Further proof reading. Fixed \emph problem!
Expanded on the DR-planning example and algorithm. Minor change to recomb notation.

@@
 proving the theorem.
-\end{proof}.
+\end{proof}
 
@@
 \begin{proof}
-    The Cartesian realization space of $(H,<\delta_E', \lambda_F>)$ is
-    computed easily with a DR-plan of size 2, and is the union of
-    $2^t$ solutions (modulo orientation preserving isometries) each
-    with a distinct orientation type, where $t$ is the number of
-    triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction
-    of the length vector $\delta$ to the edges in $S$. A desired
-    solution $p$ (or connected component of a solution space) of
-    $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a
-    subdivided binary search   of the cartesian realization space of
-    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ - as
-    $\lambda_F$ ranges over the discretized convex polytope
-    $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
-    Theorem \ref{theorem:convexcayley}. A solution $p$  is found  when
-    the lengths for nonedges in $D$ match $\delta_D$.
+    The Cartesian realization space of $(H,\left<\delta_{E'},
+    \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
+    and is the union of $2^t$ solutions (modulo orientation preserving
+    isometries) each with a distinct orientation type, where $t$ is
+    the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
+    the restriction of the length vector $\delta$ to the edges in $S$.
+    A desired solution $p$ (or connected component of a solution
+    space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
+    found by a subdivided binary search   of the cartesian realization
+    space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
+    type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
+    convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
+    described in Theorem \ref{theorem:convexcayley}. A solution $p$
+    is found  when the lengths for nonedges in $D$ match $\delta_D$.
 \end{proof}
@@
 coincidences in the realizations of $(G,\delta)$, the realization
-space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16 solutions
-$q^p_{\lambda_f}$ (modulo orientation preserving isometries), with
-distinct orientation types $\sigma_p$ (two orientation choices for
-each of the 4 triangles) that can be obtained by solving a sequence of
-4 single quadratics in 1 variable (DR-plan of size 2). By subdivided
-binary search in the interval $\lambda_f \in \Phi_f(G',
-\delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is
-found when the length of the nonedge $e$ in  the realization
-$q^p_{\lambda_f}$ is $\delta_e$.
+space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
+solutions $q^p_{\lambda_f}$ (modulo orientation preserving
+isometries), with distinct orientation types $\sigma_p$ (two
+orientation choices for each of the 4 triangles) that can be obtained
+by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
+size 2). By subdivided binary search in the interval $\lambda_f \in
+\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
+$(G,\delta)$ is found when the length of the nonedge $e$ in  the
+realization $q^p_{\lambda_f}$ is $\delta_e$.
 %

### 2014-12-18 doc/GMP2015/s09_conclusion.tex
Started a conclusion

\section{Conclusion}
We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem. To this end we have defined a new canonical DR-plan and given an effiecient ($O(n^2)$) algorithm to find the DR-plan for specific types of 2D qusecs.

We have also described methods of recombining the DR-plan to get a realization different from the normal solving of a system of equations. These results rely on the results obtained in the Cayley Configuration space of certain graphs.

We then related these theoretical results to real-world situations in which they can be applied including Examples 1-5 in Section \ref{sec:intro}.

Throughout the paper, we raised some open questions and future areas of research. These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}

### 2014-12-18 doc/GMP2015/s00_abstract.tex
Most of my job for citations.

@@
 \begin{abstract}
-A variety of quasi-uniform (aperiodic) and possibly self-similar, layered material structures are a natural consequence of the design objectives of rigidity and flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
+A variety of quasi-uniform (aperiodic) and possibly self-similar,
+layered material structures are a natural consequence of design
+objectives such as rigidity or flexibility, minimizing mass, optimally
+distributing external stresses and participating in the assembly of
+diverse and multifunctional, larger  structures.
 
-Crucial for the design and analysis of the abovementioned properties is the recursive decomposition of diverse types of underlying 2D geometric constraint systems that are isostatic or underconstrained (which we call collectively as {\it qusecs}).
-While the optimal recursive decomposition ({\it DR-planning}) problem is NP-hard even for 2D bar-joint constraint systems, we describe an $O(n^2)$ algorithm  for qusecs.
-Besides elucidating the source of NP-hardness, the algorithm outputs {\it canonical} DR-plans that meet various desirable, previously studied criteria.
-By leveraging recent results on Cayley configuration spaces, we show how to decompose the indecomposable algebraic systems that recombine the solutions to child subsystems at a node of an optimal DR-plan: we minimally modify the system to have a small DR-plan, leading to efficient realization algorithms.
-We show formal connections to well-known problems such as completion of underconstrained systems.
-We briefly discuss the modeling and analysis of two specific 2D layered materials as applications of the above algorithms, whose software implementation is available.
+Crucial for the design and analysis of the abovementioned properties
+is the recursive decomposition of diverse types of underlying 2D
+geometric constraint systems that are \vemph{isostatic or
+underconstrained (which we call collectively as \dfn{qusecs})}. While
+the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
+hard even for (possibly overconstrained) 2D bar-joint constraint
+systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
+elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
+plans that meet various desirable, previously studied criteria. We
+leverage recent results on Cayley configuration spaces to show that
+the indecomposable systems -  that are solved at node of the optimal
+DR-plan by recombining solutions to child systems - can be minimally
+modified to become decomposable and have a small DR-plan, leading to
+efficient realization algorithms. We show formal connections to well-
+known problems such as completion of underconstrained systems. We
+briefly discuss the modeling and analysis of two specific 2D layered
+materials as applications of the above theory and algorithms. We
+intend to make a software implementation and videos available upon
+request and publicly available for the final version.
 

### 2014-12-18 doc/GMP2015/s01_introduction.tex
Most of my job for citations.

@@
 Some properties of natural and engineered materials can be analyzed by
-treating them as stacked two dimensional (2D) layers. The structure
-within each layer could be self-similar \cite{Intro1} spanning
-multiple scales, generally aperiodic and quasi-uniform within any one
-scale, and consisting of a few repeated motifs appearing in disordered
-arrangements. Each layer is not necessarily planar, i.e., it consists
-of multiple, inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either \vemph{isostatic or
-under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
-definitions). These properties are consistent with a self-assembled 2D
-structure that minimizes mass, optimally distributes external stresses
-and itself participates in the assembly of diverse and
-multifunctional, larger 2D structures.
+treating them as two dimensional (2D) layers. As illustrated by the
+examples below, the structure within each layer is often self-
+similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally
+aperiodic and quasi-uniform within any one scale; and consists of a
+few repeated motifs appearing in disordered arrangements. Each layer
+is not necessarily planar, i.e., it consists of multiple, inter-
+constraining planar (genus 0) monolayers. Furthermore, a layer is
+often  either \vemph{isostatic or under-constrained, not self-
+stressed} (See Section \ref{sec:prelim} for definitions). These
+properties are consistent with a self-assembled 2D structure that
+minimizes mass, optimally distributes external stresses and itself
+participates in the assembly of diverse and multifunctional, larger 2D
+structures.
 
@@
 \begin{enumerate}
-    \item Cross-sections of microtubule structures (Figure \ref{fig:material_examples:microtubule}) \cite{Necklace1} e.g., in ciliary membranes and transitions \cite{Necklace2}
+    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}
 
@@
 
-    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{CellWalls1}~\cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
 
-    \item More recent, engineered examples, including disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.
+    \item More recent, engineered examples, including disordered
+    graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes
+    reinforced by microfibrils; and DNA assemblies including a recent
+    Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other
+    self-similar structures~\cite{Microfibrils2} within reach.
 
-    \item Silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
+    \item Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
+    materials that behave like assemblies of 2D particles under
+    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}
 \end{enumerate}
 %
-In order to study structural and mechanical properties such as
-rigidity, flexes and distribution of external stresses within the
-material layer, it is natural to model a material layer as a solution
-of a geometric constraint system of appropriate types of geometric
+In order to study structural and mechanical properties
+of a material layer, it is natural to model a material layer as a solution or
+realization of a geometric constraint system of appropriate types of geometric
 primitives, under metric or algebraic constraints. Such 2D
-\dfn{qusecs} (quasi-uniform or self-similar constraint systems) can be
+\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
+formally in Section \ref{sec:prelim}) can be
 used to understand or design material layers (their solutions) with
@@
 We now briefly survey existing techniques for studying 2D
-\dfn{qusecs},  many of which are \dfn{bar-joint} systems
-(Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
+qusecs,  many of which are \dfn{bar-joint} systems
+(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
 \dfn{body-hyperpin} systems (Example 4,5, see Section
@@
 \header{(i) Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
-locating all maximal, generically rigid subsystems \uncited. When the
-input is rigid, these algorithms compute the identity and output the
-input.
+Fast, graph-based algorithms exist (pebble-
+game~\cite{Jacobs:1997:PG}), for locating all maximal,
+\dfn{generically rigid} subsystems (defined formally in Section
+\ref{sec:prelim}) \uncited. When the input itself is rigid, these
+algorithms do nothing, i.e., compute the identity function.
 
@@
 subsystems, down to the level of geometric primitives, in order to
-understand or design properties at all scales, such as: rigidity,
-flexes, distribution of external stresses, boundary conditions for
-isostaticity, as well as behavior under constraint variations.
+understand or design properties at all scales, such as (formally
+defined in Section \ref{sec:prelim}) \dfn{rigidity}, \dfn{flexes},
+distribution of \dfn{external stresses}, boundary conditions for
+\dfn{isostaticity}, as well as behavior under constraint variations.
 
 \medskip\noindent
-\header{(ii) Optimal Recursive Decomposition (DR-planning)}
+\header{(ii) Optimal Recursive Decomposition (DR-planning).}
 Recursive decomposition of geometric constraint systems has been
 formalized \uncited and well-studied \uncited as the \dfn
-{Decomposition-Recombination (DR)-planning} problem. For the above-
+{Decomposition-Recombination (DR)-planning} problem, defined formally in
+Section \ref{sec:prelim}. For the above-
 mentioned classes of 2D qusecs, generic rigidity is a combinatorial
@@
 given in \uncited. An \dfn{optimal DR-plan} is one that minimizes
-the maximum number of child subsystems of any parent system, a
-property that is crucial for obtaining a solution to the parent
-constraint system from the solutions of the child systems.
+the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e.,
+the maximum number of child subsystems of any parent system.
+Being exponential the size, the complexity of solving the parent
+constraint system from the solutions of the child systems, is overwhelmingly
+dominated by it.
 
-However, for general 2D qusecs, even when restricted to bar-joint
-systems, the optimal DR-planning problem was shown to be NP-hard
-\uncited.
+However, for general 2D qusecs that could be overconstrained, even
+when restricted to bar-joint systems, the optimal DR-planning problem
+was shown to be NP-hard \uncited.
 
@@
 \header{(iii) DR-plans for special classes and with other
-criteria}
+criteria.}
 For a special class of 2D qusecs, namely \dfn{tree-decomposable}
@@
 requiring child subsystems to be maximal rigid proper subsystems of
-the parent system.
+the parent system. See Section \ref{sec:prelim} for formal definitions.
 
@@
 \medskip\noindent
-\header{(iv) Optimal Recombination and Solution Space Navigation}
+\header{(iv) Optimal Recombination and Solution Space Navigation.}
 For the entire DR-plan, finding all desired solutions is barely
@@
 parameterization to minimize the algebraic complexity of the
-indecomposable parent (sub)systems that occur in the DR-plan \uncited\uncited\uncited, provided the DR-plan meets some of the abovementioned
-criteria.
+indecomposable parent (sub)systems that occur in the DR-plan
+\uncited\uncited\uncited, provided the DR-plan meets some of the
+abovementioned criteria.
 
@@
 \medskip\noindent
-\header{(v) Configuration Spaces of Under-constrained Systems}
+\header{(v) Configuration Spaces of Under-constrained Systems.}
 For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
@@
   above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
-  \ref{sec:pinnedline}). The canonical DR-plan illuminates the essence
+  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
   of the NP-hardness of finding optimal DR-plans for over-constrained
@@
 
-  \item Software implementation and videos can be found at:
+  \item We intend to make software implementation and videos available
+  upon request and publicly available for the final version of the
+  paper.
   \todo{Provide Link}.

### 2014-12-18 doc/GMP2015/s03_drp.tex
Most of my job for citations.

@@
 The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
-There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \uncited to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \uncited on each node and taking the largest components that are not subgraphs of any larger ones.
+There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
 After the subgraphs have been found, it is a straightforward application of the theory.

### 2014-12-18 doc/GMP2015/s02_preliminaries.tex
Meera's changes to preliminaries.

@@
 \label{sec:prelim}
-
+%
 \subsection{Basics}
-
+%
 In this paper, a \dfn{geometric constraint system} is a multivariate
@@
 $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
-
+%
 \noindent
 \note Geometric constraint systems can also have
-inequalities, where the parameters in $\delta$ are small intervals of
+inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
 values rather than exact values.
@@
 %which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, well-constrained)} if it is
+It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is
 both rigid and independent.
@@
 %The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-It is \dfn{flexible} if it is not rigid, \dfn{under-constrained} if it is independent and not rigid, or \dfn{over-constrained} if it is not independent.
+It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
+is independent and not rigid, or \dfn{overconstrained} if it is not
+independent.
 
-Defining the independence of a subset of edges $E' \subseteq E$ to be
-the independence of corresponding rows in the appropriate rigidity
-matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a
-constraint (hyper)graph $G = (V,E)$.
-%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
+Defining the combinatorial independence of a subset of edges $E'
+\subseteq E$ to be the generic independence of corresponding rows in
+the appropriate rigidity matrix of indeterminates, we obtain the
+\dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
 %where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
@@
 There are various results on combinatorial characterization of
-independence and rigidity for different types of  rigidity matroid.
-For bar-joint rigidity matroid, the famous Laman's theorem \uncited
-states that the underlying graph is minimally rigid if and only if
+independence and rigidity  and rigidity matroids for different types
+of graphs. For bar-joint rigidity matroid, the famous Laman's theorem
+\uncited states that the underlying graph is isostatic if and only if
 $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
@@
 body rigidity matroid, Tay \uncited proved that the underlying
-multigraph is minimally rigid if and only if it can be decomposed as
-$3$ edge disjoint spanning trees. White and Whiteley \uncited gave
-the same characterization using a different technique to study the
-algebraic - geomtric conditions called pure condition. Lee, Streinu
-and Theran \uncited defined the \dfn{$(k,l)$-sparsity matroid},
-where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le
-k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2
-vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V|
-- l$. In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity
-condition is matroidal as long as $l \le dk-1$.
+multigraph is isostatic if and only if it can be decomposed as $3$
+edge disjoint spanning trees. White and Whiteley \uncited gave the
+same characterization using a different technique to study the
+algebraic-geometric conditions of genericity, called pure condition.
+Lee, Streinu and Theran \uncited defined the \dfn{$(k,l)$-sparsity
+matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if
+$|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least
+2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| =
+k|V| - l$. In general, given a $d$-uniform hypergraph, a
+$(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
 
 In this paper, a \dfn{qusec} is any \dfn{independent} geometric
-constraint system of one of the 3 types mentioned above. In the
-remainder of this section and the next two sections~\ref{sec:DRP} and
-\ref{sec:recomb} we only consider bar-joint qusecs and graphs.
-Relevant formal analogies for the other 2 types of qusecs and
-(hyper)graphs are given in the subsequent 2 sections on Applications.
+constraint system of one of the 3 types mentioned above.
+\medskip\noindent
+\note In the remainder of this section and the next three
+sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
+qusecs and graphs. Relevant formal analogies for the other 2 types of
+qusecs and (hyper)graphs are given in the subsequent 2 sections on
+Applications.
 
@@
 matrix.
+
+The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set
+$X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
 
@@
 % \end{definition}
-
-A \dfn{trivial} graph is a single vertex.
 %
@@
     The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \uncited $DRP(G)$, is defined as a forest that has the
+    $G$, \uncited {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
     following properties:
@@
         set of $C$.
-        \item A leaf node is a trivial subgraph.
+        \item A leaf node is a single vertex.
+A \dfn{trivial} graph is empty or a single vertex.
         \item A root node is a maximal rigid subgraph of $G$.
     \end{enumerate}
-
+%
     % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
 
-    An equivalent DAG can be constructed from the above forest such
-    that all nodes containing the same subgraphs are identified into a
-    single node. In this manuscript, however, we retain a forest as
-    defined above. When a graph is isostatic, its DR-plan is a tree.
-
-    %Note that nodes will be referred to interchangeably as ``the node that
-    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
-
     A DR-plan is \dfn{complete} if it satisfies an additional rule:
-        \item For a node, $C$, that is  a non-trivial graph, its $N$
+        \item For a node $C$ that is  a non-trivial graph, its $N$
         children, $C_1, \ldots, C_N$, are all of the rigid vertex-
@@
         and in fact, $C$ contains all the edges in the union of its
-        children as well. The complete DR-plan is unique.
-    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+        children as well.
+    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
+    The maximum  fan-in is called the \dfn{size} of the DR-plan
+
+%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
+%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
+%an isostatic vertex-maximal proper subgraph.
+    %Note that nodes will be referred to interchangeably as ``the node that
+    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
 \end{definition}
+%
+% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
+\begin{remark}
+    For a given graph, there could be exponentially many  DR-plans and
+    even optimal DR-plans in the size of the graph . A complete DR-
+    plan is unique but may not be (is usually not) optimal. DR-plans
+    of self-similar graphs are self-similar. More than one node (leaf)
+    in a DR-plan forest may represent the same subgraph (vertex) of
+    $G$.
+\end{remark}
 
@@
 \note In Section \ref{sec:DRP} $G$ is assumed to be isostatic.
-$Idc(G,X)$ is the subgraph of $G$ induced by the vertex set
-$X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
-In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
-$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
-an isostatic vertex-maximal proper subgraph.
 
@@
 The use of a DR-plan in designing especially self-similar qusecs is
-immediately clear, as shown in Figures \ref{XX} and \ref{XX}
-\todo{Reference?}. In fact, knowing an optimal DR-plan is crucial for
+immediately evident, as shown in Figures \ref{fig:c2c3ofk33s} and
+\ref{fig:subdivision}.
+
+In fact, knowing an optimal DR-plan is crucial for
 design and analysis of even non-self-similar qusecs and their
@@
 
-First, note that the basic task of \header{realizing given qusecs}
+\header{Realizing given qusecs.}
+Note that this basic task
 would require finding real solutions a large multivariate polynomial
@@
 
-Next, using the DR-plan we obtain a \header{recursive block
-decomposition of the rigidity matrix} with each block being the
+\header{Recursive block
+decomposition of the rigidity matrix, stress space and external stresses.}
+This can be obtained using the DR-plan we obtain a
+with each block being the
 rigidity matrix of an isostatic subgraph. The blocks intersect on
 isostatic or trival subgraphs. This recursive block decomposition of
-the rigidity matrix $R$ correspondingly \header{recursively decomposes
-the stress vectors} $s$ that balance a given external stress $t$ since
-$sR = t$, and moreover, the external stress vector $t$ now
-\header{recursively ``distributes'' as external stresses} for each
-isostatic subgraph of the DR-plan. When the rigidity matrix and stress
-vectors are indeterminates, it can be used to design  qusecs
+the rigidity matrix $R$ correspondingly recursively decomposes
+the stress vectors $s$ that balance a given external stress $t$ since
+$sR = t$. Moreover, the external stress vector $t$ is
+recursively ``distributed'' into external stresses for each
+node of the DR-plan.
+When the rigidity matrix and stress
+vectors contain indeterminates, this  can be used to design
 underlying graph and parameters $\delta$ to obtain an optimal
 distribution of stresses on the geometric primitives. When the
-realization of the qusec is given, this can be used to analyze the
+realization of qusecs is given, this can be used to analyze the
 distribution of a given external load.

### 2014-12-18 doc/GMP2015/s01_introduction.tex
White space fixes.

@@
 \label{sec:intro}
-
-% \newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}
 
@@
 structures.
+
+\noindent
+\textbf{Note on Scope:} In this manuscript we only study finite 2D structures.
+\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy
+or subdivision in an iterated scheme to generate self-similar structures.
 
@@
 desired properties.
-
+%
 \subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
@@
 systems gets bigger, with highest efficiency for underlying partial
-2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
+2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$
 minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
@@
 for non-tree-decomposable systems.
-
+%
 \subsection{Contributions and Organization}

### 2014-12-18 doc/GMP2015/s02_preliminaries.tex
White space fixes.

@@
 $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
-%
+
 \noindent
@@
 constraint system of one of the 3 types mentioned above.
+
 \medskip\noindent
@@
 \subsection{Decomposition-Recombination (DR) Plans}
-
+%
 %$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
@@
 The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
-
+%
 %Given a constant $k$, a graph $G$ is:
@@
 % \end{definition}
-
+%
 % \begin{definition}
@@
 % \end{definition}
-
+%
+%
+%
 % \begin{definition}
@@
     % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
-
+%
     A DR-plan is \dfn{complete} if it satisfies an additional rule:
@@
     A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
-    The maximum  fan-in is called the \dfn{size} of the DR-plan
-
+    The maximum  fan-in is called the \dfn{size} of the DR-plan.
+%
 %In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Corrected a figure
White space fixes.
Adding some citations
Added citations

@@
 \label{sec:bodypin}
+% \bibliography{paper}
 
-We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \uncited.
+We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}.
 Specifically, we will look into the problem when you have a mono-layer and it is under-constrained.
@@
 Not shown here are other layers stacked on this one.
-Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.}
+Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.
+Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
     \label{fig:silica_glass}
@@
 In (a), a simple subdivision is done that does not guarantee isostatic.
-In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.}
+In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.
+Credit to \cite{subdivision_paper} for (a).}
     \label{fig:subdivision}
@@
 %
-\begin{subfigure}{0.3\linewidth}\centering\scriptsize
+\begin{subfigure}{0.35\linewidth}\centering\scriptsize
     \input{img/tikz/bodypin_drp.tikz}

### 2014-12-18 doc/GMP2015/s01_introduction.tex

@@
 \header{(i) Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-
-game~\cite{Jacobs:1997:PG}), for locating all maximal,
+Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal,
 \dfn{generically rigid} subsystems (defined formally in Section
-\ref{sec:prelim}) \uncited. When the input itself is rigid, these
+\ref{sec:prelim}). When the input itself is rigid, these
 algorithms do nothing, i.e., compute the identity function.
@@
 Recursive decomposition of geometric constraint systems has been
-formalized \uncited and well-studied \uncited as the \dfn
-{Decomposition-Recombination (DR)-planning} problem, defined formally in
+formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the 
+\dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in
 Section \ref{sec:prelim}. For the above-
@@
 defining desirable or optimal DR-plans and DR-planning algorithms were
-given in \uncited. An \dfn{optimal DR-plan} is one that minimizes
+given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes
 the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e.,
@@
 when restricted to bar-joint systems, the optimal DR-planning problem
-was shown to be NP-hard \uncited.
+was shown to be NP-hard \cite{lomonosov2004graph}.
 
@@
 For a special class of 2D qusecs, namely \dfn{tree-decomposable}
-systems \uncited common in computer aided mechanical design, (which
+systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which
 includes ruler-and-compass and Henneberg-I constructible systems), all
@@
 
-While polynomial time algorithms were given \uncited to generate DR-
+While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-
 plans meeting the cluster minimality criterion, no such algorithm is
@@
 problem of finding even a single solution to the input system at the
-root of the DR-plan is NP-hard \uncited and there is a combinatorial
-explosion of solutions \uncited. Typically, however, the desired
+root of the DR-plan is NP-hard  \cite{saxe1979embeddability} 
+and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired
 solution has a given orientation type, in which case, the crux of the
@@
 indecomposable parent (sub)systems that occur in the DR-plan
-\uncited\uncited\uncited, provided the DR-plan meets some of the
+\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the
 abovementioned criteria.
@@
 principled basis, high performance heuristics and software exists
-\uncited\uncited to tame combinatorial explosion via user intervention.
+\cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
 
@@
 from various subclasses of tree-decomposable systems, algorithms have
-been developed \uncited\uncited to complete them into isostatic systems
-\uncited\uncited and to find paths within the connected components
-\uncited\uncited of standard Cartesian configuration spaces. Most of the
+been developed to complete them into isostatic systems
+\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components
+\cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. 
+Most of the
 algorithms with formal guarantees leverage Cayley configuration space
-theory \uncited\uncited to characterize subclasses of graphs and
+theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and
 additional constraints that control combinatorial explosion, and
@@
 for general 1-dof tree-decomposable graphs. While software suites
-exist \uncited\uncited, no such formal algorithms and guarantees are known
+exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known
 for non-tree-decomposable systems.

### 2014-12-18 doc/GMP2015/s02_preliminaries.tex

@@
 dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth \uncited shows a constraint
+The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint
 (hyper)graph is rigid if it is generically \dfn{infinitesimally
@@
 of graphs. For bar-joint rigidity matroid, the famous Laman's theorem
-\uncited states that the underlying graph is isostatic if and only if
+\cite{laman1970graphs} states that the underlying graph is isostatic if and only if
 $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
-at least 2 vertices. The result by Lovasz and Yemini \uncited  shows
+at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows
 that all 6-vertex-connected graphs are rigid in the plane. For bar-
-body rigidity matroid, Tay \uncited proved that the underlying
+body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying
 multigraph is isostatic if and only if it can be decomposed as $3$
-edge disjoint spanning trees. White and Whiteley \uncited gave the
+edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the
 same characterization using a different technique to study the
 algebraic-geometric conditions of genericity, called pure condition.
-Lee, Streinu and Theran \uncited defined the \dfn{$(k,l)$-sparsity
+Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity
 matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if
@@
     The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \uncited {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
+    $G$, \cite{hoffman2001decompositionI} {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
     following properties:

### 2014-12-18 doc/GMP2015/s03_drp.tex

@@
 
-\begin{enumerate}
-\item
-We show that no subset of the children can union to form $C$, thereby requiring all of them be included in an optimal plan.
+\medskip\noindent
+\vemph{Part 1:} We show that no subset of the children can union to form $C$, thereby requiring all of them be included in an optimal plan.
 
@@
 
-%%%%%%%%%%%%%%%
-
-\item
-If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
+\medskip\noindent
+\vemph{Part 2:} If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
 This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
@@
 Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
-\end{enumerate}
-%
+
 This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
@@
 \end{lemma}
-
+%
 \begin{proof}
@@
 \end{proof}
-
+%
+%
+%
 \begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
@@
 \end{proof}
-
+%
+%
+%
 \begin{lemma}\label{lemma:uc_intersection_makes_all_uc}
@@
 
-\todo{Explain just second figure now.}
 % We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.

### 2014-12-18 doc/GMP2015/s04_recombination.tex

@@
 difficulty can persist even after optimal parametrization of the
-indecomposable system $I$ as in \uncited to minimize its algebraic
+indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic
 complexity.
@@
 problem, and using a key theorem of Convex Cayley configuration spaces
-\uncited. This theorem characterizes the class of graphs $H$ and non-
+\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-
 edges $F$ (Cayley parameters), such that the set of vectors
@@
 \begin{theorem}\label{theorem:convexcayley}
-    \uncited $H$ has a convexifiable Cayley configuration space  with
+    \cite{sitharam2010convex} $H$ has a convexifiable Cayley configuration space  with
     parameters $F$ if and only if for each $f\in F$  all the minimal
@@
 completely unaffected when $\delta$ are intervals rather than exact
-values~\cite{GaoSitharam}.
-%cite my paper with Heping!!!
+values~\cite{sitharam2010convex}.
 
@@
 called \dfn{tree-decomposable graphs of low Cayley complexity}
-\uncited include the partial 2-trees and many others that are not
+\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
 partial 2-trees. These too result in DR-plans of size 2 or 3, putting

### 2014-12-18 doc/GMP2015/s08_pinnedline.tex
White space fixes.

@@
 
-\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
-\dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
+\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property \cite{buehler2008nanomechanics}.
+\dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}.
 Both of these substances consist of a large number \dfn{microfibrils},
@@
 where the \dfn{cross-linking} is like an incidence constraint that the
-two crosslinked fibrils can slide with each other  \uncited (see
+two crosslinked fibrils can slide with each other (see
 Figure~\ref{fig:crosslink}).
@@
 \end{subfigure}%
-\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \uncited. (b) Cross-linking of cellulose microfibrils \uncited.}
+\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}. (b) Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.}
 \end{figure}
@@
 %
-A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \uncited.
+A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \cite{sitharam2014incidence}.
 Note that no trivial motion exists since the pins have fixed positions on the plane.
 In particular, both a single vertex and a single edge are under-constrained graphs.
-%and the smallest over-constrained graph has $6$ vertices.
 

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Added citations

@@
 \label{sec:bodypin}
-% \bibliography{paper}
+\bibliography{paper}
 
@@
     Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
-Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
+Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
 \end{definition}
@@
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively.
+For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
 \end{definition}
@@
 
-\begin{remark}
-\label{rem:1dofcanon}
-
+\begin{remark}\label{rem:1dofcanon}
     All $(k,l)$-tight canonical DR-plans are optimal.
 We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar and joint graphs discussed in Section \ref{sec:DRP}.
-This is due tot he fact that such $(k,l)$-tight graphs are matroidal and have a pebble game.
+This is due to the fact that such $(k,l)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
 \end{remark}
@@
 (d) is the result of adding one such body to the bold-faced node.}
-
+%
 \end{figure*}
@@
 In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children.
-SO we will have to try pinning all nodes to see which gives the lowest algebraic complexity.
+So we will have to try pinning all nodes to see which gives the lowest algebraic complexity.
 This is still quadratic for the 1-dof case.
@@
 \begin{openproblem}
-    What is the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(k,l)$ characterization is no longer matroidal.
+    What is the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(k,l)$ characterization is no longer matroidal \cite{Lee:2007:PGA}.
 \end{openproblem}

### 2014-12-18 doc/GMP2015/s05_problems.tex
Further proof reading. Fixed \emph problem!
Added the open questions to the conclusion

--- 2014-12-18 doc/GMP2015/s09_problems.tex
@@
 polynomial time algorithms, the algorithms run in time exponential in
-$k$. As a result, when $k$ is part of the input, the \dfn{restricted
-OMD problem} has the potential to be difficult. For example, when the
-isostatic completion $H$ is required to be a 2-tree the restricted OMD
-problem is reducible to the maximum spanning series-parallel subgraph
-problem shown by \uncited\uncited to be NP-complete even if the input
-graph is planar of maximum degree at most 6. However, since the OMD
-problem has other input restrictions such as not having any proper
-isostatic subgraphs, it is not clear if the reverse reduction exists
-and hence it is unclear whether the OMD problem is NP-complete. The
-same holds for the restricted OMD problem where the isostatic
-completion $H$ is required to be a tree-decomposable graph of low
-Cayley complexity (i.e, have special, small DR-plans). One potential
-obstacle to an indecomposable graph $G$'s membership in the restricted
-OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
-In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
-have been constructed in \uncited.
-%
+$k$.
+
+We discuss the complexity of the restricted 
+OMD problem (when $k$ is part of the input) 
+in Section \ref{sec:conclusion}.
+
 \subsection{Optimal Modification, Completion and Recombination: formal
@@
 \begin{observation}
+\label{obs:OC_to_OMB}
     The (decision version of) the {\bf optimal completion problem
     (OC)} from \cite{bib:survey, j-a, fang} is OMD$_0$.
-\end{observation}.
+\end{observation}
 In fact, a \dfn{restricted OC} problem was studied by \cite{j-a}
-requiring the completion to be tree-decomposable. It is not clear if
-the reverse reduction holds, i.e, is OMD reducible to OC?
+requiring the completion to be tree-decomposable. 
+%It is not clear if
+%the reverse reduction holds, i.e, is OMD reducible to OC?
 
@@
 This reduction is used also for transferring  algorithms for optimal
-DR-planning, recombination, completion, OMD, and other problems and to
-so-called \dfn{multi-body pin} from so-called \dfn{bar-joint-systems},
+DR-planning, recombination, completion, OMD, and other problems 
+from bar-joint systems to 
+so-called \dfn{multi-body pin} define in Section \ref{sec:bodypin}
 by showing that the corresponding problems for the latter are reduced
@@
 
-\note This section deals only with bar-joint systems, however,
-extensions of several of these problems, approaches and solutions to
-other qusecs beyond  bar-joint - specifically pinned line incidence
-systems and multi-body-pin systems will be expanded in Sections
-\ref{sec:pinnedline} and \ref{sec:bodypin}.
+%\note This section deals only with bar-joint systems, however,
+%extensions of several of these problems, approaches and solutions to
+%other qusecs beyond  bar-joint - specifically pinned line incidence
+%systems and multi-body-pin systems will be expanded in Sections
+% \ref{sec:pinnedline} and \ref{sec:bodypin}.

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Added the open questions to the conclusion

@@
 \begin{observation}
+\label{obs:algebraic_completion}
     While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan with minimum fanin, we did not consider the algebraic complexity of realizing the DR-plan.
@@
 \end{proof}
-
-\begin{openproblem}
-    What is the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(k,l)$ characterization is no longer matroidal \cite{Lee:2007:PGA}.
-\end{openproblem}

### 2014-12-18 doc/GMP2015/s09_conclusion.tex
Added the open questions to the conclusion

@@
 \section{Conclusion}
+\bibliography{paper}
+
 We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem.
@@
 
-Throughout the paper, we raised some open questions and future areas of research.
-These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}
+\subsection{Furture Work}
+Throughout the process, we have also raised a few open questions.
+The first 2 come from Section \ref{sec:recomb}.
+The first has to do with an actual algorithm for finding an OMD that follows the theory we laid out:
+
+\begin{openproblem}    
+What is the complexity of the restricted OMD problem?
+This has the potential to be difficult. For example, when the
+isostatic completion $H$ is required to be a 2-tree the restricted OMD
+problem is reducible to the maximum spanning series-parallel subgraph
+problem shown by \cite{cai1993spanning} to be NP-complete even if the input
+graph is planar of maximum degree at most 6. However, since the OMD
+problem has other input restrictions such as not having any proper
+isostatic subgraphs, it is not clear if the reverse reduction exists
+and hence it is unclear whether the OMD problem is NP-complete.
+
+The same holds for the restricted OMD problem where the isostatic
+completion $H$ is required to be a tree-decomposable graph of low
+Cayley complexity (i.e, have special, small DR-plans). One potential
+obstacle to an indecomposable graph $G$'s membership in the restricted
+OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
+have been constructed in \cite{servatius2000rigidity}.
+\end{openproblem}
+
+The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in Section \ref{sec:table}.
+
+\begin{openproblem}
+    Is the OMD (optimal modification for decomposition) problem  
+    reducible to the OC (Optimal completion) problem?
+\end{openproblem}
+
+The final one comes about due to our limitations of the proof techniquies in Section \ref{sec:bodypin}.
+
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when the given graph has more than 2-dof?
+Our proof for the 1 and 2-dof cases relied heavily on the matroidal properties of their corresponding tightness.
+For higher number of dofs, the $(k,l)$ characterization is no longer matroidal \cite{Lee:2007:PGA}.
+Even so, if higher dofs had the same characteristics, Observation \ref{obs:algebraic_completion} raises some problems of finding a completion with small algebraic complexity.
+Unless some restrictions can be found and taken advantage of, the $k$-dof optimal completion problem would still be $O(n^{k+1})$.
+
+    The major barrier facing the porblem though is that there is no easy way of obtaining a $k$-dof DR-plan in general. 
+\end{openproblem}

### 2014-12-18 doc/GMP2015/s05_problems.tex
Citation added to Section 5

--- 2014-12-18 doc/GMP2015/s09_problems.tex
@@
 a 2-tree or a tree-decomposable graph for which simple near-linear
-time algorithms are available \uncited and the DR-plan is a
+time algorithms are available \cite{valdes1979recognition,fudos1997graph} and the DR-plan is a
 by-product output of the recognition algorithm.
@@
 problem is reducible to the maximum spanning series-parallel subgraph
-problem shown by \uncited\uncited to be NP-complete even if the input
+problem shown by \cite{cai1993spanning} to be NP-complete even if the input
 graph is planar of maximum degree at most 6. However, since the OMD
@@
 In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
-have been constructed in \uncited.
+have been constructed in \cite{servatius2000rigidity}.
 %
@@
     (OC)} from \cite{bib:survey, j-a, fang} is OMD$_0$.
-\end{observation}.
+\end{observation}
 In fact, a \dfn{restricted OC} problem was studied by \cite{j-a}

### 2014-12-19 doc/GMP2015/s01_introduction.tex
More references and citations.

@@
   \includegraphics[width=\linewidth]{img/Chlamydomonas_TEM_17}
-  \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules.}
-  % https://en.wikipedia.org/wiki/Axoneme#mediaviewer/File:Chlamydomonas_TEM_17.jpg
+  \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
   \label{fig:material_examples:microtubule}
@@
   \includegraphics[width=\linewidth]{img/ligten2}
-  \caption{Cross section of a tendon displaying the hierarchical structure.}
-  % http://www.umich.edu/~bme332/ch10ligten/ligten2.jpg
+  \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
   \label{fig:material_examples:tendon}
@@
   \includegraphics[width=\linewidth]{img/Rothemund-DNA-SierpinskiGasket}
-  \caption{A DNA array exhibiting the Sierpinski triangle.}
-  % https://en.wikipedia.org/wiki/DNA_nanotechnology#mediaviewer/File:Rothemund-DNA-SierpinskiGasket.jpg
+  \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
   \label{fig:material_examples:sierpinski}
@@
 
-    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{CellWalls1}~\cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+See Section \ref{sec:pinnedline}.
 
@@
     materials that behave like assemblies of 2D particles under
-    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}
+    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
 \end{enumerate}

### 2014-12-19 doc/GMP2015/s09_conclusion.tex
More references and citations.
Added Meeras new conclusion

--- 2014-12-18 doc/GMP2015/s09_conclusion.tex
@@
 \section{Conclusion}
-We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem.
-To this end we have defined a new canonical DR-plan and given an effiecient ($O(n^2)$) algorithm to find the DR-plan for specific types of 2D qusecs.
+\bibliography{paper}
 
-We have also described methods of recombining the DR-plan to get a realization different from the normal solving of a system of equations.
-These results rely on the results obtained in the Cayley Configuration space of certain graphs.
+We have elucidated the complexity barriers, for example overconstraints, 
+faced by the general optial DR-plan and recombination problem. 
+When there are no overconstraints, which is the case for 2D qusecs whose
+realizations are many common types of layered
+materials, we defined a canonical DR-plan and showed that such canonical
+DR-plans are guaranteed to be optimal, a strong Church-Rosser property given
+an effiecient ($O(n^2)$) algorithm to find a DR-plan that satisfies other
+desirable characteristics.
 
-We then related these theoretical results to real-world situations in which they can be applied including Examples 1-5 in Section \ref{sec:intro}.
+We have also described novel methods of efficiently realizing a 2D qusec 
+from the optimal DR-plan by modifying the otherwise nondecomposable 
+system that recombines the solved child systems at a node of a DR-plan
+These results rely on a recent theory of Convex Cayley Configuration spaces.
 
-Throughout the paper, we raised some open questions and future areas of research.
-These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}
+We then modeled specific layered materials
+using extensions of the above theoretical results
+including the motivating Examples 1-5 in the introduction.
+
+\subsection{Furture Work}
+A few natural  open questions are the following. 
+The first 2 are from Section \ref{sec:recomb}.
+
+\begin{openproblem}
+    Combinatorial rigidity for 
+    periodic structures is an active area of research. This paper motivates
+    a study of the rigidity of self-similar structures, 
+    with self-similar groups replacing periodic groups.
+\end{openproblem}
+
+\begin{openproblem}
+    The pinned line incidence structures of Section \ref{sec:pinnedline}, 
+    for example in the case of collagen microfibrils, whose function is
+    elastic contraction, should be considered
+    congruent under projective transformations. I.e, the projective group
+    should be factored out as a trivial motion in a new project for 
+    extending the combinatorial rigidity characterization of such systems.
+    (Currently we permit no trivial motions at all).
+\end{openproblem}
+
+\begin{openproblem}    
+What is the complexity of the restricted OMD (optimal modification for
+decomposition) problem?
+This has the potential to be difficult. For example, when the
+isostatic completion is required to be a 2-tree the restricted OMD
+problem is reducible to the maximum spanning series-parallel subgraph
+problem shown by \cite{cai1993spanning} to be NP-complete even if the input
+graph is planar of maximum degree at most 6. However, since the OMD
+problem has other input restrictions such as not having any proper
+isostatic subgraphs, it is not clear if the reverse reduction exists
+and hence it is unclear whether the OMD problem is NP-complete.
+
+The same holds for the restricted OMD problem where the isostatic
+completion is required to be a tree-decomposable graph of low
+Cayley complexity (i.e, have special, small DR-plans). One potential
+obstacle to an indecomposable graph $G$'s membership in the restricted
+OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
+have been constructed in \cite{servatius2000rigidity}.
+\end{openproblem}
+
+The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in Section \ref{sec:table}.
+
+\begin{openproblem}
+    Is the OMD (optimal modification for decomposition) problem  
+    reducible to the OC (Optimal completion) problem?
+\end{openproblem}
+
+The last problem is from 
+Section \ref{sec:bodypin}.
+
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when the given
+    graph has more than 2-dofs? 
+    Our proof for the 1 and 2-dof cases relied heavily on the matroidal 
+    properties of their corresponding $(k,l)$-tightness. 
+    For higher number of dofs, the $(k,l)$ characterization is 
+    no longer matroidal \cite{Lee:2007:PGA}. 
+    As a result, the major obstacle is that 
+    there is no easy way of obtaining an optimal or canonical 
+    $k$-dof DR-plan in general. 
+    Even assuming such  a DR-plan is available,
+    if higher dofs had the same characteristics, 
+    Observation \ref{obs:algebraic_completion} 
+    raises questions about the correct measure of DR-plan size that 
+    captures algebraic complexity for recombining graphs with many dofs
+    (this is not an issue in the isostatic case). Unless some restrictions 
+    can be found and taken advantage of, the $k$-dof optimal 
+    completion problem would  have complexity exponential in $k$.
+
+\end{openproblem}

### 2014-12-19 doc/GMP2015/s03_drp.tex

@@
 
-The following subsections provide the supporting proofs.
+% The following subsections provide the supporting proofs.
 
@@
 
-\subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
-\label{sec:union_intersection}
+% \subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
+% \label{sec:union_intersection}
 
@@
 
-\subsubsection{Inferring Parent Structure from Union and Intersection of Children}
-\label{sec:infer_parent}
+% \subsubsection{Inferring Parent Structure from Union and Intersection of Children}
+% \label{sec:infer_parent}
 
@@
 
-\subsection{Example}
-
 % We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.
@@
 
-% \begin{example}
+% \myexample
 % In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
@@
 Were any left out, the union of the children would not be the entire parent.
-% \end{example}
 
@@
 
+\myexample
+\textsl{[DR-Plan for self-similar structure]}
 This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
@@
 These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained.
-As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must occur 2 levels below the parent.
+As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
 Just as expected, it is on the third level, as it is a child of both of $G$'s children.

### 2014-12-19 doc/GMP2015/s04_recombination.tex

@@
 
-\medskip\noindent
-\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
+\myexample
 A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial

### 2014-12-19 doc/GMP2015/s05_problems.tex

--- 2014-12-18 doc/GMP2015/s05_problems.tex
@@
 
+%In the recombination setting, the corresponding \dfn{restricted
+%OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
+%completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
+%or to be a low Cayley complexity tree-decomposable graph as in Section
+%\ref{sec:tdecomp}. Clearly these problems have deterministic
+%polynomial time algorithms, the algorithms run in time exponential in
+%$k$. As a result, when $k$ is part of the input, the \dfn{restricted
+%OMD problem} has the potential to be difficult. For example, when the
+%isostatic completion $H$ is required to be a 2-tree the restricted OMD
+%problem is reducible to the maximum spanning series-parallel subgraph
+%problem shown by \cite{cai1993spanning} to be NP-complete even if the input
+%graph is planar of maximum degree at most 6. However, since the OMD
+%problem has other input restrictions such as not having any proper
+%isostatic subgraphs, it is not clear if the reverse reduction exists
+%and hence it is unclear whether the OMD problem is NP-complete. The
+%same holds for the restricted OMD problem where the isostatic
+%completion $H$ is required to be a tree-decomposable graph of low
+%Cayley complexity (i.e, have special, small DR-plans). One potential
+%obstacle to an indecomposable graph $G$'s membership in the restricted
+%OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+%In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
+%have been constructed in \cite{servatius2000rigidity}.
+%
 \subsection{Optimal Modification, Completion and Recombination: formal
@@
 \begin{observation}
-\label{obs:OC_to_OMB}
     The (decision version of) the {\bf optimal completion problem
-    (OC)} from \cite{bib:survey, j-a, fang} is OMD$_0$.
+    (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
 \end{observation}
-In fact, a \dfn{restricted OC} problem was studied by \cite{j-a}
+In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming}
 requiring the completion to be tree-decomposable. 
@@
 %systems and multi-body-pin systems will be expanded in Sections
-% \ref{sec:pinnedline} and \ref{sec:bodypin}.
+%\ref{sec:pinnedline} and \ref{sec:bodypin}.

### 2014-12-19 doc/GMP2015/s08_pinnedline.tex

@@
 
-The optimal qusecs DR-plan can also be applied to analyze and solve
-the structure of  cross-linking microfibrils in collagen and cellulose
-in biology, by modeling these structures as \dfn{pinned line-incidence
+The optimal DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve
+the structure of  cross-linking collagen and cellulose microfibrils 
+in animal and plant cell walls, and actin filaments in the cytoskeleton
+by modeling these structures as \dfn{pinned line-incidence
 systems}.
@@
 %It is additionally
-each of which is cross-linked at 2 places with some two other fibrils,
+each of which is cross-linked at 2 places with usually 3 other fibrils,
 where the \dfn{cross-linking} is like an incidence constraint that the
-two crosslinked fibrils can slide with each other (see
+crosslinked fibrils can slide against each other while remaining incident (see
 Figure~\ref{fig:crosslink}).
@@
 Note that no trivial motion exists since the pins have fixed positions on the plane.
+Euclidean transformations are not factored out.
 In particular, both a single vertex and a single edge are under-constrained graphs.
@@
 Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph,
-where the grey ovals stand for pins representing attachments of fibrils,
+where the grey ovals denote pins representing attachments of fibrils,
 and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
@@
 
-\begin{remark}\label{lem:pinned_union_intersection}
+\begin{lemma}\label{lem:pinned_union_intersection}
 Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$,
@@
 %\setlength{\tabcolsep}{6pt}
-\end{remark}
+\end{lemma}
 
-Given Remark~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
+Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
 proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.

### 2014-12-19 doc/GMP2015/appendix_a.tex
Pulled proofs from my section and Meera's seection into the appendix

@@
 \end{proof}
+
+\subsection{Proofs form Section \ref{sec:recomb}}
+
+\subsubsection{Proof of Theorem \ref{theorem:omdk}}
+\begin{proof}
+The proof follows from the fact that 2-trees are well decomposable and
+have simple DR-plans of size 2. We know that $G$ can be reduced by
+removing $k$ edges to create a partial-2-tree $G'$ which can then be
+completed to an (isostatic) 2-tree by adding some set of non-edges
+$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
+proving the theorem.
+\end{proof}
+
+\subsubsection{Proof of Theorem \ref{theorem:criterionc}}
+\begin{proof}
+    The Cartesian realization space of $(H,\left<\delta_{E'},
+    \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
+    and is the union of $2^t$ solutions (modulo orientation preserving
+    isometries) each with a distinct orientation type, where $t$ is
+    the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
+    the restriction of the length vector $\delta$ to the edges in $S$.
+    A desired solution $p$ (or connected component of a solution
+    space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
+    found by a subdivided binary search   of the cartesian realization
+    space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
+    type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
+    convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
+    described in Theorem \ref{theorem:convexcayley}. A solution $p$
+    is found  when the lengths for nonedges in $D$ match $\delta_D$.
+\end{proof}
+
+\subsection{Proofs from Section \ref{sec:bodypin}}
+
+\subsubsection{Proof of Remark \ref{rem:bodypin_is_barjoint}}
+\begin{proof}
+    We can replace each body that has only one pin by a single vertex.
+A body with 2 pins can be replaced by an edge.
+In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
+When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
+The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged.
+The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+\end{proof}
+
+\subsubsection{Proof of Observation \ref{obs:bodypin_drp}}
+\begin{proof}
+    \todo{Item 1 Follows from paper?}
+
+    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
+Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
+\end{proof}
+
+\subsubsection{Proof of Observation \ref{thm:1dofcase}}
+\begin{proof}
+    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
+Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.
+
+    To make the graph isostatic, we need only add one body and pin it to 2 other bodies.
+Doing so will cause $G$ to become $(3,3)$-tight.
+We choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
+From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a sub-graph.
+We pin the new body to bodies in two separate children.
+Doing so will ensure that all children of $b$ will have 1-dof and all ancestors of $b$ (including $b$) will now be isostatic.
+
+    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+In $T_b$, $fanin(b)$is the number of leaves in the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
+Similarly, for any other node $w$ that is an ancestor of $b$, $fanin(w)$ is the the number leaf nodes in the tree rooted at $w$, excluding the subtree constaining $b$, plus 1 (for the node leading to $b$).
+Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
+The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
+Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
+
+    Such a pinning covers all possible ways of adding a new body.
+Assume we add a new body $b$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
+Then, there will be some lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$.
+Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning $b$ to $b_i$ and $b_j$.
+
+    For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$.
+$|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$.
+We are looking for $b$ that minimizes $|T_b|$.
+Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$.
+Note that $fanin(b)= nl(T^b)$ because no descendant of $b$ is isostatic.
+Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$.
+All other nodes are not isostatic and so have no fanin.
+
+    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
+Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$, excluding $b'$.
+Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
+The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
+Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
+
+    The node we need to pin will always be the deepest nontrivial node of some path in $T$.
+Suppose we chose to pin a node $b$ that has a nontrivial child $v$.
+Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$.
+If we had instead chosen to pin $v$, then $fanin_v(b) = nl(T^b) - nl(T^{b'}) + 1 \leq fanin_b(v)$.
+And for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $|T_v| \leq |T_b|$.
+Thus we only have to check the deepest non-trivial nodes.
+
+    Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
+
+    For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
+    % Don't leave a blank line between last paragraph and end!
+    % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
+These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1-dof case.
+\end{proof}
+
+\subsubsection{Proof of Observation \ref{obs:2dof_case}}
+\begin{proof}
+    The only difference from the 1-dof case is that now we need to remove 2-dof from our graph.
+We build a 2-dof DR-plan $T$.
+Like above, we need to add a body and 2 pins to 2 nodes now to get to isostatic.
+
+    Suppose we pin 2 distinct nodes $v_i$ and $v_j$.
+Then, there will be some common ancestor $a$ of $v_i$ and $v_j$.
+Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$.
+However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin_v(a) = nl(T^a) - nl(T^{a'}) +1$ .
+Thus $fanin_v(a)' \leq fanin_{v_i,v_j}(a)$.
+All ancestors of $a$ are unchanged.
+So $|T_v| \leq |T_{v_i,v_j}|$.
+
+    So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
+\end{proof}
+
+\subsubsection{Proof of Observation \ref{obs:algebraic_completion}}
+\begin{proof}
+    An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton).
+These are the Euclidean motions.
+A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has.
+A 2-dof has 5 parameters, and so on.
+These paramters denote the algebraic complexity of giving a realization for the graphs.
+
+    Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}.
+When we pin a node $b$, we will have the same structure as before.
+Suppose we are looking at an isostatic node $v$ after pinning $b$.
+Then, the children of $v$ (except one if $v \neq b$) will be 1-dof.
+The complexity for that node is then just that of solving each of its children.
+In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.
+
+    Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$.
+In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children.
+So we will have to try pinning all nodes to see which gives the lowest algebraic complexity.
+This is still quadratic for the 1-dof case.
+
+    For the 2-dof case, there are more cases to consider.
+If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$.
+If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof.
+Note that solving $v'$ will result in solving $v$.
+Then, we need to consider nodes above and includiing $v'$ in our complexity.
+$np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
+The only remaining case will be if we pin two nodes that are not descendent/ancestor.
+the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.
+For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.
+
+    Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin.
+However, we cannot also assume pinning one node twice will give us the best algebraic complexity.
+Hence, we will need to check each pair of node to pin.
+This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
+\end{proof}

### 2014-12-19 doc/GMP2015/s04_recombination.tex
Pulled proofs from my section and Meera's seection into the appendix

@@
 
-\begin{proof}
-The proof follows from the fact that 2-trees are well decomposable and
-have simple DR-plans of size 2. We know that $G$ can be reduced by
-removing $k$ edges to create a partial-2-tree $G'$ which can then be
-completed to an (isostatic) 2-tree by adding some set of non-edges
-$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
-proving the theorem.
-\end{proof}
+% \begin{proof}
+% The proof follows from the fact that 2-trees are well decomposable and
+% have simple DR-plans of size 2. We know that $G$ can be reduced by
+% removing $k$ edges to create a partial-2-tree $G'$ which can then be
+% completed to an (isostatic) 2-tree by adding some set of non-edges
+% $F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
+% proving the theorem.
+% \end{proof}
 
@@
 
-\begin{proof}
-    The Cartesian realization space of $(H,\left<\delta_{E'},
-    \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
-    and is the union of $2^t$ solutions (modulo orientation preserving
-    isometries) each with a distinct orientation type, where $t$ is
-    the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
-    the restriction of the length vector $\delta$ to the edges in $S$.
-    A desired solution $p$ (or connected component of a solution
-    space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
-    found by a subdivided binary search   of the cartesian realization
-    space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
-    type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
-    convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
-    described in Theorem \ref{theorem:convexcayley}. A solution $p$
-    is found  when the lengths for nonedges in $D$ match $\delta_D$.
-\end{proof}
+% \begin{proof}
+%     The Cartesian realization space of $(H,\left<\delta_{E'},
+%     \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
+%     and is the union of $2^t$ solutions (modulo orientation preserving
+%     isometries) each with a distinct orientation type, where $t$ is
+%     the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
+%     the restriction of the length vector $\delta$ to the edges in $S$.
+%     A desired solution $p$ (or connected component of a solution
+%     space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
+%     found by a subdivided binary search   of the cartesian realization
+%     space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
+%     type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
+%     convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
+%     described in Theorem \ref{theorem:convexcayley}. A solution $p$
+%     is found  when the lengths for nonedges in $D$ match $\delta_D$.
+% \end{proof}
 

### 2014-12-19 doc/GMP2015/s07_problempinning.tex
Pulled proofs from my section and Meera's seection into the appendix

@@
 \label{sec:bodypin}
-\bibliography{paper}
+% \bibliography{paper}
 
@@
 \begin{remark}
+\label{rem:bodypin_is_barjoint}
     A body-hyperpin qusecs is a special case of bar-joint qusecs that we have been studying.
@@
 \begin{observation}
+\label{obs:bodypin_drp}
     For the 1-dof body-hyperpin graphs described above, a DR-plan node will follow one of the following:
@@
 
-\begin{proof}
-    \todo{Item 1 Follows from paper?}
+% \begin{proof}
+%     \todo{Item 1 Follows from paper?}
 
-    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
+%     For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
 Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
-\end{proof}
+% \end{proof}
 
@@
 
-\begin{proof}
-    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
+% \begin{proof}
+%     Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
 Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.
 
-    To make the graph isostatic, we need only add one body and pin it to 2 other bodies.
+%     To make the graph isostatic, we need only add one body and pin it to 2 other bodies.
 Doing so will cause $G$ to become $(3,3)$-tight.
@@
 
-    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+%     % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
 In $T_b$, $fanin(b)$is the number of leaves in the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
@@
 
-    Such a pinning covers all possible ways of adding a new body.
+%     Such a pinning covers all possible ways of adding a new body.
 Assume we add a new body $b$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
@@
 
-    For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$.
+%     For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$.
 $|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$.
@@
 
-    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
+%     % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
 In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
@@
 
-    The node we need to pin will always be the deepest nontrivial node of some path in $T$.
+%     The node we need to pin will always be the deepest nontrivial node of some path in $T$.
 Suppose we chose to pin a node $b$ that has a nontrivial child $v$.
@@
 
-    Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
+%     Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
 
-    For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
-    % Don't leave a blank line between last paragraph and end!
-    % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
+%     For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
+%     % Don't leave a blank line between last paragraph and end!
+%     % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node.
 These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1-dof case.
-\end{proof}
+% \end{proof}
 
 \begin{observation}
+\label{obs:2dof_case}
     For the 2-dof case, if a similar statement to Remark \ref{rem:1dofcanon} can be made, then Theorem \ref{thm:1dofcase} is true for 2-dof as well.
@@
 
-\begin{proof}
-    The only difference from the 1-dof case is that now we need to remove 2-dof from our graph.
+% \begin{proof} (of Observation \ref{obs:2dof_case})
+%     The only difference from the 1-dof case is that now we need to remove 2-dof from our graph.
 We build a 2-dof DR-plan $T$.
@@
 
-    Suppose we pin 2 distinct nodes $v_i$ and $v_j$.
+%     Suppose we pin 2 distinct nodes $v_i$ and $v_j$.
 Then, there will be some common ancestor $a$ of $v_i$ and $v_j$.
@@
 
-    So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
-\end{proof}
+%     So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
+% \end{proof}
 
@@
 
-\begin{proof}
-    An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton).
+% \begin{proof}
+%     An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton).
 These are the Euclidean motions.
@@
 
-    Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}.
+%     Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}.
 When we pin a node $b$, we will have the same structure as before.
@@
 
-    Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$.
+%     Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$.
 In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children.
@@
 
-    For the 2-dof case, there are more cases to consider.
+%     For the 2-dof case, there are more cases to consider.
 If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$.
@@
 
-    Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin.
+%     Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin.
 However, we cannot also assume pinning one node twice will give us the best algebraic complexity.
@@
 This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
-\end{proof}
+% \end{proof}

### 2014-12-19 doc/GMP2015/appendix_a.tex
Pulled proofs from my section and Meera's seection into the appendix
Changed DRP section once more. Moving proofs to appendix.

@@
 
-\subsection{Section \ref{sec:DRP} -- {\sffamily \textbf{Lemma \ref{lemma:wc_intersection_makes_all_wc}}}}
+\subsection{Proofs from Section \ref{sec:DRP}}
 
-% \textbf{Section \ref{sec:DRP}}
+\newcommand{\usestwod}{\todo{Uses 2D requirement:}}
+\renewcommand{\usestwod}{}
 
-% {\sffamily \textbf{Lemma \ref{lemma:main_trivial}}
+\subsubsection{Proof of Theorem \ref{theorem:main}}
+
+\begin{proof}
+Remark \ref{lemma:union_intersection} shows that the intersection of any two well-constrained subgraphs (such as $C_i$ and $C_j$) can only result in trivial or well-constrained subgraphs.
+Therefore, these are the only possibilities to consider.
+
+\medskip\noindent
+\vemph{Part 1:} We show that no subset of the children can union to form $C$, thereby requiring all of them be included, and because this is the only choice it must be optimal for this node.
+
+Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained.
+If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
+So, it must be that $U=C$.
+\usestwod
+However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
+By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
+Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$.
+Thus, the assumption is wrong and $U$ cannot be well-constrained.
+As $C$ is well-constrained, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.
+
+\medskip\noindent
+\vemph{Part 2:} If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
+This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
+Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
+
+However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in for any descendant.
+Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
+Suppose we select $i$ and $j$, where $i\neq j$, as the children.
+Since
+\[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
+the children of this node will be
+\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
+and
+\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
+% $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
+% the children of this node will be
+% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
+% and
+% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
+for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other.
+\todo{Prove these are valid children?
+Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once.
+Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
+
+This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
+\end{proof}
+
+\subsubsection{Proof of Observation \ref{lemma:union_intersection}}
+
+\begin{proof}
+For (1), simply note that if $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
+
+For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$.
+For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained.
+Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$.
+Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
+
+For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
+\end{proof}
+
+% \begin{proof}
+% Use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
+% \begin{enumerate}
+%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
+
+%     \item Observe that  Then it is straightforward application of the equation above.
+
+%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
+
+%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.
+
+%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
+Therefore $d(F_i\cap F_j)=2k-l>k$.
+This means $F_i\cap F_j$ is trivial.
+\textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial.
+By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.
+
+%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
+Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
+\textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.
+
+%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
+% \end{enumerate}
+% \end{proof}
+
+\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_is_C}}
+
+\begin{proof}
+Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
+%
+In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
+Thus, $C_i\cup C_j=C$ is well-constrained.
+\end{proof}
+
+\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc}}
+
 \begin{corollary}\label{corollary:no_edges_between_diff}
@@
 Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.
-
-\todo{Need a figure?}
+%
+% \todo{Need a figure?}
 \end{proof}
 
-\subsection{Section \ref{sec:pinnedline} -- proof of {\sffamily \textbf{Lemma \ref{lemma:wc_intersection_makes_all_wc}}} for  2-dimensional pinned line incidence graphs }
+\subsubsection{Proof of Lemma \ref{lemma:uc_intersection_makes_all_uc}}
+
+\begin{proof}
+Assume there is some $k$ such that $C_i\cap C_k$ is not trivial.
+By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be well-constrained.
+Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be well-constrained.
+This means that $C_i\cap C_j$ is well-constrained.
+Therefore, such a $k$ cannot exist and all intersections are trivial.
+\end{proof}
+
+\subsection{Proofs from Section \ref{sec:pinnedline}}
+\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
 \label{sec:appendix_pinned}
@@
 
-\subsection{Proofs form Section \ref{sec:recomb}}
+\subsection{Proofs from Section \ref{sec:recomb}}
 

### 2014-12-19 doc/GMP2015/s03_drp.tex
Changed DRP section once more. Moving proofs to appendix.

@@
-\section{The Canonical Decomposition-Recombination Plan}
+\section{The canonical DR-plan}
 \label{sec:DRP}
-
-\newcommand{\usestwod}{\todo{Uses 2D requirement:}}
-\renewcommand{\usestwod}{}
 
@@
 
-In the problem of the optimal DR-Plan there is generally not a unique plan.
-% Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
+% In the problem of the optimal DR-Plan there is generally not a unique plan.
+% % Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
 Therefore, all choices of decomposition are in some sense equivalent.
 The theorem we seek to prove is thus:
-However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems.
+% However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems.
 Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
-% To assist in showing this, we prove this core theorem throughout this section:
+% % To assist in showing this, we prove this core theorem throughout this section:
 
-In this section, we discuss 2D bar-joint graphs.
+% In this section, we discuss 2D bar-joint graphs.
 All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$.
@@
 Furthermore, 2D well-constrained graphs must be connected.
-% The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
+% % The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
 The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
 
-\begin{theorem}\label{theorem:main}
-Given a well-constrained 2-dimensional bar-joint graph $G$, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
+In the problem of the optimal DR-Plan there is generally not a unique plan.
+Furthermore, there is no obvious way to find the optimal plan.
+In this section we present a particular way to construct a plan that guarantees optimality.
+We call this the \dfn{canonical} plan.
+
+\begin{definition}
+    The \dfn{canonical DR-plan} is a DR-plan that satisfies the following two properties:
     \begin{enumerate}
-    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$
-    \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
+        \item Children are isostatic vertex-maximal subgraphs of the parent.
+        \item All of the isostatic vertex-maximal subgraphs are children if any two of them intersect trivially, otherwise exactly two are children.
     \end{enumerate}
-This covers all potential cases.
+\end{definition}
+
+The importance of this formulation is stated succinctly in the following theorem.
+
+\begin{theorem}\label{theorem:canonical_is_optimal}
+    A canonical DR-plan is an optimal DR-plan.
 \end{theorem}
 
-\begin{proof}
-Remark \ref{lemma:union_intersection} shows that the intersection of any two well-constrained subgraphs (such as $C_i$ and $C_j$) can only result in trivial or well-constrained subgraphs.
-Therefore, these are the only possibilities to consider.
+The proof of this theorem is a direct result of this more general theorem.
 
-\medskip\noindent
-\vemph{Part 1:} We show that no subset of the children can union to form $C$, thereby requiring all of them be included in an optimal plan.
+\begin{theorem}\label{theorem:main}
+Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children according to the following rules.
+\begin{enumerate}
+    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
+    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
+\end{enumerate}
+This covers all cases and the resulting plan is an optimal DR-plan.
+\end{theorem}
 
-Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained.
-If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
-So, it must be that $U=C$.
-\usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
-By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
-Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$.
-Thus, the assumption is wrong and $U$ cannot be well-constrained.
-As $C$ is well-constrained, this means no proper subset of the children in $CompleteDRP(C)$ can union to form $C$.
-
-\medskip\noindent
-\vemph{Part 2:} If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
-This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
-Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
-
-However, to guarantee that any two are \vemph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant.
-Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
-Suppose we select $i$ and $j$, where $i\neq j$, as the children.
-Since
-% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
-% the children of this node will be
-% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
-% and
-% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
-$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
-the children of this node will be
-$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
-and
-$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other.
-\todo{Prove these are valid children?
-Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once.
-Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
-
-This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
-\end{proof}
-
-% The following subsections provide the supporting proofs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
-% \label{sec:union_intersection}
-
-% This is general, not for 2d
-\begin{remark}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
+\begin{remark}
+    \todo{this captures cluster minimality}
 \end{remark}
 
-\begin{proof}
-For (1), simply note that if $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
+This theorem (proven in the Appendix) relies on the following crucial observations and lemmas.
 
-For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$.
-For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained.
-Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$.
-Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
-
-For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
-\end{proof}
-
-% \begin{proof}
-% Use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
-% \begin{enumerate}
-%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.
-
-%     \item Observe that  Then it is straightforward application of the equation above.
-
-%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
-
-%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.
-
-%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
-Therefore $d(F_i\cap F_j)=2k-l>k$.
-This means $F_i\cap F_j$ is trivial.
-\textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial.
-By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.
-
-%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
-Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
-\textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.
-
-%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
-% \end{enumerate}
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsubsection{Inferring Parent Structure from Union and Intersection of Children}
-% \label{sec:infer_parent}
+\begin{observation}\label{lemma:union_intersection}
+If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
+\end{observation}
 
@@
 \end{lemma}
-%
-\begin{proof}
-Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
-%
-In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
-Thus, $C_i\cup C_j=C$ is well-constrained.
-\end{proof}
-%
-%
-%
+
 \begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
@@
 \end{lemma}
-%
-\begin{proof}
-In appendix (\ref{sec:appendix}).
-\end{proof}
-%
-%
-%
+
 \begin{lemma}\label{lemma:uc_intersection_makes_all_uc}
@@
 \end{lemma}
-%
-\begin{proof}
-Assume there is some $k$ such that $C_i\cap C_k$ is not trivial.
-By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be well-constrained.
-Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be well-constrained.
-This means that $C_i\cap C_j$ is well-constrained.
-Therefore, such a $k$ cannot exist and all intersections are trivial.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsubsection{Further Results}
-
-% \begin{corollary}
-% Take $\bigcap_{k=1}^N{C_k}$ to be the graph $D$. If $D$ is well-constrained, then $D$ will be a descendant of every $C_k$.
-% \end{corollary}
-
-% \begin{proof}
-% Explained in Theorem \ref{theorem:main_wellconstrained}.
-% \end{proof}
-
-% \begin{corollary}
-% $\forall i,j\in [1,\ldots,N]$ if $i\neq j$ there are no edges in $C$ between the vertices in subgraphs $C\setminus C_i$ and $C\setminus C_j$.
-% \end{corollary}
-
-% \begin{proof}
-% Further constraints between them would imply $C_i \cup C_j$ is a subgraph of $C$, but lemma \ref{lemma:wc_intersection_is_C} proves is the entire graph.
-% \end{proof}
-
-% \begin{proof}
-% It follows from Theorem \ref{theorem:main_trivial} that if the intersection of any two well-constrained vertex-maximal proper subgraphs of $C$ are well-constrained then all of them are well-constrained or $\emptyset$.
-% \end{proof}
-
-% Figure idea: Large node in the middle with weight $k$.
-Several other graphs with weight $a, b, c, d\ldots$ attached radially by single edges with weights $a, b, c, d\ldots$ respectively.
-Circle each vertex-maximal subset.
-
-% $\forall k\in [1,\ldots,N]$ the subgraphs $C\setminus C_k$ will be edge disjoint. Therefore the solving of one is exclusive of the solving of the other.
-
-% Furthermore, the union of all of these subgraphs will be $C\setminus D$.
-So each subgraph will need to be ``tacked'' onto $D$ for recombination and since they are edge-disjoint the order is unimportant.
-
-% \pnplaceholder
-% \begin{lemma}
-% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
-% \end{lemma}
-
-% \begin{proof}
-% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-% \pnplaceholder
-
-% \begin{corollary}
-% There are no edges between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-% \end{corollary}
-
-% \begin{proof}
-% As a result of the discussion of the second part, it is clear that there cannot be any further constraints between $V_1 \setminus (V_1\cap V_2)$ and $V_2 \setminus (V_1\cap V_2)$.
-That would imply $C_1 \cup C_2$ is a subgraph, but we know it to be the entire graph.
-% \end{proof}
-
-% \begin{corollary}
-% There is no rigid subgraph induced by a proper, nonempty subset of vertices of $V_i \setminus (V_1\cap V_2)$ and a subset of $V_1\cap V_2$ of size at least 2.
-% \end{corollary}
-
-% \begin{proof}
-% Since $C_1\cup C_2$ is the entire graph
-% \todo{Prove this!!!!}
-% \end{proof}
-
-% \begin{corollary}
-% If $C_1$ and $C_2$ are well-constrained vertex-maximal proper subgraphs of $C$ and $C_1 \cap C_2$ is well-constrained, then $C_1 \cap C_2$ is a well-constrained vertex-maximal proper subgraph of every child of $C$.
-% \end{corollary}
-
-% \begin{proof}
-% First, it must be a vertex-maximal proper subgraph of $C_1$ and $C_2$.
-Say that there was some larger proper subgraph in $C_1$ that was well constrained.
-If this were true, then $C_2$ was not vertex-maximal because it could have included that entire subgraph and remained well-constrained.
-The same is true vice versa.
-
-% Second, the discussion in the proof of the theorem holds true for any two subgraphs. Therefore, the same intersection is present in all of them.
-% \end{proof}
-
-% We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.
-
-% \begin{figure*}\centering
-% \begin{subfigure}{.3\linewidth}\centering
-%     \input{img/tikz/two_doublets-0.tikz}
-%     \caption{}\label{fig:doublets:a}
-% \end{subfigure}%
-% \begin{subfigure}{.7\linewidth}\centering
-%     \input{img/tikz/two_doublets-1.tikz}
-%     \caption{}\label{fig:doublets:b}
-% \end{subfigure}
-
-% \caption{(\ref{fig:doublets:a}) Two doublets ($C_2\times C_3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$.
-(\ref{fig:doublets:b}) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
-% \label{fig:doublets}
-% \end{figure*}
-
-% \myexample
-% In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan.
-These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$.
-This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 1.
-Also shown is the decomposition of the individual doublets.
-These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them.
-As explained in Theorem \ref{theorem:main}, point 2, all of them must be included in the optimal DR-plan.
-Were any left out, the union of the children would not be the entire parent.
 

### 2014-12-19 doc/GMP2015/appendix_a.tex
Updated Section 7 based on new Section 3

@@
 
-First notice that since $C_i \cup C_j$ is well-constrained, by Lemma~\ref{lem:pinned_union_intersection},
+First notice that since $C_i \cup C_j$ is well-constrained, by Observation~\ref{lem:pinned_union_intersection},
 both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.

### 2014-12-19 doc/GMP2015/s08_pinnedline.tex
Updated Section 7 based on new Section 3

@@
 
-The optimal DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve
+The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve
 the structure of  cross-linking collagen and cellulose microfibrils 
 in animal and plant cell walls, and actin filaments in the cytoskeleton
-by modeling these structures as \dfn{pinned line-incidence
-systems}.
+by modeling these structures as \dfn{pinned line-incidence systems}.
 
@@
 %There are  and 10 vertices/cross-linkings in the graph.
-The graph is well-constrained, with 12 vertices and 24 edges/pins.
+The graph is isostatic, with 12 vertices and 24 edges/pins.
 
@@
   \centering
-   \includegraphics[width=.8\linewidth]{img/pinned}
-\caption{ A well-constrained pinned line-incidence graph.}
+   \includegraphics[width=.7\linewidth]{img/pinned}
+\caption{ An isostatic pinned line-incidence graph.}
 \label{fig:pinned_line}
@@
 In this section, we will adapt the results in Section~\ref{sec:DRP}
-to give the Optimal DR-plan for pinned line-incidence graphs.
+to give the canonical DR-plan for pinned line-incidence graphs.
 First, we note that %unlike bar-joint graphs,
-\vemph{a well-constrained pinned line-incidence graph can be disconnected},
-being the disjoint union of two or more well-constrained subgraphs.
+\vemph{an isostatic pinned line-incidence graph can be disconnected},
+being the disjoint union of two or more isostatic subgraphs.
 This is because the pins have fixed positions on the plane.
@@
 %
-We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of the DR-plan:
+We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of 
+the canonical DR-plan:
 
@@
 \begin{definition}
-The \dfn{DR-plan} of a well-constrained pinned line-incidence graph $G$ is one in which
-(1) each child node of a non-leaf node $C$ is either an edge not contained in any well-constrained proper subgraph of $C$, or a \dfn{connected} well-constrained vertex-maximal proper subgraph of $C$, and
-(2) a leaf node is a single edge.
+The \dfn{DR-plan} of a  pinned line-incidence graph $G$ is one in which
+(1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$,
+or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
 % Definition~\ref{def:drp} with  modified rules number 2 and 4:
@@
 %    \end{enumerate}
+
+The \dfn{canonical DR-plan} of $G$  is one in which 
+the  children containing rigid subgraphs are connected isostatic vertex-maximal subgraphs of the parent.
 \end{definition}
 
-With the above modification, we now adapt Theorem~\ref{theorem:main}
-for the optimal DR-plan for pinned line-incidence graphs.
+Theorem~\ref{theorem:canonical_is_optimal} and \ref{theorem:main}
+hold for  pinned line-incidence graphs with the above modified definition. 
+The proof is similar to the original proof (in the Appendix) using the same set of lemmas and  
+%
+%\begin{corollary}\label{cor:pinned}
+%Given a isostatic pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
+%\begin{enumerate}
+%    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$.
+%    \item if $C_i \cap C_j$ is isostatic then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
+%\end{enumerate}
+%\end{corollary}
+%
+the following modified version of Observation~\ref{lemma:union_intersection},
+% observation, which
+%serves a similar function to Observation~\ref{lemma:union_intersection} in Section~\ref{sec:DRP}
+%Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
+which  can be proved using a simple counting based argument.
 
-\begin{corollary}\label{cor:pinned}
-Given a well-constrained pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
-\begin{enumerate}
-    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$.
-    \item if $C_i \cap C_j$ is well-constrained then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
-\end{enumerate}
-\end{corollary}
-
-The following lemma serves a similar function to Remark~\ref{lemma:union_intersection} in Section~\ref{sec:union_intersection}
-%Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
-and can be proved using a similar counting based argument.
-
-\begin{lemma}\label{lem:pinned_union_intersection}
-Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$,
-where each of them can be either a single edge or a connected well-constrained subgraph.
+\begin{observation}\label{lem:pinned_union_intersection}
+Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$,
+where each of them can be either a single edge or a connected isostatic subgraph.
 There are only two possible cases:
-(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
+(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
 %The following table summarizes all bijections:
@@
 %\setlength{\tabcolsep}{6pt}
-\end{lemma}
+\end{observation}
 
 Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
-proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
+Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
 The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned
@@
 %\end{lemma}
-
-Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lemma:uc_intersection_makes_all_uc},
-it is straightforward to adapt the proof of Theorem~\ref{theorem:main}
-to  pinned line-incidence graphs,
-thus proving Corollary~\ref{cor:pinned}.
+%Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lemma:uc_intersection_makes_all_uc},
+Thus it is straightforward to adapt the proof of Theorem~\ref{theorem:main}
+to  pinned line-incidence graphs.
+%thus proving Corollary~\ref{cor:pinned}.
+%
 Consequently, we can efficiently find the optimal DR-plan for  pinned line-incidence graphs using basically the same algorithm as for  bar-joint graphs.
@@
 Since the pins are given fixed positions in the plane,
-the solutions of a well-constrained sub-system will automatically be consistent with solutions of the remaining of the system.
+the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.

### 2014-12-19 doc/GMP2015/s01_introduction.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.

@@
     Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other
-    self-similar structures~\cite{Microfibrils2} within reach.
+    self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
 

### 2014-12-19 doc/GMP2015/s02_preliminaries.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.

@@
     The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \cite{hoffman2001decompositionI} {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
+    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}), is defined as a forest that has the
     following properties:
@@
     For a given graph, there could be exponentially many  DR-plans and
-    even optimal DR-plans in the size of the graph . A complete DR-
-    plan is unique but may not be (is usually not) optimal. DR-plans
+    even optimal DR-plans in the size of the graph . A complete DR-plan
+    is unique but may not be (is usually not) optimal. DR-plans
     of self-similar graphs are self-similar. More than one node (leaf)
@@
 Without DR-planning, this requires double exponential time in the
-number of variables (even if orientation type is specified). With DR-
-planning, the complexity is dominated by the size of the largest
+number of variables (even if orientation type is specified). With
+DR-planning, the complexity is dominated by the size of the largest
 subsystem that is solved, or recombined from the solutions of its

### 2014-12-19 doc/GMP2015/s03_drp.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.

@@
 \begin{theorem}\label{theorem:main}
-Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children according to the following rules.
+Given an isostatic 2D bar-joint graph $G$ and $\comdrp{G}$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children according to the following rules.
 \begin{enumerate}
@@
 \textsl{[DR-Plan for self-similar structure]}
-This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
+This section details the canonical decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
 The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$.

### 2014-12-19 doc/GMP2015/s05_problems.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.

@@
 OMD problem (when $k$ is part of the input)
-in Section \ref{sec:conclusion}.
+in Section \ref{sec:futurework}.
 
@@
 small DR-plan.
-\begin{observation}
+\begin{observation}\label{obs:OC_to_OMD}
     The (decision version of) the {\bf optimal completion problem

### 2014-12-19 doc/GMP2015/s03_drp.tex
Meera's change to problempinning. Small change to DRP

@@
 \begin{remark}
-    \todo{this captures cluster minimality}
+    \todo{Someone} \uncited discuss cluster minimality as a desirable property for children in a DR-plan.
+The canonical DR-plan in this paper intrinsically satisfies this property.
+If the children were not cluster minimal, then the union of children that form a larger cluster is itself isostatic and the original children were not vertex-maximal.
 \end{remark}

### 2014-12-19 doc/GMP2015/s07_problempinning.tex
Meera's change to problempinning. Small change to DRP

@@
 
-We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}.
-Specifically, we will look into the problem when you have a mono-layer and it is under-constrained.
-Our goal is to make such a layer isostatic.
+We can use qusecs DR-plans to design materials such as disordered
+graphene and silica bi-layers \cite{silica_bilayers}
+\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
+a somewhat more general setting:
+the problem of finding boundary conditions (additional constraints) to
+add to an underconstrained monolayer to make it
+isostatic.
 This can be done in a number of ways.
@@
     \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
-    \caption{Example of a single mono-layer of a Silica Oxygen glassy structure.
-This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
-Not shown here are other layers stacked on this one.
-Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.
-Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
+    \caption{Example of a single monolayer of a Silica (silicon and oxygen)
+    glassy structure. This can be thought of as a triangle-multipin qusec
+where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
+Not shown here are other monolayers stacked on this one. Each Silicon atom
+has the structure seen to the right, and so binds to another Oxygen atom in
+an adjacent monolayer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
     \label{fig:silica_glass}
@@
 \begin{enumerate}
-    \item Pin together 2 under-constrained mono-layers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
-    \item Pin the boundary in such a way that the resulting system is isostatic.
-    \item Build the structures by repeated subdivision to ensure they are isostatic at each layer (see Figure \ref{fig:subdivision})
+    \item Pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
+    \item Pin the boundary of (or in general, add constraints to) a layer
+        (possibly a genus 0 monolayer) so that it becomes isostatic
+    \item Design a broader class of structures to ensure they are isostatic,
+        self-similar (via some subdivision rule) and in addition isostatic at each
+        level of the subdivision (see Figure \ref{fig:subdivision})
 \end{enumerate}
 
-We will discuss item 2 in detail in this section.
-In each case, we are specifically interested in how to they can be applied to get a resulting isostatic structure with a small DR-plan, so a realization can be found quickly.
-To answer these questions, we first introduce the structures we model the problems with:
+In all cases,
+we are specifically interested in how to add additional constraints
+resulting isostatic structure with a small DR-plan, so a realization can be
+found and stress and other property design can be efficient.
+To answer these questions, we first introduce the
+qusecs that are used to model these materials.
+We will discuss Item 2 in detail in this section.
 
@@
     \end{subfigure}
-    \caption{Examples of self-similarity in the form of repeated subdivision.
-In (a), a simple subdivision is done that does not guarantee isostatic.
-In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.
-Credit to \cite{subdivision_paper} for (a).}
+    \caption{Examples of self-similarity via repeated subdivision. In (a),
+    simple subdivision that does not guarantee isostaticity. In (b), a more
+complicated scheme is used, ensuring that the resulting graph is isostatic
+but not  tree-decomposable. Credit to \cite{subdivision_paper} for (a).}
     \label{fig:subdivision}
@@
 
-\subsection{Body-Pin}
+\subsection{Body-Hyperpin qusecs}
 
 \begin{definition}
-    A \dfn{body-hyperpin qusecs} is a qusecs where the object are rigid bodies that are pinned together by some number of pins.
+    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies
+    some subsets of which are pinned together by pins.
 \end{definition}
@@
 \label{rem:bodypin_is_barjoint}
-    A body-hyperpin qusecs is a special case of bar-joint qusecs that we have been studying.
-As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
+    A body-hyperpin qusecs is a special case of bar-joint qusecs of the
+    first 5 sections. As such, the DR-planning for isostatics discussed in Section
+    \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
 \end{remark}
@@
 \begin{proof}
-    We can replace each body that has only one pin by a single vertex.
-A body with 2 pins can be replaced by an edge.
-In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
-When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
-The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged.
-The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+    We can replace each body that has only one pin by a single vertex. A
+    body with 2 pins can be replaced by an edge. In general, a body with $n$
+    pins can be replaced by a 2-tree on $n$ vertices. When finding a
+    DR-plan, we treat each body as trivial, so they become the leaves of our
+    plan. The optimal recombination problem and approach of Section
+    \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in
+Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
 
-For the rest of this section, we will only be dealing with the DR-plan of such qusecs.
-As such, we will refer only to the underlying graph of the qusecs.
-We now go over 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.
+For the rest of this section, we will only be dealing with the DR-plan of
+such qusecs. As such, we will refer only to the underlying hypergraph of the qusecs. We now
+introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and
+5 in Section \ref{sec:intro}, for which the optimal completion problem is
+significantly easier.
 
@@
     \end{itemize}
-    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
-Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
+    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
+    the bodies are the original bodies and each pin represents 2 bars from
+    one body to another. Such body-bar graphs with 1 and 2-dof can be
+    characterized by being $(3,4)$ and $(3,5)$-tight respectively
+    \cite{Lee:2007:PGA} \cite{streinu2009sparse},
+defined in Section  \ref{sec:preliminaries.}
 \end{definition}
@@
 \begin{definition}
-    A \dfn{multi-triangle-pin graph} is a body-hyperpin graph where each body is a triangle, meaning it is pinned in at most 3 places.
+    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
+    is a triangle, i.e., it shares at most 3pins with other bodies.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
+For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
 \end{definition}
 
-Body-pin graphs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
-Multi-triangle-pin graphs can be used to represent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (see Figure \ref{fig:silica_glass}).
-Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan:
+Body-pin graphs are of particular interest to us in the context of Example 4
+in Section \ref{sec:intro}. Triangle-multipin graphs can be used to
+represent the silica bi-layers and glassy structures described in Example 5
+of Section \ref{sec:intro}, where each triangle is the junction of ``disks''
+in the plane (see Figure \ref{fig:silica_glass}). Typically, these systems
+are not isostatic, so to relate the work of this paper to the systems, we
+define a slightly different kind of DR-plan using the notion of
+$(k,l)$-sparsity and tightness.
 
@@
 
-We can develop the notion of a canonical DR-plan for certain $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}.
+We can develop the notion of a canonical DR-plan when the hypergraph is
+$(k,l)$-sparse (i.e., combinatorially independent) using certain
+notion of trivial and non-trivial intersections and
+$(k,l)$-tightness conditions similar to the way we did in Section
+\ref{sec:DRP}, provided
+    such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
+    In particular, we can define canonical DR-plans with similar properties
+    for the 1 and 2 dof body-pin and triangle-hyperpin systems defined
+    above.
 
@@
 \label{obs:bodypin_drp}
-    For the 1-dof body-hyperpin graphs described above, a DR-plan node will follow one of the following:
+    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity
+    corresponds to canonical DR-plan node
+    satisfies one of the following:
 
     \begin{itemize}
-        \item Its children will be 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph
-        \item Its children will be a number of proper maximal 1-dof sub-graphs, joined pairwise by at most one pin
+        \item Its children  are 2 proper vertex maximal 1-dof graphs that intersect on
+            another 1-dof graph
+        \item Its children are a number of proper maximal 1-dof sub-graphs,
+            sharing at most one body
     \end{itemize}
@@
 
+The strong Church-Rosser property holds making all canonical DR-plans
+optimal:
+%
 \begin{remark}\label{rem:1dofcanon}
-    All $(k,l)$-tight canonical DR-plans are optimal.
-We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar and joint graphs discussed in Section \ref{sec:DRP}.
-This is due to the fact that such $(k,l)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
+    When the input is independent,
+    all $(3,4)$-tight canonical DR-plans are optimal. We can find such a
+    DR-plan in the same time complexity as the $(2,3)$-tight case for bar
+    and joint graphs discussed in Section \ref{sec:DRP}.
+    This is because
+    such $(3,4)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
 \end{remark}
@@
 \end{figure*}
-
+The above discussion leads to the main theorem:
 \begin{theorem}
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
+    Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding
+    1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal
+    completion problem of Section \ref{sec:recomb}.
 \end{theorem}
@@
 \label{obs:algebraic_completion}
-    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan with minimum fanin, we did not consider the algebraic complexity of realizing the DR-plan.
-If that is preferred, we can no longer only consider the deepest nodes.
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
+    the isostatic completion with minimum fanin, (a reasonable
+    measure of algebraic complexity), a more nuanced
+    measure that treats solutions of 1-dof and 2-dof systems as
+    1- or 2- parameter families would no longer permit the algorithm given
+    in the proof. In particular the complexity of the standard algorithm in the k-dof
+    case would be exponential in $k$ (even if that case were matroidal and
+    an optimal DR-plan is known).
 \end{observation}

### 2014-12-19 doc/GMP2015/appendix_a.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

@@
 
-    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.
+    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.
 
@@
 
-    \item 1 case: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible.
+    \item 1 case: \usestwod $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible.
 Since $C_i\cup C_k\neq C$ (and $C_j\cup C_k\neq C$), we know by the same logic as the previous case that the $C_i\cap C_k$ must be trivial (a single node).
@@
 
-    \item 2 cases: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible.
+    \item 2 cases: \usestwod $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible.
 The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.

### 2014-12-19 doc/GMP2015/s01_introduction.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

@@
   recombining the realizations or solutions of child subsystems into a
-  solution of the parent system. Specifically, we define the problem
+  solution of the parent system
+\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
+  Specifically, we define the problem
   of minimally modifying the indecomposable recombination system so
@@
   formal connection to well known problems such as optimal completion
-  of under-constrained systems \uncited. When the required
+  of under-constrained systems
+\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
+  and to find paths within the connected components. When the
   modifications are bounded, we obtain new, efficient algorithms for
   realizing both isostatic and under-constrained qusecs by leveraging
-  recent results about Cayley parameters in \uncited\uncited (Section
+  recent results about Cayley parameters in
+\cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII}
+  (Section
   \ref{sec:recomb}).
@@
   stresses in self-similar, bar-joint systems); Example 3 (canonical
-  and optimal DR-plans for pinned line incidence systems \uncited)
+  and optimal DR-plans for pinned line incidence systems
+  \cite{sitharam2014incidence})
   and Examples 4,5 (boundary-conditions for achieving various desired
-  properties of body-hyperpin systems).
+  properties of body-hyperpin systems) \uncited.
 

### 2014-12-19 doc/GMP2015/s02_preliminaries.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

@@
     The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}), is defined as a forest that has the
+    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
     following properties:
@@
         children as well.
+        We denote a complete DR-plan of $G$ as {\it
+        ComDRP}$(G)$
     A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
     The maximum  fan-in is called the \dfn{size} of the DR-plan.
+    We denote an optimal DR-plan of $G$ as {\it OptDRP}$(G)$.
+
 %
@@
     For a given graph, there could be exponentially many  DR-plans and
-    even optimal DR-plans in the size of the graph . A complete DR-plan
+    even optimal DR-plans in the size of the graph. A complete DR-plan
     is unique but may not be (is usually not) optimal. DR-plans

### 2014-12-19 doc/GMP2015/s03_drp.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

--- 2014-12-19 doc/GMP2015/s03_drp.tex
@@
 \label{sec:DRP}
-
-\subsection{Theory}
-
 % In the problem of the optimal DR-Plan there is generally not a unique plan.
@@
 
-In the problem of the optimal DR-Plan there is generally not a unique plan.
-Furthermore, there is no obvious way to find the optimal plan.
-In this section we present a particular way to construct a plan that guarantees optimality.
-We call this the \dfn{canonical} plan.
+The NP-hardness of the DR-planning problem for 2D bar-joint graphs is partly
+the consequence of possibly exponential number of DR-plans. On
+the other hand, although the complete DR-plan is unique
+it could have large average fan-in and exponentially many nodes making it
+far from optimal.
+
+In this section we define a \dfn{canonical} plan to capture those aspects of an optimal
+DR-plan that mimic a complete DR-plan's uniqueness, while for independent
+(underconstrained or isostatic graphs), we show that the nonunique parts do not affect
+optimality.
 
@@
 
-The importance of this formulation is stated succinctly in the following theorem.
+This definition gives a canonical DR-plan  a surprisingly strong Church-Rosser property.
 
-\begin{theorem}\label{theorem:canonical_is_optimal}
-    A canonical DR-plan is an optimal DR-plan.
+\begin{theorem} \label{theorem:canonical_is_optimal}
+    \label{theorem:main}
+    Any canonical DR-plan is an optimal DR-plan.
 \end{theorem}
 
-The proof of this theorem is a direct result of this more general theorem.
+%The proof of this theorem is a direct consequence of the following  more general theorem.
 
-\begin{theorem}\label{theorem:main}
-Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children according to the following rules.
+%\begin{theorem}\label{theorem:main}
+%Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$
+%with children $C_1,\ldots,C_N$ preserve children according to the following rules.
+%\begin{enumerate}
+%    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
+%    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
+%\end{enumerate}
+%The resulting canonical DR-plan is an optimal DR-plan.
+%\end{theorem}
+
+This theorem (proven in the Appendix) relies on the following crucial observation and lemma
+that are also used for other types of qusecs in the Application sections of
+the paper.
+
+\begin{observation}\label{lemma:union_intersection}
+If $F_i$ and $F_j$ are isostatic subgraphs of the same
+isostatic graph $F$, \
+then the following hold:
+(1) $F_i\cup F_j$ is not trivial;
+(2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial;
+(3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and
+(4) $F_i\cap F_j$ is not under-constrained.
+\end{observation}
+
+Then the following key properties hold at the nodes of a canonical DR-plan.
+
+\begin{lemma}
+Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
+Then
 \begin{enumerate}
-    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
-    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
+    \item
+    \label{lemma:wc_intersection_is_C}
+$C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.
+\item
+    \label{lemma:wc_intersection_makes_all_wc}
+If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic
+Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
+\item
+    \label{lemma:uc_intersection_makes_all_uc}
+If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
 \end{enumerate}
-This covers all cases and the resulting plan is an optimal DR-plan.
-\end{theorem}
+\end{lemma}
 
 \begin{remark}
-    \todo{this captures cluster minimality}
+The first item in the above lemma generalizes to the union of any number of
+children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{Cluster
+Minimality},
+defined in \cite{hoffman2001decompositionI}
+holding for these canonical-optimal DR-plans.
 \end{remark}
-
-This theorem (proven in the Appendix) relies on the following crucial observations and lemmas.
-
-\begin{observation}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
-\end{observation}
-
-\begin{lemma}\label{lemma:wc_intersection_is_C}
-$C_i\cup C_j$ is well-constrained if and only if $C_i\cup C_j = C$.
-\end{lemma}
-
-\begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
-If $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
-Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
-\end{lemma}
-
-\begin{lemma}\label{lemma:uc_intersection_makes_all_uc}
-If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
-\end{lemma}
 

### 2014-12-19 doc/GMP2015/appendix_b.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

--- 2014-12-19 doc/GMP2015/s09_conclusion.tex
@@
-\section{Conclusion}
-\bibliography{paper}
-
-We have elucidated the complexity barriers, for example overconstraints, 
-faced by the general optial DR-plan and recombination problem. 
-When there are no overconstraints, which is the case for 2D qusecs whose
-realizations are many common types of layered
-materials, we defined a canonical DR-plan and showed that such canonical
-DR-plans are guaranteed to be optimal, a strong Church-Rosser property given
-an effiecient ($O(n^2)$) algorithm to find a DR-plan that satisfies other
-desirable characteristics.
-
-We have also described novel methods of efficiently realizing a 2D qusec 
-from the optimal DR-plan by modifying the otherwise nondecomposable 
-system that recombines the solved child systems at a node of a DR-plan
-These results rely on a recent theory of Convex Cayley Configuration spaces.
-
-We then modeled specific layered materials
-using extensions of the above theoretical results
-including the motivating Examples 1-5 in the introduction.
+\section{Appendix B}
+\label{sec:appendix:b}
 
 \subsection{Furture Work}
+\label{sec:futurework}
 A few natural  open questions are the following.
@@
     completion problem would  have complexity exponential in $k$.
-
 \end{openproblem}

### 2014-12-19 doc/GMP2015/s02_preliminaries.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

@@
 
-In this paper, we will consider 3 types of constraint systems: bar-
-joint as well as \dfn{pinned line-incidence}, and \dfn{body-hyperpin}.
-In all of these cases, a cartesian \dfn{realization} or \dfn{solution}
+In this paper, we will consider 3 types of constraint systems:
+bar-joint as well as \dfn{pinned line-incidence} and \dfn{body-hyperpin}.
+In all of these cases, a Cartesian \dfn{realization} or \dfn{solution}
 $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean
@@
 but a measure zero set of $\delta$ values). Many of these are
-properties of the jacobian $\Delta_x G(x,\delta)$, often called the
+properties of the Jacobian $\Delta_x G(x,\delta)$, often called the
 appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
 For example, the \dfn{bar-joint rigidity matrix of the graph $G =
-(V,E)$} is a matrix of indeterminates representing the jacobian of the
+(V,E)$} is a matrix of indeterminates representing the Jacobian of the
 distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$
@@
 %which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is
+It is \dfn{isostatic (minimally rigid, well-constrained)} if it is
 both rigid and independent.
@@
 %The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
-is independent and not rigid, or \dfn{overconstrained} if it is not
+It is \dfn{flexible} if it is not rigid, \dfn{under-constrained} if it
+is independent and not rigid, or \dfn{over-constrained} if it is not
 independent.
@@
 
-In this paper, a \dfn{qusec} is any \dfn{independent} geometric
+In this paper, a \dfn{qusecs} is any \dfn{independent} geometric
 constraint system of one of the 3 types mentioned above.
@@
         children as well.
-        We denote a complete DR-plan of $G$ as {\it
-        ComDRP}$(G)$
+        We denote a complete DR-plan of $G$ as $\comdrp{G}$
     A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
     The maximum  fan-in is called the \dfn{size} of the DR-plan.
-    We denote an optimal DR-plan of $G$ as {\it OptDRP}$(G)$.
+    We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
 
@@
 \noindent
-\note In Section \ref{sec:DRP} $G$ is assumed to be isostatic.
+\note In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.
 
@@
 The use of a DR-plan in designing especially self-similar qusecs is
-immediately evident, as shown in Figures \ref{fig:c2c3ofk33s} and
+immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and
 \ref{fig:subdivision}.
@@
 subsystem that is solved, or recombined from the solutions of its
-child subsystems i.e.,the maximum fan-in occuring in a DR-plan. Thus
+child subsystems i.e.,the maximum fan-in occurring in a DR-plan. Thus
 finding an optimal DR-plan is important for realization.
@@
 rigidity matrix of an isostatic subgraph. The blocks intersect on
-isostatic or trival subgraphs. This recursive block decomposition of
+isostatic or trivial subgraphs. This recursive block decomposition of
 the rigidity matrix $R$ correspondingly recursively decomposes

### 2014-12-19 doc/GMP2015/s03_drp.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

@@
     The \dfn{canonical DR-plan} is a DR-plan that satisfies the following two properties:
-    \begin{enumerate}
-        \item Children are isostatic vertex-maximal subgraphs of the parent.
-        \item All of the isostatic vertex-maximal subgraphs are children if any two of them intersect trivially, otherwise exactly two are children.
-    \end{enumerate}
+    (1) children are isostatic vertex-maximal subgraphs of the parent; and
+    (2) all of the isostatic vertex-maximal subgraphs are children if any two of them intersect trivially, otherwise exactly two are children.
 \end{definition}
 
-This definition gives a canonical DR-plan  a surprisingly strong Church-Rosser property.
+This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
 
@@
 This theorem (proven in the Appendix) relies on the following crucial observation and lemma
-that are also used for other types of qusecs in the Application sections of
+that are also used for other types of qusecs in the Application sections (\ref{sec:bodypin}, \ref{sec:pinnedline}) of
 the paper.

### 2014-12-19 doc/GMP2015/s04_recombination.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

@@
 of combining specific solutions of subsystems in a DR-plan into a
-solution of their parent system $I$, (without loss of generality, at
+solution of their parent system $I$ (without loss of generality, at
 the top level of the DR-plan). In the case of isostatic qusecs, the
@@
 \dfn{solution(s) (among a finite large number of solutions) with a
-specific orientation or chirality}. In the case of underconstrained
+specific orientation or chirality}. In the case of under-constrained
 qusecs the subsystems are the multiple roots of the DR-plan, the
-parent system $I$ is underconstrained, and we typically seek an
+parent system $I$ is under-constrained, and we typically seek an
 efficient algorithmic description of \dfn{connected component(s) of
@@
 smallest possible size or max fan-in),  is that the number of children
-(resp. number of roots) - and correspondingly the  size and complexity
-of the (indecomposable) algebraic system $I$ to be solved - could be
+(resp. number of roots) --- and correspondingly the  size and complexity
+of the (indecomposable) algebraic system $I$ to be solved --- could be
 arbitrarily large as a function of the size of the input system. This
@@
 \dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
-$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters
+$I_{\lambda_F}$ ---  one for each value $\lambda_F$ for the parameters
 $F$,  all of which have small DR-plans. Thus, given a value $v$ for
@@
 Given a generically independent graph $G = (V,E)$ with no non-trivial
-proper isostatic subgraph (indecomposable), and 2 constants $k$ and
+proper isostatic subgraph (indecomposable) and 2 constants $k$ and
 $s$, does there exist a set of at most $k$  edges $E_1$ and a set of
-nonedges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
+non-edges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
 F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
@@
 Next we provide the necessary background to describe a specific
-approach for achieving the requirements (a)-(c) mentioned above, by
+approach for achieving the requirements (a)--(c) mentioned above, by
 restricting the class of reduced graphs $G' = G\setminus E_1$ and
@@
 problem, and using a key theorem of Convex Cayley configuration spaces
-\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-
-edges $F$ (Cayley parameters), such that the set of vectors
-$\lambda_F$ of  attainable lengths of the nonedges $F$ is always
+\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$
+and non-edges $F$ (Cayley parameters), such that the set of vectors
+$\lambda_F$ of  attainable lengths of the non-edges $F$ is always
 convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
-all the realization s of the bar-joint constraint system or linkage
+all the realizations of the bar-joint constraint system or linkage
 $(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
@@
 Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
-a ``projection'' of the cartesian realization space of $(H,\delta)$ on
+a ``projection'' of the Cartesian realization space of $(H,\delta)$ on
 the Cayley parameters $F$. Such graphs $H$ are said to have
 \dfn{convexifiable Cayley configuration spaces with parameters $F$}
-(short: \dfn{convexifiable}). To state the theorem, we first have to
-define the notion of \dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and
-$H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with
-edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
-respectively.  A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
+(short: \dfn{convexifiable}).
+
+To state the theorem, we first have to
+define the notion of \dfn{2-sums} and \dfn{2-trees}.
+Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.
+% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
+A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
 obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
@@
 that cannot be further split into 2-sum components. A \dfn{2-tree} is
-recursively obtained by taking a 2-sum of two 2-trees, with the base
-case of 2-tree being a triangle. A \dfn{partial 2-tree} is a 2-tree
-minus some edges. Partial 2-trees are characterized as having $K_4$ as
+recursively obtained by taking a 2-sum of 2-trees, with the base
+case being a triangle. A \dfn{partial 2-tree} is a 2-tree
+minus some edges. Partial 2-trees characterized as having $K_4$ as a
 forbidden minor are also called series parallel graphs.
@@
 
-The idea of our approach to achieve the criteria (a)-(c) begins with
+The idea of our approach to achieve the criteria (a)--(c) begins with
 the following simple but useful theorem.
@@
 such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
-for small enough $k$. Criterion $c$ is addressed as described in the
+for small enough $k$. Criterion (c) is addressed as described in the
 following efficient search procedure which clarifies the dependence of
@@
     (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
-    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of nonedges
+    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of non-edges
     of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
@@
 \ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
-nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
+non-edge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
 configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
@@
 \Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
-$(G,\delta)$ is found when the length of the nonedge $e$ in  the
+$(G,\delta)$ is found when the length of the non-edge $e$ in  the
 realization $q^p_{\lambda_f}$ is $\delta_e$.
@@
 $G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
-Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
+Criterion (c) is met --- for example when $k=1$ --- because 1-dof Cayley
 configuration spaces of linkages based on such graphs $G'$ are known

### 2014-12-19 doc/GMP2015/s05_problems.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

@@
 have not only small, but also special DR-plans that permit easy
-solving - essentially a single quadratic at a time.
+solving --- essentially a single quadratic at a time.
 
@@
 \begin{observation}\label{obs:OC_to_OMD}
-    The (decision version of) the {\bf optimal completion problem
+    The (decision version of) the \textbf{optimal completion problem
     (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
@@
 
-In order to connect the OR problem to OMD, when the input graph for is
+In order to connect the OR problem to OMD, when the input graph is
 the isostatic graph at the DR-plan root, we do not consider the case
@@
 from bar-joint systems to
-so-called \dfn{multi-body pin} define in Section \ref{sec:bodypin}
-by showing that the corresponding problems for the latter are reduced
+so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin},
+by showing that the problems for the latter are reduced
 to the corresponding problems on bar-joint systems.

### 2014-12-19 doc/GMP2015/s07_problempinning.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

@@
 % \bibliography{paper}
-
-We can use qusecs DR-plans to design materials such as disordered
-graphene and silica bi-layers \cite{silica_bilayers}
-\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
-a somewhat more general setting:
-the problem of finding boundary conditions (additional constraints) to
-add to an underconstrained monolayer to make it
-isostatic.
-This can be done in a number of ways.
 
@@
 
-\begin{enumerate}
-    \item Pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
-    \item Pin the boundary of (or in general, add constraints to) a layer
-        (possibly a genus 0 monolayer) so that it becomes isostatic
-    \item Design a broader class of structures to ensure they are isostatic,
-        self-similar (via some subdivision rule) and in addition isostatic at each
-        level of the subdivision (see Figure \ref{fig:subdivision})
-\end{enumerate}
+We can use qusecs DR-plans to design materials such as disordered
+graphene and silica bi-layers \cite{silica_bilayers}
+\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
+a somewhat more general setting:
+the problem of finding boundary conditions (additional constraints) to
+add to an underconstrained monolayer to make it
+isostatic.
+This can be done in a number of ways:
+(1) pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass});
+(2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or
+(3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).
 
@@
 \begin{definition}
-    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies
-    some subsets of which are pinned together by pins.
+    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies some subsets of which are pinned together by pins.
 \end{definition}
@@
     A body-pin graph is a body-hyperpin graph with the following conditions:
-    \begin{itemize}
-        \item Each pin is shared by at most 2 bodies
-        \item No 2 bodies share more than one pin
-    \end{itemize}
+    (1) each pin is shared by at most two bodies; and
+    (2) no two bodies share more than one pin
+
     Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
@@
     \cite{Lee:2007:PGA} \cite{streinu2009sparse},
-defined in Section  \ref{sec:preliminaries.}
+defined in Section  \ref{sec:prelim}.
 \end{definition}
@@
     A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
-    is a triangle, i.e., it shares at most 3pins with other bodies.
+    is a triangle, i.e., it shares at most 3 pins with other bodies.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
@@
     In particular, we can define canonical DR-plans with similar properties
-    for the 1 and 2 dof body-pin and triangle-hyperpin systems defined
+    for the 1 and 2-dof body-pin and triangle-hyperpin systems defined
     above.
@@
 \label{obs:bodypin_drp}
-    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity
-    corresponds to canonical DR-plan node
-    satisfies one of the following:
-
-    \begin{itemize}
-        \item Its children  are 2 proper vertex maximal 1-dof graphs that intersect on
-            another 1-dof graph
-        \item Its children are a number of proper maximal 1-dof sub-graphs,
-            sharing at most one body
-    \end{itemize}
+    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity corresponds to a canonical DR-plan whose nodes satisfy one of the following, either
+    (a) the children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph, or
+    (b) the children are a number of proper maximal 1-dof sub-graphs, sharing at most one body.
 \end{observation}
@@
     While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
-    the isostatic completion with minimum fanin, (a reasonable
+    the isostatic completion with minimum fan-in, (a reasonable
     measure of algebraic complexity), a more nuanced

### 2014-12-19 doc/GMP2015/s09_conclusion.tex
Fixed some references and citations. Unified display of OptDRP, ComDRP, CanDRP, use \drp{G}, \optdrp{G}, etc.
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!

--- 2014-12-18 doc/GMP2015/s09_conclusion.tex
@@
 \section{Conclusion}
-We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem.
-To this end we have defined a new canonical DR-plan and given an effiecient ($O(n^2)$) algorithm to find the DR-plan for specific types of 2D qusecs.
+\label{sec:conclusion}
+% \bibliography{paper}
 
-We have also described methods of recombining the DR-plan to get a realization different from the normal solving of a system of equations.
-These results rely on the results obtained in the Cayley Configuration space of certain graphs.
+We have elucidated the complexity barriers, for example over-constraints, faced by the general optimal DR-plan and recombination problem.
+When there are no overconstraints, which is the case for 2D qusecs whose realizations are many common types of layered materials, we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property given an efficient ($O(n^2)$) algorithm to find a DR-plan that satisfies other desirable characteristics.
 
-We then related these theoretical results to real-world situations in which they can be applied including Examples 1-5 in Section \ref{sec:intro}.
+We have also described novel methods of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise non-decomposable system that recombines the solved child systems at a node of a DR-plan.
+These results rely on a recent theory of Convex Cayley Configuration spaces.
 
-Throughout the paper, we raised some open questions and future areas of research.
-These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}
+We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.
+
+In Appendix B [\ref{sec:appendix:b}] we provide open problems for future work.

### 2014-12-20 doc/GMP2015/appendix_a.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!
Discussion of algorithm refined. Added some Meera stuff.

@@
-\section{Appendix}
-\label{sec:appendix}
+\section{Appendix A}
+\label{sec:appendix:a}
 
@@
 \begin{proof}
-Remark \ref{lemma:union_intersection} shows that the intersection of any two well-constrained subgraphs (such as $C_i$ and $C_j$) can only result in trivial or well-constrained subgraphs.
+Remark \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs.
 Therefore, these are the only possibilities to consider.
@@
 
-Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained.
+Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic.
 If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
@@
 By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
-Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$.
-Thus, the assumption is wrong and $U$ cannot be well-constrained.
-As $C$ is well-constrained, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.
+Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$.
+Thus, the assumption is wrong and $U$ cannot be isostatic.
+As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.
 
 \medskip\noindent
-\vemph{Part 2:} If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained.
+\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
 This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
@@
 
-For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$.
-For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained.
+For (2), observe that under-constrained subgraphs of isostatic graphs must have density less than $k$.
+For (3), observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic.
 Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$.
@@
 
-For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
+For (4), because subgraphs of a isostatic graph can only be trivial, under-, or isostatic, all cases have already been exhausted.
 \end{proof}
@@
 
-%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
+%     % \item Observe that, given a isostatic graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
 
-%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.
+%     \item Observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic. Then it is straightforward application of the equation above.
 
-%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$.
+%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of isostatic $F$, it must be that $d(F_i\cup F_j)=l<k$.
 Therefore $d(F_i\cap F_j)=2k-l>k$.
@@
 %     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$.
-Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained.
-\textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.
+Being a subgraph of isostatic $F$, $F_i\cap F_j$ is also isostatic.
+\textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is isostatic.
 
-%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
+%     \item Subgraphs of a isostatic graph can only be trivial, under-, or isostatic. All cases are already exhausted.
 % \end{enumerate}
@@
 
-\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_is_C}}
+\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}}
 
@@
 %
-In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem).
-Thus, $C_i\cup C_j=C$ is well-constrained.
+In the reverse direction, we know $C$ is either some non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
 \end{proof}
 
-\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc}}
+\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}}
 
-\begin{corollary}\label{corollary:no_edges_between_diff}
-Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
-\end{corollary}
+\begin{observation}\label{observation:no_edges_between_diff}
+Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
+\end{observation}
 
@@
 
-\begin{proof} (Of Lemma \ref{lemma:wc_intersection_makes_all_wc})
+\begin{proof} (Of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc})
 
@@
 
-Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$).
-There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
+Then we assume that there is some third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
 
@@
 
-    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.
+    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Observation \ref{observation:no_edges_between_diff}) and cannot be isostatic.
 
     \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible.
-Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be well-constrained.
-We also know it cannot be trivial because it contains well-constrained subgraphs.
+Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be isostatic.
+We also know it cannot be trivial because it contains isostatic subgraphs.
 This means it must be under-constrained.
 From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial.
-This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.
+This is impossible because $D_{i,j}$ is isostatic, thereby contradicting the assumption that $C_k$ is isostatic.
 
@@
 
-\subsubsection{Proof of Lemma \ref{lemma:uc_intersection_makes_all_uc}}
+\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:uc_intersection_makes_all_uc}}
 
@@
 Assume there is some $k$ such that $C_i\cap C_k$ is not trivial.
-By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be well-constrained.
-Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be well-constrained.
-This means that $C_i\cap C_j$ is well-constrained.
+By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic.
+Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic.
+This means that $C_i\cap C_j$ is isostatic.
 Therefore, such a $k$ cannot exist and all intersections are trivial.
 \end{proof}
+
+\subsubsection{Proof of Theorem \ref{theorem:algo_complexity}}
+
+\begin{proof}
+The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
+Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
+The output of this will be the list of component-candidates.
+Then run the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} \todo{complexity?} on each candidate with edge $e$, which will find the minimal subgraph containing the candidate and $e$, which we call $D$.
+If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
+If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list.
+Then, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
+Otherwise, run to completion and output all items in the true-component list.
+
+This is run recursively on each node.
+If you use dynamic programming to store the decompositions of each component, it prevents repetition of work.
+In this case you will have to do work on no more than $O(V)$ nodes that get smaller at each level.
+
+Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
+\end{proof}
+
+% \begin{proof}
+% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
+There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
+After the subgraphs have been found, it is a straightforward application of the theory.
+If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
+Otherwise, all the subgraphs are the children.
+Apply recursively to the children.
+This naturally terminates when you reach trivial subgraphs.
+The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
+\todo{Think more about complexity.}
+% \end{proof}
 
@@
 
-First notice that since $C_i \cup C_j$ is well-constrained, by Observation~\ref{lem:pinned_union_intersection},
-both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.
-Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a well-constrained subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected well-constrained vertex-maximal proper subgraph of  $C$.
+First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection},
+both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
+Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a isostatic subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected isostatic vertex-maximal proper subgraph of  $C$.
 
@@
     \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be well-constrained.
+Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be isostatic.
     By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
-    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are well-constrained.
+    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.
 

### 2014-12-20 doc/GMP2015/appendix_b.tex
Discussion of algorithm refined. Added some Meera stuff.

@@
 \end{openproblem}
+
+\subsection{Open Problems}
+
+\todo{added these in newest conclusion}
+
+A few natural  open questions concern the following common theme
+that runs through the paper:
+
+\begin{openproblem}
+For fixed $k$ we have polynomial time optimal DR-planning,
+recombination (modification) in the presence of $k$ overconstraints (Section
+\ref{sec:DRP},
+optimal modification for decomposition {\sl OMD}$_k(G)$,
+when at most $k$ constraints are removed
+\ref{sec:recomb}
+and
+also optimal completion
+using at most $k\le 2$ constraints in the body-pin and
+triangle-multipin cases
+under a modified measure for optimizing the
+DR-plan \ref{sec:table}.
+However, $k$ appears in the exponent for
+the straightforward algorithms for these problems.
+\end{openproblem}
+
+One problem in the above theme is from
+Section \ref{sec:bodypin}.
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when the given
+    graph has more than 2-dofs?
+    Our proof for the 1 and 2-dof cases relied heavily on the matroidal
+    properties of their corresponding $(k,l)$-tightness.
+    For higher number of dofs, the $(k,l)$ characterization is
+    no longer matroidal \cite{Lee:2007:PGA}.
+    As a result, the major obstacle is that
+    there is no easy way of obtaining an optimal or canonical
+    $k$-dof DR-plan in general.
+    Even assuming such  a DR-plan is available,
+    if higher dofs had the same characteristics,
+    Observation \ref{obs:algebraic_completion}
+    raises questions about the correct measure of DR-plan size that
+    captures algebraic complexity for recombining graphs with many dofs
+    (this is not an issue in the isostatic case). Unless some restrictions
+    can be found and taken advantage of, the $k$-dof optimal
+    completion problem would  have complexity exponential in $k$.
+\end{openproblem}
+
+Another problem from the above theme is from Section \ref{sec:recomb}
+\begin{openproblem}
+What is the complexity of the restricted OMD (optimal modification for
+decomposition) problem?
+This has the potential to be difficult. For example, when the
+isostatic completion is required to be a 2-tree the restricted OMD
+problem is reducible to the maximum spanning series-parallel subgraph
+problem shown by \cite{cai1993spanning} to be NP-complete even if the input
+graph is planar of maximum degree at most 6. However, since the OMD
+problem has other input restrictions such as not having any proper
+isostatic subgraphs, it is not clear if the reverse reduction exists
+and hence it is unclear whether the OMD problem is NP-complete.
+
+The same holds for the restricted OMD problem where the isostatic
+completion is required to be a tree-decomposable graph of low
+Cayley complexity (i.e, have special, small DR-plans). One potential
+obstacle to an indecomposable graph $G$'s membership in the restricted
+OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
+have been constructed in \cite{servatius2000rigidity}.
+\end{openproblem}
+
+The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in
+Section \ref{sec:table}.
+
+\begin{openproblem}
+    Is the OMD (optimal modification for decomposition) problem
+    reducible to the OC (Optimal completion) problem?
+\end{openproblem}
+
+More general problem directions are the following.
+\begin{openproblem}
+    Combinatorial rigidity for
+    periodic structures is an active area of research. This paper motivates
+    a study of the rigidity of self-similar structures,
+    with self-similar groups replacing periodic groups.
+\end{openproblem}
+
+\begin{openproblem}
+    The pinned line incidence structures of Section \ref{sec:pinnedline},
+    for example in the case of collagen microfibrils, whose function is
+    elastic contraction, should be considered
+    congruent under projective transformations. I.e, the projective group
+    should be factored out as a trivial motion in a new project for
+    extending the combinatorial rigidity characterization of such systems.
+    (Currently we permit no trivial motions at all).
+\end{openproblem}

### 2014-12-20 doc/GMP2015/s00_abstract.tex
White space fixes.
Discussion of algorithm refined. Added some Meera stuff.

@@
 \begin{abstract}
-A variety of quasi-uniform (aperiodic) and possibly self-similar,
+A variety of quasi-uniform (aperiodic) and self-similar,
 layered material structures are a natural consequence of design
@@
 request and publicly available for the final version.
+\end{abstract}
 
@@
 % % Software implementation and videos can be found at: \todo{Provide Link}.
-\end{abstract}

### 2014-12-20 doc/GMP2015/s01_introduction.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!
Discussion of algorithm refined. Added some Meera stuff.

@@
 include:
+%
+\begin{enumerate}
+    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.
 
-\begin{enumerate}
-    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}
-
-    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited
+    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}).
 
@@
   of the NP-hardness of finding optimal DR-plans for over-constrained
-  systems.
+  systems. \todo{wording}
+  Our optimal/canonical DR-plan satisfies desirable properties such as cluster minimality previously studied \cite{hoffman2001decompositionI}.
 
@@
 \end{itemize}
+
+\note
+All proofs are in Appendix A (\ref{sec:appendix:a}).

### 2014-12-20 doc/GMP2015/s03_drp.tex
Discussion of algorithm refined. Added some Meera stuff.

@@
 
-\begin{lemma}
+\begin{lemma}\label{lemma:combined_lemma}
 Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
@@
 \begin{enumerate}
-    \item
-    \label{lemma:wc_intersection_is_C}
+    \item\label{lemma:wc_intersection_is_C}
     $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.
-\item
-    \label{lemma:wc_intersection_makes_all_wc}
-If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic
-Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
-\item
-    \label{lemma:uc_intersection_makes_all_uc}
+
+    \item\label{lemma:wc_intersection_makes_all_wc}
+    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
+
+    \item\label{lemma:uc_intersection_makes_all_uc}
     If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
@@
 
-\subsection{Algorithm}
+% \subsection{Algorithm}
 
-The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
-There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
-After the subgraphs have been found, it is a straightforward application of the theory.
-If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
-Otherwise, all the subgraphs are the children.
-Apply recursively to the children.
-This naturally terminates when you reach trivial subgraphs.
-The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
-\todo{Think more about complexity.}
+\begin{theorem}\label{theorem:algo_complexity}
+    There exists an $O(|V|^3)$ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
+\end{theorem}
+
+\begin{remark}
+    while the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
+However, the time complexity is exponential in $k$.
+\end{remark}

### 2014-12-20 doc/GMP2015/s09_conclusion.tex
Discussion of algorithm refined. Added some Meera stuff.

@@
 \section{Conclusion}
-\label{sec:conclusion}
 % \bibliography{paper}
 
-We have elucidated the complexity barriers, for example over-constraints, faced by the general optimal DR-plan and recombination problem.
-When there are no overconstraints, which is the case for 2D qusecs whose realizations are many common types of layered materials, we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property given an efficient ($O(n^2)$) algorithm to find a DR-plan that satisfies other desirable characteristics.
+We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
+For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient ($O(n^2)$) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.
 
-We have also described novel methods of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise non-decomposable system that recombines the solved child systems at a node of a DR-plan.
-These results rely on a recent theory of Convex Cayley Configuration spaces.
+We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan These results rely on a recent theory of Convex Cayley Configuration spaces.
+Relationships and reductions between these and previously studied problems were formally clarified.
 

### 2014-12-20 doc/GMP2015/appendix_a.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

@@
 \begin{proof}
-Remark \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs.
+Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs.
 Therefore, these are the only possibilities to consider.
@@
 \usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
+However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
 By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
@@
 \medskip\noindent
-\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
-This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
+\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
 Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
@@
 \begin{proof}
-Lemma \ref{lemma:wc_intersection_is_C} shows that $C_i\cup C_j$ must equal the entire parent $C$.
+Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the entire parent $C$.
 \end{proof}
@@
 
-The alternative phrasing is due to Lemma \ref{lemma:wc_intersection_is_C}.
+The alternative phrasing is due to Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}.
 
@@
     \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible.
-Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be isostatic.
+Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, that $C_i\cup C_k$ cannot be isostatic.
 We also know it cannot be trivial because it contains isostatic subgraphs.
 This means it must be under-constrained.
-From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial.
+From Observation \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial.
 This is impossible because $D_{i,j}$ is isostatic, thereby contradicting the assumption that $C_k$ is isostatic.
@@
 Assume there is some $k$ such that $C_i\cap C_k$ is not trivial.
-By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic.
-Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic.
+By Observation \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic.
+Then, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic.
 This means that $C_i\cap C_j$ is isostatic.
@@
 Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be the list of component-candidates.
-Then run the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} \todo{complexity?} on each candidate with edge $e$, which will find the minimal subgraph containing the candidate and $e$, which we call $D$.
+The output of this will be a list of component-candidates.
+Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
+If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list.
+If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list.
+Continue with the next candidate.
+When you exhaust the candidates, move to the next step.
+
+Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.
+
+If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$.
+The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
+We construct something that, while recombining it, results in the same systems to solve as the true DR-plan.
+Begin with the parent $C$.
+Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$.
+Recursively apply the algorithm to the child containing $R_1$.
+The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
+Etc.
+At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
+Apply the algorithm recursively to both children.
+% Although the nodes containing $R_1,\ldots,R_N$ are under-constrained, simply take the children to be the roots of the forest
+
+The reason we construct it this way is because we end up with a tree instead of a directed acyclic graph.
+The leaves of this tree will be exactly every single edge in the graph, with none repeated.
+Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.
+
+Therefore, the complexity of the algorithm as a whole is $O(|V|^3)$.
+%
+% \note In the best case, this can be $O(V^2)$ if the graph is a 2-tree
+\end{proof}
+
+% \begin{proof}
+% The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
+Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
+The output of this will be a list of component-candidates.
+Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
 If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
-If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list.
-Then, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
-Otherwise, run to completion and output all items in the true-component list.
+If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
+Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$.
+Continue with the next candidate.
+If you exhaust the candidates, output the entire true-component list.
 
-This is run recursively on each node.
-If you use dynamic programming to store the decompositions of each component, it prevents repetition of work.
-In this case you will have to do work on no more than $O(V)$ nodes that get smaller at each level.
+% %on the children
+% This is run recursively on each node.
+If you use dynamic programming to store the decompositions of each component, it prevents repetition of work (done by using the DAG representation of the DR-plan).
+In this case you will have to do work on no more than $O(|V|)$ nodes that get smaller at each level.
 
-Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
-\end{proof}
+% Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
+
+% It's a tree in only trivial case -> O(V), always order of leaves
+% If its isostatic intersection, we do the sequential extension,intersections only done once.
+
+% inductively. The number of nodes at this level will e more than the total. Number of leaves (terminals for dag) is at most the number of edges
+
+% \end{proof}
 
@@
 % \end{proof}
-
-\subsection{Proofs from Section \ref{sec:pinnedline}}
-\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
-\label{sec:appendix_pinned}
-
-\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
-We use the same notations as in the original proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} in the last section.
-
-First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection},
-both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
-Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a isostatic subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected isostatic vertex-maximal proper subgraph of  $C$.
-
-Then we analyze all the possible cases for $C_k$.
-\begin{itemize}
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be isostatic.
-    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
-    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.
-
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.
-
-    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
-    The proof is similar to the previous case.
-\end{itemize}
-
-All remaining cases are similar to the original proof.
-\end{proof}
 
@@
 In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
-When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
-The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged.
-The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+When finding a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
+The optimal recombination problem and approach of Section \ref{sec:recomb} are unchanged.
+The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
@@
 \begin{proof}
-    \todo{Item 1 Follows from paper?}
-
-    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
-Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
+    The existence of this canonical DR-plan follows from the same arguments as in the proof of Theorem \ref{theorem:main}.
+The only difference is the definition of a trivial intersection.
+In this case, when two subgraphs share more than 1 body, they become rigid (in fact over constrained).
+Sharing a pin is not considered an intersection.
+Such a structure is viewed as two subgraphs each sharing 1 body with a 3rd 1-dof subgraph which essentially just consists of those two bodies pinned together.
 \end{proof}
 
-\subsubsection{Proof of Observation \ref{thm:1dofcase}}
+\subsubsection{Proof of Remark \ref{thm:1dofcase}}
 \begin{proof}
@@
 
-    Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}.
+    Thus, say we have a $T$ as described in the proof for Remark \ref{thm:1dofcase}.
 When we pin a node $b$, we will have the same structure as before.
@@
 \end{proof}
+
+\subsection{Proofs from Section \ref{sec:pinnedline}}
+\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
+\label{sec:appendix_pinned}
+
+\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
+We use the same notations as in the original proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, in the last section.
+
+First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection},
+both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
+Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a isostatic subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected isostatic vertex-maximal proper subgraph of  $C$.
+
+Then we analyze all the possible cases for $C_k$.
+\begin{itemize}
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
+Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
+    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
+    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.
+
+    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.
+
+    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
+    The proof is similar to the previous case.
+\end{itemize}
+
+All remaining cases are similar to the original proof.
+\end{proof}

### 2014-12-20 doc/GMP2015/s00_abstract.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

@@
 hard even for (possibly overconstrained) 2D bar-joint constraint
-systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
+systems we describe an \candrpcomplexity algorithm  for qusecs. Besides
 elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-

### 2014-12-20 doc/GMP2015/s01_introduction.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

--- 2014-12-19 doc/GMP2015/s01_introduction.tex
@@
 include:
+%
+\begin{enumerate}
+    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.
 
-\begin{enumerate}
-    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}
+    \item \label{materialexample2} Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}).
 
-    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited
-
-    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+    \item \label{materialexample3} Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix.
 See Section \ref{sec:pinnedline}.
 
-    \item More recent, engineered examples, including disordered
+    \item \label{materialexample4} More recent, engineered examples, including disordered
     graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes
@@
 
-    \item Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
+    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
     materials that behave like assemblies of 2D particles under
@@
 \begin{itemize}
-  \item We navigate the NP-hardness barrier mentioned in (2) above,
+  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier mentioned in (2) above,
   for finding optimal DR-plans by defining a so-called \dfn{canonical}
@@
   canonical DR-plans for isostatic or under-constrained 2D qusecs are
-  optimal} (Section \ref{sec:DRP}).
+  optimal}.
 
-  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
+  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical
   (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
@@
   of the NP-hardness of finding optimal DR-plans for over-constrained
-  systems.
+  systems. Futhermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.
 
-  \item We give a method to deal with the algebraic complexity of
-  recombining the realizations or solutions of child subsystems into a
-  solution of the parent system
+  \item In Section \ref{sec:recomb}, we give a method to deal with the
+  algebraic complexity of recombining the realizations or solutions of
+  child subsystems into a solution of the parent system
   \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
-  Specifically, we define the problem
-  of minimally modifying the indecomposable recombination system so
-  that it becomes decomposable via a small DR-plan and yet preserves
-  the original solutions in an efficiently searchable manner. We show
-  formal connection to well known problems such as optimal completion
-  of under-constrained systems
+  Specifically, we define the problem of minimally modifying the
+  indecomposable recombination system so that it becomes decomposable
+  via a small DR-plan and yet preserves the original solutions in an
+  efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well
+  known problems such as optimal completion of under-constrained
+  systems
   \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
@@
   recent results about Cayley parameters in
-\cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII}
-  (Section
-  \ref{sec:recomb}).
+  \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
 
-  \item We briefly describe applications of the above techniques to
-  modeling, analyzing, and designing specific properties in 2D material
-  layers. For Examples 1,2 (achieving isostaticity, distribution of
-  stresses in self-similar, bar-joint systems); Example 3 (canonical
-  and optimal DR-plans for pinned line incidence systems
-  \cite{sitharam2014incidence})
-  and Examples 4,5 (boundary-conditions for achieving various desired
-  properties of body-hyperpin systems) \uncited.
+  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we
+  briefly describe applications of the above techniques to modeling,
+  analyzing, and designing specific properties in 2D material
+  layers~\cite{Jackson2008bodypin}. For Examples
+  \ref{materialexample1}, \ref{materialexample2} (achieving
+  isostaticity, distribution of stresses in self-similar, bar-joint
+  systems); Example \ref{materialexample3} (canonical and optimal DR-
+  plans for pinned line incidence systems
+  \cite{sitharam2014incidence}) and Examples \ref{materialexample4},
+  \ref{materialexample5} (boundary-conditions for achieving various
+  desired properties of body-hyperpin systems).
 
@@
   paper.
-  \todo{Provide Link}.
 \end{itemize}
+
+\noindent
+\note All proofs appear in Appendix A (\ref{sec:appendix:a}).

### 2014-12-20 doc/GMP2015/s03_drp.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

@@
 \begin{theorem}\label{theorem:algo_complexity}
-    There exists an $O(|V|^3)$ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
+    There exists an \candrpcomplexityv algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
 \end{theorem}
@@
 \begin{remark}
-    while the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
+    While the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
 However, the time complexity is exponential in $k$.

### 2014-12-20 doc/GMP2015/s07_problempinning.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

@@
 
-\begin{proof}
-    We can replace each body that has only one pin by a single vertex. A
-    body with 2 pins can be replaced by an edge. In general, a body with $n$
-    pins can be replaced by a 2-tree on $n$ vertices. When finding a
-    DR-plan, we treat each body as trivial, so they become the leaves of our
-    plan. The optimal recombination problem and approach of Section
-    \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in
-Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
-\end{proof}
-
 For the rest of this section, we will only be dealing with the DR-plan of
@@
 \label{obs:bodypin_drp}
-    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity corresponds to a canonical DR-plan whose nodes satisfy one of the following, either
-    (a) the children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph, or
-    (b) the children are a number of proper maximal 1-dof sub-graphs, sharing at most one body.
+    For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
 \end{observation}
+
+% \begin{observation}
+% \label{obs:bodypin_drp}
+%     For the 1-dof body-pin graphs described above, $(3,4)$-sparsity corresponds to a canonical DR-plan whose nodes satisfy one of the following, either
+%     (a) the children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph, or
+%     (b) the children are a number of proper maximal 1-dof sub-graphs, sharing at most one body.
+% \end{observation}
 
@@
     all $(3,4)$-tight canonical DR-plans are optimal. We can find such a
-    DR-plan in the same time complexity as the $(2,3)$-tight case for bar
-    and joint graphs discussed in Section \ref{sec:DRP}.
+    DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
     This is because
@@
     Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding
-    1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal
+    1-dof DR-plan $T$, there is an \candrpcomplexity algorithm for the 1-dof optimal
     completion problem of Section \ref{sec:recomb}.

### 2014-12-20 doc/GMP2015/s09_conclusion.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

@@
 We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
-For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient ($O(n^2)$) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.
+For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient (\candrpcomplexity) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.
 

### 2014-12-20 doc/GMP2015/appendix_a.tex
re-phrased the main proof once again.

@@
 \begin{proof}
-Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs.
-Therefore, these are the only possibilities to consider.
+Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
+Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider.
+Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.
 
 \medskip\noindent
-\vemph{Part 1:} We show that no subset of the children can union to form $C$, thereby requiring all of them be included, and because this is the only choice it must be optimal for this node.
+\vemph{Part 1:} First, we claim that if the intersection is trivial all $C_1,\ldots,C_N$ must be children of $C$ in a canonical DR-plan.
+We do this by showing that no subset of the children can union to form $C$, thereby requiring all of them be included; because this is the only choice, it is the minimum fan-in a canonical DR-plan could have for this node.
+must be optimal for this node.
 
@@
 Thus, the assumption is wrong and $U$ cannot be isostatic.
-As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.
+As $C$ is isostatic, this means no proper subset of $C_1,\ldots,C_N$ can union to form $C$.
+
+Furthermore, since a canonical DR-plan uses proper rigid \vemph{vertex-maximal} subgraphs as children, and because their pairwise intersection is trivial, it follows that any such node has at most as many children as a DR-plan without this restriction, because the union of the children have to contain all edges of the parent.
+Therefore, this is the optimal choice in this case.
 
 \medskip\noindent
-\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+\vemph{Part 2:} Second, we claim that if the intersection is isostatic any two children will suffice for a canonical DR-plan and that it will be optimal.
+
+If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
 This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
-Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
+Thus, any two children can be chosen to make a canonical DR-plan.
+% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
 
-However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in for any descendant.
+\newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
+\renewcommand{\induceonc}[1]{#1}
+\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in S_N\setminus\{#1\}}{R_k}}}
+
+However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants.
 Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
 Suppose we select $i$ and $j$, where $i\neq j$, as the children.
-Since
-\[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
-the children of this node will be
-\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
-and
-\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
-% $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
-% the children of this node will be
-% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
+For convenience, we will assume all subgraphs are induced subgraphs on $C$.
+We know that $C=\induceonc{I\cup\bigcup_{k\in S_N}{R_k}}$ and $C_i=\iunion{i}$.
+The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of which intersect on isostatic subgraphs.
+So any two of these are viable children for $C_i$.
+% Since
+% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
+% the children of $C_i$ will be
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
 % and
-% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other.
-\todo{Prove these are valid children?
-Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once.
-Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
+% % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
+% % the children of this node will be
+% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
+% % and
+% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
+% for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
+This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in S_N$, with every $k$ appearing at least once.
+Thus, regardless of the choice of $C_i,C_j$, then their children, etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
+Therefore, the choice is optimal.
 
-This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
+\medskip\noindent
+This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
 \end{proof}

### 2014-12-20 doc/GMP2015/s03_drp.tex
re-phrased the main proof once again.

@@
 
-\begin{theorem} \label{theorem:canonical_is_optimal}
+\begin{theorem}
+\label{theorem:canonical_exists_and_is_optimal}
+\label{theorem:canonical_is_optimal}
 \label{theorem:main}
-    Any canonical DR-plan is an optimal DR-plan.
+    A canonical DR-plan exists and any canonical DR-plan is optimal.
 \end{theorem}
+
+% \begin{theorem} \label{theorem:canonical_is_optimal}
+%     \label{theorem:main}
+%     Any canonical DR-plan is an optimal DR-plan.
+% \end{theorem}
 
@@
 These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained.
-As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
+As explained in the proof of Theorem \ref{theorem:main} since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
 Just as expected, it is on the third level, as it is a child of both of $G$'s children.

### 2014-12-20 doc/GMP2015/s07_problempinning.tex
re-phrased the main proof once again.

@@
     Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding
-    1-dof DR-plan $T$, there is an \candrpcomplexity algorithm for the 1-dof optimal
+    1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal
     completion problem of Section \ref{sec:recomb}.

### 2014-12-22 doc/GMP2015/appendix_a.tex
Final changes to main proof and small corrections throughout.

@@
 \begin{proof}
-Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
+We show the existence of a canonical DR-plan by constructing it as follows:
+
+Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$
+with children $C_1,\ldots,C_N$ keep children nodes according to the following rules:
+\begin{enumerate}
+   \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
+   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
+\end{enumerate}
+
+This directly satisfies properties 2 and 3 of a canonical DR-plan (see definition \ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs.
+However, we still have to show property 1 (that this makes a DR-plan).
+For rule (1), because we are beginning with a complete DR-plan, if we keep all the children it is still a DR-plan.
+For rule (2), we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we have found a larger rigid vertex-maximal proper subgraph.
+
+If we begin with an isostatic graph, rigid can be replaced with isostatic throughout the construction and remain valid. The rigid proper subgraphs of an isostatic graph must be isostatic themselves.
+
+Now we must show that a canonical DR-plan is optimal.
+The proof is by induction on the level of the plan; we show that following these rules always guarantees minimum fan-in at this level and does not increase fan-in on following levels.
+
+ % We do this inductively on the l, beginning with the root node, showing that the rules are always the optimal choice.
+% by discussing each rule and proving inductively that this is the optimal choice at each level of the DR-plan tree, starting with the root node.
+
+% Now we show that this plan is optimal by considering the two cases.
+Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
+ % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
 Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider.
@@
 \medskip\noindent
-\vemph{Part 1:} First, we claim that if the intersection is trivial all $C_1,\ldots,C_N$ must be children of $C$ in a canonical DR-plan.
-We do this by showing that no subset of the children can union to form $C$, thereby requiring all of them be included; because this is the only choice, it is the minimum fan-in a canonical DR-plan could have for this node.
-must be optimal for this node.
+\vemph{Rule 1:}
+% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
+%
+First, we claim that if the intersection is trivial all $C_1,\ldots,C_N$ must be children of $C$ in an optimal DR-plan.
+We do this by showing that no subset of the children can union to form $C$, thereby requiring all of them be included; because this is the only choice, it is the minimum fan-in any DR-plan could have for this node.
+% must be optimal for this node.
 
@@
 \medskip\noindent
-\vemph{Part 2:} Second, we claim that if the intersection is isostatic any two children will suffice for a canonical DR-plan and that it will be optimal.
+\vemph{Rule 2:} Second, we claim that if the intersection is isostatic, the choice of any two children will result in an optimal DR-plan.
 
@@
 Therefore, the choice is optimal.
-
-\medskip\noindent
-This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
+%
+% \medskip\noindent
+% This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
+%
+% -- say this at top `proof is by induction on the level of the dr=plan'
 \end{proof}
@@
 
-For (2), observe that under-constrained subgraphs of isostatic graphs must have density less than $k$.
+For (2), observe that underconstrained subgraphs of isostatic graphs must have density less than $k$.
 For (3), observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic.
@@
 
-%     % \item Observe that, given a isostatic graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
+%     % \item Observe that, given a isostatic graph, an underconstrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.
 
@@
 
-%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of isostatic $F$, it must be that $d(F_i\cup F_j)=l<k$.
+%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is underconstrained and a subgraph of isostatic $F$, it must be that $d(F_i\cup F_j)=l<k$.
 Therefore $d(F_i\cap F_j)=2k-l>k$.
@@
 \textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial.
-By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.
+By the same math, we find that $d(F_i\cup F_j)<k$, showing it is underconstrained.
 
@@
 We also know it cannot be trivial because it contains isostatic subgraphs.
-This means it must be under-constrained.
+This means it must be underconstrained.
 From Observation \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial.
@@
 The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
-We construct something that, while recombining it, results in the same systems to solve as the true DR-plan.
+We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan.
 Begin with the parent $C$.
@@
 The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
-Etc.
+And so on.
 At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
 Apply the algorithm recursively to both children.
-% Although the nodes containing $R_1,\ldots,R_N$ are under-constrained, simply take the children to be the roots of the forest
+% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
 
-The reason we construct it this way is because we end up with a tree instead of a directed acyclic graph.
-The leaves of this tree will be exactly every single edge in the graph, with none repeated.
+The reason for this construction is that the result is a tree with no duplicate nodes.
+The leaves of this tree will exactly be the edges of this graph.
 Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.
@@
 % \begin{proof}
-% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
+% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the wellconstrained vertex-maximal proper subgraphs of the input wellconstrained graph.
 There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
 After the subgraphs have been found, it is a straightforward application of the theory.
-If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
+If the intersection of any two of the subgraphs is wellconstrained, then take those two to be children of the graph.
 Otherwise, all the subgraphs are the children.
@@
 Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
-    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
+    By Lemma~\ref{lem:pinned_union_intersection}, it must be underconstrained,
     so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

### 2014-12-22 doc/GMP2015/s04_recombination.tex
Final changes to main proof and small corrections throughout.

@@
 \label{sec:recomb}
-
+%
 In this section, we consider the \dfn{optimal recombination} problem
@@
 parent system $I$ is isostatic (the root of the DR-plan), and we seek
-\dfn{solution(s) (among a finite large number of solutions) with a
-specific orientation or chirality}. In the case of under-constrained
+\vemph{solution(s) (among a finite large number of solutions) with a
+specific orientation or chirality}. In the case of underconstrained
 qusecs the subsystems are the multiple roots of the DR-plan, the
-parent system $I$ is under-constrained, and we typically seek an
-efficient algorithmic description of \dfn{connected component(s) of
+parent system $I$ is underconstrained, and we typically seek an
+efficient algorithmic description of \vemph{connected component(s) of
 solutions with a specific orientation or chirality}.
@@
 approach to the optimal modification problem achieves the following:
-\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
+
+\medskip\noindent
+\header{(a) Small DR-plan.} We obtain a  parameterized family of systems
 $I_{\lambda_F}$ --- one for each value $\lambda_F$ for the parameters
@@
 easily once the orientation type of the solution is known  (when the
-DR-plan size is small enough). \dfn{(b) Solution preservation.}
+DR-plan size is small enough).
+
+\medskip\noindent
+\header{(b) Solution preservation.}
 Moreover, the union of solution spaces of the systems in the family
-$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
+$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.
+
+\medskip\noindent
+\header{(c) Efficient search.} Finally, this union can be projected into the
 so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
@@
 problem, and using a key theorem of Convex Cayley configuration spaces
-\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$
-and non-edges $F$ (Cayley parameters), such that the set of vectors
+\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors
 $\lambda_F$ of  attainable lengths of the non-edges $F$ is always
@@
 Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
-a ``projection'' of the Cartesian realization space of $(H,\delta)$ on
+a ``projection'' of the cartesian realization space of $(H,\delta)$ on
 the Cayley parameters $F$. Such graphs $H$ are said to have
@@
 
-To state the theorem, we first have to
-define the notion of \dfn{2-sums} and \dfn{2-trees}.
-Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.
-% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
-A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
-obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
-and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
-components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
-that cannot be further split into 2-sum components. A \dfn{2-tree} is
-recursively obtained by taking a 2-sum of 2-trees, with the base
-case being a triangle. A \dfn{partial 2-tree} is a 2-tree
-minus some edges. Partial 2-trees characterized as having $K_4$ as a
-forbidden minor are also called series parallel graphs.
+To state the theorem, we first have to define the notion of
+\dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and $H_2$ be two graphs on
+disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and
+$E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
+A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the
+union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$. In this
+case, $H_1$ and $H_2$ are called \dfn{2-sum components} of $H$. A
+\dfn{minimal 2-sum component} of $H$ is  one that cannot be further
+split into 2-sum components. A \dfn{2-tree} is recursively obtained by
+taking a 2-sum of 2-trees, with the base case of a 2-tree being a
+triangle. A \dfn{partial 2-tree} is a 2-tree minus some edges. Partial
+2-trees have an alternate characterization as the graphs that avoid
+$K_4$ minors, and are also called series parallel graphs.
 
@@
 % \end{proof}
-
+%
 \begin{figure*}\centering
@@
 \noindent
-\note A huge advantage of the convex Cayley method is that it is
+\note A major advantage of the convex Cayley method is that it is
 completely unaffected when $\delta$ are intervals rather than exact
-values~\cite{sitharam2010convex}.
+values~~\cite{sitharam2010convex}.
+%cite my paper with Heping!!!
 
 \myexample
+[Using Cayley configuration space].
 A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
@@
 simultaneously solving this system, we could instead use the fact that
-$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
+$G=K_{3,3}$ is in OMD$_2$: remove the edges $e_1,e_2$ in Figure
 \ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
-non-edge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
+nonedges $f_1,f_2$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
 configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
@@
 \Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
-$(G,\delta)$ is found when the length of the non-edge $e$ in  the
+$(G,\delta)$ is found when the length of the nonedge $e$ in  the
 realization $q^p_{\lambda_f}$ is $\delta_e$.
+
+In fact, we can show that $G=K_{3,3}$ is in OMD$_1$ by removing a single
+edge to reduce (as shown in Figure \ref{fig:omd_k33_example}) to a
+tree-decomposable graph of low Cayley complexity (which includes the class
+of partial 2-trees).
+In the next Section \ref{sec:tdecomp}, we discuss this issue
+of why the largest class of reduced graphs is desirable.
 %
@@
 \cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
-partial 2-trees. These too result in DR-plans of size 2 or 3, putting
+partial 2-trees. See an illustration in Figure \ref{fig:omd_k33_example}.
+These too result in DR-plans of size 2 or 3, putting
 $G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The

### 2014-12-22 doc/GMP2015/s08_pinnedline.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.
Final changes to main proof and small corrections throughout.

@@
 Euclidean transformations are not factored out.
-In particular, both a single vertex and a single edge are under-constrained graphs.
+In particular, both a single vertex and a single edge are underconstrained graphs.
 
@@
 %since the pins have fixed positions on the plane,
-%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
-%being the disjoint union of two or more well-constrained subgraphs.
+%{\em a wellconstrained vertex-maximal proper subgraph can be disconnected},
+%being the disjoint union of two or more wellconstrained subgraphs.
 %
@@
 %since the pins have fixed positions on the plane,
-%no trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
-%In particular, note that both a single vertex and a single edge are under-constrained graphs.
+%no trivial motion exists, therefore {\em no trivial overconstrained graphs exist} (the smallest overconstrained graph has $6$ vertices).\\
+%In particular, note that both a single vertex and a single edge are underconstrained graphs.
 %
@@
 %since the pins have fixed positions on the plane,
-%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
-%being the disjoint union of two or more well-constrained subgraphs. \\
+%{\em a wellconstrained vertex-maximal proper subgraph can be disconnected},
+%being the disjoint union of two or more wellconstrained subgraphs. \\
 %%Without loss of generality,
@@
 %\begin{itemize}
-%\item For any $C$, its $N$ children, $C_1, \ldots, C_N$, are {\em edges} or {\em connected} well-constrained vertex-maximal proper subgraphs of that node $C$.
+%\item For any $C$, its $N$ children, $C_1, \ldots, C_N$, are {\em edges} or {\em connected} wellconstrained vertex-maximal proper subgraphs of that node $C$.
 %\item A node containing a single edge is a leaf.
@@
 %        %\item The root of the tree `contains' $G$.
-%        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
+%        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any wellconstrained proper subgraph of $C$, or a {\em connected} wellconstrained vertex-maximal proper subgraph of  $C$.
 %        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
@@
 
-Theorem~\ref{theorem:canonical_is_optimal} and \ref{theorem:main}
-hold for  pinned line-incidence graphs with the above modified definition. 
+Theorem~\ref{theorem:main}
+holds for  pinned line-incidence graphs with the above modified definition.
 The proof is similar to the original proof (in the Appendix) using the same set of lemmas and
@@
 There are only two possible cases:
-(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
+(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is underconstrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
 %The following table summarizes all bijections:
@@
 %\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
-% \centering--- &  \sout{over-constrained}          &  \centering --- \tabularnewline \hline
-%at least one of $F_i,F_j$ is an edge & under-constrained       &  \centering a vertex (under-constrained)  \tabularnewline \hline
-%both $F_i$ and $F_j$ are well-constrained & well-constrained        &  \centering well-constrained  \tabularnewline \hline
-% \centering --- & ---                     &  \centering \sout{under-constrained} \tabularnewline \hline
+% \centering--- &  \sout{overconstrained}          &  \centering --- \tabularnewline \hline
+%at least one of $F_i,F_j$ is an edge & underconstrained       &  \centering a vertex (underconstrained)  \tabularnewline \hline
+%both $F_i$ and $F_j$ are wellconstrained & wellconstrained        &  \centering wellconstrained  \tabularnewline \hline
+% \centering --- & ---                     &  \centering \sout{underconstrained} \tabularnewline \hline
 %\end{tabular}

### 2015-02-15 src/GUI/MainGuiManager.cpp
Can now display the drplan tree.

@@ -225,7 +225,14 @@ void MainGuiManager::poll_for_events(){
 
 
 void MainGuiManager::wait_for_events(){
+    std::cout << "\tAbout to wait." << std::endl;
     glfwWaitEvents();
+    std::cout << "\tHandled." << std::endl;
+    // try {
+    //     glfwWaitEvents();
+    // } catch(...) {
+    //     std::cout <<"Error!" << std::endl;
+    // }
 }
 
 

### 2015-02-15 src/GUI/program.cpp

@@ -29,7 +29,8 @@ namespace color {
 
   // gray/green/black
   const gl_obj::color_vec bg   = gl_obj::color_vec(0.85f, 0.85f, 0.85f, 1.f);
-  const gl_obj::color_vec vert = gl_obj::color_vec(0.2f, 0.6f, 0.2f, 0.f);
+  const gl_obj::color_vec h_vert = gl_obj::color_vec(0.2f, 0.6f, 0.2f, 0.f);
+  const gl_obj::color_vec vert = gl_obj::color_vec(0.6f, 0.2f, 0.2f, 0.f);
   // const gl_obj::color_vec vert = gl_obj::color_vec(0.6f, 0.2f, 0.6f, 0.f);
   const gl_obj::color_vec edge = gl_obj::color_vec(0.f, 0.f, 0.f, 0.f);
 
@@ -227,36 +228,34 @@ void Program::check_for_GL_errors(const char* from) {
 }
 
 
-void Program::draw_graph_vertices(
-  vector<gl_obj::pos_vec> &p)
+void Program::_draw_graph_vertices_aux(
+  vector<gl_obj::pos_vec> &p,
+  float radius,
+  gl_obj::color_vec color,
+  bool border)
 {
-  float radius = 0.05;
-  bool border = true;
-
   check_for_GL_errors("Program::draw_graph_vertices - 0");
 
   glUniform1i(render_mode_location, Render_Mode::basic2D);
   check_for_GL_errors("Program::draw_graph_vertices - 1");
 
-
   if (border) {
     // glLineWidth(10);
-
 
     gl_obj::Circle c_outline = gl_obj::Circle(radius, color::edge);
     gl_obj::VertexGroup *vg = &(c_outline.tf->vg);
 
     // initialize the vertex buffer with the vertex data
     glBindBuffer(GL_ARRAY_BUFFER, vbo);
-  check_for_GL_errors("Program::draw_graph_vertices - 2");
+    check_for_GL_errors("Program::draw_graph_vertices - 2");
     // glBufferData(GL_ARRAY_BUFFER, vg.size() * sizeof(Vertex), &vg[0] , GL_STATIC_DRAW);
     glBufferData(GL_ARRAY_BUFFER, vg->size() * sizeof(gl_obj::Vertex), &vg->at(0) , GL_STREAM_DRAW);
-  check_for_GL_errors("Program::draw_graph_vertices - 3");
+    check_for_GL_errors("Program::draw_graph_vertices - 3");
 
[…]

### 2015-02-15 src/GUI/program.hpp

@@ -37,6 +37,7 @@
 
 #include <string>
 #include <vector>
+#include <set>
 
 
 
@@ -75,7 +76,7 @@ public:
   ////
   //// BEGIN DRAWING FUNCTIONS
 
-  void draw_graph_vertices(std::vector<gl_obj::pos_vec> &p);
+  void draw_graph_vertices(std::vector<gl_obj::pos_vec> &p, std::vector<gl_obj::pos_vec> &highlight);
   void draw_graph_edges(std::vector<gl_obj::pos_vec> &p);
 
   void draw_letter(const GLfloat pos[4], const GLfloat tex_coord[4], GLuint tex_id);
@@ -119,6 +120,13 @@ private:
   GLuint translate_location,
          render_mode_location,
          tex_sampler_location;
+
+
+  void _draw_graph_vertices_aux(
+    std::vector<gl_obj::pos_vec> &p,
+    float radius,
+    gl_obj::color_vec color,
+    bool border = true);
 };
 
 

### 2015-02-15 src/Graph/Pebbled_Graph.cpp

@@ -7,24 +7,189 @@
 #include <iostream>
 
 
+Cluster::Cluster(bool f) {
+    finished = f;
+    _parent = NULL;
+    _first_child = NULL; _last_child = NULL;
+    _next = NULL; _prev = NULL;
+}
+Cluster::Cluster(std::set<Vertex_ID> vs, bool f) {
+    vertices = vs;
+    finished = f;
+    _parent = NULL;
+    _first_child = NULL; _last_child = NULL;
+    _next = NULL; _prev = NULL;
+}
+Cluster::Cluster(std::set<Vertex_ID> vs, std::set<Cluster*> cs, bool f) {
+    vertices = vs;
+    finished = f;
+    _parent = NULL;
+    _first_child = NULL; _last_child = NULL;
+    _next = NULL; _prev = NULL;
+
+    add_children(cs);
+}
 
 
-
-void Cluster::print_tree(Graph &g, unsigned int tabs) {
-    for (int i = 0; i < tabs; i++) std::cout << "\t";
-    std::cout << this << " " << finished << ": ";
-    for (std::set<Vertex_ID>::iterator v_it = vertices.begin(); v_it != vertices.end(); v_it++) {
-        std::cout << g[*v_it].name << " ";
+std::set<Cluster*> Cluster::make_children_set() {
+    std::set<Cluster*> ret;
+    for (Cluster* c = _first_child; c != NULL; c = c->next()) {
+        ret.insert(c);
     }
-    std::cout << std::endl;
+    return ret;
+}
 
-    for (std::set<Cluster*>::iterator c_it = children.begin(); c_it != children.end(); c_it++) {
-        (*c_it)->print_tree(g, tabs+1);
+Cluster* Cluster::parent() const {
+    return _parent;
+}
+
+// std::set<Cluster*> Cluster::children() {
+//     return _children;
+// }
+
+Cluster* Cluster::first_child() const {
+    return _first_child;
+    // return (_children.begin() == _children.end())? NULL: *_children.begin();
+}
+Cluster* Cluster::next() const {
+    return _next;
+}
+Cluster* Cluster::prev() const {
+    return _prev;
+}
+
+
+void Cluster::add_vertex(Vertex_ID v) {
+    vertices.insert(v);
+}
+
+
+void Cluster::add_child(Cluster* c) {
+    // _children.insert(c);
+    c->_parent = this;
+    c->_next = NULL; c->_prev = NULL;
+
+    if (_first_child == NULL) {
[…]

### 2015-02-15 src/Graph/Pebbled_Graph.hpp

@@ -11,27 +11,44 @@
 
 struct Cluster {
     std::set<Vertex_ID> vertices;
-    std::set<Cluster*> children;
-    // std::set<Edge_ID> edges;
-
     bool finished;
 
+    // std::set<Cluster*> _children;
 
-    Cluster(bool f=false)
-        {finished = f;}
-    Cluster(std::set<Vertex_ID> vs, bool f=false)
-        {vertices = vs; finished = f;}
-    Cluster(std::set<Vertex_ID> vs, std::set<Cluster*> cs, bool f=false)
-        {vertices = vs; children = cs; finished = f;}
+    Cluster(bool f=false);
+    Cluster(std::set<Vertex_ID> vs, bool f=false);
+    Cluster(std::set<Vertex_ID> vs, std::set<Cluster*> cs, bool f=false);
+
+    std::set<Cluster*> make_children_set();
+
+    Cluster* parent() const;
+
+    // std::set<Cluster*> children();
+
+    Cluster* first_child() const;
+    Cluster* next() const;
+    Cluster* prev() const;
 
 
-    void add_vertex(Vertex_ID v) {vertices.insert(v);}
+    void add_vertex(Vertex_ID v);
 
-    // void finished(bool f) {_finished = f}
-    // bool finished() {return finished;}
-    // void add_edge(Edge_ID e) {edges.insert(e);}
 
-    void print_tree(Graph &g, unsigned int tabs = 0);
+    void add_child(Cluster* c);
+    void add_children(std::set<Cluster*> cs);
+
+
+    unsigned int height();
+    unsigned int width(); // of the largest level
+    std::vector<unsigned int> width_per_level();
+
+    void print_tree(const Graph *g, unsigned int tabs=0);
+private:
+    Cluster
+        *_parent,
+        *_first_child,
+        *_last_child,
+        *_next,
+        *_prev;
 };
 
 
@@ -46,7 +63,7 @@ public:
     std::set<Cluster*> component_pebble_game_2D(Vertex_ID *exclude = NULL);
 
     // If optimal is false, you find completeDRP
-    Cluster DRP_2D(bool optimalDRP = true);
+    Cluster* DRP_2D();
 
 
 private:

### 2015-02-15 src/main.cpp

@@ -8,6 +8,7 @@
 
 #include <iostream>
 #include <vector>
+#include <map>
 using namespace std;
 
 
@@ -32,7 +33,9 @@ using namespace std;
 
 
 namespace global {
-    Program *myProg;
+    // Program *myProg;
+
+    MainGuiManager mgm;
 
     namespace window {
         const int width_screen_coords = 1000, height_screen_coords = 1000;
@@ -42,134 +45,351 @@ namespace global {
 }
 
 
-class Demowindow : public Window {
-    void key_callback(int key, int scancode, int action, int mods) {
-        if (key == GLFW_KEY_ESCAPE && action == GLFW_PRESS)
-            cout << "HERE!" << endl;
-            // glfwSetWindowShouldClose(window, GL_TRUE);
-    };
+
+class DrpDisplayWindow : public Window {
+public:
+    Program *program;
+
+    Cluster *drp;
+    std::vector<gl_obj::pos_vec> vertices, vertices_highlight;
+    std::vector<gl_obj::pos_vec> edges;
+
+    // unsigned int height;
+    // std::vector<unsigned int> width_per_level;
+
+    DrpDisplayWindow(Cluster *drp) {
+        this->drp = drp;
+
+
+        // this->graph = graph;
+        // this->current_drp_node = NULL;
+
+        update_graph_positions();
+        // display_graph();
+    }
+
+    void init_program() {
+        set_as_context();
+        this->program = new Program("src/GUI/shaders/vs.glsl", "src/GUI/shaders/fs.glsl");
+    }
+
+
+    void update_graph_positions() {
+        // erase old
+        vertices.clear();
+        vertices_highlight.clear();
+        edges.clear();
+
+
+        std::vector<unsigned int> width_per_level = drp->width_per_level();
+        unsigned int height = width_per_level.size();
+
+        std::map<Cluster*, gl_obj::pos_vec> node_position;
+
+
+        std::vector<Cluster*> this_level, next_level;
+        this_level.push_back(drp);
+
+        gl_obj::pos_vec parent(0, 1);
+
+        std::vector<unsigned int> position(height);
+
+        for (unsigned int i = 0; i < height; i++) {
+            position[i] = 0;
+            float y = (height == 1)? 0 : 1 - ((float)i)/(height-1)*2;
+
+            for (std::vector<Cluster*>::iterator c = this_level.begin(); c != this_level.end(); c++) {
[…]

### 2015-02-15 doc/GMP2015/original_submission/appendix_b.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-19 doc/GMP2015/appendix_b.tex
@@
 
-\subsection{Furture Work}
+\subsection{Open Problems}
 \label{sec:futurework}
-A few natural  open questions are the following.
-The first 2 are from Section \ref{sec:recomb}.
+
+One open problem comes from Section \ref{sec:DRP}:
+\begin{openproblem}
+    Is there a lower complexity algorithm than \candrpcomplexityv to find the canonical DR-plan of isostatic 2D bar-joint graphs?
+\end{openproblem}
+
+A few natural  open questions concern the following common theme that runs through the paper:
 
 \begin{openproblem}
-    Combinatorial rigidity for
-    periodic structures is an active area of research. This paper motivates
-    a study of the rigidity of self-similar structures,
-    with self-similar groups replacing periodic groups.
+    For fixed $k$ we have polynomial time optimal DR-planning, recombination (modification) in the presence of $k$ overconstraints (Section \ref{sec:DRP}, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed \ref{sec:recomb} and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan \ref{sec:table}.
+However, $k$ appears in the exponent for the straightforward algorithms for these problems.
+\end{openproblem}
+
+One problem in the above theme is from Section \ref{sec:bodypin}.
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when the given graph has more than 2-dofs?
+Our proof for the 1 and 2-dof cases relied heavily on the matroidal properties of their corresponding $(k,l)$-tightness.
+For higher number of dofs, the $(k,l)$ characterization is no longer matroidal \cite{Lee:2007:PGA}.
+As a result, the major obstacle is that there is no easy way of obtaining an optimal or canonical $k$-dof DR-plan in general.
+Even assuming such  a DR-plan is available, if higher dofs had the same characteristics, Observation \ref{obs:algebraic_completion} raises questions about the correct measure of DR-plan size that captures algebraic complexity for recombining graphs with many dofs (this is not an issue in the isostatic case).
+Unless some restrictions can be found and taken advantage of, the $k$-dof optimal completion problem would  have complexity exponential in $k$.
+\end{openproblem}
+
+Another problem from the above theme is from Section \ref{sec:recomb}
+\begin{openproblem}
+    What is the complexity of the restricted OMD (optimal modification for decomposition) problem?
+This has the potential to be difficult.
+For example, when the isostatic completion is required to be a 2-tree the restricted OMD problem is reducible to the maximum spanning series-parallel subgraph problem shown by \cite{cai1993spanning} to be NP-complete even if the input graph is planar of maximum degree at most 6.
+However, since the OMD problem has other input restrictions such as not having any proper isostatic subgraphs, it is not clear if the reverse reduction exists and hence it is unclear whether the OMD problem is NP-complete.
+
+    The same holds for the restricted OMD problem where the isostatic completion is required to be a tree-decomposable graph of low Cayley complexity (i.e, have special, small DR-plans).
+One potential obstacle to an indecomposable graph $G$'s membership in the restricted OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth have been constructed in \cite{servatius2000rigidity}.
+\end{openproblem}
+
+The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in Section \ref{sec:table}.
+\begin{openproblem}
+    Is the OMD (optimal modification for decomposition) problem reducible to the OC (Optimal completion) problem?
+\end{openproblem}
+
+More general problem directions are the following.
+\begin{openproblem}
+    Combinatorial rigidity for periodic structures is an active area of research.
+This paper motivates a study of the rigidity of self-similar structures, with self-similar groups replacing periodic groups.
 \end{openproblem}
@@
 \begin{openproblem}
-    The pinned line incidence structures of Section \ref{sec:pinnedline},
-    for example in the case of collagen microfibrils, whose function is
-    elastic contraction, should be considered
-    congruent under projective transformations. I.e, the projective group
-    should be factored out as a trivial motion in a new project for
-    extending the combinatorial rigidity characterization of such systems.
+    The pinned line incidence structures of Section \ref{sec:pinnedline}, for example in the case of collagen microfibrils, whose function is elastic contraction, should be considered congruent under projective transformations.
+I.e, the projective group should be factored out as a trivial motion in a new project for extending the combinatorial rigidity characterization of such systems.
 (Currently we permit no trivial motions at all).
@@
 \begin{openproblem}
-What is the complexity of the restricted OMD (optimal modification for
-decomposition) problem?
-This has the potential to be difficult. For example, when the
-isostatic completion is required to be a 2-tree the restricted OMD
-problem is reducible to the maximum spanning series-parallel subgraph
-problem shown by \cite{cai1993spanning} to be NP-complete even if the input
-graph is planar of maximum degree at most 6. However, since the OMD
-problem has other input restrictions such as not having any proper
-isostatic subgraphs, it is not clear if the reverse reduction exists
-and hence it is unclear whether the OMD problem is NP-complete.
-
-The same holds for the restricted OMD problem where the isostatic
-completion is required to be a tree-decomposable graph of low
-Cayley complexity (i.e, have special, small DR-plans). One potential
-obstacle to an indecomposable graph $G$'s membership in the restricted
-OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
-In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
-have been constructed in \cite{servatius2000rigidity}.
+    Experimental validation (either computationally or physical experiments) of predications based on the material model and theory used in this paper.
+This can be done by modeling known materials and putting stresses on them, seeing if the prediction is observed in the real material.
+Or, our theory could be used to design new materials, which can be tested to see if they possess predicted properties.
 \end{openproblem}
-
-The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in Section \ref{sec:table}.
-
-\begin{openproblem}
-    Is the OMD (optimal modification for decomposition) problem
-    reducible to the OC (Optimal completion) problem?
-\end{openproblem}
-
-The last problem is from
-Section \ref{sec:bodypin}.
-
-\begin{openproblem}
-    What is the complexity of the optimal completion problem when the given
-    graph has more than 2-dofs?
-    Our proof for the 1 and 2-dof cases relied heavily on the matroidal
-    properties of their corresponding $(k,l)$-tightness.
-    For higher number of dofs, the $(k,l)$ characterization is
-    no longer matroidal \cite{Lee:2007:PGA}.
-    As a result, the major obstacle is that
-    there is no easy way of obtaining an optimal or canonical
-    $k$-dof DR-plan in general.
-    Even assuming such  a DR-plan is available,
-    if higher dofs had the same characteristics,
-    Observation \ref{obs:algebraic_completion}
-    raises questions about the correct measure of DR-plan size that
-    captures algebraic complexity for recombining graphs with many dofs
-    (this is not an issue in the isostatic case). Unless some restrictions
-    can be found and taken advantage of, the $k$-dof optimal
-    completion problem would  have complexity exponential in $k$.
-\end{openproblem}

### 2015-02-15 doc/GMP2015/original_submission/s01_introduction.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-20 doc/GMP2015/s01_introduction.tex
@@
 constraining planar (genus 0) monolayers. Furthermore, a layer is
-often  either \vemph{isostatic or under-constrained, not self-
+often  either \vemph{isostatic or underconstrained, not self-
 stressed} (See Section \ref{sec:prelim} for definitions). These
@@
 \medskip\noindent
-\header{(v) Configuration Spaces of Under-constrained Systems.}
-For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
+\header{(v) Configuration Spaces of Underconstrained Systems.}
+For underconstrained 2D bar-joint and body-hyperpin qusecs obtained
 from various subclasses of tree-decomposable systems, algorithms have
@@
   DR-plan and showing a strong Church-Rosser property: \vemph{all
-  canonical DR-plans for isostatic or under-constrained 2D qusecs are
+  canonical DR-plans for isostatic or underconstrained 2D qusecs are
   optimal}.
@@
   efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well
-  known problems such as optimal completion of under-constrained
+  known problems such as optimal completion of underconstrained
   systems
@@
   modifications are bounded, we obtain new, efficient algorithms for
-  realizing both isostatic and under-constrained qusecs by leveraging
+  realizing both isostatic and underconstrained qusecs by leveraging
   recent results about Cayley parameters in
@@
 \noindent
-\note All proofs appear in Appendix A (\ref{sec:appendix:a}).
+\note All proofs appear in the Appendix (submitted an a permitted attachment via EasyChair).

### 2015-02-15 doc/GMP2015/original_submission/s03_drp.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-20 doc/GMP2015/s03_drp.tex
@@
 % In the problem of the optimal DR-Plan there is generally not a unique plan.
-% % Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
+% % Indeed, we will prove a union of $N$ isostatic subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
 Therefore, all choices of decomposition are in some sense equivalent.
@@
 Trivial graphs are a single vertex and empty set.
-Furthermore, 2D well-constrained graphs must be connected.
-% % The greatest density of a 2D well-constrained graph is $-2$ (the vertex).
-The other disconnected part of the graph would need to have a density of $-1$, which is over-constrained and not possible in a well-constrained graph (because there is no trivial graph with that density).
+Furthermore, 2D isostatic graphs must be connected.
+% % The greatest density of a 2D isostatic graph is $-2$ (the vertex).
+The other disconnected part of the graph would need to have a density of $-1$, which is overconstrained and not possible in a isostatic graph (because there is no trivial graph with that density).
 
@@
 
-\begin{definition}
-    The \dfn{canonical DR-plan} is a DR-plan that satisfies the following two properties:
-    (1) children are isostatic vertex-maximal subgraphs of the parent; and
-    (2) all of the isostatic vertex-maximal subgraphs are children if any two of them intersect trivially, otherwise exactly two are children.
+\begin{definition}\label{def:canonical_drplan}
+    The \dfn{canonical DR-plan} satisfies the following three properties:
+    (1) it is a DR-plan;
+    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
+    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise take as children exactly two that intersect non-trivially.
 \end{definition}
@@
 %\end{enumerate}
-%The resulting canonical DR-plan is an optimal DR-plan.
+%This is a canonical DR-plan.
 %\end{theorem}
@@
 (1) $F_i\cup F_j$ is not trivial;
-(2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial;
-(3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and
-(4) $F_i\cap F_j$ is not under-constrained.
+(2) $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial;
+(3) $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic; and
+(4) $F_i\cap F_j$ is not underconstrained.
 \end{observation}
@@
 The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$.
-These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained.
+These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic.
 As explained in the proof of Theorem \ref{theorem:main} since there are only 2 possible children, their intersection must be a node 2 levels below the parent.

### 2015-02-15 doc/GMP2015/original_submission/s05_problems.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-19 doc/GMP2015/s05_problems.tex
@@
 The OMD problem is closely related to a well-studied problem of
-completion of an under-constrained system to an isostatic one with a
+completion of an underconstrained system to an isostatic one with a
 small DR-plan.
@@
 child solved subgraphs (resp. two root subgraphs when the input graph
-is under-constrained) share more than 1 vertex. We replace such solved
+is underconstrained) share more than 1 vertex. We replace such solved
 subgraphs  by isostatic graphs as follows. If a solved subgraph shares

### 2015-02-15 doc/GMP2015/original_submission/s07_problempinning.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-20 doc/GMP2015/s07_problempinning.tex
@@
-\section{Application: Finding completions of under-constrained glassy structures to isostatic}
+\section{Application: Finding completions of underconstrained glassy structures to isostatic}
 \label{sec:bodypin}
@@
 This can be done in a number of ways:
-(1) pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass});
+(1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass});
 (2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or

### 2015-02-15 doc/GMP2015/revision_1/s00_abstract.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-20 doc/GMP2015/s00_abstract.tex
@@
 leverage recent results on Cayley configuration spaces to show that
-the indecomposable systems -  that are solved at node of the optimal
-DR-plan by recombining solutions to child systems - can be minimally
+the indecomposable systems ---  that are solved at node of the optimal
+DR-plan by recombining solutions to child systems --- can be minimally
 modified to become decomposable and have a small DR-plan, leading to

### 2015-02-15 doc/GMP2015/revision_1/s02_preliminaries.tex
Made dir for original submission. Made dir for revision one and got it to compile with the new class file.

--- 2014-12-19 doc/GMP2015/s02_preliminaries.tex
@@
 %where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
-%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
-It is \dfn{flexible} if it is not rigid, \dfn{under-constrained} if it
-is independent and not rigid, or \dfn{over-constrained} if it is not
+%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
+It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
+is independent and not rigid, or \dfn{overconstrained} if it is not
 independent.
@@
 %
-%    \item \textbf{over-constrained} if it is not independent.
+%    \item \textbf{overconstrained} if it is not independent.
 %
-%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
+%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
 %
-%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
 %
-%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
 %
-%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
+%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
 %\end{itemize}
@@
 % \begin{definition}
-%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
+%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
 %     % induced
@@
 % \begin{definition}
-%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
+%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
 %     % induced
@@
 % \begin{definition}
-%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
+%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
 %     \todo{Maybe leave the following out until it's needed later?}
@@
         % subgraphs of $C$.
-        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
-        set of $C$.
+        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
         \item A leaf node is a single vertex.
@@
 %
-    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
+    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
 %
@@
 % \begin{definition}
-%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
 By this, rule 3 is implicit for a complete DR-plan.
@@
 
-%\emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
+%\emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.
 % \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
@@
 
+\medskip\noindent
 \header{Realizing given qusecs.}
@@
 
+\medskip\noindent
 \header{Recursive block
@@
 %%
-%%    \item \textbf{over-constrained} if it is not independent.
+%%    \item \textbf{overconstrained} if it is not independent.
 %%
-%%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
+%%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
 %%
-%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
+%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
 %%
-%%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
+%%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
 %%
-%%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
+%%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
 %%\end{itemize}
 %%
-%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.
+%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must overconstrained and (2) all of its subgraphs are also trivial.
 %
 %% \begin{definition}
-%%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
+%%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
 %%     % induced
@@
 %% \begin{definition}
-%%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
+%%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
 %%     % induced
@@
 %% \begin{definition}
-%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
+%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
 %%     \todo{Maybe leave the following out until it's needed later?}
@@
 %        \item The root of the tree `contains' $G$.
-%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
 %        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
@@
 %
-%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
+%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
 %
@@
 %
-%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
 This makes rule 3 implicit.
@@
 %% \begin{definition}
-%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or well-constrained vertex-maximal proper subgraphs of that node $C$.
+%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
 By this, rule 3 is implicit for a complete DR-plan.
@@
 %
-%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{well-constrained} bar and joint graph with values $(V,E,w_V,w_E)$.
+%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.
 %% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
@@
 $C_i$ is the $i^{\text{th}}$ child of $C$.
-By definition of a DR-plan, it is implied that $C_i$ is a well-constrained vertex-maximal proper subgraph.
+By definition of a DR-plan, it is implied that $C_i$ is a wellconstrained vertex-maximal proper subgraph.

### 2015-02-16 doc/GMP2015/revision_1/s00_abstract.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2014-12-20 doc/GMP2015/s00_abstract.tex
@@
 \begin{abstract}
-A variety of quasi-uniform (aperiodic) and self-similar,
-layered material structures are a natural consequence of design
-objectives such as rigidity or flexibility, minimizing mass, optimally
-distributing external stresses and participating in the assembly of
-diverse and multifunctional, larger  structures.
+A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
 
-Crucial for the design and analysis of the abovementioned properties
-is the recursive decomposition of diverse types of underlying 2D
-geometric constraint systems that are \vemph{isostatic or
-underconstrained (which we call collectively as \dfn{qusecs})}. While
-the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
-hard even for (possibly overconstrained) 2D bar-joint constraint
-systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
-elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
-plans that meet various desirable, previously studied criteria. We
-leverage recent results on Cayley configuration spaces to show that
-the indecomposable systems -  that are solved at node of the optimal
-DR-plan by recombining solutions to child systems - can be minimally
-modified to become decomposable and have a small DR-plan, leading to
-efficient realization algorithms. We show formal connections to well-
-known problems such as completion of underconstrained systems. We
-briefly discuss the modeling and analysis of two specific 2D layered
-materials as applications of the above theory and algorithms. We
-intend to make a software implementation and videos available upon
-request and publicly available for the final version.
+Crucial for the design and analysis of the abovementioned properties is the recursive decomposition of diverse types of underlying 2D geometric constraint systems that are \vemph{isostatic or underconstrained (which we call collectively as \dfn{qusecs})}.
+While the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-hard even for (possibly overconstrained) 2D bar-joint constraint systems we describe an \candrpcomplexity\ algorithm  for qusecs.
+Besides elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-plans that meet various desirable, previously studied criteria.
+We leverage recent results on Cayley configuration spaces to show that the indecomposable systems---that are solved at node of the optimal DR-plan by recombining solutions to child systems---can be minimally modified to become decomposable and have a small DR-plan, leading to efficient realization algorithms.
+We show formal connections to well-known problems such as completion of underconstrained systems.
+We briefly discuss the modeling and analysis of two specific 2D layered materials as applications of the above theory and algorithms.
+We intend to make a software implementation and videos available upon request and publicly available for the final version.
 \end{abstract}
+
+\begin{keyword}
+    rigidity \sep
+    geometric constraint solving \sep
+    configuration spaces \sep
+    self-similar structures \sep
+    layered materials
+    %
+    % \PACS 71.35.-y \sep 71.35.Lk \sep 71.36.+c
+\end{keyword}
+
+% Easychair keyphrases
+% canonical dr plan (380), optimal dr plan (364), constraint system (236), pinned line incidence (200), cayley configuration space (198), geometric constraint system (190), pinned line incidence graph (180), geometric constraint (166), pinned line incidence system (120), self similar (115), bar joint (113), cayley configuration (90), parent system (90), pinned line (85), strong church rosser property (80), external stress (80), rigidity matrix (80), tree decomposable graph (79), low cayley complexity (79), bar joint system (79), connected component (70), line incidence (70), orientation type (70), wikimedia common (70), cross linking (70), computational geometry (70), rigid vertex maximal proper subgraph (69), bar joint graph (63), appropriate rigidity matrix (63), plant cell wall (63)
 

### 2015-02-16 doc/GMP2015/revision_1/s01_introduction.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2015-02-15 doc/GMP2015/original_submission/s01_introduction.tex
@@
 
-Some properties of natural and engineered materials can be analyzed by
-treating them as two dimensional (2D) layers. As illustrated by the
-examples below, the structure within each layer is often self-
-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally
-aperiodic and quasi-uniform within any one scale; and consists of a
-few repeated motifs appearing in disordered arrangements. Each layer
-is not necessarily planar, i.e., it consists of multiple, inter-
-constraining planar (genus 0) monolayers. Furthermore, a layer is
-often  either \vemph{isostatic or underconstrained, not self-
-stressed} (See Section \ref{sec:prelim} for definitions). These
-properties are consistent with a self-assembled 2D structure that
-minimizes mass, optimally distributes external stresses and itself
-participates in the assembly of diverse and multifunctional, larger 2D
-structures.
+Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers.
+As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements.
+Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers.
+Furthermore, a layer is often  either \vemph{isostatic or underconstrained, not self-stressed} (See Section \ref{sec:prelim} for definitions).
+These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.
 
@@
 \textbf{Note on Scope:} In this manuscript we only study finite 2D structures.
-\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy
-or subdivision in an iterated scheme to generate self-similar structures.
+\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.
 
-\begin{figure*}\centering
-% \begin{subfigure}{.31\linewidth}\centering
-%   \includegraphics[width=\linewidth]{img/Axopodium_Mikrotubuli}
-%   \caption{Cross section of a Heliozoa's psuedopod, formed by a spiral structure of microtubules.}
-%   % https://de.wikipedia.org/wiki/Mikrotubulus#mediaviewer/File:Axopodium_Mikrotubuli.jpg
-%   \label{fig:material_examples:microtubule}
-% \end{subfigure}\hfill
-% \begin{subfigure}{.31\linewidth}\centering
-%   \includegraphics[width=\linewidth]{img/Transverse_Section_Of_Bone}
-%   \caption{Cross section of an osteon (the fundamental unit of compact bone) exhibiting its hierarchical structure.}
-%   % https://en.wikipedia.org/wiki/Osteon#mediaviewer/File:Transverse_Section_Of_Bone.png
-%   \label{fig:material_examples:osteon}
-% \end{subfigure}\hfill
-\begin{subfigure}{.28\linewidth}\centering
-  \includegraphics[width=\linewidth]{img/Chlamydomonas_TEM_17}
-  \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
-  \label{fig:material_examples:microtubule}
-\end{subfigure}\hfill
-\begin{subfigure}{.40\linewidth}\centering
-  \includegraphics[width=\linewidth]{img/ligten2}
-  \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
-  \label{fig:material_examples:tendon}
-\end{subfigure}\hfill
-\begin{subfigure}{.30\linewidth}\centering
-  \includegraphics[width=\linewidth]{img/Rothemund-DNA-SierpinskiGasket}
-  \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
-  \label{fig:material_examples:sierpinski}
-\end{subfigure}
+% \FigInit{}{fig:material_examples}
+% \FigThreeSubfig%
+%   {img/Chlamydomonas_TEM_17}
+%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
+%   {fig:material_examples:microtubule}%
+%   %
+%   {img/ligten2}
+%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
+%   {fig:material_examples:tendon}%
+%   %
+%   {img/Rothemund-DNA-SierpinskiGasket}
+%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
+%   {fig:material_examples:sierpinski}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.3}{img/Chlamydomonas_TEM_17}
+\SetMyMinHeight{.3}{img/ligten2}
+\SetMyMinHeight{.3}{img/Rothemund-DNA-SierpinskiGasket}
+
+\begin{figure*}\centering%
+  %
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/Chlamydomonas_TEM_17}
+    \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}\label{fig:material_examples:microtubule}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/ligten2}
+    \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}\label{fig:material_examples:tendon}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/Rothemund-DNA-SierpinskiGasket}
+    \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples:sierpinski}
+  \end{subfigure}%
+  %
   \caption{}\label{fig:material_examples}
-\end{figure*}
+\end{figure*}%
+
+% \FigAddSubfig%
+%   {0.3}
+%   {img/Chlamydomonas_TEM_17}
+%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
+%   {fig:material_examples:microtubule}%
+% \FigAddSubfig%
+%   {0.3}
+%   {img/ligten2}
+%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
+%   {fig:material_examples:tendon}%
+% \FigAddSubfig%
+%   {0.3}
+%   {img/Rothemund-DNA-SierpinskiGasket}
+%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
+%   {fig:material_examples:sierpinski}
+% \FigDisplay{}{fig:material_examples}
 
@@
 
-    \item \label{materialexample4} More recent, engineered examples, including disordered
-    graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes
-    reinforced by microfibrils; and DNA assemblies including a recent
-    Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other
-    self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
+    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
 
-    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
-    materials that behave like assemblies of 2D particles under
-    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
+    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and materials that behave like assemblies of 2D particles under non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}.
+See Section \ref{sec:bodypin}.
 \end{enumerate}
 %
-In order to study structural and mechanical properties
-of a material layer, it is natural to model a material layer as a solution or
-realization of a geometric constraint system of appropriate types of geometric
-primitives, under metric or algebraic constraints. Such 2D
-\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
-formally in Section \ref{sec:prelim}) can be
-used to understand or design material layers (their solutions) with
-desired properties.
+In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints.
+Such 2D \dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined formally in Section \ref{sec:prelim}) can be used to understand or design material layers (their solutions) with desired properties.
 %
 \subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
-We now briefly survey existing techniques for studying 2D
-qusecs,  many of which are \dfn{bar-joint} systems
-(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
-\dfn{body-hyperpin} systems (Example 4,5, see Section
-\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
-3, see Section \ref{sec:pinnedline}). The limitations of these
-techniques directly motivate the contributions of this paper.
+We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4,5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+The limitations of these techniques directly motivate the contributions of this paper.
 
@@
 \header{(i) Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal,
-\dfn{generically rigid} subsystems (defined formally in Section
-\ref{sec:prelim}). When the input itself is rigid, these
-algorithms do nothing, i.e., compute the identity function.
+Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems (defined formally in Section \ref{sec:prelim}).
+When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.
 
-However, both for self-similar or just aperiodic 2D qusecs, it is
-imperative to recursively decompose rigid systems into their rigid
-subsystems, down to the level of geometric primitives, in order to
-understand or design properties at all scales, such as (formally
-defined in Section \ref{sec:prelim}) \dfn{rigidity}, \dfn{flexes},
-distribution of \dfn{external stresses}, boundary conditions for
-\dfn{isostaticity}, as well as behavior under constraint variations.
+However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as (formally defined in Section \ref{sec:prelim}) \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
 
@@
 \header{(ii) Optimal Recursive Decomposition (DR-planning).}
-Recursive decomposition of geometric constraint systems has been
-formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the
-\dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in
-Section \ref{sec:prelim}. For the above-
-mentioned classes of 2D qusecs, generic rigidity is a combinatorial
-property and hence each level of the decomposition should, in
-principle, be achievable by a graph-based algorithm as in (1), without
-involving geometric information in the constraint system. Since many
-such decompositions can exist for a given constraint system, criteria
-defining desirable or optimal DR-plans and DR-planning algorithms were
-given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes
-the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e.,
-the maximum number of child subsystems of any parent system.
-Being exponential the size, the complexity of solving the parent
-constraint system from the solutions of the child systems, is overwhelmingly
-dominated by it.
+Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in Section \ref{sec:prelim}.
+For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system.
+Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}.
+An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e., the maximum number of child subsystems of any parent system.
+Being exponential the size, the complexity of solving the parent constraint system from the solutions of the child systems, is overwhelmingly dominated by it.
 
-However, for general 2D qusecs that could be overconstrained, even
-when restricted to bar-joint systems, the optimal DR-planning problem
-was shown to be NP-hard \cite{lomonosov2004graph}.
+However, for general 2D qusecs that could be overconstrained, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph}.
 
@@
 criteria.}
-For a special class of 2D qusecs, namely \dfn{tree-decomposable}
-systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which
-includes ruler-and-compass and Henneberg-I constructible systems), all
-DR-plans turn out to be optimal. This satisfies the so-called \dfn
-{Church-Rosser} criterion, leading to highly efficient DR-planning
-algorithms. For general 2D qusecs, alternate criteria were suggested
-such as \dfn{cluster minimality} requiring parent systems to be
-composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
-no proper subset forms a rigid system); and \dfn{proper maximality},
-requiring child subsystems to be maximal rigid proper subsystems of
-the parent system. See Section \ref{sec:prelim} for formal definitions.
+For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.
+This satisfies the so-called \dfn {Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
+For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
+See Section \ref{sec:prelim} for formal definitions.
 
-While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-
-plans meeting the cluster minimality criterion, no such algorithm is
-known for the latter criterion.
+While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-plans meeting the cluster minimality criterion, no such algorithm is known for the latter criterion.
 
@@
 \header{(iv) Optimal Recombination and Solution Space Navigation.}
-For the entire DR-plan, finding all desired solutions is barely
-tractable even if recombination of solved subsystems is easy for each
-indecomposable parent system in the DR-plan. This is because even for
-the simplest, highly decomposable systems with small DR-plans, the
-problem of finding even a single solution to the input system at the
-root of the DR-plan is NP-hard  \cite{saxe1979embeddability}
-and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired
-solution has a given orientation type, in which case, the crux of the
-difficulty is concentrated in the algebraic complexity of
-(re)combining child system solutions, to give a solution to the parent
-system at any given level of the DR-plan. For fairly general 3D
-constraint systems, there are algorithms with formal guarantees that
-uncover underlying matroids to combinatorially obtain an optimal
-parameterization to minimize the algebraic complexity of the
-indecomposable parent (sub)systems that occur in the DR-plan
-\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the
-abovementioned criteria.
+For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
+This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
+Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions, to give a solution to the parent system at any given level of the DR-plan.
+For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
 
-However, the generality of these algorithms trades-off against
-efficiency, and despite the optimization, the best algorithms can
-still take exponential time in the number of child subsystems (which
-can be arbitrarily large even for optimal DR-plans) in order to
-guarantee all solutions of a given orientation type, even for a single
-(sub)system in a DR-plan. They are prohibitively slow in practice. We
-note that, utilizing the DR-plan and optimal recombination as a
-principled basis, high performance heuristics and software exists
-\cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
+However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan.
+They are prohibitively slow in practice.
+We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
 
@@
 \header{(v) Configuration Spaces of Underconstrained Systems.}
-For underconstrained 2D bar-joint and body-hyperpin qusecs obtained
-from various subclasses of tree-decomposable systems, algorithms have
-been developed to complete them into isostatic systems
-\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components
-\cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
-Most of the
-algorithms with formal guarantees leverage Cayley configuration space
-theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and
-additional constraints that control combinatorial explosion, and
-provide faithful bijective representation of connected components and
-paths. These algorithms have decreasing efficiency as the subclass of
-systems gets bigger, with highest efficiency for underlying partial
-2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$
-minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
-graphs with low Cayley complexity (which include common linkages such
-as the Strandbeest, Limacon and Cardioid), and decreased efficiency
-for general 1-dof tree-decomposable graphs. While software suites
-exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known
-for non-tree-decomposable systems.
+For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
+Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
+While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
 %
@@
 \begin{itemize}
-  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier mentioned in (2) above,
-  for finding optimal DR-plans by defining a so-called \dfn{canonical}
-  DR-plan and showing a strong Church-Rosser property: \vemph{all
-  canonical DR-plans for isostatic or underconstrained 2D qusecs are
-  optimal}.
+  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.
 
-  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical
-  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
-  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
-  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
-  of the NP-hardness of finding optimal DR-plans for over-constrained
-  systems. Futhermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.
+  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
+The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
+Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.
 
-  \item In Section \ref{sec:recomb}, we give a method to deal with the
-  algebraic complexity of recombining the realizations or solutions of
-  child subsystems into a solution of the parent system
-  \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
-  Specifically, we define the problem of minimally modifying the
-  indecomposable recombination system so that it becomes decomposable
-  via a small DR-plan and yet preserves the original solutions in an
-  efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well
-  known problems such as optimal completion of underconstrained
-  systems
-  \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
-  and to find paths within the connected components. When the
-  modifications are bounded, we obtain new, efficient algorithms for
-  realizing both isostatic and underconstrained qusecs by leveraging
-  recent results about Cayley parameters in
-  \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
+  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
+Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
+In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components.
+When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
 
-  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we
-  briefly describe applications of the above techniques to modeling,
-  analyzing, and designing specific properties in 2D material
-  layers~\cite{Jackson2008bodypin}. For Examples
-  \ref{materialexample1}, \ref{materialexample2} (achieving
-  isostaticity, distribution of stresses in self-similar, bar-joint
-  systems); Example \ref{materialexample3} (canonical and optimal DR-
-  plans for pinned line incidence systems
-  \cite{sitharam2014incidence}) and Examples \ref{materialexample4},
-  \ref{materialexample5} (boundary-conditions for achieving various
-  desired properties of body-hyperpin systems).
+  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
+For Examples \ref{materialexample1}, \ref{materialexample2} (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example \ref{materialexample3} (canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}) and Examples \ref{materialexample4}, \ref{materialexample5} (boundary-conditions for achieving various desired properties of body-hyperpin systems).
 
-  \item We intend to make software implementation and videos available
-  upon request and publicly available for the final version of the
-  paper.
+  \item We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
 \end{itemize}
 
-\noindent
-\note All proofs appear in the Appendix (submitted an a permitted attachment via EasyChair).
+% \medskip\noindent
+% \note All proofs appear in the Appendix.
+% \medskip
+
+% \noindent
+% \note All proofs appear in the Appendix.
+
+\sidenote{All proofs appear in \ref{sec:appendix:a}.}

### 2015-02-16 doc/GMP2015/revision_1/s02_preliminaries.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

@@
 %
-In this paper, a \dfn{geometric constraint system} is a multivariate
-polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$,
-representing constraints with parameters $\delta$ between geometric
-primitives  in $\mathbb{R}^2$ represented collectively as $x\in
-\mathbb{R}^n$.
+In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
 %
-When the type of constraint (system) is fixed, the system is simply
-represented as $(G,\delta)$, where $G$ is the underlying constraint
-(hyper)graph $G = (V,E)$ with the vertices $V$ representing the
-geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$
-representing the constraints, each with an associated parameter
-$\delta$.
+When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where $G$ is the underlying constraint (hyper)graph $G = (V,E)$ with the vertices $V$ representing the geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$ representing the constraints, each with an associated parameter $\delta$.
 %
-For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a
-graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E
-\rightarrow \mathbb{R}$; this represents the distance constraint
-system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where
-$x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
+For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a graph $G=(V,E)$ with fixed length bars as edges, i.e.
+$\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
 
-\noindent
-\note Geometric constraint systems can also have
-inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
-values rather than exact values.
+% \noindent
+% \note Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.
 
-In this paper, we will consider 3 types of constraint systems:
-bar-joint as well as \dfn{pinned line-incidence} and \dfn{body-hyperpin}.
-In all of these cases, a Cartesian \dfn{realization} or \dfn{solution}
-$G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean
-transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or
-$\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with
-parameters $\delta$, modulo orientation preserving isometries
-(Euclidean rigid body motions for distance constraint).
+\sidenote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}
 
-Although the realization space itself depends on the constraint
-parameters $\delta$, many relevant \dfn{generic} properties of the
-constraint system $G(x,\delta)$ are properties of the constraint
-(hyper)graph $G$ and do not depend on $\delta$ (or they hold for all
-but a measure zero set of $\delta$ values). Many of these are
-properties of the Jacobian $\Delta_x G(x,\delta)$, often called the
-appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
-For example, the \dfn{bar-joint rigidity matrix of the graph $G =
-(V,E)$} is a matrix of indeterminates representing the Jacobian of the
-distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$
-columns per vertex in $V$ and one row per edge in $E$, where the row
-corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates
-for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\
-$v$), that is 4 non-zero entries per row.
+In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{pinned line-incidence} and \dfn{body-hyperpin}.
+In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
+
+Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values).
+Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
+For example, the \dfn{bar-joint rigidity matrix of the graph $G = (V,E)$} is a matrix of indeterminates representing the Jacobian of the distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$.
+The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp.
+$x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.
 
@@
 %that depends on the appropriate rigidity matrix of indeterminates
-is \dfn{rigidity} (\note we refer to these as properties of
-the constraint system or as properties of the underlying (hyper)graph
-interchangeably), i.e., the realizations or solutions of the
-corresponding constraint system being generically isolated and zero-
-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint
-(hyper)graph is rigid if it is generically \dfn{infinitesimally
-rigid}, i.e.\ the number of independent rows of its appropriate
-rigidity matrix is at least the number of columns less the number of
-rigid body motions, which is 3 for bar-joint systems.
+is \dfn{rigidity} (\note we refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably), i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
+The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
 %This condition is called generic {\em infinitesimal rigidity}.
 
-\noindent
-\note If the constraint system has inequalities, i.e, if the $\delta$
-lie in an interval, the definition of rigidity is approximate: the
-solutions are isolated small, full-dimensional  connected components.
+% \noindent
+% \note If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.
+
+\sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
 % generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
 
-Other generic constraint system or (hyper)graph properties are the
-following:
+Other generic constraint system or (hyper)graph properties are the following:
 % (\note we refer to these as properties of the constraint
@@
 % interchangeably).
-A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate
-rigidity matrix of indeterminates has independent rows (i.e, the
-determinant of some square submatrix is not identically zero).
+A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
 %It is {\em rigid} if the number of independent rows of the rigidity matrix is
@@
 %which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is
-both rigid and independent.
+It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
 %if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
@@
 %The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
-It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
-is independent and not rigid, or \dfn{overconstrained} if it is not
-independent.
+It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it is independent and not rigid, or \dfn{overconstrained} if it is not independent.
 
-Defining the combinatorial independence of a subset of edges $E'
-\subseteq E$ to be the generic independence of corresponding rows in
-the appropriate rigidity matrix of indeterminates, we obtain the
-\dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
+Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
 %where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
 %if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-There are various results on combinatorial characterization of
-independence and rigidity  and rigidity matroids for different types
-of graphs. For bar-joint rigidity matroid, the famous Laman's theorem
-\cite{laman1970graphs} states that the underlying graph is isostatic if and only if
-$|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
-at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows
-that all 6-vertex-connected graphs are rigid in the plane. For bar-
-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying
-multigraph is isostatic if and only if it can be decomposed as $3$
-edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the
-same characterization using a different technique to study the
-algebraic-geometric conditions of genericity, called pure condition.
-Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity
-matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if
-$|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least
-2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| =
-k|V| - l$. In general, given a $d$-uniform hypergraph, a
-$(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
+There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs.
+For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices.
+The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane.
+For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees.
+White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition.
+Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$.
+In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
 
-In this paper, a \dfn{qusecs} is any \dfn{independent} geometric
-constraint system of one of the 3 types mentioned above.
+In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.
 
-\medskip\noindent
-\note In the remainder of this section and the next three
-sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
-qusecs and graphs. Relevant formal analogies for the other 2 types of
-qusecs and (hyper)graphs are given in the subsequent 2 sections on
-Applications.
+% \medskip\noindent
+% \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.
+
+\sidenote{In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.}
 
@@
 %TODO
-Given a bar-joint graph, a vector (of indeterminates) in the right
-null space of its rigidity matrix is called a \dfn{flex vector}. It
-has 2 entries per vertex and represents the internal motions of the
-system (i.e.\ modulo rigid body motions). For rigid graphs, all flex
-vectors are identically zero.
+Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}.
+It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions).
+For rigid graphs, all flex vectors are identically zero.
 %
-A vector (of indeterminates) in the left null space of the rigidity
-matrix is called a \dfn{stress vector}. It has one entry per edge and
-represents a \dfn{self stress} of the system. For independent graphs,
-all stress vectors are identically zero.
+A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}.
+It has one entry per edge and represents a \dfn{self stress} of the system.
+For independent graphs, all stress vectors are identically zero.
 %
-An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of
-indeterminates) which specifies a 2D vector acting at each vertex such
-that the stress balance equation $sR = t$ holds, where $s$ is a vector
-of internal stresses, one for each bar, and $R$ is the rigidity
-matrix.
+An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
 
-The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set
-$X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
+The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set $X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
 
@@
 \begin{definition}\label{def:drp}
-    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
-    following properties:
+    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the following properties:
     \begin{enumerate}
-        \item Each node represents/contains/is a rigid vertex-induced
-        subgraph of $G$.
+        \item Each node represents/contains/is a rigid vertex-induced subgraph of $G$.
         % \item For a node, $C$, that is  a non-trivial graph, its $N$
@@
         \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
-        \item A leaf node is a single vertex.
-A \dfn{trivial} graph is empty or a single vertex.
+        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
         \item A root node is a maximal rigid subgraph of $G$.
@@
     A DR-plan is \dfn{complete} if it satisfies an additional rule:
-        \item For a node $C$ that is  a non-trivial graph, its $N$
-        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
-        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
-        and in fact, $C$ contains all the edges in the union of its
-        children as well.
+        \item For a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$.
+This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well.
 We denote a complete DR-plan of $G$ as $\comdrp{G}$
@@
 \begin{remark}
-    For a given graph, there could be exponentially many  DR-plans and
-    even optimal DR-plans in the size of the graph. A complete DR-plan
-    is unique but may not be (is usually not) optimal. DR-plans
-    of self-similar graphs are self-similar. More than one node (leaf)
-    in a DR-plan forest may represent the same subgraph (vertex) of
-    $G$.
+    For a given graph, there could be exponentially many  DR-plans and even optimal DR-plans in the size of the graph.
+A complete DR-plan is unique but may not be (is usually not) optimal.
+DR-plans of self-similar graphs are self-similar.
+More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
 \end{remark}
@@
 
-\noindent
-\note In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.
+\sidenote{In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.}
 
 \subsection{Importance of an Optimal DR-plan}
-The use of a DR-plan in designing especially self-similar qusecs is
-immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and
-\ref{fig:subdivision}.
+The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:subdivision}.
 
-In fact, knowing an optimal DR-plan is crucial for
-design and analysis of even non-self-similar qusecs and their
-realizations with various desirable properties.
+In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
 
@@
 \header{Realizing given qusecs.}
-Note that this basic task
-would require finding real solutions a large multivariate polynomial
-system (of inequalities and equalities representing the constraints).
-Without DR-planning, this requires double exponential time in the
-number of variables (even if orientation type is specified). With
-DR-planning, the complexity is dominated by the size of the largest
-subsystem that is solved, or recombined from the solutions of its
-child subsystems i.e.,the maximum fan-in occurring in a DR-plan. Thus
-finding an optimal DR-plan is important for realization.
+Note that this basic task would require finding real solutions a large multivariate polynomial system (of inequalities and equalities representing the constraints).
+Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
+With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems i.e.,the maximum fan-in occurring in a DR-plan.
+Thus finding an optimal DR-plan is important for realization.
 
 \medskip\noindent
-\header{Recursive block
-decomposition of the rigidity matrix, stress space and external stresses.}
-This can be obtained using the DR-plan we obtain a
-with each block being the
-rigidity matrix of an isostatic subgraph. The blocks intersect on
-isostatic or trivial subgraphs. This recursive block decomposition of
-the rigidity matrix $R$ correspondingly recursively decomposes
-the stress vectors $s$ that balance a given external stress $t$ since
-$sR = t$. Moreover, the external stress vector $t$ is
-recursively ``distributed'' into external stresses for each
-node of the DR-plan.
-When the rigidity matrix and stress
-vectors contain indeterminates, this  can be used to design
-underlying graph and parameters $\delta$ to obtain an optimal
-distribution of stresses on the geometric primitives. When the
-realization of qusecs is given, this can be used to analyze the
-distribution of a given external load.
+\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
+This can be obtained using the DR-plan we obtain a with each block being the rigidity matrix of an isostatic subgraph.
+The blocks intersect on isostatic or trivial subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
+Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
+When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
+When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.
 

### 2015-02-16 doc/GMP2015/revision_1/s03_drp.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2015-02-15 doc/GMP2015/original_submission/s03_drp.tex
@@
 
-The NP-hardness of the DR-planning problem for 2D bar-joint graphs is partly
-the consequence of possibly exponential number of DR-plans. On
-the other hand, although the complete DR-plan is unique
-it could have large average fan-in and exponentially many nodes making it
-far from optimal.
+The NP-hardness of the DR-planning problem for 2D bar-joint graphs is partly the consequence of possibly exponential number of DR-plans.
+On the other hand, although the complete DR-plan is unique it could have large average fan-in and exponentially many nodes making it far from optimal.
 
-In this section we define a \dfn{canonical} plan to capture those aspects of an optimal
-DR-plan that mimic a complete DR-plan's uniqueness, while for independent
-(underconstrained or isostatic graphs), we show that the nonunique parts do not affect
-optimality.
+In this section we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.
 
@@
 
-This theorem (proven in the Appendix) relies on the following crucial observation and lemma
-that are also used for other types of qusecs in the Application sections (\ref{sec:bodypin}, \ref{sec:pinnedline}) of
-the paper.
+This theorem (proven in the \ref{theorem:main:proof}) relies on the following crucial observation and lemma that are also used for other types of qusecs in the Application sections (\ref{sec:bodypin}, \ref{sec:pinnedline}) of the paper.
 
@@
     \item\label{lemma:wc_intersection_makes_all_wc}
-    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
+    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
 
@@
 \begin{remark}
-The first item in the above lemma generalizes to the union of any number of
-children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{Cluster
-Minimality},
-defined in \cite{hoffman2001decompositionI}
-holding for these canonical-optimal DR-plans.
+The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{Cluster Minimality}, defined in \cite{hoffman2001decompositionI} holding for these canonical-optimal DR-plans.
 \end{remark}
 
-\begin{figure*}\centering
-\begin{subfigure}{.3\linewidth}\centering
-    \input{img/tikz/c2c3_of_k33s-0.tikz}
+% \begin{figure*}\centering
+% \begin{subfigure}{.3\linewidth}\centering
+%     \input{img/tikz/c2c3_of_k33s-0.tikz}
+%     \caption{}\label{fig:c2c3ofk33s:a}
+% \end{subfigure}%
+% \begin{subfigure}{.7\linewidth}\centering
+%     \input{img/tikz/c2c3_of_k33s-1.tikz}
+%     \caption{}\label{fig:c2c3ofk33s:b}
+% \end{subfigure}
+
+% \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
+This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
+(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
+The dashed lines indicated that this exact structure is repeated.}
+% \label{fig:c2c3ofk33s}
+% \end{figure*}
+
+% \FigInit
+%     {(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
+This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
+(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
+The dashed lines indicated that this exact structure is repeated.}
+%     {fig:c2c3ofk33s}
+% \FigTwoSubfigWithWidth
+%     {.3}
+%     {img/epsfromtikz/c2c3_of_k33s-0}
+%     {}
+%     {fig:c2c3ofk33s:a}
+%     %
+%     {.7}
+%     {img/epsfromtikz/c2c3_of_k33s-1}
+%     {}
+%     {fig:c2c3ofk33s:b}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.3}{img/epsfromtikz/c2c3_of_k33s-0}
+\SetMyMinHeight{.7}{img/epsfromtikz/c2c3_of_k33s-1}
+
+\begin{figure*}\centering%
+    %
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-0}
         \caption{}\label{fig:c2c3ofk33s:a}
     \end{subfigure}%
-\begin{subfigure}{.7\linewidth}\centering
-    \input{img/tikz/c2c3_of_k33s-1.tikz}
+    %
+    \hfill
+    \begin{subfigure}{0.7\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-1}
         \caption{}\label{fig:c2c3ofk33s:b}
-\end{subfigure}
-
+    \end{subfigure}%
+    %
     \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
@@
     \label{fig:c2c3ofk33s}
-\end{figure*}
+\end{figure*}%
 

### 2015-02-16 doc/GMP2015/revision_1/s04_recombination.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2014-12-22 doc/GMP2015/s04_recombination.tex
@@
 %
-In this section, we consider the \dfn{optimal recombination} problem
-of combining specific solutions of subsystems in a DR-plan into a
-solution of their parent system $I$ (without loss of generality, at
-the top level of the DR-plan). In the case of isostatic qusecs, the
-parent system $I$ is isostatic (the root of the DR-plan), and we seek
-\vemph{solution(s) (among a finite large number of solutions) with a
-specific orientation or chirality}. In the case of underconstrained
-qusecs the subsystems are the multiple roots of the DR-plan, the
-parent system $I$ is underconstrained, and we typically seek an
-efficient algorithmic description of \vemph{connected component(s) of
-solutions with a specific orientation or chirality}.
+In this section, we consider the \dfn{optimal recombination} problem of combining specific solutions of subsystems in a DR-plan into a solution of their parent system $I$ (without loss of generality, at the top level of the DR-plan).
+In the case of isostatic qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we seek \vemph{solution(s) (among a finite large number of solutions) with a specific orientation or chirality}.
+In the case of underconstrained qusecs the subsystems are the multiple roots of the DR-plan, the parent system $I$ is underconstrained, and we typically seek an efficient algorithmic description of \vemph{connected component(s) of solutions with a specific orientation or chirality}.
 
-The main barrier in recombination when given an optimal DR-plan (of
-smallest possible size or max fan-in),  is that the number of children
-(resp. number of roots) --- and correspondingly the  size and complexity
-of the (indecomposable) algebraic system $I$ to be solved --- could be
-arbitrarily large as a function of the size of the input system. This
-difficulty can persist even after optimal parametrization of the
-indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic
-complexity.
+The main barrier in recombination when given an optimal DR-plan (of smallest possible size or max fan-in),  is that the number of children (resp.
+number of roots)---and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved---could be arbitrarily large as a function of the size of the input system.
+This difficulty can persist even after optimal parametrization of the indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic complexity.
 
-In the following, we formulate the problem of \dfn{optimal
-modification}, of the indecomposable algebraic system $I$ into a
-decomposable system with a small DR-plan (low algebraic complexity).
-Leveraging recent results on \dfn{Cayley configuration spaces}, our
-approach to the optimal modification problem achieves the following:
+In the following, we formulate the problem of \dfn{optimal modification}, of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
+Leveraging recent results on \dfn{Cayley configuration spaces}, our approach to the optimal modification problem achieves the following:
 
 \medskip\noindent
-\header{(a) Small DR-plan.} We obtain a  parameterized family of systems
-$I_{\lambda_F}$ --- one for each value $\lambda_F$ for the parameters
-$F$,  all of which have small DR-plans. Thus, given a value $v$ for
-$\lambda_F$, the system $I_v$ can potentially be solved or realized
-easily once the orientation type of the solution is known  (when the
-DR-plan size is small enough).
+\header{(a) Small DR-plan.}
+We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans.
+Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can potentially be solved or realized easily once the orientation type of the solution is known  (when the DR-plan size is small enough).
 
@@
 \header{(b) Solution preservation.}
-Moreover, the union of solution spaces of the systems in the family
-$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.
+Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.
 
 \medskip\noindent
-\header{(c) Efficient search.} Finally, this union can be projected into the
-so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
-convex or otherwise easy to traverse in order to search for $I$'s
-solution (or connected component) of the desired orientation type. For
-the case when the modification is bounded, this approach provides an
-efficient algorithm for recombination. We first define the decision
-version of the problem of optimal modification for decomposition. The
-standard optimization versions are straightforward.
+\header{(c) Efficient search.}
+Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type.
+For the case when the modification is bounded, this approach provides an efficient algorithm for recombination.
+We first define the decision version of the problem of optimal modification for decomposition.
+The standard optimization versions are straightforward.
 
 \problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
-Given a generically independent graph $G = (V,E)$ with no non-trivial
-proper isostatic subgraph (indecomposable) and 2 constants $k$ and
-$s$, does there exist a set of at most $k$  edges $E_1$ and a set of
-non-edges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
-F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
-where $k$ is a fixed bound (not part of the input).   We say that such
-a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to
-graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
-with appropriately small $s$.
+Given a generically independent graph $G = (V,E)$ with no non-trivial proper isostatic subgraph (indecomposable) and 2 constants $k$ and $s$, does there exist a set of at most $k$  edges $E_1$ and a set of non-edges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup F)$ has a DR-plan of size at most $s$?
+The OMD$_k$ problem is  OMD where $k$ is a fixed bound (not part of the input).
+We say that such a tuple $(G,s)$ is a member of the set OMD$_k$.
+We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs with appropriately small $s$.
 
-It is immediately clear that indecomposable graphs $G$ that belong in
-OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
-decomposable graphs satisfying Criteria (a) and (b) above. However, it
-is not clear how Criterion (c) is met by OMD graphs.
+It is immediately clear that indecomposable graphs $G$ that belong in OMD$_k$ for small $k$ and $s$  lend themselves to modification  into decomposable graphs satisfying Criteria (a) and (b) above.
+However, it is not clear how Criterion (c) is met by OMD graphs.
 %
@@
 %
-Next we provide the necessary background to describe a specific
-approach for achieving the requirements (a)--(c) mentioned above, by
-restricting the class of reduced graphs $G' = G\setminus E_1$ and
-their isostatic completions $H$ in the above definition of the OMD
-problem, and using a key theorem of Convex Cayley configuration spaces
-\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors
-$\lambda_F$ of  attainable lengths of the non-edges $F$ is always
-convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
-all the realizations of the bar-joint constraint system or linkage
-$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
-\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
-Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
-a ``projection'' of the cartesian realization space of $(H,\delta)$ on
-the Cayley parameters $F$. Such graphs $H$ are said to have
-\dfn{convexifiable Cayley configuration spaces with parameters $F$}
-(short: \dfn{convexifiable}).
+Next we provide the necessary background to describe a specific approach for achieving the requirements (a)--(c) mentioned above, by restricting the class of reduced graphs $G' = G\setminus E_1$ and their isostatic completions $H$ in the above definition of the OMD problem, and using a key theorem of Convex Cayley configuration spaces \cite{sitharam2010convex}.
+This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of the non-edges $F$ is always convex for any given lengths $\delta$ for the edges of $H$ (i.e, over all the realizations of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
+This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$.
+Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces with parameters $F$} (short: \dfn{convexifiable}).
 
-To state the theorem, we first have to define the notion of
-\dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and $H_2$ be two graphs on
-disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and
-$E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
-A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the
-union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$. In this
-case, $H_1$ and $H_2$ are called \dfn{2-sum components} of $H$. A
-\dfn{minimal 2-sum component} of $H$ is  one that cannot be further
-split into 2-sum components. A \dfn{2-tree} is recursively obtained by
-taking a 2-sum of 2-trees, with the base case of a 2-tree being a
-triangle. A \dfn{partial 2-tree} is a 2-tree minus some edges. Partial
-2-trees have an alternate characterization as the graphs that avoid
-$K_4$ minors, and are also called series parallel graphs.
+To state the theorem, we first have to define the notion of \dfn{2-sums} and \dfn{2-trees}.
+Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
+A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$.
+In this case, $H_1$ and $H_2$ are called \dfn{2-sum components} of $H$.
+A \dfn{minimal 2-sum component} of $H$ is  one that cannot be further split into 2-sum components.
+A \dfn{2-tree} is recursively obtained by taking a 2-sum of 2-trees, with the base case of a 2-tree being a triangle.
+A \dfn{partial 2-tree} is a 2-tree minus some edges.
+Partial 2-trees have an alternate characterization as the graphs that avoid $K_4$ minors, and are also called series parallel graphs.
 
 \begin{theorem}\label{theorem:convexcayley}
-    \cite{sitharam2010convex} $H$ has a convexifiable Cayley configuration space  with
-    parameters $F$ if and only if for each $f\in F$  all the minimal
-    2-sum components of $H\cup F$ that contain both endpoints of $f$
-    are partial 2-trees. The Cayley configuration space
-    $\Phi_F(H,\delta)$ of a bar-joint system or linkage $(H,\delta)$
-    is a convex polytope. When $H\cup F$ is a 2-tree, the bounding
-    hyperplanes of this polytope are triangle inequalities relating
-    the lengths of edges of the triangles in $H\cup F$.
+    \cite{sitharam2010convex} $H$ has a convexifiable Cayley configuration space  with parameters $F$ if and only if for each $f\in F$  all the minimal 2-sum components of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees.
+The Cayley configuration space $\Phi_F(H,\delta)$ of a bar-joint system or linkage $(H,\delta)$ is a convex polytope.
+When $H\cup F$ is a 2-tree, the bounding hyperplanes of this polytope are triangle inequalities relating the lengths of edges of the triangles in $H\cup F$.
 \end{theorem}
 
-The idea of our approach to achieve the criteria (a)--(c) begins with
-the following simple but useful theorem.
+The idea of our approach to achieve the criteria (a)--(c) begins with the following simple but useful theorem.
 
 \begin{theorem}\label{theorem:omdk}
-    Given an indecomposable graph $G$, let $G'$ be a spanning partial
-    2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then
-    $(G,2)$ belongs in the set OMD$_k$.
+    Given an indecomposable graph $G$, let $G'$ be a spanning partial 2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then $(G,2)$ belongs in the set OMD$_k$.
 \end{theorem}
@@
 
-We refer to such graphs $G$ in short as \dfn{$k$-approximately
-convexifiable}, where the reduced graphs $G'$ and isostatic
-completions $H$ are convexifiable. As observed earlier, since graphs
-such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
-for small enough $k$. Criterion (c) is addressed as described in the
-following efficient search procedure which clarifies the dependence of
-the complexity on the number and ranges  of Cayley parameters $F$.
+We refer to such graphs $G$ in short as \dfn{$k$-approximately convexifiable}, where the reduced graphs $G'$ and isostatic completions $H$ are convexifiable.
+As observed earlier, since graphs such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met for small enough $k$.
+Criterion (c) is addressed as described in the following efficient search procedure which clarifies the dependence of the complexity on the number and ranges  of Cayley parameters $F$.
 
@@
     [Efficient search]
-    For an indecomposable, $k$-approximately convexifiable graph $G =
-    (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
-    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of non-edges
-    of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
-    $p$ (or connected component of a solution space) of $(G,\delta)$
-    of an orientation type $\sigma_p$ can be found in time
-    $O(\log(W))$ where $W$ is the number of cells of desired accuracy
-    (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
-    The (discrete) volume $W$ is exponential in $|F|$ and polynomial
-    in the (discrete range) of the parameters in $F$.
+    For an indecomposable, $k$-approximately convexifiable graph $G = (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph where $|D| \le  k$.
+Let  $F$ be a set of non-edges of $G$ such that $H = (V, E'\cup F)$ is a 2-tree.
+Each solution $p$ (or connected component of a solution space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be found in time $O(\log(W))$ where $W$ is the number of cells of desired accuracy (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
+The (discrete) volume $W$ is exponential in $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
 \end{theorem}
@@
 %
+% \begin{figure*}\centering
+% \begin{subfigure}{.2\linewidth}\centering
+%     \input{img/tikz/omd_k33_example_new-0.tikz}
+%     \caption{}\label{fig:omd_k33_example:a}
+% \end{subfigure}\hfill
+% \begin{subfigure}{.15\linewidth}\centering
+%     \input{img/tikz/omd_k33_example_new-1.tikz}
+%     \caption{}\label{fig:omd_k33_example:b}
+% \end{subfigure}
+% \begin{subfigure}{.15\linewidth}\centering
+%     \input{img/tikz/omd_k33_example_new-2.tikz}
+%     \caption{}\label{fig:omd_k33_example:c}
+% \end{subfigure}\hfill
+% \begin{subfigure}{.15\linewidth}\centering
+%     \input{img/tikz/omd_k33_example-1.tikz}
+%     \caption{}\label{fig:omd_k33_example:d}
+% \end{subfigure}
+% \begin{subfigure}{.15\linewidth}\centering
+%     \input{img/tikz/omd_k33_example-2.tikz}
+%     \caption{}\label{fig:omd_k33_example:e}
+% \end{subfigure}
+
+% \caption{
+% (\ref{fig:omd_k33_example:a}) The $K_{3,3}$ with two labeled edges, $e_1$ and $e_2$.
+(\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree.
+(\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$.
+(\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line).
+(\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
+% }
+% \label{fig:omd_k33_example}
+% \end{figure*}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.2}{img/epsfromtikz/omd_k33_example_new-0}
+\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-1}
+\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-2}
+\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-1}
+\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-2}
+
 \begin{figure*}\centering
     \begin{subfigure}{.2\linewidth}\centering
-    \input{img/tikz/omd_k33_example_new-0.tikz}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-0}
         \caption{}\label{fig:omd_k33_example:a}
@@
     \begin{subfigure}{.15\linewidth}\centering
-    \input{img/tikz/omd_k33_example_new-1.tikz}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-1}
         \caption{}\label{fig:omd_k33_example:b}
@@
     \begin{subfigure}{.15\linewidth}\centering
-    \input{img/tikz/omd_k33_example_new-2.tikz}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-2}
         \caption{}\label{fig:omd_k33_example:c}
@@
     \begin{subfigure}{.15\linewidth}\centering
-    \input{img/tikz/omd_k33_example-1.tikz}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-1}
         \caption{}\label{fig:omd_k33_example:d}
@@
     \begin{subfigure}{.15\linewidth}\centering
-    \input{img/tikz/omd_k33_example-2.tikz}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-2}
         \caption{}\label{fig:omd_k33_example:e}
@@
 
-\noindent
-\note A major advantage of the convex Cayley method is that it is
-completely unaffected when $\delta$ are intervals rather than exact
-values~~\cite{sitharam2010convex}.
+% \noindent
+% \note A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.
+
+\sidenote{A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.}
 %cite my paper with Heping!!!
@@
 [Using Cayley configuration space].
-A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
-isostatic graphs, i.e, its DR-plan has a root and 9 children
-corresponding to the 9 edges. Solving or recombining the system
-$(G,\delta)$ corresponding to the root of this DR-plan involves
-solving a quadratic system with 8 equations and variables. Instead of
-simultaneously solving this system, we could instead use the fact that
-$G=K_{3,3}$ is in OMD$_2$: remove the edges $e_1,e_2$ in Figure
-\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
-nonedges $f_1,f_2$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
-configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
-interval of attainable length values $\lambda_F$ for the edge $f$.
-When $\delta$ is generic, i.e, does not admit collinearities or
-coincidences in the realizations of $(G,\delta)$, the realization
-space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
-solutions $q^p_{\lambda_f}$ (modulo orientation preserving
-isometries), with distinct orientation types $\sigma_p$ (two
-orientation choices for each of the 4 triangles) that can be obtained
-by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
-size 2). By subdivided binary search in the interval $\lambda_f \in
-\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
-$(G,\delta)$ is found when the length of the nonedge $e$ in  the
-realization $q^p_{\lambda_f}$ is $\delta_e$.
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e, its DR-plan has a root and 9 children corresponding to the 9 edges.
+Solving or recombining the system $(G,\delta)$ corresponding to the root of this DR-plan involves solving a quadratic system with 8 equations and variables.
+Instead of simultaneously solving this system, we could instead use the fact that $G=K_{3,3}$ is in OMD$_2$: remove the edges $e_1,e_2$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
+Now add the non-edges $f_1,f_2$ to give a 2-tree $H$ with a DR-plan of size 2.
+The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single interval of attainable length values $\lambda_F$ for the edge $f$.
+When $\delta$ is generic, i.e, does not admit collinearities or coincidences in the realizations of $(G,\delta)$, the realization space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16 solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries), with distinct orientation types $\sigma_p$ (two orientation choices for each of the 4 triangles) that can be obtained by solving a sequence of 4 single quadratics in 1 variable (DR-plan of size 2).
+By subdivided binary search in the interval $\lambda_f \in \Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is found when the length of the nonedge $e$ in  the realization $q^p_{\lambda_f}$ is $\delta_e$.
 
-In fact, we can show that $G=K_{3,3}$ is in OMD$_1$ by removing a single
-edge to reduce (as shown in Figure \ref{fig:omd_k33_example}) to a
-tree-decomposable graph of low Cayley complexity (which includes the class
-of partial 2-trees).
-In the next Section \ref{sec:tdecomp}, we discuss this issue
-of why the largest class of reduced graphs is desirable.
+In fact, we can show that $G=K_{3,3}$ is in OMD$_1$ by removing a single edge to reduce (as shown in Figure \ref{fig:omd_k33_example}) to a tree-decomposable graph of low Cayley complexity (which includes the class of partial 2-trees).
+In the next Section \ref{sec:tdecomp}, we discuss this issue of why the largest class of reduced graphs is desirable.
 %
@@
 %
-It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
-the number of edges to be removed to ensure an isostatic completion
-that is decomposable with a small DR-plan) by considering reduced
-graphs $G'$ (and modified graphs $H$) that come from a larger class
-than partial 2-trees but nevertheless have convex Cayley configuration
-spaces at least when the realization space is restricted to a
-sufficiently comprehensive orientation type. In particular, the so-
-called \dfn{tree-decomposable graphs of low Cayley complexity}
-\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
-partial 2-trees. See an illustration in Figure \ref{fig:omd_k33_example}.
-These too result in DR-plans of size 2 or 3, putting
-$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
-Criterion (c) is met --- for example when $k=1$ --- because 1-dof Cayley
-configuration spaces of linkages based on such graphs $G'$ are known
-to be single intervals when a comprehensive orientation type
-$\sigma_p$ of the sought solution $p$ is given. In addition, the
-bounds of these intervals are of low algebraic complexity. More
-precisely, the bounds  can themselves be computed using a DR-plan of
-size 2 or 3, i.e, the computation of these bounds is tree-
-decomposable. Alternatively, the bounds are in a simple quadratic or
-radically solvable extension field over the rationals, or can be
-computed by solving a triangularized system of quadratics.
+It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e, the number of edges to be removed to ensure an isostatic completion that is decomposable with a small DR-plan) by considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than partial 2-trees but nevertheless have convex Cayley configuration spaces at least when the realization space is restricted to a sufficiently comprehensive orientation type.
+In particular, the so-called \dfn{tree-decomposable graphs of low Cayley complexity} \cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not partial 2-trees.
+See an illustration in Figure \ref{fig:omd_k33_example}.
+These too result in DR-plans of size 2 or 3, putting $G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b).
+The Criterion (c) is met---for example when $k=1$---because 1-dof Cayley configuration spaces of linkages based on such graphs $G'$ are known to be single intervals when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
+In addition, the bounds of these intervals are of low algebraic complexity.
+More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e, the computation of these bounds is tree-decomposable.
+Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or can be computed by solving a triangularized system of quadratics.

### 2015-02-16 doc/GMP2015/revision_1/s05_problems.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2015-02-15 doc/GMP2015/original_submission/s05_problems.tex
@@
 \label{sec:table}
-In this section we provide a unified view of  the various problems
-studied in the previous 2 sections, along with formal reductions
-between them, and relationships to other known problems and results as
-well as open questions.
+In this section we provide a unified view of  the various problems studied in the previous 2 sections, along with formal reductions between them, and relationships to other known problems and results as well as open questions.
 %
 \subsection{Special classes of small DR-plans}
-As seen in the previous section, 2-trees and tree-decomposable graphs
-have not only small, but also special DR-plans that permit easy
-solving --- essentially a single quadratic at a time.
+As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially a single quadratic at a time.
 
-The \dfn{restricted optimal DR-planning problem} requires DR-plans of
-one of these types, which reduces to recognizing if the input graph is
-a 2-tree or a tree-decomposable graph for which simple near-linear
-time algorithms are available \cite{valdes1979recognition,fudos1997graph} and the DR-plan is a
-by-product output of the recognition algorithm.
+The \dfn{restricted optimal DR-planning problem} requires DR-plans of one of these types, which reduces to recognizing if the input graph is a 2-tree or a tree-decomposable graph for which simple near-linear time algorithms are available \cite{valdes1979recognition,fudos1997graph} and the DR-plan is a by-product output of the recognition algorithm.
 
-In the recombination setting, the corresponding \dfn{restricted
-OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
-completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
-or to be a low Cayley complexity tree-decomposable graph as in Section
-\ref{sec:tdecomp}. Clearly these problems have deterministic
-polynomial time algorithms, the algorithms run in time exponential in
-$k$.
+In the recombination setting, the corresponding \dfn{restricted OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction} or to be a low Cayley complexity tree-decomposable graph as in Section \ref{sec:tdecomp}.
+Clearly these problems have deterministic polynomial time algorithms, the algorithms run in time exponential in $k$.
 
-We discuss the complexity of the restricted
-OMD problem (when $k$ is part of the input)
-in Section \ref{sec:futurework}.
+We discuss the complexity of the restricted OMD problem (when $k$ is part of the input) in \ref{sec:futurework}.
 
@@
 %
-The OMD problem is closely related to a well-studied problem of
-completion of an underconstrained system to an isostatic one with a
-small DR-plan.
+The OMD problem is closely related to a well-studied problem of completion of an underconstrained system to an isostatic one with a small DR-plan.
 \begin{observation}\label{obs:OC_to_OMD}
-    The (decision version of) the \textbf{optimal completion problem
-    (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
+    The (decision version of) the \textbf{optimal completion problem (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
 \end{observation}
-In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming}
-requiring the completion to be tree-decomposable.
+In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming} requiring the completion to be tree-decomposable.
 %It is not clear if
@@
 
-We now connect the OMD problem to the informal \dfn{optimal
-recombination (OR)} problem mentioned as motivation at the beginning
-of Section \ref{sec:recomb}.
+We now connect the OMD problem to the informal \dfn{optimal recombination (OR)} problem mentioned as motivation at the beginning of Section \ref{sec:recomb}.
 
-In order to connect the OR problem to OMD, when the input graph is
-the isostatic graph at the DR-plan root, we do not consider the case
-where the two child \dfn{solved subgraphs} (corresponding to already
-solved subsystems) have a nontrivial intersection (in this case the
-recombination is trivial). We only consider the case where no two
-child solved subgraphs (resp. two root subgraphs when the input graph
-is underconstrained) share more than 1 vertex. We replace such solved
-subgraphs  by isostatic graphs as follows. If a solved subgraph shares
-at most one vertex with the remainder of the graph, simply replace it
-by an edge one of whose endpoints  is the shared vertex. Otherwise,
-replace it by  a 2-tree graph of the shared vertices. Finally, we add
-the additional restriction to the OM Problem that when any edge in a
-solved subgraph is chosen among the $k$ edges to be removed, in fact
-the entire solved subgraph must be removed  and all of its edges must
-be counted in $k$.
+In order to connect the OR problem to OMD, when the input graph is the isostatic graph at the DR-plan root, we do not consider the case where the two child \dfn{solved subgraphs} (corresponding to already solved subsystems) have a nontrivial intersection (in this case the recombination is trivial).
+We only consider the case where no two child solved subgraphs (resp.
+two root subgraphs when the input graph is underconstrained) share more than 1 vertex.
+We replace such solved subgraphs  by isostatic graphs as follows.
+If a solved subgraph shares at most one vertex with the remainder of the graph, simply replace it by an edge one of whose endpoints  is the shared vertex.
+Otherwise, replace it by  a 2-tree graph of the shared vertices.
+Finally, we add the additional restriction to the OM Problem that when any edge in a solved subgraph is chosen among the $k$ edges to be removed, in fact the entire solved subgraph must be removed  and all of its edges must be counted in $k$.
 
-This reduction is used also for transferring  algorithms for optimal
-DR-planning, recombination, completion, OMD, and other problems
-from bar-joint systems to
-so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin},
-by showing that the problems for the latter are reduced
-to the corresponding problems on bar-joint systems.
+This reduction is used also for transferring  algorithms for optimal DR-planning, recombination, completion, OMD, and other problems from bar-joint systems to so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin}, by showing that the problems for the latter are reduced to the corresponding problems on bar-joint systems.
 

### 2015-02-16 doc/GMP2015/revision_1/s06_problempinning.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2015-02-15 doc/GMP2015/original_submission/s07_problempinning.tex
@@
 \begin{figure}\centering
-    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
+    \includegraphics[width=0.4\linewidth]{img/Silica}\hspace{0.5cm}
     \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
-    \caption{Example of a single monolayer of a Silica (silicon and oxygen)
-    glassy structure. This can be thought of as a triangle-multipin qusec
-where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
-Not shown here are other monolayers stacked on this one. Each Silicon atom
-has the structure seen to the right, and so binds to another Oxygen atom in
-an adjacent monolayer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
+    \caption{Example of a single monolayer of a Silica (silicon and oxygen) glassy structure.
+This can be thought of as a triangle-multipin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
+Not shown here are other monolayers stacked on this one.
+Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent monolayer.
+Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
     \label{fig:silica_glass}
@@
 
-We can use qusecs DR-plans to design materials such as disordered
-graphene and silica bi-layers \cite{silica_bilayers}
-\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
-a somewhat more general setting:
-the problem of finding boundary conditions (additional constraints) to
-add to an underconstrained monolayer to make it
-isostatic.
-This can be done in a number of ways:
-(1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass});
-(2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or
-(3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).
+We can use qusecs DR-plans to design materials such as disordered graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}.
+We will investigate a more specific problem in a somewhat more general setting: the problem of finding boundary conditions (additional constraints) to add to an underconstrained monolayer to make it isostatic.
+This can be done in a number of ways: (1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}); (2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or (3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).
 
-In all cases,
-we are specifically interested in how to add additional constraints
-resulting isostatic structure with a small DR-plan, so a realization can be
-found and stress and other property design can be efficient.
-To answer these questions, we first introduce the
-qusecs that are used to model these materials.
+In all cases, we are specifically interested in how to add additional constraints resulting isostatic structure with a small DR-plan, so a realization can be found and stress and other property design can be efficient.
+To answer these questions, we first introduce the qusecs that are used to model these materials.
 We will discuss Item 2 in detail in this section.
@@
     \end{subfigure}
-    \caption{Examples of self-similarity via repeated subdivision. In (a),
-    simple subdivision that does not guarantee isostaticity. In (b), a more
-complicated scheme is used, ensuring that the resulting graph is isostatic
-but not  tree-decomposable. Credit to \cite{subdivision_paper} for (a).}
+    \caption{Examples of self-similarity via repeated subdivision.
+In (a), simple subdivision that does not guarantee isostaticity.
+In (b), a more complicated scheme is used, ensuring that the resulting graph is isostatic but not  tree-decomposable.
+Credit to \cite{subdivision_paper} for (a).}
     \label{fig:subdivision}
@@
 
-\begin{remark}
-\label{rem:bodypin_is_barjoint}
+\begin{remark}\label{rem:bodypin_is_barjoint}
     A body-hyperpin qusecs is a special case of bar-joint qusecs of the
@@
 
-For the rest of this section, we will only be dealing with the DR-plan of
-such qusecs. As such, we will refer only to the underlying hypergraph of the qusecs. We now
-introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and
-5 in Section \ref{sec:intro}, for which the optimal completion problem is
-significantly easier.
+For the rest of this section, we will only be dealing with the DR-plan of such qusecs.
+As such, we will refer only to the underlying hypergraph of the qusecs.
+We now introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}, for which the optimal completion problem is significantly easier.
 
-\begin{definition}
-\label{def:body-pin}
+\begin{definition}\label{def:body-pin}
     A body-pin graph is a body-hyperpin graph with the following conditions:
@@
 
-    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
-    the bodies are the original bodies and each pin represents 2 bars from
-    one body to another. Such body-bar graphs with 1 and 2-dof can be
-    characterized by being $(3,4)$ and $(3,5)$-tight respectively
-    \cite{Lee:2007:PGA} \cite{streinu2009sparse},
-defined in Section  \ref{sec:prelim}.
+    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
+Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}, defined in Section  \ref{sec:prelim}.
 \end{definition}
@@
 \begin{definition}
-    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
-    is a triangle, i.e., it shares at most 3 pins with other bodies.
+    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares at most 3 pins with other bodies.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
@@
 
-Body-pin graphs are of particular interest to us in the context of Example 4
-in Section \ref{sec:intro}. Triangle-multipin graphs can be used to
-represent the silica bi-layers and glassy structures described in Example 5
-of Section \ref{sec:intro}, where each triangle is the junction of ``disks''
-in the plane (see Figure \ref{fig:silica_glass}). Typically, these systems
-are not isostatic, so to relate the work of this paper to the systems, we
-define a slightly different kind of DR-plan using the notion of
-$(k,l)$-sparsity and tightness.
+Body-pin graphs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}.
+Triangle-multipin graphs can be used to represent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (see Figure \ref{fig:silica_glass}).
+Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan using the notion of $(k,l)$-sparsity and tightness.
 
@@
 
-We can develop the notion of a canonical DR-plan when the hypergraph is
-$(k,l)$-sparse (i.e., combinatorially independent) using certain
-notion of trivial and non-trivial intersections and
-$(k,l)$-tightness conditions similar to the way we did in Section
-\ref{sec:DRP}, provided
-    such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
-    In particular, we can define canonical DR-plans with similar properties
-    for the 1 and 2-dof body-pin and triangle-hyperpin systems defined
-    above.
+We can develop the notion of a canonical DR-plan when the hypergraph is $(k,l)$-sparse (i.e., combinatorially independent) using certain notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}, provided such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
+In particular, we can define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.
 
@@
 
-The strong Church-Rosser property holds making all canonical DR-plans
-optimal:
+The strong Church-Rosser property holds making all canonical DR-plans optimal:
 %
 \begin{remark}\label{rem:1dofcanon}
-    When the input is independent,
-    all $(3,4)$-tight canonical DR-plans are optimal. We can find such a
-    DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
-    This is because
-    such $(3,4)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
+    When the input is independent, all $(3,4)$-tight canonical DR-plans are optimal.
+We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
+This is because such $(3,4)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
 \end{remark}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.2}{img/bodypin}
+\SetMyMinHeight{.2}{img/epsfromtikz/bodypin_graph}
+\SetMyMinHeight{.35}{img/epsfromtikz/bodypin_drp}
+\SetMyMinHeight{.2}{img/bodypin2}
 
@@
     \begin{subfigure}{0.2\linewidth}\centering
-    \includegraphics[width=\linewidth]{img/bodypin}
-    \caption{}
+        \includegraphics[height=\myMinHeight]{img/bodypin}
+        \caption{}\label{fig:bodypindrp:a}
     \end{subfigure}
@@
     \begin{subfigure}{0.2\linewidth}\centering
-    \input{img/tikz/bodypin_graph.tikz}
-    \caption{}
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/bodypin_graph}
+        \caption{}\label{fig:bodypindrp:b}
     \end{subfigure}
     %
-\begin{subfigure}{0.35\linewidth}\centering\scriptsize
-    \input{img/tikz/bodypin_drp.tikz}
-    \caption{}
+    \begin{subfigure}{0.35\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{img/epsfromtikz/bodypin_drp}
+        \caption{}\label{fig:bodypindrp:c}
     \end{subfigure}
@@
     \begin{subfigure}{0.2\linewidth}\centering
-    \includegraphics[width=\linewidth]{img/bodypin2}
-    \caption{}
+        \includegraphics[height=\myMinHeight]{img/bodypin2}
+        \caption{}\label{fig:bodypindrp:d}
     \end{subfigure}
     %
-\caption{(a) is a 1-dof body-pin graph.
-(b) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}.
-(c) is the 1-dof DR-plan for the graph.
+    \caption{(\ref{fig:bodypindrp:a}) is a 1-dof body-pin graph.
+(\ref{fig:bodypindrp:b}) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}.
+(\ref{fig:bodypindrp:c}) is the 1-dof DR-plan for the graph.
 In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level.
-(d) is the result of adding one such body to the bold-faced node.}
+(\ref{fig:bodypindrp:d}) is the result of adding one such body to the bold-faced node.}
     %
 \end{figure*}
+
 The above discussion leads to the main theorem:
@@
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding
-    1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal
-    completion problem of Section \ref{sec:recomb}.
+    Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb}.
 \end{theorem}
@@
 \label{obs:algebraic_completion}
-    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
-    the isostatic completion with minimum fan-in, (a reasonable
-    measure of algebraic complexity), a more nuanced
-    measure that treats solutions of 1-dof and 2-dof systems as
-    1- or 2- parameter families would no longer permit the algorithm given
-    in the proof. In particular the complexity of the standard algorithm in the k-dof
-    case would be exponential in $k$ (even if that case were matroidal and
-    an optimal DR-plan is known).
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1- or 2- parameter families would no longer permit the algorithm given in the proof.
+In particular the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if that case were matroidal and an optimal DR-plan is known).
 \end{observation}

### 2015-02-16 doc/GMP2015/revision_1/s07_pinnedline.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

--- 2014-12-22 doc/GMP2015/s08_pinnedline.tex
@@
 
-The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve
-the structure of  cross-linking collagen and cellulose microfibrils
-in animal and plant cell walls, and actin filaments in the cytoskeleton
-by modeling these structures as \dfn{pinned line-incidence systems}.
+The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of  cross-linking collagen and cellulose microfibrils in animal and plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as \dfn{pinned line-incidence systems}.
 
@@
 %It is additionally
-each of which is cross-linked at 2 places with usually 3 other fibrils,
-where the \dfn{cross-linking} is like an incidence constraint that the
-crosslinked fibrils can slide against each other while remaining incident (see
-Figure~\ref{fig:crosslink}).
+each of which is cross-linked at 2 places with usually 3 other fibrils, where the \dfn{cross-linking} is like an incidence constraint that the crosslinked fibrils can slide against each other while remaining incident (see Figure~\ref{fig:crosslink}).
 
@@
 
-\begin{figure}
-\centering
+% \begin{figure}
+% \centering
 
-\begin{subfigure}{.49\linewidth}
-  \centering
-  \includegraphics[width=.8\linewidth]{img/AFM_Innventia_nanocellulose}
-  \caption{}
-  \label{fig:cellulose}
-\end{subfigure}
-\begin{subfigure}{.49\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{img/crosslink}
-  \caption{}
-  \label{fig:crosslink}
+% \begin{subfigure}{.49\linewidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{img/AFM_Innventia_nanocellulose}
+%   \caption{}
+%   \label{fig:cellulose}
+% \end{subfigure}
+% \begin{subfigure}{.49\linewidth}
+%   \centering
+%   \includegraphics[width=\linewidth]{img/crosslink}
+%   \caption{}
+%   \label{fig:crosslink}
+% \end{subfigure}%
+% \caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
+(b) Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.}
+% \end{figure}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.45}{img/AFM_Innventia_nanocellulose}
+\SetMyMinHeight{.45}{img/crosslink}
+
+\begin{figure}\centering%
+    %
+    \begin{subfigure}{0.45\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{img/AFM_Innventia_nanocellulose}
+        \caption{}\label{fig:cellulose}
     \end{subfigure}%
-\caption{(a) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}. (b) Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.}
-\end{figure}
+    %
+    \hfill
+    \begin{subfigure}{0.45\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{img/crosslink}
+        \caption{}\label{fig:crosslink}
+    \end{subfigure}%
+    %
+    \caption{(\ref{fig:cellulose}) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
+(\ref{fig:crosslink}) Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.}
+\end{figure}%
 
@@
 
-The cross-linking microfibrils can be modeled as a pinned line-
-incidence constraint system in $\mathbb{R}^2$, where incidence
-constraints are used instead of distance constraints.
+The cross-linking microfibrils can be modeled as a pinned line-incidence constraint system in $\mathbb{R}^2$, where incidence constraints are used instead of distance constraints.
 
-A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
-together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
-fixed positions in $\mathbb{R}^2$, such that each edge is constrained
-to lie on a line passing through the corresponding pin, i.e.\ $\delta:
-E \rightarrow \mathbb{R}^2$.
+A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$, such that each edge is constrained to lie on a line passing through the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
 %
@@
 
-In the case of microfibril cross-linking, each fibril is
-attached to some fixed larger organelle/membrane at one spot.
-Consequently, each fibril can be modeled as an edge of the graph,
-with the attachment being the corresponding pin.
+In the case of microfibril cross-linking, each fibril is attached to some fixed larger organelle/membrane at one spot.
+Consequently, each fibril can be modeled as an edge of the graph, with the attachment being the corresponding pin.
 The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge.
 
-Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph,
-where the grey ovals denote pins representing attachments of fibrils,
-and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
+Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph, where the grey ovals denote pins representing attachments of fibrils, and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
 % and the black dots stand for vertices/cross-linkings.
@@
 
-\begin{figure}
-  \centering
+\begin{figure}\centering
     \includegraphics[width=.7\linewidth]{img/pinned}
-\caption{ An isostatic pinned line-incidence graph.}
+    \caption{An isostatic pinned line-incidence graph.}
     \label{fig:pinned_line}
@@
 
-In this section, we will adapt the results in Section~\ref{sec:DRP}
-to give the canonical DR-plan for pinned line-incidence graphs.
-First, we note that %unlike bar-joint graphs,
-\vemph{an isostatic pinned line-incidence graph can be disconnected},
-being the disjoint union of two or more isostatic subgraphs.
-This is because the pins have fixed positions on the plane.
+In this section, we will adapt the results in Section~\ref{sec:DRP} to give the canonical DR-plan for pinned line-incidence graphs. First, we note that
+%unlike bar-joint graphs,
+\vemph{an isostatic pinned line-incidence graph can be disconnected}, being the disjoint union of two or more isostatic subgraphs. This is because the pins have fixed positions on the plane.
 %
@@
 %
-We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of
-the canonical DR-plan:
+We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of the canonical DR-plan:
 
@@
 \begin{definition}
-The \dfn{DR-plan} of a  pinned line-incidence graph $G$ is one in which
-(1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$,
-or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
+    The \dfn{DR-plan} of a  pinned line-incidence graph $G$ is one in which (1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$, or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
     % Definition~\ref{def:drp} with  modified rules number 2 and 4:
@@
 
-The \dfn{canonical DR-plan} of $G$  is one in which
-the  children containing rigid subgraphs are connected isostatic vertex-maximal subgraphs of the parent.
+    The \dfn{canonical DR-plan} of $G$  is one in which the  children containing rigid subgraphs are connected isostatic vertex-maximal subgraphs of the parent.
 \end{definition}
 
-Theorem~\ref{theorem:main}
-holds for  pinned line-incidence graphs with the above modified definition.
-The proof is similar to the original proof (in the Appendix) using the same set of lemmas and
+Theorem~\ref{theorem:main} holds for  pinned line-incidence graphs with the above modified definition.
+The proof is similar to the original proof (in \ref{theorem:main:proof}) using the same set of lemmas and
 %
@@
 \begin{observation}\label{lem:pinned_union_intersection}
-Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$,
-where each of them can be either a single edge or a connected isostatic subgraph.
-There are only two possible cases:
+    Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$, where each of them can be either a single edge or a connected isostatic subgraph. There are only two possible cases:
     (1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is underconstrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
@@
 Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
-The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned
-line-incidence graphs is given in Appendix~\ref{sec:appendix_pinned}.
+The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs is given in~\ref{sec:appendix_pinned}.
 %The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.
@@
 %Given Lemma~\ref{lemma:wc_intersection_makes_all_wc} and Lemma~\ref{lemma:uc_intersection_makes_all_uc},
-Thus it is straightforward to adapt the proof of Theorem~\ref{theorem:main}
-to  pinned line-incidence graphs.
+Thus it is straightforward to adapt the proof of Theorem~\ref{theorem:main} to  pinned line-incidence graphs.
 %thus proving Corollary~\ref{cor:pinned}.
@@
 
-\noindent
-\note The recombination problem for  pinned line-incidence systems is trivial.
-Since the pins are given fixed positions in the plane,
-the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.
+% \noindent
+% \note The recombination problem for  pinned line-incidence systems is trivial.
+Since the pins are given fixed positions in the plane, the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.
+
+\sidenote{The recombination problem for  pinned line-incidence systems is trivial.
+Since the pins are given fixed positions in the plane, the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.}

### 2015-02-21 doc/GMP2015/revision_1/appendix_b.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.
Finished and submitted revision 1.

--- 2015-02-15 doc/GMP2015/original_submission/appendix_b.tex
@@
-\section{Appendix B}
+\section{Open Problems}
 \label{sec:appendix:b}
 
-\subsection{Open Problems}
+% \subsection{Open Problems}
 \label{sec:futurework}
@@
 \begin{openproblem}
-    Is there a lower complexity algorithm than \candrpcomplexityv to find the canonical DR-plan of isostatic 2D bar-joint graphs?
+    Is there a lower complexity algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
 \end{openproblem}
@@
 \begin{openproblem}
-    For fixed $k$ we have polynomial time optimal DR-planning, recombination (modification) in the presence of $k$ overconstraints (Section \ref{sec:DRP}, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed \ref{sec:recomb} and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan \ref{sec:table}.
+    For fixed $k$ we have polynomial time optimal DR-planning, recombination (modification) in the presence of $k$ overconstraints \todo{(Section \ref{sec:DRP},} optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed \ref{sec:recomb} and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan \ref{sec:table}.
 However, $k$ appears in the exponent for the straightforward algorithms for these problems.

### 2015-02-21 doc/GMP2015/revision_1/appendix_defs.tex
Finished and submitted revision 1.

--- 2015-02-16 doc/GMP2015/revision_1/s02_preliminaries.tex
@@
-\section{Preliminaries}
-\label{sec:prelim}
-%
-\subsection{Basics}
-%
+\section{Basic Definitions and Theory}
+\label{sec:appendix:defs}
+
+\notetoreviewer{The following section was moved in from the Preliminaries section.}
+
 In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
@@
 
-In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{pinned line-incidence} and \dfn{body-hyperpin}.
+In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} \addin{(defined formally in Section \ref{sec:bodypin})} and \dfn{pinned line-incidence} \addin{(defined formally in Section \ref{sec:pinnedline})}.
 In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
@@
 
-Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
+Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
+%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
 %where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
@@
 
-In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.
+\cutout{In this paper, a qusecs is any independent geometric constraint system of one of the 3 types mentioned above.} \notetoreviewer{This now only in the Preliminaries section.}
 
-% \medskip\noindent
-% \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+% In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.
+
+% % \medskip\noindent
+% % \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
 Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.
 
-\sidenote{In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
-Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.}
+% \sidenote{In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications (\ref{sec:bodypin} and \ref{sec:pinnedline}).}
 
@@
 An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
-
-The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set $X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
-
-\subsection{Decomposition-Recombination (DR) Plans}
-%
-%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
-Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
-The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
-%
-%Given a constant $k$, a graph $G$ is:
-%\begin{itemize}
-%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
-%
-%    \item \textbf{overconstrained} if it is not independent.
-%
-%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
-%
-%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
-%
-%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
-%
-%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
-%\end{itemize}
-%
-% \begin{definition}
-%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
-%     % induced
-%     subgraph $S\subseteq G$ such that $d(S) > k$.
-% \end{definition}
-%
-% \begin{definition}
-%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
-%     % induced
-%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
-% \end{definition}
-%
-%
-%
-% \begin{definition}
-%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
-%     \todo{Maybe leave the following out until it's needed later?}
-%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
-Trivial graphs for 2D would be a single vertex.
-Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between).
-Etc.
-These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
-% \end{definition}
-%
-\begin{definition}\label{def:drp}
-    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the following properties:
-    \begin{enumerate}
-        \item Each node represents/contains/is a rigid vertex-induced subgraph of $G$.
-        % \item For a node, $C$, that is  a non-trivial graph, its $N$
-        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
-        % subgraphs of $C$.
-        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
-        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
-        \item A root node is a maximal rigid subgraph of $G$.
-    \end{enumerate}
-%
-    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
-%
-    A DR-plan is \dfn{complete} if it satisfies an additional rule:
-        \item For a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$.
-This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well.
-We denote a complete DR-plan of $G$ as $\comdrp{G}$
-    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
-The maximum  fan-in is called the \dfn{size} of the DR-plan.
-We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
-
-%
-%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
-%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
-%an isostatic vertex-maximal proper subgraph.
-    %Note that nodes will be referred to interchangeably as ``the node that
-    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
-\end{definition}
-%
-% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
-\begin{remark}
-    For a given graph, there could be exponentially many  DR-plans and even optimal DR-plans in the size of the graph.
-A complete DR-plan is unique but may not be (is usually not) optimal.
-DR-plans of self-similar graphs are self-similar.
-More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
-\end{remark}
-
-% \begin{definition}
-%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
-By this, rule 3 is implicit for a complete DR-plan.
-% \end{definition}
-
-% \begin{definition}
-%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-% \end{definition}
-
-%\emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.
-% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
-It is not useful to solve for two bodies simultaneously if there are no constraints between them.
-% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.
-
-\sidenote{In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.}
-
-\subsection{Importance of an Optimal DR-plan}
-The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:subdivision}.
-
-In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
-
-\medskip\noindent
-\header{Realizing given qusecs.}
-Note that this basic task would require finding real solutions a large multivariate polynomial system (of inequalities and equalities representing the constraints).
-Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
-With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems i.e.,the maximum fan-in occurring in a DR-plan.
-Thus finding an optimal DR-plan is important for realization.
-
-\medskip\noindent
-\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
-This can be obtained using the DR-plan we obtain a with each block being the rigidity matrix of an isostatic subgraph.
-The blocks intersect on isostatic or trivial subgraphs.
-This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
-Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
-When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
-When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.
-
-%\subsection{DR-plan}
-%%\subsection{Definitions}
-%%\label{sec:prelim_defs}
-%%
-%%% TODO
-%%
-%%% A paragraph of preliminaries such as:
-%%%     - rigidity matrix R,
-%%%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
-%%%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
-%%% Should precede the definition of
-%%%     + independent,
-%%%     + overconstrained (not independent),
-%%%     - rigid (contains an independent set with sufficiently many eedges/constraints)
-%%%     - isostatic (wellconstrained/minimally rigid), ,
-%%%     - flexible (not rigid),
-%%%     + underconstrained (independent and not rigid).
-%%% After that, define DR plans etc.
-%%
-%%In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system.
-$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected).
-Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$.
-The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
-%%
-%%Given a constant $k$, a graph $G$ is:
-%%\begin{itemize}
-%%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
-%%
-%%    \item \textbf{overconstrained} if it is not independent.
-%%
-%%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
-%%
-%%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
-%%
-%%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
-%%
-%%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
-%%\end{itemize}
-%%
-%\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must overconstrained and (2) all of its subgraphs are also trivial.
-%
-%% \begin{definition}
-%%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
-%%     % induced
-%%     subgraph $S\subseteq G$ such that $d(S) > k$.
-%% \end{definition}
-%
-%% \begin{definition}
-%%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
-%%     % induced
-%%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
-%% \end{definition}
-%
-%
-%
-%% \begin{definition}
-%%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
-%%     \todo{Maybe leave the following out until it's needed later?}
-%%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$.
-Trivial graphs for 2D would be a single vertex.
-Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between).
-Etc.
-These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
-%% \end{definition}
-%
-%
-%\begin{definition}\label{def:drp}
-%    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, $DRP(G)$, is defined as a tree that has the following properties:
-%    \begin{enumerate}
-%        \item The root of the tree `contains' $G$.
-%        \item For all nodes, $C$, that `contain' non-trivial graphs, its $N$ children, $C_1, \ldots, C_N$, are trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
-%        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
-%        \item A node with a trivial subgraph is a leaf.
-%    \end{enumerate}
-%
-%    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
-%
-%    An equivalent DAG can be constructed from this tree such that all nodes containing the same subgraphs are combined into one, with all edges preserved.
-%
-%    Note that nodes will be referred to interchangeably as ``the node that contains the (sub)graph $C$'' and as simply ``$C$''.
-%
-%    A DR-plan is \textbf{complete} if it satisfies the modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
-This makes rule 3 implicit.
-The complete DR-plan is unique.
-%
-%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-%\end{definition}
-%
-%% \begin{definition}
-%%     The \textbf{complete DR-plan} of graph $G$, $CompleteDRP(G)$, is the unique DR-plan but with a modified rule number 2: the children are \emph{all} of the trivial or wellconstrained vertex-maximal proper subgraphs of that node $C$.
-By this, rule 3 is implicit for a complete DR-plan.
-%% \end{definition}
-%
-%% \begin{definition}
-%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
-%% \end{definition}
-%
-%
-%\subsection{Notation}
-%
-%In Sections \ref{sec:DRP} and \ref{sec:recomb}, $G$ is taken to be a \emph{wellconstrained} bar and joint graph with values $(V,E,w_V,w_E)$.
-%% \todo{I don't think the following is actually necessary to mention...} To keep problems meaningful we assume that $G$ is connected, as the methods discussed are used to solve systems of equations to establish linkages.
-It is not useful to solve for two bodies simultaneously if there are no constraints between them.
-%% THIS MEANS NO EDGES MAY BE LEFT OUT, ALSO.
-%
-%$Idc(G,X)$ is the graph induced on $G$ with the vertex set $X\subseteq V$. This can also be overloaded such that, using graph $H=(W,F)$, $Idc(G,H)=Idc(G,W)$.
-%
-%$C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
-$C_i$ is the $i^{\text{th}}$ child of $C$.
-By definition of a DR-plan, it is implied that $C_i$ is a wellconstrained vertex-maximal proper subgraph.

### 2015-02-21 doc/GMP2015/revision_1/appendix_proofs.tex
Finished and submitted revision 1.

--- 2014-12-20 doc/GMP2015/appendix_a.tex
@@
-\section{Appendix A}
+\section{Proofs}
 \label{sec:appendix:a}
+\label{sec:appendix:proofs}
 
@@
 
-\newcommand{\usestwod}{\todo{Uses 2D requirement:}}
-\renewcommand{\usestwod}{}
-
-\subsubsection{Proof of Theorem \ref{theorem:main}}
-
-\begin{proof}
-Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs.
-Therefore, these are the only possibilities to consider.
-
-\medskip\noindent
-\vemph{Part 1:} We show that no subset of the children can union to form $C$, thereby requiring all of them be included, and because this is the only choice it must be optimal for this node.
-
-Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic.
-If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
-So, it must be that $U=C$.
-\usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
-By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
-Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$.
-Thus, the assumption is wrong and $U$ cannot be isostatic.
-As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.
-
-\medskip\noindent
-\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
-This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
-Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
-
-However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in for any descendant.
-Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
-Suppose we select $i$ and $j$, where $i\neq j$, as the children.
-Since
-\[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
-the children of this node will be
-\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
-and
-\[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
-% $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
-% the children of this node will be
-% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
-% and
-% $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
-for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other.
-\todo{Prove these are valid children?
-Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once.
-Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
-
-This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
-\end{proof}
+% \subsubsection{Proof of Theorem \ref{theorem:main}}
+% \label{theorem:main:proof}
 
@@
 
-For (4), because subgraphs of a isostatic graph can only be trivial, under-, or isostatic, all cases have already been exhausted.
+For (4), because subgraphs of a isostatic graph can only be trivial, under\addin{constrained}, or isostatic, all cases have already been exhausted.
 \end{proof}
@@
 For convenience, it is implied that all graphs are the induced graph on $C$.
+
+% The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set $X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
+
 \newcommand{\inducedOnC}[1]{#1}
-
 % \newcommand{\inducedOnC}[1]{Idc\left(C,#1\right)}
@@
 
-\subsubsection{Proof of Theorem \ref{theorem:algo_complexity}}
-
-\begin{proof}
-The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
-Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of component-candidates.
-Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list.
-If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list.
-Continue with the next candidate.
-When you exhaust the candidates, move to the next step.
-
-Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.
-
-If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$.
-The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
-We construct something that, while recombining it, results in the same systems to solve as the true DR-plan.
-Begin with the parent $C$.
-Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$.
-Recursively apply the algorithm to the child containing $R_1$.
-The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
-Etc.
-At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
-Apply the algorithm recursively to both children.
-% Although the nodes containing $R_1,\ldots,R_N$ are under-constrained, simply take the children to be the roots of the forest
-
-The reason we construct it this way is because we end up with a tree instead of a directed acyclic graph.
-The leaves of this tree will be exactly every single edge in the graph, with none repeated.
-Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.
-
-Therefore, the complexity of the algorithm as a whole is $O(|V|^3)$.
-%
-% \note In the best case, this can be $O(V^2)$ if the graph is a 2-tree
-\end{proof}
-
-% \begin{proof}
-% The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
-Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of component-candidates.
-Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
-If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
-Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$.
-Continue with the next candidate.
-If you exhaust the candidates, output the entire true-component list.
-
-% %on the children
-% This is run recursively on each node.
-If you use dynamic programming to store the decompositions of each component, it prevents repetition of work (done by using the DAG representation of the DR-plan).
-In this case you will have to do work on no more than $O(|V|)$ nodes that get smaller at each level.
-
-% Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
-
-% It's a tree in only trivial case -> O(V), always order of leaves
-% If its isostatic intersection, we do the sequential extension,intersections only done once.
-
-% inductively. The number of nodes at this level will e more than the total. Number of leaves (terminals for dag) is at most the number of edges
-
-% \end{proof}
-
-% \begin{proof}
-% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph.
-There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
-After the subgraphs have been found, it is a straightforward application of the theory.
-If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph.
-Otherwise, all the subgraphs are the children.
-Apply recursively to the children.
-This naturally terminates when you reach trivial subgraphs.
-The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
-\todo{Think more about complexity.}
-% \end{proof}
+% \subsubsection{Proof of Theorem \ref{theorem:algo_complexity}}
 
@@
 
-\subsubsection{Proof of Theorem \ref{theorem:omdk}}
-\begin{proof}
-The proof follows from the fact that 2-trees are well decomposable and
-have simple DR-plans of size 2. We know that $G$ can be reduced by
-removing $k$ edges to create a partial-2-tree $G'$ which can then be
-completed to an (isostatic) 2-tree by adding some set of non-edges
-$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
-proving the theorem.
-\end{proof}
+% \subsubsection{Proof of Theorem \ref{theorem:omdk}}
 
@@
 \begin{proof}
-    The Cartesian realization space of $(H,\left<\delta_{E'},
-    \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
-    and is the union of $2^t$ solutions (modulo orientation preserving
-    isometries) each with a distinct orientation type, where $t$ is
-    the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
-    the restriction of the length vector $\delta$ to the edges in $S$.
-    A desired solution $p$ (or connected component of a solution
-    space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
-    found by a subdivided binary search   of the cartesian realization
-    space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
-    type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
-    convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
-    described in Theorem \ref{theorem:convexcayley}. A solution $p$
-    is found  when the lengths for nonedges in $D$ match $\delta_D$.
+    The Cartesian realization space of $(H,\left<\delta_{E'}, \lambda_F\right>)$ is computed easily with a DR-plan of size 2, and is the union of $2^t$ solutions (modulo orientation preserving isometries) each with a distinct orientation type, where $t$ is the number of triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction of the length vector $\delta$ to the edges in $S$.
+A desired solution $p$ (or connected component of a solution space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a subdivided binary search   of the Cartesian realization space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation type $\sigma_p$ - as $\lambda_F$ ranges over the discretized convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in Theorem \ref{theorem:convexcayley}.
+A solution $p$ is found  when the lengths for nonedges in $D$ match $\delta_D$.
 \end{proof}
@@
 
-\subsubsection{Proof of Remark \ref{thm:1dofcase}}
+\subsubsection{Proof of Theorem \ref{thm:1dofcase}}
 \begin{proof}
@@
 We use the same notations as in the original proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, in the last section.
+For convenience, it is implied that all graphs are the induced graph on $C$.
 
-First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection},
-both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
+First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection}, both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
 Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a isostatic subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected isostatic vertex-maximal proper subgraph of  $C$.
+
+\newcommand{\inducedOnC}[1]{#1}
+% \newcommand{\inducedOnC}[1]{Idc\left(C,#1\right)}
 
@@
 \begin{itemize}
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible.
-Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
-    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
+    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible.
+Since $C_k\cup C_i = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
+    By Lemma~\ref{lem:pinned_union_intersection}, it must be underconstrained,
     so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.
 
-    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D'_{i,j})$ is not possible. The proof is similar to the previous case.
+    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible. The proof is similar to the previous case.
 
-    \item 2 cases: $C_k=Idc(C,R'_i\cup R_j\cup D'_{i,j})$ and $C_k=Idc(C,R_i\cup R'_j\cup D'_{i,j})$ are not possible.
+    \item 2 cases: $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible.
     The proof is similar to the previous case.

### 2015-02-21 doc/GMP2015/revision_1/img/tikz/demo_graph_candrp_seq.tikz.tex
Finished and submitted revision 1.

\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
    \tikzstyle{b}=[align=center]
    \tikzstyle{e}=[]

    \node[regular polygon, regular polygon sides=3, fill=white, inner sep=4, dashed] (center) at (0,0-.125-.125) {};

% LEVEL 1
    \node[b] (vG) at ( 0,  0) {$G_{\Delta}$};

% LEVEL 2
    \def\x{0}
    \def\y{0}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    \tedge{n1vb}{n1vd}{solid}{}{};
    \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    \tedge{n1vc}{n1vf}{solid}{}{};
    \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{2.5}
    \def\y{-.866}
    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \tedge{n1va}{n1vb}{solid}{}{};

    \def\x{3.5}
    \def\y{-.866}
    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \tedge{n1vc}{n1va}{solid}{}{};

% LEVEL 3
    \def\x{0}
    \def\y{.5-2.35}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    \tedge{n1vb}{n1vd}{solid}{}{};
    \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    % \tedge{n1vc}{n1vf}{solid}{}{};
    % \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{2}
    \def\y{.5-2.35}
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    \tedge{n1vf}{n1ve}{solid}{}{};

    \def\x{3}
    \def\y{.5-2.35}
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    \tedge{n1vc}{n1vf}{solid}{}{};

% LEVEL 4
    \def\x{0}
    \def\y{.5-4.3}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    % \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    % \tedge{n1vb}{n1vd}{solid}{}{};
    % \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    % \tedge{n1vc}{n1vf}{solid}{}{};
    % \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{3}
    \def\y{.5-4.3}
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \tedge{n1vb}{n1vd}{solid}{}{};

    \def\x{4}
    \def\y{.5-4.3}
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \tedge{n1vd}{n1ve}{solid}{}{};

% DRP edges
    \draw (vG)--(0,  -.75);
    \draw (vG)--(2.1,-.65);
    \draw (vG)--(3.2,-.55);

    \draw (0,  -.2-1.85)--(0,  -.75-1.85);
    \draw (0,  -.2-1.85)--(2.1,-.65-1.85);
    \draw (0,  -.2-1.85)--(3.2,-.55-1.85);

    \draw (0,  -.2-1.85-1.85)--(0,  -.75-1.85-1.85);
    \draw (0,  -.2-1.85-1.85)--(2.1,-.65-1.85-1.85);
    \draw (0,  -.2-1.85-1.85)--(3.2,-.55-1.85-1.85);

\end{tikzpicture}

### 2015-02-21 doc/GMP2015/revision_1/img/tikz/demo_graph_clustmindrp.tikz.tex
Finished and submitted revision 1.

--- 2015-02-21 doc/GMP2015/revision_1/img/tikz/demo_graph_candrp_seq.tikz.tex
@@
 % LEVEL 2
-    \def\x{0}
-    \def\y{0}
-    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
+    \def\x{-2.5}
+    \def\y{-.75}
+    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
     \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
     \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
-    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
-    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
-    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
-    % \tedge{n1va}{n1vb}{solid}{}{};
+    \tedge{n1va}{n1vb}{solid}{}{};
     \tedge{n1vb}{n1vc}{solid}{}{};
-    % \tedge{n1vc}{n1va}{solid}{}{};
+    \tedge{n1vc}{n1va}{solid}{}{};
+
+    \def\x{0}
+    \def\y{-.75}
+    \node[c] (n1vb) at (\x+0, \y+0)  [label={left, inner sep=.555}:$b$]{};
+    \node[c] (n1vd) at (\x-.5,\y-.866) [label={left, inner sep=.555}:$d$]{};
+    \node[c] (n1ve) at (\x+.5,\y-.866) [label={below,inner sep=.555}:$e$]{};
     \tedge{n1vb}{n1vd}{solid}{}{};
@@
     \tedge{n1ve}{n1vb}{solid}{}{};
+
+    \def\x{2.5}
+    \def\y{-.75}
+    \node[c] (n1vc) at (\x+0, \y+0)  [label={right,inner sep=.555}:$c$]{};
+    \node[c] (n1ve) at (\x-.5,\y-.866) [label={below,inner sep=.555}:$e$]{};
+    \node[c] (n1vf) at (\x+.5,\y-.866) [label={right,inner sep=.555}:$f$]{};
     \tedge{n1vc}{n1vf}{solid}{}{};
@@
 
-    \def\x{2.5}
-    \def\y{-.866}
+% LEVEL 3
+    \def\x{-2.75}
+    \def\y{-2.25}
     \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
@@
 
-    \def\x{3.5}
-    \def\y{-.866}
-    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
+    \def\x{-2.25}
+    \def\y{-2.25}
     \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
-    \tedge{n1vc}{n1va}{solid}{}{};
+%
+%
+    \def\x{-.25}
+    \def\y{-2.25}
+    \node[c] (n1vb) at (\x+0, \y+0)  [label={left, inner sep=.555}:$b$]{};
+    \node[c] (n1vd) at (\x-.5,\y-.866) [label={left, inner sep=.555}:$d$]{};
+    \tedge{n1vb}{n1vd}{solid}{}{};
 
-% LEVEL 3
-    \def\x{0}
-    \def\y{.5-2.35}
-    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
-    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
-    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
-    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
-    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
-    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
-    % \tedge{n1va}{n1vb}{solid}{}{};
-    \tedge{n1vb}{n1vc}{solid}{}{};
-    % \tedge{n1vc}{n1va}{solid}{}{};
-    \tedge{n1vb}{n1vd}{solid}{}{};
-    \tedge{n1vd}{n1ve}{solid}{}{};
-    \tedge{n1ve}{n1vb}{solid}{}{};
-    % \tedge{n1vc}{n1vf}{solid}{}{};
-    % \tedge{n1vf}{n1ve}{solid}{}{};
+    \def\x{.25}
+    \def\y{-2.25}
+    \node[c] (n1ve) at (\x+.5,\y-.866) [label={below,inner sep=.555}:$e$]{};
+%
+%
+    \def\x{2.25}
+    \def\y{-2.25}
+    \node[c] (n1vc) at (\x+0, \y+0)  [label={right,inner sep=.555}:$c$]{};
+    \node[c] (n1ve) at (\x-.5,\y-.866) [label={below,inner sep=.555}:$e$]{};
     \tedge{n1ve}{n1vc}{solid}{}{};
 
-    \def\x{2}
-    \def\y{.5-2.35}
-    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
-    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
-    \tedge{n1vf}{n1ve}{solid}{}{};
-
-    \def\x{3}
-    \def\y{.5-2.35}
-    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
-    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
-    \tedge{n1vc}{n1vf}{solid}{}{};
-
-% LEVEL 4
-    \def\x{0}
-    \def\y{.5-4.3}
-    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
-    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
-    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
-    % \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
-    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
-    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
-    % \tedge{n1va}{n1vb}{solid}{}{};
-    \tedge{n1vb}{n1vc}{solid}{}{};
-    % \tedge{n1vc}{n1va}{solid}{}{};
-    % \tedge{n1vb}{n1vd}{solid}{}{};
-    % \tedge{n1vd}{n1ve}{solid}{}{};
-    \tedge{n1ve}{n1vb}{solid}{}{};
-    % \tedge{n1vc}{n1vf}{solid}{}{};
-    % \tedge{n1vf}{n1ve}{solid}{}{};
-    \tedge{n1ve}{n1vc}{solid}{}{};
-
-    \def\x{3}
-    \def\y{.5-4.3}
-    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
-    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
-    \tedge{n1vb}{n1vd}{solid}{}{};
-
-    \def\x{4}
-    \def\y{.5-4.3}
-    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
-    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
-    \tedge{n1vd}{n1ve}{solid}{}{};
+    \def\x{2.75}
+    \def\y{-2.25}
+    \node[c] (n1vf) at (\x+.5,\y-.866) [label={right,inner sep=.555}:$f$]{};
 
 % DRP edges
-    \draw (vG)--(0,  -.75);
-    \draw (vG)--(2.1,-.65);
-    \draw (vG)--(3.2,-.55);
+    \draw (vG)--(-2.3,-.55);
+    \draw (vG)--( 0,  -.55);
+    \draw (vG)--( 2.3,-.55);
 
-    \draw (0,  -.2-1.85)--(0,  -.75-1.85);
-    \draw (0,  -.2-1.85)--(2.1,-.65-1.85);
-    \draw (0,  -.2-1.85)--(3.2,-.55-1.85);
+    \draw (-2.5-.1, -1.716)--(-2.65,-1.95);
+    \draw (-2.5+.1, -1.716)--(-2.05,-2.55);
 
-    \draw (0,  -.2-1.85-1.85)--(0,  -.75-1.85-1.85);
-    \draw (0,  -.2-1.85-1.85)--(2.1,-.65-1.85-1.85);
-    \draw (0,  -.2-1.85-1.85)--(3.2,-.55-1.85-1.85);
+    \draw (-2.5-.1+2.5, -1.716)--(-2.65+2.5,-1.95);
+    \draw (-2.5+.1+2.5, -1.716)--(-2.05+2.5,-2.55);
+
+    \draw (-2.5-.1+2.5+2.5, -1.716)--(-2.65+2.5+2.5,-1.95);
+    \draw (-2.5+.1+2.5+2.5, -1.716)--(-2.05+2.5+2.5,-2.55);
 

### 2015-02-21 doc/GMP2015/revision_1/s00_abstract.tex
Finished and submitted revision 1.

@@
 \begin{abstract}
-A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
+\cutout{A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.}
 
-Crucial for the design and analysis of the abovementioned properties is the recursive decomposition of diverse types of underlying 2D geometric constraint systems that are \vemph{isostatic or underconstrained (which we call collectively as \dfn{qusecs})}.
-While the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-hard even for (possibly overconstrained) 2D bar-joint constraint systems we describe an \candrpcomplexity\ algorithm  for qusecs.
-Besides elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-plans that meet various desirable, previously studied criteria.
-We leverage recent results on Cayley configuration spaces to show that the indecomposable systems---that are solved at node of the optimal DR-plan by recombining solutions to child systems---can be minimally modified to become decomposable and have a small DR-plan, leading to efficient realization algorithms.
+Crucial for the design and analysis of \cutout{the abovementioned properties} \addin{quasi-uniform (aperiodic) and self-similar, layered material structures} is the recursive decomposition of diverse types of underlying 2D geometric constraint systems \cutout{that are isostatic or underconstrained (} \addin{,} which we call \dfn{qusecs}\cutout{)}.
+While the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-hard \cutout{even} for \cutout{(possibly overconstrained)} \addin{general} 2D bar-joint constraint systems\addin{,} we describe an \candrpcomplexity\ algorithm  for qusecs \addin{that are defined to be isostatic or underconstrained}.
+\cutout{Besides elucidating NP-hardness, the algorithm output} \addin{The algorithm relies on the new notion of a} canonical DR-plan\cutout{s} that meets various desirable, previously studied criteria.
+\addin{In addition,} we leverage recent results on Cayley configuration spaces to show that the indecomposable systems---that are solved at the nodes of the optimal DR-plan by recombining solutions to child systems---can be minimally modified to become decomposable and have a small DR-plan, leading to efficient realization algorithms.
 We show formal connections to well-known problems such as completion of underconstrained systems.
-We briefly discuss the modeling and analysis of two specific 2D layered materials as applications of the above theory and algorithms.
+We briefly discuss the modeling and analysis of two specific 2D layered materials, modeled as body-hyperpin and pinned line-incidence systems, as applications of the above theory and algorithms.
 We intend to make a software implementation and videos available upon request and publicly available for the final version.

### 2015-02-21 doc/GMP2015/revision_1/s01_introduction.tex
Finished and submitted revision 1.

@@
+\section{Response to reviewers}
+
+\cutout{This means} the text was removed.
+\addin{This means} the text was added.
+\notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
+% \movedfrom{This means} the text was moved from another part of the document.
+
+\subsection{Reviewer 1}
+No suggestions made.
+
+\subsection{Reviewer 2}
+% \noindent
+% Changes made:
+\begin{itemize}
+  \item We added a new example in figures to illustrate definitions/proofs/comparisons.
+And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
+  \item Abstract was shortened and clarified.
+  \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}.
+Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}.
+Slight modification to definition of DR-plan.
+  \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
+  \item Section \ref{sec:DRP} had some proofs moved back in from the appendix.
+Section \ref{sec:recomb} had a proof moved back in from the appendix.
+Section \ref{sec:conclusion} had future work moved back in from the appendix.
+  \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
+  \item Images were properly aligned and cropped.
+\end{itemize}
+
+\subsection{Reviewer 3}
+We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.
+
 \section{Introduction}
 \label{sec:intro}
+
+\newcommand{\seedefs}{(formally defined in \ref{sec:appendix:defs})}
+\newcommand{\seedefsb}{(see \ref{sec:appendix:defs} for definitions)}
+\newcommand{\seedefsc}{See \ref{sec:appendix:defs} for formal definitions}
+\newcommand{\seedefsd}{see \ref{sec:appendix:defs} for definitions}
+\newcommand{\seedefsprelim}{(formally defined in Section \ref{sec:prelim})}
 
@@
 Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either \vemph{isostatic or underconstrained, not self-stressed} (See Section \ref{sec:prelim} for definitions).
-These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.
+Furthermore, a layer is often  either
+\vemph{isostatic or underconstrained}\cutout{,} \addin{(}\vemph{not self-stressed}\addin{,}
+\cutout{(}\seedefsd).
+\cutout{These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.}
+\addin{A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.}
 
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.3}{img/Chlamydomonas_TEM_17}
-\SetMyMinHeight{.3}{img/ligten2}
-\SetMyMinHeight{.3}{img/Rothemund-DNA-SierpinskiGasket}
+\SetMyMinHeight{.32}{img/Chlamydomonas_TEM_17}
+\SetMyMinHeight{.32}{img/ligten2}
+\SetMyMinHeight{.32}{img/Rothemund-DNA-SierpinskiGasket}
 
@@
   %
-  \begin{subfigure}{0.3\linewidth}\centering
+  \begin{subfigure}{0.32\linewidth}\centering
     \includegraphics[height=\myMinHeight]{img/Chlamydomonas_TEM_17}
-    \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}\label{fig:material_examples:microtubule}
+    \caption{}\label{fig:material_examples:microtubule}
   \end{subfigure}%
@@
   \hfill
-  \begin{subfigure}{0.3\linewidth}\centering
+  \begin{subfigure}{0.32\linewidth}\centering
     \includegraphics[height=\myMinHeight]{img/ligten2}
-    \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}\label{fig:material_examples:tendon}
+    \caption{}\label{fig:material_examples:tendon}
   \end{subfigure}%
@@
   \hfill
-  \begin{subfigure}{0.3\linewidth}\centering
+  \begin{subfigure}{0.32\linewidth}\centering
     \includegraphics[height=\myMinHeight]{img/Rothemund-DNA-SierpinskiGasket}
-    \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples:sierpinski}
+    \caption{}\label{fig:material_examples:sierpinski}
   \end{subfigure}%
   %
-  \caption{}\label{fig:material_examples}
+  \caption{(\ref{fig:material_examples:microtubule}) Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.
+(\ref{fig:material_examples:tendon}) Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.
+(\ref{fig:material_examples:sierpinski}) A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples}
 \end{figure*}%
@@
 In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints.
-Such 2D \dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined formally in Section \ref{sec:prelim}) can be used to understand or design material layers (their solutions) with desired properties.
-%
+Such 2D \dfn{qusecs}, quasi-uniform or self-similar constraint systems \seedefsprelim, can be used to understand or design material layers (their solutions) with desired properties.
+
+\subsection{Contributions and Organization}
+\label{sec:cont}
+
+\notetoreviewer{The position of this section and following section were swapped.}
+
+The contributions of this paper are the following.
+\begin{itemize}
+  \item \addin{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
+This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}
+
+  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier (discussed in the following subsection), for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.
+
+  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
+The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
+Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} \addin{(see Figure~\ref{fig:demo_graph:clustmindrp})}.
+
+  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
+Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
+In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components.
+When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
+
+  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
+  \cutout{For Examples 1 and 2 (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example 3 (canonical and optimal DR-plans for pinned line incidence systems [X]) and Examples 4 and 5 (boundary-conditions for achieving various desired properties of body-hyperpin systems).}
+  \addin{For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems.
+For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).}
+
+  \item We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
+\end{itemize}
+
 \subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
-We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4,5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+\label{sec:intro:prevwork}
+
+We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1\cutout{,} \addin{and} 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \addin{and} \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4\cutout{,} \addin{and} 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
 The limitations of these techniques directly motivate the contributions of this paper.
@@
 \header{(i) Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems (defined formally in Section \ref{sec:prelim}).
+Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs.
 When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.
 
-However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as (formally defined in Section \ref{sec:prelim}) \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
+However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
 
@@
 \header{(ii) Optimal Recursive Decomposition (DR-planning).}
-Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in Section \ref{sec:prelim}.
+Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim.
 For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system.
 Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}.
-An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e., the maximum number of child subsystems of any parent system.
-Being exponential the size, the complexity of solving the parent constraint system from the solutions of the child systems, is overwhelmingly dominated by it.
+An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.
+Being exponential \addin{in} the size, the complexity of solving the parent constraint system \cutout{from the solutions of the of the child systems, is overwhelmingly dominated by it} \addin{is overwhelmingly dominated by the complexity of solving the child systems}.
 
@@
 For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
-See Section \ref{sec:prelim} for formal definitions.
+\seedefsc.
 
-While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-plans meeting the cluster minimality criterion, no such algorithm is known for the latter criterion.
+While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion \cite{hoffman2001decompositionI}, no such algorithm is known for the latter criterion.
 
@@
 This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
-Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions, to give a solution to the parent system at any given level of the DR-plan.
+Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions\cutout{,} to give a solution to the parent system at any given level of the DR-plan.
 For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
@@
 Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called\cutout{,} tree-width 2, series-parallel, \addin{and} $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
 While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
-%
-\subsection{Contributions and Organization}
-\label{sec:cont}
-
-The contributions of this paper are the following.
-\begin{itemize}
-  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier mentioned in (2) above, for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.
-
-  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
-The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
-Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.
-
-  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
-Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
-In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components.
-When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
-
-  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
-For Examples \ref{materialexample1}, \ref{materialexample2} (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example \ref{materialexample3} (canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}) and Examples \ref{materialexample4}, \ref{materialexample5} (boundary-conditions for achieving various desired properties of body-hyperpin systems).
-
-  \item We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
-\end{itemize}
 
@@
 
-% \noindent
-% \note All proofs appear in the Appendix.
-
-\sidenote{All proofs appear in \ref{sec:appendix:a}.}
+\cutout{
+\noindent
+\textbf{Note:} All proofs appear in the Appendix.
+% \sidenote{All proofs appear in X.}
+}

### 2015-02-21 doc/GMP2015/revision_1/s02_preliminaries.tex
Finished and submitted revision 1.

--- 2015-02-15 doc/GMP2015/revision_1/s02_preliminaries.tex
@@
 \label{sec:prelim}
-%
-\subsection{Basics}
-%
-In this paper, a \dfn{geometric constraint system} is a multivariate
-polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$,
-representing constraints with parameters $\delta$ between geometric
-primitives  in $\mathbb{R}^2$ represented collectively as $x\in
-\mathbb{R}^n$.
-%
-When the type of constraint (system) is fixed, the system is simply
-represented as $(G,\delta)$, where $G$ is the underlying constraint
-(hyper)graph $G = (V,E)$ with the vertices $V$ representing the
-geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$
-representing the constraints, each with an associated parameter
-$\delta$.
-%
-For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a
-graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E
-\rightarrow \mathbb{R}$; this represents the distance constraint
-system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where
-$x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
 
-\noindent
-\note Geometric constraint systems can also have
-inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
-values rather than exact values.
+\addin{
+Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.
 
-In this paper, we will consider 3 types of constraint systems:
-bar-joint as well as \dfn{pinned line-incidence} and \dfn{body-hyperpin}.
-In all of these cases, a Cartesian \dfn{realization} or \dfn{solution}
-$G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean
-transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or
-$\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with
-parameters $\delta$, modulo orientation preserving isometries
-(Euclidean rigid body motions for distance constraint).
+Throughout this paper, a \textbf{\textsf{Theorem}}, \textbf{\textsf{Lemma}}, \textbf{\textsf{Observation}}, or \textbf{\textsf{Remark}} with an asterisk after it means that the proof appears in~\ref{sec:appendix:proofs}.
 
-Although the realization space itself depends on the constraint
-parameters $\delta$, many relevant \dfn{generic} properties of the
-constraint system $G(x,\delta)$ are properties of the constraint
-(hyper)graph $G$ and do not depend on $\delta$ (or they hold for all
-but a measure zero set of $\delta$ values). Many of these are
-properties of the Jacobian $\Delta_x G(x,\delta)$, often called the
-appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
-For example, the \dfn{bar-joint rigidity matrix of the graph $G =
-(V,E)$} is a matrix of indeterminates representing the Jacobian of the
-distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$
-columns per vertex in $V$ and one row per edge in $E$, where the row
-corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates
-for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\
-$v$), that is 4 non-zero entries per row.
+\subsection{Qusecs}
 
-% XXXXXXXXX
-% something has to be said along with asimow roth
-% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
-% appropriate rank, the realizations or solutions of the corresponding
-% constraint system are generically isolated and zero-dimensional if the
-% constraint system has equalities and exact values for the parameters
-% $\delta$. If the constraint system has inequalities, i.e, if the
-% $\delta$ lie in an interval, this claim is approximate: the solutions
-% are isolated small, full-dimensional  connected components.
+In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:appendix:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
 
-One important property of generic constraint system or (hyper)graph
-%that depends on the appropriate rigidity matrix of indeterminates
-is \dfn{rigidity} (\note we refer to these as properties of
-the constraint system or as properties of the underlying (hyper)graph
-interchangeably), i.e., the realizations or solutions of the
-corresponding constraint system being generically isolated and zero-
-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint
-(hyper)graph is rigid if it is generically \dfn{infinitesimally
-rigid}, i.e.\ the number of independent rows of its appropriate
-rigidity matrix is at least the number of columns less the number of
-rigid body motions, which is 3 for bar-joint systems.
-%This condition is called generic {\em infinitesimal rigidity}.
+% \medskip\noindent
+% \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.
 
-\noindent
-\note If the constraint system has inequalities, i.e, if the $\delta$
-lie in an interval, the definition of rigidity is approximate: the
-solutions are isolated small, full-dimensional  connected components.
-% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
-
-Other generic constraint system or (hyper)graph properties are the
-following:
-% (\note we refer to these as properties of the constraint
-% system or as properties of the underlying (hyper)graph
-% interchangeably).
-A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate
-rigidity matrix of indeterminates has independent rows (i.e, the
-determinant of some square submatrix is not identically zero).
-%It is {\em rigid} if the number of independent rows of the rigidity matrix is
-%at least the number of columns less the number of rigid body motions,
-%which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is
-both rigid and independent.
-%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
-%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
-%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
-%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
-It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it
-is independent and not rigid, or \dfn{overconstrained} if it is not
-independent.
-
-Defining the combinatorial independence of a subset of edges $E'
-\subseteq E$ to be the generic independence of corresponding rows in
-the appropriate rigidity matrix of indeterminates, we obtain the
-\dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
-%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
-%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-There are various results on combinatorial characterization of
-independence and rigidity  and rigidity matroids for different types
-of graphs. For bar-joint rigidity matroid, the famous Laman's theorem
-\cite{laman1970graphs} states that the underlying graph is isostatic if and only if
-$|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
-at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows
-that all 6-vertex-connected graphs are rigid in the plane. For bar-
-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying
-multigraph is isostatic if and only if it can be decomposed as $3$
-edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the
-same characterization using a different technique to study the
-algebraic-geometric conditions of genericity, called pure condition.
-Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity
-matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if
-$|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least
-2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| =
-k|V| - l$. In general, given a $d$-uniform hypergraph, a
-$(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
-
-In this paper, a \dfn{qusecs} is any \dfn{independent} geometric
-constraint system of one of the 3 types mentioned above.
-
-\medskip\noindent
-\note In the remainder of this section and the next three
-sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
-qusecs and graphs. Relevant formal analogies for the other 2 types of
-qusecs and (hyper)graphs are given in the subsequent 2 sections on
-Applications.
-
-%XXXNow comes the definition of stress vector and flex vector.
-%NOTE: something has to be said about what flexes and stresses mean
-
-%TODO
-Given a bar-joint graph, a vector (of indeterminates) in the right
-null space of its rigidity matrix is called a \dfn{flex vector}. It
-has 2 entries per vertex and represents the internal motions of the
-system (i.e.\ modulo rigid body motions). For rigid graphs, all flex
-vectors are identically zero.
-%
-A vector (of indeterminates) in the left null space of the rigidity
-matrix is called a \dfn{stress vector}. It has one entry per edge and
-represents a \dfn{self stress} of the system. For independent graphs,
-all stress vectors are identically zero.
-%
-An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of
-indeterminates) which specifies a 2D vector acting at each vertex such
-that the stress balance equation $sR = t$ holds, where $s$ is a vector
-of internal stresses, one for each bar, and $R$ is the rigidity
-matrix.
-
-The notation $Idc(G,X)$ is the subgraph of $G$ induced by the vertex set
-$X\subseteq V$. For a subgraph $H=(W,F)$ of $G$, $Idc(G,H)=Idc(G,W)$.
+In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in their corresponding sections on applications.
+}
 
@@
 %
+
+\ClearMyMinHeight
+\SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
+\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
+\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}
+
+\begin{figure*}\centering%
+  %
+  \begin{subfigure}{0.4\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
+    \caption{}\label{fig:demo_graph:graph}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
+    \caption{}\label{fig:demo_graph:comdrp}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
+    \caption{}\label{fig:demo_graph:candrp}
+  \end{subfigure}%
+  %
+  \caption{\notetoreviewer{This is a new figure.} \addin{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
+(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
+Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
+The children of triangles (3 edges) are omitted.
+(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
+The children of triangles are omitted.}}
+  \label{fig:demo_graph}
+\end{figure*}%
+
+\begin{figure*}\centering%
+  %
+  \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
+  \caption{
+  \notetoreviewer{This is a new figure.}
+  \addin{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
+Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge.
+Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2.
+With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
+  }
+  \label{fig:demo_graph:clustmindrp}
+\end{figure*}%
+
 \begin{definition}\label{def:drp}
-    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
-    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
-    following properties:
+    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, \cutout{(see Figure X)} is defined as a forest that has the following properties:
     \begin{enumerate}
-        \item Each node represents/contains/is a rigid vertex-induced
-        subgraph of $G$.
+        \item Each node represents/contains/is a rigid \cutout{vertex-induced} subgraph of $G$.
         % \item For a node, $C$, that is  a non-trivial graph, its $N$
@@
 %
-    A DR-plan is \dfn{complete} if it satisfies an additional rule:
-        \item For a node $C$ that is  a non-trivial graph, its $N$
-        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
-        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
-        and in fact, $C$ contains all the edges in the union of its
-        children as well.
-        We denote a complete DR-plan of $G$ as $\comdrp{G}$
+    A DR-plan is \dfn{complete} if it satisfies an additional rule: for a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$.
+This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well.
+We denote a complete DR-plan of $G$ as $\comdrp{G}$.
+
     A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
@@
 We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
+
+    \addin{See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.}
 
@@
 \begin{remark}
-    For a given graph, there could be exponentially many  DR-plans and
-    even optimal DR-plans in the size of the graph. A complete DR-plan
-    is unique but may not be (is usually not) optimal. DR-plans
-    of self-similar graphs are self-similar. More than one node (leaf)
-    in a DR-plan forest may represent the same subgraph (vertex) of
-    $G$.
+    For a given graph, there could be exponentially many  DR-plans and even optimal DR-plans in the size of the graph.
+A complete DR-plan is unique but may not be (is usually not) optimal.
+DR-plans of self-similar graphs are self-similar.
+More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
 \end{remark}
+
+\cutout{
+\noindent
+\textbf{Note:} In Section X, $G$ is assumed to be isostatic.
+}
 
@@
 
-\noindent
-\note In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.
+\subsection{Importance of an Optimal DR-plan}
+The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
 
-\subsection{Importance of an Optimal DR-plan}
-The use of a DR-plan in designing especially self-similar qusecs is
-immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and
-\ref{fig:subdivision}.
-
-In fact, knowing an optimal DR-plan is crucial for
-design and analysis of even non-self-similar qusecs and their
-realizations with various desirable properties.
+In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
 
@@
 \header{Realizing given qusecs.}
-Note that this basic task
-would require finding real solutions a large multivariate polynomial
-system (of inequalities and equalities representing the constraints).
-Without DR-planning, this requires double exponential time in the
-number of variables (even if orientation type is specified). With
-DR-planning, the complexity is dominated by the size of the largest
-subsystem that is solved, or recombined from the solutions of its
-child subsystems i.e.,the maximum fan-in occurring in a DR-plan. Thus
-finding an optimal DR-plan is important for realization.
+Note that this basic task would require finding real solutions \addin{to} a large multivariate polynomial system (of inequalities and equalities representing the constraints).
+Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
+With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems\addin{,} i.e., the maximum fan-in occurring in a DR-plan.
+Thus\addin{,} finding an optimal DR-plan is important for realization.
 
 \medskip\noindent
-\header{Recursive block
-decomposition of the rigidity matrix, stress space and external stresses.}
-This can be obtained using the DR-plan we obtain a
-with each block being the
-rigidity matrix of an isostatic subgraph. The blocks intersect on
-isostatic or trivial subgraphs. This recursive block decomposition of
-the rigidity matrix $R$ correspondingly recursively decomposes
-the stress vectors $s$ that balance a given external stress $t$ since
-$sR = t$. Moreover, the external stress vector $t$ is
-recursively ``distributed'' into external stresses for each
-node of the DR-plan.
-When the rigidity matrix and stress
-vectors contain indeterminates, this  can be used to design
-underlying graph and parameters $\delta$ to obtain an optimal
-distribution of stresses on the geometric primitives. When the
-realization of qusecs is given, this can be used to analyze the
-distribution of a given external load.
+\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
+This can be obtained using the DR-plan\addin{,} \cutout{we obtain a} with each block being the rigidity matrix of an isostatic subgraph.
+The blocks intersect on isostatic or trivial subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
+Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
+When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
+When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.
 

### 2015-02-21 doc/GMP2015/revision_1/s03_drp.tex
Finished and submitted revision 1.

@@
 
-In this section we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.
+In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.
+
+\addin{
+\sidenote{In this section, $G$ is always assumed to be isostatic.}
+}
 
@@
 
-This theorem (proven in the \ref{theorem:main:proof}) relies on the following crucial observation and lemma that are also used for other types of qusecs in the Application sections (\ref{sec:bodypin}, \ref{sec:pinnedline}) of the paper.
+\notetoreviewer{The following proof was moved in from the appendix.}
 
-\begin{observation}\label{lemma:union_intersection}
+\begin{proof}
+We show the existence of a canonical DR-plan by constructing it as follows:
+
+Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$
+with children $C_1,\ldots,C_N$ keep children nodes according to the following rules:
+\begin{enumerate}
+   \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
+   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
+\end{enumerate}
+
+This directly satisfies properties 2 and 3 of a canonical DR-plan (see definition \ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs.
+However, we still have to show property 1 (that this makes a DR-plan).
+For rule (1), because we are beginning with a complete DR-plan, if we keep all the children it is still a DR-plan.
+For rule (2), we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we have found a larger rigid vertex-maximal proper subgraph.
+
+If we begin with an isostatic graph, rigid can be replaced with isostatic throughout the construction and remain valid. The rigid proper subgraphs of an isostatic graph must be isostatic themselves.
+
+Now we must show that a canonical DR-plan is optimal.
+The proof is by induction on the level of the plan; we show that following these rules always guarantees minimum fan-in at this level and does not increase fan-in on following levels.
+
+ % We do this inductively on the l, beginning with the root node, showing that the rules are always the optimal choice.
+% by discussing each rule and proving inductively that this is the optimal choice at each level of the DR-plan tree, starting with the root node.
+
+% Now we show that this plan is optimal by considering the two cases.
+Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
+ % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
+Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider.
+Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.
+
+\medskip\noindent
+\vemph{Rule 1:}
+% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
+%
+First, we claim that if the intersection is trivial all $C_1,\ldots,C_N$ must be children of $C$ in an optimal DR-plan.
+We do this by showing that no subset of the children can union to form $C$, thereby requiring all of them be included; because this is the only choice, it is the minimum fan-in any DR-plan could have for this node.
+% must be optimal for this node.
+
+Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic.
+If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
+So, it must be that $U=C$.
+\usestwod
+However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
+By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
+Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$.
+Thus, the assumption is wrong and $U$ cannot be isostatic.
+As $C$ is isostatic, this means no proper subset of $C_1,\ldots,C_N$ can union to form $C$.
+
+Furthermore, since a canonical DR-plan uses proper rigid \vemph{vertex-maximal} subgraphs as children, and because their pairwise intersection is trivial, it follows that any such node has at most as many children as a DR-plan without this restriction, because the union of the children have to contain all edges of the parent.
+Therefore, this is the optimal choice in this case.
+
+\medskip\noindent
+\vemph{Rule 2:} Second, we claim that if the intersection is isostatic, the choice of any two children will result in an optimal DR-plan.
+
+If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
+Thus, any two children can be chosen to make a canonical DR-plan.
+% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
+
+\newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
+\renewcommand{\induceonc}[1]{#1}
+\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in S_N\setminus\{#1\}}{R_k}}}
+
+However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants.
+Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
+Suppose we select $i$ and $j$, where $i\neq j$, as the children.
+For convenience, we will assume all subgraphs are induced subgraphs on $C$.
+We know that $C=\induceonc{I\cup\bigcup_{k\in S_N}{R_k}}$ and $C_i=\iunion{i}$.
+The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of which intersect on isostatic subgraphs.
+So any two of these are viable children for $C_i$.
+% Since
+% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
+% the children of $C_i$ will be
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
+% and
+% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
+% % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
+% % the children of this node will be
+% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
+% % and
+% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
+% for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
+This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in S_N$, with every $k$ appearing at least once.
+Thus, regardless of the choice of $C_i,C_j$, then their children, etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
+Therefore, the choice is optimal.
+%
+% \medskip\noindent
+% This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
+%
+% -- say this at top `proof is by induction on the level of the dr=plan'
+\end{proof}
+
+This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the \cutout{A}\addin{a}pplication sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.
+
+\begin{observation*}\label{lemma:union_intersection}
 If $F_i$ and $F_j$ are isostatic subgraphs of the same
@@
 (4) $F_i\cap F_j$ is not underconstrained.
-\end{observation}
+\end{observation*}
 
@@
 
-\begin{lemma}\label{lemma:combined_lemma}
+\begin{lemma*}\label{lemma:combined_lemma}
 Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
@@
 \end{enumerate}
-\end{lemma}
+\end{lemma*}
 
@@
 
+% \ClearMyMinHeight
+% \SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
+% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
+% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}
+
+% \begin{figure*}\centering%
+%   %
+%   \begin{subfigure}{0.4\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
+%     \caption{}\label{fig:demo_graph:graph}
+%   \end{subfigure}%
+%   %
+%   \hfill
+%   \begin{subfigure}{0.3\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
+%     \caption{}\label{fig:demo_graph:comdrp}
+%   \end{subfigure}%
+%   %
+%   \hfill
+%   \begin{subfigure}{0.3\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
+%     \caption{}\label{fig:demo_graph:candrp}
+%   \end{subfigure}%
+%   %
+%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
+(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
+Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
+The children of triangles (3 edges) are omitted.
+(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
+The children of triangles or omitted.}
+%   \label{fig:demo_graph}
+% \end{figure*}%
+
+\begin{figure*}\centering%
+  %
+  \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
+  \caption{\notetoreviewer{This is a new figure.} \addin{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
+The children of the triangle are omitted.
+Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
+Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}}
+  \label{fig:demo_graph:candrpseq}
+\end{figure*}%
+
 \begin{theorem}\label{theorem:algo_complexity}
-    There exists an \candrpcomplexityv algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
+    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
 \end{theorem}
 
+\notetoreviewer{The following proof was moved in from the appendix.}
+
+\begin{proof}
+The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
+Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
+The output of this will be a list of component-candidates.
+Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
+If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list.
+If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list.
+Continue with the next candidate.
+When you exhaust the candidates, move to the next step.
+
+Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.
+
+If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$.
+The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
+We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan.
+Begin with the parent $C$.
+Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$.
+Recursively apply the algorithm to the child containing $R_1$.
+The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
+And so on.
+At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
+Apply the algorithm recursively to both children.
+See Figure~\ref{fig:demo_graph:candrpseq} for an example.
+% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
+
+The reason for this construction is that the result is a tree with no duplicate nodes.
+The leaves of this tree will exactly be the edges of this graph.
+Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.
+
+Therefore, the complexity of the algorithm as a whole is $O(|V|^3)$.
+%
+% \note In the best case, this can be $O(V^2)$ if the graph is a 2-tree
+\end{proof}
+
+% \begin{proof}
+% The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
+Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
+The output of this will be a list of component-candidates.
+Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
+If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
+If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
+Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$.
+Continue with the next candidate.
+If you exhaust the candidates, output the entire true-component list.
+
+% %on the children
+% This is run recursively on each node.
+If you use dynamic programming to store the decompositions of each component, it prevents repetition of work (done by using the DAG representation of the DR-plan).
+In this case you will have to do work on no more than $O(|V|)$ nodes that get smaller at each level.
+
+% Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
+
+% It's a tree in only trivial case -> O(V), always order of leaves
+% If its isostatic intersection, we do the sequential extension,intersections only done once.
+
+% inductively. The number of nodes at this level will e more than the total. Number of leaves (terminals for dag) is at most the number of edges
+
+% \end{proof}
+
+% \begin{proof}
+% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the wellconstrained vertex-maximal proper subgraphs of the input wellconstrained graph.
+There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
+After the subgraphs have been found, it is a straightforward application of the theory.
+If the intersection of any two of the subgraphs is wellconstrained, then take those two to be children of the graph.
+Otherwise, all the subgraphs are the children.
+Apply recursively to the children.
+This naturally terminates when you reach trivial subgraphs.
+The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
+\todo{Think more about complexity.}
+% \end{proof}
+
 \begin{remark}
-    While the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
+    While the canonical DR-plan is optimal only if the input graph is independent, if there are only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
 However, the time complexity is exponential in $k$.
 \end{remark}
+
+\addin{
+\begin{conjecture}
+\label{conj:mfaisoptimal}
+    The Modified Frontier Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
+\end{conjecture}
+
+The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA is a bottom-up algorithm, involving complex datastructures.
+However, such a proof, even if it exists, would not be possible without the new notion of a canonical DR-plan at hand.
+The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the sequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.
+The strength of MFA compared to other DR-planning algorithms is illustrated in the paper~\cite{lomonosov2004graph}.
+While MFA also offers complexity $O(n^3)$, it does not guarantee optimality.
+}

### 2015-02-21 doc/GMP2015/revision_1/s04_recombination.tex
Finished and submitted revision 1.

@@
 
-In the following, we formulate the problem of \dfn{optimal modification}, of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
+In the following, we formulate the problem of \dfn{optimal modification}\cutout{,} of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
 Leveraging recent results on \dfn{Cayley configuration spaces}, our approach to the optimal modification problem achieves the following:
@@
 However, it is not clear how Criterion (c) is met by OMD graphs.
-%
+
 \subsection{Using Convex Cayley Configuration Spaces}
@@
 To state the theorem, we first have to define the notion of \dfn{2-sums} and \dfn{2-trees}.
-Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
+Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.
+% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
 A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$.
@@
 
+\notetoreviewer{The following proof was moved in from the appendix.}
+
+\begin{proof}
+    The proof follows from the fact that 2-trees are well decomposable and have simple DR-plans of size 2.
+We know that $G$ can be reduced by removing $k$ edges to create a partial-2-tree $G'$ which can then be completed to an (isostatic) 2-tree by adding some set of non-edges $F$.
+Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2, proving the theorem.
+\end{proof}
+
 % \begin{proof}
@@
 
-\begin{theorem}\label{theorem:criterionc}
+\begin{theorem*}\label{theorem:criterionc}
     [Efficient search]
@@
 The (discrete) volume $W$ is exponential in $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
-\end{theorem}
+\end{theorem*}
 
@@
 [Using Cayley configuration space].
-A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e, its DR-plan has a root and 9 children corresponding to the 9 edges.
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e\addin{.}, its DR-plan has a root and 9 children corresponding to the 9 edges.
 Solving or recombining the system $(G,\delta)$ corresponding to the root of this DR-plan involves solving a quadratic system with 8 equations and variables.
@@
 In the next Section \ref{sec:tdecomp}, we discuss this issue of why the largest class of reduced graphs is desirable.
-%
+
 \subsection{Optimized Modification by Enlarging the class of Reduced
@@
 In addition, the bounds of these intervals are of low algebraic complexity.
-More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e, the computation of these bounds is tree-decomposable.
+More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e\addin{.}, the computation of these bounds is tree-decomposable.
 Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or can be computed by solving a triangularized system of quadratics.

### 2015-02-21 doc/GMP2015/revision_1/s05_problems.tex
Finished and submitted revision 1.

@@
 \label{sec:table}
-In this section we provide a unified view of  the various problems studied in the previous 2 sections, along with formal reductions between them, and relationships to other known problems and results as well as open questions.
-%
+
+In this section we provide a unified view of the various problems studied in the previous 2 sections, along with formal reductions between them\cutout{, and}\addin{.
+We discuss their} relationship\cutout{s} to other known problems and results as well as open questions.
+
 \subsection{Special classes of small DR-plans}
@@
 In the recombination setting, the corresponding \dfn{restricted OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction} or to be a low Cayley complexity tree-decomposable graph as in Section \ref{sec:tdecomp}.
-Clearly these problems have deterministic polynomial time algorithms, the algorithms run in time exponential in $k$.
+Clearly these problems have deterministic polynomial time algorithms \addin{in $n$}, \addin{but} the algorithms run in time exponential in $k$.
 
-We discuss the complexity of the restricted OMD problem (when $k$ is part of the input) in \ref{sec:futurework}.
+We discuss the complexity of the restricted OMD problem (when $k$ is part of the input) in Section \ref{sec:futurework}.
 

### 2015-02-21 doc/GMP2015/revision_1/s06_problempinning.tex
Finished and submitted revision 1.

@@
-\section{Application: Finding completions of underconstrained glassy structures to isostatic}
+\section{Application: Finding completions of underconstrained glassy structures from underconstrained to isostatic}
 \label{sec:bodypin}
+
+\notetoreviewer{Changed section title from ``Application: Finding completions of underconstrained glassy structures to isostatic''}
 % \bibliography{paper}
 
+\ClearMyMinHeight
+\SetMyMinHeight{.3}{img/silica}
+\SetMyMinHeight{.2}{img/Silicon_tetrahedron}
+
 \begin{figure}\centering
-    \includegraphics[width=0.4\linewidth]{img/Silica}\hspace{0.5cm}
-    \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
-    \caption{Example of a single monolayer of a Silica (silicon and oxygen) glassy structure.
-This can be thought of as a triangle-multipin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
+    \includegraphics[height=\myMinHeight]{img/Silica}
+    \hspace{0.5cm}
+    \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
+    % \raisebox{0.25\height}{
+    %     \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
+    % }
+    \caption{Example of a single monolayer of a silica (silicon and oxygen) glassy structure.
+This can be thought of as a triangle-multipin qusecs where the silicon atoms are the triangles and the oxygen atoms are the pins.
 Not shown here are other monolayers stacked on this one.
-Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent monolayer.
+Each silicon atom has the structure seen to the right, and so binds to another oxygen atom in an adjacent monolayer.
 Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
@@
 
-In all cases, we are specifically interested in how to add additional constraints resulting isostatic structure with a small DR-plan, so a realization can be found and stress and other property design can be efficient.
+In all cases, we are specifically interested in how to add additional constraints \addin{such that} the resulting isostatic structure \cutout{with} \addin{has} a small DR-plan\cutout{, so} \addin{; this way} a realization can be found\addin{, allowing efficient} \cutout{and} stress and other property design\cutout{ can be efficient}.
 To answer these questions, we first introduce the qusecs that are used to model these materials.
-We will discuss Item 2 in detail in this section.
+We will discuss \cutout{Item 2} \addin{item (2)} in detail in this section.
 
-\begin{figure}\centering
-    (a)
-    \begin{subfigure}{0.2\linewidth}
-        \includegraphics[width=\linewidth]{img/pentl1}
-    \end{subfigure}
-    %
-    $\rightarrow$
-    \begin{subfigure}{0.2\linewidth}
-        \includegraphics[width=\linewidth]{img/pentl2}
-    \end{subfigure}
-    %
+% \begin{figure}\centering
+%     (a)
+%     \begin{subfigure}{0.2\linewidth}
+%         \includegraphics[width=\linewidth]{img/pentl1}
+%     \end{subfigure}
+%     %
+%     $\rightarrow$
+%     \begin{subfigure}{0.2\linewidth}
+%         \includegraphics[width=\linewidth]{img/pentl2}
+%     \end{subfigure}
+%     %
 
-    (b)
-    \begin{subfigure}{0.8\linewidth}
-        \includegraphics[width=\linewidth]{img/pentawesome}
-    \end{subfigure}
-    \caption{Examples of self-similarity via repeated subdivision.
+%     (b)
+%     \begin{subfigure}{0.8\linewidth}
+%         \includegraphics[width=\linewidth]{img/pentawesome}
+%     \end{subfigure}
+%     \caption{Examples of self-similarity via repeated subdivision.
 In (a), simple subdivision that does not guarantee isostaticity.
@@
 Credit to \cite{subdivision_paper} for (a).}
+%     \label{fig:subdivision}
+% \end{figure}
+
+\begin{figure}\centering
+    \begin{subfigure}{0.35\linewidth}
+        $
+        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{img/pentl1}}}
+        % \raisebox{-.5\height}{\scalebox{1}{$\rightarrow$}}
+        \vcenter{\hbox{$\rightarrow$}}
+        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{img/pentl2}}}
+        $
+        \caption{}\label{fig:subdivision:simple}
+    \end{subfigure}
+    \begin{subfigure}{0.55\linewidth}
+        \includegraphics[width=\linewidth]{img/pentawesome}
+        \caption{}\label{fig:subdivision:complicated}
+    \end{subfigure}
+    \caption{Examples of self-similarity via repeated subdivision.
+In (\ref{fig:subdivision:simple}) \cite{subdivision_paper}, there is a simple subdivision scheme that does not guarantee isostaticity.
+In (\ref{fig:subdivision:complicated}), a more complicated scheme is used, ensuring that the resulting graph is isostatic but not  tree-decomposable.}
     \label{fig:subdivision}
@@
 \begin{definition}
-    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies some subsets of which are pinned together by pins.
+    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies, some subsets of which are pinned together by pins.
 \end{definition}
 
-\begin{remark}\label{rem:bodypin_is_barjoint}
-    A body-hyperpin qusecs is a special case of bar-joint qusecs of the
-    first 5 sections. As such, the DR-planning for isostatics discussed in Section
-    \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
-\end{remark}
+\begin{remark*}\label{rem:bodypin_is_barjoint}
+    A body-hyperpin qusecs is a special case of bar-joint qusecs of the first 5 sections.
+As such, the DR-planning for isostatics discussed in Section \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
+\end{remark*}
 
@@
 
-    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another.
-Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}, defined in Section  \ref{sec:prelim}.
+    Such a body-pin graph\addin{, $G_{BP}$,} can also be seen as a \dfn{body-bar graph}\addin{, $G_{BB}$,} where the bodies \addin{of $G_{BB}$} are the original bodies \addin{of $G_{BP}$} and each pin \cutout{represents 2 bars from one body to another} \addin{between bodies in $G_{BP}$ are replaced with 2 bars in $G_{BB}$ between the same bodies}.
+    Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse} (defined in \ref{sec:appendix:defs}).
+\addin{See Figure \ref{fig:bodypindrp:b}.}
 \end{definition}
@@
     A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
-In our case, the trivial nodes are just the bodies.
+In our case, the trivial nodes are \cutout{just} the bodies.
 \end{definition}
 
-We can develop the notion of a canonical DR-plan when the hypergraph is $(k,l)$-sparse (i.e., combinatorially independent) using certain notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}, provided such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
+We can develop the notion of a canonical DR-plan when the hypergraph is $(k,l)$-sparse (i.e., combinatorially independent) using certain notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}, provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}).
 In particular, we can define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.
 
-\begin{observation}
+\begin{observation*}
 \label{obs:bodypin_drp}
     For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
-\end{observation}
+\end{observation*}
 
@@
 
-The strong Church-Rosser property holds making all canonical DR-plans optimal:
+The strong Church-Rosser property holds\addin{,} making all canonical DR-plans optimal:
 %
@@
 We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
-This is because such $(3,4)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
+This is because such $(3,4)$-tight graphs are matroidal and have a pebble game\cutout{, shown in} \cite{Lee:2007:PGA}.
 \end{remark}
@@
     %
+    \hfill
     \begin{subfigure}{0.2\linewidth}\centering
@@
     %
+    \hfill
     \begin{subfigure}{0.35\linewidth}\centering
@@
     %
-    \caption{(\ref{fig:bodypindrp:a}) is a 1-dof body-pin graph.
-(\ref{fig:bodypindrp:b}) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}.
-(\ref{fig:bodypindrp:c}) is the 1-dof DR-plan for the graph.
+    \caption{(\ref{fig:bodypindrp:a}) A 1-dof body-pin graph.
+(\ref{fig:bodypindrp:b}) The corresponding body-bar graph, explained in Definition \ref{def:body-pin}.
+(\ref{fig:bodypindrp:c}) The 1-dof DR-plan for the graph.
 In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level.
-(\ref{fig:bodypindrp:d}) is the result of adding one such body to the bold-faced node.}
+(\ref{fig:bodypindrp:d}) The result of adding one such body to the bold-faced node.}
+    \label{fig:bodypindrp}
     %
@@
 The above discussion leads to the main theorem:
-\begin{theorem}
+\begin{theorem*}
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb}.
-\end{theorem}
+    Given a 1-dof body-pin or triangle-multipin graph \cutout{$G$} and corresponding 1-dof DR-plan \cutout{$T$}, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb}.
+\end{theorem*}
 
@@
 
-\begin{observation}
+\begin{observation*}
 \label{obs:2dof_case}
     For the 2-dof case, if a similar statement to Remark \ref{rem:1dofcanon} can be made, then Theorem \ref{thm:1dofcase} is true for 2-dof as well.
-\end{observation}
+\end{observation*}
 
@@
 
-\begin{observation}
+\begin{observation*}
 \label{obs:algebraic_completion}
-    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1- or 2- parameter families would no longer permit the algorithm given in the proof.
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer permit the algorithm given in the proof.
 In particular the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if that case were matroidal and an optimal DR-plan is known).
-\end{observation}
+\end{observation*}
 

### 2015-02-21 doc/GMP2015/revision_1/s07_pinnedline.tex
Finished and submitted revision 1.

@@
-\section{Application: finding optimal DR-plans and realizations for cross-linking microfibrils }
+\section{Application: Finding optimal DR-plans and realizations for cross-linking microfibrils }
 \label{sec:pinnedline}
 
-The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of  cross-linking collagen and cellulose microfibrils in animal and plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as \dfn{pinned line-incidence systems}.
+The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of
+\cutout{cross-linking collagen and cellulose microfibrils in animal and plant cell walls, and actin filaments in the cytoskeleton}
+\addin{cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton}
+by modeling these structures as \dfn{pinned line-incidence systems}.
 
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.45}{img/AFM_Innventia_nanocellulose}
-\SetMyMinHeight{.45}{img/crosslink}
+\SetMyMinHeight{.32}{img/AFM_Innventia_nanocellulose}
+\SetMyMinHeight{.32}{img/crosslink}
+\SetMyMinHeight{.32}{img/pinned}
 
 \begin{figure}\centering%
-    %
-    \begin{subfigure}{0.45\linewidth}\centering
+\begin{subfigure}{0.32\linewidth}\centering
     \includegraphics[height=\myMinHeight]{img/AFM_Innventia_nanocellulose}
     \caption{}\label{fig:cellulose}
-    \end{subfigure}%
-    %
-    \hfill
-    \begin{subfigure}{0.45\linewidth}\centering
+\end{subfigure}\hfill%
+\begin{subfigure}{0.32\linewidth}\centering
     \includegraphics[height=\myMinHeight]{img/crosslink}
     \caption{}\label{fig:crosslink}
+\end{subfigure}\hfill%
+\begin{subfigure}{0.32\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/pinned}
+    \caption{}
+    \label{fig:pinned_line}
 \end{subfigure}%
-    %
-    \caption{(\ref{fig:cellulose}) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
-(\ref{fig:crosslink}) Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.}
+\caption{\notetoreviewer{Combined 2 figures.} (\ref{fig:cellulose}) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
+\ref{fig:crosslink} Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.
+\ref{fig:pinned_line} An isostatic pinned line-incidence graph\addin{, representing fibrils and their attachments.}}
 \end{figure}%
@@
 
-\begin{figure}\centering
-    \includegraphics[width=.7\linewidth]{img/pinned}
-    \caption{An isostatic pinned line-incidence graph.}
-    \label{fig:pinned_line}
-\end{figure}
+% \begin{figure}\centering
+%     \includegraphics[width=.7\linewidth]{img/pinned}
+%     \caption{An isostatic pinned line-incidence graph.}
+%     \label{fig:pinned_line}
+% \end{figure}
 
@@
 
-Theorem~\ref{theorem:main} holds for  pinned line-incidence graphs with the above modified definition.
-The proof is similar to the original proof (in \ref{theorem:main:proof}) using the same set of lemmas and
+Theorem~\ref{theorem:main} holds for pinned line-incidence graphs with \cutout{the above} \addin{this} modified definition.
+The proof is similar to the original proof (in Section~\ref{sec:DRP}) using the same set of lemmas and
 %
@@
     Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$, where each of them can be either a single edge or a connected isostatic subgraph. There are only two possible cases:
-    (1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is underconstrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
+    (1) at least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is underconstrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
     %The following table summarizes all bijections:
@@
 
-Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
-Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
-The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs is given in~\ref{sec:appendix_pinned}.
+Given \cutout{Lemma} \addin{Observation}~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
+Lemma \addin{\ref{lemma:combined_lemma}, Point} \ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc}\addin{,} straightforwardly extend to pinned line-incidence graphs.
+The proof of Point~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs is given in~\ref{sec:appendix_pinned}.
 %The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.

### 2015-02-21 doc/GMP2015/revision_1/s08_conclusion.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.
Finished and submitted revision 1.

--- 2014-12-20 doc/GMP2015/s09_conclusion.tex
@@
 \section{Conclusion}
-% \bibliography{paper}
+\label{sec:conclusion}
 
 We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
-For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient ($O(n^2)$) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.
+For the former problem, when there are no overconstraints\cutout{,} (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that \cutout{such} \addin{any} canonical DR-plan\cutout{s are} \addin{is} guaranteed to be optimal, \addin{which is} a strong Church-Rosser property\addin{.
+This gives} \cutout{giving} an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
 
-We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan These results rely on a recent theory of Convex Cayley Configuration spaces.
+We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan.
+These results rely on a recent theory of Convex Cayley Configuration spaces.
 Relationships and reductions between these and previously studied problems were formally clarified.
@@
 
-In Appendix B [\ref{sec:appendix:b}] we provide open problems for future work.
+% \section{Open Problems}
+\subsection{Open Problems}
+\label{sec:appendix:b}
+\label{sec:futurework}
+
+\notetoreviewer{The following section was moved in from the appendix.}
+
+One open problem comes from Section \ref{sec:DRP}:
+\begin{openproblem}
+    Is there a lower complexity algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
+\end{openproblem}
+
+A few natural  open questions concern the following common theme that runs through the paper:
+
+\begin{openproblem}
+    For fixed $k$ we have polynomial time optimal DR-planning \addin{(Section~\ref{sec:DRP})}, recombination (modification) in the presence of $k$ overconstraints\cutout{ (Section X}, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed \addin{(Section~\ref{sec:recomb})} and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan \addin{(Section~\ref{sec:table})}.
+However, \addin{in the running time of all of these algorithms,} $k$ appears in the exponent\cutout{ for the straightforward algorithms for these problems}.
+\end{openproblem}
+
+One problem in the above theme is from Section \ref{sec:bodypin}.
+\begin{openproblem}
+    What is the complexity of the optimal completion problem when the given graph has more than 2-dofs?
+Our proof for the 1 and 2-dof cases relied heavily on the matroidal properties of their corresponding $(k,l)$-tightness.
+For higher number of dofs, the $(k,l)$ characterization is no longer matroidal \cite{Lee:2007:PGA}.
+As a result, the major obstacle is that there is no easy way of obtaining an optimal or canonical $k$-dof DR-plan in general.
+Even assuming such  a DR-plan is available, if higher dofs had the same characteristics, Observation \ref{obs:algebraic_completion} raises questions about the correct measure of DR-plan size that captures algebraic complexity for recombining graphs with many dofs (this is not an issue in the isostatic case).
+Unless some restrictions can be found and taken advantage of, the $k$-dof optimal completion problem would  have complexity exponential in $k$.
+\end{openproblem}
+
+Another problem from the above theme is from Section \ref{sec:recomb}
+\begin{openproblem}
+    What is the complexity of the restricted OMD (optimal modification for decomposition) problem?
+This has the potential to be difficult.
+For example, when the isostatic completion is required to be a 2-tree the restricted OMD problem is reducible to the maximum spanning series-parallel subgraph problem shown by \cite{cai1993spanning} to be NP-complete even if the input graph is planar of maximum degree at most 6.
+However, since the OMD problem has other input restrictions such as not having any proper isostatic subgraphs, it is not clear if the reverse reduction exists and hence it is unclear whether the OMD problem is NP-complete.
+
+    The same holds for the restricted OMD problem where the isostatic completion is required to be a tree-decomposable graph of low Cayley complexity (i.e, have special, small DR-plans).
+One potential obstacle to an indecomposable graph $G$'s membership in the restricted OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
+In fact, 6-connected (hence rigid) graphs with arbitrarily large girth have been constructed in \cite{servatius2000rigidity}.
+\end{openproblem}
+
+The next is the reverse direction of Observation \ref{obs:OC_to_OMD} in Section \ref{sec:table}.
+\begin{openproblem}
+    Is the OMD (optimal modification for decomposition) problem reducible to the OC (Optimal completion) problem?
+\end{openproblem}
+
+More general problem directions are the following.
+\begin{openproblem}
+    Combinatorial rigidity for periodic structures is an active area of research.
+This paper motivates a study of the rigidity of self-similar structures, with self-similar groups replacing periodic groups.
+\end{openproblem}
+
+\begin{openproblem}
+    The pinned line incidence structures of Section \ref{sec:pinnedline}, for example in the case of collagen microfibrils, whose function is elastic contraction, should be considered congruent under projective transformations.
+I.e, the projective group should be factored out as a trivial motion in a new project for extending the combinatorial rigidity characterization of such systems.
+(Currently we permit no trivial motions at all).
+\end{openproblem}
+
+\begin{openproblem}
+    Experimental validation (either computationally or physical experiments) of predications based on the material model and theory used in this paper.
+This can be done by modeling known materials and putting stresses on them, seeing if the prediction is observed in the real material.
+Or, our theory could be used to design new materials, which can be tested to see if they possess predicted properties.
+\end{openproblem}

### 2015-03-25 src/GUI/MainGuiManager.cpp
Minor changes to code, update to TODO

@@ -22,6 +22,10 @@ void Window::set_as_context() {
     MainGuiManager::set_window_context(this->glfw_window());
 }
 
+void Window::close_window() {
+    glfwSetWindowShouldClose(glfw_window(), GL_TRUE);
+}
+
 bool Window::should_close() {
     return glfwWindowShouldClose(this->glfw_window());
 }

### 2015-03-25 src/GUI/MainGuiManager.hpp

@@ -23,6 +23,7 @@ public:
     Window();
 
     void set_as_context();
+    void close_window();
     bool should_close();
     void swap_buffers();
 

### 2015-03-25 src/Graph/Pebbled_Graph.cpp

@@ -6,6 +6,15 @@
 
 #include <iostream>
 
+// #define FOR_EACH_IN(TYPE, X, Y) \
+//     for (TYPE::iterator Y = X.begin(); Y != X.end(); Y++)
+
+// #define FOR_EACH_EDGE_IN_SUBGRAPH(X, Y) \
+//     std::pair<Sg_Edge_Iterator, Sg_Edge_Iterator> _FOR_EACH_EDGE_IN_SUBGRAPH_TEMP_PAIR = X->edges(); \
+//     for (Sg_Edge_Iterator Y = _FOR_EACH_EDGE_IN_SUBGRAPH_TEMP_PAIR.first; Y != _FOR_EACH_EDGE_IN_SUBGRAPH_TEMP_PAIR.second; Y++)
+// #define FOR_EACH_EDGE_IN_SUBGRAPH(X, Y) \
+//     for (Sg_Edge_Iterator Y = X->edges().first; Y != X->edges().second; Y++)
+
 
 Cluster::Cluster(bool f) {
     finished = f;
@@ -483,7 +492,7 @@ void erase_and_delete_all_subseteqs(std::set<Cluster*> &super, std::set<Cluster*
 
 
 // get all wellconstrained vertex-maximal subgraphs
-std::set<Cluster*> Pebbled_Graph::get_all_wcvmps(Cluster* known_cluster) {
+std::set<Cluster*> Pebbled_Graph::get_all_wcvmps() {
     std::set<Cluster*> current_clusters;
 
     std::pair<Sg_Vertex_Iterator, Sg_Vertex_Iterator> vs = this->in_subgraph->vertices();
@@ -496,9 +505,9 @@ std::set<Cluster*> Pebbled_Graph::get_all_wcvmps(Cluster* known_cluster) {
         std::vector<std::set<Cluster*>::iterator> to_erase;
 
 
-        // remove anything from potential that's a subseteq of known
-        if (known_cluster != NULL)
-            erase_and_delete_all_subseteqs(known_cluster, potential_clusters);
+        // // remove anything from potential that's a subseteq of known
+        // if (known_cluster != NULL)
+        //     erase_and_delete_all_subseteqs(known_cluster, potential_clusters);
 
 
         // remove anything from potential that's a subseteq of something from current
@@ -525,8 +534,6 @@ std::set<Cluster*> Pebbled_Graph::get_all_wcvmps(Cluster* known_cluster) {
 
 
 
[…]

### 2015-03-25 src/Graph/Pebbled_Graph.hpp

@@ -106,15 +106,17 @@ private:
     };
 
 
-    std::set<Cluster*> _DRP_2D_aux(
-        Cluster* known_cluster,
-        bool optimalDRP);
+    // std::set<Cluster*> _DRP_2D_aux(
+    //     Cluster* known_cluster,
+    //     bool optimalDRP);
 
+    // The linear decomposition version of the DR-plan. It will be optimal,
+    // but it isn't the vertex-maximal decomposition. It's equivalent.
     std::set<Cluster*> _DRP_2D_linear_aux();
 
 
     // get all wellconstrained vertex-maximal subgraphs
-    std::set<Cluster*> get_all_wcvmps(Cluster* known_cluster = NULL);
+    std::set<Cluster*> get_all_wcvmps();
 
 
 

### 2015-03-25 src/main.cpp

@@ -300,8 +300,7 @@ public:
         if (action == GLFW_PRESS) {
             // cout << "\tPressed" << endl;
             if (key == GLFW_KEY_ESCAPE) {
-                cout << "HERE!" << endl;
-                // glfwSetWindowShouldClose(window, GL_TRUE);
+                this->close_window();
             } else if (key == GLFW_KEY_UP) {
                 if (this->current_drp_node != NULL) {
                     highlight_cluster(current_drp_node->parent());
@@ -319,11 +318,16 @@ public:
                     highlight_cluster(current_drp_node->prev());
                 }
             } else if (key == GLFW_KEY_SPACE) {
-                // cout << "\t\tSpace" << endl;
                 if (drp_display_window == NULL) {
                     drp_display_window = new DrpDisplayWindow(drp);
                     float scale = 3;
-                    global::mgm.create_window(drp_display_window, global::window::width_screen_coords/scale, global::window::height_screen_coords/scale, "DR-Plan", 3, 2);
+                    global::mgm.create_window(
+                        drp_display_window,
+                        global::window::width_screen_coords/scale,
+                        global::window::height_screen_coords/scale,
+                        "DR-Plan",
+                        3,
+                        2);
                     drp_display_window->init_program();
 
                     drp_display_window->update_graph_positions();

### 2015-03-25 doc/GMP2015/revision_2/appendix_defs.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/appendix_defs.tex
@@
 \label{sec:appendix:defs}
-
-\notetoreviewer{The following section was moved in from the Preliminaries section.}
 
@@
 
-In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} \addin{(defined formally in Section \ref{sec:bodypin})} and \dfn{pinned line-incidence} \addin{(defined formally in Section \ref{sec:pinnedline})}.
+In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
 In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
@@
 
-\cutout{In this paper, a qusecs is any independent geometric constraint system of one of the 3 types mentioned above.} \notetoreviewer{This now only in the Preliminaries section.}
-
 % In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.

### 2015-03-25 doc/GMP2015/revision_2/appendix_proofs.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/appendix_proofs.tex
@@
 
-For (4), because subgraphs of a isostatic graph can only be trivial, under\addin{constrained}, or isostatic, all cases have already been exhausted.
+For (4), because subgraphs of a isostatic graph can only be trivial, underconstrained, or isostatic, all cases have already been exhausted.
 \end{proof}
@@
 \begin{proof}
-    An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton).
+    An isostatic graph has 3 parameters that define it (in terms of coming up with a realization).
 These are the Euclidean motions.
@@
 A 2-dof has 5 parameters, and so on.
-These paramters denote the algebraic complexity of giving a realization for the graphs.
+These parameters denote the algebraic complexity of giving a realization for the graphs.
 
@@
 Note that solving $v'$ will result in solving $v$.
-Then, we need to consider nodes above and includiing $v'$ in our complexity.
+Then, we need to consider nodes above and including $v'$ in our complexity.
 $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
-The only remaining case will be if we pin two nodes that are not descendent/ancestor.
+The only remaining case will be if we pin two nodes that are not descendant/ancestor.
 the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.

### 2015-03-25 doc/GMP2015/revision_2/s01_introduction.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s01_introduction.tex
@@
-\section{Response to reviewers}
+% \section{Response to reviewers}
 
-\cutout{This means} the text was removed.
-\addin{This means} the text was added.
-\notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
-% \movedfrom{This means} the text was moved from another part of the document.
+% \cutout{This means} the text was removed.
+% \addin{This means} the text was added.
+% \notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
+% % \movedfrom{This means} the text was moved from another part of the document.
 
-\subsection{Reviewer 1}
-No suggestions made.
+% \subsection{Reviewer 1}
+% No suggestions made.
 
-\subsection{Reviewer 2}
-% \noindent
-% Changes made:
-\begin{itemize}
-  \item We added a new example in figures to illustrate definitions/proofs/comparisons.
+% \subsection{Reviewer 2}
+% % \noindent
+% % Changes made:
+% \begin{itemize}
+%   \item We added a new example in figures to illustrate definitions/proofs/comparisons.
 And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
-  \item Abstract was shortened and clarified.
-  \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}.
+%   \item Abstract was shortened and clarified.
+%   \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}.
 Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}.
 Slight modification to definition of DR-plan.
-  \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
-  \item Section \ref{sec:DRP} had some proofs moved back in from the appendix.
+%   \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
+%   \item Section \ref{sec:DRP} had some proofs moved back in from the appendix.
 Section \ref{sec:recomb} had a proof moved back in from the appendix.
 Section \ref{sec:conclusion} had future work moved back in from the appendix.
-  \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
-  \item Images were properly aligned and cropped.
-\end{itemize}
+%   \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
+%   \item Images were properly aligned and cropped.
+% \end{itemize}
 
-\subsection{Reviewer 3}
-We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.
+% \subsection{Reviewer 3}
+% We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.
 
@@
 Furthermore, a layer is often  either
-\vemph{isostatic or underconstrained}\cutout{,} \addin{(}\vemph{not self-stressed}\addin{,}
-\cutout{(}\seedefsd).
-\cutout{These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.}
-\addin{A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.}
+\vemph{isostatic or underconstrained} (\vemph{not self-stressed},\seedefsd).
+A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
 
@@
 
-\notetoreviewer{The position of this section and following section were swapped.}
-
 The contributions of this paper are the following.
 \begin{itemize}
-  \item \addin{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
-This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}
+  \item In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
+This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.
 
@@
 The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
-Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} \addin{(see Figure~\ref{fig:demo_graph:clustmindrp})}.
+Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
 
@@
   \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
-  \cutout{For Examples 1 and 2 (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example 3 (canonical and optimal DR-plans for pinned line incidence systems [X]) and Examples 4 and 5 (boundary-conditions for achieving various desired properties of body-hyperpin systems).}
-  \addin{For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems.
-For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).}
+For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems.
+For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).
 
@@
 
-We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1\cutout{,} \addin{and} 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \addin{and} \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4\cutout{,} \addin{and} 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
 The limitations of these techniques directly motivate the contributions of this paper.
@@
 An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.
-Being exponential \addin{in} the size, the complexity of solving the parent constraint system \cutout{from the solutions of the of the child systems, is overwhelmingly dominated by it} \addin{is overwhelmingly dominated by the complexity of solving the child systems}.
+Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.
 
@@
 For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.
-This satisfies the so-called \dfn {Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
+This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
 For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
@@
 This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
-Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions\cutout{,} to give a solution to the parent system at any given level of the DR-plan.
+Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan.
 For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
@@
 Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called\cutout{,} tree-width 2, series-parallel, \addin{and} $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
 While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
@@
 % \medskip
-
-\cutout{
-\noindent
-\textbf{Note:} All proofs appear in the Appendix.
-% \sidenote{All proofs appear in X.}
-}

### 2015-03-25 doc/GMP2015/revision_2/s02_preliminaries.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s02_preliminaries.tex
@@
 
-\addin{
 Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.
@@
 Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in their corresponding sections on applications.
-}
 
@@
   %
-  \caption{\notetoreviewer{This is a new figure.} \addin{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
+  \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
 (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
@@
 (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
-The children of triangles are omitted.}}
+The children of triangles are omitted.}
   \label{fig:demo_graph}
@@
   \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
-  \caption{
-  \notetoreviewer{This is a new figure.}
-  \addin{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
+  \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
 Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge.
@@
 With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
-  }
   \label{fig:demo_graph:clustmindrp}
@@
 \begin{definition}\label{def:drp}
-    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, \cutout{(see Figure X)} is defined as a forest that has the following properties:
+    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
     \begin{enumerate}
-        \item Each node represents/contains/is a rigid \cutout{vertex-induced} subgraph of $G$.
+        \item Each node represents/contains/is a rigid subgraph of $G$.
         % \item For a node, $C$, that is  a non-trivial graph, its $N$
@@
 
-    \addin{See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.}
+    See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.
 
@@
 \end{remark}
-
-\cutout{
-\noindent
-\textbf{Note:} In Section X, $G$ is assumed to be isostatic.
-}
 
@@
 \header{Realizing given qusecs.}
-Note that this basic task would require finding real solutions \addin{to} a large multivariate polynomial system (of inequalities and equalities representing the constraints).
+Note that this basic task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints).
 Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
-With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems\addin{,} i.e., the maximum fan-in occurring in a DR-plan.
-Thus\addin{,} finding an optimal DR-plan is important for realization.
+With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
+Thus, finding an optimal DR-plan is important for realization.
 
@@
 \header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
-This can be obtained using the DR-plan\addin{,} \cutout{we obtain a} with each block being the rigidity matrix of an isostatic subgraph.
+This can be obtained using the DR-plan, with each block being the rigidity matrix of an isostatic subgraph.
 The blocks intersect on isostatic or trivial subgraphs.

### 2015-03-25 doc/GMP2015/revision_2/s03_drp.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s03_drp.tex
@@
 
-\addin{
 \sidenote{In this section, $G$ is always assumed to be isostatic.}
-}
 
@@
 
-\notetoreviewer{The following proof was moved in from the appendix.}
-
 \begin{proof}
@@
 
-Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$
-with children $C_1,\ldots,C_N$ keep children nodes according to the following rules:
+Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children nodes according to the following rules:
 \begin{enumerate}
@@
 
-This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the \cutout{A}\addin{a}pplication sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.
+This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.
 
@@
   \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
-  \caption{\notetoreviewer{This is a new figure.} \addin{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
+  \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
 The children of the triangle are omitted.
 Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
-Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}}
+Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
   \label{fig:demo_graph:candrpseq}
@@
 \end{theorem}
-
-\notetoreviewer{The following proof was moved in from the appendix.}
 
@@
 
-\addin{
 \begin{conjecture}
@@
 While MFA also offers complexity $O(n^3)$, it does not guarantee optimality.
-}

### 2015-03-25 doc/GMP2015/revision_2/s04_recombination.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s04_recombination.tex
@@
 
-In the following, we formulate the problem of \dfn{optimal modification}\cutout{,} of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
+In the following, we formulate the problem of \dfn{optimal modification} of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
 Leveraging recent results on \dfn{Cayley configuration spaces}, our approach to the optimal modification problem achieves the following:
@@
 [Using Cayley configuration space].
-A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e\addin{.}, its DR-plan has a root and 9 children corresponding to the 9 edges.
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e., its DR-plan has a root and 9 children corresponding to the 9 edges.
 Solving or recombining the system $(G,\delta)$ corresponding to the root of this DR-plan involves solving a quadratic system with 8 equations and variables.
@@
 In addition, the bounds of these intervals are of low algebraic complexity.
-More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e\addin{.}, the computation of these bounds is tree-decomposable.
+More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e., the computation of these bounds is tree-decomposable.
 Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or can be computed by solving a triangularized system of quadratics.

### 2015-03-25 doc/GMP2015/revision_2/s05_problems.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s05_problems.tex
@@
 
-In this section we provide a unified view of the various problems studied in the previous 2 sections, along with formal reductions between them\cutout{, and}\addin{.
-We discuss their} relationship\cutout{s} to other known problems and results as well as open questions.
+In this section we provide a unified view of the various problems studied in the previous 2 sections, along with formal reductions between them.
+We discuss their relationship to other known problems and results as well as open questions.
 
@@
 In the recombination setting, the corresponding \dfn{restricted OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction} or to be a low Cayley complexity tree-decomposable graph as in Section \ref{sec:tdecomp}.
-Clearly these problems have deterministic polynomial time algorithms \addin{in $n$}, \addin{but} the algorithms run in time exponential in $k$.
+Clearly these problems have deterministic polynomial time algorithms in $n$, but the algorithms run in time exponential in $k$.
 

### 2015-03-25 doc/GMP2015/revision_2/s06_problempinning.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s06_problempinning.tex
@@
 
-\notetoreviewer{Changed section title from ``Application: Finding completions of underconstrained glassy structures to isostatic''}
 % \bibliography{paper}
@@
 
-In all cases, we are specifically interested in how to add additional constraints \addin{such that} the resulting isostatic structure \cutout{with} \addin{has} a small DR-plan\cutout{, so} \addin{; this way} a realization can be found\addin{, allowing efficient} \cutout{and} stress and other property design\cutout{ can be efficient}.
+In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress and other property design.
 To answer these questions, we first introduce the qusecs that are used to model these materials.
-We will discuss \cutout{Item 2} \addin{item (2)} in detail in this section.
+We will discuss item (2) in detail in this section.
 
@@
 
-    Such a body-pin graph\addin{, $G_{BP}$,} can also be seen as a \dfn{body-bar graph}\addin{, $G_{BB}$,} where the bodies \addin{of $G_{BB}$} are the original bodies \addin{of $G_{BP}$} and each pin \cutout{represents 2 bars from one body to another} \addin{between bodies in $G_{BP}$ are replaced with 2 bars in $G_{BB}$ between the same bodies}.
+    Such a body-pin graph, $G_{BP}$, can also be seen as a \dfn{body-bar graph}, $G_{BB}$, where the bodies of $G_{BB}$ are the original bodies of $G_{BP}$ and each pin between bodies in $G_{BP}$ are replaced with 2 bars in $G_{BB}$ between the same bodies.
     Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse} (defined in \ref{sec:appendix:defs}).
-\addin{See Figure \ref{fig:bodypindrp:b}.}
+See Figure \ref{fig:bodypindrp:b}.
 \end{definition}
@@
     A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
-In our case, the trivial nodes are \cutout{just} the bodies.
+In our case, the trivial nodes are the bodies.
 \end{definition}
@@
 
-The strong Church-Rosser property holds\addin{,} making all canonical DR-plans optimal:
+The strong Church-Rosser property holds, making all canonical DR-plans optimal:
 %
@@
 We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
-This is because such $(3,4)$-tight graphs are matroidal and have a pebble game\cutout{, shown in} \cite{Lee:2007:PGA}.
+This is because such $(3,4)$-tight graphs are matroidal and have a pebble game \cite{Lee:2007:PGA}.
 \end{remark}
@@
 \label{thm:1dofcase}
-    Given a 1-dof body-pin or triangle-multipin graph \cutout{$G$} and corresponding 1-dof DR-plan \cutout{$T$}, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb}.
+    Given a 1-dof body-pin or triangle-multipin graph and corresponding 1-dof DR-plan, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb}.
 \end{theorem*}

### 2015-03-25 doc/GMP2015/revision_2/s07_pinnedline.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s07_pinnedline.tex
@@
 
-The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of
-\cutout{cross-linking collagen and cellulose microfibrils in animal and plant cell walls, and actin filaments in the cytoskeleton}
-\addin{cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton}
-by modeling these structures as \dfn{pinned line-incidence systems}.
+The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as \dfn{pinned line-incidence systems}.
 
@@
 \end{subfigure}%
-\caption{\notetoreviewer{Combined 2 figures.} (\ref{fig:cellulose}) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
+\caption{(\ref{fig:cellulose}) Microfibrils of carboxymethylated nanocellulose adsorbed on a silica surface \cite{wikimediacommons2010afm}.
 \ref{fig:crosslink} Cross-linking of cellulose microfibrils \cite{wikimediacommons2007plant}.
-\ref{fig:pinned_line} An isostatic pinned line-incidence graph\addin{, representing fibrils and their attachments.}}
+\ref{fig:pinned_line} An isostatic pinned line-incidence graph, representing fibrils and their attachments.}
 \end{figure}%
@@
 
-Theorem~\ref{theorem:main} holds for pinned line-incidence graphs with \cutout{the above} \addin{this} modified definition.
-The proof is similar to the original proof (in Section~\ref{sec:DRP}) using the same set of lemmas and
+Theorem~\ref{theorem:main} holds for pinned line-incidence graphs with this modified definition. The proof is similar to the original proof (in Section~\ref{sec:DRP}) using the same set of lemmas and
 %
@@
 
-Given \cutout{Lemma} \addin{Observation}~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
-Lemma \addin{\ref{lemma:combined_lemma}, Point} \ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc}\addin{,} straightforwardly extend to pinned line-incidence graphs.
+Given Observation~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
+Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc}, straightforwardly extend to pinned line-incidence graphs.
 The proof of Point~\ref{lemma:wc_intersection_makes_all_wc} for pinned line-incidence graphs is given in~\ref{sec:appendix_pinned}.

### 2015-03-25 doc/GMP2015/revision_2/s08_conclusion.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

--- 2015-02-21 doc/GMP2015/revision_1/s08_conclusion.tex
@@
 We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
-For the former problem, when there are no overconstraints\cutout{,} (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that \cutout{such} \addin{any} canonical DR-plan\cutout{s are} \addin{is} guaranteed to be optimal, \addin{which is} a strong Church-Rosser property\addin{.
-This gives} \cutout{giving} an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
+For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, which is a strong Church-Rosser property.
+This gives an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
 
@@
 
-\notetoreviewer{The following section was moved in from the appendix.}
-
 One open problem comes from Section \ref{sec:DRP}:
@@
 \begin{openproblem}
-    For fixed $k$ we have polynomial time optimal DR-planning \addin{(Section~\ref{sec:DRP})}, recombination (modification) in the presence of $k$ overconstraints\cutout{ (Section X}, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed \addin{(Section~\ref{sec:recomb})} and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan \addin{(Section~\ref{sec:table})}.
-However, \addin{in the running time of all of these algorithms,} $k$ appears in the exponent\cutout{ for the straightforward algorithms for these problems}.
+    For fixed $k$ we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan (Section~\ref{sec:table}).
+However, in the running time of all of these algorithms, $k$ appears in the exponent.
 \end{openproblem}

### 2015-04-23 src/Graph/Graph.cpp
Can recognize series parallel graphs (can do edge contractions)

@@ -3,6 +3,9 @@
 #include "Graph.hpp"
 
 
+#include "MinPriorityQueue.hpp"
+
+
 
 // #include <boost/graph/topological_sort.hpp>
 
@@ -83,6 +86,11 @@
 
 
 
+
+
+
+
+
 Vertex_ID Graph::add_vertex(Vertex_Properties vp) {
     return this->Graph_Type::add_vertex(vp);
 }
@@ -94,13 +102,13 @@ Vertex_ID Graph::add_vertex() {
 Edge_ID Graph::add_edge(Vertex_ID v0, Vertex_ID v1, Edge_Properties ep) {
     std::pair<Edge_ID, bool> p = this->Graph_Type::add_edge(v0, v1, ep);
     Edge_ID e = p.first;
-    bool success = p.second;
+    // bool success = p.second;
 
-    if (success) {
-        //
-    } else {
-        //
-    }
+    // if (success) {
+    //     //
+    // } else {
+    //     //
+    // }
 
     return e;
 }
@@ -109,12 +117,268 @@ Edge_ID Graph::add_edge(Vertex_ID v0, Vertex_ID v1, double distance) {
     return this->add_edge(v0, v1, Edge_Properties(distance));
 }
 
+void Graph::remove_vertex(Vertex_ID v) {
+    boost::remove_vertex(v, *this);
+}
+
+void Graph::remove_edge(Edge_ID e) {
+    boost::remove_edge(e, *this);
+}
+
+void Graph::remove_edge(Vertex_ID v0, Vertex_ID v1) {
+    boost::remove_edge(v0, v1, *this);
+}
+
+
+bool Graph::has_edge(Vertex_ID v0, Vertex_ID v1) {
+    return boost::edge(v0, v1, *this).second;
+}
+
+void Graph::contract_edge(Vertex_ID v0, Vertex_ID v1) {
+    std::cout
+        << "Graph::contract_edge: "
+        << (*this)[v0].name << "--" << (*this)[v1].name
+        << std::endl;
+
+    // boost::graph_traits<Graph>::adjacency_iterator v, v_end;
+    // for (boost::tie(v, v_end) = boost::adjacent_vertices(vertex, *_graph);
+    //     v != v_end; v++)
+    // {
+    //     if (_vertices.find(*v) != _vertices.end()) {
+    //         Edge_ID e = boost::edge(vertex, *v, *_graph).first;
+    //         _edges.insert(e);
[…]

### 2015-04-23 src/Graph/Graph.hpp

@@ -38,6 +38,16 @@ struct Vertex_Properties {
     }
 };
 
+// struct vertex_copier {
+//   template <typename S, typename D>
+//   void operator()(const S& /*src*/, D& /*dest*/) {
+//     // Nothing for now, deduced types to try and just make it compile
+//     // TODO: set values for pos to reflect position on grid.
+//   }
+// };
+// void vertex_copier(Vertex_Properties v0, Vertex_Properties v1);
+
+
 
 struct Edge_Properties {
     double length;
@@ -53,6 +63,9 @@ struct Edge_Properties {
         // pebbles = 0;
     }
 };
+
+// void edge_copier(Edge_Properties e0, Edge_Properties e1);
+
 
 // typedef boost::adjacency_list<boost::listS,
 //                 boost::listS,
@@ -72,7 +85,7 @@ typedef boost::graph_traits<Graph_Type>::edge_descriptor Edge_ID;
 typedef boost::graph_traits<Graph_Type>::vertex_iterator Vertex_Iterator;
 typedef boost::graph_traits<Graph_Type>::edge_iterator Edge_Iterator;
 
-
+typedef boost::graph_traits<Graph_Type>::adjacency_iterator Graph_Adj_Iterator;
 
 
 
@@ -86,10 +99,22 @@ public:
     Vertex_ID add_vertex();
 
     Edge_ID add_edge(Vertex_ID v0, Vertex_ID v1, Edge_Properties ep);
-    Edge_ID add_edge(Vertex_ID v0, Vertex_ID v1, double distance);
+    Edge_ID add_edge(Vertex_ID v0, Vertex_ID v1, double length);
+
+    void remove_vertex(Vertex_ID v);
+    void remove_edge(Edge_ID e);
+    void remove_edge(Vertex_ID v0, Vertex_ID v1);
+
+    bool has_edge(Vertex_ID v0, Vertex_ID v1);
+
+    // v0 disappears, v1 remains
+    void contract_edge(Vertex_ID v0, Vertex_ID v1);
+    // bool is_series_parallel();
 
     std::pair<Vertex_Iterator, Vertex_Iterator> vertices() const;
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
+
+    unsigned int degree_of_vertex(Vertex_ID v);
 
     Vertex_Iterator find_vertex(const char *name) const;
 
@@ -103,6 +128,8 @@ public:
 
     void write_to_file(const char* filename);
[…]

### 2015-04-23 src/Graph/Isostatic_Graph_Realizer.cpp

@@ -1,0 +1,175 @@
+#include "Isostatic_Graph_Realizer.hpp"
+
+#include "MinPriorityQueue.hpp"
+
+#include <iostream>
+
+
+
+graph_copy::graph_copy(Graph *g) {
+    orig = g;
+
+    Vertex_Iterator v, v_end;
+    for (boost::tie(v, v_end) = orig->vertices(); v != v_end; v++) {
+        // Vertex_ID newv = copy.add_vertex();
+        Vertex_ID newv = copy.add_vertex((*orig)[*v]);
+        orig_to_copy[*v] = newv;
+        copy_to_orig[newv] = *v;
+    }
+
+    Edge_Iterator e, e_end;
+    for (boost::tie(e, e_end) = orig->edges(); e != e_end; e++) {
+        std::pair<Vertex_ID, Vertex_ID> vs = orig->verts_on_edge(*e);
+        copy.add_edge(orig_to_copy[vs.first], orig_to_copy[vs.second], (*orig)[*e].length);
+    }
+}
+
+
+Isostatic_Graph_Realizer::Isostatic_Graph_Realizer(Graph *g) {
+    in_graph = g;
+}
+
+void Isostatic_Graph_Realizer::realize() {
+    // graph_copy gc(in_graph);
+    // bool partial_2_tree = is_partial_2_tree(gc);
+
+    make_partial_2_tree();
+
+    // std::cout
+    //     << "Partial 2 tree? "
+    //     << (partial_2_tree? "true": "false")
+    //     << std::endl;
+}
+
+bool Isostatic_Graph_Realizer::is_partial_2_tree(graph_copy &gc) {
+    MinPriorityQueue<Vertex_ID, unsigned int> mpq;
+    Vertex_Iterator v, v_end;
+    for (boost::tie(v, v_end) = gc.copy.vertices(); v != v_end; v++) {
+        mpq.insert(*v, gc.copy.degree_of_vertex(*v));
+    }
+
+    // while (!mpq.empty()) {
+    //     std::pair<Vertex_ID, unsigned int> top = mpq.pop_top();
+    //     std::cout << "(" << (*gc.copy)[top.first].name << ", " << top.second << ") ";
+    // }
+    // std::cout << std::endl;
+
+    while (!mpq.empty()) {
+        std::pair<Vertex_ID, unsigned int> top = mpq.pop_top();
+        std::cout << "(" << (gc.copy)[top.first].name << ", " << top.second << ")" << std::endl;
+
[…]

### 2015-04-23 src/Graph/Isostatic_Graph_Realizer.hpp

@@ -1,0 +1,46 @@
+#ifndef ISOSTATIC_GRAPH_REALIZER_HPP
+#define ISOSTATIC_GRAPH_REALIZER_HPP
+
+#include "Graph.hpp"
+
+#include <vector>
+#include <map>
+#include <utility>
+
+
+
+struct graph_copy {
+    Graph *orig;
+
+    Graph copy;
+    std::map<Vertex_ID, Vertex_ID> orig_to_copy, copy_to_orig;
+
+    graph_copy(Graph *g);
+};
+
+
+class Isostatic_Graph_Realizer {
+public:
+    // assumes that input is isostatic
+    Isostatic_Graph_Realizer(Graph *g);
+
+    void realize();
+
+private:
+    Graph *in_graph;
+
+    std::vector<std::pair<Vertex_ID, Vertex_ID> > contractions;
+
+    // returns a list of dropped edges
+    std::vector<std::pair<Vertex_ID, Vertex_ID> > make_partial_2_tree();
+
+    // assumes in_graph is now a partial 2 tree
+    // returns a list of added edges
+    std::vector<Edge_ID> make_2_tree();
+
+    // makes a copy of in_graph,
+    // destructively determines if it's partial 2-tree (series parallel, k4 minor avoiding...)
+    bool is_partial_2_tree(graph_copy &gc);
+};
+
+#endif

### 2015-04-23 src/Graph/MinPriorityQueue.hpp

@@ -1,0 +1,128 @@
+#ifndef MINPRIORITYQUEUE_HPP
+#define MINPRIORITYQUEUE_HPP
+
+#include <vector>
+#include <map>
+// #include <iostream>
+
+template<typename ID, typename Key>
+class MinPriorityQueue {
+private:
+    std::vector< std::pair<ID, Key> > heap;
+    std::map<ID, unsigned int> map;
+
+    unsigned int parent_idx(unsigned int idx) {
+        return ((int) (idx+1)/2) - 1;
+    }
+
+    unsigned int left_child_idx(unsigned int idx) {
+        return (int) idx*2+1;
+    }
+
+    unsigned int right_child_idx(unsigned int idx) {
+        return (int) idx*2+2;
+    }
+
+    void swap(unsigned int a, unsigned int b) {
+        std::pair<ID, Key> temp = heap[a];
+        heap[a] = heap[b];
+        heap[b] = temp;
+
+        map[heap[a].first] = a;
+        map[heap[b].first] = b;
+    }
+
+    void erase_last() {
+        map.erase(heap.back().first);
+        heap.pop_back();
+    }
+
+    // assumes left and right children are a heap, but from the parent it’s not a heap
+    void heapify_down(unsigned int idx) {
+        while (true) {
+            unsigned int lc = left_child_idx(idx);
+            unsigned int rc = right_child_idx(idx);
+
+            // determine which is the smallest: parent, left or right child
+            unsigned int smallest = idx;
+            if ( lc < heap.size()  &&  heap[lc].second <= heap[idx].second )
+                smallest = lc;
+            if ( rc < heap.size()  &&  heap[rc].second <= heap[smallest].second )
+                smallest = rc;
+
+            if (smallest != idx) {
+                swap(idx, smallest);
+                idx = smallest;
+            } else {
+                break;
+            }
+        }
+
+
+        // if (smallest != idx) {
+        //     swap(idx, smallest);
+        //     heapify_down(smallest); // Now the child that swapped may not be a heap
+        // }
+    }
+
+    void heapify_up(unsigned int idx) {
+        unsigned int parent = parent_idx(idx);
+        while (idx != 0 && heap[idx].second < heap[parent].second) {
+            swap(parent, idx);
+            idx = parent;
+            parent = parent_idx(idx);
+        }
+    }
+public:
+    MinPriorityQueue() {}
+
+    bool empty() {
[…]

### 2015-04-23 src/main.cpp

@@ -1,5 +1,6 @@
 #include "Graph/Graph.hpp"
 #include "Graph/Pebbled_Graph.hpp"
+#include "Graph/Isostatic_Graph_Realizer.hpp"
 
 #include "GUI/MainGuiManager.hpp"
 #include "GUI/Vision_Handler_2D.hpp"
@@ -670,7 +671,44 @@ int main(int argc, char **argv) {
 
         cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
     } else if (runtime_option == ro_test) {
-        cout << "Nothing." << endl;
+        cout << "Testing edge contraction." << endl;
+
+        Isostatic_Graph_Realizer isg(&g);
+        isg.realize();
+
+        // int i = 0;
+
+        // cout << "Original Vertices (" << g.num_vertices() << " vertices, "<< g.num_edges() << " edges): " << endl;
+        // std::pair<Vertex_Iterator, Vertex_Iterator> vs = g.vertices();
+        // for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
+        //     // cout << "\t" << i++ << ": " << *v_it << endl;
+        //     cout << "\t" << i++ << ": " << g[*v_it].name << endl;
+        // }
+
+        // // Vertex_ID v0 = *vs.first;
+        // // vs.first++;
+        // // Vertex_ID v1 = *vs.first;
+        // g.contract_edge(*(g.find_vertex("e")), *(g.find_vertex("c")));
+        // g.contract_edge(*(g.find_vertex("g")), *(g.find_vertex("c")));
+        // g.contract_edge(*(g.find_vertex("a")), *(g.find_vertex("b")));
+        // g.contract_edge(*(g.find_vertex("h")), *(g.find_vertex("i")));
+        // g.contract_edge(*(g.find_vertex("f")), *(g.find_vertex("d")));
+
+
+        // i = 0;
+        // cout << "New Vertices (" << g.num_vertices() << " vertices, "<< g.num_edges() << " edges): " << endl;
+        // vs = g.vertices();
+        // for (Vertex_Iterator v_it = vs.first; v_it != vs.second; v_it++) {
+        //     // cout << "\t" << i++ << ": " << *v_it << endl;
+        //     cout << "\t" << i++ << ": " << g[*v_it].name << endl;
+        // }
+
+        myWindow.update_graph_positions();
+        myWindow.update_display();
[…]

### 2015-04-23 doc/GMP2015/revision_2/s01_introduction.tex
Working on the paper revision.

@@
-% \section{Response to reviewers}
-
-% \cutout{This means} the text was removed.
-% \addin{This means} the text was added.
-% \notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
-% % \movedfrom{This means} the text was moved from another part of the document.
-
-% \subsection{Reviewer 1}
-% No suggestions made.
-
-% \subsection{Reviewer 2}
-% % \noindent
-% % Changes made:
-% \begin{itemize}
-%   \item We added a new example in figures to illustrate definitions/proofs/comparisons.
-And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
-%   \item Abstract was shortened and clarified.
-%   \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}.
-Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}.
-Slight modification to definition of DR-plan.
-%   \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
-%   \item Section \ref{sec:DRP} had some proofs moved back in from the appendix.
-Section \ref{sec:recomb} had a proof moved back in from the appendix.
-Section \ref{sec:conclusion} had future work moved back in from the appendix.
-%   \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
-%   \item Images were properly aligned and cropped.
-% \end{itemize}
-
-% \subsection{Reviewer 3}
-% We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.
-
 \section{Introduction}
@@
 
+\todo{Main overview/intro.}
+Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields.
+However, the design and analysis of these systems is made difficult by i
+
+\subsection{Recursive Decomposition Problem}
+The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph.
+We call these systems \nameit.
+This characterization has broad applicability in practice, and we provide several concrete examples.
+
+Such decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
+In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which, in general, is NP-hard.
+Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan.
+By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan; we show that all canonical DR-plans are optimal for systems that are not overconstrained.
+This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties.
+In fact, we conjecture that one of the earlier algorithms also produces an optimal DR-plan and that every optimal DR-plan is in some formal sense canonical.
+
+% Knowing an optimal DR-plan is crucial when working with isostatic geometric constraint systems with an underlying sparsity condition \nameit.
+It allows for efficient design and analysis of these systems with various desirable properties.
+The importance of an optimal plan is particularly evident when these systems \nameit are self-similar or quasi-uniform, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}, as the complexity of the plan is even further reduced.
+
+When realizing such a system \nameit, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints).
+Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
+With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
+Thus, finding an optimal DR-plan is important for realization.
+
+\subsection{Recombination Problem}
+\todo{The second problem we consider...}
+We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination.
+Once a subsystem is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
+This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.
+
+\subsection{Introducing Qusecs}
+When working on these problems, there is a large, natural class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint systems''.
+%
 Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers.
-As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements.
-Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either
-\vemph{isostatic or underconstrained} (\vemph{not self-stressed},\seedefsd).
+As illustrated by the examples below, the structure within each layer is often: self-similar\footnote{In this manuscript we only study finite 2D structures.
+\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.}~\cite{2012arXiv1204.6389G}, spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and composed of a few repeated motifs appearing in disordered arrangements.
+Note that a layer is not necessarily planar, it can consist of multiple, inter-constraining planar (genus 0) monolayers.
+Furthermore, a layer is often  either \vemph{isostatic} or \vemph{underconstrained} (\vemph{not self-stressed}, \seedefsd).
 A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
-
-\noindent
-\textbf{Note on Scope:} In this manuscript we only study finite 2D structures.
-\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.
+%
+The importance of an optimal DR-plan is particularly evident when a system is a qusecs.
+The quasi-uniform or self-similar properties mean that solutions for one subsystem can solve other subsystems, thus causing further reduction in the complexity of the plan.
+This is shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
 
@@
 
-Examples of such materials (See Figure \ref{fig:material_examples})
-include:
+% Examples of such materials (See Figure \ref{fig:material_examples}) include:
+% In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints.
+% Such 2D \dfn{qusecs}, quasi-uniform or self-similar constraint systems \seedefsprelim, can be used to understand or design material layers (their solutions) with desired properties.
+Some materials that are readily modeled as qusecs include:
 %
@@
 
-    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
+    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski} (Figure \ref{fig:material_examples:sierpinski}), bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
 
@@
 \end{enumerate}
-%
-In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints.
-Such 2D \dfn{qusecs}, quasi-uniform or self-similar constraint systems \seedefsprelim, can be used to understand or design material layers (their solutions) with desired properties.
 
-\subsection{Contributions and Organization}
+% \subsection{old}
+
+\subsection{Organization and Contributions}
 \label{sec:cont}
 
-The contributions of this paper are the following.
-\begin{itemize}
-  \item In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
-This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.
+\todo{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
+This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}
 
-  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier (discussed in the following subsection), for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.
+% In Section \ref{sec:DRP}, we navigate the NP-hardness barrier (discussed in the following subsection), for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.
 
-  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
-The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
+% Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
+% The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
+% Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
+
+In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal.
+In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
 Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
+Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
 
-  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
+% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
+% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
+
+In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
 Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
@@
 
-  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
+\todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification.
+By dropping edges to get a convex configuration space, realizations of the original linkage can be found.
+Also problem relations...}
+
+While both of these sections focus on bar-joint graphs, the theory is easily extended to other qusecs.
+The following sections focus on explicit applications by modeling rigidity in certain materials as qusecs.
+
+In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
 For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems.
@@
 
-  \item We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
-\end{itemize}
+We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
 
-\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
-\label{sec:intro:prevwork}
+\subsection{The Importance of an Optimal DR-plan}
+% \subsection{Importance of an Optimal DR-plan}
+% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
 
-We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
-The limitations of these techniques directly motivate the contributions of this paper.
+% In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
 
-\medskip\noindent
-\header{(i) Finding (vertex)-maximal generically rigid subsystems}
-Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs.
-When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.
-
-However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
-
-\medskip\noindent
-\header{(ii) Optimal Recursive Decomposition (DR-planning).}
-Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim.
-For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system.
-Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}.
-An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.
-Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.
-
-However, for general 2D qusecs that could be overconstrained, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph}.
-
-\medskip\noindent
-\header{(iii) DR-plans for special classes and with other
-criteria.}
-For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.
-This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
-For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
-\seedefsc.
-
-While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion \cite{hoffman2001decompositionI}, no such algorithm is known for the latter criterion.
-
-\medskip\noindent
-\header{(iv) Optimal Recombination and Solution Space Navigation.}
-For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
-This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
-Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan.
-For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
-
-However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan.
-They are prohibitively slow in practice.
-We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
-
-\medskip\noindent
-\header{(v) Configuration Spaces of Underconstrained Systems.}
-For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
-Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
-While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
-
-% \medskip\noindent
-% \note All proofs appear in the Appendix.
-% \medskip
+Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan.
+Each block is the rigidity matrix of an isostatic subgraph.
+The blocks intersect on isostatic or trivial subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
+Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
+When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
+When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.

### 2015-04-23 doc/GMP2015/revision_2/s02_preliminaries.tex
Working on the paper revision.

@@
 
-\subsection{Importance of an Optimal DR-plan}
-The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
-
-In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
-
-\medskip\noindent
-\header{Realizing given qusecs.}
-Note that this basic task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints).
-Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
-With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
-Thus, finding an optimal DR-plan is important for realization.
-
-\medskip\noindent
-\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
-This can be obtained using the DR-plan, with each block being the rigidity matrix of an isostatic subgraph.
-The blocks intersect on isostatic or trivial subgraphs.
-This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
-Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
-When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
-When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.
-
 %\subsection{DR-plan}

### 2015-04-23 doc/GMP2015/revision_2/s03_drp.tex
Working on the paper revision.

@@
 
+\subsection{Previous Work}
+\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+The limitations of these techniques directly motivate the contributions of this paper.}
+
+\begin{enumerate}
+    \item
+    \header{Finding (vertex)-maximal generically rigid subsystems}
+    Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs.
+When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.
+
+    However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
+
+    \item
+    \header{Optimal Recursive Decomposition (DR-planning).}
+    Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim.
+For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system.
+Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}.
+An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.
+Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.
+
+    However, for general 2D qusecs that could be overconstrained, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph}.
+
+    \item
+    \header{DR-plans for special classes and with other criteria.}
+    For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.
+This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms.
+For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
+\seedefsc.
+
+    \indent
+    While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion \cite{hoffman2001decompositionI}, no such algorithm is known for the latter criterion.
+\end{enumerate}
+
+\subsection{Theory}
 \sidenote{In this section, $G$ is always assumed to be isostatic.}
@@
 
-% \subsection{Algorithm}
+\subsection{Algorithm}
 

### 2015-04-23 doc/GMP2015/revision_2/s04_recombination.tex
Working on the paper revision.

@@
-\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
+% \section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
+\section{Recombination and Problem Relationships}
 \label{sec:recomb}
@@
 
-\medskip\noindent
-\header{(a) Small DR-plan.}
+\begin{enumerate}[(a)]
+    \item
+    \header{Small DR-plan.}
     We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans.
@@
 
-\medskip\noindent
-\header{(b) Solution preservation.}
+    \item
+    \header{Solution preservation.}
     Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.
 
-\medskip\noindent
-\header{(c) Efficient search.}
+    \item
+    \header{Efficient search.}
     Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type.
@@
 The standard optimization versions are straightforward.
+\end{enumerate}
 
@@
 However, it is not clear how Criterion (c) is met by OMD graphs.
+
+\subsection{Previous Work}
+\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+The limitations of these techniques directly motivate the contributions of this paper.}
+
+\begin{enumerate}
+    \item
+    \header{Optimal Recombination and Solution Space Navigation.}
+    For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
+This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
+Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan.
+For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
+
+    However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan.
+They are prohibitively slow in practice.
+We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
+
+    \item
+    \header{Configuration Spaces of Underconstrained Systems.}
+    For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
+Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
+While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
+\end{enumerate}
 

### 2015-04-23 doc/GMP2015/revision_2/s05_problems.tex
Working on the paper revision.

@@
-\section{Problem Relationships and Unification}
+\subsection{Problem Relationships}
 \label{sec:table}
@@
 
-\subsection{Special classes of small DR-plans}
+\subsubsection{Special classes of small DR-plans}
 As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially a single quadratic at a time.
@@
 %
-\subsection{Optimal Modification, Completion and Recombination: formal
+\subsubsection{Optimal Modification, Completion and Recombination: formal
 connections}

### 2015-05-07 doc/GMP2015/revision_2/appendix_defs.tex
Major changes to paper (revision 2)

@@
 \label{sec:appendix:defs}
+% \label{sec:prelim:defs}
 
-In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
-%
-When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where $G$ is the underlying constraint (hyper)graph $G = (V,E)$ with the vertices $V$ representing the geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$ representing the constraints, each with an associated parameter $\delta$.
-%
-For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a graph $G=(V,E)$ with fixed length bars as edges, i.e.
-$\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
-
-% \noindent
-% \note Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.
-
-\sidenote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}
-
-In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
-In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
-
-Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values).
-Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
-For example, the \dfn{bar-joint rigidity matrix of the graph $G = (V,E)$} is a matrix of indeterminates representing the Jacobian of the distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$.
-The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp.
-$x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.
-
-% XXXXXXXXX
-% something has to be said along with asimow roth
-% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
-% appropriate rank, the realizations or solutions of the corresponding
-% constraint system are generically isolated and zero-dimensional if the
-% constraint system has equalities and exact values for the parameters
-% $\delta$. If the constraint system has inequalities, i.e, if the
-% $\delta$ lie in an interval, this claim is approximate: the solutions
-% are isolated small, full-dimensional  connected components.
-
-One important property of generic constraint system or (hyper)graph
-%that depends on the appropriate rigidity matrix of indeterminates
-is \dfn{rigidity} (\note we refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably), i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
-%This condition is called generic {\em infinitesimal rigidity}.
-
-% \noindent
-% \note If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.
-
-\sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
-% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
-
-Other generic constraint system or (hyper)graph properties are the following:
-% (\note we refer to these as properties of the constraint
-% system or as properties of the underlying (hyper)graph
-% interchangeably).
-A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
-%It is {\em rigid} if the number of independent rows of the rigidity matrix is
-%at least the number of columns less the number of rigid body motions,
-%which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
-%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
-%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
-%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
-%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
-It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it is independent and not rigid, or \dfn{overconstrained} if it is not independent.
-
-Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
-%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
-%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
-%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs.
-For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices.
-The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane.
-For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees.
-White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition.
-Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$.
-In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
-
-% In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.
-
-% % \medskip\noindent
-% % \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
-Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.
-
-% \sidenote{In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
-Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications (\ref{sec:bodypin} and \ref{sec:pinnedline}).}
-
-%XXXNow comes the definition of stress vector and flex vector.
-%NOTE: something has to be said about what flexes and stresses mean
-
-%TODO
-Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}.
-It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions).
-For rigid graphs, all flex vectors are identically zero.
-%
-A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}.
-It has one entry per edge and represents a \dfn{self stress} of the system.
-For independent graphs, all stress vectors are identically zero.
-%
-An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
+Moved into prelims section.

### 2015-05-07 doc/GMP2015/revision_2/s01_introduction.tex
Major changes to paper (revision 2)

@@
 
-\todo{Main overview/intro.}
+% \todo{Regarding qusecs... We use the standard way of dealing w/ these problems, namely modeling it as a geometric constraint system...
+% \todo{Restructure intro so that we have...
+problems first...
+closing with ``Concerning these issue...
+this is what we do in relation to them...
+the specifics we address..'' and go into summary of our contributions}
+
 Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields.
-However, the design and analysis of these systems is made difficult by i
+Some systems have an underlying sparsity condition that characterize rigidity, allowing for matroidal techniques of rigidity detection and decomposition.
+However, the design and analysis of these systems is complicated by the NP-hardness barrier of creating optimal decomposition-recombination (DR-) plans.
+While other papers have focused on a few desirable criteria and made use of hueristics, by focusing on the class of independent (isostatic or underconstrained) systems we have found an efficient way to construct the optimal DR-plan.
+Additionally, in real world applications, many of these systems exhibit degrees of self-similarity and quasi-uniformity.
+Our formulation of the optimal DR-plan can leverage these properties, further reducing the complexity of the DR-plan.
+Furthermore, we address the fact that indecomposable subsystems in a system can be arbitrarily large, representing a bottle-neck in the computational efficiency in the analysis of these systems.
+We also address this problem with an efficient technique of optimal modification for further decomposition.
+
+% In this paper, we introduce new techniques for the recursive decomposition and recombination of two dimensional independent (isostatic or underconstrained) constraint systems.
+The main contributions are a method to generate an optimal decomposition-recombination (DR-) plan
+
+%  leveraging the quasi-uniformity and self-similarity present in many real world applications,
 
@@
 The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph.
-We call these systems \nameit.
 This characterization has broad applicability in practice, and we provide several concrete examples.
 
-Such decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
-In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which, in general, is NP-hard.
+% Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
+When realizing these sparsity-condition constraint systems, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints).
+Without decomposition to plan the recombination (DR-planning), this requires double exponential time in the number of variables (even if orientation type is specified).
+With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
+Thus, finding an optimal DR-plan is important for realization.
+
+Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
+When realizing constraint systems, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints); this requires double exponential time in the number of variable, even if orientation type is specified.
+% Without decomposition to plan the recombination (DR-planning), this requires double exponential time in the number of variables (even if orientation type is specified).
+With the use of decomposition-recomination (DR-) planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
+Thus, finding an optimal DR-plan is important for realization.
+
+% \subsection{The Importance of an Optimal DR-plan}
+% \subsection{Importance of an Optimal DR-plan}
+% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
+
+% In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
+
+% Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan.
+Each block is the rigidity matrix of an isostatic subgraph.
+The blocks intersect on isostatic or trivial subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
+Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
+When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
+When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.
+
+Furthermore, the DR-plan can be used to obtain recursive block decomposition of the rigidity matrix, stress space, and external stresses.
+Each block is the rigidity matrix of an isostatic subgraph.
+The blocks intersect on isostatic or trivial subgraphs.
+This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
+Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
+When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
+When the realization of the system is given, this can be used to analyze the distribution of a given external load.
+
+% In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which,
+An optimal DR-plan decreases the computation involved in such tasks; however, in general, finding an optimal DR-plan is NP-hard.
 Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan.
-By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan; we show that all canonical DR-plans are optimal for systems that are not overconstrained.
+By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan.
+Then we show that all canonical DR-plans are optimal for systems that are not overconstrained.
 This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties.
@@
 
-When realizing such a system \nameit, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints).
-Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified).
-With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
-Thus, finding an optimal DR-plan is important for realization.
+\subsection{Recombination Problem}
+The second problem we consider is the complexity of recombination.
+Na{\"i}vely, recombination can be done by finding the solutions to the polynomial system of equations describing the constraints between the children; however, even in an optimal DR-plan, there can be arbitrarily many children at a node.
 
-\subsection{Recombination Problem}
-\todo{The second problem we consider...}
-We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination.
+% Instead, we propose the modification of the subsystem to find a convex configuration space that can be efficiently searched for a solution to the original system.
+
+Instead, we make use of the notion of Cayley complexity to quantify the difficulty of each step of recombination.
+Once a subsystem of a node is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
+This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.
+
+% We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination.
 Once a subsystem is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
-This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.
+% This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.
 
@@
 
-% \subsection{old}
-
 \subsection{Organization and Contributions}
@@
 
-\todo{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan.
-This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}
+In Section \ref{sec:prelim}, we provide basic definitions in rigidity theory, including what it means for a DR-plan to be complete or optimal. We also elaborate on the new notion of qusecs.
+This section is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.
+\todo{We don't model them as bar-joint...
+What did this mean originally?}
 
@@
 Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
+%
+When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
+%
 In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components.
-When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
 
-\todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification.
+% \todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification.
 By dropping edges to get a convex configuration space, realizations of the original linkage can be found.
@@
 
-We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
-
-\subsection{The Importance of an Optimal DR-plan}
-% \subsection{Importance of an Optimal DR-plan}
-% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
-
-% In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.
-
-Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan.
-Each block is the rigidity matrix of an isostatic subgraph.
-The blocks intersect on isostatic or trivial subgraphs.
-This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
-Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
-When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
-When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.
+\todo{We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.}

### 2015-05-07 doc/GMP2015/revision_2/s02_preliminaries.tex
Major changes to paper (revision 2)

@@
 
-Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.
+% Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.
 
@@
 
+\subsection{Basic Definitions and Theory}
+\label{sec:prelim:defs}
+\label{sec:appendix:defs}
+
+In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
+%
+When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where $G$ is the underlying constraint (hyper)graph $G = (V,E)$ with the vertices $V$ representing the geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$ representing the constraints, each with an associated parameter $\delta$.
+%
+For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a graph $G=(V,E)$ with fixed length bars as edges,
+% \footnote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}
+i.e.
+$\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
+
+In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
+In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
+
+Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values).
+Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
+For example, the \dfn{bar-joint rigidity matrix of the graph $G = (V,E)$} is a matrix of indeterminates representing the Jacobian of the distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$.
+The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp.
+$x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.
+
+% XXXXXXXXX
+% something has to be said along with asimow roth
+% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
+% appropriate rank, the realizations or solutions of the corresponding
+% constraint system are generically isolated and zero-dimensional if the
+% constraint system has equalities and exact values for the parameters
+% $\delta$. If the constraint system has inequalities, i.e, if the
+% $\delta$ lie in an interval, this claim is approximate: the solutions
+% are isolated small, full-dimensional  connected components.
+
+One important property of generic constraint system or (hyper)graph\footnote{We refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably.}
+%that depends on the appropriate rigidity matrix of indeterminates
+is \dfn{rigidity},
+% (\note footnote now...),
+i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
+The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
+%This condition is called generic {\em infinitesimal rigidity}.
+
+% \sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
+% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system
+
+Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of rather than exact values.
+In this case, the definition of rigidity is approximate; the solutions are isolated, small, full-dimensional connected components.
+
+Other generic constraint system or (hyper)graph properties are the following:
+% (\note we refer to these as properties of the constraint
+% system or as properties of the underlying (hyper)graph
+% interchangeably).
+A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
+%It is {\em rigid} if the number of independent rows of the rigidity matrix is
+%at least the number of columns less the number of rigid body motions,
+%which is 3 for distance constraint systems.
+It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
+%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
+%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
+%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
+%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
+It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it is independent and not rigid, or \dfn{overconstrained} if it is not independent.
+
+Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
+%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
+%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
+%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
+There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs.
+For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices.
+The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane.
+For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees.
+White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition.
+Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$.
+In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
+
+% In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.
+
+% % \medskip\noindent
+% % \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.
+
+% \sidenote{In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications (\ref{sec:bodypin} and \ref{sec:pinnedline}).}
+
+%XXXNow comes the definition of stress vector and flex vector.
+%NOTE: something has to be said about what flexes and stresses mean
+
+%TODO
+Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}.
+It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions).
+For rigid graphs, all flex vectors are identically zero.
+%
+A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}.
+It has one entry per edge and represents a \dfn{self stress} of the system.
+For independent graphs, all stress vectors are identically zero.
+%
+An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
+
 \subsection{Qusecs}
 
-In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:appendix:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
+In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:prelim:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
 
@@
 
+% \ClearMyMinHeight
+% \SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
+% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
+% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}
+
+% \begin{figure*}\centering%
+%   %
+%   \begin{subfigure}{0.4\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
+%     \caption{}\label{fig:demo_graph:graph}
+%   \end{subfigure}%
+%   %
+%   \hfill
+%   \begin{subfigure}{0.3\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
+%     \caption{}\label{fig:demo_graph:comdrp}
+%   \end{subfigure}%
+%   %
+%   \hfill
+%   \begin{subfigure}{0.3\linewidth}\centering
+%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
+%     \caption{}\label{fig:demo_graph:candrp}
+%   \end{subfigure}%
+%   %
+%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
+(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
+Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
+The children of triangles (3 edges) are omitted.
+(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
+The children of triangles are omitted.}
+%   \label{fig:demo_graph}
+% \end{figure*}%
+
+% \begin{figure*}\centering%
+%   %
+%   \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
+%   \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
+Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge.
+Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2.
+With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
+%   \label{fig:demo_graph:clustmindrp}
+% \end{figure*}%
+
 \ClearMyMinHeight
-\SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
-\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
-\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}
+\SetMyMinHeight{.4}{img/svg/3xc2c3}
+\SetMyMinHeight{.3}{img/svg/3xc2c3_comdrp}
+\SetMyMinHeight{.3}{img/svg/3xc2c3_candrp}
 
@@
   \begin{subfigure}{0.4\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
+    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3}
     \caption{}\label{fig:demo_graph:graph}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
+    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_comdrp}
     \caption{}\label{fig:demo_graph:comdrp}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
+    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_candrp}
     \caption{}\label{fig:demo_graph:candrp}
@@
   %
-  \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
-(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
+  \caption{(\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section.
+(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$, i.e.\ $ComDRP(G_{demo})$.
 Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
 The children of triangles (3 edges) are omitted.
-(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
+(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{demo})$.
 The children of triangles are omitted.}
@@
 
+\ClearMyMinHeight
+\SetMyMinHeight{.4}{img/svg/triangle}
+\SetMyMinHeight{.3}{img/svg/triangle_candrp}
+\SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}
+
 \begin{figure*}\centering%
   %
-  \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
-  \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
+  \begin{subfigure}{0.4\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/svg/triangle}
+    \caption{}\label{fig:demo_graph_tri:graph}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
+    \caption{}\label{fig:demo_graph_tri:candrp}
+  \end{subfigure}%
+  %
+  \hfill
+  \begin{subfigure}{0.3\linewidth}\centering
+    \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
+    \caption{}\label{fig:demo_graph_tri:clustmindrp}
+  \end{subfigure}%
+  %
+  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially.
+(\ref{fig:demo_graph_tri:candrp}) The canonical DR-plan of $G_{tri}$.
+(\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality.
 Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge.
-Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2.
+Thus, the fan-in of this graph is 3, whereas the canonical/optimal DR-plan has fan-in of 2.
 With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
   \label{fig:demo_graph:clustmindrp}
+  \label{fig:demo_graph_tri}
 \end{figure*}%

### 2015-05-07 doc/GMP2015/revision_2/s03_drp.tex
Major changes to paper (revision 2)

@@
 
-In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.
+In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the uniqueness of a complete DR-plan, and we show that the nonunique parts do not affect optimality for independent (underconstrained or isostatic) graphs.
+Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph.
+The definition is as follows:
+
+\begin{definition}\label{def:canonical_drplan}
+    The \dfn{canonical DR-plan} satisfies the following three properties:
+    (1) it is a DR-plan;
+    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
+    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise take as children exactly two that intersect non-trivially.
+\end{definition}
 
 \subsection{Previous Work}
-\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
 The limitations of these techniques directly motivate the contributions of this paper.}
+We now briefly survey existing techniques for detecting rigidity and creating DR-plans of 2D constraint systems.
+The limitations of these techniques directly motivate the contributions in this section.
 
@@
 \subsection{Theory}
-\sidenote{In this section, $G$ is always assumed to be isostatic.}
+In this section and in section~\ref{sec:recomb}, we deal exclusively with isostatic constraint systems.
+Any reference to some graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+% \sidenote{In this section, $G$ is always assumed to be isostatic.}
 
-\begin{definition}\label{def:canonical_drplan}
-    The \dfn{canonical DR-plan} satisfies the following three properties:
-    (1) it is a DR-plan;
-    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise take as children exactly two that intersect non-trivially.
-\end{definition}
+% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
 
-This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
+Definition~\ref{def:canonical_drplan} gives the canonical DR-plan a surprisingly strong Church-Rosser property, which is made explicit in Theorem~\ref{theorem:main}, the main result of this section.
 
@@
 
-This directly satisfies properties 2 and 3 of a canonical DR-plan (see definition \ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs.
+This directly satisfies properties 2 and 3 of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs.
 However, we still have to show property 1 (that this makes a DR-plan).
@@
 
-This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.
+This proof of this theorem relied on the following crucial observation and lemma.
+These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.
 
@@
 \begin{lemma*}\label{lemma:combined_lemma}
-Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
+Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$.
 Then
@@
 
-\begin{figure*}\centering%
-  %
-  \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
-  \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
+% \begin{figure*}\centering%
+%   %
+%   \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
+%   \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
 The children of the triangle are omitted.
@@
 Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
+%   \label{fig:demo_graph:candrpseq}
+% \end{figure*}%
+\begin{figure*}\centering%
+  %
+  \includegraphics[width=0.3\linewidth]{img/svg/3xc2c3_candrp_seq}
+  \caption{The sequential canonical DR-plan of $G_{demo}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
+The children of the triangle are omitted.
+Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
+Also, note that the bottom-left node, the triangle, is the intersection of the 3 children of $G_{demo}$ in $ComDRP(G_{demo})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
   \label{fig:demo_graph:candrpseq}
@@
 \begin{proof}
-The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
+The first step of the algorithm, which we call $CanDRP(G)$, is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
 Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
@@
 
-Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.
-
-If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$.
-The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
+Test the intersection of any two of the true-components.
+If the intersection is trivial, the entire list become the children; recursively apply the algorithm to each child.
+If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k=1}^{N}{C_k}$.
+The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}.
 We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan.
 Begin with the parent $C$.
-Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$.
-Recursively apply the algorithm to the child containing $R_1$.
-The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
+Its children will be $S_1$ and $CanDRP(T_1)$, where $S_1=C\setminus R_1$ and $T_1$ is $R_1$ plus all incident edges (observe that these edges will add nodes from $I$ and that $T_1$ is underconstrained so it will be a forest and each root becomes a child of $C$).
+The children of $S_1$ will be $S_2$ and $CanDRP(T_2)$, where $S_2=C\setminus (R_1\cup R_2)$ and $T_2$ is $R_2$ plus all incident edges.
 And so on.
-At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
+At $N$ levels down, the children will be $S_N=I$ and $CanDRP(T_N)$, where $T_N$ is $R_N$ plus all incident edges.
 Apply the algorithm recursively to both children.
 See Figure~\ref{fig:demo_graph:candrpseq} for an example.
+\todo{should this more elaborate discussion be moved outside the proof?}
+
+The previous discussion was a detailed explanation with an implied efficient method to solve for the case of non-trivial intersections, which avoids recomputation of the isostatic components of $C$.
+This stage of the algorithm could be stated more succinctly:
+if the intersection is trivial, recursively apply $CanDRP$ to the entire list of true-components and set the resulting trees as the children.
+Else, the intersection is non-trivial, and the children of the node are $CanDRP(C_1)$ and $CanDRP(T)$, where $C_1$ is the first true-component of the parent $C$ (chosen arbitrarily, could be any child) and $T$ is the underconstrained graph formed from $C\setminus C_1$ plus all incident edges (also adding the associated nodes in $C_1$).
+
 % Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
 
-The reason for this construction is that the result is a tree with no duplicate nodes.
-The leaves of this tree will exactly be the edges of this graph.
+The reason for this construction is that it makes analysis of the algorithm convenient.
+The result of the algorithm is DR-plan that is a tree with no duplicate nodes.
+The leaves of this tree will be exactly the edges of this graph.
 Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.

### 2015-05-07 doc/GMP2015/revision_2/s04_recombination.tex
Major changes to paper (revision 2)

@@
 \subsection{Previous Work}
-\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
 The limitations of these techniques directly motivate the contributions of this paper.}
+We now briefly survey existing techniques for handling the complexity of recombination of 2D DR-plans. The limitations of these techniques directly motivate the contributions in this section.
 

### 2015-05-09 src/Graph/Graph.cpp
Can make a graph into a partial 2 tree (series parallel). Can add in edges to make that into a 2 tree. Can realize 2 trees (chooses orientation itself).

@@ -118,7 +118,32 @@ Edge_ID Graph::add_edge(Vertex_ID v0, Vertex_ID v1, double distance) {
 }
 
 void Graph::remove_vertex(Vertex_ID v) {
+    std::vector<Vertex_ID> vs;
+
+    for (std::pair<Graph_Adj_Iterator, Graph_Adj_Iterator> v_adj = boost::adjacent_vertices(v, *this);
+        v_adj.first != v_adj.second;
+        v_adj.first++)
+    {
+        vs.push_back(*(v_adj.first));
+    }
+
+    for (std::vector<Vertex_ID>::iterator v_it = vs.begin(); v_it != vs.end(); v_it++) {
+        // std::cout << "Removing edge: (" << (*this)[v].name << ", " << (*this)[*v_it].name << ")" << std::endl;
+        // std::cout << "\tDegree of " << (*this)[*v_it].name << " before: " << degree_of_vertex(*v_it) << std::endl;
+        remove_edge(v, *v_it);
+        // std::cout << "\tDegree of " << (*this)[*v_it].name << " after: " << degree_of_vertex(*v_it) << std::endl;
+    }
+
+    // for (std::vector<Vertex_ID>::iterator v_it = vs.begin(); v_it != vs.end(); v_it++) {
+    //     std::cout << "\tDegree of " << (*this)[*v_it].name << " before: " << degree_of_vertex(*v_it) << std::endl;
+    // }
+
+    // std::cout << "Removing vertex: " << (*this)[v].name << std::endl;
     boost::remove_vertex(v, *this);
+
+    // for (std::vector<Vertex_ID>::iterator v_it = vs.begin(); v_it != vs.end(); v_it++) {
+    //     std::cout << "\tDegree of " << (*this)[*v_it].name << " after: " << degree_of_vertex(*v_it) << std::endl;
+    // }
 }
 
 void Graph::remove_edge(Edge_ID e) {
@@ -134,11 +159,17 @@ bool Graph::has_edge(Vertex_ID v0, Vertex_ID v1) {
     return boost::edge(v0, v1, *this).second;
 }
 
+Edge_ID Graph::edge(Vertex_ID v0, Vertex_ID v1) {
+    return boost::edge(v0, v1, *this).first;
+}
+
 void Graph::contract_edge(Vertex_ID v0, Vertex_ID v1) {
-    std::cout
-        << "Graph::contract_edge: "
-        << (*this)[v0].name << "--" << (*this)[v1].name
-        << std::endl;
+    // std::cout
[…]

### 2015-05-09 src/Graph/Graph.hpp

@@ -36,6 +36,11 @@ struct Vertex_Properties {
         y = 0;
         color = 0x000000;
     }
+
+    void set_position(double x, double y) {
+        this->x = x; this->y = y;
+        point[0] = x; point[1] = y;
+    }
 };
 
 // struct vertex_copier {
@@ -106,6 +111,7 @@ public:
     void remove_edge(Vertex_ID v0, Vertex_ID v1);
 
     bool has_edge(Vertex_ID v0, Vertex_ID v1);
+    Edge_ID edge(Vertex_ID v0, Vertex_ID v1); // can be unsafe... edge may not exist
 
     // v0 disappears, v1 remains
     void contract_edge(Vertex_ID v0, Vertex_ID v1);
@@ -115,6 +121,8 @@ public:
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
     unsigned int degree_of_vertex(Vertex_ID v);
+    std::set<Vertex_ID> vertices_adjacent(Vertex_ID v);
+    std::set<Vertex_ID> vertices_adjacent(std::set<Vertex_ID> &v_set);
 
     Vertex_Iterator find_vertex(const char *name) const;
 
@@ -128,6 +136,9 @@ public:
 
     void write_to_file(const char* filename);
     void read_from_file(const char* filename);
+
+    void print_vertices(unsigned int indent = 0);
+    void print_edges(unsigned int indent = 0);
 private:
     // bool _is_series_parallel_aux();
 };

### 2015-05-09 src/Graph/Isostatic_Graph_Realizer.cpp

@@ -2,6 +2,7 @@
 
 #include "MinPriorityQueue.hpp"
 
+#include <cmath>
 #include <iostream>
 
 
@@ -17,6 +18,10 @@ graph_copy::graph_copy(Graph *g) {
         copy_to_orig[newv] = *v;
     }
 
+    // std::cout << "========graph_copy:" <<std::endl;
+    // print_copy_to_orig(1);
+    // copy.print_vertices(2);
+
     Edge_Iterator e, e_end;
     for (boost::tie(e, e_end) = orig->edges(); e != e_end; e++) {
         std::pair<Vertex_ID, Vertex_ID> vs = orig->verts_on_edge(*e);
@@ -24,16 +29,30 @@ graph_copy::graph_copy(Graph *g) {
     }
 }
 
+void graph_copy::print_copy_to_orig(unsigned int indent) {
+    for (int i = 0; i < indent; i++) std::cout << "\t";
+    std::cout << "copy_to_orig mapping (new, old):" << std::endl;
+
+    for (std::map<Vertex_ID, Vertex_ID>::iterator v = copy_to_orig.begin(); v!= copy_to_orig.end(); v++) {
+        for (int i = 0; i < indent+1; i++) std::cout << "\t";
+        // std::cout << "(" << v->first << ", " << v->second << ")" << std::endl;
+        std::cout << "(" << v->first << ": " << copy[v->first].name << ", " << v->second << ": " << (*orig)[v->second].name << ")" << std::endl;
+    }
+}
+
+
 
 Isostatic_Graph_Realizer::Isostatic_Graph_Realizer(Graph *g) {
     in_graph = g;
 }
 
-void Isostatic_Graph_Realizer::realize() {
+bool Isostatic_Graph_Realizer::realize() {
     // graph_copy gc(in_graph);
     // bool partial_2_tree = is_partial_2_tree(gc);
 
     make_partial_2_tree();
+    make_2_tree();
+    return realize_2_tree();
 
     // std::cout
     //     << "Partial 2 tree? "
@@ -56,7 +75,7 @@ bool Isostatic_Graph_Realizer::is_partial_2_tree(graph_copy &gc) {
 
     while (!mpq.empty()) {
         std::pair<Vertex_ID, unsigned int> top = mpq.pop_top();
-        std::cout << "(" << (gc.copy)[top.first].name << ", " << top.second << ")" << std::endl;
[…]

### 2015-05-09 src/Graph/Isostatic_Graph_Realizer.hpp

@@ -16,6 +16,8 @@ struct graph_copy {
     std::map<Vertex_ID, Vertex_ID> orig_to_copy, copy_to_orig;
 
     graph_copy(Graph *g);
+
+    void print_copy_to_orig(unsigned int indent = 0);
 };
 
 
@@ -24,23 +26,32 @@ public:
     // assumes that input is isostatic
     Isostatic_Graph_Realizer(Graph *g);
 
-    void realize();
+    bool realize();
 
 private:
     Graph *in_graph;
 
+    // used by methods
     std::vector<std::pair<Vertex_ID, Vertex_ID> > contractions;
+    std::map<Vertex_ID, bool> realized;
+
+    // makes a copy of in_graph,
+    // destructively determines if it's partial 2-tree (series parallel, k4 minor avoiding...)
+    bool is_partial_2_tree(graph_copy &gc);
 
     // returns a list of dropped edges
     std::vector<std::pair<Vertex_ID, Vertex_ID> > make_partial_2_tree();
 
     // assumes in_graph is now a partial 2 tree
     // returns a list of added edges
-    std::vector<Edge_ID> make_2_tree();
+    // std::vector<Edge_ID> make_2_tree();
+    std::vector<std::pair<Vertex_ID, Vertex_ID> > make_2_tree();
 
-    // makes a copy of in_graph,
-    // destructively determines if it's partial 2-tree (series parallel, k4 minor avoiding...)
-    bool is_partial_2_tree(graph_copy &gc);
+    // assumes in_graph is a 2 tree
+    // uses edge lengths and directly edits vertex properties (x and y)
+    // returns true if it can be realized, false otherwise
+    bool realize_2_tree();
+    bool _realize_2_tree_aux(Vertex_ID v0, Vertex_ID v1);
 };
 
 #endif

### 2015-05-09 src/Graph/MinPriorityQueue.hpp

@@ -80,6 +80,10 @@ public:
         return heap.size() == 0? true: false;
     }
 
+    unsigned int size() {
+        return heap.size();
+    }
+
     void insert(ID id, Key key) {
         if (map.find(id) != map.end()) // TODO: error
             return;
@@ -91,13 +95,19 @@ public:
         heapify_up(inserted);
     }
 
-    std::pair<ID, Key> pop_top() {
-        std::pair<ID, Key> ret = heap[0];
+    void pop() {
         swap(0, heap.size()-1);
         erase_last();
-
         heapify_down(0);
+    }
 
+    std::pair<ID, Key> top() {
+        return heap[0];
+    }
+
+    std::pair<ID, Key> pop_top() {
+        std::pair<ID, Key> ret = top();
+        pop();
         return ret;
     }
 

### 2015-05-16 doc/GMP2015/revision_2/response_to_reviewers.txt
Revision 2. 24 hours to submission.

Response to Reviewers
=====================

Reviewer #1
-----------
Overall I think this paper is timely and interesting.  It is very nice that the authors have included a number of open questions to emphasize that there is more work to be done in this area.  The notation and terminology is rather dense, however, this is probably unavoidable given the intricacy of the questions being addressed.

I don't think the abstract is significantly improved from the original version; there is still a lot of area-specific terminology that makes it hard to understand what the paper is about.  Perhaps a simple motivating example explained in detail over a couple sentences would help, especially explaining why one would want to model a particular ``quasi-uniform, self-similar, layered material structure'' using a decomposition of geometric constraint systems.

To that end, I would also suggest bringing the definition of geometric constraint system out of Appendix A and putting it right at the beginning of section 2, or perhaps section 2.1.

This line at the end of the abstract and at the end of section 2.1 should be modified for the final version of the paper (and applicable files should be included accordingly): ``We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.''

This paper seems essentially ready for publication.

Response:
- Abstract had substantial rework. While we don't give a complete example in the abstract, we do mention which materials we model in this paper. In the introduction, we do explain much more clearly why we would want to use qusecs.
- We moved the definitions into the body of the paper.
- We provide a link where the video and code will be made available. Currently a work in progress, but we are committed to providing a well-documented, open source, highly functional application. The video will be made sooner.

Reviewer #2
-----------
This paper proposes a recursive decomposition planning algorithm to model a material layer as a solution of a 2D geometric constraint system. While the work seems important in material modeling, the topic is a kind of out of focus of GMP audiences. Besides, I'm not quite satisfied with the organization and structure of the paper, and I feel it is not suitable for publication in current form.

1. The paper is far out of a 10 pages limit.
2. The basis concepts should be better put in the main context, but was put in an appendix (Appendix A).
3. The main algorithm in Section 4 should be stood out seperately and explained in details.
4. The examples should be enumerated.
5. The texts contain various formats and styles: Captioned, Bold fonts, colored, underlined, .... This unusual in scientific papers.

Response:
1. We are now submitting to the main journal.
2. We moved preliminaries into the main body of the paper (section 2).
3. We moved the algorithm into its own subsection and revised the proof to be more clear.
4. ...
5. We have addressed this. We have reformatted our unusual headings and such.

Further Changes
===============

Here we discuss additional changes the reviewers should be aware of.

One of the main issues with our paper was that it would not be a good fit for GMP. After much discussion, we decided that we agree and are now submitting the paper for consideration to be published in the standard journal (see next section in this file). The changes made in this revision are discussed below.

Title: Because of the substantial changes to the paper, we decided the title no longer made sense in the context of the revised paper. As such, we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'. We believe this also makes the intent of our paper much more clear.

Throughout: Many minor corrections to spelling, grammar, and punctuation. Additionally, many sentences were rephrased/restructured to improve clarity. Also, because our paper is no longer being considered for the special issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).

Two concepts were further fleshed out and discussed in this iteration: (a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and (b) the extension of our theory to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.. This is reflected throughout the paper, but is especially prevalent in (1), (2), (3), and (8).

Abstract: Substantially altered for clarity. Similar content.

(1) Introduction: Entirely rewritten. A major complaint was that it was difficult to get a big picture understanding of the paper without reading the entire thing. We believe we addressed this with our revised intro (and abstract). This also includes a thorough explanation of `qusecs' and why we believe this notion is important to define and give a name, directly addressing reviewer criticism. Also, the discussion of previous work was moved into the corresponding sections (2) and (4).

(2) Preliminaries: Moved back in the definitions from the appendix (as per reviewer request). Moved in discussion of previous work on DR-plans from (1).

(3) Canonical DR-plan: Improved the section introduction. Clarified/improved the definition of a canonical DR-plan. There are new figures to demonstrate the concepts; instead of the previous trivial (in the sense that it was tree-decomposable) example of three triangles, we use three $C2\times C3$s intersecting on a single triangle. Also, there is a new figure that corrects the previous, erroneous figure discussing cluster minimality in the previous revision. Revised proofs for clarity, but there were no fundamental changes. Put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request. Moved out cluster minimality conjecture to the conclusion. Added a discussion on the overconstrained problem and NP-hardness (3.3).

(4) Recombination: Improved the section introduction. Moved in discussion of previous work on recombination from (1). Made the section that followed in the previous revision into a subsection (4.5).

(5) and (6) Application sections: No changes.

(8) Conclusion: Moved in conjectures about cluster minimality from (3). Added new open problem (OP2).

(App. A) Proofs: Revised DRP proofs for clarity, but there were no fundamental changes.

Journal Change
==============

Originally the paper was submitted for consideration of publication in the special issue of CAGD for GMP2015. However, pending our major revision, this was changed to the regular issue of CAGD. We received confirmation from the GMP co-chairs that this paper would now be handled by the CAGD editors-in-chief. Below, we have copied the email we received from the conference chairs, stating the change.

Upon submission, the exact content of this section will also be emailed to editors-in-chief, along with our submission number: XX.

If there are additional questions in regard to this issue, please send an email.

Email
-----
GMP2015 <gmp2015@easychair.org> Wed, Mar 25, 2015 at 7:17 AM
To: Troy Baker <tbaker@cise.ufl.edu>
Dear Troy Baker,

even though your GMP submission will not appear in the conference special
issue, but - after a major revision - in a forthcoming regular issue of
CAGD, you still have the chance to present your work at the GMP 2015
conference, and we would be very glad if you accept this invitation.

More information about the conference can be found here:
http://gmp2015.inf.usi.ch/

Note that you may require a visa to come to Switzerland and that the
application process can take some time. So it is best to start early. More
information about visa can be found here: http://gmp2015.inf.usi.ch/visa

Please inform us about your decision whether or not to come to GMP 2015
and to present your work by sending an e-mail to "gmp2015@usi.ch" before
*April 2, 2015* to help us plan the conference program.

With kind regards

Elaine Cohen & Kai Hormann
[GMP 2015 conference chairs]

### 2015-05-16 doc/GMP2015/revision_2/appendix_proofs.tex
Revision 2. 24 hours to submission.

@@
 
-For (2), observe that underconstrained subgraphs of isostatic graphs must have density less than $k$.
-For (3), observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic.
-Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$.
-Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
+For the next parts, we use the quantity $d(G)=2|V|-|E|$, which we call \dfn{density}.
+For (2), observe that underconstrained subgraphs of isostatic graphs must have density less than 3.
+For (3), observe that, given an isostatic graph, a subgraph with density $3$ must also be isostatic.
+Then, use the fact that, by definition, density of $F_i$ denoted $d(F_i)=3$ and $d(F_j)=3$.
+Then it is straightforward application of the inclusion-exclusion $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
 
@@
 \begin{proof}
-Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
+Assume $C_i\cup C_j \neq C$. This would contradict the proper vertex-maximality of  $C_i,C_j$.
 %
-In the reverse direction, we know $C$ is either some non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
+In the reverse direction, we know $C$ is either a non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
 \end{proof}
@@
 \begin{observation}\label{observation:no_edges_between_diff}
-Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
+Take $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
 \end{observation}
@@
 \begin{proof}
-Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the entire parent $C$.
+Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the parent graph $C$.
 \end{proof}
@@
 
-The alternative phrasing is due to Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}.
+The alternative phrasing is as in Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}.
 
-Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
-Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.
+Take $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
+Furthermore, take the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ that are non-empty.
 
-Then we assume that there is some third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
+Assume that there is a third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
 
-For convenience, it is implied that all graphs are the induced graph on $C$.
+Without loss of generality, all graphs are the induced subgraphs of $C$.
 
@@
 
-    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal.
+    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they are not vertex-maximal.
 These are the graphs $\inducedOnC{R'_i\cup R_j\cup D_{i,j}}$ and $\inducedOnC{R_i\cup R'_j\cup D_{i,j}}$ respectively.
@@
 In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices.
-When finding a DR-plan, we treat each body as trivial, so they become the leaves of our plan.
+When finding a DR-plan, we treat each body as trivial, so they become the leaves of the DR-plan.
 The optimal recombination problem and approach of Section \ref{sec:recomb} are unchanged.
-The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
+The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional restriction that all edges in the 2-tree representation of the bodies must be removed together, not individually.
 \end{proof}
@@
 \begin{proof}
-    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
-Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.
+    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ is then a vertex-maximal proper 1-dof subgraph of $G$.
 
-    To make the graph isostatic, we need only add one body and pin it to 2 other bodies.
-Doing so will cause $G$ to become $(3,3)$-tight.
-We choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
-From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a sub-graph.
-We pin the new body to bodies in two separate children.
+    To make the graph isostatic, we need only add one body and pin it to 2 other bodies. Doing so will cause $G$ to become $(3,3)$-tight.
+
+We adopt the following algorithm.
+Choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
+From Observation \ref{rem:1dofcanon}, we know that the children can only share  a single pin or a sub-graph.
+Pin the new body to bodies in two separate children.
 Doing so will ensure that all children of $b$ will have 1-dof and all ancestors of $b$ (including $b$) will now be isostatic.
@@
     Such a pinning covers all possible ways of adding a new body.
-Assume we add a new body $b$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic.
-Then, there will be some lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$.
+Assume a new body $b$ is added to the input  graph and pin it to $b_i$ and $b_j$ to make it isostatic.
+Then, there is a lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$.
 Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning $b$ to $b_i$ and $b_j$.
 
-    For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$.
+    For each node $b$, assign a size of the $T_b$ denoted $|T_b|$.
 $|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$.
@@
 Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$.
-All other nodes are not isostatic and so have no fanin.
-
+All other nodes are not isostatic and hence do not appear in the isostatic DR-plan.
     % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
@@
 
-    The node we need to pin will always be the deepest nontrivial node of some path in $T$.
-Suppose we chose to pin a node $b$ that has a nontrivial child $v$.
+    The node to be pinned is always the the deepest nontrivial node of some path in $T$.
+Suppose a node $b$ is pinned that has a nontrivial child $v$.
 Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$.
@@
 
-    Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.
+    Running the above algorithm brute force  gives running time quadratic in the number of bodies of the given body-pin system.
 
-    For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
+    For the multi-triangle pin graphs, we can do the same thing, except we need to add a single triangle to one of the nodes to cause it to become isostatic.
     % Don't leave a blank line between last paragraph and end!
@@
 \begin{proof}
-    The only difference from the 1-dof case is that now we need to remove 2-dof from our graph.
-We build a 2-dof DR-plan $T$.
-Like above, we need to add a body and 2 pins to 2 nodes now to get to isostatic.
+    The only difference from the 1-dof case is that now we need to remove
+2-dofs from our graph. Start with a 2-dof DR-plan $T$. Like in the previous proof, we need to add a body and 2 pins to 2 nodes to obtain an isostatic DR-plan.
 
     Suppose we pin 2 distinct nodes $v_i$ and $v_j$.
-Then, there will be some common ancestor $a$ of $v_i$ and $v_j$.
+Then, there must exist a common ancestor $a$ of $v_i$ and $v_j$.
 Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$.
@@
 
-    So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
+    Thus the only choice is to pin a single node twice. Hence, we can run the same algorithm as the 1-dof case and simply pin twice instead of once.
 \end{proof}
@@
 \begin{proof}
-    An isostatic graph has 3 parameters that define it (in terms of coming up with a realization).
+    An isostatic graph has 3 parameters that define its position and orientation.
 These are the Euclidean motions.
-A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has.
-A 2-dof has 5 parameters, and so on.
-These parameters denote the algebraic complexity of giving a realization for the graphs.
+A 1-dof graph has  4 parameters: the 3 Euclidean motions and a dof parameter.
+A 2-dof has 5 parameters.
+The number of parameters roughly correlates with the algebraic complexity of obtaining a realization.
 
-    Thus, say we have a $T$ as described in the proof for Remark \ref{thm:1dofcase}.
-When we pin a node $b$, we will have the same structure as before.
-Suppose we are looking at an isostatic node $v$ after pinning $b$.
-Then, the children of $v$ (except one if $v \neq b$) will be 1-dof.
-The complexity for that node is then just that of solving each of its children.
-In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.
+    Thus, starting with a $T$ as described in the proof for Remark \ref{thm:1dofcase}: when a node $b$ is pinned, the same structure is preserved as before.
+Suppose  $v$ is an isostatic node after pinning $b$.
+Then, the children of $v$ (except one if $v \neq b$) have 1-dof.
+The realization complexity for $v$ is simply that of realizing each of its children.
+In general, the number of parameters for $v$  will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.
 
-    Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$.
-In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children.
-So we will have to try pinning all nodes to see which gives the lowest algebraic complexity.
-This is still quadratic for the 1-dof case.
+    Minimizing the algebraic complexity  requires minimizing the maximum $np(v)$ for any node $v$.
+In this case, it is not possible to always choose to pin a node closest to a leaf in the tree, because it could have high fan-in.
+So we  try
+brute force by pinning all nodes to  pick the one with the lowest algebraic complexity. This algorithm is still quadratic for the 1-dof case.
 
-    For the 2-dof case, there are more cases to consider.
+    For the 2-dof situation, there are more cases to consider.
 If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$.
 If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof.
-Note that solving $v'$ will result in solving $v$.
-Then, we need to consider nodes above and including $v'$ in our complexity.
-$np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
-The only remaining case will be if we pin two nodes that are not descendant/ancestor.
-the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.
+Note that solving or realizing $v'$ will also realize $v$.
+Next, we need to consider nodes above and including $v'$ in our complexity: $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
+
+The only remaining case is pinning two nodes that are incomparable, i.e., do not have a descendant/ancestor relationship.
+The only change from the previous case is that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.
 For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.
 
-    Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin.
-However, we cannot also assume pinning one node twice will give us the best algebraic complexity.
-Hence, we will need to check each pair of node to pin.
+    Like the 1-dof case, we again cannot simply choose the nodes deepest in the tree to pin.
+However, neither can we assume pinning one node twice will give us the best algebraic complexity.
+Hence, we will need to check each pair of nodes to pin.
 This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
@@
 \subsection{Proofs from Section \ref{sec:pinnedline}}
-\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
+\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- for 2-dimensional pinned line incidence graphs}
 \label{sec:appendix_pinned}
@@
 \begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
-We use the same notations as in the original proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, in the last section.
-For convenience, it is implied that all graphs are the induced graph on $C$.
+We use the same notation as in the original proof of  Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, given above.
+Without loss of generality, all graphs are the induced graphs on $C$.
 
@@
 
-Then we analyze all the possible cases for $C_k$.
+We analyze all the possible cases for $C_k$.
 \begin{itemize}
@@
 
-All remaining cases are similar to the original proof.
+All remaining cases are similar to the original proof for 2D bar-joint graphs.
 \end{proof}

### 2015-05-16 doc/GMP2015/revision_2/s01_introduction.tex
Revision 2. 24 hours to submission.

@@
 \newcommand{\seedefsd}{see \ref{sec:appendix:defs} for definitions}
-\newcommand{\seedefsprelim}{(formally defined in Section \ref{sec:prelim})}
+\newcommand{\seedefsprelim}{(formally defined in Section \ref{sec:prelim:defs})}
 
@@
 
-Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields.
-Some systems have an underlying sparsity condition that characterize rigidity, allowing for matroidal techniques of rigidity detection and decomposition.
-However, the design and analysis of these systems is complicated by the NP-hardness barrier of creating optimal decomposition-recombination (DR-) plans.
-While other papers have focused on a few desirable criteria and made use of hueristics, by focusing on the class of independent (isostatic or underconstrained) systems we have found an efficient way to construct the optimal DR-plan.
-Additionally, in real world applications, many of these systems exhibit degrees of self-similarity and quasi-uniformity.
-Our formulation of the optimal DR-plan can leverage these properties, further reducing the complexity of the DR-plan.
-Furthermore, we address the fact that indecomposable subsystems in a system can be arbitrarily large, representing a bottle-neck in the computational efficiency in the analysis of these systems.
-We also address this problem with an efficient technique of optimal modification for further decomposition.
+Geometric constraint systems have well-established, mature applications in mechanical engineering and  robotics, and they continue to find emerging applications in diverse fields from machine learning to molecular modeling.
+Solving or realizing geometric constraint systems requires finding real solutions to a large multivariate polynomial system (of equalities and inequalities representing the constraints); this requires double exponential time in the number of variables, even if the type or orientation of the solution is specified.
+% Thus, recursive decomposition into locally rigid subsystems (which have finitely many solutions) is crucial for realizing a geometric constraint system by the reverse process of recombining the subsystem solutions.
+Thus, to realize a geometric constraint system, it is crucial to perform recursive decomposition into locally rigid subsystems (which have finitely many solutions), and then apply the reverse process of recombining the subsystem solutions.
+With the use of \dfn{decomposition-recombination (DR-) planning}, the complexity is dominated by the size of the largest subsystem that is solved, or recombined, from the solutions of its child subsystems, i.e.\ the maximum fan-in occurring in a DR-plan.
+In addition, navigating and analyzing the solution spaces, as well as designing constraint systems with desired solution spaces leads to the \dfn{optimal decomposition-recombination (DR-) planning} problem~\cite{sitharam2005combinatorial, hoffman2001decompositionI, hoffman2001decompositionII}.
+\todo{Feels very clunky.}
+
+For a broad class of geometric constraint systems, local rigidity is characterized generically as a sparsity and tightness condition of the underlying constraint (hyper)graph~\cite{laman1970graphs,streinu2009sparse,tay1976rigidity,white1987algebraic}.
+This allows the generic DR-planning problem to be stated and treated as a combinatorial or (hyper)graph problem as we do in this paper.
+
+We additionally conjecture that versions of \frontier, a previously given, bottom-up, polynomial time method for obtaining DR-plans with various desired properties~\cite{hoffman2001decompositionI, hoffman2001decompositionII,lomonosov2004graph,sitharam2005combinatorial}, in fact generate an optimal DR-plan for non-overconstrained systems.
+
+Na{\"i}vely, the optimal DR-plan is used as follows.
+Each decomposed subsystem---a node of the DR-plan---is treated and solved as a polynomial system of constraints between its child subsystems.
+However, even in an optimal DR-plan, there can be arbitrarily many children at a node.
+In other words, even in the recursive decomposition given by an optimal DR-plan, the  size of the maximal indecomposable subsystem could be arbitrarily large.
+It represents a bottleneck that dictates the complexity of solving or realizing the constraint system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
+We address this problem using the recently developed concept of \dfn{convex Cayley configuration spaces} \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII,sitharam2014beast,sitharam2013caymos,wang2014cayley}.
+This allows for even greater reduction of the complexity by realizing large, indecomposable systems in a manner that avoids working with large systems of equations.
+Specifically, we give an efficient technique for \dfn{optimally modifying} large indecomposable subsystems in a manner that reduces their complexity while preserving desired solutions; the modification ensures a convex Cayley configuration space, and the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
+This optimal modification  problem is a generalization of the previously studied problem of optimal completion of underconstrained systems \cite{sitharam2005combinatorial,joan-arinyo2003transforming}.
+
+DR-plans are especially useful for constraint systems that exhibit some level of \dfn{self-similarity} and \dfn{quasi-uniformity}, in addition to isostaticity.
+These properties can be leveraged to further reduce the complexity of both optimal DR-plan construction and recombination.
+We consider 3 different types of constraint systems---which we collectively call \dfn{qusecs}---that are used to model, design, and analyze quasi-uniform or self-similar materials.
+In the remainder of this section, we motivate the materials application and give the contributions and organization of the paper.
 
@@
 The main contributions are a method to generate an optimal decomposition-recombination (DR-) plan
-
 %  leveraging the quasi-uniformity and self-similarity present in many real world applications,
-
-\subsection{Recursive Decomposition Problem}
-The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph.
-This characterization has broad applicability in practice, and we provide several concrete examples.
-
 % Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
@@
 
-Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads.
-When realizing constraint systems, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints); this requires double exponential time in the number of variable, even if orientation type is specified.
 % Without decomposition to plan the recombination (DR-planning), this requires double exponential time in the number of variables (even if orientation type is specified).
-With the use of decomposition-recomination (DR-) planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan.
-Thus, finding an optimal DR-plan is important for realization.
 
@@
 
-Furthermore, the DR-plan can be used to obtain recursive block decomposition of the rigidity matrix, stress space, and external stresses.
-Each block is the rigidity matrix of an isostatic subgraph.
-The blocks intersect on isostatic or trivial subgraphs.
-This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$.
-Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan.
-When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives.
-When the realization of the system is given, this can be used to analyze the distribution of a given external load.
+%highly general geometric constraint systems with the  This characterization has broad applicability in practice, and we provide several concrete examples.
 
 % In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which,
-An optimal DR-plan decreases the computation involved in such tasks; however, in general, finding an optimal DR-plan is NP-hard.
+%An optimal DR-plan decreases the computation involved in such tasks; however, in general, finding an optimal DR-plan is NP-hard.
 Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan.
-By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan.
-Then we show that all canonical DR-plans are optimal for systems that are not overconstrained.
-This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties.
-In fact, we conjecture that one of the earlier algorithms also produces an optimal DR-plan and that every optimal DR-plan is in some formal sense canonical.
-
 % Knowing an optimal DR-plan is crucial when working with isostatic geometric constraint systems with an underlying sparsity condition \nameit.
@@
 
-\subsection{Recombination Problem}
-The second problem we consider is the complexity of recombination.
-Na{\"i}vely, recombination can be done by finding the solutions to the polynomial system of equations describing the constraints between the children; however, even in an optimal DR-plan, there can be arbitrarily many children at a node.
-
 % Instead, we propose the modification of the subsystem to find a convex configuration space that can be efficiently searched for a solution to the original system.
-
-Instead, we make use of the notion of Cayley complexity to quantify the difficulty of each step of recombination.
-Once a subsystem of a node is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
-This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.
-
+%Instead, we make use of the notion of Cayley complexity to quantify the difficulty of each step of recombination.
 % We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination.
@@
 \subsection{Introducing Qusecs}
-When working on these problems, there is a large, natural class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint systems''.
+A large class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint system'', (a) can be treated combinatorially as described above and (b) occur as independent (isostatic or underconstrained) systems in materials applications.
+We discuss these next.
 %
-Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers.
+Some natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers.
 As illustrated by the examples below, the structure within each layer is often: self-similar\footnote{In this manuscript we only study finite 2D structures.
 \dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.}~\cite{2012arXiv1204.6389G}, spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and composed of a few repeated motifs appearing in disordered arrangements.
-Note that a layer is not necessarily planar, it can consist of multiple, inter-constraining planar (genus 0) monolayers.
-Furthermore, a layer is often  either \vemph{isostatic} or \vemph{underconstrained} (\vemph{not self-stressed}, \seedefsd).
-A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
+Note that a 2D layer is not necessarily planar (genus 0), it can consist of multiple, inter-constraining planar monolayers.
+Furthermore, a layer is often  either \dfn{isostatic} or \dfn{underconstrained} (not \dfn{self-stressed}/\dfn{overconstrained}, \seedefsd).
+\todo{consider...} These properties, as well as quasi-uniformity, aperiodicity, self-similarity, and layered structure, are natural consequences of evolutionary pressures or design objectives such as stability, minimizing mass, optimally distributing external stresses, and participating in the assembly of diverse and multifunctional, larger structures.
 %
-The importance of an optimal DR-plan is particularly evident when a system is a qusecs.
-The quasi-uniform or self-similar properties mean that solutions for one subsystem can solve other subsystems, thus causing further reduction in the complexity of the plan.
+
+The importance of an optimal DR-plan is particularly evident for a qusecs.
+The quasi-uniform or self-similar properties mean that the decomposition and solution for one subsystem can be used as the decomposition and solution for other subsystems, thus causing further reduction in the complexity of both DR-planning and recombination.
 This is shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.
@@
 \begin{enumerate}
-    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.
+    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}), e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.
 
-    \item \label{materialexample2} Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}).
+    \item \label{materialexample2} Cross-sections of organic tissue with hierarchical structure, e.g., compact bone and tendon (Figure \ref{fig:material_examples:tendon}).
 
-    \item \label{materialexample3} Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix.
+    \item \label{materialexample3} Crosslinked cellulose or collagen microfibril monolayers, e.g., in cell-walls~\cite{wikimediacommons2010afm, wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix.
 See Section \ref{sec:pinnedline}.
 
-    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski} (Figure \ref{fig:material_examples:sierpinski}), bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
+    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1, Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski} (Figure \ref{fig:material_examples:sierpinski}), bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.
 
-    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and materials that behave like assemblies of 2D particles under non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}.
+    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and materials that behave like assemblies of 2D particles under non-overlap constraints, i.e.\ like jammed disks on the plane~\cite{jammed_disks}.
 See Section \ref{sec:bodypin}.
@@
 
-In Section \ref{sec:prelim}, we provide basic definitions in rigidity theory, including what it means for a DR-plan to be complete or optimal. We also elaborate on the new notion of qusecs.
-This section is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.
+In Section \ref{sec:prelim}, we provide basic definitions in combinatorial rigidity theory,  and formalize the new notion of qusecs~\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.
+In addition, we define DR-plans and what it means for a DR-plan to be complete or optimal.
+We survey previous work on DR-planning algorithms, discussing other desirable criteria of DR-plans and their relation to the NP-hard optimality property of DR-plans.
+%This section is relevant to Examples \ref{materialexample1} and \ref{materialexample2},  model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.
 \todo{We don't model them as bar-joint...
@@
 
-In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal.
-In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
-Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
+In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained qusecs are optimal.
+In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained.
+Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied {\em cluster minimality}  \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
 Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
-
+While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
 % that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
@@
 
-While both of these sections focus on bar-joint graphs, the theory is easily extended to other qusecs.
-The following sections focus on explicit applications by modeling rigidity in certain materials as qusecs.
+ % The next sections focus on explicit applications by modeling certain materials as qusecs.
 
-In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
-For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems.
-For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).
+In Section~\ref{sec:bodypin} and~\ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
+We explicitly model these materials as qusecs.
+For Examples~\ref{materialexample4} and~\ref{materialexample5}, we discuss boundary conditions for achieving various desired properties of body-hyperpin systems.
+For Example~\ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems~\cite{sitharam2014incidence}.
+
+The last Section~\ref{sec:conclusion} concludes the paper, and Section \ref{sec:open} lists open problems and conjectures.
+In particular, we conjecture that the methods of Section~\ref{sec:DRP} extend in fact to a large class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.
+
+Throughout this paper,  an asterisk after a formal statement indicates that its proof appears in~\ref{sec:appendix:proofs}.
 

### 2015-05-16 doc/GMP2015/revision_2/s02_preliminaries.tex
Revision 2. 24 hours to submission.

@@
-\section{Preliminaries}
+\section{Preliminaries and Background}
 \label{sec:prelim}
@@
 
-Throughout this paper, a \textbf{\textsf{Theorem}}, \textbf{\textsf{Lemma}}, \textbf{\textsf{Observation}}, or \textbf{\textsf{Remark}} with an asterisk after it means that the proof appears in~\ref{sec:appendix:proofs}.
+We first give basic definitions and concepts in combinatorial rigidity, leading to a definition of a DR-plan, its properties, and how they relate.
+The section ends with a discussion of previous work on DR-plans.
 
-\subsection{Basic Definitions and Theory}
+\subsection{Geometric Constraint systems and Combinatorial Rigidity}
 \label{sec:prelim:defs}
@@
 % \footnote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}
-i.e.
-$\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
+i.e.\ $\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
 
-In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
-In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).
+In all types of geometric constraint systems we consider in this paper, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions).
 
-Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values).
+Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are defined to be properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values).
 Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates).
@@
 The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp.
-$x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.
+$x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), i.e.\ 4 non-zero entries per row.
 
@@
 
-One important property of generic constraint system or (hyper)graph\footnote{We refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably.}
+One important property of a generic constraint system or (hyper)graph\footnote{We refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably}
 %that depends on the appropriate rigidity matrix of indeterminates
@@
 % (\note footnote now...),
-i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
-The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
+i.e.\ the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.
+%(if the constraint system has equalities and exact values for the parameters $\delta$).
+The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if and only if  it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for 2D bar-joint systems.
 %This condition is called generic {\em infinitesimal rigidity}.
@@
 
-Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of rather than exact values.
+Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals rather than exact values.
 In this case, the definition of rigidity is approximate; the solutions are isolated, small, full-dimensional connected components.
 
-Other generic constraint system or (hyper)graph properties are the following:
+Other generic constraint system or (hyper)graph properties are mentioned here.
 % (\note we refer to these as properties of the constraint
@@
 % interchangeably).
-A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
+A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e.\ the determinant of some square submatrix is not identically zero).
 %It is {\em rigid} if the number of independent rows of the rigidity matrix is
@@
 %which is 3 for distance constraint systems.
-It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
+It is \dfn{isostatic} (\dfn{minimally rigid}, \dfn{wellconstrained}) if it is both rigid and independent.
 %if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
@@
 
-Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
+Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the independence of corresponding rows in the rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
 %The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
@@
 %if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
-There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs.
-For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices.
+There are various results on combinatorial characterization of independence, rigidity, and rigidity matroids for different types of (hyper)graphs.
+For bar-joint rigidity matroids, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices.
 The result by Lovasz and Yemini \cite{lovasz1982generic} shows that all 6-vertex-connected graphs are rigid in the plane.
-For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees.
+For bar-body rigidity matroids, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees.
 White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition.
-Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$.
+Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if it is $(k,l)$-sparse and $|E| = k|V| - l$.
 In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.
+
+In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:prelim:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
 
@@
 %TODO
-Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}.
+%Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}.
 It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions).
@@
 %
-A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}.
+%A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}.
 It has one entry per edge and represents a \dfn{self stress} of the system.
@@
 %
-An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
+%An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.
 
-\subsection{Qusecs}
-
-In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:prelim:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).
+%\subsection{Qusecs}
 
@@
 
-In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
-Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in their corresponding sections on applications.
+We note that  the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs.
+Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given along with their materials applications in Sections \ref{sec:bodypin} and \ref{sec:pinnedline}.
 
@@
 
-\begin{figure*}\centering%
+%\begin{figure*}\centering%
 %
-  \begin{subfigure}{0.4\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/triangle}
-    \caption{}\label{fig:demo_graph_tri:graph}
-  \end{subfigure}%
+%\begin{subfigure}{0.4\linewidth}\centering
+%\includegraphics[height=\myMinHeight]{img/svg/triangle}
+%    \caption{}\label{fig:demo_graph_tri:graph}
+% \end{subfigure}%
 %
-  \hfill
-  \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
-    \caption{}\label{fig:demo_graph_tri:candrp}
-  \end{subfigure}%
+%  \hfill
+% \begin{subfigure}{0.3\linewidth}\centering
+%   \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
+%    \caption{}\label{fig:demo_graph_tri:candrp}
+%  \end{subfigure}%
 %
-  \hfill
-  \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
-    \caption{}\label{fig:demo_graph_tri:clustmindrp}
-  \end{subfigure}%
+%  \hfill
+%  \begin{subfigure}{0.3\linewidth}\centering
+%   \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
+%   \caption{}\label{fig:demo_graph_tri:clustmindrp}
+%  \end{subfigure}%
 %
-  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially.
+%  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially.
 (\ref{fig:demo_graph_tri:candrp}) The canonical DR-plan of $G_{tri}$.
-(\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality.
+(\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality  (See Section \ref{sec:prev}).
 Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge.
@@
 With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
-  \label{fig:demo_graph:clustmindrp}
-  \label{fig:demo_graph_tri}
-\end{figure*}%
+% \label{fig:demo_graph:clustmindrp}
+%  \label{fig:demo_graph_tri}
+%\end{figure*}%
 
@@
         % subgraphs of $C$.
-        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
-        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
-        \item A root node is a maximal rigid subgraph of $G$.
+        \item The children $C_1,\ldots,C_N$ of a node $C$ satisfy $\bigcup_{i=1}^N{C_i}=C$.
+        \item A leaf node is a single edge. A \dfn{trivial} graph is empty or a single vertex. Note that a trivial graph is not isostatic.
+        \item A root node is a vertex-maximal rigid subgraph of $G$.
     \end{enumerate}
@@
 %
-    A DR-plan is \dfn{complete} if it satisfies an additional rule: for a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$.
-This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well.
+    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$.
+This makes Property 2 implicit.
+  %and in fact, $C$ contains all the edges in the union of its children as well.
   We denote a complete DR-plan of $G$ as $\comdrp{G}$.
@@
 We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
-
-    See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.
-
 %
@@
 \begin{remark}
-    For a given graph, there could be exponentially many  DR-plans and even optimal DR-plans in the size of the graph.
-A complete DR-plan is unique but may not be (is usually not) optimal.
+More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
+For a given graph, there could be exponentially many DR-plans---and even optimal DR-plans---in the size of the graph.
+A complete DR-plan is unique but may not be (and is usually not) optimal.
 DR-plans of self-similar graphs are self-similar.
-More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
 \end{remark}
+
+See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
+
+\subsection{Previous Work on DR-plans}
+\label{sec:prev}
+% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+The limitations of these techniques directly motivate the contributions of this paper.}
+We now briefly survey existing techniques for detecting rigidity and creating DR-plans of 2D constraint systems.
+The limitations of these techniques directly motivate the contributions of the next  section.
+
+\subsubsection{Finding (vertex)-maximal generically rigid subsystems}
+Fast, graph-based algorithms exist (pebble-game~\cite{Jacobs:1997:PG,hoffmann1997solvablesubsets,jermann2006decomposition,Lee:2007:PGA}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs.
+When the input itself is rigid, these algorithms do nothing, i.e.\ compute the identity function.
+
+However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.
+
+\subsubsection{Optimal Recursive Decomposition (DR-planning)}
+Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{lomonosov2004graph,sitharam2005combinatorial,jermann2006decomposition} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim.
+For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm \todo{rephrase?
+maybe not} without involving geometric information in the constraint system.
+Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in~\cite{hoffman2001decompositionI}.
+%An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.
+Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.
+
+However, for overconstrained 2D qusecs, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph, sitharam2005combinatorial}.
+The NP-hardness of the optimal DR-planning problem for 2D bar-joint graphs is partly the consequence of possibly exponential number of DR-plans.
+On the other hand, although the complete DR-plan is unique it could have large average fan-in and exponentially many nodes making it far from optimal.
+
+\subsubsection{DR-plans for special classes and with other criteria}
+For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems~\cite{owen1991algebraic,fudos1997graph,joan-arinyo2004revisiting} common in computer aided mechanical design (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.
+This satisfies the Church-Rosser property, leading to highly efficient DR-planning algorithms.
+For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to have a minimal set of at least 2 rigid proper subsystems as children (i.e.\ the union of no proper subset of size at least 2 child subsystems forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system.
+\seedefsc.
+
+While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion~\cite{lomonosov2004graph}, no such algorithm is known for the latter criterion.
+%Moreover, cluster minimality does not imply optimality as shown in Figure \ref{fig:demo_graph_tri}
 

### 2015-05-16 doc/GMP2015/revision_2/s03_drp.tex
Revision 2. 24 hours to submission.

--- 2015-03-25 doc/GMP2015/revision_2/s03_drp.tex
@@
-\section{The canonical DR-plan}
+\section{Main Result: canonical DR-plan, Optimality, and Algorithm}
 \label{sec:DRP}
@@
 
-The NP-hardness of the DR-planning problem for 2D bar-joint graphs is partly the consequence of possibly exponential number of DR-plans.
-On the other hand, although the complete DR-plan is unique it could have large average fan-in and exponentially many nodes making it far from optimal.
-
-In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.
-
-\sidenote{In this section, $G$ is always assumed to be isostatic.}
+\subsection{Canonical DR-plan}
+In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
+Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph.
+The definition is as follows:
 
 \begin{definition}\label{def:canonical_drplan}
-    The \dfn{canonical DR-plan} satisfies the following three properties:
-    (1) it is a DR-plan;
+    The \dfn{canonical DR-plan} \addin{of a graph $G$} satisfies the following three properties:
+    (1) it is a DR-plan \addin{of $G$};
     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise take as children exactly two that intersect non-trivially.
+    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
 \end{definition}
 
-This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
+In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+% \sidenote{In this section, $G$ is always assumed to be isostatic.}
+
+% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
+
+Definition~\ref{def:canonical_drplan} gives the canonical DR-plan a surprisingly strong Church-Rosser property, which is made explicit in Theorem~\ref{theorem:main}, the main result of this section.
 
@@
 \label{theorem:main}
-    A canonical DR-plan exists and any canonical DR-plan is optimal.
+    A canonical DR-plan exists \addin{for a graph $G$} and any canonical DR-plan is optimal \addin{if $G$ is independent.}
 \end{theorem}
@@
 
-Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children nodes according to the following rules:
+Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$  retain children nodes according to the following rules:
 \begin{enumerate}
-   \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
+   \item If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
    \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
@@
 
-This directly satisfies properties 2 and 3 of a canonical DR-plan (see definition \ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs.
-However, we still have to show property 1 (that this makes a DR-plan).
-For rule (1), because we are beginning with a complete DR-plan, if we keep all the children it is still a DR-plan.
-For rule (2), we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we have found a larger rigid vertex-maximal proper subgraph.
+This directly satisfies Properties (2) and (3) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs,  which we shorten to {\em clusters}.
+To show Property (1) holds (that this constitutes a DR-plan):
+for Case 1 above,  since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Case 2 above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.
 
-If we begin with an isostatic graph, rigid can be replaced with isostatic throughout the construction and remain valid. The rigid proper subgraphs of an isostatic graph must be isostatic themselves.
+Note that if we begin with an isostatic graph, ``rigid'' can be replaced with ``isostatic'' throughout the construction and preserve the above properties.
+The rigid proper subgraphs of an isostatic graph must be isostatic themselves.
 
-Now we must show that a canonical DR-plan is optimal.
-The proof is by induction on the level of the plan; we show that following these rules always guarantees minimum fan-in at this level and does not increase fan-in on following levels.
+Next we show that a canonical DR-plan is optimal.
+
+First, note  that any  DR-plan \addin{$R$} without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan \addin{$R$})
+is the subgraph of some \addin{cluster} of $C$.
+Thus without loss of generality, we can assume that \cutout{the} \addin{an} optimal DR-plan satisfies Property (2) of a canonical DR-plan.
+
+\addin{The proof of optimality of a canonical DR-plan is by induction on its height.
+The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges.
+The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.}
+\addin{For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.}
+Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
+Thus, from the induction hypothesis, we know that the $R(C_i)$ is optimal for $C_i$.
+
+Thus it is sufficient to demonstrate a set of nodes $S$  that must be present in any  DR-plan $R$ for $C$ that satisfies Property (2), including a known optimal one; and furthermore, for any such DR-plan $R$,  either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $R$ has the minimum possible fan-in of 2.
+
+We show the two claims below.
+The first claim is that \addin{for a node $C$ whose clusters have trivial pairwise intersections}, any DR-plan \cutout{$R$} \addin{of $C$} that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
+\addin{Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.}
+The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, \addin{every canonical DR-plan of $C$ that uses any possible choice} of two such subgraphs of $C$ as children \cutout{ (minimum possible fan-in at $C$)} results in \addin{a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the}     {\em same maximal antichain $S$ of descendants $D$ of $C$}.
+\addin{The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment.} I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; \addin{this follows from properties of maximal antichains} that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$.
+Thus any  DR-plan  \cutout{$R$} that satisfies Property (2) \addin{and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children} must \addin{also} contain every element of $S$.
+The two claims complete the proof that every canonical
+DR-plan is optimal.
 
@@
 \medskip\noindent
-\vemph{Rule 1:}
+\vemph{Claim 1:}
 % If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
 %
-First, we claim that if the intersection is trivial all $C_1,\ldots,C_N$ must be children of $C$ in an optimal DR-plan.
-We do this by showing that no subset of the children can union to form $C$, thereby requiring all of them be included; because this is the only choice, it is the minimum fan-in any DR-plan could have for this node.
+A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial, must be children of $C$ in an optimal DR-plan.
+
+We prove this claim  by showing that the union of no subset of the children can be $C$, thereby requiring all of them to be included as children.
+
 % must be optimal for this node.
 
-Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic.
-If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with.
+We prove by contradiction.
+Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic.
+If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$.
 So, it must be that $U=C$.
 \usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs.
-By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
-Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$.
-Thus, the assumption is wrong and $U$ cannot be isostatic.
-As $C$ is isostatic, this means no proper subset of $C_1,\ldots,C_N$ can union to form $C$.
+However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices.
+By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
+Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible.
+As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.
 
-Furthermore, since a canonical DR-plan uses proper rigid \vemph{vertex-maximal} subgraphs as children, and because their pairwise intersection is trivial, it follows that any such node has at most as many children as a DR-plan without this restriction, because the union of the children have to contain all edges of the parent.
-Therefore, this is the optimal choice in this case.
+Furthermore, since a canonical DR-plan has nodes with proper rigid \vemph{vertex-maximal} subgraphs as children, if, as in this case, their pairwise intersection is trivial, it follows that any node has at most as many children as a DR-plan without this restriction, because the union of the children must contain all edges of the parent.
+Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.
 
 \medskip\noindent
-\vemph{Rule 2:} Second, we claim that if the intersection is isostatic, the choice of any two children will result in an optimal DR-plan.
+\vemph{Claim 2:}
+If some pair in the set of child clusters $C_1,\ldots,C_N$ of $C$ has an isostatic (nontrivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of $C$.
 
-If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
-This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself.
-Thus, any two children can be chosen to make a canonical DR-plan.
+\addin{To prove Claim 2}, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+This means that, by Lemma \ref{lemma:combined_lemma}, \addin{Point} \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself.
+Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
 % are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
@@
 \renewcommand{\induceonc}[1]{#1}
-\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in S_N\setminus\{#1\}}{R_k}}}
+\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in [N]\setminus\{#1\}}{R_k}}}
 
-However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants.
-Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$.
-Suppose we select $i$ and $j$, where $i\neq j$, as the children.
-For convenience, we will assume all subgraphs are induced subgraphs on $C$.
-We know that $C=\induceonc{I\cup\bigcup_{k\in S_N}{R_k}}$ and $C_i=\iunion{i}$.
-The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of which intersect on isostatic subgraphs.
+However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants leading up to a common maximal antichain $S$ of subgraphs.
+
+To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$.
+Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children.
+For convenience, we will assume all subgraphs are induced subgraphs of $C$.
+We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
+The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs.
 So any two of these are viable children for $C_i$.
@@
 % for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
-This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in S_N$, with every $k$ appearing at least once.
-Thus, regardless of the choice of $C_i,C_j$, then their children, etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
-Therefore, the choice is optimal.
+This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
+At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
+Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for  $N$ levels,  and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
 %
@@
 
-This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.
+This proof of this theorem relies on the following crucial observation and lemma.
+These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.
 
 \begin{observation*}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are isostatic subgraphs of the same
-isostatic graph $F$, \
-then the following hold:
+If $F_i$ and $F_j$ are non-empty isostatic graphs then the following hold: \\
 (1) $F_i\cup F_j$ is not trivial;
@@
 
-Then the following key properties hold at the nodes of a canonical DR-plan.
+The following key properties hold at the nodes of a canonical DR-plan.
 
 \begin{lemma*}\label{lemma:combined_lemma}
-Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
+Let $C$ be an isostatic node of a canonical DR-plan, with distinct children $C_1,C_2,\ldots, C_m$. Assume $i\ne j$.
 Then
@@
 \begin{remark}
-The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{Cluster Minimality}, defined in \cite{hoffman2001decompositionI} holding for these canonical-optimal DR-plans.
+The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{cluster minimality} (defined in \cite{hoffman2001decompositionI} and in Section \ref{sec:prev}) holding for \cutout{these} canonical-optimal DR-plans.
 \end{remark}
@@
 \textsl{[DR-Plan for self-similar structure]}
-This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
+This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
 The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$.
@@
 \begin{theorem}\label{theorem:algo_complexity}
-    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
+    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for \addin{an input graph}.
 \end{theorem}
@@
 \begin{proof}
-The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
-Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of component-candidates.
-Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list.
-If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list.
-Continue with the next candidate.
-When you exhaust the candidates, move to the next step.
+We first describe the algorithm.
+The first step of the algorithm, which we call $CanDRP(G)$, is finding the isostatic vertex-maximal proper subgraphs (clusters) of the input isostatic graph $G$.
+One of many ways to do this is  by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
+The output of this will be a list of cluster-candidates.
+The next step is to isolate the set of true-clusters by running the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, to find the minimal subgraph $D$ that contains the candidate and $e$.
+If $D=G$, the candidate is moved (before adding $e$) to the true-cluster list and removed from the candidate-cluster list.
+If $D\subsetneq G$, all candidates that are subgraphs of $D$ are removed from the candidate-cluster list and $D$ is added back to the list.
+The next candidate is considered until all candidates are exhausted.
 
-Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.
+The next step is to check the intersection of any two of the true-components.
+If the intersection is trivial, the entire list becomes the children, and the algorithm is recursively applied to each child.
+If the intersection is non-trivial (necessarily isostatic), the  decomposition is computed down to $I=\bigcap_{k=1}^{N}{C_k}$.
+(This is discussed in detail in the proof of Theorem \ref{theorem:main}).
+We construct a DR-plan whose node set is that of a canonical/optimal DR-plan.
+Begin with the parent $C$.
+Its children are $S_1$ and $CanDRP(T_1)$, where $S_1=C\setminus R_1$ and $T_1$ is $R_1$, together with all incident edges.
+Observe that these edges are incident on vertices from $I$ and that $T_1$ is underconstrained so its DR-plan is a forest where each root becomes a child of $C$.
+The children of $S_1$ are $S_2$ and $CanDRP(T_2)$, where $S_2=C\setminus (R_1\cup R_2)$ and $T_2$ is $R_2$ together with all incident edges.
+The process iterates.
+At $N$ levels down, the children are  $S_N=I$ and $CanDRP(T_N)$, where $T_N$ is $R_N$ together with all incident edges.
+Apply the algorithm recursively to both children.
+Thus we obtain an efficient method for the case of DR-plan nodes whose rigid proper vertex-maximal subgraphs have non-trivial intersections, avoiding recomputation of the isostatic components of $C$.
+See Figure~\ref{fig:demo_graph:candrpseq} for an example.
 
-If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$.
-The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here.
-We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan.
-Begin with the parent $C$.
-Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$.
-Recursively apply the algorithm to the child containing $R_1$.
-The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$.
-And so on.
-At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$.
-Apply the algorithm recursively to both children.
-See Figure~\ref{fig:demo_graph:candrpseq} for an example.
+To summarize the above step of the algorithm:
+if the intersection is trivial, recursively apply $CanDRP$ to the entire list of true-components and set the resulting trees as the children.
+Else, the intersection is non-trivial, and the children of the node are $CanDRP(C_1)$ and $CanDRP(T)$, where $C_1$ is the first true-component of the parent $C$ (chosen arbitrarily, could be any child) and $T$ is the underconstrained graph formed by $C\setminus C_1$ together with all incident edges  and the associated nodes in $C_1$.
+
 % Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
 
-The reason for this construction is that the result is a tree with no duplicate nodes.
-The leaves of this tree will exactly be the edges of this graph.
-Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.
-
-Therefore, the complexity of the algorithm as a whole is $O(|V|^3)$.
+The DR-plan  output by this algorithm is a tree where each node is a distinct subgraph.
+The leaves of this tree are the edges of the starting graph $G$.
+Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$ and the complexity of the algorithm is $O(|V|^3)$.
 %
@@
 The output of this will be a list of component-candidates.
-Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
+Then run the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
 If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
@@
 % The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the wellconstrained vertex-maximal proper subgraphs of the input wellconstrained graph.
-There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
+There are two approaches: (1) the bottom-up approach, using the \frontier\ algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
 After the subgraphs have been found, it is a straightforward application of the theory.
@@
 
-\begin{remark}
-    While the canonical DR-plan is optimal only if the input graph is independent, if there are only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan.
+\addin{
+
+\subsection{Overconstrained Graphs and NP-hardness of Optimal DR-planning}
+\label{sec:drp:overconstrained}
+
+For overconstrained (not independent) graphs, a canonical DR-plan is still well-defined.
+However, it may be far from optimal. The proofs of Theorem \ref{theorem:main}, Observation \ref{lemma:union_intersection}, and Lemma \ref{lemma:combined_lemma} all fail for overconstrained graphs.
+It is important to note that, regardless whether the graph is overconstrained, if every node in a canonical DR-plan $R$ has clusters whose pairwise intersection is trivial, then the DR-plan is the unique one satisfying Property (2), and since we know that there is an optimal DR-plan that satisfies Property (2), $R$ is in fact optimal.
+The problem arises when some node in a DR-plan has clusters whose pairwise intersection is non-trivial.
+In this case, an arbitrary choice of a pair of clusters as children of the node in a canonical DR-plan may not result in an optimal DR-plan if the graph is overconstrained.
+\todo{rephrase?
+Theorem 4 shows this.
+If it's independent it \vemph{will} be optimal (the church rosser property), no longer holds if it's overconstrained.}
+A good source of examples of overconstrained graphs with canonical DR-plans that are not optimal are graphs whose cluster-minimal DR-plans that are not optimal.
+The example shown in Figure \ref{fig:canonicalnotoptimal} is a canonical, cluster-minimal DR-plan that is not optimal; an optimal DR-plan is also shown in the figure.
+
+As mentioned earlier, the Modified \frontier\ algorithm
+version given in \cite{lomonosov2004graph} runs in polynomial time and finds a cluster-minimal DR-plan for any graph.
+Similarly, the algorithm given above finds a canonical DR-plan also for any input graph.
+However neither of these DR-plans may be optimal for overconstrained graphs as shown in Figure \ref{fig:canonicalnotoptimal}.
+
+While the canonical DR-plan is optimal only if the input graph is independent, when there are only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan using a straightforward modification of the above algorithm.
 However, the time complexity is exponential in $k$.
-\end{remark}
 
-\begin{conjecture}
-\label{conj:mfaisoptimal}
-    The Modified Frontier Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
-\end{conjecture}
-
-The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA is a bottom-up algorithm, involving complex datastructures.
-However, such a proof, even if it exists, would not be possible without the new notion of a canonical DR-plan at hand.
-The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the sequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.
-The strength of MFA compared to other DR-planning algorithms is illustrated in the paper~\cite{lomonosov2004graph}.
-While MFA also offers complexity $O(n^3)$, it does not guarantee optimality.
+This exponential growth of time complexity for overconstrained graphs is in fact captured in the proof of NP-hardness of optimal DR-planning in \cite{sitharam2005combinatorial, lomonosov2004graph}.
+}

### 2015-05-16 doc/GMP2015/revision_2/s04_recombination.tex
Revision 2. 24 hours to submission.

@@
 
-The main barrier in recombination when given an optimal DR-plan (of smallest possible size or max fan-in),  is that the number of children (resp.
-number of roots)---and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved---could be arbitrarily large as a function of the size of the input system.
+The main barrier in recombination when given an optimal DR-plan  (of smallest possible size or max fan-in) for a system $S$,  is that the number of children of the root  (resp.
+number of roots of the DR-plan)---and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved---could be arbitrarily large as a function of the size of $S$.
 This difficulty can persist even after optimal parametrization of the indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic complexity.
 
-In the following, we formulate the problem of \dfn{optimal modification} of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity).
+\subsection{Previous Work}
+% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
+The limitations of these techniques directly motivate the contributions of this paper.}
+We now briefly survey existing techniques for handling the complexity of recombination of DR-plans for qusecs. The limitations of these techniques directly motivate the contributions in this section.
+
+\subsubsection{Optimal Recombination and Solution Space Navigation.}
+For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
+This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
+Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan.
+For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
+
+However, the generality of these algorithms trades-off against efficiency, and, despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan.
+They are prohibitively slow in practice.
+We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
+
+\subsubsection{Configuration Spaces of Underconstrained Systems.}
+For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems
+\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
+Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
+These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon, and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
+While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
+
+\subsection{Optimal Modification for Recombination}
+In the following, we formulate the problem of \dfn{optimal modification} of an indecomposable algebraic system $I$ at some node of a (possibly optimal) DR-plan into a decomposable system with a small DR-plan (low algebraic complexity).
 Leveraging recent results on \dfn{Cayley configuration spaces}, our approach to the optimal modification problem achieves the following:
@@
     \item
-    \header{Small DR-plan.}
+    {\sl Small DR-plan.}
     We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans.
@@
     \item
-    \header{Solution preservation.}
+    {\sl Solution preservation.}
     Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.
@@
     \item
-    \header{Efficient search.}
-    Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type.
-For the case when the modification is bounded, this approach provides an efficient algorithm for recombination.
+    {\sl Efficient search.}
+    Finally,  the so-called \dfn{Cayley} or distance parameter space $\lambda_F$  is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type.
+For the case when the modification (number of Cayley parameters) is bounded, this approach provides an efficient algorithm for recombination.
 We first define the decision version of the problem of optimal modification for decomposition.
@@
 We say that such a tuple $(G,s)$ is a member of the set OMD$_k$.
-We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs with appropriately small $s$.
+We loosely refer to graphs $G$ as OMD with appropriately small $k$ and $s$ or OMD$_k$ with appropriately small $s$.
 
@@
 However, it is not clear how Criterion (c) is met by OMD graphs.
-
-\subsection{Previous Work}
-% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}).
-The limitations of these techniques directly motivate the contributions of this paper.}
-We now briefly survey existing techniques for handling the complexity of recombination of 2D DR-plans. The limitations of these techniques directly motivate the contributions in this section.
-
-\begin{enumerate}
-    \item
-    \header{Optimal Recombination and Solution Space Navigation.}
-    For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
-This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}.
-Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan.
-For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.
-
-    However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan.
-They are prohibitively slow in practice.
-We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.
-
-    \item
-    \header{Configuration Spaces of Underconstrained Systems.}
-    For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
-Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths.
-These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs.
-While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
-\end{enumerate}
+Before we consider this question, we discuss previous work on recombination of DR-plans.
 
@@
 Next we provide the necessary background to describe a specific approach for achieving the requirements (a)--(c) mentioned above, by restricting the class of reduced graphs $G' = G\setminus E_1$ and their isostatic completions $H$ in the above definition of the OMD problem, and using a key theorem of Convex Cayley configuration spaces \cite{sitharam2010convex}.
-This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of the non-edges $F$ is always convex for any given lengths $\delta$ for the edges of $H$ (i.e, over all the realizations of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
+This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of the non-edges $F$ is always convex for any given lengths $\delta$ for the edges of $H$ (i.e.\ over all the realizations of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
 This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$.
-Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces with parameters $F$} (short: \dfn{convexifiable}).
+Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces for some parameters $F$} (short: $H$ is \dfn{convexifiable} \todo{This shortening doesn't exactly make sense}).
 
@@
 A \dfn{partial 2-tree} is a 2-tree minus some edges.
-Partial 2-trees have an alternate characterization as the graphs that avoid $K_4$ minors, and are also called series parallel graphs.
+Partial 2-trees have an alternate characterization as the graphs that avoid $K_4$ minors, and are also called series-parallel graphs.
 
@@
     The proof follows from the fact that 2-trees are well decomposable and have simple DR-plans of size 2.
-We know that $G$ can be reduced by removing $k$ edges to create a partial-2-tree $G'$ which can then be completed to an (isostatic) 2-tree by adding some set of non-edges $F$.
+We know that $G$ can be reduced by removing $k$ edges to create a partial 2-tree $G'$ which can then be completed to an (isostatic) 2-tree by adding some set of non-edges $F$.
 Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2, proving the theorem.
@@
 
-\sidenote{A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.}
+\sidenote{A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals of values rather than exact values~~\cite{sitharam2010convex}.}
 %cite my paper with Heping!!!
@@
 [Using Cayley configuration space].
-A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e., its DR-plan has a root and 9 children corresponding to the 9 edges.
+A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e.\ its DR-plan has a root and 9 children corresponding to the 9 edges.
 Solving or recombining the system $(G,\delta)$ corresponding to the root of this DR-plan involves solving a quadratic system with 8 equations and variables.
@@
 The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single interval of attainable length values $\lambda_F$ for the edge $f$.
-When $\delta$ is generic, i.e, does not admit collinearities or coincidences in the realizations of $(G,\delta)$, the realization space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16 solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries), with distinct orientation types $\sigma_p$ (two orientation choices for each of the 4 triangles) that can be obtained by solving a sequence of 4 single quadratics in 1 variable (DR-plan of size 2).
+When $\delta$ is generic, i.e.\ does not admit collinearities or coincidences in the realizations of $(G,\delta)$, the realization space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16 solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries), with distinct orientation types $\sigma_p$ (two orientation choices for each of the 4 triangles) that can be obtained by solving a sequence of 4 single quadratics in 1 variable (DR-plan of size 2).
 By subdivided binary search in the interval $\lambda_f \in \Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is found when the length of the nonedge $e$ in  the realization $q^p_{\lambda_f}$ is $\delta_e$.
@@
 %
-It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e, the number of edges to be removed to ensure an isostatic completion that is decomposable with a small DR-plan) by considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than partial 2-trees but nevertheless have convex Cayley configuration spaces at least when the realization space is restricted to a sufficiently comprehensive orientation type.
+It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e.\ the number of edges to be removed to ensure an isostatic completion that is decomposable with a small DR-plan) by considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than partial 2-trees but nevertheless have convex Cayley configuration spaces at least when the realization space is restricted to a sufficiently comprehensive orientation type.
 In particular, the so-called \dfn{tree-decomposable graphs of low Cayley complexity} \cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not partial 2-trees.
@@
 In addition, the bounds of these intervals are of low algebraic complexity.
-More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e., the computation of these bounds is tree-decomposable.
-Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or can be computed by solving a triangularized system of quadratics.
+More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e.\ the computation of these bounds is tree-decomposable.
+Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or they can be computed by solving a triangularized system of quadratics.

### 2015-05-16 doc/GMP2015/revision_2/s05_problems.tex
Revision 2. 24 hours to submission.

@@
 \subsubsection{Special classes of small DR-plans}
-As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially a single quadratic at a time.
+As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially by solving a single quadratic at a time.
 
@@
 
-We discuss the complexity of the restricted OMD problem (when $k$ is part of the input) in Section \ref{sec:futurework}.
+We discuss the complexity of the restricted OMD problem (when $k$ is part of the input) in the open-problem Section \ref{sec:futurework}.
 
@@
 %
-\subsubsection{Optimal Modification, Completion and Recombination: formal
+\subsubsection{Optimal Modification, Completion and Recombination: previous work and formal
 connections}
@@
 Otherwise, replace it by  a 2-tree graph of the shared vertices.
-Finally, we add the additional restriction to the OM Problem that when any edge in a solved subgraph is chosen among the $k$ edges to be removed, in fact the entire solved subgraph must be removed  and all of its edges must be counted in $k$.
+Finally, we add the additional restriction to the OM problem that when any edge in a solved subgraph is chosen among the $k$ edges to be removed, in fact the entire solved subgraph must be removed  and all of its edges must be counted in $k$.
 
-This reduction is used also for transferring  algorithms for optimal DR-planning, recombination, completion, OMD, and other problems from bar-joint systems to so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin}, by showing that the problems for the latter are reduced to the corresponding problems on bar-joint systems.
+This reduction is used also for adapting  algorithms for optimal DR-planning, recombination, completion, OMD, and other problems from bar-joint systems to so-called \dfn{body-hyperpin}, defined in Section \ref{sec:bodypin}, by showing that the problems for the latter are reduced to the corresponding problems on bar-joint systems.
 

### 2015-05-16 doc/GMP2015/revision_2/s06_problempinning.tex
Revision 2. 24 hours to submission.

@@
     \caption{Example of a single monolayer of a silica (silicon and oxygen) glassy structure.
-This can be thought of as a triangle-multipin qusecs where the silicon atoms are the triangles and the oxygen atoms are the pins.
+This can be viewed as a triangle-multipin qusecs where the silicon atoms are the triangles and the oxygen atoms are the pins.
 Not shown here are other monolayers stacked on this one.
@@
 We can use qusecs DR-plans to design materials such as disordered graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}.
-We will investigate a more specific problem in a somewhat more general setting: the problem of finding boundary conditions (additional constraints) to add to an underconstrained monolayer to make it isostatic.
+We investigate a more specific problem in a somewhat more general setting: the problem of finding boundary conditions (additional constraints) to add to an underconstrained monolayer to make it isostatic.
 This can be done in a number of ways: (1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}); (2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or (3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).
 
-In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress and other property design.
-To answer these questions, we first introduce the qusecs that are used to model these materials.
-We will discuss item (2) in detail in this section.
+In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress, flex and other property design related to the rigidity matrix.
+To answer these questions, we first introduce the qusecs' that are used to model these materials.
+In this section, we discuss Item (2) in detail.
 
@@
 \begin{definition}
-    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies, some subsets of which are pinned together by pins.
+    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies,  subsets of which are pinned together by pins; i.e, are incident at a common point.
 \end{definition}
@@
 \begin{remark*}\label{rem:bodypin_is_barjoint}
-    A body-hyperpin qusecs is a special case of bar-joint qusecs of the first 5 sections.
-As such, the DR-planning for isostatics discussed in Section \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
+    A body-hyperpin qusecs is a special case of bar-joint qusecs of the previous sections of the paper.
+As such, the DR-planning for isostatic systems discussed in Section \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} still go through with minor modifications.
 \end{remark*}
 
-For the rest of this section, we will only be dealing with the DR-plan of such qusecs.
-As such, we will refer only to the underlying hypergraph of the qusecs.
+For the remainder of this section, we deal only with the DR-plan of such qusecs.
+Hence, we refer only to the combinatorics or underlying hypergraph of the qusecs.
 We now introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}, for which the optimal completion problem is significantly easier.
@@
 \begin{definition}
-    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares at most 3 pins with other bodies.
+    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares  pins with at most 3 other bodies.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
-For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
+For such hypergraphs, {\em 1 and 2-dof} can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
 \end{definition}
@@
 
-We can develop the notion of a canonical DR-plan when the hypergraph is $(k,l)$-sparse (i.e., combinatorially independent) using certain notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}, provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}).
-In particular, we can define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.
+Provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}),
+the notion of a canonical DR-plan extends directly to the case when the hypergraph is $(k,l)$-sparse (i.e., independent) using the straightforward notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions as in Section \ref{sec:DRP}.
+In particular, we define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.
 
@@
 \label{obs:bodypin_drp}
-    For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
+For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex-maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
 \end{observation*}
@@
 
-The strong Church-Rosser property holds, making all canonical DR-plans optimal:
+As in Section \ref{sec:DRP},  a strong Church-Rosser property holds, making all canonical DR-plans optimal:
 %
-\begin{remark}\label{rem:1dofcanon}
+\begin{observation*}\label{rem:1dofcanon}
     When the input is independent, all $(3,4)$-tight canonical DR-plans are optimal.
 We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
-This is because such $(3,4)$-tight graphs are matroidal and have a pebble game \cite{Lee:2007:PGA}.
-\end{remark}
+\end{observation*}
+
+The above-mentioned algorithm exists because such $(3,4)$-tight graphs are matroidal and have a pebble game \cite{Lee:2007:PGA}.
 
@@
 \label{obs:2dof_case}
-    For the 2-dof case, if a similar statement to Remark \ref{rem:1dofcanon} can be made, then Theorem \ref{thm:1dofcase} is true for 2-dof as well.
+    For the 2-dof case, provided an analogous statement to Observation \ref{rem:1dofcanon} holds, then Theorem \ref{thm:1dofcase} holds for the 2-dof  systems.
 \end{observation*}
@@
 
-\begin{observation*}
+\begin{remark}
 \label{obs:algebraic_completion}
-    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer permit the algorithm given in the proof.
-In particular the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if that case were matroidal and an optimal DR-plan is known).
-\end{observation*}
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer be optimized by the algorithm given in that proof.
+In particular, the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if the case were matroidal and an optimal DR-plan is known).
+\end{remark}
 

### 2015-05-16 doc/GMP2015/revision_2/s07_pinnedline.tex
Revision 2. 24 hours to submission.

@@
 
-The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as \dfn{pinned line-incidence systems}.
+The canonical DR-plan of Section \ref{sec:DRP} can additionally be applied to analyze and solve the structure of cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as a third type of qusecs, \dfn{pinned line-incidence systems}.
 
-\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property \cite{buehler2008nanomechanics}.
+\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical properties \cite{buehler2008nanomechanics}.
 \dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}.
@@
 
-In the case of microfibril cross-linking, each fibril is attached to some fixed larger organelle/membrane at one spot.
+In the case of microfibril cross-linking, each fibril is attached to some fixed larger organelle/membrane at one site.
 Consequently, each fibril can be modeled as an edge of the graph, with the attachment being the corresponding pin.
-The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge.
+The two cross-linkings in which the fibril participates are modeled as the two vertices in $V$ defining the edge.
 
@@
 
-    The \dfn{canonical DR-plan} of $G$  is one in which the  children containing rigid subgraphs are connected isostatic vertex-maximal subgraphs of the parent.
+    The \dfn{canonical DR-plan}  of $G$  is one in which the  child rigid subgraphs are connected, isostatic vertex-maximal subgraphs of the parent.
 \end{definition}

### 2015-05-16 doc/GMP2015/revision_2/s08_conclusion.tex
Revision 2. 24 hours to submission.

@@
 We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
-For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, which is a strong Church-Rosser property.
+For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, a strong Church-Rosser property.
 This gives an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
 
-We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan.
+We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise indecomposable systems  at nodes of a DR-plan.
 These results rely on a recent theory of Convex Cayley Configuration spaces.
@@
 \label{sec:futurework}
+\label{sec:open}
 
-One open problem comes from Section \ref{sec:DRP}:
+The first set of problems are from Section \ref{sec:DRP}:
 \begin{openproblem}
-    Is there a lower complexity algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
+    Is there a more efficient algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
 \end{openproblem}
 
-A few natural  open questions concern the following common theme that runs through the paper:
+\begin{conjecture}
+\label{conj:mfaisoptimal}
+    The Modified Frontier Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
+\end{conjecture}
+
+The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA, although running in time $O(n^3)$, is a bottom-up algorithm, involving complex datastructures.
+However, a proof of optimality, even if it exists, would not be possible without the new notion of a canonical DR-plan at hand.
+The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the sequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.
+Since it is known \cite{lomonosov2004graph} that the DR-plan generated by MFA is cluster-minimal, an alternate conjecture is the following.
+
+\begin{conjecture}
+\label{conj:mfaisoptimal:rephrase}
+For independent graphs, cluster-minimal DR-plans are optimal. In fact, for independent graphs, cluster-minimality and canonical are equivalent properties of a DR-plan.
+\end{conjecture}
 
 \begin{openproblem}
-    For fixed $k$ we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan (Section~\ref{sec:table}).
+\label{open:sparsitymatroid}
+Although generic rigidity is a property of graphs, and moreover,  in the case of qusecs, generic rigidity has a combinatorial sparsity and tightness-based characterization, the original definition of independence in the rigidity matroid requires an algebraic notion of independence of vectors of indeterminates over $\mathbf{R}$).
+Thus the definition of the DR-plan requires algebra over the reals.
+In fact, the recursive decomposition problem is not tied to geometric constraint graphs or
+an algebraic-geometric or mechanical notion of rigidity, and can be defined for any graph using
+the notion of an abstract rigidity matroid \cite{graver93book}.
+This is  type of matroid with two additional matroid axioms,  abstract rigidity matroids can be defined in a purely graph-theoretic manner,  with no need for algebra in their definition.
+However, such abstract rigidity need not have a sparsity characterization.
+On the other hand, there are sparsity matroids that do not correspond to any notion of abstract rigidity.
+However, when an abstract rigidity matroid is also a sparsity matroid, then the techniques of this
+paper  directly apply and we can obtain purely combinatorially defined recursive decompositions of graphs.
+\end{openproblem}
+
+A few natural  open questions concern the following common theme that runs through the optimal recombination and later sections of the paper:
+
+\begin{openproblem}
+    For fixed $k$, we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases for a somewhat different optimization of the DR-plan (Section~\ref{sec:table}).
 However, in the running time of all of these algorithms, $k$ appears in the exponent.
+Can $k$ be removed from the exponent?
 \end{openproblem}

### 2015-05-16 src/Graph/Graph.cpp
Does coarse, uniform sampling over entire Cayley configuration space.

@@ -155,11 +155,11 @@ void Graph::remove_edge(Vertex_ID v0, Vertex_ID v1) {
 }
 
 
-bool Graph::has_edge(Vertex_ID v0, Vertex_ID v1) {
+bool Graph::has_edge(Vertex_ID v0, Vertex_ID v1) const {
     return boost::edge(v0, v1, *this).second;
 }
 
-Edge_ID Graph::edge(Vertex_ID v0, Vertex_ID v1) {
+Edge_ID Graph::edge(Vertex_ID v0, Vertex_ID v1) const {
     return boost::edge(v0, v1, *this).first;
 }
 
@@ -408,12 +408,12 @@ std::pair<Edge_Iterator, Edge_Iterator> Graph::edges() const {
     return boost::edges(*this);
 }
 
-unsigned int Graph::degree_of_vertex(Vertex_ID v) {
+unsigned int Graph::degree_of_vertex(Vertex_ID v) const {
     return boost::degree(v, *this);
 }
 
 
-std::set<Vertex_ID> Graph::vertices_adjacent(Vertex_ID v) {
+std::set<Vertex_ID> Graph::vertices_adjacent(Vertex_ID v) const {
     std::set<Vertex_ID> ret;
 
     boost::graph_traits<Graph>::adjacency_iterator v_it, v_end;
@@ -426,7 +426,7 @@ std::set<Vertex_ID> Graph::vertices_adjacent(Vertex_ID v) {
     return ret;
 }
 
-std::set<Vertex_ID> Graph::vertices_adjacent(std::set<Vertex_ID> &v_set) {
+std::set<Vertex_ID> Graph::vertices_adjacent(std::set<Vertex_ID> &v_set) const {
     std::set<Vertex_ID> ret;
     for (std::set<Vertex_ID>::iterator set_v = v_set.begin(); set_v != v_set.end(); set_v++) {
         boost::graph_traits<Graph>::adjacency_iterator v, v_end;
@@ -609,7 +609,7 @@ void Graph::read_from_file(const char* filename) {
 
 
 
-void Graph::print_vertices(unsigned int indent) {
+void Graph::print_vertices(unsigned int indent) const {
     std::string tabs = "";
     for (int i = 0; i < indent; i++) tabs+="\t";
 
@@ -620,13 +620,14 @@ void Graph::print_vertices(unsigned int indent) {
         std::cout << tabs
             << "ID: " << v << " | "
             << "Name: " << (*this)[v].name  << " | "
[…]

### 2015-05-16 src/Graph/Graph.hpp

@@ -110,8 +110,8 @@ public:
     void remove_edge(Edge_ID e);
     void remove_edge(Vertex_ID v0, Vertex_ID v1);
 
-    bool has_edge(Vertex_ID v0, Vertex_ID v1);
-    Edge_ID edge(Vertex_ID v0, Vertex_ID v1); // can be unsafe... edge may not exist
+    bool has_edge(Vertex_ID v0, Vertex_ID v1) const;
+    Edge_ID edge(Vertex_ID v0, Vertex_ID v1) const; // can be unsafe... edge may not exist
 
     // v0 disappears, v1 remains
     void contract_edge(Vertex_ID v0, Vertex_ID v1);
@@ -120,9 +120,9 @@ public:
     std::pair<Vertex_Iterator, Vertex_Iterator> vertices() const;
     std::pair<Edge_Iterator, Edge_Iterator> edges() const;
 
-    unsigned int degree_of_vertex(Vertex_ID v);
-    std::set<Vertex_ID> vertices_adjacent(Vertex_ID v);
-    std::set<Vertex_ID> vertices_adjacent(std::set<Vertex_ID> &v_set);
+    unsigned int degree_of_vertex(Vertex_ID v) const;
+    std::set<Vertex_ID> vertices_adjacent(Vertex_ID v) const;
+    std::set<Vertex_ID> vertices_adjacent(std::set<Vertex_ID> &v_set) const;
 
     Vertex_Iterator find_vertex(const char *name) const;
 
@@ -137,10 +137,35 @@ public:
     void write_to_file(const char* filename);
     void read_from_file(const char* filename);
 
-    void print_vertices(unsigned int indent = 0);
-    void print_edges(unsigned int indent = 0);
+    void print_vertices(unsigned int indent = 0) const;
+    void print_edges(unsigned int indent = 0) const;
 private:
     // bool _is_series_parallel_aux();
+};
+
+
+
+
+
+class Mapped_Graph_Copy : public Graph {
+public:
+    Mapped_Graph_Copy(const Graph *g);
+
+    // This "copies through"... Makes a copy of the input mapped graph copy 'g'
+    // but instead of mapping to 'g' it maps to 'g.orig'
+    Mapped_Graph_Copy(const Mapped_Graph_Copy &g);
+
+
+    // input is a copy Vertex_ID
+    Vertex_ID original_vertex(Vertex_ID v) const;
+    // input is an input Vertex_ID
[…]

### 2015-05-16 src/Graph/Isostatic_Graph_Realizer.cpp

@@ -7,93 +7,83 @@
 
 
 
-graph_copy::graph_copy(Graph *g) {
-    orig = g;
 
-    Vertex_Iterator v, v_end;
-    for (boost::tie(v, v_end) = orig->vertices(); v != v_end; v++) {
-        // Vertex_ID newv = copy.add_vertex();
-        Vertex_ID newv = copy.add_vertex((*orig)[*v]);
-        orig_to_copy[*v] = newv;
-        copy_to_orig[newv] = *v;
-    }
-
-    // std::cout << "========graph_copy:" <<std::endl;
-    // print_copy_to_orig(1);
-    // copy.print_vertices(2);
-
-    Edge_Iterator e, e_end;
-    for (boost::tie(e, e_end) = orig->edges(); e != e_end; e++) {
-        std::pair<Vertex_ID, Vertex_ID> vs = orig->verts_on_edge(*e);
-        copy.add_edge(orig_to_copy[vs.first], orig_to_copy[vs.second], (*orig)[*e].length);
-    }
+Isostatic_Graph_Realizer::Isostatic_Graph_Realizer(const Graph *g) {
+    this->in_graph = g;
+    this->working_copy = new Mapped_Graph_Copy(g);
 }
 
-void graph_copy::print_copy_to_orig(unsigned int indent) {
-    for (int i = 0; i < indent; i++) std::cout << "\t";
-    std::cout << "copy_to_orig mapping (new, old):" << std::endl;
+Isostatic_Graph_Realizer::~Isostatic_Graph_Realizer() {
+    delete(working_copy);
+}
 
-    for (std::map<Vertex_ID, Vertex_ID>::iterator v = copy_to_orig.begin(); v!= copy_to_orig.end(); v++) {
-        for (int i = 0; i < indent+1; i++) std::cout << "\t";
-        // std::cout << "(" << v->first << ", " << v->second << ")" << std::endl;
-        std::cout << "(" << v->first << ": " << copy[v->first].name << ", " << v->second << ": " << (*orig)[v->second].name << ")" << std::endl;
+std::list<Mapped_Graph_Copy*> Isostatic_Graph_Realizer::realize() {
+    std::vector<std::pair<Vertex_ID, Vertex_ID> > dropped = make_partial_2_tree();
+    std::list<Edge_ID> nonedges_ordered = make_2_tree();
+
+    std::set<Edge_ID> nonedges;
[…]

### 2015-05-16 src/Graph/Isostatic_Graph_Realizer.hpp

@@ -4,54 +4,64 @@
 #include "Graph.hpp"
 
 #include <vector>
+#include <list>
 #include <map>
 #include <utility>
-
-
-
-struct graph_copy {
-    Graph *orig;
-
-    Graph copy;
-    std::map<Vertex_ID, Vertex_ID> orig_to_copy, copy_to_orig;
-
-    graph_copy(Graph *g);
-
-    void print_copy_to_orig(unsigned int indent = 0);
-};
 
 
 class Isostatic_Graph_Realizer {
 public:
     // assumes that input is isostatic
-    Isostatic_Graph_Realizer(Graph *g);
+    Isostatic_Graph_Realizer(const Graph *g);
+    ~Isostatic_Graph_Realizer();
 
-    bool realize();
+    std::list<Mapped_Graph_Copy*> realize();
+
+
+    Mapped_Graph_Copy *working_copy;
 
 private:
-    Graph *in_graph;
+    const Graph *in_graph;
 
     // used by methods
-    std::vector<std::pair<Vertex_ID, Vertex_ID> > contractions;
-    std::map<Vertex_ID, bool> realized;
+    // std::vector<std::pair<Vertex_ID, Vertex_ID> > contractions;
+    // std::map<Edge_ID, bool> is_nonedge;
+    // std::map<Edge_ID, std::pair<double, double> > nonedge_interval;
+    // std::map<Vertex_ID, bool> realized;
 
-    // makes a copy of in_graph,
-    // destructively determines if it's partial 2-tree (series parallel, k4 minor avoiding...)
-    bool is_partial_2_tree(graph_copy &gc);
+    // destructively determines if gc.copy is a partial 2-tree (series parallel, k4 minor avoiding...)
+    static bool is_partial_2_tree(Mapped_Graph_Copy &gc);
 
-    // returns a list of dropped edges
+    // returns a list of dropped edges from in_graph, in terms of working_copy vertex_id's
     std::vector<std::pair<Vertex_ID, Vertex_ID> > make_partial_2_tree();
 
-    // assumes in_graph is now a partial 2 tree
-    // returns a list of added edges
+    // assumes working_copy is now a partial 2 tree
+    // returns a list of added edges in working_copy
     // std::vector<Edge_ID> make_2_tree();
-    std::vector<std::pair<Vertex_ID, Vertex_ID> > make_2_tree();
+    std::list<Edge_ID> make_2_tree();
[…]

### 2015-05-16 src/main.cpp

@@ -550,7 +550,8 @@ public:
 
 enum runtime_options {
     ro_drp_2d,
-    ro_display_realized_graph,
+    ro_display_linkage,
+    ro_display_framework,
     ro_test
 };
 
@@ -597,8 +598,10 @@ int main(int argc, char **argv) {
             if (!val.compare("0")) {
                 runtime_option = ro_drp_2d;
             } else if (!val.compare("1")) {
-                runtime_option = ro_display_realized_graph;
+                runtime_option = ro_display_linkage;
             } else if (!val.compare("2")) {
+                runtime_option = ro_display_framework;
+            } else if (!val.compare("3")) {
                 runtime_option = ro_test;
             } else {
                 cerr << "Invalid value for flag " << flag << "." << endl;
@@ -631,7 +634,7 @@ int main(int argc, char **argv) {
     // copy.write_to_file("test_files/out.dot");
 
     // get layout... x/y coords
-    if (runtime_option != ro_display_realized_graph) {
+    if (runtime_option != ro_display_framework) {
         g.set_layout();
         g.get_graph_in_range(-0.87, 0.87, -0.87, 0.87);
     } else {
@@ -661,7 +664,16 @@ int main(int argc, char **argv) {
         cout << "Running DRP_2D on \"" << dot_file << "\"." << endl;
 
         myWindow.get_drp();
-    } else if (runtime_option == ro_display_realized_graph) {
+    } else if (runtime_option == ro_display_linkage) {
+        cout << "Displaying realized graph \"" << dot_file << "\"." << endl;
+
+        // Subgraph sg(&g);
+        // std::pair<Vertex_Iterator, Vertex_Iterator> vs = g.vertices();
+        // sg.induce(vs.first, vs.second);
+        // Pebbled_Graph pg(&sg);
+
+        // cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
+    } else if (runtime_option == ro_display_framework) {
         cout << "Displaying realized graph \"" << dot_file << "\"." << endl;
 
         Subgraph sg(&g);
@@ -671,10 +683,12 @@ int main(int argc, char **argv) {
 
         cout << "Pebbles remaining: " << pg.pebble_game_2D() << endl;
     } else if (runtime_option == ro_test) {
-        cout << "Testing edge contraction." << endl;
+        cout << "Testing realization." << endl;
 
         Isostatic_Graph_Realizer isg(&g);
         isg.realize();
+
+        g = *isg.working_copy;
 
         // int i = 0;
 

### 2015-05-17 doc/GMP2015/revision_2/response_to_reviewers.txt
Revision 2, finished.

@@
 Response to Reviewers
+
 =====================
@@
 Reviewer #1
+
 -----------
+
 Overall I think this paper is timely and interesting.
@@
 
-Response:
-- Abstract had substantial rework.
-While we don't give a complete example in the abstract, we do mention which materials we model in this paper.
-In the introduction, we do explain much more clearly why we would want to use qusecs.
+RESPONSE TO REVIEWER # 1:
+
+- Abstract had substantial rework. We felt it would be nontraditional to give a complete example in the abstract. But we do mention in the abstract which materials we model in the paper.
+
+In the introduction, we explain much more clearly why we would want to use qusecs.
+
 - We moved the definitions into the body of the paper.
-- We provide a link where the video and code will be made available.
-Currently a work in progress, but we are committed to providing a well-documented, open source, highly functional application.
-The video will be made sooner.
+
+- We cite a link where the video and code will be made available.
+We do not intend the code to be part of the submission.
+If the journal publishes videos linked to the paper, and the editors find it appropriate, we will upload the video files along with the final version.
 
 Reviewer #2
+
 -----------
+
 This paper proposes a recursive decomposition planning algorithm to model a material layer as a solution of a 2D geometric constraint system.
@@
 
-Response:
-1. We are now submitting to the main journal.
+RESPONSE TO REVIEWER #2:
+
+1. Our revision is being sent to the main journal rather than the
+conference special issue and we were
+told that the page limit does not apply.
+
 2. We moved preliminaries into the main body of the paper (section 2).
+
 3. We moved the algorithm into its own subsection and revised the proof to be more clear.
-4. ...
-5. We have addressed this. We have reformatted our unusual headings and such.
+
+4. We have enumerated the examples.
+
+5. We have addressed this and made the style uniform. We have reformatted our unusual headings etc..
 
 Further Changes
+
 ===============
@@
 
-One of the main issues with our paper was that it would not be a good fit for GMP.
-After much discussion, we decided that we agree and are now submitting the paper for consideration to be published in the standard journal (see next section in this file).
+After much discussion,  we are now submitting the revision to appear in the regular journal (see next section in this file) rather than the conference special issue.
 The changes made in this revision are discussed below.
 
-Title: Because of the substantial changes to the paper, we decided the title no longer made sense in the context of the revised paper.
-As such, we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'.
+Title: Because of the substantial changes to the paper, we decided the title was no longer appropriately indicative of the revised paper.
+Hence we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'.
 We believe this also makes the intent of our paper much more clear.
@@
 Additionally, many sentences were rephrased/restructured to improve clarity.
-Also, because our paper is no longer being considered for the special issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).
+Also, because our paper is being considered to appear in the regular issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).
 
-Two concepts were further fleshed out and discussed in this iteration: (a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and (b) the extension of our theory to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition..
-This is reflected throughout the paper, but is especially prevalent in (1), (2), (3), and (8).
+Some concepts were further fleshed out and discussed in this iteration:
+(a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and
+(b) the reasons why neither canonical DRP nor cluster-minimal DRP's are optimal when the original constraint graph is overconstrained (not independent).
+(c) We have also modified the writing so that it is clear that our theory extends to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.
 
-Abstract: Substantially altered for clarity. Similar content.
+This is reflected throughout the paper, but is especially  in (1), (2), (3), and (8).
+
+Abstract: Substantially altered for clarity.
 
@@
 
-(2) Preliminaries: Moved back in the definitions from the appendix (as per reviewer request). Moved in discussion of previous work on DR-plans from (1).
+(2) Preliminaries:  The definitions from the appendix (as per reviewer request) have been moved to this section.  Also reorganized the discussion of previous work on DR-plans.
 
@@
 There are new figures to demonstrate the concepts; instead of the previous trivial (in the sense that it was tree-decomposable) example of three triangles, we use three $C2\times C3$s intersecting on a single triangle.
-Also, there is a new figure that corrects the previous, erroneous figure discussing cluster minimality in the previous revision.
-Revised proofs for clarity, but there were no fundamental changes.
-Put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request.
-Moved out cluster minimality conjecture to the conclusion.
-Added a discussion on the overconstrained problem and NP-hardness (3.3).
+Also, there is a new figure illustrating cluster minimality or canonical does not imply optimality in the case of overconstrained graphs.
+This was not present in the previous revision.
+We revised proofs for clarity, but there were no fundamental changes.
+
+We put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request.
+We moved the cluster minimality conjecture to the conclusion.
+We added a discussion on the overconstrained problem and NP-hardness (3.3).
 
@@
 
-(5) and (6) Application sections: No changes.
+(5) and (6) Application sections: No substantial changes other than rewordings.
 
-(8) Conclusion: Moved in conjectures about cluster minimality from (3). Added new open problem (OP2).
+(8) Conclusion: Moved in conjectures about MFA and cluster minimality from (3). Added new open problem (OP2).
 
@@
 
-Journal Change
+Journal Issue Change
+
 ==============
@@
 Originally the paper was submitted for consideration of publication in the special issue of CAGD for GMP2015.
-However, pending our major revision, this was changed to the regular issue of CAGD.
+
+However, for various reasons, this was changed to the regular issue of CAGD (and although our paper was invited to be presented at GMP15, we were unable to attend).
 We received confirmation from the GMP co-chairs that this paper would now be handled by the CAGD editors-in-chief.
+
 Below, we have copied the email we received from the conference chairs, stating the change.
 
-Upon submission, the exact content of this section will also be emailed to editors-in-chief, along with our submission number: XX.
+If there are additional questions in regard to this issue, please contact the corresponding author.
 
-If there are additional questions in regard to this issue, please send an email.
+-----------
 
-Email
------
-GMP2015 <gmp2015@easychair.org> Wed, Mar 25, 2015 at 7:17 AM
+Reference email from GMP15 co-chairs
+
+-----------
+
+GMP2015 <gmp2015@easychair.org>
+
+Wed, Mar 25, 2015 at 7:17 AM
+
 To: Troy Baker <tbaker@cise.ufl.edu>
+
 Dear Troy Baker,

### 2015-05-17 doc/GMP2015/revision_2/s01_introduction.tex
Revision 2, finished.

@@
 With the use of \dfn{decomposition-recombination (DR-) planning}, the complexity is dominated by the size of the largest subsystem that is solved, or recombined, from the solutions of its child subsystems, i.e.\ the maximum fan-in occurring in a DR-plan.
-In addition, navigating and analyzing the solution spaces, as well as designing constraint systems with desired solution spaces leads to the \dfn{optimal decomposition-recombination (DR-) planning} problem~\cite{sitharam2005combinatorial, hoffman2001decompositionI, hoffman2001decompositionII}.
-\todo{Feels very clunky.}
+In addition, navigating and analyzing the solution spaces, as well as designing constraint systems with desired solution spaces, leads to the \dfn{optimal decomposition-recombination (DR-) planning} problem~\cite{sitharam2005combinatorial, hoffman2001decompositionI, hoffman2001decompositionII}.
 
@@
 Furthermore, a layer is often  either \dfn{isostatic} or \dfn{underconstrained} (not \dfn{self-stressed}/\dfn{overconstrained}, \seedefsd).
-\todo{consider...} These properties, as well as quasi-uniformity, aperiodicity, self-similarity, and layered structure, are natural consequences of evolutionary pressures or design objectives such as stability, minimizing mass, optimally distributing external stresses, and participating in the assembly of diverse and multifunctional, larger structures.
+These properties, as well as quasi-uniformity, aperiodicity, self-similarity, and layered structure, are natural consequences of evolutionary pressures or design objectives such as stability, minimizing mass, optimally distributing external stresses, and participating in the assembly of diverse and multifunctional, larger structures.
 %
@@
 In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained.
-Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied {\em cluster minimality}  \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).
+Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied \dfn{cluster minimality}~\cite{hoffman2001decompositionI}.
 Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
@@
 
-\todo{We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.}
+A software implementation of our algorithms and videos demonstrating the software are publicly available online\footnote{See footnote 1.}.

### 2015-05-17 doc/GMP2015/revision_2/s02_preliminaries.tex
Revision 2, finished.

@@
 
-\subsection{Geometric Constraint systems and Combinatorial Rigidity}
+\subsection{Geometric Constraint Systems and Combinatorial Rigidity}
 \label{sec:prelim:defs}
@@
 
-\subsection{Decomposition-Recombination (DR) Plans}
+\subsection{Decomposition-Recombination (DR-) Plans}
 %
@@
 
-\ClearMyMinHeight
-\SetMyMinHeight{.4}{img/svg/triangle}
-\SetMyMinHeight{.3}{img/svg/triangle_candrp}
-\SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}
+% \ClearMyMinHeight
+% \SetMyMinHeight{.4}{img/svg/triangle}
+% \SetMyMinHeight{.3}{img/svg/triangle_candrp}
+% \SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}
 
@@
 
-See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
+See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
 
-\subsection{Previous Work on DR-plans}
+\subsection{Previous Work on DR-Plans}
 \label{sec:prev}
@@
 
-\subsubsection{Finding (vertex)-maximal generically rigid subsystems}
+\subsubsection{Finding (Vertex)-Maximal, Generically Rigid Subsystems}
 Fast, graph-based algorithms exist (pebble-game~\cite{Jacobs:1997:PG,hoffmann1997solvablesubsets,jermann2006decomposition,Lee:2007:PGA}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs.
@@
 
-\subsubsection{Optimal Recursive Decomposition (DR-planning)}
+\subsubsection{Optimal Recursive Decomposition (DR-Planning)}
 Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{lomonosov2004graph,sitharam2005combinatorial,jermann2006decomposition} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim.
-For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm \todo{rephrase?
-maybe not} without involving geometric information in the constraint system.
+For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm without involving the geometric information in the constraint system.
 Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in~\cite{hoffman2001decompositionI}.
@@
 
-\subsubsection{DR-plans for special classes and with other criteria}
+\subsubsection{DR-plans for Special Classes and with Other Criteria}
 For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems~\cite{owen1991algebraic,fudos1997graph,joan-arinyo2004revisiting} common in computer aided mechanical design (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal.

### 2015-05-17 doc/GMP2015/revision_2/s03_drp.tex
Revision 2, finished.

@@
-\section{Main Result: canonical DR-plan, Optimality, and Algorithm}
+\section{Main Result: Canonical DR-Plan, Optimality, and Algorithm}
 \label{sec:DRP}
@@
 
-\subsection{Canonical DR-plan}
+\subsection{Canonical DR-Plan}
 In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
@@
 \begin{definition}\label{def:canonical_drplan}
-    The \dfn{canonical DR-plan} \addin{of a graph $G$} satisfies the following three properties:
-    (1) it is a DR-plan \addin{of $G$};
+    The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
+    (1) it is a DR-plan of $G$;
     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
@@
 \label{theorem:main}
-    A canonical DR-plan exists \addin{for a graph $G$} and any canonical DR-plan is optimal \addin{if $G$ is independent.}
+    A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.
 \end{theorem}
@@
 
-First, note  that any  DR-plan \addin{$R$} without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan \addin{$R$})
-is the subgraph of some \addin{cluster} of $C$.
-Thus without loss of generality, we can assume that \cutout{the} \addin{an} optimal DR-plan satisfies Property (2) of a canonical DR-plan.
+First, note  that any  DR-plan $R$ without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$)
+is the subgraph of some cluster of $C$.
+Thus without loss of generality, we can assume that an optimal DR-plan satisfies Property (2) of a canonical DR-plan.
 
-\addin{The proof of optimality of a canonical DR-plan is by induction on its height.
+The proof of optimality of a canonical DR-plan is by induction on its height.
 The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges.
-The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.}
-\addin{For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.}
+The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
+For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.
 Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
@@
 We show the two claims below.
-The first claim is that \addin{for a node $C$ whose clusters have trivial pairwise intersections}, any DR-plan \cutout{$R$} \addin{of $C$} that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
-\addin{Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.}
-The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, \addin{every canonical DR-plan of $C$ that uses any possible choice} of two such subgraphs of $C$ as children \cutout{ (minimum possible fan-in at $C$)} results in \addin{a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the}     {\em same maximal antichain $S$ of descendants $D$ of $C$}.
-\addin{The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment.} I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; \addin{this follows from properties of maximal antichains} that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$.
-Thus any  DR-plan  \cutout{$R$} that satisfies Property (2) \addin{and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children} must \addin{also} contain every element of $S$.
-The two claims complete the proof that every canonical
-DR-plan is optimal.
+The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
+Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.
+The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the {\em same maximal antichain $S$ of descendants $D$ of $C$}.
+The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment.
+I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$.
+Thus any  DR-plan that satisfies Property (2) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$.
+The two claims complete the proof that every canonical DR-plan is optimal.
 
@@
 
-\addin{To prove Claim 2}, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
-This means that, by Lemma \ref{lemma:combined_lemma}, \addin{Point} \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself.
+To prove Claim 2, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself.
 Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
@@
 \begin{remark}
-The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{cluster minimality} (defined in \cite{hoffman2001decompositionI} and in Section \ref{sec:prev}) holding for \cutout{these} canonical-optimal DR-plans.
+The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{cluster minimality} (defined in \cite{hoffman2001decompositionI} and in Section \ref{sec:prev}) holding for canonical-optimal DR-plans.
 \end{remark}
@@
 
+% \ClearMyMinHeight
+% \SetMyMinHeight{.3}{img/epsfromtikz/c2c3_of_k33s-0}
+% \SetMyMinHeight{.7}{img/epsfromtikz/c2c3_of_k33s-1}
+
+% \begin{figure*}\centering%
+%     %
+%     \begin{subfigure}{0.3\linewidth}\centering
+%         \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-0}
+%         \caption{}\label{fig:c2c3ofk33s:a}
+%     \end{subfigure}%
+%     %
+%     \hfill
+%     \begin{subfigure}{0.7\linewidth}\centering
+%         \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-1}
+%         \caption{}\label{fig:c2c3ofk33s:b}
+%     \end{subfigure}%
+%     %
+%     \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
+This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
+(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
+The dashed lines indicated that this exact structure is repeated.}
+%     \label{fig:c2c3ofk33s}
+% \end{figure*}%
+
 \ClearMyMinHeight
-\SetMyMinHeight{.3}{img/epsfromtikz/c2c3_of_k33s-0}
-\SetMyMinHeight{.7}{img/epsfromtikz/c2c3_of_k33s-1}
+\SetMyMinHeight{.3}{img/svg/revised_c2c3_of_k33s}
+\SetMyMinHeight{.7}{img/svg/revised_c2c3_of_k33s_candrp}
 
@@
     \begin{subfigure}{0.3\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-0}
+        \includegraphics[height=\myMinHeight]{img/svg/revised_c2c3_of_k33s}
         \caption{}\label{fig:c2c3ofk33s:a}
@@
     \begin{subfigure}{0.7\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-1}
+        \includegraphics[height=\myMinHeight]{img/svg/revised_c2c3_of_k33s_candrp}
         \caption{}\label{fig:c2c3ofk33s:b}
@@
     %
-    \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
-This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
-(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
+    \caption{
+    (\ref{fig:c2c3ofk33s:a}) A sequence of doublets ($C_2 \times C_3$) intersecting on triangles, where the edges of the triangles are replaced by $K_{3,3}$'s.
+This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows.
+(\ref{fig:c2c3ofk33s:b}) The canonical DR-plan of $G$, drawn as a DAG.
 $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-The dashed lines indicated that this exact structure is repeated.}
+Below the third level, the obvious pattern continues until only the individual doublets are present (fourth level) with the ellipses indicating the remaining doublets between those shown.
+Decomposition of one of these doublets is shown.
+The dashed lines indicated that this exact decomposition (of the similar nodes on the level) is repeated.
+Further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges is omitted from the figure.
+    }
     \label{fig:c2c3ofk33s}
@@
 
-\myexample
-\textsl{[DR-Plan for self-similar structure]}
-This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}.
-The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$.
+\begin{example}[DR-plan for self-similar structure]
+% \myexample
+% \textsl{[DR-plan for self-similar structure]}
+This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, the canonical DR-plan of $G$.
+It begins with the whole (isostatic) graph as the root.
+The graph $G$ has only two isostatic vertex-maximal subgraphs: $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle (triangle $N$).
 These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic.
-As explained in the proof of Theorem \ref{theorem:main} since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
-Just as expected, it is on the third level, as it is a child of both of $G$'s children.
-Furthermore, it intersects trivially with the edges that connect triangle $1$ and $N$ to $2$ and $N-1$, respectively; these necessarily intersect with triangle $1$ and $N$ themselves trivially (by Theorem \ref{theorem:main}, part 1) making all of these children.
-The triangles then decompose into their constituent $K_{3,3}$ graphs, which then decompose into 9 edges.
+As explained in the proof of Theorem \ref{theorem:main}, since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
+As expected, it is on the third level, as a child of both of $G$'s children.
 
-Now, the self-similar nature of the graph is clear; $G$ without triangle $1$ and $N$ has the same structure, and thus the same pattern is repeated until only triangle $(N+1)/2$ remains in the case that $N$ is odd, or until only triangles $N/2$ and $N/2+1$ remain if $N$ is even.
+Both of $G$'s children are similar to $G$, but containing only $N-1$ triangles.
+Therefore, the canonical DR-plans of these children follow the same pattern.
+This continues downward until the individual doublets are reached (there will be multiple occurrences of the same doublets at this level, but they can be represented as the same node in a DAG).
 
-% \begin{example}
-%     In
-% \end{example}
+Further decomposition of one of these doublets is shown.
+The three edges between the triangles and the triangles themselves all intersect trivially pairwise.
+By Theorem \ref{theorem:main}, part 1, they must all be children in the DR-plan.
+Similarly, the triangles decompose into their three trivially intersecting $K_{3,3}$'s.
+Then the $K_{3,3}$ subgraphs decompose into their separate 9 edges.
+
+The self-similar nature of this graph is evident in the canonical DR-plan. Many structures are repeated throughout the DR-plan, allowing for shared computation in both decomposition and recombination.
+\end{example}
 
@@
 \begin{theorem}\label{theorem:algo_complexity}
-    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for \addin{an input graph}.
+    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for an input graph.
 \end{theorem}
@@
 
-\addin{
+% \begin{figure*}\centering%
+%   \includegraphics[width=\linewidth]{img/svg/overconstrained_choices}
+%   \caption{The first graph is over constrained by one edge (a $K_{3,3}$ with an extra vertex attached by 3 edges).
+The second graph is a minimally rigid subgraph of the first.
+Any DR-plan of this graph will have fan-in of at least 9, occurring at the node that decomposes the $K_{3,3}$.
+The third graph is a different subgraph of the first.
+A canonical and cluster minimal DR-plan of size 2 can be made for this graph (it is a Henneberg-I construction).
+Thus, we can see the first graph has both non-optimal (second graph) and optimal (third graph) canonical, cluster minimal DR-plans.}
+%   \label{fig:overconstrained_choices}
+% \end{figure*}%
 
-\subsection{Overconstrained Graphs and NP-hardness of Optimal DR-planning}
+\ClearMyMinHeight
+\SetMyMinHeight{.6}{img/svg/new_overconstrained_not_optimal}
+\SetMyMinHeight{.4}{img/svg/new_overconstrained_optimal}
+
+\begin{figure*}\centering%
+    %
+    \begin{subfigure}{0.5\linewidth}\centering
+        \scalebox{1}[.9]{
+        \includegraphics[height=\myMinHeight]{img/svg/new_overconstrained_optimal}
+        }
+        \caption{}\label{fig:overconstrained:optimal}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.5\linewidth}\centering
+        \scalebox{1}[.9]{
+        \includegraphics[height=\myMinHeight]{img/svg/new_overconstrained_not_optimal}
+        }
+        \caption{}\label{fig:overconstrained:not_optimal}
+    \end{subfigure}%
+    %
+    \caption{
+    % (\ref{fig:overconstrained:optimal}) An optimal DR-plan of a singly overconstrained rigid graph with a fan-in of 5.
+Decomposition of triangles is omitted.
+(\ref{fig:overconstrained:not_optimal}) A canonical and cluster-minimal DR-plan of the same graph.
+The DR-plan has a fan-in of 9 and is non-optimal, shown by the preceding counter-example.
+Since the nontrivial intersection of the two children of the root is underconstrained, their decompositions are separate, causing the large fan-in.
+    Both figures are canonical and cluster-minimal DR-plans of the same singly overconstrained rigid graph.
+Decomposition of triangles are omitted and dashed lines indicate a decomposition similar to the other nodes on the same level.
+(\ref{fig:overconstrained:optimal}) is an optimal DR-plan, with a fan-in of 5.
+(\ref{fig:overconstrained:not_optimal}) has a fan-in of 9 and is non-optimal, shown by the preceding counter-example.
+    }
+    \label{fig:overconstrained}
+\end{figure*}%
+
+\subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
 \label{sec:drp:overconstrained}
@@
 The problem arises when some node in a DR-plan has clusters whose pairwise intersection is non-trivial.
-In this case, an arbitrary choice of a pair of clusters as children of the node in a canonical DR-plan may not result in an optimal DR-plan if the graph is overconstrained.
-\todo{rephrase?
-Theorem 4 shows this.
-If it's independent it \vemph{will} be optimal (the church rosser property), no longer holds if it's overconstrained.}
+In this case, an arbitrary choice of a pair of clusters as children of an overconstrained node in a canonical DR-plan may not result in an optimal DR-plan.
+This is in contrast to independent graphs, which, as shown in Theorem~\ref{theorem:main}, exhibit the strong Church-Rosser property that any choice yields an optimal DR-plan.
+% \todo{rephrase? Theorem 4 shows this. If it's independent it \vemph{will} be optimal (the church rosser property), no longer holds if it's overconstrained.}
 A good source of examples of overconstrained graphs with canonical DR-plans that are not optimal are graphs whose cluster-minimal DR-plans that are not optimal.
-The example shown in Figure \ref{fig:canonicalnotoptimal} is a canonical, cluster-minimal DR-plan that is not optimal; an optimal DR-plan is also shown in the figure.
+The example shown in Figure \ref{fig:overconstrained} is a canonical, cluster-minimal DR-plan that is not optimal; an optimal DR-plan is also shown in the figure.
+The root cause of the NP-hardness is encapsulated in this figure: because the different choices of vertex-maximal subgraphs for overconstrained input do not incur the same fan-in, finding the optimal DR-plan becomes a search problem with a combinatorial explosion of options.
 
@@
 Similarly, the algorithm given above finds a canonical DR-plan also for any input graph.
-However neither of these DR-plans may be optimal for overconstrained graphs as shown in Figure \ref{fig:canonicalnotoptimal}.
+However neither of these DR-plans may be optimal for overconstrained graphs as shown in Figure \ref{fig:overconstrained}.
 
@@
 This exponential growth of time complexity for overconstrained graphs is in fact captured in the proof of NP-hardness of optimal DR-planning in \cite{sitharam2005combinatorial, lomonosov2004graph}.
-}

### 2015-05-17 doc/GMP2015/revision_2/s04_recombination.tex
Revision 2, finished.

@@
 
-\subsubsection{Optimal Recombination and Solution Space Navigation.}
+\subsubsection{Optimal Recombination and Solution Space Navigation}
 For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan.
@@
 
-\subsubsection{Configuration Spaces of Underconstrained Systems.}
+\subsubsection{Configuration Spaces of Underconstrained Systems}
 For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems
@@
 This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$.
-Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces for some parameters $F$} (short: $H$ is \dfn{convexifiable} \todo{This shortening doesn't exactly make sense}).
+Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces for some parameters $F$} (short: $H$ is \dfn{convexifiable}).
 
@@
 
-\begin{theorem*}\label{theorem:criterionc}
-    [Efficient search]
+\begin{theorem*}[Efficient search]
+\label{theorem:criterionc}
+    % [Efficient search]
     For an indecomposable, $k$-approximately convexifiable graph $G = (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph where $|D| \le  k$.
@@
 (\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree.
-(\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$.
+(\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least OMD$_2$.
 (\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line).
-(\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
+(\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is OMD$_1$.
     }
@@
 
-\myexample
-[Using Cayley configuration space].
+\begin{example}[Using Cayley configuration space]
+% \myexample
+% [Using Cayley configuration space].
 A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e.\ its DR-plan has a root and 9 children corresponding to the 9 edges.
@@
 In the next Section \ref{sec:tdecomp}, we discuss this issue of why the largest class of reduced graphs is desirable.
+\end{example}
 
-\subsection{Optimized Modification by Enlarging the class of Reduced
+\subsection{Optimized Modification by Enlarging the Class of Reduced
 Graphs}

### 2015-05-17 doc/GMP2015/revision_2/s05_problems.tex
Revision 2, finished.

@@
 
-\subsubsection{Special classes of small DR-plans}
+\subsubsection{Special Classes of Small DR-Plans}
 As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially by solving a single quadratic at a time.
@@
 %
-\subsubsection{Optimal Modification, Completion and Recombination: previous work and formal
-connections}
+\subsubsection{Optimal Modification, Completion and Recombination: Previous Work and Formal Connections}
 %

### 2015-05-17 doc/GMP2015/revision_2/s07_pinnedline.tex
Revision 2, finished.

@@
-\section{Application: Finding optimal DR-plans and realizations for cross-linking microfibrils }
+\section{Application: Finding Optimal DR-Plans and Realizations for Cross-Linking Microfibrils }
 \label{sec:pinnedline}
@@
 
-\subsection{Modeling the fibrils as a pinned line-incidence system}
+\subsection{Modeling the Fibrils as a Pinned Line-Incidence System}
 
@@
 
-\subsection{Optimal DR-plan for pinned line-incidence systems}
+\subsection{Optimal DR-Plan for Pinned Line-Incidence Systems}
 

### 2015-07-19 doc/GMP2015/revision_3/appendix_proofs.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore

--- 2015-05-16 doc/GMP2015/revision_2/appendix_proofs.tex
@@
 For (3), observe that, given an isostatic graph, a subgraph with density $3$ must also be isostatic.
-Then, use the fact that, by definition, density of $F_i$ denoted $d(F_i)=3$ and $d(F_j)=3$.
+Then, use the fact that, by definition, $d(F_i)=3$ and $d(F_j)=3$.
 Then it is straightforward application of the inclusion-exclusion $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
@@
     The Cartesian realization space of $(H,\left<\delta_{E'}, \lambda_F\right>)$ is computed easily with a DR-plan of size 2, and is the union of $2^t$ solutions (modulo orientation preserving isometries) each with a distinct orientation type, where $t$ is the number of triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction of the length vector $\delta$ to the edges in $S$.
-A desired solution $p$ (or connected component of a solution space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a subdivided binary search   of the Cartesian realization space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation type $\sigma_p$ - as $\lambda_F$ ranges over the discretized convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in Theorem \ref{theorem:convexcayley}.
+A desired solution $p$ (or connected component of a solution space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a subdivided binary search of the Cartesian realization space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation type $\sigma_p$, as $\lambda_F$ ranges over the discretized convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in Theorem \ref{theorem:convexcayley}.
 A solution $p$ is found  when the lengths for nonedges in $D$ match $\delta_D$.
@@
 Sharing a pin is not considered an intersection.
-Such a structure is viewed as two subgraphs each sharing 1 body with a 3rd 1-dof subgraph which essentially just consists of those two bodies pinned together.
+Such a structure is viewed as two subgraphs each sharing 1 body with a third 1-dof subgraph which essentially just consists of those two bodies pinned together.
 \end{proof}
@@
 
-    Thus, starting with a $T$ as described in the proof for Remark \ref{thm:1dofcase}: when a node $b$ is pinned, the same structure is preserved as before.
+    Thus, starting with a $T$ as described in the proof for Remark \ref{thm:1dofcase}, when a node $b$ is pinned, the same structure is preserved as before.
 Suppose  $v$ is an isostatic node after pinning $b$.
@@
 
-The only remaining case is pinning two nodes that are incomparable, i.e., do not have a descendant/ancestor relationship.
+The only remaining case is pinning two nodes that are incomparable, i.e.\ do not have a descendant/ancestor relationship.
 The only change from the previous case is that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.

### 2015-07-19 doc/GMP2015/revision_3/s03_drp.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore

--- 2015-05-17 doc/GMP2015/revision_2/s03_drp.tex
@@
 % \begin{subfigure}{.3\linewidth}\centering
-%     \input{img/tikz/c2c3_of_k33s-0.tikz}
+%     \input{../../img/tikz/c2c3_of_k33s-0.tikz}
 %     \caption{}\label{fig:c2c3ofk33s:a}
@@
 % \begin{subfigure}{.7\linewidth}\centering
-%     \input{img/tikz/c2c3_of_k33s-1.tikz}
+%     \input{../../img/tikz/c2c3_of_k33s-1.tikz}
 %     \caption{}\label{fig:c2c3ofk33s:b}
@@
 %     {.3}
-%     {img/epsfromtikz/c2c3_of_k33s-0}
+%     {../../img/epsfromtikz/c2c3_of_k33s-0}
 %     {}
@@
 %     {.7}
-%     {img/epsfromtikz/c2c3_of_k33s-1}
+%     {../../img/epsfromtikz/c2c3_of_k33s-1}
 %     {}
@@
 % \ClearMyMinHeight
-% \SetMyMinHeight{.3}{img/epsfromtikz/c2c3_of_k33s-0}
-% \SetMyMinHeight{.7}{img/epsfromtikz/c2c3_of_k33s-1}
+% \SetMyMinHeight{.3}{../../img/epsfromtikz/c2c3_of_k33s-0}
+% \SetMyMinHeight{.7}{../../img/epsfromtikz/c2c3_of_k33s-1}
 
@@
 %     \begin{subfigure}{0.3\linewidth}\centering
-%         \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-0}
+%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-0}
 %         \caption{}\label{fig:c2c3ofk33s:a}
@@
 %     \begin{subfigure}{0.7\linewidth}\centering
-%         \includegraphics[height=\myMinHeight]{img/epsfromtikz/c2c3_of_k33s-1}
+%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-1}
 %         \caption{}\label{fig:c2c3ofk33s:b}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.3}{img/svg/revised_c2c3_of_k33s}
-\SetMyMinHeight{.7}{img/svg/revised_c2c3_of_k33s_candrp}
+\SetMyMinHeight{.3}{../../img/svg/revised_c2c3_of_k33s}
+\SetMyMinHeight{.7}{../../img/svg/revised_c2c3_of_k33s_candrp}
 
@@
     \begin{subfigure}{0.3\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/svg/revised_c2c3_of_k33s}
+        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s}
         \caption{}\label{fig:c2c3ofk33s:a}
@@
     \begin{subfigure}{0.7\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/svg/revised_c2c3_of_k33s_candrp}
+        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s_candrp}
         \caption{}\label{fig:c2c3ofk33s:b}
@@
 % \ClearMyMinHeight
-% \SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
-% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
-% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}
+% \SetMyMinHeight{.4}{../../img/epsfromtikz/demo_graph}
+% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_comdrp}
+% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_candrp}
 
@@
 %   \begin{subfigure}{0.4\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
+%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph}
 %     \caption{}\label{fig:demo_graph:graph}
@@
 %   \begin{subfigure}{0.3\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
+%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_comdrp}
 %     \caption{}\label{fig:demo_graph:comdrp}
@@
 %   \begin{subfigure}{0.3\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
+%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_candrp}
 %     \caption{}\label{fig:demo_graph:candrp}
@@
 %   %
-%   \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
+%   \includegraphics[width=0.3\linewidth]{../../img/epsfromtikz/demo_graph_candrp_seq}
 %   \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
@@
   %
-  \includegraphics[width=0.3\linewidth]{img/svg/3xc2c3_candrp_seq}
+  \includegraphics[width=0.3\linewidth]{../../img/svg/3xc2c3_candrp_seq}
   \caption{The sequential canonical DR-plan of $G_{demo}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
@@
 % \begin{figure*}\centering%
-%   \includegraphics[width=\linewidth]{img/svg/overconstrained_choices}
+%   \includegraphics[width=\linewidth]{../../img/svg/overconstrained_choices}
 %   \caption{The first graph is over constrained by one edge (a $K_{3,3}$ with an extra vertex attached by 3 edges).
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.6}{img/svg/new_overconstrained_not_optimal}
-\SetMyMinHeight{.4}{img/svg/new_overconstrained_optimal}
+\SetMyMinHeight{.6}{../../img/svg/new_overconstrained_not_optimal}
+\SetMyMinHeight{.4}{../../img/svg/new_overconstrained_optimal}
 
@@
         \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{img/svg/new_overconstrained_optimal}
+        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_optimal}
         }
@@
         \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{img/svg/new_overconstrained_not_optimal}
+        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_not_optimal}
         }

### 2015-07-19 doc/GMP2015/revision_3/s04_recombination.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore

--- 2015-05-17 doc/GMP2015/revision_2/s04_recombination.tex
@@
 % \begin{subfigure}{.2\linewidth}\centering
-%     \input{img/tikz/omd_k33_example_new-0.tikz}
+%     \input{../../img/tikz/omd_k33_example_new-0.tikz}
 %     \caption{}\label{fig:omd_k33_example:a}
@@
 % \begin{subfigure}{.15\linewidth}\centering
-%     \input{img/tikz/omd_k33_example_new-1.tikz}
+%     \input{../../img/tikz/omd_k33_example_new-1.tikz}
 %     \caption{}\label{fig:omd_k33_example:b}
@@
 % \begin{subfigure}{.15\linewidth}\centering
-%     \input{img/tikz/omd_k33_example_new-2.tikz}
+%     \input{../../img/tikz/omd_k33_example_new-2.tikz}
 %     \caption{}\label{fig:omd_k33_example:c}
@@
 % \begin{subfigure}{.15\linewidth}\centering
-%     \input{img/tikz/omd_k33_example-1.tikz}
+%     \input{../../img/tikz/omd_k33_example-1.tikz}
 %     \caption{}\label{fig:omd_k33_example:d}
@@
 % \begin{subfigure}{.15\linewidth}\centering
-%     \input{img/tikz/omd_k33_example-2.tikz}
+%     \input{../../img/tikz/omd_k33_example-2.tikz}
 %     \caption{}\label{fig:omd_k33_example:e}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.2}{img/epsfromtikz/omd_k33_example_new-0}
-\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-1}
-\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-2}
-\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-1}
-\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-2}
+\SetMyMinHeight{.2}{../../img/epsfromtikz/omd_k33_example_new-0}
+\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-1}
+\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-2}
+\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-1}
+\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-2}
 
@@
     \begin{subfigure}{.2\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-0}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-0}
         \caption{}\label{fig:omd_k33_example:a}
@@
     \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-1}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-1}
         \caption{}\label{fig:omd_k33_example:b}
@@
     \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-2}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-2}
         \caption{}\label{fig:omd_k33_example:c}
@@
     \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-1}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example-1}
         \caption{}\label{fig:omd_k33_example:d}
@@
     \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-2}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example-2}
         \caption{}\label{fig:omd_k33_example:e}

### 2015-07-19 doc/GMP2015/revision_3/s08_conclusion.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore

--- 2015-05-16 doc/GMP2015/revision_2/s08_conclusion.tex
@@
 We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise indecomposable systems  at nodes of a DR-plan.
-These results rely on a recent theory of Convex Cayley Configuration spaces.
+These results rely on a recent theory of convex Cayley configuration spaces.
 Relationships and reductions between these and previously studied problems were formally clarified.
@@
 \label{conj:mfaisoptimal}
-    The Modified Frontier Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
+    The Modified \frontier\ Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
 \end{conjecture}
@@
 \label{open:sparsitymatroid}
-Although generic rigidity is a property of graphs, and moreover,  in the case of qusecs, generic rigidity has a combinatorial sparsity and tightness-based characterization, the original definition of independence in the rigidity matroid requires an algebraic notion of independence of vectors of indeterminates over $\mathbf{R}$).
+Although generic rigidity is a property of graphs, and moreover,  in the case of qusecs, generic rigidity has a combinatorial sparsity and tightness-based characterization, the original definition of independence in the rigidity matroid requires an algebraic notion of independence of vectors of indeterminates over $\RR$.
 Thus the definition of the DR-plan requires algebra over the reals.
-In fact, the recursive decomposition problem is not tied to geometric constraint graphs or
-an algebraic-geometric or mechanical notion of rigidity, and can be defined for any graph using
-the notion of an abstract rigidity matroid \cite{graver93book}.
-This is  type of matroid with two additional matroid axioms,  abstract rigidity matroids can be defined in a purely graph-theoretic manner,  with no need for algebra in their definition.
+In fact, the recursive decomposition problem is not tied to geometric constraint graphs or an algebraic-geometric or mechanical notion of rigidity, and can be defined for any graph using the notion of an abstract rigidity matroid~\cite{graver93book}.
+This is a type of matroid with two additional matroid axioms; abstract rigidity matroids can be defined in a purely graph-theoretic manner,  with no need for algebra in their definition.
 However, such abstract rigidity need not have a sparsity characterization.
 On the other hand, there are sparsity matroids that do not correspond to any notion of abstract rigidity.
-However, when an abstract rigidity matroid is also a sparsity matroid, then the techniques of this
-paper  directly apply and we can obtain purely combinatorially defined recursive decompositions of graphs.
+However, when an abstract rigidity matroid is also a sparsity matroid, then the techniques of this paper directly apply and we can obtain purely combinatorially defined recursive decompositions of graphs.
 \end{openproblem}
@@
 \begin{openproblem}
-    For fixed $k$, we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases for a somewhat different optimization of the DR-plan (Section~\ref{sec:table}).
-However, in the running time of all of these algorithms, $k$ appears in the exponent.
-Can $k$ be removed from the exponent?
+    For fixed $k$, we have
+    %
+    polynomial time optimal DR-planning (Section~\ref{sec:DRP}),
+    %
+    recombination (modification) in the presence of $k$ overconstraints,
+    %
+    optimal modification for decomposition OMD$_k(G)$ when at most $k$ constraints are removed (Section~\ref{sec:recomb}),
+    %
+    and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases for a somewhat different optimization of the DR-plan (Section~\ref{sec:table}).
+    %
+    However, in the running time of all of these algorithms, $k$ appears in the exponent. Can $k$ be removed from the exponent?
 \end{openproblem}
@@
 
-    The same holds for the restricted OMD problem where the isostatic completion is required to be a tree-decomposable graph of low Cayley complexity (i.e, have special, small DR-plans).
+    The same holds for the restricted OMD problem where the isostatic completion is required to be a tree-decomposable graph of low Cayley complexity (i.e.\ have special, small DR-plans).
 One potential obstacle to an indecomposable graph $G$'s membership in the restricted OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
@@
 \begin{openproblem}
-    Is the OMD (optimal modification for decomposition) problem reducible to the OC (Optimal completion) problem?
+    Is the OMD (optimal modification for decomposition) problem reducible to the OC (optimal completion) problem?
 \end{openproblem}
@@
     The pinned line incidence structures of Section \ref{sec:pinnedline}, for example in the case of collagen microfibrils, whose function is elastic contraction, should be considered congruent under projective transformations.
-I.e, the projective group should be factored out as a trivial motion in a new project for extending the combinatorial rigidity characterization of such systems.
+I.e.\ the projective group should be factored out as a trivial motion in a new project for extending the combinatorial rigidity characterization of such systems.
 (Currently we permit no trivial motions at all).

### 2015-07-19 doc/img/tikz/demo_graph.tikz.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore

% $
% \vcenter{\hbox{$G_{\Delta}=$}}
% \vcenter{\hbox{\input{img/tikz/demo_graph.tikz}}}
% $

\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
    \tikzstyle{e}=[]

    \node[regular polygon, regular polygon sides=3, fill=white, inner sep=4, dashed] (center) at (0,0-.125-.125) {};

    \node[c] (va) at ( 0,  0) [label={left, inner sep=.555}:$a$]{};
    \node[c] (vb) at (-.5, -.866) [label={left, inner sep=.555}:$b$]{};
    \node[c] (vc) at ( .5, -.866) [label={right,inner sep=.555}:$c$]{};
    \node[c] (vd) at (-1, -1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (ve) at ( 0, -1.732) [label={below,inner sep=.555}:$e$]{};
    \node[c] (vf) at ( 1, -1.732) [label={right,inner sep=.555}:$f$]{};

    \tedge{va}{vb}{solid}{}{};
    \tedge{vb}{vc}{solid}{}{};
    \tedge{vc}{va}{solid}{}{};

    \tedge{vb}{vd}{solid}{}{};
    \tedge{vd}{ve}{solid}{}{};
    \tedge{ve}{vb}{solid}{}{};

    \tedge{vc}{vf}{solid}{}{};
    \tedge{vf}{ve}{solid}{}{};
    \tedge{ve}{vc}{solid}{}{};
\end{tikzpicture}

### 2015-08-04 doc/GMP2015/revision_3/s02_preliminaries.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

--- 2015-05-17 doc/GMP2015/revision_2/s02_preliminaries.tex
@@
 %   %
-%   \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
+%   \includegraphics[width=0.4\linewidth]{../../img/epsfromtikz/demo_graph_clustmindrp}
 %   \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality.
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.4}{img/svg/3xc2c3}
-\SetMyMinHeight{.3}{img/svg/3xc2c3_comdrp}
-\SetMyMinHeight{.3}{img/svg/3xc2c3_candrp}
+\SetMyMinHeight{.4}{../../img/svg/3xc2c3}
+\SetMyMinHeight{.3}{../../img/svg/3xc2c3_comdrp}
+\SetMyMinHeight{.3}{../../img/svg/3xc2c3_candrp}
 
@@
   \begin{subfigure}{0.4\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3}
     \caption{}\label{fig:demo_graph:graph}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_comdrp}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_comdrp}
     \caption{}\label{fig:demo_graph:comdrp}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_candrp}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_candrp}
     \caption{}\label{fig:demo_graph:candrp}
@@
 % \ClearMyMinHeight
-% \SetMyMinHeight{.4}{img/svg/triangle}
-% \SetMyMinHeight{.3}{img/svg/triangle_candrp}
-% \SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}
+% \SetMyMinHeight{.4}{../../img/svg/triangle}
+% \SetMyMinHeight{.3}{../../img/svg/triangle_candrp}
+% \SetMyMinHeight{.3}{../../img/svg/triangle_clustmindrp}
 
@@
 %\begin{subfigure}{0.4\linewidth}\centering
-%\includegraphics[height=\myMinHeight]{img/svg/triangle}
+%\includegraphics[height=\myMinHeight]{../../img/svg/triangle}
 %    \caption{}\label{fig:demo_graph_tri:graph}
@@
 % \begin{subfigure}{0.3\linewidth}\centering
-%   \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
+%   \includegraphics[height=\myMinHeight]{../../img/svg/triangle_candrp}
 %    \caption{}\label{fig:demo_graph_tri:candrp}
@@
 %  \begin{subfigure}{0.3\linewidth}\centering
-%   \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
+%   \includegraphics[height=\myMinHeight]{../../img/svg/triangle_clustmindrp}
 %   \caption{}\label{fig:demo_graph_tri:clustmindrp}
@@
 
-\begin{definition}\label{def:drp}
-    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
-    \begin{enumerate}
-        \item Each node represents/contains/is a rigid subgraph of $G$.
-        % \item For a node, $C$, that is  a non-trivial graph, its $N$
-        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
-        % subgraphs of $C$.
-        \item The children $C_1,\ldots,C_N$ of a node $C$ satisfy $\bigcup_{i=1}^N{C_i}=C$.
-        \item A leaf node is a single edge. A \dfn{trivial} graph is empty or a single vertex. Note that a trivial graph is not isostatic.
-        \item A root node is a vertex-maximal rigid subgraph of $G$.
-    \end{enumerate}
-%
-    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
-%
-    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$.
+% \begin{definition}\label{def:drp}
+%     The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
+%     \begin{enumerate}
+%         \item Each node represents/contains/is a rigid subgraph of $G$.
+%         % \item For a node, $C$, that is  a non-trivial graph, its $N$
+%         % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
+%         % subgraphs of $C$.
+%         \item The children $C_1,\ldots,C_N$ of a node $C$ satisfy $\bigcup_{i=1}^N{C_i}=C$.
+%         \item A leaf node is a single edge. A \dfn{trivial} graph is empty or a single vertex. Note that a trivial graph is not isostatic.
+%         \item A root node is a vertex-maximal rigid subgraph of $G$.
+%     \end{enumerate}
+% %
+%     % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
+% %
+%     A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$.
 This makes Property 2 implicit.
-  %and in fact, $C$ contains all the edges in the union of its children as well.
-  We denote a complete DR-plan of $G$ as $\comdrp{G}$.
+%   %and in fact, $C$ contains all the edges in the union of its children as well.
+%   We denote a complete DR-plan of $G$ as $\comdrp{G}$.
 
-  A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
+%   A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest.
 The maximum  fan-in is called the \dfn{size} of the DR-plan.
 We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
-%
-%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
-%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
-%an isostatic vertex-maximal proper subgraph.
-    %Note that nodes will be referred to interchangeably as ``the node that
-    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
+% %
+% %In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
+% %$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
+% %an isostatic vertex-maximal proper subgraph.
+%     %Note that nodes will be referred to interchangeably as ``the node that
+%     %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
+% \end{definition}
+
+% We begin with the definition of a general DR-plan and also introduce two specific kinds of DR-plans.
+
+\begin{definition}\label{def:drp}
+    A \dfn{decomposition-recombination \mbox{(DR-)} plan}~\cite{hoffman2001decompositionI} of graph $G$ is defined as a forest that has the following properties:
+    \begin{enumerate}
+        \item Each node represents a rigid subgraph of $G$.
+        \item A root node is a vertex-maximal rigid subgraph of $G$.
+        \item A node is the subgraph of $G$ induced by the union of its children.
+        \item A leaf node is a single edge.
+    \end{enumerate}
 \end{definition}
+
+\begin{definition}
+    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$ (which makes Property 3 of a DR-plan implicit.)
+\end{definition}
+
+\begin{definition}
+    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest. The maximum fan-in is called the \dfn{size} of the DR-plan.
+\end{definition}
+
+We denote these different kinds of DR-plans as $\drp{G}$, $\comdrp{G}$, and $\optdrp{G}$, respectively.
+Note that nodes will be referred to interchangeably as ``the node that represents or contains the (sub)graph $C$'' and as simply ``$C$''.
+
 %
@@
 % \begin{definition}
-%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.
 % \end{definition}
@@
 %
-%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.
 %\end{definition}
@@
 %% \begin{definition}
-%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
+%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.
 %% \end{definition}

### 2015-08-04 doc/GMP2015/revision_3/s03_drp.tex
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

@@
 In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
-Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph.
+Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find the canonical DR-plan of any independent graph.
 The definition is as follows:
 
+% \begin{definition}\label{def:canonical_drplan}
+%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
+%     (1) it is a DR-plan of $G$;
+%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
+%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+% \end{definition}
+
 \begin{definition}\label{def:canonical_drplan}
-    The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
-    (1) it is a DR-plan of $G$;
-    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+    A \dfn{canonical DR-plan} is a DR-plan that satisfies the additional two properties:
+    \begin{enumerate}
+        \item Children are rigid vertex-maximal proper subgraphs of the parent.
+        \item If all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+    \end{enumerate}
 \end{definition}
@@
 % \end{figure*}%
+
+To find the canonical DR-plan of an isostatic graph, we first need the following subroutine.
+
+% \begin{theorem}\label{theorem:algo_find_wcvmps_complexity}
+%     There exists an $O(|V|^2)$ time complexity algorithm to find the set of isostatic vertex-maximal proper subgraphs (the \dfn{clusters}) of an isostatic input graph.
+% \end{theorem}
+
+% \begin{proof}
+%     Call the input graph $G=(V,E)$.
+%     % If the input is not isostatic (i.e.\ does not satisfy $|E| = 2|V|-3$), run the component pebble game algorithm
+%     Take an arbitrary edge $e\in E$.
+Run the component pebble game algorithm \cite{Jacobs:1997:PG} on the subgraph $G\setminus e$, which takes time $O(|V|^2)$.
+The output of this will be all of the maximal isostatic components of $G\setminus e$, which we will call the list of candidate-clusters, or $CC$.
+It will contain at most $|E|-1 = O(|V|)$ elements.
+
+%     The next step is to isolate the set of true-clusters.
+Begin by initializing the list true-clusters, or $TC$, as an empty list.
+Then, for each subgraph $C\in CC$, run the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} on $C+e$ to find $D$ (which is the minimal isostatic subgraph containing $C+e$.) In case 1, where $D=G$, $C$ is removed from $CC$ and added to $TC$.
+In case 2, where $D\subsetneq G$, for all $H\in CC$ that are subgraphs of $D$, $H$ is removed from $CC$; then $D$ is added to $TC$.
+If every element in $CC$ results in case 1, add $e$ to $TC$.
+Thus, we do $O(|V|)$ work on $O(|V|)$ subgraphs for a total time of $O(|V|^2)$.
+% \end{proof}
+
+\begin{theorem}\label{theorem:algo_find_wcvmps_complexity}
+    There exists an \complexityAllClustersV\ time complexity algorithm to find the set of isostatic vertex-maximal proper subgraphs (the \dfn{clusters}) of an isostatic input graph.
+\end{theorem}
+
+\begin{proof}
+    Call the input graph $G=(V,E)$.
+Initialize the empty list of current-clusters, or $CC$.
+For each edge $e\in E$, run the component pebble game algorithm \cite{Jacobs:1997:PG} on the subgraph $G\setminus e$.
+The output of this will be all of the maximal isostatic components of $G\setminus e$, which we will call potential-clusters, or $PC$.
+Remove any element from $PC$ that is a subgraph of an element in $CC$, then remove any element from $CC$ that is a subgraph of an element in $PC$.
+Add all remaining elements of $PC$ to $CC$.
+Return $CC$ after running all edges.
+
+    The time complexity is $O(|V|^3)$.
+For each edge, of which there are $O(|V|)$, we run the component pebble game, which takes $O(|V|^2)$, and then check its $PC$ against the $CC$, which also takes $O(|V|^2)$.
+This subgraph comparison can be done by labeling each edge in $E$ with a pointer to the element that contains it in $PC$ and $CC$ while running the component pebble games.
+Then, traverse edge set $E$ and check if the corresponding elements of $PC$ and $CC$ are subgraphs.
+\end{proof}
+
 \begin{figure*}\centering%
@@
 \begin{theorem}\label{theorem:algo_complexity}
-    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for an input graph.
+    There exists an \ComplexityCanDRPV\ algorithm to find a canonical DR-plan for an independent input graph.
 \end{theorem}
@@
 \begin{proof}
-We first describe the algorithm.
-The first step of the algorithm, which we call $CanDRP(G)$, is finding the isostatic vertex-maximal proper subgraphs (clusters) of the input isostatic graph $G$.
-One of many ways to do this is  by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of cluster-candidates.
-The next step is to isolate the set of true-clusters by running the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, to find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, the candidate is moved (before adding $e$) to the true-cluster list and removed from the candidate-cluster list.
-If $D\subsetneq G$, all candidates that are subgraphs of $D$ are removed from the candidate-cluster list and $D$ is added back to the list.
-The next candidate is considered until all candidates are exhausted.
+    Call the input graph $G=(V,E)$. We will call the result of the algorithm $CanDRP(G)$.
 
-The next step is to check the intersection of any two of the true-components.
-If the intersection is trivial, the entire list becomes the children, and the algorithm is recursively applied to each child.
-If the intersection is non-trivial (necessarily isostatic), the  decomposition is computed down to $I=\bigcap_{k=1}^{N}{C_k}$.
-(This is discussed in detail in the proof of Theorem \ref{theorem:main}).
-We construct a DR-plan whose node set is that of a canonical/optimal DR-plan.
-Begin with the parent $C$.
-Its children are $S_1$ and $CanDRP(T_1)$, where $S_1=C\setminus R_1$ and $T_1$ is $R_1$, together with all incident edges.
-Observe that these edges are incident on vertices from $I$ and that $T_1$ is underconstrained so its DR-plan is a forest where each root becomes a child of $C$.
-The children of $S_1$ are $S_2$ and $CanDRP(T_2)$, where $S_2=C\setminus (R_1\cup R_2)$ and $T_2$ is $R_2$ together with all incident edges.
-The process iterates.
-At $N$ levels down, the children are  $S_N=I$ and $CanDRP(T_N)$, where $T_N$ is $R_N$ together with all incident edges.
-Apply the algorithm recursively to both children.
-Thus we obtain an efficient method for the case of DR-plan nodes whose rigid proper vertex-maximal subgraphs have non-trivial intersections, avoiding recomputation of the isostatic components of $C$.
+    If $G$ is a single edge, return $G$.
+
+    % Either $G$ is isostatic (i.e.\ satisfies $|E|=2|V|-3$)
+    If $G$ is not isostatic (i.e.\ does not satisfy $|E|=2|V|-3$), run the component pebble game on $G$ to get the set of isostatic vertex-maximal subgraphs.
+Return the forest formed by recursively running $CanDRP$ on each element.
+
+    If $G$ is isostatic, run the isostatic vertex-maximal proper subgraph detection algorithm (Theorem \ref{theorem:algo_find_wcvmps_complexity}) on $G$ to get what we call the list of clusters.
+This can be done in \complexityAllClustersV\ time.
+Choose two arbitrary clusters, called $C_1$ and $C_2$.
+If $C_1\cap C_2$ is trivial (a single vertex), recursively apply $CanDRP$ on all clusters and set the results as the children of $G$; return $G$.
+Else if $C_1\cap C_2$ is non-trivial (an isostatic subgraph), apply $CanDRP$ to $C_1$ (note that this choice is still arbitrary, any cluster will suffice) and $T$ (the underconstrained graph formed by $C\setminus C_1$ together with all incident edges and the associated vertices in $C$) and set the results as the children of $G$; return $G$.
+
+    For the sake of increased clarity (and performance in implementation), we expand on the case of isostatic input with clusters that intersect non-trivially.
+Instead of immediate recursive application of $CanDRP$, we propose an equivalent method to first compute the decomposition down to $I=\bigcap_{k=1}^{N}{C_k}$.
+(The rationale and nomenclature is discussed in detail in the proof of Theorem \ref{theorem:main}.) Let $R_i= C\setminus C_i$; note that each $R_i$ is pairwise disjoint.
+Let $T_i$ be $R_i$ plus all incident edges; note that these edges are incident on vertices in $I$ and that $T_i$ is underconstrained.
+Let $S_i = C\setminus \bigcup_{j=1}^{i}{R_j}$.
+Set the children of $G$ as $S_1=C\setminus R_1=C_1$ and $CanDRP(T_1)$.
+Set the children of $S_1$ as $S_2=C\setminus (R_1\cup R_2)$ and $CanDRP(T_2)$.
+Etc.
+At step $N$, set the children of $S_{N-1}$ as $CanDRP(S_N)$ and $CanDRP(T_N)$.
+Thus we can avoid the $N$ recomputations of the isostatic components of $G$.
+
+    We call this intermediate result of the algorithm the ``sequential'' canonical DR-plan.
 See Figure~\ref{fig:demo_graph:candrpseq} for an example.
+This plan is a tree, where each node is a distinct subgraph, each is smaller than its parent, and the leaves comprise the edge set of $G$.
+Therefore, the number of nodes in the tree is $O(|E|)$ which, for independent input graphs, is $O(|V|)$.
+The cluster finding algorithm is run on each internal node, resulting in an overall time complexity of \ComplexityCanDRPV.
 
-To summarize the above step of the algorithm:
-if the intersection is trivial, recursively apply $CanDRP$ to the entire list of true-components and set the resulting trees as the children.
-Else, the intersection is non-trivial, and the children of the node are $CanDRP(C_1)$ and $CanDRP(T)$, where $C_1$ is the first true-component of the parent $C$ (chosen arbitrarily, could be any child) and $T$ is the underconstrained graph formed by $C\setminus C_1$ together with all incident edges  and the associated nodes in $C_1$.
-
-% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
-
-The DR-plan  output by this algorithm is a tree where each node is a distinct subgraph.
-The leaves of this tree are the edges of the starting graph $G$.
-Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$ and the complexity of the algorithm is $O(|V|^3)$.
+    The sequential canonical DR-plan can then be used to find the canonical DR-plan in
+    % sub-cubic
+    sub-quartic
+    time.
+Simply alter any subtree that is the result of non-trivial intersection in the obvious way (adding nodes for the clusters implied by the decomposition) to get the canonical DAG.
+Although, in practice, the sequential plan may be sufficient.
     %
@@
 \end{proof}
+
+% Note that in the case of each stage of decomposition resulting in trivial intersections, this algorithm can be run in time $O(|V|^3)$.
+We can find all clusters in time $O(|V|^2)$, because we only have to look at a constant number of edges.
+Take two arbitrary edges to get the two sets of potential-clusters, $PC_1$ and $PC_2$.
+
+% \begin{proof}
+%     Choose an arbitrary ordering of the edges, such that $e_i$ is the $i^\text{th}$ edge in the sequence.
+
+%     Run component pebble game on each edge to get the map $M$ that takes an edge, $e$, and returns a set of subgraphs that are the output of the component pebble game on $G\setminus e$.
+On each run of the pebble game, also attach a pointer to each edge that points to the subgraph containing the edge.
+%     Call this map $P$, where $P(i, j)$ (which is not defined where $i=j$) returns a pointer to the element containing $e_i$ from $M(e_j)$.
+%     This is $O(|V|^3)$.
+
+%     Dropping an edge will never result in some other edge being in some new cluster, the other edge will only ever be in a subset of its true cluster.
+Therefore, we can look at each edge in $E$ and find the proper maximal subgraph containing it by following all the cluster pointers and comparing the cardinality of the vertex sets, taking the largest.
+This can be done in $O(|V|^2)$.
+Save any subgraph found in this manner.
+%     HAVE TO CHECK FOR DUPLICATES... CAN BE ROLLED INTO NEXT STEP, DIFFERENT BASED ON INTERSECTION TYPE. JUST DO AN EDGE FROM THE COMPLEMENT ($G\setminus C$) SECOND.
+
+%     Now check the intersections between clusters (pick an arbitrary two).
+
+%     In the case of trivial intersections, all of the found clusters will be edge-disjoint.
+%     For each cluster $C$, take the submatrix of $P$ only on the intersections of elements in $C$.
+That is, if $C$ contains $i$ and $j$, keep element $P(i,j)$.
+Delete any element in $M$ whose pointer is lost from $P$ after this operation.
+%     % For each edge $e_i$, for every edge $e_j$ not in the cluster to which $e_i$ belongs, delete $P(i, j)$ and the element it points to in $M(e_j)$ (if it isn't already deleted).
+%     This can be done in $O(|V|^2)$ time.
+%     The result will be a new $P$ and $M$ for each cluster, containing only the edges in the cluster. Furthermore, all clusters from this stage will be removed from $M(e_i)$ (for all $i$).
+%     % Divide the edges into sets based on their containing cluster (one edge is contained in only one cluster) and repeat the previous step.
+%     Recursively run the algorithm on these clusters, but we can now skip the first step, avoiding $O(|V|^3)$ work at each node...
+So with an initial step of $O(|V|^3)$ and then $O(|V|^2)$ work at $O(|V|)$ nodes we get an overall complexity of $O(|V|^3)$.
+
+%     In the case of isostatic intersections, you will find 2 clusters (the one with the most vertices, $C_1$, and the largest one containing the vertex set omitted from the first, $C_2$).
+You will also need to be careful of different clusters with the same size (just be aware it might happen, shouldn't affect anything, we'll want them later).
+Go to an arbitrary edge not in $C_1$.
+Look at the clusters containing the edge with the largest cardinality.
+The largest will be $C_2$, ignore it.
+Keep going to the next largest, adding it to the list of found clusters if it has an isostatic intersection with the $C_1$.
+This will take $O(|V|^2)$.
+Now we have all of the isostatic vertex-maximal proper subgraphs, which we call $C_1, C_2, \ldots , C_N$.
+Compute the intersection of all of these, which we call the ``core'' or $I$.
+Let $R_i = G\setminus C_i$.
+
+%     Choose an arbitrary edge $e\in I$.
+Take the set $M(e)$, remove all elements that are subgraphs of $I$.
+Of the remainder, the elements that are subgraphs of $R_i$ are the first level of decomposition of $R_i$ and we call this subset $T_i$.
+Now we are ready to build the ``sequential'' canonical DR-plan of $G$ down to node $I$.
+
+%     Set the children of $G$ as $S_1=C\setminus R_1=C_1$ and the elements of $T_1$.
+Set the children of $S_1$ as $S_2=C\setminus (R_1\cup R_2)$ and the elements of $T_2$.
+Etc.
+At step $N$, set the children of $S_{N-1}$ as $S_N=I$ and the elements of $T_N$.
+
+%     For node $I$ and each element of $T_1, T_2, \ldots , T_N$ in the DR-plan, get the submatrix of $P$ using the edges of the subgraph in the node, and update $M$ accordingly.
+%     % For each edge $e_i$ in the core, for every edge $e_j$ not in the core, delete $P(e_i, j)$ and the element it points to in $M(e_j)$ (if it isn't already deleted).
+
+%     The sequential canonical DR-plan can then be used to find the canonical DR-plan in
+%     sub-cubic
+%     % sub-quartic
+%     time.
+Simply alter any subtree that is the result of non-trivial intersection in the obvious way (adding nodes for the clusters implied by the decomposition) to get the canonical DAG.
+Although, in practice, the sequential plan may be sufficient.
+% \end{proof}
+
+First, we remind the reader of the component pebble game~\cite{Jacobs:1997:PG}.
+Given an independent input graph $G$ with vertex weight 2 and edge weight 1, the output will be all of the 2D maximal isostatic subgraphs.
+We will denote this algorithm with $COMPONENTS(G)$.
+For independent input, this runs in time $O(|V|^2)$.
+
+We also introduce the term \dfn{clusters} of graph $G$ to refer to the set of isostatic vertex-maximal proper subgraphs of $G$.
+
+Next, we define an operation called \dfn{modularization}.
+The function $MZ(M, I)$ takes a square matrix $M$ and a set of indices $I$ and returns $(M', A)$.
+$M'$ is the square matrix formed by keeping all elements $M_{ij}$ if $i,j\in I$.
+$A$ is the set of elements $M_{ij}$ and $M_{ji}$ where $i\in I$ and $j\notin I$.
+% This operation takes $O(|I|^2)$.
+
+% Next, we make the observation that dropping an edge will only cause other edges in the graph to be in components (as found by pebble games) that are subgraphs of their cluster ()
+
+% Consider dropping edge $e_i$ and looking at the effect on $e_j$.
+In the case that all clusters intersect trivially, no edge is in multiple clusters.
+Therefore, $e_i$ could come from a different cluster than $e_j$, in which case dropping $e_i$ only affects its own cluster.
+Or $e_i$ could come from the same cluster as $e_j$, in which case all other clusters are found as components, but additionally edge-disjoint subgraphs of this cluster will be found as components.
+% In the case that clusters intersect non-trivially, edges do com from multiple clusters.
+
+\begin{proof}
+    %% INITIALIZATION
+    Let the input be $G=(E,V)$.
+If the input is an edge, terminate.
+If the input is not isostatic, set the roots of the forest as $COMPONENTS(G)$ then recursively call the algorithm on each root.
+Otherwise, do the following.
+
+    Choose an arbitrary ordering of the edges, such that $e_i$ is the $i^\text{th}$ edge in the sequence.
+    Initialize a map $M$ and a map $P$.
+For each $i$, set $M(e_i)=COMPONENTS(G\setminus e_i)$ and, for all $j\neq i$, set $P(i,j)$ as a pointer to the element in $M(e_i)$ whose subgraph contains $e_j$.
+This initialization takes time $O(|V|^3)$.
+
+    %% STEP 1: Find the clusters
+    % Choose some $i\leq |E|$, then find the
+    Now we want to find the clusters of $G$.
+    For each $i$, if we have not already found a cluster containing $e_i$, then find the $j$ such that $P(j,i)$ points to the component with the largest cardinality and add it to the list of clusters.
+This takes time $O(|V|^2)$.
+In the case that the clusters intersect trivially, we will find all of the clusters.
+In the case that they intersect non-trivially, we will find exactly two clusters.
+We will return to the latter case after continuing the proof for the former.
+
+    %% STEP 2a: Trivial intersections
+    In the case of trivial intersections, all of the found clusters will be edge-disjoint.
+For each cluster edge set $C_i$, compute $MZ(P, C_i)=(P_i, A_i)$.
+For element in each $A_i$, delete the component it points to in $M$.
+This can be done in $O(|V|^2)$ time.
+    Note that, for any $i$, $M(e_i)$ contained only subgraphs of the cluster containing $e_i$ and the other clusters.
+By doing the deletions, we have removed all of the clusters found in this stage.
+This means $M$ can now be broken into an $M_i$ for each $C_i$, where $M_i$ contains only proper subgraphs of $C_i$.
+Now the algorithm can be run recursively on each $C_i$, while avoiding the recomputation of $M$ and $P$ by using the $M_i$ and $P_i$ we found.
+
+    %% STEP 2b: Non-trivial intersections
+    In the case of non-trivial intersections, we have only found 2 clusters (the one with the most vertices, $C_1$, and the largest one containing the vertex set omitted from the first, $C_2$) and we need to find the remainder.
+Go to an arbitrary edge $e_i\notin C_1$.
+For all $j\neq i$, if the component pointed to by $P(j,i)$ intersects non-trivially with $C_1$ add it to the list of clusters (ignore duplicates).
+This will take $O(|V|^2)$.
+Now we have all of the isostatic vertex-maximal proper subgraphs, which we call $C_1, C_2, \ldots , C_N$.
+Let $I=\bigcap_{k=1}^{N}{C_k}$.
+Let $R_i = G\setminus C_i$.
+    Let $T_i = COMPONENTS(R_i)$. All of these can be computed in $O(|V|^2)$.
+
+    % Choose an arbitrary edge $e\in I$.
+Take the set $A=M(e)$, remove all elements from $A$ that are subgraphs of $I$.
+For each $R_i$, find the set $T_i$ which contains the elements of $A$ that are subgraphs of $R_i$.
+Note that $T_i=COMPONENTS(R_i)$, and will be the first level of decomposition of $R_i$ since it is underconstrained.
+    % COULD WE JUST RUN $COMPONENTS$???
+
+    Now we build the ``sequential'' canonical DR-plan of $G$ down to node $I$.
+    Set the children of $G$ as $S_1=C\setminus R_1=C_1$ and the elements of $T_1$.
+Set the children of $S_1$ as $S_2=C\setminus (R_1\cup R_2)$ and the elements of $T_2$.
+Etc.
+At step $N$, set the children of $S_{N-1}$ as $S_N=I$ and the elements of $T_N$.
+    For each node $N$ of the DR-plan in the set $\{I\}\cup T_1\cup T_2\cup \cdots\cup T_N$, compute $MZ(P, N)=(P', A')$ and update $M$ by deleting all elements pointed to by $A'$.
+Then, for each $N$, run the algorithm recursively on (avoiding recomputation of $M$ and $P$).
+    % For node $I$ and each element of $T_1, T_2, \ldots , T_N$ in the DR-plan, compute $MZ(P, \cdot)$, update $M$ in the same fashion as above, and run the algorithm recursively on $(\cdot)$ (avoiding recomputation of $M$ and $P$).
+
+    %% Proof of complexity
+    The decomposition will ultimately terminate as a tree with $|E|$ leaves, each containing a unique edge.
+With an initial step of $O(|V|^3)$ and then $O(|V|^2)$ work at $O(|V|)$ nodes we get an overall complexity of $O(|V|^3)$.
+
+    The sequential canonical DR-plan can then be used to find the canonical DR-plan in
+    sub-cubic
+    % sub-quartic
+    time.
+Simply alter any subtree that is the result of non-trivial intersection in the obvious way (adding nodes for the clusters implied by the decomposition) to get the canonical DAG.
+Although, in practice, the sequential plan may be sufficient.
+\end{proof}
+
+% In the case of trivial intersections, all of the found clusters will be unique.
+If the cardinality is greater than 2, we can save computation by not checking the other edges in the found cluster.
+But we can only know that they are trivial intersections after finding 2 different clusters.
+So do that first...
+
+% In the case of isostatic intersections, if we begin with an edge in the core, there are several clusters containing this edge (and the algo will pick the largest among them).
+We could only safely skip the other edges in the core, but we don't know what it is yet.
+So after finding the first cluster, go to an edge not in it and find the next cluster.
+Check the intersection.
+Now we can skip the others in the core.
+WAIT....
+THIS WILL ONLY PICK THE LARGEST, AND ITS ``COMPLEMENT'', SO TO SPEAK.
+WHICH TECHNICALLY WORKS....
+
+% For each found cluster $c$, for each edge $e$ \emph{not in} $c$, delete $c$ from $M(e)$ and delete the pointer to $M(e)$ from the pointer list of $e$.
 
@@
 Then run the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
-If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
+If $D=G$, move $D$ to the true-cluster list and remove the candidate from the candidate-component list.
+If $D\subsetneq G$, check the union of $D$ with the all items in the true-cluster list; if the union is isostatic, halt and output just these two components as children.
 Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$.
 Continue with the next candidate.
-If you exhaust the candidates, output the entire true-component list.
+If you exhaust the candidates, output the entire true-cluster list.
 

### 2015-08-04 doc/GMP2015/revision_3/s06_problempinning.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

--- 2015-05-16 doc/GMP2015/revision_2/s06_problempinning.tex
@@
-\section{Application: Finding completions of underconstrained glassy structures from underconstrained to isostatic}
+\section{Application: Finding Completions of Underconstrained Glassy Structures from Underconstrained to Isostatic}
 \label{sec:bodypin}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.3}{img/silica}
-\SetMyMinHeight{.2}{img/Silicon_tetrahedron}
+\SetMyMinHeight{.3}{../../img/silica}
+\SetMyMinHeight{.2}{../../img/Silicon_tetrahedron}
 
 \begin{figure}\centering
-    \includegraphics[height=\myMinHeight]{img/Silica}
+    \includegraphics[height=\myMinHeight]{../../img/Silica}
     \hspace{0.5cm}
-    \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
+    \includegraphics[height=\myMinHeight]{../../img/Silicon_tetrahedron}
     % \raisebox{0.25\height}{
-    %     \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
+    %     \includegraphics[height=\myMinHeight]{../../img/Silicon_tetrahedron}
     % }
@@
 In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress, flex and other property design related to the rigidity matrix.
-To answer these questions, we first introduce the qusecs' that are used to model these materials.
+To answer these questions, we first introduce the qusecs that are used to model these materials.
 In this section, we discuss Item (2) in detail.
@@
 %     \begin{subfigure}{0.2\linewidth}
-%         \includegraphics[width=\linewidth]{img/pentl1}
+%         \includegraphics[width=\linewidth]{../../img/pentl1}
 %     \end{subfigure}
@@
 %     \begin{subfigure}{0.2\linewidth}
-%         \includegraphics[width=\linewidth]{img/pentl2}
+%         \includegraphics[width=\linewidth]{../../img/pentl2}
 %     \end{subfigure}
@@
 %     \begin{subfigure}{0.8\linewidth}
-%         \includegraphics[width=\linewidth]{img/pentawesome}
+%         \includegraphics[width=\linewidth]{../../img/pentawesome}
 %     \end{subfigure}
@@
         $
-        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{img/pentl1}}}
+        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{../../img/pentl1}}}
         % \raisebox{-.5\height}{\scalebox{1}{$\rightarrow$}}
         \vcenter{\hbox{$\rightarrow$}}
-        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{img/pentl2}}}
+        \vcenter{\hbox{\includegraphics[width=.4\linewidth]{../../img/pentl2}}}
         $
@@
     \begin{subfigure}{0.55\linewidth}
-        \includegraphics[width=\linewidth]{img/pentawesome}
+        \includegraphics[width=\linewidth]{../../img/pentawesome}
         \caption{}\label{fig:subdivision:complicated}
@@
 
-\subsection{Body-Hyperpin qusecs}
+\subsection{Body-Hyperpin Qusecs}
+
+% \begin{definition}
+%     A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies, subsets of which are pinned together by pins (i.e.\ they are incident at a common point.)
+% \end{definition}
 
 \begin{definition}
-    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies,  subsets of which are pinned together by pins; i.e, are incident at a common point.
+    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies and the constraints are incidences of object subsets at a common point, i.e.\ a pinning of bodies.
 \end{definition}
@@
 \begin{definition}
-    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares  pins with at most 3 other bodies.
+    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e.\ it shares  pins with at most 3 other bodies.
 This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge.
@@
 \begin{definition}
-    A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial.
+    A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
 In our case, the trivial nodes are the bodies.
@@
 Provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}),
-the notion of a canonical DR-plan extends directly to the case when the hypergraph is $(k,l)$-sparse (i.e., independent) using the straightforward notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions as in Section \ref{sec:DRP}.
+the notion of a canonical DR-plan extends directly to the case when the hypergraph is $(k,l)$-sparse (i.e.\ independent) using the straightforward notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions as in Section \ref{sec:DRP}.
 In particular, we define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.
@@
 \label{obs:bodypin_drp}
-For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex-maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
+For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex-maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof subgraphs, pairwise sharing at most one body.
 \end{observation*}
@@
 %     (a) the children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph, or
-%     (b) the children are a number of proper maximal 1-dof sub-graphs, sharing at most one body.
+%     (b) the children are a number of proper maximal 1-dof subgraphs, sharing at most one body.
 % \end{observation}
@@
 
-%     For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$.
+%     For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof subgraphs $s_1, ..., s_k, k > 2$.
 Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.2}{img/bodypin}
-\SetMyMinHeight{.2}{img/epsfromtikz/bodypin_graph}
-\SetMyMinHeight{.35}{img/epsfromtikz/bodypin_drp}
-\SetMyMinHeight{.2}{img/bodypin2}
+\SetMyMinHeight{.2}{../../img/bodypin}
+\SetMyMinHeight{.2}{../../img/epsfromtikz/bodypin_graph}
+\SetMyMinHeight{.35}{../../img/epsfromtikz/bodypin_drp}
+\SetMyMinHeight{.2}{../../img/bodypin2}
 
@@
     \begin{subfigure}{0.2\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/bodypin}
+        \includegraphics[height=\myMinHeight]{../../img/bodypin}
         \caption{}\label{fig:bodypindrp:a}
@@
     \begin{subfigure}{0.2\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/bodypin_graph}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/bodypin_graph}
         \caption{}\label{fig:bodypindrp:b}
@@
     \begin{subfigure}{0.35\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/epsfromtikz/bodypin_drp}
+        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/bodypin_drp}
         \caption{}\label{fig:bodypindrp:c}
@@
     \begin{subfigure}{0.2\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{img/bodypin2}
+        \includegraphics[height=\myMinHeight]{../../img/bodypin2}
         \caption{}\label{fig:bodypindrp:d}
@@
 %     Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$.
-Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.
+Each node of $T$ will then be a vertex maximal proper 1-dof subgraph of $G$.
 
@@
 We choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
-From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a sub-graph.
+From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a subgraph.
 We pin the new body to bodies in two separate children.
@@
 \label{obs:algebraic_completion}
-    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer be optimized by the algorithm given in that proof.
+    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer be optimized by the algorithm given in that proof.
 In particular, the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if the case were matroidal and an optimal DR-plan is known).

### 2015-08-04 doc/GMP2015/revision_3/s07_pinnedline.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

--- 2015-05-17 doc/GMP2015/revision_2/s07_pinnedline.tex
@@
 %   \centering
-%   \includegraphics[width=.8\linewidth]{img/AFM_Innventia_nanocellulose}
+%   \includegraphics[width=.8\linewidth]{../../img/AFM_Innventia_nanocellulose}
 %   \caption{}
@@
 %   \centering
-%   \includegraphics[width=\linewidth]{img/crosslink}
+%   \includegraphics[width=\linewidth]{../../img/crosslink}
 %   \caption{}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.32}{img/AFM_Innventia_nanocellulose}
-\SetMyMinHeight{.32}{img/crosslink}
-\SetMyMinHeight{.32}{img/pinned}
+\SetMyMinHeight{.32}{../../img/AFM_Innventia_nanocellulose}
+\SetMyMinHeight{.32}{../../img/crosslink}
+\SetMyMinHeight{.32}{../../img/pinned}
 
@@
 \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/AFM_Innventia_nanocellulose}
+    \includegraphics[height=\myMinHeight]{../../img/AFM_Innventia_nanocellulose}
     \caption{}\label{fig:cellulose}
@@
 \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/crosslink}
+    \includegraphics[height=\myMinHeight]{../../img/crosslink}
     \caption{}\label{fig:crosslink}
@@
 \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/pinned}
+    \includegraphics[height=\myMinHeight]{../../img/pinned}
     \caption{}
@@
 
-A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$, such that each edge is constrained to lie on a line passing through the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
+\begin{definition}
+    A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \dfn{pins} with fixed positions in $\mathbb{R}^2$, such that each edge is constrained to lie on a line passing through the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
+\end{definition}
+
 %
@@
 % \begin{figure}\centering
-%     \includegraphics[width=.7\linewidth]{img/pinned}
+%     \includegraphics[width=.7\linewidth]{../../img/pinned}
 %     \caption{An isostatic pinned line-incidence graph.}
@@
 
-% \noindent
-% \note The recombination problem for  pinned line-incidence systems is trivial.
-Since the pins are given fixed positions in the plane, the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.
-
-\sidenote{The recombination problem for  pinned line-incidence systems is trivial.
-Since the pins are given fixed positions in the plane, the solutions of a isostatic sub-system will automatically be consistent with solutions of the remaining of the system.}
+% \sidenote{
+Note that the recombination problem for pinned line-incidence systems is trivial.
+Since the pins are given fixed positions in the plane, the solutions of an isostatic subsystem will automatically be consistent with the solutions of the other subsystems.
+% }

### 2015-08-04 doc/GMP2015/revision_3/s08_conclusion.tex
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

@@
 For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, a strong Church-Rosser property.
-This gives an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
+This gives an efficient (\ComplexityCanDRP) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
 
@@
 \begin{openproblem}
-    Is there a more efficient algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
+    Is there a more efficient algorithm than \ComplexityCanDRPV\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
 \end{openproblem}

### 2015-08-05 src/Application/Graph_Display_Window.cpp
Forced GLM_RADIANS where is needed to be

@@ -3,7 +3,6 @@
 #include "../Graph/Pebbled_Graph.hpp"
 
 
-
 Graph_Display_Window::Graph_Display_Window(Main_GUI_Manager *mgm, Graph *graph) :
     Movable_App_Window_2D(mgm),
     graph(graph),
@@ -86,6 +85,7 @@ void Graph_Display_Window::update_graph_positions() {
     edges_highlight.clear();
     edges_dashed.clear();
     // non_edges.clear();
+    
 
     if (drp)
         _update_graph_positions_drp();
@@ -291,8 +291,8 @@ void Graph_Display_Window::update_display() {
 
     // string printthis = "Hello,\nthis\nis\na\nfont\ntest!";
     // print_string(0,0,printthis.c_str(),0,0,0);
-    // cout << "DISPLAY: Number of verts = " << vertices.size() << endl;
-    // cout << "DISPLAY: Number of edges = " << edges.size() << endl;
+    // std::cout << "DISPLAY: Number of verts = " << vertices.size() << std::endl;
+    // std::cout << "DISPLAY: Number of edges = " << edges.size() << std::endl;
     drawer->draw_graph_edges(edges, edges_highlight, edges_dashed);
     drawer->draw_graph_vertices(vertices, vertices_highlight);
     drawer->draw_graph_vertices_names(vertices_names, vertices, vertices_highlight_names, vertices_highlight);

### 2015-08-05 src/GUI/Camera_2D.hpp

@@ -1,7 +1,7 @@
 #ifndef CAMERA_2D_HPP
 #define CAMERA_2D_HPP
 
-
+#define GLM_FORCE_RADIANS
 #include "../../ext/glm/glm.hpp"
 #include "../../ext/glm/gtc/matrix_transform.hpp"
 #include "../../ext/glm/gtx/rotate_vector.hpp"

### 2015-08-05 src/GUI/gl_objects.hpp

@@ -1,10 +1,10 @@
 #ifndef GL_OBJECTS_HPP
 #define GL_OBJECTS_HPP
 
+#define GLM_FORCE_RADIANS
 
 #include "../../ext/glm/glm.hpp"
 
-#define GLM_FORCE_RADIANS
 #include "../../ext/glm/gtx/rotate_vector.hpp"
 
 // #include <GLFW/glfw3.h>

### 2015-08-23 doc/GMP2015/revision_3/appendix_proofs.tex
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.
Lots of work on Revision 3, getting close.

@@
 
+\begin{proof}
+    % (Of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc})
+    %
+    Take two children of node $C$, called $C_i$ and $C_j$.
+Let $R_i$ be the graph induced in node $C$ by the edge set of $C$ minus the edge set of child $C_i$ ($C\setminus C_i$ for convenience); let $R_j=C\setminus C_j$.
+Let $D_{i,j}=C_i\cap C_j=C\setminus (R_i\cup R_j)$.
+Let $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$, and take these proper subgraphs to be non-empty.
+
+    If there are two children ($N=2$) then the proof is simple, it follows from the definition of a DR-plan: the union of the children is the parent which is isostatic.
+Assume that $N>2$ and take a third child, called $C_k$.
+    % Let $R_k=C\setminus C_k$.
+    We want to determine what $C_k$ can be, in terms of $i$ and $j$.
+$C_k$ can possibly be composed of an element from $\{\emptyset, D'_{i,j}, D_{i,j}\}$, an element from $\{\emptyset, R'_i, R_i\}$, and an element from $\{\emptyset, R'_j, R_j\}$, for a total of 27 cases.
+We will exhaustively show that it must be $R_i\cup R_j\cup D'_{i,j}$.
+First, we make the following observation:
+
     \begin{observation}\label{observation:no_edges_between_diff}
-Take $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
+        If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
     \end{observation}
@@
 
-\begin{proof} (Of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc})
-
-The alternative phrasing is as in Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}.
-
-Take $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
+    % Take $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$).
 Furthermore, take the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ that are non-empty.
 
-Assume that there is a third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
+    % Assume that there is a third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.
 
-Without loss of generality, all graphs are the induced subgraphs of $C$.
+    % Without loss of generality, all graphs are the induced subgraphs of $C$.
 
@@
 
-Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving the Lemma.
+    Since $D_{i,j}\subset C_i, C_j$ it means that $C_k\cup C_i = C_k \cup C_j = C$, thus proving Point \ref{lemma:wc_intersection_makes_all_wc}.
+    The alternative phrasing of this Point is straightforward from Point \ref{lemma:wc_intersection_is_C}.
     %
@@
 Choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children.
-From Observation \ref{rem:1dofcanon}, we know that the children can only share  a single pin or a sub-graph.
+From Observation \ref{rem:1dofcanon}, we know that the children can only share  a single pin or a subgraph.
 Pin the new body to bodies in two separate children.

### 2015-08-23 doc/GMP2015/revision_3/old_algo_sections.tex
Lots of work on Revision 3, getting close.

--- 2015-08-04 doc/GMP2015/revision_3/s03_drp.tex
@@
-\section{Main Result: Canonical DR-Plan, Optimality, and Algorithm}
-\label{sec:DRP}
-% In the problem of the optimal DR-Plan there is generally not a unique plan.
-% % Indeed, we will prove a union of $N$ isostatic subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same.
-Therefore, all choices of decomposition are in some sense equivalent.
-The theorem we seek to prove is thus:
-% However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems.
-Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
-% % To assist in showing this, we prove this core theorem throughout this section:
-
-% In this section, we discuss 2D bar-joint graphs.
-All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$.
-Trivial graphs are a single vertex and empty set.
-Furthermore, 2D isostatic graphs must be connected.
-% % The greatest density of a 2D isostatic graph is $-2$ (the vertex).
-The other disconnected part of the graph would need to have a density of $-1$, which is overconstrained and not possible in a isostatic graph (because there is no trivial graph with that density).
-
-\subsection{Canonical DR-Plan}
-In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
-Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find the canonical DR-plan of any independent graph.
-The definition is as follows:
-
-% \begin{definition}\label{def:canonical_drplan}
-%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
-%     (1) it is a DR-plan of $G$;
-%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
-% \end{definition}
-
-\begin{definition}\label{def:canonical_drplan}
-    A \dfn{canonical DR-plan} is a DR-plan that satisfies the additional two properties:
-    \begin{enumerate}
-        \item Children are rigid vertex-maximal proper subgraphs of the parent.
-        \item If all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
-    \end{enumerate}
-\end{definition}
-
-In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
-% \sidenote{In this section, $G$ is always assumed to be isostatic.}
-
-% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
-
-Definition~\ref{def:canonical_drplan} gives the canonical DR-plan a surprisingly strong Church-Rosser property, which is made explicit in Theorem~\ref{theorem:main}, the main result of this section.
-
-\begin{theorem}
-\label{theorem:canonical_exists_and_is_optimal}
-\label{theorem:canonical_is_optimal}
-\label{theorem:main}
-    A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.
-\end{theorem}
-
-% \begin{theorem} \label{theorem:canonical_is_optimal}
-%     \label{theorem:main}
-%     Any canonical DR-plan is an optimal DR-plan.
-% \end{theorem}
-
-%The proof of this theorem is a direct consequence of the following  more general theorem.
-
-%\begin{theorem}\label{theorem:main}
-%Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$
-%with children $C_1,\ldots,C_N$ preserve children according to the following rules.
-%\begin{enumerate}
-%    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
-%    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
-%\end{enumerate}
-%This is a canonical DR-plan.
-%\end{theorem}
-
-\begin{proof}
-We show the existence of a canonical DR-plan by constructing it as follows:
-
-Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$  retain children nodes according to the following rules:
-\begin{enumerate}
-   \item If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
-   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
-\end{enumerate}
-
-This directly satisfies Properties (2) and (3) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs,  which we shorten to {\em clusters}.
-To show Property (1) holds (that this constitutes a DR-plan):
-for Case 1 above,  since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Case 2 above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.
-
-Note that if we begin with an isostatic graph, ``rigid'' can be replaced with ``isostatic'' throughout the construction and preserve the above properties.
-The rigid proper subgraphs of an isostatic graph must be isostatic themselves.
-
-Next we show that a canonical DR-plan is optimal.
-
-First, note  that any  DR-plan $R$ without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$)
-is the subgraph of some cluster of $C$.
-Thus without loss of generality, we can assume that an optimal DR-plan satisfies Property (2) of a canonical DR-plan.
-
-The proof of optimality of a canonical DR-plan is by induction on its height.
-The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges.
-The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
-For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.
-Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
-Thus, from the induction hypothesis, we know that the $R(C_i)$ is optimal for $C_i$.
-
-Thus it is sufficient to demonstrate a set of nodes $S$  that must be present in any  DR-plan $R$ for $C$ that satisfies Property (2), including a known optimal one; and furthermore, for any such DR-plan $R$,  either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $R$ has the minimum possible fan-in of 2.
-
-We show the two claims below.
-The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
-Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.
-The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the {\em same maximal antichain $S$ of descendants $D$ of $C$}.
-The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment.
-I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$.
-Thus any  DR-plan that satisfies Property (2) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$.
-The two claims complete the proof that every canonical DR-plan is optimal.
-
- % We do this inductively on the l, beginning with the root node, showing that the rules are always the optimal choice.
-% by discussing each rule and proving inductively that this is the optimal choice at each level of the DR-plan tree, starting with the root node.
-
-% Now we show that this plan is optimal by considering the two cases.
-Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
- % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
-Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider.
-Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.
-
-\medskip\noindent
-\vemph{Claim 1:}
-% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
-%
-A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial, must be children of $C$ in an optimal DR-plan.
-
-We prove this claim  by showing that the union of no subset of the children can be $C$, thereby requiring all of them to be included as children.
-
-% must be optimal for this node.
-
-We prove by contradiction.
-Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic.
-If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$.
-So, it must be that $U=C$.
-\usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices.
-By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
-Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible.
-As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.
-
-Furthermore, since a canonical DR-plan has nodes with proper rigid \vemph{vertex-maximal} subgraphs as children, if, as in this case, their pairwise intersection is trivial, it follows that any node has at most as many children as a DR-plan without this restriction, because the union of the children must contain all edges of the parent.
-Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.
-
-\medskip\noindent
-\vemph{Claim 2:}
-If some pair in the set of child clusters $C_1,\ldots,C_N$ of $C$ has an isostatic (nontrivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of $C$.
-
-To prove Claim 2, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
-This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself.
-Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
-% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
-
-\newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
-\renewcommand{\induceonc}[1]{#1}
-\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in [N]\setminus\{#1\}}{R_k}}}
-
-However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants leading up to a common maximal antichain $S$ of subgraphs.
-
-To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$.
-Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children.
-For convenience, we will assume all subgraphs are induced subgraphs of $C$.
-We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
-The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs.
-So any two of these are viable children for $C_i$.
-% Since
-% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
-% the children of $C_i$ will be
-% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
-% and
-% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
-% % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
-% % the children of this node will be
-% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
-% % and
-% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
-% for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
-This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
-At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
-Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for  $N$ levels,  and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
-%
-% \medskip\noindent
-% This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
-%
-% -- say this at top `proof is by induction on the level of the dr=plan'
-\end{proof}
-
-This proof of this theorem relies on the following crucial observation and lemma.
-These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.
-
-\begin{observation*}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are non-empty isostatic graphs then the following hold: \\
-(1) $F_i\cup F_j$ is not trivial;
-(2) $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial;
-(3) $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic; and
-(4) $F_i\cap F_j$ is not underconstrained.
-\end{observation*}
-
-The following key properties hold at the nodes of a canonical DR-plan.
-
-\begin{lemma*}\label{lemma:combined_lemma}
-Let $C$ be an isostatic node of a canonical DR-plan, with distinct children $C_1,C_2,\ldots, C_m$. Assume $i\ne j$.
-Then
-\begin{enumerate}
-    \item\label{lemma:wc_intersection_is_C}
-    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.
-
-    \item\label{lemma:wc_intersection_makes_all_wc}
-    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
-
-    \item\label{lemma:uc_intersection_makes_all_uc}
-    If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
-\end{enumerate}
-\end{lemma*}
-
-\begin{remark}
-The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{cluster minimality} (defined in \cite{hoffman2001decompositionI} and in Section \ref{sec:prev}) holding for canonical-optimal DR-plans.
-\end{remark}
-
-% \begin{figure*}\centering
-% \begin{subfigure}{.3\linewidth}\centering
-%     \input{../../img/tikz/c2c3_of_k33s-0.tikz}
-%     \caption{}\label{fig:c2c3ofk33s:a}
-% \end{subfigure}%
-% \begin{subfigure}{.7\linewidth}\centering
-%     \input{../../img/tikz/c2c3_of_k33s-1.tikz}
-%     \caption{}\label{fig:c2c3ofk33s:b}
-% \end{subfigure}
-
-% \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
-This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
-(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
-$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-The dashed lines indicated that this exact structure is repeated.}
-% \label{fig:c2c3ofk33s}
-% \end{figure*}
-
-% \FigInit
-%     {(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
-This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
-(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
-$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-The dashed lines indicated that this exact structure is repeated.}
-%     {fig:c2c3ofk33s}
-% \FigTwoSubfigWithWidth
-%     {.3}
-%     {../../img/epsfromtikz/c2c3_of_k33s-0}
-%     {}
-%     {fig:c2c3ofk33s:a}
-%     %
-%     {.7}
-%     {../../img/epsfromtikz/c2c3_of_k33s-1}
-%     {}
-%     {fig:c2c3ofk33s:b}
-
-% \ClearMyMinHeight
-% \SetMyMinHeight{.3}{../../img/epsfromtikz/c2c3_of_k33s-0}
-% \SetMyMinHeight{.7}{../../img/epsfromtikz/c2c3_of_k33s-1}
-
-% \begin{figure*}\centering%
-%     %
-%     \begin{subfigure}{0.3\linewidth}\centering
-%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-0}
-%         \caption{}\label{fig:c2c3ofk33s:a}
-%     \end{subfigure}%
-%     %
-%     \hfill
-%     \begin{subfigure}{0.7\linewidth}\centering
-%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-1}
-%         \caption{}\label{fig:c2c3ofk33s:b}
-%     \end{subfigure}%
-%     %
-%     \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$.
-This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines.
-(\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes.
-$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-The dashed lines indicated that this exact structure is repeated.}
-%     \label{fig:c2c3ofk33s}
-% \end{figure*}%
-
-\ClearMyMinHeight
-\SetMyMinHeight{.3}{../../img/svg/revised_c2c3_of_k33s}
-\SetMyMinHeight{.7}{../../img/svg/revised_c2c3_of_k33s_candrp}
-
-\begin{figure*}\centering%
-    %
-    \begin{subfigure}{0.3\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s}
-        \caption{}\label{fig:c2c3ofk33s:a}
-    \end{subfigure}%
-    %
-    \hfill
-    \begin{subfigure}{0.7\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s_candrp}
-        \caption{}\label{fig:c2c3ofk33s:b}
-    \end{subfigure}%
-    %
-    \caption{
-    (\ref{fig:c2c3ofk33s:a}) A sequence of doublets ($C_2 \times C_3$) intersecting on triangles, where the edges of the triangles are replaced by $K_{3,3}$'s.
-This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows.
-(\ref{fig:c2c3ofk33s:b}) The canonical DR-plan of $G$, drawn as a DAG.
-$G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
-Below the third level, the obvious pattern continues until only the individual doublets are present (fourth level) with the ellipses indicating the remaining doublets between those shown.
-Decomposition of one of these doublets is shown.
-The dashed lines indicated that this exact decomposition (of the similar nodes on the level) is repeated.
-Further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges is omitted from the figure.
-    }
-    \label{fig:c2c3ofk33s}
-\end{figure*}%
-
-\begin{example}[DR-plan for self-similar structure]
-% \myexample
-% \textsl{[DR-plan for self-similar structure]}
-This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, the canonical DR-plan of $G$.
-It begins with the whole (isostatic) graph as the root.
-The graph $G$ has only two isostatic vertex-maximal subgraphs: $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle (triangle $N$).
-These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic.
-As explained in the proof of Theorem \ref{theorem:main}, since there are only 2 possible children, their intersection must be a node 2 levels below the parent.
-As expected, it is on the third level, as a child of both of $G$'s children.
-
-Both of $G$'s children are similar to $G$, but containing only $N-1$ triangles.
-Therefore, the canonical DR-plans of these children follow the same pattern.
-This continues downward until the individual doublets are reached (there will be multiple occurrences of the same doublets at this level, but they can be represented as the same node in a DAG).
-
-Further decomposition of one of these doublets is shown.
-The three edges between the triangles and the triangles themselves all intersect trivially pairwise.
-By Theorem \ref{theorem:main}, part 1, they must all be children in the DR-plan.
-Similarly, the triangles decompose into their three trivially intersecting $K_{3,3}$'s.
-Then the $K_{3,3}$ subgraphs decompose into their separate 9 edges.
-
-The self-similar nature of this graph is evident in the canonical DR-plan. Many structures are repeated throughout the DR-plan, allowing for shared computation in both decomposition and recombination.
-\end{example}
-
-% \subsection{Extensions}
-% This framework immediately pushes through for body-pin systems via a simple reduction.
-If there are $N$ pins on a body, it can be represented as a 2-tree with $N$ vertices, each corresponding to a pin, making sure to select edge distances such that the distance between pins is preserved.
-E.g.\ a body with two pins is an edge, three pins is a triangle, etc.
-Any bodies that share a pin now intersect on their vertex that corresponds to that pin.
-Now we have a bar-joint representation of the body-pin system in 2D and all proofs follow.
-
-% With more effort, it can be shown that pinned line-incidence systems can also use this framework. This is done in section \ref{XXX}.
-
-\subsection{Algorithm}
-
 % \ClearMyMinHeight
@@
 
-\begin{theorem}\label{theorem:algo_complexity}
+\begin{theorem}
+\label{theorem:algo_complexity}
     There exists an \ComplexityCanDRPV\ algorithm to find a canonical DR-plan for an independent input graph.
@@
 
-% In the case of trivial intersections, all of the found clusters will be unique.
-If the cardinality is greater than 2, we can save computation by not checking the other edges in the found cluster.
-But we can only know that they are trivial intersections after finding 2 different clusters.
-So do that first...
-
-% In the case of isostatic intersections, if we begin with an edge in the core, there are several clusters containing this edge (and the algo will pick the largest among them).
-We could only safely skip the other edges in the core, but we don't know what it is yet.
-So after finding the first cluster, go to an edge not in it and find the next cluster.
-Check the intersection.
-Now we can skip the others in the core.
-WAIT....
-THIS WILL ONLY PICK THE LARGEST, AND ITS ``COMPLEMENT'', SO TO SPEAK.
-WHICH TECHNICALLY WORKS....
-
-% For each found cluster $c$, for each edge $e$ \emph{not in} $c$, delete $c$ from $M(e)$ and delete the pointer to $M(e)$ from the pointer list of $e$.
-
 % \begin{proof}
@@
 
-% \begin{figure*}\centering%
-%   \includegraphics[width=\linewidth]{../../img/svg/overconstrained_choices}
-%   \caption{The first graph is over constrained by one edge (a $K_{3,3}$ with an extra vertex attached by 3 edges).
-The second graph is a minimally rigid subgraph of the first.
-Any DR-plan of this graph will have fan-in of at least 9, occurring at the node that decomposes the $K_{3,3}$.
-The third graph is a different subgraph of the first.
-A canonical and cluster minimal DR-plan of size 2 can be made for this graph (it is a Henneberg-I construction).
-Thus, we can see the first graph has both non-optimal (second graph) and optimal (third graph) canonical, cluster minimal DR-plans.}
-%   \label{fig:overconstrained_choices}
-% \end{figure*}%
-
-\ClearMyMinHeight
-\SetMyMinHeight{.6}{../../img/svg/new_overconstrained_not_optimal}
-\SetMyMinHeight{.4}{../../img/svg/new_overconstrained_optimal}
+While the canonical DR-plan is clear conceptually, it is difficult to reason with directly in the design and analysis of algorithms.
+Therefore, we introduce the following modification to get a similar but more useful structure.
 
@@
   %
-    \begin{subfigure}{0.5\linewidth}\centering
-        \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_optimal}
-        }
-        \caption{}\label{fig:overconstrained:optimal}
-    \end{subfigure}%
-    %
-    \hfill
-    \begin{subfigure}{0.5\linewidth}\centering
-        \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_not_optimal}
-        }
-        \caption{}\label{fig:overconstrained:not_optimal}
-    \end{subfigure}%
-    %
-    \caption{
-    % (\ref{fig:overconstrained:optimal}) An optimal DR-plan of a singly overconstrained rigid graph with a fan-in of 5.
-Decomposition of triangles is omitted.
-(\ref{fig:overconstrained:not_optimal}) A canonical and cluster-minimal DR-plan of the same graph.
-The DR-plan has a fan-in of 9 and is non-optimal, shown by the preceding counter-example.
-Since the nontrivial intersection of the two children of the root is underconstrained, their decompositions are separate, causing the large fan-in.
-    Both figures are canonical and cluster-minimal DR-plans of the same singly overconstrained rigid graph.
-Decomposition of triangles are omitted and dashed lines indicate a decomposition similar to the other nodes on the same level.
-(\ref{fig:overconstrained:optimal}) is an optimal DR-plan, with a fan-in of 5.
-(\ref{fig:overconstrained:not_optimal}) has a fan-in of 9 and is non-optimal, shown by the preceding counter-example.
-    }
-    \label{fig:overconstrained}
+  \includegraphics[width=0.3\linewidth]{../../img/svg/3xc2c3_candrp_seq}
+  \caption{The sequential canonical DR-plan of $G_{demo}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
+The children of the triangle are omitted.
+Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
+Also, note that the bottom-left node, the triangle, is the intersection of the 3 children of $G_{demo}$ in $ComDRP(G_{demo})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
+  \label{fig:demo_graph:candrpseq}
 \end{figure*}%
 
-\subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
-\label{sec:drp:overconstrained}
+\begin{definition}
+    A \dfn{sequential} canonical DR-plan (or simply sequential DR-plan) is a canonical DR-plan with the following modification:
 
-For overconstrained (not independent) graphs, a canonical DR-plan is still well-defined.
-However, it may be far from optimal. The proofs of Theorem \ref{theorem:main}, Observation \ref{lemma:union_intersection}, and Lemma \ref{lemma:combined_lemma} all fail for overconstrained graphs.
-It is important to note that, regardless whether the graph is overconstrained, if every node in a canonical DR-plan $R$ has clusters whose pairwise intersection is trivial, then the DR-plan is the unique one satisfying Property (2), and since we know that there is an optimal DR-plan that satisfies Property (2), $R$ is in fact optimal.
-The problem arises when some node in a DR-plan has clusters whose pairwise intersection is non-trivial.
-In this case, an arbitrary choice of a pair of clusters as children of an overconstrained node in a canonical DR-plan may not result in an optimal DR-plan.
-This is in contrast to independent graphs, which, as shown in Theorem~\ref{theorem:main}, exhibit the strong Church-Rosser property that any choice yields an optimal DR-plan.
-% \todo{rephrase? Theorem 4 shows this. If it's independent it \vemph{will} be optimal (the church rosser property), no longer holds if it's overconstrained.}
-A good source of examples of overconstrained graphs with canonical DR-plans that are not optimal are graphs whose cluster-minimal DR-plans that are not optimal.
-The example shown in Figure \ref{fig:overconstrained} is a canonical, cluster-minimal DR-plan that is not optimal; an optimal DR-plan is also shown in the figure.
-The root cause of the NP-hardness is encapsulated in this figure: because the different choices of vertex-maximal subgraphs for overconstrained input do not incur the same fan-in, finding the optimal DR-plan becomes a search problem with a combinatorial explosion of options.
+    In the case that a node has isostatic vertex-maximal proper subgraphs (or clusters) that intersect non-trivially.
+    we construct a \dfn{sequential extension} in the following manner.
+    %
+    Let the current node be $G$.
+    Let the $N$ clusters be $C_1, C_2, \ldots, C_N$.
+    Let the \dfn{core} be $I=\bigcap_{k=1}^{N}{C_k}$.
+    Let the $N$ \dfn{appendages} be $R_i = G\setminus C_i$.
+    %
+    Choose arbitrary appendage $R_i$ and set the children of $G$ as $I\cup \bigcup_{k=1}^{i}{R_k}\cup \bigcup_{k=i}^{N}{R_k}$ and the clusters of $R_i$.
+Construct the sequential canonical DR-plan of each child.
+\end{definition}
 
-As mentioned earlier, the Modified \frontier\ algorithm
-version given in \cite{lomonosov2004graph} runs in polynomial time and finds a cluster-minimal DR-plan for any graph.
-Similarly, the algorithm given above finds a canonical DR-plan also for any input graph.
-However neither of these DR-plans may be optimal for overconstrained graphs as shown in Figure \ref{fig:overconstrained}.
+See Figure~\ref{fig:demo_graph:candrpseq} for an example.
+Note the pattern will continue for $N$ levels, where $N$ is the number of appendages.
+At the top is $G$, the core with all appendages; at the next level, the children are the core with all but one appendage and the clusters of the removed appendage; at the next level, it has the core with all but two appendages and the clusters of the appendage removed second; etc.; at the bottom level is the core.
 
-While the canonical DR-plan is optimal only if the input graph is independent, when there are only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan using a straightforward modification of the above algorithm.
-However, the time complexity is exponential in $k$.
+\begin{observation}
+    A sequential canonical DR-plan of independent graph $G=(V,E)$ is a tree with $O(|V|)$ nodes.
+\end{observation}
 
-This exponential growth of time complexity for overconstrained graphs is in fact captured in the proof of NP-hardness of optimal DR-planning in \cite{sitharam2005combinatorial, lomonosov2004graph}.
+\begin{observation}
+    A sequential canonical DR-plan of an independent graph is optimal.
+\end{observation}
+
+\begin{proof}
+    A sequential DR-plan behaves the same as a canonical DR-plan in the case of a node whose clusters have trivial intersections.
+    In a regular canonical DR-plan, at a node with $N$ clusters that intersect non-trivially, there is fan-in of 2 until level $N$.
+At level $N-1$, there are $N$ distinct nodes, one for each appendage together with the core.
+The children of these nodes are the core and the clusters of the corresponding appendage.
+Thus, the maximum fan-in over this part of the canonical DR-plan is one plus the maximum number of clusters in an appendage.
+By construction, this is exactly the maximum fan-in of a sequential extension.
+\end{proof}
+
+\begin{observation}\label{obs:seq-to-reg-canonical}
+    Given a sequential canonical DR-plan of an independent graph $G=(V,E)$, a canonical DR-plan can be found in time $O(|V|^3)$.
+\end{observation}
+
+\begin{proof}
+    In the case of a node whose children intersect trivially, nothing needs to be done.
+    In the case of non-trivial intersections, we will need to construct the canonical DR-plan subtree from the sequential extension.
+
+    At such a node $G$ with $N$ clusters, we will have to build $\sum_{i=1}^N{i}=O(N^2)$ new nodes.
+For convenience, $C_{[i,j]}$ will represent the core plus appendages $i$ through $j$.
+We set the children of $G$ as $C_{[0,N-1]}$ and $C_{[1,N]}$.
+The children of $C_{[0,N-1]}$ are set as $C_{[0,N-2]}$ and $C_{[1,N-1]}$.
+The children of $C_{[1,N]}$ are set as $C_{[1,N-1]}$ (not unique) and $C_{[2,N]}$.
+This pattern continues for $N$ levels, with level $i$ containing $i$ unique nodes.
+
+    Since there are at most $O(|V|)$ nodes with $O(|V|)$ clusters that need this treatment, the complexity is $O(|V|^3)$.
+\end{proof}
+
+Before discussing the algorithm, we remind the reader of the component pebble game~\cite{Jacobs:1997:PG}.
+Given an independent input graph $G$ with vertex weight 2 and edge weight 1, the output will be all of the 2D maximal isostatic subgraphs (the components).
+We will denote this algorithm with $\components{G}$.
+For independent input, this runs in time $O(|V|^2)$.
+
+The algorithm for finding the sequential DR-plan is going to rely on the observation that the set of $\components{G\setminus e}$ over all $e\in E$ will (essentially) encompass all the nodes of the sequential DR-plan for $G$, with minor caveats that we will explain later.
+
+\begin{lemma}\label{lemma:seq-nontriv-Gtoe}
+    When the intersections of clusters are trivial at all nodes in the (sequential) canonical DR-plan, then the set $\components{G\setminus e}$ will contain all of the siblings in the path from $G$ to $e$.
+\end{lemma}
+
+\begin{proof}
+    The sequential and canonical DR-plans are identical in this case.
+Suppose $G$ has $N$ trivially intersecting components, and $e$ is in component $C_i$.
+Removing edge $e$ will leave all components besides $C_i$ intact, thus they will be isostatic components in $G\setminus e$ and will be in the set.
+Therefore, $\components{G\setminus e}$ contains
+    % $C_1, \ldots, C_{i-1}, C_{i+1}, \ldots, C_N$
+    all clusters besides $C_i$
+    and the set $\components{C_i\setminus e}$. By induction, the proof is complete.
+\end{proof}
+
+\begin{lemma}\label{lemma:seq-Gtoe}
+    In the (sequential) canonical DR-plan, the set $\components{G\setminus e}$ will contain all of the siblings in the path from $G$ to $e$ excluding the following: in the case of a node whose $N$ clusters have non-trivial intersections, those nodes that contain the core plus any $i$ appendages where $0\leq i<N-1$.
+\end{lemma}
+
+\begin{proof}
+    By Lemma~\ref{lemma:seq-nontriv-Gtoe}, this is true for trivial intersections.
+For non-trivial intersections, consider an edge $e$ that comes from appendage $R_i$.
+Then $\components{G\setminus e}$ will contain the element `core plus all appendages but $R_i$' and the set $\components{R_i\setminus e}$, so the statement is true.
+Now, consider an edge $e$ that comes from $I$, the core.
+Then $\components{G\setminus e}$ will contain $\components{R_i}$ for all appendages $R_i$ and $\components{I\setminus e}$.
+\end{proof}
+
+Note that in the sequential DR-plan, the number of missing nodes is at worst $O(|V|^2)$.
+There can be as many as $O(|V|)$ nodes with non-trivial intersections on $O(|V|)$ clusters, each cluster contributing one appendage and one missing node (excluding the first in the sequential extension).
+
+\begin{theorem}\label{theorem:algo_complexity}
+    There exists an \ComplexityCanDRPV\ *** algorithm to find a canonical DR-plan for an independent graph.
+\end{theorem}
+
+\begin{proof}
+    Let the input graph be $G=(V,E)$.
+Begin by computing $\components{G\setminus e}$ for all $e\in E$ in $O(|V|^3)$.
+By Lemmas~\ref{lemma:seq-nontriv-Gtoe} and \ref{lemma:seq-Gtoe}, this will contain all but a select group of nodes in the sequential DR-plan.
+Because the sequential DR-plan is a tree whose leaves are unique edges, there are $O(|V|)$ nodes.
+At each node, if the children can be determined in $O(|V|^2)$, the sequential DR-plan can be found in time $O(|V|^3)$ as a whole.
+By Observation~\ref{obs:seq-to-reg-canonical}, this would mean a canonical DR-plan can be found in $O(|V|^3)$.
+
+    First, we find the children.
+For each edge $e_i$, find the largest component from all $\components{G\setminus e_j}$ containing $e_i$.
+This is a cluster containing $e_i$ (in the case of trivial intersections, it is the only cluster containing $e_i$).
+
+    If the clusters intersect trivially, every cluster $C_1, C_2, \ldots, C_N$ will be found this way.
+For each cluster $C_i$, for each edge $e_j\in C_i$, compute $\components{C_i\setminus e_j}$ by removing the other clusters from $\components{G\setminus e_j}$; set $C_i$ as a child of $G$ and run the algorithm recursively without the $O(|V|^3)$ time recomputation of the components.
+
+    If the clusters intersect non-trivially, exactly two clusters will be found this way.
+Cluster $C_1$ will be the largest cluster overall and $C_2$ will be the largest cluster that contains appendage $R_1$.
+To find the remaining clusters, choose an arbitrary edge $e_i\in R_1$, for each edge $e_j\neq e_i$, if the component containing $e_i$ in $\components{G\setminus e_j}$ intersects non-trivially with $C_1$ then it is a cluster and is added to the list, otherwise $e_j$ came from the core.
+Once all clusters are found, the core and the appendages can be determined, and the sequential extension from $G$ can be computed.
+The incomplete children will be the sets $\components{R_i}$ for all appendage $R_i$ and the node containing the core.
+Before recursion, for each incomplete child $P$, for each $e\in P$, find $\components{P\setminus e}$ by removing the appropriate elements from $\components{G\setminus e}$.
+\end{proof}

### 2015-08-23 doc/GMP2015/revision_3/s02_preliminaries.tex
Lots of work on Revision 3, getting close.

@@
 \ClearMyMinHeight
-\SetMyMinHeight{.4}{../../img/svg/3xc2c3}
-\SetMyMinHeight{.3}{../../img/svg/3xc2c3_comdrp}
-\SetMyMinHeight{.3}{../../img/svg/3xc2c3_candrp}
+\SetMyMinHeight{.4}{../../img/svg/3xc2c3_new}
+\SetMyMinHeight{.3}{../../img/svg/3xc2c3_new_comdrp}
+\SetMyMinHeight{.3}{../../img/svg/3xc2c3_new_candrp}
 
 \begin{figure*}\centering%
-  %
+
   \begin{subfigure}{0.4\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_new}
     \caption{}\label{fig:demo_graph:graph}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_comdrp}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_new_comdrp}
     \caption{}\label{fig:demo_graph:comdrp}
@@
   \begin{subfigure}{0.3\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_candrp}
+    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_new_candrp}
     \caption{}\label{fig:demo_graph:candrp}
@@
   %
-  \caption{(\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section.
-(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$, i.e.\ $ComDRP(G_{demo})$.
-Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
-The children of triangles (3 edges) are omitted.
-(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{demo})$.
+  % \begin{subfigure}[c]{0.3\linewidth}\centering
+  %   \includegraphics[width=\linewidth]{../../img/svg/3xc2c3_new}
+  %   \caption{}\label{fig:demo_graph:graph}
+  % \end{subfigure}%
+  % %
+  % \hfill
+  % \begin{subfigure}[t]{0.345\linewidth}\centering
+  %   \includegraphics[width=\linewidth]{../../img/svg/3xc2c3_new_comdrp}
+  %   \caption{}\label{fig:demo_graph:comdrp}
+  % \end{subfigure}%
+  % %
+  % \hfill
+  % \begin{subfigure}[t]{0.345\linewidth}\centering
+  %   \includegraphics[width=\linewidth]{../../img/svg/3xc2c3_new_candrp}
+  %   \caption{}\label{fig:demo_graph:candrp}
+  % \end{subfigure}%
+  %
+  \caption{
+  (\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section.
+  %
+  (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$.
+  % i.e.\ $ComDRP(G_{demo})$.
+  Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted.
+  %
+  (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}).
+  % i.e.\ $OptDRP(G_{demo})$.
   The children of triangles are omitted.}
@@
 % %
-%     A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$.
+%     A DR-plan is \dfn{complete} if it satisfies an additional property: for an internal node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$.
 This makes Property 2 implicit.
@@
 
-% We begin with the definition of a general DR-plan and also introduce two specific kinds of DR-plans.
-
-\begin{definition}\label{def:drp}
+\begin{definition}
+[DR-plan]
+\label{def:drp}
     A \dfn{decomposition-recombination \mbox{(DR-)} plan}~\cite{hoffman2001decompositionI} of graph $G$ is defined as a forest that has the following properties:
     \begin{enumerate}
-        \item Each node represents a rigid subgraph of $G$.
+        \item Each node represents a rigid subgraph of $G$.\footnote{Nodes will be referred to interchangeably as ``the node that represents or contains the (sub)graph $C$'' and as simply ``$C$''.}
         \item A root node is a vertex-maximal rigid subgraph of $G$.
@@
 
+Note that this definition permits the same rigid subgraph to appear in multiple nodes of the DR-plan.
+Not permitting such duplication would, in general, require the DR-plan to be defined as a directed acyclic graph instead of a forest.
+
 \begin{definition}
-    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$ (which makes Property 3 of a DR-plan implicit.)
+[Complete DR-plan]
+    A DR-plan is \dfn{complete} if it satisfies an additional property: for an internal node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$ (which makes Property 3 of a DR-plan implicit).
 \end{definition}
@@
 \begin{definition}
-    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest. The maximum fan-in is called the \dfn{size} of the DR-plan.
+[Optimal DR-plan]
+    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest.
+    % The maximum fan-in is called the \dfn{size} of the DR-plan.
 \end{definition}
 
-We denote these different kinds of DR-plans as $\drp{G}$, $\comdrp{G}$, and $\optdrp{G}$, respectively.
-Note that nodes will be referred to interchangeably as ``the node that represents or contains the (sub)graph $C$'' and as simply ``$C$''.
+% We denote these different kinds of DR-plans as $\drp{G}$, $\comdrp{G}$, and $\optdrp{G}$, respectively.
 
@@
 
-See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
+See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \todo{we lost this figure, replaced with \ref{fig:seqdrp_nonuniqueness}, should it be cited here now or should the other be added back in somewhere?} \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
 

### 2015-08-23 doc/GMP2015/revision_3/s03_drp.tex
Lots of work on Revision 3, getting close.

--- 2015-07-19 doc/GMP2015/revision_3/s03_drp.tex
@@
+\newcommand{\drpx}[1]{P_{#1}}
+\newcommand{\comdrpx}[1]{P_{#1}^{com}}
+\newcommand{\candrpx}[1]{P_{#1}^{can}}
+\newcommand{\seqdrpx}[1]{P_{#1}^{seq}}
+
+% \newcommand{\componentsx}[1]{CMPNTS(#1)}
+% \newcommand{\componentsx}[1]{COMPONENTS(#1)}
+% \newcommand{\componentsx}[1]{\text{\emph{COMPONENTS}}(#1)}
+\newcommand{\componentsx}[1]{\textsc{Components}(#1)}
+
+\newcommand{\Branch}{Branch}
+\newcommand{\branch}{branch}
+\newcommand{\branches}{branches}
+
+% \newcommand{\branchx}[1]{\text{\emph{branch}}(#1)}
+\newcommand{\branchx}[1]{\textsc{Branch}(#1)}
+\newcommand{\branchGePG}{$\branchx{G,e,P_G}$}
+\newcommand{\branchLfPL}{$\branchx{L,f,P_L}$}
+
 \section{Main Result: Canonical DR-Plan, Optimality, and Algorithm}
@@
 In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
-Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph.
-The definition is as follows:
+Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
+% The definition is as follows:
 
-\begin{definition}\label{def:canonical_drplan}
-    The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
-    (1) it is a DR-plan of $G$;
-    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$.
+In this case unions and intersections are well defined.
+For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).
+
+% When using intersections and unions of graphs, we are always considering subgraphs of a single parent graph.
+In this case unions and intersections are well defined.
+For example, the union of $G_1=(V_1, E_1)$ and $G_2=(V_2, E_2)$ is the subgraph $(V_1\cup V_2, E_1\cup E_2)$; similarly, $G_1\cap G_2$ is the subgraph $(V_1\cap V_2, E_1\cap E_2)$.
+
+% \begin{definition}\label{def:canonical_drplan}
+%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
+%     (1) it is a DR-plan of $G$;
+%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
+%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+% \end{definition}
+
+\begin{definition}
+[Canonical DR-plan]
+\label{def:canonical_drplan}
+    A \dfn{canonical DR-plan} is a DR-plan that satisfies the additional two properties:
+    \begin{enumerate}
+        \item \label{def:canonical_drplan:prop1} Children are rigid vertex-maximal proper subgraphs of the parent.
+        \item \label{def:canonical_drplan:prop2} If all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+    \end{enumerate}
 \end{definition}
-
-In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
-% \sidenote{In this section, $G$ is always assumed to be isostatic.}
-
-% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.
 
@@
 \begin{theorem}
+[Canonical is optimal]
 \label{theorem:canonical_exists_and_is_optimal}
@@
 
-Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$  retain children nodes according to the following rules:
-\begin{enumerate}
-   \item If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
-   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
+    Let $\comdrpx{G}$ be the complete DR-plan of the rigid 2D bar-joint graph $G$.
+    % Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$.
+    For all nodes $C$ with children $C_1,\ldots,C_N$ retain children nodes according to the following rules:
+    \begin{enumerate}[(a)]
+       \item \label{thm:main:proof:rule1} If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
+       \item \label{thm:main:proof:rule2} If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
     \end{enumerate}
 
-This directly satisfies Properties (2) and (3) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs,  which we shorten to {\em clusters}.
-To show Property (1) holds (that this constitutes a DR-plan):
-for Case 1 above,  since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Case 2 above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.
+    This directly satisfies Properties (\ref{def:canonical_drplan:prop1}) and (\ref{def:canonical_drplan:prop2}) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrpx{G}$ are rigid vertex-maximal proper subgraphs, which we shorten to \dfn{clusters}.
+To show that a canonical DR-plan is, in fact, a DR-plan:
+    for Rule (\ref{thm:main:proof:rule1}) above, since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Rule (\ref{thm:main:proof:rule2}) above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.
 
@@
 
-First, note  that any  DR-plan $R$ without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$)
-is the subgraph of some cluster of $C$.
-Thus without loss of generality, we can assume that an optimal DR-plan satisfies Property (2) of a canonical DR-plan.
+    % First, note that any DR-plan $R$ without the Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (\ref{def:canonical_drplan:prop1}) without increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$) is the subgraph of some cluster of $C$.
+    First, note that any DR-plan of $G$, $\drpx{G}$, without the Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (\ref{def:canonical_drplan:prop1}) without increasing the max fan-in, since any child (a rigid proper subgraph) of node $C$ in $\drpx{G}$  is the subgraph of some cluster of $C$.
+    Thus, without loss of generality, we can assume that an optimal DR-plan satisfies Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan.
 
@@
 The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
-For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.
-Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
-Thus, from the induction hypothesis, we know that the $R(C_i)$ is optimal for $C_i$.
+    For the induction step consider a canonical DR-plan $\candrpx{C}$ rooted at node $C$ with height $t+1$.
+    Notice that $\candrpx{C}$ contains a canonical DR-plan $\candrpx{C_i}$ for the graphs $C_i$ corresponding to each of $C$'s descendent nodes.
+    Thus, from the induction hypothesis, we know that the $\candrpx{C_i}$ is optimal for $C_i$.
 
-Thus it is sufficient to demonstrate a set of nodes $S$  that must be present in any  DR-plan $R$ for $C$ that satisfies Property (2), including a known optimal one; and furthermore, for any such DR-plan $R$,  either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $R$ has the minimum possible fan-in of 2.
+    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $\drpx{A}$ has the minimum possible fan-in of 2 \todo{Shouldn't it be `node $A$ in $\drpx{C}$ has the minimum...'}.
 
     We show the two claims below.
-The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
-Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.
-The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the {\em same maximal antichain $S$ of descendants $D$ of $C$}.
+    The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}) must also satisfy Property (\ref{def:canonical_drplan:prop2}) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
+    Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (\ref{def:canonical_drplan:prop1}), including a known optimal one.
+    The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the \emph{same maximal antichain $S$ of descendants $D$ of $C$}.
 The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment.
 I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$.
-Thus any  DR-plan that satisfies Property (2) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$.
-The two claims complete the proof that every canonical DR-plan is optimal.
+Thus any DR-plan that satisfies Property (\ref{def:canonical_drplan:prop1}) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$.
+The two claims complete the proof of the induction step and thus the proof that every canonical DR-plan is optimal.
 
@@
      % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs.
-Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider.
-Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.
+Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to considerthat a canonical DR-plan subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.
 
@@
     \vemph{Claim 1:}
-% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
-%
-A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial, must be children of $C$ in an optimal DR-plan.
+    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If the pairwise intersection of clusters is trivial, all of the clusters must be children of $C$ in an optimal DR-plan
+    % A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial must all be children of $C$ in an optimal DR-plan.
 
@@
 
-% must be optimal for this node.
-
     We prove by contradiction.
-Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic.
+    % Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic.
 If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$.
 So, it must be that $U=C$.
+    % \usestwod
+    % However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices.
+By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
+Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible.
+As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.
+    Assume to the contrary that there is a strict subset $S$ of the clusters such that $U$, the union of all elements in $S$, is isostatic.
+If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the clusters in set $S$.
+So, it must be that $U=C$.
     \usestwod
-However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices.
+    However, since $C_i \cap C_j$ is trivial then for $C_k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more vertices, i.e.\ disconnected trivial subgraphs.
 By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$.
@@
     \vemph{Claim 2:}
-If some pair in the set of child clusters $C_1,\ldots,C_N$ of $C$ has an isostatic (nontrivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of $C$.
+    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If some pair of clusters
+    % If some pair in the set of clusters $C_1,\ldots,C_N$ of $C$
+    has an isostatic (non-trivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of node $C$.
 
-To prove Claim 2, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic.
+    To prove Claim 2, notice that if $C_i\cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i\cup C_j$ is also isostatic.
 This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself.
 Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
-% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.
 
@@
 
-To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$.
+    % To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$.
 Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children.
@@
 So any two of these are viable children for $C_i$.
+    Let $I$ denote the intersection of all the clusters; we call this the \dfn{core}.
+Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
+    % To prove this holds,
+    Suppose $C_i$ and $C_j$, where $i\neq j$, are taken to be the children of node $C$.
+    % We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
+    The $N-1$ clusters of $C_i$ are the core plus all appendages except $R_i$ and $R_j$, for each $j\neq i$.
+The pairwise intersection of any of these clusters of $C_i$ will clearly be isostatic, so any two of them are viable children of node $C_i$.
+     % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
     % Since
@@
     % for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
-This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
-At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
-Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for  $N$ levels,  and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
+    Beginning with node $C$, this pattern repeats for $N-1$ levels.
+Every node in this subtree rooted at $C$ has a fan-in of two (the minimum possible) up through this level.
+At level $N-1$, we have a set of nodes where each node is the core plus some appendage (with every appendage appearing at least once).
+    % This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
+    % At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
+    Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for $N-1$ levels, and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
     %
@@
 
-\begin{observation*}\label{lemma:union_intersection}
-If $F_i$ and $F_j$ are non-empty isostatic graphs then the following hold: \\
+\begin{observation*}
+\label{lemma:union_intersection}
+    If $F_i$ and $F_j$ are isostatic subgraphs of an independent graph then the following hold:
+    % \begin{enumerate}
+    %     \item $F_i\cup F_j$ is not trivial.
+    %     \item $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial.
+    %     \item $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic.
+    %     \item $F_i\cap F_j$ is not underconstrained.
+    % \end{enumerate}
     (1) $F_i\cup F_j$ is not trivial;
@@
 
-\begin{lemma*}\label{lemma:combined_lemma}
-Let $C$ be an isostatic node of a canonical DR-plan, with distinct children $C_1,C_2,\ldots, C_m$. Assume $i\ne j$.
+\begin{lemma*}
+\label{lemma:combined_lemma}
+    Let $C$ be an isostatic node of a canonical DR-plan with distinct children $C_1,C_2,\ldots, C_N$. Assume $i\ne j$.
     Then
@@
 This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows.
-(\ref{fig:c2c3ofk33s:b}) The canonical DR-plan of $G$, drawn as a DAG.
+(\ref{fig:c2c3ofk33s:b}) A canonical DR-plan of $G$, drawn as a DAG.
 $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs.
@@
 
-\begin{example}[DR-plan for self-similar structure]
-% \myexample
-% \textsl{[DR-plan for self-similar structure]}
-This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, the canonical DR-plan of $G$.
+\begin{example}
+[DR-plan for self-similar structure]
+    This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, a canonical DR-plan of $G$.
 It begins with the whole (isostatic) graph as the root.
@@
 
-The self-similar nature of this graph is evident in the canonical DR-plan. Many structures are repeated throughout the DR-plan, allowing for shared computation in both decomposition and recombination.
+    The self-similar nature of this graph is evident in the canonical DR-plan.
+Many structures are repeated throughout the DR-plan, allowing for shared computation in both decomposition and recombination.
 \end{example}
@@
 \subsection{Algorithm}
+\label{sec:DRP:algo}
 
-% \ClearMyMinHeight
-% \SetMyMinHeight{.4}{../../img/epsfromtikz/demo_graph}
-% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_comdrp}
-% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_candrp}
+The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}.
+We list these properties before describing the algorithm.
+Furthermore, in order to deal with the apparent non-uniqueness of a canonical DR-plan (in the case when the children of a node have non-trivial intersections), and to make the canonical DR-plan more malleable for algorithmic purposes, we define a key notion of a \dfn{sequential DR-plan} that is obtained from a canonical DR-plan, and also turns out to be an optimal DR-plan.
+A sequential DR-plan further clarifies the essential uniqueness of canonical DR-plans.
+First, we present a recursive definition of the canonical DR-plan, which is more useful in this section, and make an important observation.
 
-% \begin{figure*}\centering%
-%   %
-%   \begin{subfigure}{0.4\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph}
-%     \caption{}\label{fig:demo_graph:graph}
+\begin{definition}
+[Canonical DR-plan]
+\label{def:canonical_drp_rec}
+    A \dfn{canonical DR-plan} of $G$ is recursively defined as follows:
+    \begin{enumerate}
+        \item Base case: When $G$ is a single edge, the canonical DR-plan for $G$ is $G$ itself.
+        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$.
+        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$ and $C_j$.
+    \end{enumerate}
+    % When $G$ is a single edge (base case), the canonical DR-plan for $G$ is $G$ itself.
+\end{definition}
+\todo{What if input is not isostatic? What roots?}
+
+\begin{observation}
+[Canonical DR-plan size]
+    While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes).
+If we want the canonical DR-plan to have small size (linear \todo{or just polynomial?} in the number of vertices) we can choose the children in the following fashion: In Case (3) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp.
+$C_j\setminus C_i$) (which are not rigid, by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}).
+
+    Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
+\end{observation}
+
+% Observation XX[recursive definition of canonical DR-plan]:
+% \begin{observation}
+    % In a canonical DR-plan of an independent graph, the portion of the DR-plan rooted at any node $C_i$ is a canonical DR-plan for $C_i$. \todo{unneccessary sentence.}
+
+    % In other words, a canonical DR-plan of $C$ is recursively defined by taking the children of $C$ to be the roots of canonical DR-plans for (a) each proper vertex-maximal rigid  subgraph of $C$, in case their pairwise intersection is trivial or (b) two proper vertex-maximal rigid  subgraphs $C_i$ and $C_j$ of $C$, in case their intersection is non-trivial.
+When $C$ is trivial (base case), the DR-plan for $C$ is $C$ itself.
+    % \todo{delete everything above, move definition 6 here.}
+
+    % In the latter Case (b) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp.
+$C_j\setminus C_i$) (which are not rigid, by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}).
+
+    % Note:
+    % a canonical DR-plan has more than one node representing the same rigid subgraph if and only if Case (b) occurs.
+% \end{observation}
+
+Now we define the sequential DR-plan in a manner analogous to Definition~\ref{def:canonical_drp_rec}.
+
+\begin{definition}
+[(Pseudo-)Sequential DR-plan, appendage, partner, core]
+\label{def:seqdrp}
+    A \dfn{pseudo-sequential DR-plan} of $G$ is recursively defined as follows:
+    \begin{enumerate}
+        \item Base case: When $G$ is an edge, the sequential DR-plan for $G$ is $G$ itself.
+        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the sequential DR-plans for $C_i$.
+        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of sequential DR-plans for $C_j\setminus C_i$ (called an \dfn{appendage}), and $C_i$ (called its \dfn{partner}).
+        % If the decomposition of the partner $C_i$ results in Case (2), one of the children will be $C_i\cap C_j$, which we call the \dfn{core}.\footnote{Which can be thought of as partner to appendage $C_i\setminus C_j$.}\todo{Is this okay?
+I don't know how to say this succinctly.}
+    \end{enumerate}
+
+    % DR-plans satisfying Requirements 1,2, and 3 are called \dfn{pseudo-sequential.}
+
+    \dfn{Sequential DR-plans} must satisfy another requirement:
+
+    \begin{enumerate}
+        \setcounter{enumi}{3}
+        \item Let $C$ be a node and $C_s$ the set of its siblings in a sequential DR-plan.
+If there is a descendant $D$ of $C$, with siblings $D_s$  possessing the property that $C\cup C_s\setminus D_s$ is rigid, then for the contiguous sequence of nodes $D'$ on the path from $C$ to $D$, we require that $C\cup C_s \setminus D'_s$ be rigid, where $D'_s$ is the set of siblings of $D'$.
+Here $C$, $D$ and $D'$ are the partners of the appendages $C_s$, $D_s$ and $D'_s$.
+        %
+    \end{enumerate}
+
+    When the graph is independent and the DR-plan is sequential (Property (4) holds), if the parent of $C\cup C_s$ falls under Property (2), while $C\cup C_s$ falls under Property (3), then the lowest descendant $D$ as above is called the \dfn{core} of $C\cup C_s$.
+\end{definition}
+
+\begin{figure}\centering
+    \includegraphics[width=\linewidth]{../../img/svg/drp_triv_and_non}
+    \caption{Two sequential DR-plans of the same isostatic input graph. \todo{More caption.}}
+    \label{fig:seqdrp_nonuniqueness}
+\end{figure}
+
+\begin{remark}
+[Essential uniqueness of sequential DR-plans]
+\label{remark:seq_is_essentially_unique}
+    For sequential DR-plans of independent graphs, Property (3) of the above definition appears asymmetric with respect to $i$ and $j$; but in fact, $i$ and $j$ can be switched, using the appendage $C_i\setminus C_j$ instead.
+Let $C_1, C_2, \ldots, C_N$ be a complete list of proper vertex-maximal rigid subgraphs of $C$.
+Their pairwise intersections must all be nontrivial, and their common intersection, called a \dfn{core}, is isostatic by Lemma~\ref{lemma:combined_lemma}.
+Denote by $A_i$ the $i^\text{th}$ appendage, where $A_i = C_j\setminus C_i$, for any $j\neq i$.
+Note that $C_i = C\setminus A_i$; and $C_i = \bigcap_i{C_i} \bigcup_{j\ne i}{A_j}$.
+Choosing a particular ordering of the $C_i$, i.e.\ choosing the maximal rigid components of a particular appendage, say $A_1$, and its partner $C_1$ to be the children of $C$ simply pushes down the nodes corresponding to the appendages $A_2, A_3, \ldots, A_N$ to a lower level of the sequential DR-plan and the corresponding partners are created as $C_1\cap C_2, C_1\cap C_2\cap C_3, \ldots$; the last appendage, $A_N$, will always have the core, $\bigcap_i{C_i}$, as its partner.
+Thus, modulo the ordering of appendages, the sequential DR-plan is in fact unique.
+See Figure~\ref{fig:seqdrp_nonuniqueness}.
+
+    Furthermore, any rigid subgraph appears in at most one node of a sequential DR-plan.
+Also, if an edge $e$ of $G$ belongs in the rigid subgraph at any two nodes $A$ and $B$ of a sequential DR-plan, then either $A\subset B$ or $B\subset A$.
+Since the leaves of sequential DR-plan tree are the $O(|V(G)|)$ edges of the independent input graph, this implies the size of the tree is $O(|V(G)|)$.
+\end{remark}
+
+\begin{lemma}
+[Pseudo-sequential to sequential DR-plan]
+    Any pseudo-sequential DR-plan for an independent graph can be converted to a sequential DR-plan (satisfying Property (4)) in time $O(|V(G)|)$.
+\end{lemma}
+
+\begin{proof}
+    For independent graphs, note that Property (2) and (3) of Definition~\ref{def:seqdrp} automatically imply that the following holds for a pseudo-sequential DR-plan: for a node $G$ that falls under Property (2), there is no child $C$ of $G = C\cup C_s$ for which there is a descendant $D$ with siblings $D_s$ possessing the property that $C\cup C_s \setminus D_s$ is rigid; for a node $G$ that falls under Property (3), such a descendant $D$ exists for a unique child $C$ of the node $G=C\cup C_s$, as well as for unique children of all $D'$ on the path from $C$ to $D$.
+Furthermore, for all such $D'$, it automatically holds that $D'\cup D'_s \setminus D_s$ is rigid.
+We call this \emph{Property (4')} (since it is a parallel property of Property (4) of Definition~\ref{def:seqdrp}).
+
+    Given a pseudo-sequential DR-plan of an independent graph where nodes with Property (3) of Definition~\ref{def:seqdrp} are labeled, we show how to enforce Property (4) of Definition~\ref{def:seqdrp}, for a node $C\cup C_s$ with Property (3), i.e.\ when there is a descendant $D$ with siblings $D_s$ possessing the property that $C\setminus D_s \cup C_s$ is rigid.
+
+    Let $D$ be any such descendant where Property (4) does not hold and let $D^*$ (with siblings $D^*_s$) be the highest node along the path from $C$ to $D$ where $C\cup C_s \setminus D^*_s$ is not rigid.
+    Since we know that $C\cup C_s \setminus D_s$ is in fact rigid by the hypothesis of Property (4), and since Property (4') holds for $D^*$, i.e.\ $D^*\cup D^*_s \setminus D_s$, we ``switch $D^*_s$ with $D_s$'' by setting up the partners and appendages along the path from $D^*$ to $D$ as follows.
+
+    % At the level of the node $D^*$, the appendage is $D_s$ and the partner is $D^*\cup  D^*_s \setminus D_s$ (which we know to be rigid by Property (4)); At the level of each node $D'$ on the path starting from a child of $D^*$ down to $D$, the appendage is parent$(D')_s$ \todo{Is this correct?
+Or is it just $D'_s$?} and its partner is $D'\cup D'_s\setminus D_s$ (which we know to be rigid by Property (4')).
+    At the level of each partner node $D'$ on the path starting from $D$ to a child of $D^*$ the appendage becomes parent$(D')_s$ and the new partner becomes $D'\cup D'_s\setminus D_s$ (which we know to be rigid by Property (4')).
+In particular, at the level of the partner node $D$ the appendage becomes parent$(D)_s$ and the new partner remains $D$.
+However, at the next higher level the old partner node parent$(D) = D\cup D_s$ becomes the new partner $D\cup$ parent$(D)_s$ of the new appendage is parent$($parent$(D))_s$.
+
+    At the level of the node $D^*$, the appendage becomes $D_s$ and the partner becomes $D^*\cup D^*_s\setminus D_s$ (which we know to be rigid by Property (4)).
+
+    Once the above switch has been completed, moving $D_s$ up the path from $D$ to $C$,
+    % Once the above switch has been completed, moving $D_s$ up the path from $C$ to $D$,
+    it is unnecessary to inspect any of the siblings of any of the nodes along this path.
+This is because $C$ is the unique child of $C\cup C_s$ for which such a descendant $D$ even exists, and the same uniqueness holds for every node that is on the path from $C$ to $D$.
+
+    Note that in the same iteration, the algorithm can simultaneously perform the above process on all such descendants of $C$ for which Property (4) does not hold, since they must all be on the same descending path from $C$.
+Once Property (4) has been enforced for all such descendants of $C$, the algorithm has found the core for $C\cup C_s$, namely the last node in a contiguous path of nodes $D'$ starting down from $C$, for which $C\cup C_s\setminus D'_s$ is rigid.
+The order of appendages of all of these nodes are interchangeable (with appropriate partners) as indicated in Remark~\ref{remark:seq_is_essentially_unique}.
+
+    The algorithm then proceeds to the next node $C\cup C_s$ with Property (3), for which there is a descendant $D$ with siblings $D_s$ possessing the property that $C\setminus D_s \cup C_s$ is rigid and for which Property (4) does not hold.
+
+    This process continues until Property (4) always holds, resulting in a sequential DR-plan.
+It is clear that the algorithm visits any given node of the sequential DR-plan at most $O(1)$ times resulting in a linear time complexity.
+
+    The next observations follow from the abovementioned structure of a sequential DR-plan and from the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}.
+\end{proof}
+
+% \begin{observation}
+% [Sequential-canonical]
+%     A sequential DR-plan for an independent graph $G$ can be converted to a canonical DR-plan in time linear in the number of its nodes, which is $O(|V(G)|)$.
+\todo{This isn't true...
+in O(nodes) of which plan?...
+there could be $n^2$ nodes in the canonical...}
+% \end{observation}
+
+\begin{observation}
+[Sequential is optimal]
+    A \todo{psuedo-}sequential DR-plan of an independent graph is optimal.
+    \todo{(Psuedo-)sequential DR-plans of independent graphs are optimal.}
+\end{observation}
+
+\begin{proof}
+    We need to show the statement:
+    (*) the max fan-in (i.e.\ number of children) of any sequential DR-plan for an independent graph $G$ is no larger that of some canonical DR-plan for $G$.
+By Theorem~\ref{theorem:canonical_is_optimal}, this would imply that the sequential DR-plan is optimal.
+
+    Recall that both sequential and canonical DR-plans are defined recursively.
+We will prove the statement (*) by induction on the height of a sequential DR-plan for $G$.
+Induction hypothesis: (*) holds for independent graphs $G$ with sequential DR-plans of height $h$.
+Induction step: a sequential DR-plan of height $h+1$ rooted at a node $G$ consists of sequential DR-plans (of height at most $h$) rooted at its children $C$.
+    The base case (height of 0, i.e.\ single edges) trivially holds.
+
+    It is sufficient to show (a) the children $C$ will exist somewhere in some canonical DR-plan, and (b) $|C|$ is less than or equal to the max fan-in of some canonical DR-plan.
+Given (a), then by the recursive definition of canonical DR-plans, these nodes are the roots of canonical DR-plans, thus the induction hypothesis applies to the nodes in $C$.
+Additionally given (b), the proof of the induction step of (*) is complete.
+
+    Case (1):
+    When the isostatic vertex-maximal proper subgraphs of $G$ have trivial intersections, the children of $G$ in any canonical DR-plan are the same set $C$ of children in any sequential DR-plan.
+Thus both conditions (a) and (b) are immediately satisfied.
+
+    Case (2):
+    When the isostatic vertex-maximal proper subgraphs of $G$ have non-trivial intersections, the children $C$ are the isostatic components of some appendage, $A_i$, and its partner.
+The partner will be a child of $G$ in some canonical DR-plan.
+Furthermore, in any canonical DR-plan there will be some node $N-1$ levels below $G$ (where $N$ is the number of appendages) containing the core plus appendage $A_i$.
+In some canonical DR-plan, the children of this node will be the components of $A_i$ and the core.
+This node has the same fan-in as $G$ in the sequential DR-plan, satisfying (b), and, along with the partner (a child of $G$), shows that some canonical DR-plan has all the nodes in $C$, thereby satisfying (a).
+\end{proof}
+
+% \begin{proof}
+%     Any sequential DR-plan, $\seqdrpx{G}$, has maximum fan-in no greater than that of some canonical DR-plan, $\candrpx{G}$, of the same independent graph $G$.
+By Theorem~\ref{theorem:canonical_is_optimal}, this implies a sequential DR-plan is also optimal.
+Consider node $C$ that appears in both $\seqdrpx{G}$ and $\candrpx{G}$ with isostatic vertex-maximal proper subgraphs (clusters) $C_1, C_2, \ldots, C_N$.
+%     We prove this recursively by showing that node $C$ in $\seqdrpx{G}$ and any descendant of $C$ that is not in $\candrpx{G}$ has fan-in less than or equal to the maximum fan-in of $\candrpx{G}$.
+
+%     Case 1: If the clusters of $C$ have pairwise trivial intersections, then node $C$ has the same set of children and fan-in in both $\seqdrpx{G}$ and $\candrpx{G}$.
+
+%     Case 2: Otherwise, the clusters of $C$ have pairwise non-trivial intersection, and we call the common intersection the core.
+The children of node $C$ in $\seqdrpx{G}$ will be the components of the appendage, $A_1$, and its partner, $T_1$.
+In $\candrpx{G}$, there will be some node $D$ that is $A_1$ union the core; $D$ can have clusters with either trivial or non-trivial intersections.
+
+%     Case 2a: In the trivial case, node $D$ decomposes into the core and the components of $A_1$; thus, the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
+
+%     Case 2b: In the non-trivial case, there is some new core of $D$, so there will be some node $D'$ in $\candrpx{G}$ that is $A_1$ union the core of $D$; this pattern will continue until we eventually find $D^*$ with clusters that intersect trivially.
+Similar to Case (2b), the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D^*$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
+
+%     The decomposition of $T_1$ in $\seqdrpx{G}$ will be the components of appendage $A_2$ and its partner $T_2$.
+Similarly, $\candrpx{G}$ will contain core union $A_2$, and so we must now show $T_3$ is decomposed optimally in a sequential DR-plan.
+This continues to the node containing $T_{N-1}$, the core union appendage $A_N$, in $\seqdrpx{G}$, which we know to be a node in $\candrpx{G}$.
+% \end{proof}
+
+\begin{definition}
+[\Branch]
+    Given a sequential DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the leaf containing $e$ to the root containing $G$, together with the children of all the nodes on this path, which are the leaves of \branchGePG.
+\end{definition}
+
+\begin{observation}
+[Sequential DR-plan recursively from \branches]
+\label{obs:seqplan_rec_from_branches}
+    A sequential DR-plan $P_G$ for an isostatic graph $G$ is obtained from \branchGePG\ by recursively attaching to each of its leaves $L$ a sequential DR-plan $P_L$ for $L$.
+\end{observation}
+
+The next two Lemmas are the crux of our $O(|V|^3)$ algorithm to find a sequential DR-plan.
+
+\begin{lemma}
+[\Branch\ leaves from components]
+    Let $G$ be isostatic, $e$ be an edge in $G$, and $\componentsx{G\setminus e}$ be the set of maximal rigid components of $G\setminus e$.
+Then there is a sequential DR-plan $P_G$ for $G$ such that $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (minus $e$ itself).
+\end{lemma}
+
+\begin{proof}
+    Follows from the structure of a sequential DR-plan and the definition of \branchGePG.
+\end{proof}
+
+\ClearMyMinHeight
+\SetMyMinHeight{0.240}{../../img/svg/seq_proof_figs_a}
+\SetMyMinHeight{0.24}{../../img/svg/seq_proof_figs_b}
+\SetMyMinHeight{0.240}{../../img/svg/seq_proof_figs_c}
+\SetMyMinHeight{0.24}{../../img/svg/seq_proof_figs_d}
+
+\begin{figure*}\centering%
+    % \begin{subfigure}[t]{0.27\linewidth}
+    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_a}
+    %     \caption{}\label{fig:seq_proof:a}
     % \end{subfigure}%
@@
     % \hfill
-%   \begin{subfigure}{0.3\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_comdrp}
-%     \caption{}\label{fig:demo_graph:comdrp}
+    % \begin{subfigure}[t]{0.17\linewidth}
+    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_b}
+    %     \caption{}\label{fig:seq_proof:b}
     % \end{subfigure}%
@@
     % \hfill
-%   \begin{subfigure}{0.3\linewidth}\centering
-%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_candrp}
-%     \caption{}\label{fig:demo_graph:candrp}
+    % \begin{subfigure}[t]{0.25\linewidth}
+    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_c}
+    %     \caption{}\label{fig:seq_proof:c}
     % \end{subfigure}%
     % %
-%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section.
-(\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$.
-Dashed lines indicate that the children repeat the same pattern as the others shown on this level.
-The children of triangles (3 edges) are omitted.
-(\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$.
-The children of triangles or omitted.}
-%   \label{fig:demo_graph}
-% \end{figure*}%
+    % \hfill
+    % \begin{subfigure}[t]{0.23\linewidth}
+    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_d}
+    %     \caption{}\label{fig:seq_proof:d}
+    % \end{subfigure}%
+    \begin{subfigure}{0.240\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
+        \caption{}\label{fig:seq_proof:a}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.24\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
+        \caption{}\label{fig:seq_proof:b}
+    \end{subfigure}%
+    %
+    \begin{subfigure}{0.240\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
+        \caption{}\label{fig:seq_proof:c}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.24\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
+        \caption{}\label{fig:seq_proof:d}
+    \end{subfigure}%
 
-% \begin{figure*}\centering%
-%   %
-%   \includegraphics[width=0.3\linewidth]{../../img/epsfromtikz/demo_graph_candrp_seq}
-%   \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
-The children of the triangle are omitted.
-Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
-Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
-%   \label{fig:demo_graph:candrpseq}
-% \end{figure*}%
-\begin{figure*}\centering%
+    \caption{
+    Figure (\ref{fig:seq_proof:a}) depicts $\branchx{G,e,P_G}$, the black nodes are the $\componentsx{G\setminus e}$.
+The node labeled `Core+A$_1$+A$_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially.
     %
-  \includegraphics[width=0.3\linewidth]{../../img/svg/3xc2c3_candrp_seq}
-  \caption{The sequential canonical DR-plan of $G_{demo}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}).
-The children of the triangle are omitted.
-Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}.
-Also, note that the bottom-left node, the triangle, is the intersection of the 3 children of $G_{demo}$ in $ComDRP(G_{demo})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
-  \label{fig:demo_graph:candrpseq}
-\end{figure*}%
+    Figure (\ref{fig:seq_proof:b}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f\in L_2$ (from Figure (\ref{fig:seq_proof:a})).
+Similarly, the black nodes are the $\componentsx{G\setminus f}$.
+These leaves are labeled with the cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}.
+    %
+    Figure (\ref{fig:seq_proof:c}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the core in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
+Note that based on the choice of $e$ as the first edge, we get a sequential DR-plan that looks like Figure (\ref{fig:seq_proof:c}), if we had chosen an edge from appendage A$_2$ the second and third level would be flipped.
+    %
+    Figure (\ref{fig:seq_proof:d}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the appendage A$_1$ in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
+Note that this contains a node $D$ that has Case 3.
+With this $D$, $D\cup L_1$ is `Core+A$_1$+A$_2$' and $D\cap L_1$ is `Core'.
+The leaves labeled L$_1'$ in Figure (\ref{fig:seq_proof:a}) are the leaves that will allow us to find the node on the path from $G$ to $e$ that is their sibling (the parent of $e$).
+    %
+    % Take edge $a\in G$, the $\componentsx{G\setminus a}$ is the set of nodes marked with an `x' in figure xx.
+The entire figure shows $\branchx{G,a,P_G}$, where $P_G$ is the sequential DR-plan of $G$, however, we do not yet know the contents of the `circle' nodes or the order of the nodes in the \branch.
+    %
+    % So we take leaf $L$ (an element of $\componentsx{G\setminus a}$) and edge $b\in L$, and find $\components{G\setminus b}$.
+The this set is the leaves of the $\branchx{G,b,P_G}$, shown in figure xx.
+    }
+    \label{fig:seq_proof}
+\end{figure*}
 
-\begin{theorem}\label{theorem:algo_complexity}
-    There exists an \candrpcomplexityv\ algorithm to find a canonical DR-plan for an input graph.
+\begin{lemma}
+[\Branch\ from \branch\ leaves]
+\label{lem:branch-from-branch-leaves}
+    For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a sequential DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$, i.e.\ from the set of leaves of \branchGePG, by carrying out---for each leaf $L$---one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $L$.
+\end{lemma}
+
+\begin{proof}
+    First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $e$ to $G$ in $P_G$.
+Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
+
+    Now observe that for a leaf $L$ and an edge $f\in L$, each component $D\in \componentsx{G\setminus f}$ can be classified as one of the following cases.
+
+    Each case either discovers at least one node $C$ on the path from $e$ to $G$, and a sibling of $D$ in $P_G$, and partitions the leaves of \branchGePG\ into two sets, namely descendants of $C$ or children of ancestors of $C$; or finds a leaf of \branchLfPL\ ($\componentsx{L\setminus f}$ for the next recursion level of the algorithm).
+
+    \todo{
+    Case (0):
+    $D$ is exactly some component $C$ of $\componentsx{G\setminus e}$....
+The only true 1a is the leftmost in subfig 2, everything else is 0...
+In this case, we know $C=D$ to be some child of an ancestor of $C$.
+    }
+
+    Case (1):
+    $D$ contains some components $C$ of $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) and intersects all leaves of \branchGePG\ trivially \todo{besides the components $C$ it contains}.
+One occurrence of this case (Case 1a) is when $L$'s siblings in (any sequential DR-plan) $P_G$ have trivial pairwise intersection.
+Another occurrence of this case (Case 1b) is when $L$'s siblings in (any sequential DR-plan) $P_G$ have non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core.
+In both cases, $D$ is a node along the path from $e$ to $G$ in $P_G$, with the components $C$ being its descendants, $L$ being its sibling and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
+
+    Case (2):
+    $D$ is contained in $L$ and intersects all \todo{other} leaves of \branchGePG\ trivially. In this case, $D$ is a leaf of \branchLfPL, to be used in the next level recursion of the algorithm.
+
+    Case (3):
+    $D$ intersects $L$ non-trivially, neither contains nor is contained in $L$, but contains some components $C$ of $\componentsx{G\setminus e}$ (leaves of \branchGePG).
+This happens when the siblings of $L$ in a canonical DR-plan for $G$ intersect $L$ non-trivially.
+In this case, $D\cup L$ is a node along the path from $e$ to $G$ with the components $C$ being its descendants, $L$ being its  child and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
+In addition, $D\cap L$ is a leaf of \branchLfPL\ to be used in the next recursion level of the algorithm.
+
+    One subtle obstacle to overcome in Case 3 is that the newly found node along the path from $e$ to $G$ is the parent of $L$ as opposed to a sibling of $L$ as in Case (1).
+On the surface, this is problematic because the sibling $D'$ of $L$ along the path from $e$ to $G$ may never be found.
+However a closer inspection of Case 3 reveals that $L$ is the partner of an appendage $A$ containing $e$, and $A$, being underconstrained by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}, must have maximal rigid components with nontrivial intersections, which means that there must be another sibling of $D'$ that is a leaf $L'$ of \branchGePG.
+Hence for some edge $f'$ in $L'$, $\componentsx{G\setminus f'}$ will find $D'$ within Case 1.
+
+    Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
+\end{proof}
+
+\begin{theorem}
+[Complexity of the algorithm]
+\label{theorem:algo_complexity}
+    Computing a sequential DR-plan for a graph $G$ has time complexity $O(|V(G)|^3)$.
 \end{theorem}
@@
 \begin{proof}
-We first describe the algorithm.
-The first step of the algorithm, which we call $CanDRP(G)$, is finding the isostatic vertex-maximal proper subgraphs (clusters) of the input isostatic graph $G$.
-One of many ways to do this is  by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of cluster-candidates.
-The next step is to isolate the set of true-clusters by running the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, to find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, the candidate is moved (before adding $e$) to the true-cluster list and removed from the candidate-cluster list.
-If $D\subsetneq G$, all candidates that are subgraphs of $D$ are removed from the candidate-cluster list and $D$ is added back to the list.
-The next candidate is considered until all candidates are exhausted.
-
-The next step is to check the intersection of any two of the true-components.
-If the intersection is trivial, the entire list becomes the children, and the algorithm is recursively applied to each child.
-If the intersection is non-trivial (necessarily isostatic), the  decomposition is computed down to $I=\bigcap_{k=1}^{N}{C_k}$.
-(This is discussed in detail in the proof of Theorem \ref{theorem:main}).
-We construct a DR-plan whose node set is that of a canonical/optimal DR-plan.
-Begin with the parent $C$.
-Its children are $S_1$ and $CanDRP(T_1)$, where $S_1=C\setminus R_1$ and $T_1$ is $R_1$, together with all incident edges.
-Observe that these edges are incident on vertices from $I$ and that $T_1$ is underconstrained so its DR-plan is a forest where each root becomes a child of $C$.
-The children of $S_1$ are $S_2$ and $CanDRP(T_2)$, where $S_2=C\setminus (R_1\cup R_2)$ and $T_2$ is $R_2$ together with all incident edges.
-The process iterates.
-At $N$ levels down, the children are  $S_N=I$ and $CanDRP(T_N)$, where $T_N$ is $R_N$ together with all incident edges.
-Apply the algorithm recursively to both children.
-Thus we obtain an efficient method for the case of DR-plan nodes whose rigid proper vertex-maximal subgraphs have non-trivial intersections, avoiding recomputation of the isostatic components of $C$.
-See Figure~\ref{fig:demo_graph:candrpseq} for an example.
-
-To summarize the above step of the algorithm:
-if the intersection is trivial, recursively apply $CanDRP$ to the entire list of true-components and set the resulting trees as the children.
-Else, the intersection is non-trivial, and the children of the node are $CanDRP(C_1)$ and $CanDRP(T)$, where $C_1$ is the first true-component of the parent $C$ (chosen arbitrarily, could be any child) and $T$ is the underconstrained graph formed by $C\setminus C_1$ together with all incident edges  and the associated nodes in $C_1$.
-
-% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest
-
-The DR-plan  output by this algorithm is a tree where each node is a distinct subgraph.
-The leaves of this tree are the edges of the starting graph $G$.
-Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$ and the complexity of the algorithm is $O(|V|^3)$.
-%
-% \note In the best case, this can be $O(V^2)$ if the graph is a 2-tree
+    The previous two Lemmas and the $O(|V(G)|^2)$ complexity of the pebble game algorithm for computing $\componentsx{G\setminus e}$ show that the procedure for computing \branchGePG\ given $G$ and $e$ takes time $O(M|V(G)|^2)$, using $M$ $\componentsx{}$ computations, where $M$ is the number of leaves $L$ of \branchGePG.
+Furthermore, in the process, the leaves of all the \branches\ \branchLfPL\ have already been computed.
+Recursive computation of a sequential DR-plan $P_G$ as in Observation~\ref{obs:seqplan_rec_from_branches} now proceeds by computing \branches\ \branchLfPL\ for each leaf $L$ of \branchGePG.
+Since each node of $P_G$ appears as the leaf $L$ of a \branch\ exactly once during the above recursive procedure, overall one $O(|V(L)|^2)$ computation of $\componentsx{}$ is carried out for each of the $O(|V(G)|)$ nodes $L$ of $P_G$, resulting in an overall complexity of $O(|V(G)|^3)$.
 \end{proof}
 
-% \begin{proof}
-% The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$.
-Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$.
-The output of this will be a list of component-candidates.
-Then run the \frontier\ algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$.
-If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list.
-If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children.
-Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$.
-Continue with the next candidate.
-If you exhaust the candidates, output the entire true-component list.
-
-% %on the children
-% This is run recursively on each node.
-If you use dynamic programming to store the decompositions of each component, it prevents repetition of work (done by using the DAG representation of the DR-plan).
-In this case you will have to do work on no more than $O(|V|)$ nodes that get smaller at each level.
-
-% Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
-
-% It's a tree in only trivial case -> O(V), always order of leaves
-% If its isostatic intersection, we do the sequential extension,intersections only done once.
-
-% inductively. The number of nodes at this level will e more than the total. Number of leaves (terminals for dag) is at most the number of edges
-
-% \end{proof}
-
-% \begin{proof}
-% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the wellconstrained vertex-maximal proper subgraphs of the input wellconstrained graph.
-There are two approaches: (1) the bottom-up approach, using the \frontier\ algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones.
-After the subgraphs have been found, it is a straightforward application of the theory.
-If the intersection of any two of the subgraphs is wellconstrained, then take those two to be children of the graph.
-Otherwise, all the subgraphs are the children.
-Apply recursively to the children.
-This naturally terminates when you reach trivial subgraphs.
-The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time).
-\todo{Think more about complexity.}
-% \end{proof}
+\subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
+\label{sec:drp:overconstrained}
 
@@
 
-\subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
-\label{sec:drp:overconstrained}
-
 For overconstrained (not independent) graphs, a canonical DR-plan is still well-defined.

### 2015-08-23 doc/GMP2015/revision_3/s04_recombination.tex
Lots of work on Revision 3, getting close.

@@
 
-\begin{theorem}\label{theorem:convexcayley}
+\begin{theorem}
+\label{theorem:convexcayley}
     \cite{sitharam2010convex} $H$ has a convexifiable Cayley configuration space with parameters $F$ if and only if for each $f\in F$ all the minimal 2-sum components of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees.
@@
 
-\begin{theorem}\label{theorem:omdk}
+\begin{theorem}
+\label{theorem:omdk}
     Given an indecomposable graph $G$, let $G'$ be a spanning partial 2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then $(G,2)$ belongs in the set OMD$_k$.
@@
 
-\begin{theorem*}[Efficient search]
+\begin{theorem*}
+[Efficient search]
 \label{theorem:criterionc}
@@
 
+% \ClearMyMinHeight
+% \SetMyMinHeight{.2}{../../img/epsfromtikz/omd_k33_example_new-0}
+% \SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-1}
+% \SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-2}
+% \SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-1}
+% \SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-2}
 \ClearMyMinHeight
-\SetMyMinHeight{.2}{../../img/epsfromtikz/omd_k33_example_new-0}
-\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-1}
-\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example_new-2}
-\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-1}
-\SetMyMinHeight{.15}{../../img/epsfromtikz/omd_k33_example-2}
+\SetMyMinHeight{.08}{../../img/svg/k33_omd_a}
+\SetMyMinHeight{.08}{../../img/svg/k33_omd_b}
+\SetMyMinHeight{.08}{../../img/svg/k33_omd_c}
+\SetMyMinHeight{.08}{../../img/svg/k33_omd_d}
+\SetMyMinHeight{.08}{../../img/svg/k33_omd_e}
 
 \begin{figure*}\centering
-    \begin{subfigure}{.2\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-0}
+    \hspace{1cm}
+    \begin{subfigure}{.08\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/k33_omd_a}
         \caption{}\label{fig:omd_k33_example:a}
     \end{subfigure}\hfill
-    \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-1}
+    \begin{subfigure}{.08\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/k33_omd_b}
         \caption{}\label{fig:omd_k33_example:b}
-    \end{subfigure}
-    \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example_new-2}
+    \end{subfigure}\hspace{1cm}
+    \begin{subfigure}{.08\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/k33_omd_c}
         \caption{}\label{fig:omd_k33_example:c}
     \end{subfigure}\hfill
-    \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example-1}
+    \begin{subfigure}{.08\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/k33_omd_d}
         \caption{}\label{fig:omd_k33_example:d}
-    \end{subfigure}
-    \begin{subfigure}{.15\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/omd_k33_example-2}
+    \end{subfigure}\hspace{1cm}
+    \begin{subfigure}{.08\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/k33_omd_e}
         \caption{}\label{fig:omd_k33_example:e}
     \end{subfigure}
+    \hspace{1cm}
 
@@
 
-% \noindent
-% \note A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.
-
-\sidenote{A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals of values rather than exact values~~\cite{sitharam2010convex}.}
-%cite my paper with Heping!!!
+Note that a major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals of values rather than exact values~~\cite{sitharam2010convex}.
 

### 2015-08-23 doc/GMP2015/revision_3/s06_problempinning.tex
Lots of work on Revision 3, getting close.

@@
 \begin{definition}
+[Body-hyperpin qusecs]
     A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies and the constraints are incidences of object subsets at a common point, i.e.\ a pinning of bodies.
@@
 
-\begin{remark*}\label{rem:bodypin_is_barjoint}
+\begin{remark*}
+\label{rem:bodypin_is_barjoint}
     A body-hyperpin qusecs is a special case of bar-joint qusecs of the previous sections of the paper.
@@
 
-\begin{definition}\label{def:body-pin}
-    A body-pin graph is a body-hyperpin graph with the following conditions:
+\begin{definition}
+[Body-pin graph]
+\label{def:body-pin}
+    A \dfn{body-pin graph} is a body-hyperpin graph with the following conditions:
     (1) each pin is shared by at most two bodies; and
@@
 \begin{definition}
+[Triangle-hyperpin graph]
     A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e.\ it shares  pins with at most 3 other bodies.
@@
 \begin{definition}
+[$(k,l)$-tight DR-plan]
     A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight subgraph of the parent node or it is trivial.
@@
 %
-\begin{observation*}\label{rem:1dofcanon}
+\begin{observation*}
+\label{rem:1dofcanon}
     When the input is independent, all $(3,4)$-tight canonical DR-plans are optimal.

### 2015-08-23 doc/GMP2015/revision_3/s07_pinnedline.tex
Lots of work on Revision 3, getting close.

@@
 \begin{definition}
+[Pinned line-incidence system]
     A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$ together with parameters $\delta$ specifying $|E|$ \dfn{pins} with fixed positions in $\mathbb{R}^2$, such that each edge is constrained to lie on a line passing through the corresponding pin, i.e.\ $\delta: E \rightarrow \mathbb{R}^2$.
@@
 \begin{definition}
+[DR-plans of pinned line-incidence systems]
     The \dfn{DR-plan} of a pinned line-incidence graph $G$ is one in which (1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$, or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
@@
 
-\begin{observation}\label{lem:pinned_union_intersection}
+\begin{observation}
+\label{lem:pinned_union_intersection}
     Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$, where each of them can be either a single edge or a connected isostatic subgraph. There are only two possible cases:

### 2015-08-28 doc/GMP2015/revision_3/s01_introduction.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.
Lots of work on Revision 3, getting close.
Revision 3, close to finished.

--- 2015-05-17 doc/GMP2015/revision_2/s01_introduction.tex
@@
 Na{\"i}vely, the optimal DR-plan is used as follows.
-Each decomposed subsystem---a node of the DR-plan---is treated and solved as a polynomial system of constraints between its child subsystems.
+Each decomposed subsystem -- a node of the DR-plan -- is treated and solved as a polynomial system of constraints between its child subsystems.
 However, even in an optimal DR-plan, there can be arbitrarily many children at a node.
@@
 These properties can be leveraged to further reduce the complexity of both optimal DR-plan construction and recombination.
-We consider 3 different types of constraint systems---which we collectively call \dfn{qusecs}---that are used to model, design, and analyze quasi-uniform or self-similar materials.
+We consider 3 different types of constraint systems -- which we collectively call \dfn{qusecs} -- that are used to model, design, and analyze quasi-uniform or self-similar materials.
 In the remainder of this section, we motivate the materials application and give the contributions and organization of the paper.
@@
 % \FigThreeSubfig%
-%   {img/Chlamydomonas_TEM_17}
+%   {../../img/Chlamydomonas_TEM_17}
 %   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
@@
 %   %
-%   {img/ligten2}
+%   {../../img/ligten2}
 %   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
@@
 %   %
-%   {img/Rothemund-DNA-SierpinskiGasket}
+%   {../../img/Rothemund-DNA-SierpinskiGasket}
 %   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.32}{img/Chlamydomonas_TEM_17}
-\SetMyMinHeight{.32}{img/ligten2}
-\SetMyMinHeight{.32}{img/Rothemund-DNA-SierpinskiGasket}
+\SetMyMinHeight{.32}{../../img/Chlamydomonas_TEM_17}
+\SetMyMinHeight{.32}{../../img/ligten2}
+\SetMyMinHeight{.32}{../../img/Rothemund-DNA-SierpinskiGasket}
 
@@
   \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/Chlamydomonas_TEM_17}
+    \includegraphics[height=\myMinHeight]{../../img/Chlamydomonas_TEM_17}
     \caption{}\label{fig:material_examples:microtubule}
@@
   \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/ligten2}
+    \includegraphics[height=\myMinHeight]{../../img/ligten2}
     \caption{}\label{fig:material_examples:tendon}
@@
   \begin{subfigure}{0.32\linewidth}\centering
-    \includegraphics[height=\myMinHeight]{img/Rothemund-DNA-SierpinskiGasket}
+    \includegraphics[height=\myMinHeight]{../../img/Rothemund-DNA-SierpinskiGasket}
     \caption{}\label{fig:material_examples:sierpinski}
@@
 %   {0.3}
-%   {img/Chlamydomonas_TEM_17}
+%   {../../img/Chlamydomonas_TEM_17}
 %   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
@@
 %   {0.3}
-%   {img/ligten2}
+%   {../../img/ligten2}
 %   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
@@
 %   {0.3}
-%   {img/Rothemund-DNA-SierpinskiGasket}
+%   {../../img/Rothemund-DNA-SierpinskiGasket}
 %   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
@@
 
-% Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
+% Also in Section \ref{sec:DRP}, we give an efficient (\ComplexityCanDRP) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
 % The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
@@
 Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied \dfn{cluster minimality}~\cite{hoffman2001decompositionI}.
-Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
+Also in this section, an \ComplexityCanDRP\ time algorithm is provided to find an optimal DR-plan for independent bar-joint graphs.
 While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
 % that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
-% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
+% Additionally, a polynomial time (\ComplexityCanDRP) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.
 

### 2015-08-28 doc/GMP2015/revision_3/s02_preliminaries.tex
Revision 3, close to finished.

@@
   More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
-  For a given graph, there could be exponentially many DR-plans---and even optimal DR-plans---in the size of the graph.
+  For a given graph, there could be exponentially many DR-plans -- and even optimal DR-plans -- in the size of the graph.
 A complete DR-plan is unique but may not be (and is usually not) optimal.
@@
 
-See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \todo{we lost this figure, replaced with \ref{fig:seqdrp_nonuniqueness}, should it be cited here now or should the other be added back in somewhere?} \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
+See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
 

### 2015-08-28 doc/GMP2015/revision_3/s03_drp.tex
Revision 3, close to finished.

@@
 \newcommand{\seqdrpx}[1]{P_{#1}^{seq}}
+
+\newcommand{\pseudosequential}{pseudosequential}
+\newcommand{\Pseudosequential}{Pseudosequential}
 
@@
 
-\subsection{Canonical DR-Plan}
-In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
-Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
-% The definition is as follows:
+The goal of this section is to develop an \ComplexityCanDRP\ time complexity algorithm for finding an optimal DR-plan.
+To this end, we first introduce a canonical DR-plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
+While useful for proving optimality, the canonical DR-plan is difficult to work with algorithmically.
+Therefore, we define the \pseudosequential\ DR-plan which is derived from the canonical and is still optimal.
+However, the \pseudosequential\ DR-plan lacks the essential uniqueness of the complete DR-plan.
+From the \pseudosequential, we derive the sequential DR-plan, which can be found in the same time complexity and is essentially unique.
 
-In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+In this section and in Section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
 Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$.
@@
 
-% When using intersections and unions of graphs, we are always considering subgraphs of a single parent graph.
+\subsection{Canonical DR-Plan}
+
+% In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
+Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
+% % The definition is as follows:
+
+% In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$.
 In this case unions and intersections are well defined.
-For example, the union of $G_1=(V_1, E_1)$ and $G_2=(V_2, E_2)$ is the subgraph $(V_1\cup V_2, E_1\cup E_2)$; similarly, $G_1\cap G_2$ is the subgraph $(V_1\cap V_2, E_1\cap E_2)$.
-
-% \begin{definition}\label{def:canonical_drplan}
-%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
-%     (1) it is a DR-plan of $G$;
-%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
-%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
-% \end{definition}
+For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).
 
@@
 \end{definition}
+
+% \begin{definition}\label{def:canonical_drplan}
+%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
+%     (1) it is a DR-plan of $G$;
+%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
+%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
+% \end{definition}
 
@@
 
-    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $\drpx{A}$ has the minimum possible fan-in of 2 \todo{Shouldn't it be `node $A$ in $\drpx{C}$ has the minimum...'}.
+    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) all the ancestors $A$ of $S$ that are descendants of $C$ have the minimum possible fan-in of 2.
 
@@
 
-    \medskip\noindent
-    \vemph{Claim 1:}
+    \begin{proof}[Proof of Claim 1]
         Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If the pairwise intersection of clusters is trivial, all of the clusters must be children of $C$ in an optimal DR-plan
@@
 Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.
+    \end{proof}
 
-    \medskip\noindent
-    \vemph{Claim 2:}
+    % \medskip\noindent
+    % \vemph{Claim 1:}
+
+    \begin{proof}[Proof of Claim 2]
         Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If some pair of clusters
@@
         Let $I$ denote the intersection of all the clusters; we call this the \dfn{core}.
-Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
+Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:pseudoseqdrp} and~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
         % To prove this holds,
@@
 The pairwise intersection of any of these clusters of $C_i$ will clearly be isostatic, so any two of them are viable children of node $C_i$.
-     % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
+         % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs.
+So any two of these are viable children for $C_i$.
         % Since
@@
         Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for $N-1$ levels, and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
+    \end{proof}
+    %
+    % \medskip\noindent
+    % \vemph{Claim 2:}
     %
@@
     % -- say this at top `proof is by induction on the level of the dr=plan'
+    % The proofs of Claim 1 and Claim 2 prove the theorem.
 \end{proof}
 
-This proof of this theorem relies on the following crucial observation and lemma.
+This proof of the theorem relies on the following crucial observation and lemma.
 These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.
@@
         \item\label{lemma:wc_intersection_makes_all_wc}
-        If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
+        If $C_i\cup C_j$ is isostatic, then for all $k$, $C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then for all $k$, $C_i\cup C_k=C$.
 
         \item\label{lemma:uc_intersection_makes_all_uc}
-        If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
+        If $C_i\cap C_j$ is trivial, then for all $k$, $C_i\cap C_k$ is trivial.
     \end{enumerate}
@@
 
-The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}.
+% The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}.
 We list these properties before describing the algorithm.
@@
 
+The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}.
+We begin by redefining the canonical DR-plan in a recursive manner and also by recursively defining the new \dfn{\pseudosequential\ DR-plan} to highlight the similarities.
+We show that this \pseudosequential\ DR-plan is optimal (i.e.\ has the smallest maximum fan-in over all nodes) and has at most the same overall size (i.e.\ number of unique nodes) as a canonical DR-plan, and is in general smaller.
+% By taming the large number of nodes with redundant subgraphs (in the case when the children of a node have non-trivial intersections),
+By ensuring that no two nodes contain the same subgraph,
+the \pseudosequential\ DR-plan is more malleable for algorithmic purposes.
+
 \begin{definition}
@@
 \label{def:canonical_drp_rec}
-    A \dfn{canonical DR-plan} of $G$ is recursively defined as follows:
+    A \dfn{canonical DR-plan} of an isostatic $G$ is recursively defined as follows:
     \begin{enumerate}
@@
 \end{definition}
-\todo{What if input is not isostatic? What roots?}
+% \todo{What if input is not isostatic? What roots?}
 
-\begin{observation}
-[Canonical DR-plan size]
-    While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes).
+% \begin{observation}
+% [Canonical DR-plan size]
+%     While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes).
 If we want the canonical DR-plan to have small size (linear \todo{or just polynomial?} in the number of vertices) we can choose the children in the following fashion: In Case (3) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp.
@@
 
-    Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
-\end{observation}
+%     Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
+% \end{observation}
 
@@
 
-Now we define the sequential DR-plan in a manner analogous to Definition~\ref{def:canonical_drp_rec}.
+Now we define the \pseudosequential\ DR-plan in a manner analogous to Definition~\ref{def:canonical_drp_rec}.
 
 \begin{definition}
-[(Pseudo-)Sequential DR-plan, appendage, partner, core]
+[\Pseudosequential\ DR-plan, appendage, partner]
+\label{def:pseudoseqdrp}
+    A \dfn{\pseudosequential\ DR-plan} of an isostatic $G$ is recursively defined as follows:
+    \begin{enumerate}
+        \item Base case: When $G$ is an edge, the \pseudosequential\ DR-plan for $G$ is $G$ itself.
+        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the \pseudosequential\ DR-plans for $C_i$.
+        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of \pseudosequential\ DR-plans for $C_j\setminus C_i$ (called an \dfn{appendage}), and $C_i$ (called its \dfn{partner}).
+    \end{enumerate}
+\end{definition}
+
+\begin{remark}
+[Size of \pseudosequential\ DR-plan]
+    Any rigid subgraph appears in at most one node of a \pseudosequential\ DR-plan.
+Also, if an edge $e$ of $G$ belongs in the rigid subgraph at any two nodes $A$ and $B$ of a \pseudosequential\ DR-plan, then either $A\subset B$ or $B\subset A$.
+Since the leaves of \pseudosequential\ DR-plan tree are the $O(|V(G)|)$ edges of the independent input graph, this implies the size of the DR-plan is $O(|V(G)|)$.
+\end{remark}
+
+\begin{observation}
+[Sequential is optimal]
+    Any \pseudosequential\ DR-plan of an independent graph is optimal.
+\end{observation}
+
+\begin{proof}
+    We need to show the statement:
+    (*) the max fan-in (i.e.\ number of children) of any \pseudosequential\ DR-plan for an independent graph $G$ is no larger that of some canonical DR-plan for $G$.
+By Theorem~\ref{theorem:canonical_is_optimal}, this would imply that the \pseudosequential\ DR-plan is optimal.
+
+    Recall that both \pseudosequential\ and canonical DR-plans are defined recursively.
+We will prove the statement (*) by induction on the height of a \pseudosequential\ DR-plan for $G$.
+The base case (height of 0, i.e.\ single edges) trivially holds.
+Induction hypothesis: (*) holds for independent graphs $G$ with \pseudosequential\ DR-plans of height $h$.
+
+    Induction step: a \pseudosequential\ DR-plan of height $h+1$ rooted at a node $G$ consists of \pseudosequential\ DR-plans (of height at most $h$) rooted at the children of $G$, which we call the set $C$.
+It is sufficient to show (a) the children $C$ will exist somewhere in some canonical DR-plan, and (b) $|C|$ is less than or equal to the max fan-in of some canonical DR-plan.
+Given (a), then by the recursive definition of canonical DR-plans, these nodes are the roots of canonical DR-plans, thus the induction hypothesis applies to the nodes in $C$.
+Additionally given (b), the proof of the induction step of (*) is complete.
+
+    Case (1):
+    When the isostatic vertex-maximal proper subgraphs of $G$ have trivial intersections, the children of $G$ in any canonical DR-plan are the same set $C$ of children in any \pseudosequential\ DR-plan.
+Thus both conditions (a) and (b) are immediately satisfied.
+
+    Case (2):
+    When the isostatic vertex-maximal proper subgraphs of $G$ have non-trivial intersections, the children $C$ are the isostatic components of some appendage, $A_i$, and its partner.
+The partner will be a child of $G$ in some canonical DR-plan.
+Furthermore, in any canonical DR-plan there will be some node $N-1$ levels below $G$ (where $N$ is the number of appendages) containing the core (the common intersection of all the isostatic vertex-maximal proper subgraphs of $G$) plus appendage $A_i$.
+In some canonical DR-plan, the children of this node will be the components of $A_i$ and the core.
+This node has the same fan-in as $G$ in the \pseudosequential\ DR-plan, satisfying (b), and, along with the partner (a child of $G$), shows that some canonical DR-plan has all the nodes in $C$, thereby satisfying (a).
+\end{proof}
+
+% \begin{proof}
+%     Any sequential DR-plan, $\seqdrpx{G}$, has maximum fan-in no greater than that of some canonical DR-plan, $\candrpx{G}$, of the same independent graph $G$.
+By Theorem~\ref{theorem:canonical_is_optimal}, this implies a sequential DR-plan is also optimal.
+Consider node $C$ that appears in both $\seqdrpx{G}$ and $\candrpx{G}$ with isostatic vertex-maximal proper subgraphs (clusters) $C_1, C_2, \ldots, C_N$.
+%     We prove this recursively by showing that node $C$ in $\seqdrpx{G}$ and any descendant of $C$ that is not in $\candrpx{G}$ has fan-in less than or equal to the maximum fan-in of $\candrpx{G}$.
+
+%     Case 1: If the clusters of $C$ have pairwise trivial intersections, then node $C$ has the same set of children and fan-in in both $\seqdrpx{G}$ and $\candrpx{G}$.
+
+%     Case 2: Otherwise, the clusters of $C$ have pairwise non-trivial intersection, and we call the common intersection the core.
+The children of node $C$ in $\seqdrpx{G}$ will be the components of the appendage, $A_1$, and its partner, $T_1$.
+In $\candrpx{G}$, there will be some node $D$ that is $A_1$ union the core; $D$ can have clusters with either trivial or non-trivial intersections.
+
+%     Case 2a: In the trivial case, node $D$ decomposes into the core and the components of $A_1$; thus, the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
+
+%     Case 2b: In the non-trivial case, there is some new core of $D$, so there will be some node $D'$ in $\candrpx{G}$ that is $A_1$ union the core of $D$; this pattern will continue until we eventually find $D^*$ with clusters that intersect trivially.
+Similar to Case (2b), the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D^*$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
+
+%     The decomposition of $T_1$ in $\seqdrpx{G}$ will be the components of appendage $A_2$ and its partner $T_2$.
+Similarly, $\candrpx{G}$ will contain core union $A_2$, and so we must now show $T_3$ is decomposed optimally in a sequential DR-plan.
+This continues to the node containing $T_{N-1}$, the core union appendage $A_N$, in $\seqdrpx{G}$, which we know to be a node in $\candrpx{G}$.
+% \end{proof}
+
+\begin{definition}
+[\Branch]
+    Given a \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the leaf containing $e$ to the root containing $G$, together with the children of all the nodes on this path, which are the leaves of \branchGePG.
+See Figure~\ref{fig:seq_proof} for examples.
+\end{definition}
+
+\begin{observation}
+[\Pseudosequential\ DR-plan recursively from \branches]
+\label{obs:seqplan_rec_from_branches}
+    A \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$ is obtained from \branchGePG\ by recursively attaching to each of its leaves $L$ a \pseudosequential\ DR-plan $P_L$ for $L$.
+\end{observation}
+
+The next two Lemmas are the crux of our $O(|V|^3)$ algorithm to find a \pseudosequential\ DR-plan.
+
+\begin{lemma}
+[\Branch\ leaves from components]
+    Let $G$ be isostatic, $e$ be an edge in $G$, and $\componentsx{G\setminus e}$ be the set of maximal rigid components of $G\setminus e$.
+Then there is a \pseudosequential\ DR-plan $P_G$ for $G$ such that $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (minus $e$ itself).
+\end{lemma}
+
+\begin{proof}
+    Follows from the structure of a \pseudosequential\ DR-plan and the definition of \branchGePG.
+\end{proof}
+
+% \ClearMyMinHeight
+% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_a}
+% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_b}
+% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_c}
+% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_d}
+% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_e}
+\ClearMyMinHeight
+\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_a}
+\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_b}
+\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_c}
+\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_d}
+\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_e}
+
+\begin{figure*}\centering%
+    % \begin{subfigure}[t]{0.27\linewidth}
+    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_a}
+    %     \caption{}\label{fig:seq_proof:a}
+    % \end{subfigure}%
+
+    % \begin{subfigure}{0.195\linewidth}\centering
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
+    %     \caption{}\label{fig:seq_proof:a}
+    % \end{subfigure}%
+    % %
+    % \hfill
+    % \begin{subfigure}{0.195\linewidth}\centering
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
+    %     \caption{}\label{fig:seq_proof:b}
+    % \end{subfigure}%
+    % %
+    % \begin{subfigure}{0.195\linewidth}\centering
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
+    %     \caption{}\label{fig:seq_proof:c}
+    % \end{subfigure}%
+    % %
+    % \hfill
+    % \begin{subfigure}{0.195\linewidth}\centering
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
+    %     \caption{}\label{fig:seq_proof:d}
+    % \end{subfigure}%
+    % %
+    % \hfill
+    % \begin{subfigure}{0.195\linewidth}\centering
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_e}
+    %     \caption{}\label{fig:seq_proof:e}
+    % \end{subfigure}%
+
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
+        \caption{}\label{fig:seq_proof:a}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
+        \caption{}\label{fig:seq_proof:b}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
+        \caption{}\label{fig:seq_proof:c}
+    \end{subfigure}%
+
+    % \hspace{1cm}
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
+        \caption{}\label{fig:seq_proof:d}
+    \end{subfigure}%
+    %
+    \hspace{2cm}
+    \begin{subfigure}{0.3\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_e}
+        \caption{}\label{fig:seq_proof:e}
+    \end{subfigure}%
+    % \hspace{1cm}
+
+    \caption{
+    This figure illustrates the notion of a \branch, how it relates to the components of $G\setminus e$, and the variety of cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}.
+Each subfigure is $\branchx{G,e,P_G}$ for some edge $e$ (resp.\ $\branchx{G,f,P_G}$), the white nodes are the path from $G$ to $e$ (resp.\ $f$), and the black nodes are the set $\componentsx{G\setminus e}$ (resp.\ $G\setminus f$).
+In subfigures (\ref{fig:seq_proof:b}), (\ref{fig:seq_proof:c}), and (\ref{fig:seq_proof:d}), nodes are labeled with the case they fall under in the proof of Lemma~\ref{lem:branch-from-branch-leaves} if they are chosen to be $D$ (given the choice of $L$ discussed below), any unlabeled node falls under Case (1b).
+    %
+    Figure (\ref{fig:seq_proof:a}) has nodes labeled $L_1$ and $L_2$ ($L_1'$ will be discussed later), from which different edges $f$ are taken in subsequent figures.
+Note that the node labeled `Core+$A_1$+$A_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially.
+The core is the intersection of these subgraphs and $A_i$ is an appendage.
+    %
+    Figure (\ref{fig:seq_proof:b}) uses $f=L_1$.
+    %
+    Figure (\ref{fig:seq_proof:c}) uses an arbitrary $f\in L_2$.
+    %
+    Figure (\ref{fig:seq_proof:d}) uses an $f$ in the core in $L_3$.
+    % Note that based on the choice of $e$ as the first edge, we get a \pseudosequential\ DR-plan that looks like Figure (\ref{fig:seq_proof:c}), if we had chosen an edge from appendage $A_2$ the second and third level would be flipped.
+\todo{Shows cases...}
+    %
+    Figure (\ref{fig:seq_proof:e}) uses an $f$ in the appendage $A_1$ in $L_3$.
+Note that this contains a node $D$ that has Case (3).
+With this $D$, $D\cup L_3$ is `Core+$A_1$+$A_2$' and $D\cap L_3$ is `Core'.
+The leaves labeled $L_3'$ in Figure (\ref{fig:seq_proof:a}) are the leaves that will allow us to find the node on the path from $G$ to $e$ that is their sibling (the parent of $e$).
+    %
+    % Figure (\ref{fig:seq_proof:a}) depicts $\branchx{G,e,P_G}$, the black nodes are the $\componentsx{G\setminus e}$.
+The node labeled `Core+$A_1$+$A_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially.
+    %
+    % Figure (\ref{fig:seq_proof:b}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f\in L_2$ (from Figure (\ref{fig:seq_proof:a})).
+Similarly, the black nodes are the $\componentsx{G\setminus f}$.
+These leaves are labeled with the cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}.
+\todo{emphasize in beginning.}
+    %
+    % Figure (\ref{fig:seq_proof:c}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the core in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
+Note that based on the choice of $e$ as the first edge, we get a \pseudosequential\ DR-plan that looks like Figure (\ref{fig:seq_proof:c}), if we had chosen an edge from appendage $A_2$ the second and third level would be flipped.
+    %
+    % Figure (\ref{fig:seq_proof:d}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the appendage $A_1$ in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
+Note that this contains a node $D$ that has Case 3.
+With this $D$, $D\cup L_1$ is `Core+$A_1$+$A_2$' and $D\cap L_1$ is `Core'.
+The leaves labeled $L_1'$ in Figure (\ref{fig:seq_proof:a}) are the leaves that will allow us to find the node on the path from $G$ to $e$ that is their sibling (the parent of $e$).
+    %
+    % Take edge $a\in G$, the $\componentsx{G\setminus a}$ is the set of nodes marked with an `x' in figure xx.
+The entire figure shows $\branchx{G,a,P_G}$, where $P_G$ is the \pseudosequential\ DR-plan of $G$, however, we do not yet know the contents of the `circle' nodes or the order of the nodes in the \branch.
+    %
+    % So we take leaf $L$ (an element of $\componentsx{G\setminus a}$) and edge $b\in L$, and find $\components{G\setminus b}$.
+The this set is the leaves of the $\branchx{G,b,P_G}$, shown in figure xx.
+    }
+    \label{fig:seq_proof}
+\end{figure*}
+
+\begin{lemma}
+[\Branch\ from \branch\ leaves]
+\label{lem:branch-from-branch-leaves}
+    For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a \pseudosequential\ DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$, i.e.\ from the set of leaves of \branchGePG, by carrying out -- for each leaf $L$ -- one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $L$.
+\end{lemma}
+
+% \begin{proof}
+%     First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $e$ to $G$ in $P_G$.
+Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
+
+%     Now observe that for a leaf $L$ and edge $f\in L$, there exists component $D\in \componentsx{G\setminus f}$ can be classified as one of the following cases.
+
+%     Each case either discovers at least one node $C$ on the path from $e$ to $G$, and a sibling of $D$ in $P_G$, and partitions the leaves of \branchGePG\ into two sets, namely descendants of $C$ or children of ancestors of $C$; or finds a leaf of \branchLfPL\ ($\componentsx{L\setminus f}$ for the next recursion level of the algorithm).
+
+%     % \todo{
+%     % Case (0):
+%     % $D$ is exactly some component $C$ of $\componentsx{G\setminus e}$....
+The only true 1a is the leftmost in subfig 2, everything else is 0...
+In this case, we know $C=D$ to be some child of an ancestor of $C$.
+%     % }
+
+%     Case (1):
+%     $D$ contains some components $C$ of $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) and intersects all other leaves of \branchGePG\ trivially.
+One occurrence of this case (Case 1a) is when $L$'s siblings in (any \pseudosequential\ DR-plan) $P_G$ have trivial pairwise intersection.
+Another occurrence of this case (Case 1b) is when $L$'s siblings in (any \pseudosequential\ DR-plan) $P_G$ have non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core \todo{mentions core!}.
+In both cases, $D$ is a node along the path from $e$ to $G$ in $P_G$, with the components $C$ being its descendants, $L$ being its sibling and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
+%     \todo{(1c), it contains no components.}
+
+%     Case (2):
+%     $D$ is contained in $L$ and intersects all other leaves of \branchGePG\ trivially. In this case, $D$ is a leaf of \branchLfPL, to be used in the next level recursion of the algorithm.
+
+%     Case (3):
+%     $D$ intersects $L$ non-trivially, neither contains nor is contained in $L$, but contains some components $C$ of $\componentsx{G\setminus e}$ (leaves of \branchGePG).
+This happens when the siblings of $L$ in a canonical DR-plan for $G$ intersect $L$ non-trivially.
+In this case, $D\cup L$ is a node along the path from $e$ to $G$ with the components $C$ being its descendants, $L$ being its  child and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
+In addition, $D\cap L$ is a leaf of \branchLfPL\ to be used in the next recursion level of the algorithm.
+
+%     One subtle obstacle to overcome in Case 3 is that the newly found node along the path from $e$ to $G$ is the parent of $L$ as opposed to a sibling of $L$ as in Case (1).
+On the surface, this is problematic because the sibling $D'$ of $L$ along the path from $e$ to $G$ may never be found.
+However a closer inspection of Case 3 reveals that $L$ is the partner of an appendage $A$ containing $e$, and $A$, being underconstrained by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}, must have maximal rigid components with nontrivial intersections, which means that there must be another sibling of $D'$ that is a leaf $L'$ of \branchGePG.
+Hence for some edge $f'$ in $L'$, $\componentsx{G\setminus f'}$ will find $D'$ within Case 1.
+
+%     Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
+% \end{proof}
+
+\begin{proof}
+    First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $G$ to $e$ in $P_G$.
+Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
+
+    For a leaf $L$ and edge $f\in L$, we can classify the component $D\in \componentsx{G\setminus f}$ as being one of the following cases:
+
+    Case (1):
+    $D$ and $L$ are edge disjoint, and $D$ contains
+    (1a) no elements,
+    (1b) exactly one element, or
+    % originally 1a and 1b
+    (1c) more than one element
+    of $\componentsx{G\setminus e}$.
+    %
+    Case (2):
+    $D$ and $L$ have non-empty intersection of edge sets, and $D$ is
+    % originally 2
+    (2a)
+    a proper subgraph or
+    % originally 3
+    (2b)
+    not a subgraph
+    of $L$.
+    %
+    Note that this list is exhaustive.
+If $D$ and $L$ are edge disjoint, $D$ cannot contain subgraphs of elements of $\componentsx{G\setminus e}$.
+If $D$ and $L$ have non-empty intersection of edge sets, $D$ cannot be exactly $L$.
+    See Figure~\ref{fig:seq_proof} for examples of the cases.
+
+    Now observe that it is impossible for all $D\in \componentsx{G\setminus f}$ to be Case (1b), as this would imply $f=e$.
+
+    Case (1a) implies that $D=e$, and is therefore a sibling of $L$ and an element on the path from $G$ to $e$.
+Furthermore, we can classify the rest of the leaves (i.e.\ $\componentsx{G\setminus e}$) as children of the ancestors of $D$.
+
+    Similarly, Case (1c) implies that $D$ is a sibling of $L$ on the path from $G$ to $e$.
+It also allows us to partition the leaves as either children of $L$ (those components which are subgraphs of $D$) or as children of the ancestors of $D$.
+This case occurs when the siblings of $L$ in (any \pseudosequential\ DR-plan) $P_G$ have either (1c1) trivial pairwise intersection or (1c2) non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core.
+
+    Case (2a) implies that $D$ is a leaf in \branchLfPL\ to be used in the next level of recursion of the algorithm.
+
+    Case (2b) implies that $D$ is the partner to the appendage containing $f$.
+Node $D$ will also contain some other leaves from $\componentsx{G\setminus e}$, namely the siblings of $L$ and their descendants.
+In this case, $D\cup L$ is a node along the path from $G$ to $e$ with the leaves contained in $L$ being its descendants, $L$ being its child, and all other components in $\componentsx{G\setminus e}$ being children of the ancestors of $D\cup L$ in \branchGePG.
+In addition, $D\cap L$ is a leaf of \branchLfPL\ to be used in the next recursion level of the algorithm.
+
+    One subtle obstacle to overcome in Case (2b) is that the newly found node along the path from $G$ to $e$ is the parent of $L$ as opposed to a sibling of $L$ as in Case (1a) and (1c).
+On the surface, this is problematic because the sibling $D'$ of $L$ along the path from $G$ to $e$ may never be found.
+However a closer inspection of Case (2b) reveals that $L$ is the partner of an appendage $A$ containing $e$, and $A$, being underconstrained by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}, must have maximal rigid components with nontrivial intersections, which means that there must be another sibling of $D'$ that is a leaf $L'$ of \branchGePG.
+Hence for some edge $f'$ in $L'$, $\componentsx{G\setminus f'}$ will find $D'$ within Case (1a) or (1c).
+
+    % Since each level of \branchGePG\ will contain some $L$ and $f\in L$ with a $D\in \componentsx{G\setminus f}$ that is either of Case (1a), (1c), or (2b) all nodes in the path from $G$ to $e$ can be found, which completes the proof.
+    Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
+\end{proof}
+
+\begin{theorem}
+[Complexity of the algorithm]
+\label{theorem:algo_complexity}
+    Computing a \pseudosequential\ DR-plan for a graph $G$ has time complexity $O(|V(G)|^3)$.
+\end{theorem}
+
+\begin{proof}
+    The previous two Lemmas and the $O(|V(G)|^2)$ complexity of the pebble game algorithm for computing $\componentsx{G\setminus e}$ show that the procedure for computing \branchGePG\ given $G$ and $e$ takes time $O(M|V(G)|^2)$, using $M$ $\componentsx{}$ computations, where $M$ is the number of leaves $L$ of \branchGePG.
+Furthermore, in the process, the leaves of all the \branches\ \branchLfPL\ have already been computed.
+Recursive computation of a \pseudosequential\ DR-plan $P_G$ as in Observation~\ref{obs:seqplan_rec_from_branches} now proceeds by computing \branches\ \branchLfPL\ for each leaf $L$ of \branchGePG.
+Since each node of $P_G$ appears as the leaf $L$ of a \branch\ exactly once during the above recursive procedure, overall one $O(|V(L)|^2)$ computation of $\componentsx{}$ is carried out for each of the $O(|V(G)|)$ nodes $L$ of $P_G$, resulting in an overall complexity of $O(|V(G)|^3)$.
+\end{proof}
+
+\subsection{The `Essentially Unique' Sequential DR-Plan}
+
+The class of sequential DR-plans is defined as a subset of \pseudosequential\ and satisfies all properties discussed in the previous section.
+Furthermore, we show that a sequential DR-plan can always be obtained from a \pseudosequential\ DR-plan.
+A sequential DR-plan is desirable because it is `essentially' unique, and illustrates that the canonical DR-plan in fact retains the essential uniqueness of a complete DR-plan.
+ % which will be elaborated upon following the definition.
+
+\begin{definition}
+[Sequential DR-plan, core]
 \label{def:seqdrp}
-    A \dfn{pseudo-sequential DR-plan} of $G$ is recursively defined as follows:
-    \begin{enumerate}
-        \item Base case: When $G$ is an edge, the sequential DR-plan for $G$ is $G$ itself.
-        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the sequential DR-plans for $C_i$.
-        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of sequential DR-plans for $C_j\setminus C_i$ (called an \dfn{appendage}), and $C_i$ (called its \dfn{partner}).
-        % If the decomposition of the partner $C_i$ results in Case (2), one of the children will be $C_i\cap C_j$, which we call the \dfn{core}.\footnote{Which can be thought of as partner to appendage $C_i\setminus C_j$.}\todo{Is this okay?
-I don't know how to say this succinctly.}
-    \end{enumerate}
-
-    % DR-plans satisfying Requirements 1,2, and 3 are called \dfn{pseudo-sequential.}
-
-    \dfn{Sequential DR-plans} must satisfy another requirement:
+    A \dfn{sequential DR-plan} is a \pseudosequential\ DR-plan that additionally satisfies the following requirement:
 
@@
 
-    When the graph is independent and the DR-plan is sequential (Property (4) holds), if the parent of $C\cup C_s$ falls under Property (2), while $C\cup C_s$ falls under Property (3), then the lowest descendant $D$ as above is called the \dfn{core} of $C\cup C_s$.
+    % When the graph is independent and the DR-plan is sequential (Property (4) holds), if the parent of $C\cup C_s$ falls under Property (2), while $C\cup C_s$ falls under Property (3), then the lowest descendant $D$ as above is called the \dfn{core} of $C\cup C_s$.
+    Take an independent graph with a sequential DR-plan (i.e.\ Property (4) holds).
+Now, if the parent of $C\cup C_s$ falls under Property (2) of Definition~\ref{def:pseudoseqdrp}, while $C\cup C_s$ falls under Property (3) of Definition~\ref{def:pseudoseqdrp}, then the lowest descendant $D$ as above is called the \dfn{core} of $C\cup C_s$.
 \end{definition}
 
-\begin{figure}\centering
-    \includegraphics[width=\linewidth]{../../img/svg/drp_triv_and_non}
-    \caption{Two sequential DR-plans of the same isostatic input graph. \todo{More caption.}}
-    \label{fig:seqdrp_nonuniqueness}
-\end{figure}
-
-\begin{remark}
-[Essential uniqueness of sequential DR-plans]
-\label{remark:seq_is_essentially_unique}
-    For sequential DR-plans of independent graphs, Property (3) of the above definition appears asymmetric with respect to $i$ and $j$; but in fact, $i$ and $j$ can be switched, using the appendage $C_i\setminus C_j$ instead.
+For sequential DR-plans of independent graphs, Property (3) of Definition~\ref{def:pseudoseqdrp} appears asymmetric with respect to $i$ and $j$; but in fact, $i$ and $j$ can be switched, using the appendage $C_i\setminus C_j$ instead.
 Let $C_1, C_2, \ldots, C_N$ be a complete list of proper vertex-maximal rigid subgraphs of $C$.
@@
 Choosing a particular ordering of the $C_i$, i.e.\ choosing the maximal rigid components of a particular appendage, say $A_1$, and its partner $C_1$ to be the children of $C$ simply pushes down the nodes corresponding to the appendages $A_2, A_3, \ldots, A_N$ to a lower level of the sequential DR-plan and the corresponding partners are created as $C_1\cap C_2, C_1\cap C_2\cap C_3, \ldots$; the last appendage, $A_N$, will always have the core, $\bigcap_i{C_i}$, as its partner.
-Thus, modulo the ordering of appendages, the sequential DR-plan is in fact unique.
-See Figure~\ref{fig:seqdrp_nonuniqueness}.
+Thus, we come to the following conclusion.
 
-    Furthermore, any rigid subgraph appears in at most one node of a sequential DR-plan.
-Also, if an edge $e$ of $G$ belongs in the rigid subgraph at any two nodes $A$ and $B$ of a sequential DR-plan, then either $A\subset B$ or $B\subset A$.
-Since the leaves of sequential DR-plan tree are the $O(|V(G)|)$ edges of the independent input graph, this implies the size of the tree is $O(|V(G)|)$.
+\begin{remark}
+[Essential uniqueness of sequential DR-plans]
+\label{remark:seq_is_essentially_unique}
+    Modulo the ordering of appendages and the corresponding partners, the sequential DR-plan is unique.
 \end{remark}
 
+% \begin{figure}\centering
+%     \includegraphics[width=\linewidth]{../../img/svg/drp_triv_and_non}
+%     \caption{Two sequential DR-plans of the same isostatic input graph. \todo{More caption. Show \pseudosequential\ that isn't sequential.}}
+%     \label{fig:seqdrp_nonuniqueness}
+% \end{figure}
+
+\ClearMyMinHeight
+\SetMyMinHeight{.3}{../../img/svg/pseudoseq_vs_seq_seq}
+\SetMyMinHeight{.3}{../../img/svg/pseudoseq_vs_seq_pseudoseq}
+
+\begin{figure}\centering
+    \begin{subfigure}{0.48\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/pseudoseq_vs_seq_seq}
+        \caption{}\label{fig:pseudoseq_vs_seq:seq}
+    \end{subfigure}%
+    %
+    \hfill
+    \begin{subfigure}{0.48\linewidth}\centering
+        \includegraphics[height=\myMinHeight]{../../img/svg/pseudoseq_vs_seq_pseudoseq}
+        \caption{}\label{fig:pseudoseq_vs_seq:pseudoseq}
+    \end{subfigure}%
+    %
+    % \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
+
+    \caption{Two \pseudosequential\ DR-plans of the same isostatic input graph, further decomposition of the diamonds is omitted.
+Figure (\ref{fig:pseudoseq_vs_seq:seq})) also satisfies Property (4) of Definition~\ref{def:seqdrp} and is therefore a sequential DR-plan as well.
+The core and the appendages are marked in the figure.
+Figure (\ref{fig:pseudoseq_vs_seq:pseudoseq}) uses an alternative decomposition of the partner of appendage 1; since the isostatic vertex-maximal proper subgraphs of this node have non-trivial intersections as well, there is a choice in which appendage to decompose first.
+This choice pushes appendage 2 of the input graph further down the plan and the core is not present.}
+    \label{fig:pseudoseq_vs_seq}
+\end{figure}
+
+For an example of a \pseudosequential\ DR-plan that is not sequential (i.e.\ Property (4) does not hold), see Figure~\ref{fig:pseudoseq_vs_seq}.
+Note that the core is not present in the \pseudosequential\ DR-plan (Figure~\ref{fig:pseudoseq_vs_seq:pseudoseq}) and the order of the appendages cannot be easily changed.
+Nevertheless, we will now
+% Now we wish to
+show that a sequential DR-plan can be found from any \pseudosequential\ DR-plan in linear time (in the size of the input graph), making it a worthwhile tool in practice.
+
 \begin{lemma}
-[Pseudo-sequential to sequential DR-plan]
-    Any pseudo-sequential DR-plan for an independent graph can be converted to a sequential DR-plan (satisfying Property (4)) in time $O(|V(G)|)$.
+[\Pseudosequential\ to sequential DR-plan]
+    Any \pseudosequential\ DR-plan for an independent graph can be converted to a sequential DR-plan (satisfying Property (4)) in time $O(|V(G)|)$.
 \end{lemma}
@@
 \begin{proof}
-    For independent graphs, note that Property (2) and (3) of Definition~\ref{def:seqdrp} automatically imply that the following holds for a pseudo-sequential DR-plan: for a node $G$ that falls under Property (2), there is no child $C$ of $G = C\cup C_s$ for which there is a descendant $D$ with siblings $D_s$ possessing the property that $C\cup C_s \setminus D_s$ is rigid; for a node $G$ that falls under Property (3), such a descendant $D$ exists for a unique child $C$ of the node $G=C\cup C_s$, as well as for unique children of all $D'$ on the path from $C$ to $D$.
+    For independent graphs, note that Property (2) and (3) of Definition~\ref{def:pseudoseqdrp} automatically imply that the following holds for a \pseudosequential\ DR-plan: for a node $G$ that falls under Property (2), there is no child $C$ of $G = C\cup C_s$ for which there is a descendant $D$ with siblings $D_s$ possessing the property that $C\cup C_s \setminus D_s$ is rigid; for a node $G$ that falls under Property (3), such a descendant $D$ exists for a unique child $C$ of the node $G=C\cup C_s$, as well as for unique children of all $D'$ on the path from $C$ to $D$.
 Furthermore, for all such $D'$, it automatically holds that $D'\cup D'_s \setminus D_s$ is rigid.
@@
 
-    Given a pseudo-sequential DR-plan of an independent graph where nodes with Property (3) of Definition~\ref{def:seqdrp} are labeled, we show how to enforce Property (4) of Definition~\ref{def:seqdrp}, for a node $C\cup C_s$ with Property (3), i.e.\ when there is a descendant $D$ with siblings $D_s$ possessing the property that $C\setminus D_s \cup C_s$ is rigid.
+    Given a \pseudosequential\ DR-plan of an independent graph where nodes with Property (3) of Definition~\ref{def:pseudoseqdrp} are labeled, we show how to enforce Property (4) of Definition~\ref{def:seqdrp}, for a node $C\cup C_s$ with Property (3), i.e.\ when there is a descendant $D$ with siblings $D_s$ possessing the property that $C\setminus D_s \cup C_s$ is rigid.
 
@@
 
-\begin{observation}
-[Sequential is optimal]
-    A \todo{psuedo-}sequential DR-plan of an independent graph is optimal.
-    \todo{(Psuedo-)sequential DR-plans of independent graphs are optimal.}
-\end{observation}
-
-\begin{proof}
-    We need to show the statement:
-    (*) the max fan-in (i.e.\ number of children) of any sequential DR-plan for an independent graph $G$ is no larger that of some canonical DR-plan for $G$.
-By Theorem~\ref{theorem:canonical_is_optimal}, this would imply that the sequential DR-plan is optimal.
-
-    Recall that both sequential and canonical DR-plans are defined recursively.
-We will prove the statement (*) by induction on the height of a sequential DR-plan for $G$.
-Induction hypothesis: (*) holds for independent graphs $G$ with sequential DR-plans of height $h$.
-Induction step: a sequential DR-plan of height $h+1$ rooted at a node $G$ consists of sequential DR-plans (of height at most $h$) rooted at its children $C$.
-    The base case (height of 0, i.e.\ single edges) trivially holds.
-
-    It is sufficient to show (a) the children $C$ will exist somewhere in some canonical DR-plan, and (b) $|C|$ is less than or equal to the max fan-in of some canonical DR-plan.
-Given (a), then by the recursive definition of canonical DR-plans, these nodes are the roots of canonical DR-plans, thus the induction hypothesis applies to the nodes in $C$.
-Additionally given (b), the proof of the induction step of (*) is complete.
-
-    Case (1):
-    When the isostatic vertex-maximal proper subgraphs of $G$ have trivial intersections, the children of $G$ in any canonical DR-plan are the same set $C$ of children in any sequential DR-plan.
-Thus both conditions (a) and (b) are immediately satisfied.
-
-    Case (2):
-    When the isostatic vertex-maximal proper subgraphs of $G$ have non-trivial intersections, the children $C$ are the isostatic components of some appendage, $A_i$, and its partner.
-The partner will be a child of $G$ in some canonical DR-plan.
-Furthermore, in any canonical DR-plan there will be some node $N-1$ levels below $G$ (where $N$ is the number of appendages) containing the core plus appendage $A_i$.
-In some canonical DR-plan, the children of this node will be the components of $A_i$ and the core.
-This node has the same fan-in as $G$ in the sequential DR-plan, satisfying (b), and, along with the partner (a child of $G$), shows that some canonical DR-plan has all the nodes in $C$, thereby satisfying (a).
-\end{proof}
-
-% \begin{proof}
-%     Any sequential DR-plan, $\seqdrpx{G}$, has maximum fan-in no greater than that of some canonical DR-plan, $\candrpx{G}$, of the same independent graph $G$.
-By Theorem~\ref{theorem:canonical_is_optimal}, this implies a sequential DR-plan is also optimal.
-Consider node $C$ that appears in both $\seqdrpx{G}$ and $\candrpx{G}$ with isostatic vertex-maximal proper subgraphs (clusters) $C_1, C_2, \ldots, C_N$.
-%     We prove this recursively by showing that node $C$ in $\seqdrpx{G}$ and any descendant of $C$ that is not in $\candrpx{G}$ has fan-in less than or equal to the maximum fan-in of $\candrpx{G}$.
-
-%     Case 1: If the clusters of $C$ have pairwise trivial intersections, then node $C$ has the same set of children and fan-in in both $\seqdrpx{G}$ and $\candrpx{G}$.
-
-%     Case 2: Otherwise, the clusters of $C$ have pairwise non-trivial intersection, and we call the common intersection the core.
-The children of node $C$ in $\seqdrpx{G}$ will be the components of the appendage, $A_1$, and its partner, $T_1$.
-In $\candrpx{G}$, there will be some node $D$ that is $A_1$ union the core; $D$ can have clusters with either trivial or non-trivial intersections.
-
-%     Case 2a: In the trivial case, node $D$ decomposes into the core and the components of $A_1$; thus, the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
-
-%     Case 2b: In the non-trivial case, there is some new core of $D$, so there will be some node $D'$ in $\candrpx{G}$ that is $A_1$ union the core of $D$; this pattern will continue until we eventually find $D^*$ with clusters that intersect trivially.
-Similar to Case (2b), the fan-in of node $C$ in $\seqdrpx{G}$ is equal to the fan-in of $D^*$ in $\candrpx{G}$ and they both have the components of $A_1$ as children, leaving $T_1$ to be shown to be optimally decomposed in $\seqdrpx{G}$.
-
-%     The decomposition of $T_1$ in $\seqdrpx{G}$ will be the components of appendage $A_2$ and its partner $T_2$.
-Similarly, $\candrpx{G}$ will contain core union $A_2$, and so we must now show $T_3$ is decomposed optimally in a sequential DR-plan.
-This continues to the node containing $T_{N-1}$, the core union appendage $A_N$, in $\seqdrpx{G}$, which we know to be a node in $\candrpx{G}$.
-% \end{proof}
-
-\begin{definition}
-[\Branch]
-    Given a sequential DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the leaf containing $e$ to the root containing $G$, together with the children of all the nodes on this path, which are the leaves of \branchGePG.
-\end{definition}
-
-\begin{observation}
-[Sequential DR-plan recursively from \branches]
-\label{obs:seqplan_rec_from_branches}
-    A sequential DR-plan $P_G$ for an isostatic graph $G$ is obtained from \branchGePG\ by recursively attaching to each of its leaves $L$ a sequential DR-plan $P_L$ for $L$.
-\end{observation}
-
-The next two Lemmas are the crux of our $O(|V|^3)$ algorithm to find a sequential DR-plan.
-
-\begin{lemma}
-[\Branch\ leaves from components]
-    Let $G$ be isostatic, $e$ be an edge in $G$, and $\componentsx{G\setminus e}$ be the set of maximal rigid components of $G\setminus e$.
-Then there is a sequential DR-plan $P_G$ for $G$ such that $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (minus $e$ itself).
-\end{lemma}
-
-\begin{proof}
-    Follows from the structure of a sequential DR-plan and the definition of \branchGePG.
-\end{proof}
-
-\ClearMyMinHeight
-\SetMyMinHeight{0.240}{../../img/svg/seq_proof_figs_a}
-\SetMyMinHeight{0.24}{../../img/svg/seq_proof_figs_b}
-\SetMyMinHeight{0.240}{../../img/svg/seq_proof_figs_c}
-\SetMyMinHeight{0.24}{../../img/svg/seq_proof_figs_d}
-
-\begin{figure*}\centering%
-    % \begin{subfigure}[t]{0.27\linewidth}
-    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_a}
-    %     \caption{}\label{fig:seq_proof:a}
-    % \end{subfigure}%
-    % %
-    % \hfill
-    % \begin{subfigure}[t]{0.17\linewidth}
-    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_b}
-    %     \caption{}\label{fig:seq_proof:b}
-    % \end{subfigure}%
-    % %
-    % \hfill
-    % \begin{subfigure}[t]{0.25\linewidth}
-    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_c}
-    %     \caption{}\label{fig:seq_proof:c}
-    % \end{subfigure}%
-    % %
-    % \hfill
-    % \begin{subfigure}[t]{0.23\linewidth}
-    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_d}
-    %     \caption{}\label{fig:seq_proof:d}
-    % \end{subfigure}%
-    \begin{subfigure}{0.240\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
-        \caption{}\label{fig:seq_proof:a}
-    \end{subfigure}%
-    %
-    \hfill
-    \begin{subfigure}{0.24\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
-        \caption{}\label{fig:seq_proof:b}
-    \end{subfigure}%
-    %
-    \begin{subfigure}{0.240\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
-        \caption{}\label{fig:seq_proof:c}
-    \end{subfigure}%
-    %
-    \hfill
-    \begin{subfigure}{0.24\linewidth}\centering
-        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
-        \caption{}\label{fig:seq_proof:d}
-    \end{subfigure}%
-
-    \caption{
-    Figure (\ref{fig:seq_proof:a}) depicts $\branchx{G,e,P_G}$, the black nodes are the $\componentsx{G\setminus e}$.
-The node labeled `Core+A$_1$+A$_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially.
-    %
-    Figure (\ref{fig:seq_proof:b}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f\in L_2$ (from Figure (\ref{fig:seq_proof:a})).
-Similarly, the black nodes are the $\componentsx{G\setminus f}$.
-These leaves are labeled with the cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}.
-    %
-    Figure (\ref{fig:seq_proof:c}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the core in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
-Note that based on the choice of $e$ as the first edge, we get a sequential DR-plan that looks like Figure (\ref{fig:seq_proof:c}), if we had chosen an edge from appendage A$_2$ the second and third level would be flipped.
-    %
-    Figure (\ref{fig:seq_proof:d}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f$ in the appendage A$_1$ in $L_1$ (from Figure (\ref{fig:seq_proof:a})).
-Note that this contains a node $D$ that has Case 3.
-With this $D$, $D\cup L_1$ is `Core+A$_1$+A$_2$' and $D\cap L_1$ is `Core'.
-The leaves labeled L$_1'$ in Figure (\ref{fig:seq_proof:a}) are the leaves that will allow us to find the node on the path from $G$ to $e$ that is their sibling (the parent of $e$).
-    %
-    % Take edge $a\in G$, the $\componentsx{G\setminus a}$ is the set of nodes marked with an `x' in figure xx.
-The entire figure shows $\branchx{G,a,P_G}$, where $P_G$ is the sequential DR-plan of $G$, however, we do not yet know the contents of the `circle' nodes or the order of the nodes in the \branch.
-    %
-    % So we take leaf $L$ (an element of $\componentsx{G\setminus a}$) and edge $b\in L$, and find $\components{G\setminus b}$.
-The this set is the leaves of the $\branchx{G,b,P_G}$, shown in figure xx.
-    }
-    \label{fig:seq_proof}
-\end{figure*}
-
-\begin{lemma}
-[\Branch\ from \branch\ leaves]
-\label{lem:branch-from-branch-leaves}
-    For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a sequential DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$, i.e.\ from the set of leaves of \branchGePG, by carrying out---for each leaf $L$---one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $L$.
-\end{lemma}
-
-\begin{proof}
-    First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $e$ to $G$ in $P_G$.
-Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
-
-    Now observe that for a leaf $L$ and an edge $f\in L$, each component $D\in \componentsx{G\setminus f}$ can be classified as one of the following cases.
-
-    Each case either discovers at least one node $C$ on the path from $e$ to $G$, and a sibling of $D$ in $P_G$, and partitions the leaves of \branchGePG\ into two sets, namely descendants of $C$ or children of ancestors of $C$; or finds a leaf of \branchLfPL\ ($\componentsx{L\setminus f}$ for the next recursion level of the algorithm).
-
-    \todo{
-    Case (0):
-    $D$ is exactly some component $C$ of $\componentsx{G\setminus e}$....
-The only true 1a is the leftmost in subfig 2, everything else is 0...
-In this case, we know $C=D$ to be some child of an ancestor of $C$.
-    }
-
-    Case (1):
-    $D$ contains some components $C$ of $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) and intersects all leaves of \branchGePG\ trivially \todo{besides the components $C$ it contains}.
-One occurrence of this case (Case 1a) is when $L$'s siblings in (any sequential DR-plan) $P_G$ have trivial pairwise intersection.
-Another occurrence of this case (Case 1b) is when $L$'s siblings in (any sequential DR-plan) $P_G$ have non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core.
-In both cases, $D$ is a node along the path from $e$ to $G$ in $P_G$, with the components $C$ being its descendants, $L$ being its sibling and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
-
-    Case (2):
-    $D$ is contained in $L$ and intersects all \todo{other} leaves of \branchGePG\ trivially. In this case, $D$ is a leaf of \branchLfPL, to be used in the next level recursion of the algorithm.
-
-    Case (3):
-    $D$ intersects $L$ non-trivially, neither contains nor is contained in $L$, but contains some components $C$ of $\componentsx{G\setminus e}$ (leaves of \branchGePG).
-This happens when the siblings of $L$ in a canonical DR-plan for $G$ intersect $L$ non-trivially.
-In this case, $D\cup L$ is a node along the path from $e$ to $G$ with the components $C$ being its descendants, $L$ being its  child and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
-In addition, $D\cap L$ is a leaf of \branchLfPL\ to be used in the next recursion level of the algorithm.
-
-    One subtle obstacle to overcome in Case 3 is that the newly found node along the path from $e$ to $G$ is the parent of $L$ as opposed to a sibling of $L$ as in Case (1).
-On the surface, this is problematic because the sibling $D'$ of $L$ along the path from $e$ to $G$ may never be found.
-However a closer inspection of Case 3 reveals that $L$ is the partner of an appendage $A$ containing $e$, and $A$, being underconstrained by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}, must have maximal rigid components with nontrivial intersections, which means that there must be another sibling of $D'$ that is a leaf $L'$ of \branchGePG.
-Hence for some edge $f'$ in $L'$, $\componentsx{G\setminus f'}$ will find $D'$ within Case 1.
-
-    Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
-\end{proof}
-
-\begin{theorem}
-[Complexity of the algorithm]
-\label{theorem:algo_complexity}
-    Computing a sequential DR-plan for a graph $G$ has time complexity $O(|V(G)|^3)$.
-\end{theorem}
-
-\begin{proof}
-    The previous two Lemmas and the $O(|V(G)|^2)$ complexity of the pebble game algorithm for computing $\componentsx{G\setminus e}$ show that the procedure for computing \branchGePG\ given $G$ and $e$ takes time $O(M|V(G)|^2)$, using $M$ $\componentsx{}$ computations, where $M$ is the number of leaves $L$ of \branchGePG.
-Furthermore, in the process, the leaves of all the \branches\ \branchLfPL\ have already been computed.
-Recursive computation of a sequential DR-plan $P_G$ as in Observation~\ref{obs:seqplan_rec_from_branches} now proceeds by computing \branches\ \branchLfPL\ for each leaf $L$ of \branchGePG.
-Since each node of $P_G$ appears as the leaf $L$ of a \branch\ exactly once during the above recursive procedure, overall one $O(|V(L)|^2)$ computation of $\componentsx{}$ is carried out for each of the $O(|V(G)|)$ nodes $L$ of $P_G$, resulting in an overall complexity of $O(|V(G)|^3)$.
-\end{proof}
-
 \subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
@@
 \ClearMyMinHeight
-\SetMyMinHeight{.6}{../../img/svg/new_overconstrained_not_optimal}
-\SetMyMinHeight{.4}{../../img/svg/new_overconstrained_optimal}
+\SetMyMinHeight{.6}{../../img/svg/new2_overconstrained_not_optimal}
+\SetMyMinHeight{.4}{../../img/svg/new2_overconstrained_optimal}
 
@@
     %
+    % \begin{subfigure}{0.5\linewidth}\centering
+    %     \scalebox{1}[.9]{
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/new2_overconstrained_optimal}
+    %     }
+    %     \caption{}\label{fig:overconstrained:optimal}
+    % \end{subfigure}%
+    % %
+    % \hfill
+    % \begin{subfigure}{0.5\linewidth}\centering
+    %     \scalebox{1}[.9]{
+    %     \includegraphics[height=\myMinHeight]{../../img/svg/new2_overconstrained_not_optimal}
+    %     }
+    %     \caption{}\label{fig:overconstrained:not_optimal}
+    % \end{subfigure}%
+    %
     \begin{subfigure}{0.5\linewidth}\centering
-        \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_optimal}
-        }
+        \includegraphics[height=\myMinHeight]{../../img/svg/new2_overconstrained_optimal}
         \caption{}\label{fig:overconstrained:optimal}
@@
     \begin{subfigure}{0.5\linewidth}\centering
-        \scalebox{1}[.9]{
-        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_not_optimal}
-        }
+        \includegraphics[height=\myMinHeight]{../../img/svg/new2_overconstrained_not_optimal}
         \caption{}\label{fig:overconstrained:not_optimal}
@@
     Both figures are canonical and cluster-minimal DR-plans of the same singly overconstrained rigid graph.
-Decomposition of triangles are omitted and dashed lines indicate a decomposition similar to the other nodes on the same level.
+Further decomposition of the bottom level is omitted (indicated by tree edges) and dashed lines indicate a decomposition similar to the other nodes on the same level.
 (\ref{fig:overconstrained:optimal}) is an optimal DR-plan, with a fan-in of 5.

### 2015-08-28 doc/GMP2015/revision_3/s04_recombination.tex
Revision 3, close to finished.

@@
 The main barrier in recombination when given an optimal DR-plan  (of smallest possible size or max fan-in) for a system $S$,  is that the number of children of the root  (resp.
-number of roots of the DR-plan)---and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved---could be arbitrarily large as a function of the size of $S$.
+number of roots of the DR-plan) -- and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved -- could be arbitrarily large as a function of the size of $S$.
 This difficulty can persist even after optimal parametrization of the indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic complexity.
@@
     {\sl Small DR-plan.}
-    We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans.
+    We obtain a  parameterized family of systems $I_{\lambda_F}$ -- one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans.
 Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can potentially be solved or realized easily once the orientation type of the solution is known  (when the DR-plan size is small enough).
@@
 These too result in DR-plans of size 2 or 3, putting $G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b).
-The Criterion (c) is met---for example when $k=1$---because 1-dof Cayley configuration spaces of linkages based on such graphs $G'$ are known to be single intervals when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
+The Criterion (c) is met -- for example when $k=1$ -- because 1-dof Cayley configuration spaces of linkages based on such graphs $G'$ are known to be single intervals when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
 In addition, the bounds of these intervals are of low algebraic complexity.

### 2015-08-28 doc/GMP2015/revision_3/s05_problems.tex
Revision 3, close to finished.

--- 2015-05-17 doc/GMP2015/revision_2/s05_problems.tex
@@
 \subsubsection{Special Classes of Small DR-Plans}
-As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving---essentially by solving a single quadratic at a time.
+As seen in the previous section, 2-trees and tree-decomposable graphs have not only small, but also special DR-plans that permit easy solving -- essentially by solving a single quadratic at a time.
 

### 2015-09-02 doc/GMP2015/revision_3/s01_introduction.tex
The submitted revision 3.

@@
 This allows the generic DR-planning problem to be stated and treated as a combinatorial or (hyper)graph problem as we do in this paper.
-
-We additionally conjecture that versions of \frontier, a previously given, bottom-up, polynomial time method for obtaining DR-plans with various desired properties~\cite{hoffman2001decompositionI, hoffman2001decompositionII,lomonosov2004graph,sitharam2005combinatorial}, in fact generate an optimal DR-plan for non-overconstrained systems.
 
@@
 
-The last Section~\ref{sec:conclusion} concludes the paper, and Section \ref{sec:open} lists open problems and conjectures.
+The last Section~\ref{sec:probs_and_conclusion} concludes the paper, and Section \ref{sec:open} lists open problems and conjectures.
 In particular, we conjecture that the methods of Section~\ref{sec:DRP} extend in fact to a large class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.

### 2015-09-02 doc/GMP2015/revision_3/s03_drp.tex
The submitted revision 3.

@@
 
-In this section and in Section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
+In this section and in Section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ wellconstrained or $(k,l)$-tight).
 Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$.
@@
     For the induction step consider a canonical DR-plan $\candrpx{C}$ rooted at node $C$ with height $t+1$.
-    Notice that $\candrpx{C}$ contains a canonical DR-plan $\candrpx{C_i}$ for the graphs $C_i$ corresponding to each of $C$'s descendent nodes.
+    Notice that $\candrpx{C}$ contains a canonical DR-plan $\candrpx{C_i}$ for the graphs $C_i$ corresponding to each of $C$'s descendant nodes.
     Thus, from the induction hypothesis, we know that the $\candrpx{C_i}$ is optimal for $C_i$.
@@
 
-        \newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
-        \renewcommand{\induceonc}[1]{#1}
-        \newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in [N]\setminus\{#1\}}{R_k}}}
+        % \newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
+        % \renewcommand{\induceonc}[1]{#1}
+        % \newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in [N]\setminus\{#1\}}{R_k}}}
 
@@
 \begin{observation}
-[Sequential is optimal]
+[\Pseudosequential\ is optimal]
     Any \pseudosequential\ DR-plan of an independent graph is optimal.
@@
 [\Branch]
-    Given a \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the leaf containing $e$ to the root containing $G$, together with the children of all the nodes on this path, which are the leaves of \branchGePG.
+    Given a \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the root containing $G$ to the leaf containing $e$, together with the children of all the nodes on this path, which are the leaves of \branchGePG.
 See Figure~\ref{fig:seq_proof} for examples.
@@
 If $D$ and $L$ have non-empty intersection of edge sets, $D$ cannot be exactly $L$.
-    See Figure~\ref{fig:seq_proof} for examples of the cases.
 
@@
     % Since each level of \branchGePG\ will contain some $L$ and $f\in L$ with a $D\in \componentsx{G\setminus f}$ that is either of Case (1a), (1c), or (2b) all nodes in the path from $G$ to $e$ can be found, which completes the proof.
-    Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
+    Since each node along the path from $G$ to $e$ is found by the above procedure, the proof is complete.
+    See Figure~\ref{fig:seq_proof} for examples of the cases.
 \end{proof}

### 2016-04-16 doc/GMP2015/revision_4/appendix_proofs.tex
Prepping for addendum

--- 2015-08-23 doc/GMP2015/revision_3/appendix_proofs.tex
@@
+\newcommand{\faninx}[1]{\textsc{FanIn}(#1)}
+\newcommand{\faninxy}[2]{\textsc{FanIn}_{#1}(#2)}
+\newcommand{\numberofleavesx}[1]{\textsc{NL}(#1)}
+\newcommand{\numberofparametersx}[1]{\textsc{NP}(#1)}
+\newcommand{\numberofchildrenxy}[2]{\textsc{NC}_{#1}(#2)}
+
 \section{Proofs}
@@
     % Then, we can form a valid isostatic DR-plan $T_b$ from $T$.
-In $T_b$, $fanin(b)$is the number of leaves in the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
-Similarly, for any other node $w$ that is an ancestor of $b$, $fanin(w)$ is the the number leaf nodes in the tree rooted at $w$, excluding the subtree constaining $b$, plus 1 (for the node leading to $b$).
+In $T_b$, $\faninx{b}$is the number of leaves in the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic.
+Similarly, for any other node $w$ that is an ancestor of $b$, $\faninx{w}$ is the the number leaf nodes in the tree rooted at $w$, excluding the subtree containing $b$, plus 1 (for the node leading to $b$).
 Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
-The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
+The size of $T_b$ will just be the maximum fan-in of all nodes in $T_b$.
 Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
@@
     For each node $b$, assign a size of the $T_b$ denoted $|T_b|$.
-$|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$.
+$|T_b| = \displaystyle\max_{v \in T_b} \faninx{v}$.
 We are looking for $b$ that minimizes $|T_b|$.
-Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$.
-Note that $fanin(b)= nl(T^b)$ because no descendant of $b$ is isostatic.
-Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$.
+Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $\numberofleavesx{T}$.
+Note that $\faninx{b}= \numberofleavesx{T^b}$ because no descendant of $b$ is isostatic.
+Similarly, for any ancestor $w$ of $b$, $\faninx{w} = \numberofleavesx{T^w}-\numberofleavesx{T^{b'}}+1$, where $b'$ is the child leading to $b$.
 All other nodes are not isostatic and hence do not appear in the isostatic DR-plan.
@@
 Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan.
-The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$.
+The size of $T_b$ will just be the maximum fan-in of all nodes in $T_b$.
 Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.
@@
 Suppose a node $b$ is pinned that has a nontrivial child $v$.
-Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$.
-If we had instead chosen to pin $v$, then $fanin_v(b) = nl(T^b) - nl(T^{b'}) + 1 \leq fanin_b(v)$.
-And for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $|T_v| \leq |T_b|$.
+Then, $\faninxy{b}{b} = \numberofleavesx{T^b} = \numberofleavesx{T^v} + n$, where $n$ is essentially the number of leaves between $b$ and $v$.
+If we had instead chosen to pin $v$, then $\faninxy{v}{b} = \numberofleavesx{T^b} - \numberofleavesx{T^{b'}} + 1 \leq \faninxy{b}{v}$.
+And for each ancestor $w$ of $b$, $\faninx{w}$ is unchanged, meaning $|T_v| \leq |T_b|$.
 Thus we only have to check the deepest non-trivial nodes.
@@
 Then, there must exist a common ancestor $a$ of $v_i$ and $v_j$.
-Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$.
-However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin_v(a) = nl(T^a) - nl(T^{a'}) +1$ .
-Thus $fanin_v(a)' \leq fanin_{v_i,v_j}(a)$.
+Then, in $T_{v_i,v_j}$, $\faninxy{v_i,v_j}{a} = \numberofleavesx{T^a}$.
+However, if we chose to pin one of $v_i$ and $v_j$ twice, then $\faninxy{v}{a} = \numberofleavesx{T^a} - \numberofleavesx{T^{a'}} +1$ .
+Thus $\faninxy{v}{a}' \leq \faninxy{v_i,v_j}{a}$.
 All ancestors of $a$ are unchanged.
@@
 The realization complexity for $v$ is simply that of realizing each of its children.
-In general, the number of parameters for $v$  will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.
+In general, the number of parameters for $v$ will be $\numberofparametersx{v} = 4\numberofchildrenxy{1}{v}+3$, if $v \neq b$ and $\numberofparametersx{b} = 4\numberofchildrenxy{1}{b}$, where $\numberofchildrenxy{k}{v}$ is the number of $k$-dof children of $v$.
 
-    Minimizing the algebraic complexity  requires minimizing the maximum $np(v)$ for any node $v$.
+    Minimizing the algebraic complexity requires minimizing the maximum $\numberofparametersx{v}$ for any node $v$.
 In this case, it is not possible to always choose to pin a node closest to a leaf in the tree, because it could have high fan-in.
-So we  try
-brute force by pinning all nodes to  pick the one with the lowest algebraic complexity. This algorithm is still quadratic for the 1-dof case.
+So we try brute force by pinning all nodes to  pick the one with the lowest algebraic complexity.
+This algorithm is still quadratic for the 1-dof case.
 
     For the 2-dof situation, there are more cases to consider.
-If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$.
+If we pin the same node twice as above, we have $\numberofparametersx{v} = 5\numberofchildrenxy{2}{v}+3$ for any ancestor $v \neq b$ and $\numberofparametersx{b} = 5\numberofchildrenxy{2}{b}$.
 If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof.
 Note that solving or realizing $v'$ will also realize $v$.
-Next, we need to consider nodes above and including $v'$ in our complexity: $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.
+Next, we need to consider nodes above and including $v'$ in our complexity: $\numberofparametersx{v'} = 5\numberofchildrenxy{2}{v'} + 4$ and $\numberofparametersx{a} = 3 + 5\numberofchildrenxy{2}{a}$ for $a$ an ancestor of $v'$.
 
     The only remaining case is pinning two nodes that are incomparable, i.e.\ do not have a descendant/ancestor relationship.
-The only change from the previous case is that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$.
-For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.
+The only change from the previous case is that for the lowest common ancestor of the nodes $v'$, $\numberofparametersx{v'} = 2*4+5\numberofchildrenxy{2}{v'}$.
+For any ancestor $a$ of $v'$, we still have $\numberofparametersx{a} = 3 + 5\numberofchildrenxy{2}{a}$.
 

### 2016-04-16 doc/GMP2015/revision_4/s02_preliminaries.tex
Prepping for addendum

--- 2015-08-28 doc/GMP2015/revision_3/s02_preliminaries.tex
@@
 Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in~\cite{hoffman2001decompositionI}.
+%
+We conjecture (in Section~\ref{sec:open}) that one such decomposition, a version of \frontier~\cite{hoffman2001decompositionI, hoffman2001decompositionII,lomonosov2004graph,sitharam2005combinatorial}, which is a bottom-up, polynomial time method, also generates optimal DR-plans for independent systems.
+% We additionally conjecture in Section~\ref{sec:open} that versions of \frontier, a previously given, bottom-up, polynomial time method for obtaining DR-plans with various desired properties~\cite{hoffman2001decompositionI, hoffman2001decompositionII,lomonosov2004graph,sitharam2005combinatorial}, in fact generate an optimal DR-plan for non-overconstrained systems.
+
 %An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system.

### 2016-04-16 doc/GMP2015/revision_4/s03_drp.tex
Prepping for addendum

--- 2015-09-02 doc/GMP2015/revision_3/s03_drp.tex
@@
 [\Branch\ leaves from components]
+\label{lem:leaves-from-comps}
     Let $G$ be isostatic, $e$ be an edge in $G$, and $\componentsx{G\setminus e}$ be the set of maximal rigid components of $G\setminus e$.
-Then there is a \pseudosequential\ DR-plan $P_G$ for $G$ such that $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (minus $e$ itself).
+Then there is a \pseudosequential\ DR-plan $P_G$ for $G$ such that each element in the set of leaves of \branchGePG\ (except $e$ itself) is a subgraph of exactly one element in $\componentsx{G\setminus e}$.
 \end{lemma}
@@
 \begin{proof}
-    Follows from the structure of a \pseudosequential\ DR-plan and the definition of \branchGePG.
+    By definition, the leaves of \branchGePG\ are pairwise disjoint, as are the elements of $\componentsx{G\setminus e}$.
+Furthermore, $E(G)$ is equal to the union of the leaves of \branchGePG\ is equal to the union of $\componentsx{G\setminus e}$ and $\{ e \}$.
+
+    We show that there is no leaf that has a non-empty intersection with two or more elements of $\componentsx{G\setminus e}$, proving the lemma.
+Assume that such a leaf $L$ exists.
+If there were some component $C$ such that $L\cap C$ were non-trivial, then $L\cup C\neq C$ would be isostatic, contradicting the assumption that $C$ was maximal.
+If, for all components $C$ that $L$ intersects, $C\cap L$ is trivial, then none of the edges in $L$ are contained in $\componentsx{G\setminus e}$, again a contradiction.
 \end{proof}
+
+\begin{observation}
+     If there is no proper isostatic subgraph of $G$ that (i) contains $e$ and at least one component of $G\setminus e$ and (ii) intersects another component non-trivially, then $\componentsx{G\setminus e}\cup \{ e\}$ is exactly the set of leaves of \branchGePG\
+\end{observation}
 
@@
 
+% \begin{lemma}
+% [\Branch\ from \branch\ leaves]
+% \label{lem:branch-from-branch-leaves}
+%     For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a \pseudosequential\ DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$, i.e.\ from the set of leaves of \branchGePG, by carrying out -- for each leaf $L$ -- one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $L$.
+% \end{lemma}
+
 \begin{lemma}
@@
 \label{lem:branch-from-branch-leaves}
-    For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a \pseudosequential\ DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$, i.e.\ from the set of leaves of \branchGePG, by carrying out -- for each leaf $L$ -- one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $L$.
+    For an isostatic graph $G$ containing an edge $e$, the \branchGePG\ for a \pseudosequential\ DR-plan $P_G$ can be constructed from $\componentsx{G\setminus e}$ by carrying out -- for each component $C$ -- one computation of $\componentsx{G\setminus f}$, where $f$ is any edge in $C$.
 \end{lemma}
 
-% \begin{proof}
-%     First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $e$ to $G$ in $P_G$.
+\begin{proof}
+    First, note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $G$ to $e$ in $P_G$.
 Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
 
-%     Now observe that for a leaf $L$ and edge $f\in L$, there exists component $D\in \componentsx{G\setminus f}$ can be classified as one of the following cases.
+    Second, note that by Lemma~\ref{lem:leaves-from-comps}, exactly two cases can arise: either (A) $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (except $e$ itself); or, (B) there are some elements of $\componentsx{G\setminus e}$ that are exactly the union of two or more leaves (except $e$).
+We begin by describing case (A).
 
-%     Each case either discovers at least one node $C$ on the path from $e$ to $G$, and a sibling of $D$ in $P_G$, and partitions the leaves of \branchGePG\ into two sets, namely descendants of $C$ or children of ancestors of $C$; or finds a leaf of \branchLfPL\ ($\componentsx{L\setminus f}$ for the next recursion level of the algorithm).
-
-%     % \todo{
-%     % Case (0):
-%     % $D$ is exactly some component $C$ of $\componentsx{G\setminus e}$....
-The only true 1a is the leftmost in subfig 2, everything else is 0...
-In this case, we know $C=D$ to be some child of an ancestor of $C$.
-%     % }
-
-%     Case (1):
-%     $D$ contains some components $C$ of $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) and intersects all other leaves of \branchGePG\ trivially.
-One occurrence of this case (Case 1a) is when $L$'s siblings in (any \pseudosequential\ DR-plan) $P_G$ have trivial pairwise intersection.
-Another occurrence of this case (Case 1b) is when $L$'s siblings in (any \pseudosequential\ DR-plan) $P_G$ have non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core \todo{mentions core!}.
-In both cases, $D$ is a node along the path from $e$ to $G$ in $P_G$, with the components $C$ being its descendants, $L$ being its sibling and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
-%     \todo{(1c), it contains no components.}
-
-%     Case (2):
-%     $D$ is contained in $L$ and intersects all other leaves of \branchGePG\ trivially. In this case, $D$ is a leaf of \branchLfPL, to be used in the next level recursion of the algorithm.
-
-%     Case (3):
-%     $D$ intersects $L$ non-trivially, neither contains nor is contained in $L$, but contains some components $C$ of $\componentsx{G\setminus e}$ (leaves of \branchGePG).
-This happens when the siblings of $L$ in a canonical DR-plan for $G$ intersect $L$ non-trivially.
-In this case, $D\cup L$ is a node along the path from $e$ to $G$ with the components $C$ being its descendants, $L$ being its  child and all other components in $\componentsx{G\setminus e}$ (i.e.\ leaves of \branchGePG) being children of $D$'s ancestors in \branchGePG.
-In addition, $D\cap L$ is a leaf of \branchLfPL\ to be used in the next recursion level of the algorithm.
-
-%     One subtle obstacle to overcome in Case 3 is that the newly found node along the path from $e$ to $G$ is the parent of $L$ as opposed to a sibling of $L$ as in Case (1).
-On the surface, this is problematic because the sibling $D'$ of $L$ along the path from $e$ to $G$ may never be found.
-However a closer inspection of Case 3 reveals that $L$ is the partner of an appendage $A$ containing $e$, and $A$, being underconstrained by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}, must have maximal rigid components with nontrivial intersections, which means that there must be another sibling of $D'$ that is a leaf $L'$ of \branchGePG.
-Hence for some edge $f'$ in $L'$, $\componentsx{G\setminus f'}$ will find $D'$ within Case 1.
-
-%     Since each node along the path from $e$ to $G$ is found by the above procedure, the proof is complete.
-% \end{proof}
-
-\begin{proof}
-    First note that in order to obtain \branchGePG\ from the set of subgraphs at its leaves, it is sufficient to find the subgraphs at the nodes along the path from $G$ to $e$ in $P_G$.
-Once these non-leaf nodes of the \branch\ are known, the \branch\ leaves can be organized by parent node (and level) thereby obtaining the \branch.
-
-    For a leaf $L$ and edge $f\in L$, we can classify the component $D\in \componentsx{G\setminus f}$ as being one of the following cases:
+    For a component $C$ and edge $f\in C$, we can classify the component $D\in \componentsx{G\setminus f}$ as being one of the following cases:
 
     Case (1):
-    $D$ and $L$ are edge disjoint, and $D$ contains
+    $C$ and $D$ are edge disjoint, and $D$ contains
     (1a) no elements,
-    (1b) exactly one element, or
+    (1b) exactly one element (and is equal to this element), or
     % originally 1a and 1b
-    (1c) more than one element
+    (1c) more than one element (and is equal to the union of these elements)
     of $\componentsx{G\setminus e}$.
@@
     Case (2):
-    $D$ and $L$ have non-empty intersection of edge sets, and $D$ is
+    $C$ and $D$ have non-empty intersection of edge sets, and $D$ is
     % originally 2
-    (2a)
-    a proper subgraph or
+    (2a) a proper subgraph or
     % originally 3
-    (2b)
-    not a subgraph
-    of $L$.
+    (2b) not a subgraph
+    of $C$.
     %
-    Note that this list is exhaustive.
-If $D$ and $L$ are edge disjoint, $D$ cannot contain subgraphs of elements of $\componentsx{G\setminus e}$.
+    Note that this list is exhaustive. If $C$ and $D$ are edge disjoint, $D$ cannot contain a subgraph of an element of $\componentsx{G\setminus e}$.
+    % Suppose there were some other component $C'$ such that $C'\cap D$ was non-empty and did not equal $C'$. If the intersection were non-trivial
+    % intersected f the intersection This would imply a larger isostatic graph
     If $D$ and $L$ have non-empty intersection of edge sets, $D$ cannot be exactly $L$.
+
+    % Now observe that, given case (1), it is implied that $C$ is in fact a leaf in \branchGePG. If it were the union of several leaves, then the $f$ we chose would come from one such leaf
 
@@
 
-    Case (1a) implies that $D=e$, and is therefore a sibling of $L$ and an element on the path from $G$ to $e$.
-Furthermore, we can classify the rest of the leaves (i.e.\ $\componentsx{G\setminus e}$) as children of the ancestors of $D$.
+    Case (1a) implies that $D=e$, and is therefore a sibling of $C$ and an element on the path from $G$ to $e$.
+Furthermore, we can classify the rest of $\componentsx{G\setminus e}$ as components that contain only leaves that are children of the ancestors of $D$.
 
-    Similarly, Case (1c) implies that $D$ is a sibling of $L$ on the path from $G$ to $e$.
-It also allows us to partition the leaves as either children of $L$ (those components which are subgraphs of $D$) or as children of the ancestors of $D$.
-This case occurs when the siblings of $L$ in (any \pseudosequential\ DR-plan) $P_G$ have either (1c1) trivial pairwise intersection or (1c2) non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core.
+    Similarly, Case (1c) implies that $D$ is a sibling of $C$ on the path from $G$ to $e$.
+It also allows us to partition the rest of $\componentsx{G\setminus e}$ as either descendants of $C$ (those components which are subgraphs of $D$) or as children of the ancestors of $D$.
+This case occurs when the siblings of $C$ in (any \pseudosequential\ DR-plan) $P_G$ have either (1c1) trivial pairwise intersection or (1c2) non-trivial pairwise intersections, but the edge $f$ is in their common intersection, i.e.\ the core.
 
-    Case (2a) implies that $D$ is a leaf in \branchLfPL\ to be used in the next level of recursion of the algorithm.
+    Case (2a) implies that $D$ is a component whose constituent leaves are leaves in \branchLfPL\ to be used in the next level of recursion of the algorithm.
+    % Case (2a) implies that $D$ is in $\componentsx{C\setminus f}$, to be used in the next level of recursion of the algorithm.
 
@@
 
+    % !!!!!!!
+    However, if $D$ contains $e$...
+    This implies that $C\cap D$ is a leaf of \branchGePG.
+Now choose a new $f'\in (C\setminus D)$ and repeat the above process (classifying all the $D'\in \componentsx{G\setminus f'}$ based on their intersections with $C'=(C\setminus D)$).
+
     % Since each level of \branchGePG\ will contain some $L$ and $f\in L$ with a $D\in \componentsx{G\setminus f}$ that is either of Case (1a), (1c), or (2b) all nodes in the path from $G$ to $e$ can be found, which completes the proof.
-    Since each node along the path from $G$ to $e$ is found by the above procedure, the proof is complete.
+    Since each node along the path from $G$ to $e$ is found by the above procedure, the proof of case (A) is complete.
     See Figure~\ref{fig:seq_proof} for examples of the cases.
+
+    Case (B).
+We apply the process recursively to the component that is the union of some subset of the leaves.
+This is guaranteed to terminate (at some level of recursion we will be dealing with only case (A)).
+Once we have the DR-plan of the leaf, we rearrange it such that the relevant piece is on the last level.
+Then we merge it into the branch we were trying to build.
+
+    Since either case requires the decomposition of each leaf, we can detect at this time whether or not this is an instance of case (A) or (B).
 \end{proof}

### 2016-04-16 doc/GMP2015/revision_4/s04_recombination.tex
Prepping for addendum

--- 2015-08-28 doc/GMP2015/revision_3/s04_recombination.tex
@@
 This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of the non-edges $F$ is always convex for any given lengths $\delta$ for the edges of $H$ (i.e.\ over all the realizations of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
-This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$.
+This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the Cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$.
 Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces for some parameters $F$} (short: $H$ is \dfn{convexifiable}).
@@
 %     space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
-%     found by a subdivided binary search   of the cartesian realization
+%     found by a subdivided binary search   of the Cartesian realization
 %     space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation

### 2016-04-16 doc/GMP2015/revision_4/s07_pinnedline.tex
Prepping for addendum

--- 2015-08-23 doc/GMP2015/revision_3/s07_pinnedline.tex
@@
-\section{Application: Finding Optimal DR-Plans and Realizations for Cross-Linking Microfibrils }
+\section{Application: Finding Optimal DR-Plans and Realizations for Cross-Linking Microfibrils}
 \label{sec:pinnedline}
@@
 
-Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph, where the grey ovals denote pins representing attachments of fibrils, and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
+Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph, where the gray ovals denote pins representing attachments of fibrils, and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
 % and the black dots stand for vertices/cross-linkings.

### 2016-04-16 doc/GMP2015/revision_4/s08_conclusion.tex
Prepping for addendum

--- 2015-08-04 doc/GMP2015/revision_3/s08_conclusion.tex
@@
-\section{Conclusion}
-\label{sec:conclusion}
-
-We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
-For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, a strong Church-Rosser property.
-This gives an efficient (\ComplexityCanDRP) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
-
-We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise indecomposable systems  at nodes of a DR-plan.
-These results rely on a recent theory of convex Cayley configuration spaces.
-Relationships and reductions between these and previously studied problems were formally clarified.
-
-We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.
+\section{Open Problems and Conclusion}
+\label{sec:probs_and_conclusion}
 
@@
 
+The results of this paper lead to a number of open problems.
 The first set of problems are from Section \ref{sec:DRP}:
@@
 
-The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA, although running in time $O(n^3)$, is a bottom-up algorithm, involving complex datastructures.
+The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA, although running in time $O(n^3)$, is a bottom-up algorithm, involving complex data structures.
 However, a proof of optimality, even if it exists, would not be possible without the new notion of a canonical DR-plan at hand.
-The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the sequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.
+The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the pseudosequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.
 Since it is known \cite{lomonosov2004graph} that the DR-plan generated by MFA is cluster-minimal, an alternate conjecture is the following.
@@
 
-More general problem directions are the following.
-\begin{openproblem}
-    Combinatorial rigidity for periodic structures is an active area of research.
+% More general problem directions are the following.
+% \begin{openproblem}
+%     Combinatorial rigidity for periodic structures is an active area of research.
 This paper motivates a study of the rigidity of self-similar structures, with self-similar groups replacing periodic groups.
-\end{openproblem}
+% \end{openproblem}
 
-\begin{openproblem}
-    The pinned line incidence structures of Section \ref{sec:pinnedline}, for example in the case of collagen microfibrils, whose function is elastic contraction, should be considered congruent under projective transformations.
+% \begin{openproblem}
+%     The pinned line incidence structures of Section \ref{sec:pinnedline}, for example in the case of collagen microfibrils, whose function is elastic contraction, should be considered congruent under projective transformations.
 I.e.\ the projective group should be factored out as a trivial motion in a new project for extending the combinatorial rigidity characterization of such systems.
 (Currently we permit no trivial motions at all).
-\end{openproblem}
+% \end{openproblem}
 
-\begin{openproblem}
-    Experimental validation (either computationally or physical experiments) of predications based on the material model and theory used in this paper.
+% \begin{openproblem}
+%     Experimental validation (either computationally or physical experiments) of predications based on the material model and theory used in this paper.
 This can be done by modeling known materials and putting stresses on them, seeing if the prediction is observed in the real material.
 Or, our theory could be used to design new materials, which can be tested to see if they possess predicted properties.
-\end{openproblem}
+% \end{openproblem}
+
+\subsection{Conclusion}
+We have clarified the main source of complexity for the optimal DR-plan and recombination problems.
+For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, a strong Church-Rosser property.
+This gives an efficient (\ComplexityCanDRP) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.
+
+We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise indecomposable systems  at nodes of a DR-plan.
+These results rely on a recent theory of convex Cayley configuration spaces.
+Relationships and reductions between these and previously studied problems were formally clarified.
+
+We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.
+
+Finally, we detailed a number of open problems that were motivated by the work in this paper.

### 2016-04-16 doc/Talks/GMP2016/GMP2016_25m.tex
New talk for GMP2016.

% --------------------------------------------------------------
%                       Initialize
% --------------------------------------------------------------

\documentclass{mySlides}
% \usepackage{mathHW}

% \usepackage{amsmath}
% % \usepackage{amsthm}
% \usepackage{amssymb}
% % \usepackage{algpseudocode}
% % \usepackage{breqn}

% \usepackage{enumerate}

% \usepackage{color}

\usepackage{graphicx}
\usepackage[compatibility=false]{caption}
% \usepackage{caption}
\usepackage{subcaption}

% \usepackage[backend=biber,style=authoryear]{biblatex}
% \usepackage{biblatex}
% \usepackage[backend=biber]{biblatex}
% \bibliography{references.bib}

% \usepackage{multicol}

% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}

% \def\n{\vskip\baselineskip}
\newcommand{\n}{\vspace{1em}}

\newcommand{\todo}[1]{\alert{#1}}

% \newcommand{\term}[1]{\alert{\textit{#1}}}
\newcommand{\term}[1]{\textit{#1}}

% \newcommand{\drp}[1]{\text{DRP}(#1)}
% \newcommand{\comdrp}[1]{\text{ComDRP}(#1)}
% \newcommand{\optdrp}[1]{\text{OptDRP}(#1)}
% \newcommand{\candrp}[1]{\text{CanDRP}(#1)}
\newcommand{\rvmps}{rvmps}

% \newcommand{\closure}[1]{\left<#1\right>}

% \def\matroid{\mathcal{M}}
\def\RR{\mathbb{R}}

\newtheorem{openproblem}{\sffamily Open Problem}
\newtheorem{conjecture}{\sffamily Conjecture}

% --------------------------------------------------------------
%                          Info
% --------------------------------------------------------------

\title[Optimal\\Decomposition]{Optimal decomposition of geometric constraint graphs with application to materials modeling}

\author[Troy Baker]{Troy Baker\\
\n
\footnotesize Joint work\cite{baker2015optimal} with: Meera Sitharam, Menghan Wang, Joel Willoughby}
% \author[Troy Baker]{Troy Baker}
% \author[Troy Baker]{Troy Baker\footfullcite{baker2015arxiv}}

\institute{University of Florida}

\date{April 12, 2016}

% --------------------------------------------------------------
%                    Beginning of Document
% --------------------------------------------------------------

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

% --------------------------------------------------------------
%                           Body
% --------------------------------------------------------------

\section{Introduction}
\begin{frame}{Geometric Constraint Systems}
    \begin{definition}[Geometric constraint systems]
        Collection of \term{geometric primitives} and \term{constraints} between them.
    \end{definition}

    % talk about bar-joint, body-a.... systems and show pictures.

    % \n

    Can be represented as a polynomial system.

    \pause
    \n

    \begin{definition}[Realization]
        A real solution to this system. Typically, a realization is an entire congruence class, modulo the trivial motions.
    \end{definition}
    Ambient dimension of realizations defined by polynomial system.
\end{frame}

\begin{frame}{Geometric Constraint System Examples}
% This slide does three things,
% (1) defines a few system types we use,
% (2) gives applications that use them, and
% (3) shows a specific realization of a system of each type.

% Naturally suited to recursive decomp, especially those with hierarchical structures. But more general than that
\small
    \begin{columns}
    \begin{column}{0.725\textwidth}
        \begin{definition}[Bar-joint system]
            Primitives: Points.
            Constraints: Distances.
        \end{definition}
        E.g., cross-sections of microtubule structures and organic tissue with hierarchical structure.

        \uncover<2->{
        \begin{definition}[Body-hyperpin system]
            Primitives: Rigid bodies.
            Constraints: Incidences between any number of bodies.
        \end{definition}
        E.g., silica bilayers.
         % with 0-, 1-, and 2-dof.
        }

        \uncover<3->{
        \begin{definition}[Pinned line-incidence system]
            Primitives: Lines.
            Constraints: Incidences with fixed points in the space.
        \end{definition}
        E.g.\ crosslinked microtubules.
        }
    \end{column}

    \begin{column}{0.225\textwidth}
        \includegraphics[width=\linewidth]{../../img/Chlamydomonas_TEM_17}

        \includegraphics[width=\linewidth]{../../img/ligten2.jpg}

        \uncover<2->{
        \includegraphics[width=\linewidth]{../../img/Silica}
        }

        \uncover<3->{
        \includegraphics[width=\linewidth]{../../img/crosslink}
        }
    \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Terminology}
% \footnotesize
    \begin{definition}[Over-constrained system]
        More constraint equations than primitive variables, modulo trivial motions of the ambient space. Generically, this means no realizations.
    \end{definition}
    % E.g., for 2D bar-joint, $|E|>2|V|-3$.

    \pause
    \begin{definition}[Independent system]
        No subsystem is over-constrained. Generically,  there exists some realization.
    \end{definition}

    \pause
    \begin{definition}[Rigid system]
        % Generically, there are finitely many realizations.
        There are finitely many realizations.
    \end{definition}

    \pause
    Can we efficiently find a realization?

    Doubly exponential in the size of the largest subsystem that must be solved simultaneously.

    % Questions:
    % \begin{itemize}
    % \footnotesize
    %     \item Can we classify a system as one of the above?
    %     % \item Given a constraint system, does there exist a realization? If the solution is real.
    %     % \item How many realizations? Finitely many if it's rigid.
    %     \item Can we find it efficiently?
    % \end{itemize}

    % The cost of solving is doubly exponential to the size of the largest subsystem that must be simultaneously solved.
\end{frame}

\begin{frame}{Combinatorics}
    Motivation: recursively decompose a system into independently solvable subsystems.

    \pause
    \n

    To do this we no longer care about the system, but rather the underlying combinatorics.

    \begin{definition}[Constraint graph]
        The underlying hypergraph of a geometric constraint system. Vertices represent the \term{geometric primitives} and hyperedges represent the \term{constraints}.
    \end{definition}

    \pause
    Combinatorial rigidity translates the properties of systems (over-constrained, independent, rigid) into properties of the underlying constraint graph, generically.
\end{frame}

\begin{frame}{Bar-Joint}
    Our results apply in the general setting, but for ease of exposition, we'll be restricting ourselves to bar-joint.

    \begin{definition}[Bar-joint system]
        Primitives: Points.
        Constraints: Distances.
    \end{definition}

    Distance is a binary relation on points, therefore the constraint graph is simply a graph.

    \begin{figure}
        \includegraphics[width=\linewidth]{../../img/svg/barjoint_illustration_simple}
    \end{figure}
\end{frame}

\section{DR-Planning}
\begin{frame}{DR-Planning}
    \begin{definition}[Decomposition-recombination (DR-) plan]
        A \emph{DR-plan} of constraint graph $G$ is a forest where:
        \begin{itemize}
            \item Each node is a rigid subgraph of $G$.
            \item A root node is a maximal rigid subgraph.
            \item An internal node is the union of its children.
            \item A leaf node is a single constraint and the involved primitives.
        \end{itemize}
    \end{definition}

    % \todo{Probably irrelevant, examples, two for the same graph}
    % This is purely combinatorial, dealing with the underlying graph. The actual constraint don't change the output.
\end{frame}

\begin{frame}{Example DR-Plans: $C_2\times C_3$}
    \begin{figure}
        \includegraphics[width=0.9\linewidth]{../../img/svg/drplanning_illustration_a}

        \pause
        \n

        \includegraphics[width=0.9\linewidth]{../../img/svg/drplanning_illustration_b}
    \end{figure}
\end{frame}

\begin{frame}{Optimal DR-Plan}
    An \emph{optimal DR-plan} minimizes the maximum fan-in.
    % Corresponds to the largest subsystem that needs to be solved simultaneously.

    \pause
    \n

    In general, finding optimal is NP-hard.

    % If the graph is not independent, finding an optimal DR-plan is NP-hard in most cases, \emph{even} if rigidity is captured by an underlying sparsity condition (as in 2D bar-joint.)
    % \cite{lomonosov2004thesis}
    % % \footfullcite{lomonosov2004graph}

    \begin{figure}\centering
        \begin{subfigure}{.35\linewidth}\centering
            \includegraphics[width=\linewidth]{../../img/svg/new_overconstrained_optimal}
        \end{subfigure}
        %
        \hspace{.1\linewidth}
        \begin{subfigure}{.35\linewidth}\centering
            \includegraphics[width=\linewidth]{../../img/svg/new_overconstrained_not_optimal}
        \end{subfigure}
    \end{figure}
    % \begin{figure}\centering
    %     \includegraphics[width=.5\linewidth]{../../img/svg/3xc2c3_candrp_full}
    % \end{figure}
\end{frame}

\begin{frame}{Uses of DR-Planning}
    \begin{itemize}
    \setlength\itemsep{1em}
        \item Determining complexity of solving.
        \item Decomposition of the stress and flex matrices.
        \item Under-constrained completion.
        \item Interactive removal of over-constraints.
    \end{itemize}

    % Determining complexity of solving.

    % \n

    % Under-constrained completion.

    % \n

    % Interactive removal of over-constraints.

    % \n

    % Decomposition of the stress and flex matrices.

    % ------------

    % \begin{itemize}
    %     \item Size of max fan-in is the size of the largest subsystem that has to be simultaneously solved, thus determining the complexity of solving.
    %     \item Figuring out how to complete under-constrained systems.
    %     \item Offering choices to remove over-constraints to get well-constrained.
    %     \item For those of you familiar, corresponds to a decomposition of the underlying rigidity matrix as well, giving decompositions of the flexes and self-stresses.
    % \end{itemize}
\end{frame}

\begin{frame}{History}
    2001: Formalized by Hoffman, Lomonosov, and Sitharam.~\cite{hoffman2001decompositionI}

    \n

    Late 1980's: Began with triangle-decomposable graphs. Corresponds to systems that can be triangularized and therefore have quadratic radical solutions (QRS).

    \n

    1990's-2000's: Older algorithms were \emph{bottom-up} and were based on maximum matching. E.g., Frontier. Made plans with some property other than optimality.

    \n

    2015: Our paper~\cite{baker2015optimal} contains a top-down $O(|V|^3)$ algorithm with a formal guarantee to find an optimal DR-plan, given some conditions.
\end{frame}

\section{Main Result: Optimal DR-Plan}
\begin{frame}{Summary of Results}
    If the system we're considering\ldots
    \begin{itemize}[<+->]
        \item Has an underlying abstract rigidity matroid $\to$ We can push the conceptual structures through.
        \item Is independent $\to$ We achieve optimality.
        \item Has an underlying sparsity matroid $\to$ We get a polynomial time algorithm.
    \end{itemize}

    \n

    \uncover<3->{
    For 2D bar-joint we have $O(|V|^3)$ time algorithm.
    }
\end{frame}

\begin{frame}{Canonical DR-Plan}
    % We introduce a canonical DR-plan, which is guaranteed to exist and to be optimal for independent graphs.

    \begin{columns}
    \begin{column}{0.495\textwidth}
        \begin{definition}[Canonical DR-plan]
            A DR-plan that satisfies the additional two properties:
            \begin{enumerate}
                \item Children are rigid vertex-maximal proper subgraphs (\rvmps) of the parent.
                \item If all pairs of \rvmps\ intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
            \end{enumerate}
        \end{definition}

    \end{column}
    \begin{column}{0.475\textwidth}
        \begin{figure}\centering
            \includegraphics[width=\linewidth]{../../img/svg/3xc2c3_candrp_full}
            % \includegraphics[width=\linewidth]{../../img/svg/3xc2c3_new_candrp}
        \end{figure}

    \end{column}
    \end{columns}

\end{frame}

\begin{frame}{Importance of Canonical DR-Plan}
    We restrict the space of DR-plans for the input to this special class of canonical DR-plans.

    \n

    \begin{theorem}
        A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.$^*$
    \end{theorem}
    $^*$Applies when $G$ has an underlying abstract rigidity matroid.

    \n

    Proof is non-trivial.
\end{frame}

\begin{frame}{Algorithmic Result}
    \begin{theorem}
        Computing an optimal DR-plan for an independent graph $G$ has time complexity $O(|V|^3)$.$^*$
    \end{theorem}
    $^*$Provided there exists underlying sparsity matroid.

    % \n

    % As a byproduct, this also finds the minimum and maximum (non-trivial) rigid subgraphs of the input, which is NP-hard in general.

    \pause
    \n

    Proof outline:
    \begin{enumerate}
        \item We define a new class of DR-plans.
        \item We show it has fan-in no larger than a canonical DR-plan. (Non-trivial proof.)
        \item We show how to build it in time $O(|V|^3)$.
    \end{enumerate}
\end{frame}

\begin{frame}{Pseudosequential DR-plan}
    \begin{definition}[Pseudosequential DR-plan]
        A DR-plan where, if all pairs of \rvmps\ of a node intersect
        \begin{itemize}
            \item Trivially: then all of them are children.
            \item Non-trivially: then exactly two that intersect non-trivially, $C_1$ and $C_2$, are used to find the children; they are $C_1$ and the pseudosequential DR-plan of $C_2\setminus C_1$.
        \end{itemize}
    \end{definition}
\end{frame}

\begin{frame}{Example DR-Plans: Canonical vs. Pseudosequential}
    \begin{tabular}{c | c}
        Canonical & Psuedosequential \\
        \includegraphics[width=.475\linewidth]{../../img/svg/3xc2c3_candrp_full} & \includegraphics[width=.475\linewidth]{../../img/svg/3xc2c3_seqdrp_full}
    \end{tabular}
\end{frame}

\begin{frame}{Branches}
    \begin{definition}[Branch]
        Branch$(T,a,b)$ of tree $T$ is every node on the path from $a$ to $b$ and their children.
    \end{definition}

    % The leaves of the branch is exactly$^*$ the set of \rvmps\ of $a\setminus b$.

    % \n

    % The \rvmps\ can be found in time $O(|V|^2)$.

    \begin{figure}\centering
        \includegraphics[width=.475\linewidth]{../../img/svg/branch_illustration}
    \end{figure}
\end{frame}

\begin{frame}{A Pseudosequential DR-Plan Branch from the Leaves}
    Given $G$ and $e\in G$, there exists a pseudosequential DR-plan $P_G$ where the leaves of branch$(P_G,G,e)$ is exactly$^*$ the \rvmps\ of $G\setminus e$.

    \pause
    \n

    We can find the \rvmps\ of $G\setminus e$ in $O(|V|^2)$.

    \pause
    \n

    Given a preprocessing step of finding the \rvmps\ of $G\setminus f$ for all $f$, branch$(P_G,G,e)$ can be built in time $O(|V|^2)$.

    \pause
    \n

    Building the branch (from $G$ to $e$):
    \begin{enumerate}
        \item Compute the \rvmps\ of $G\setminus e$, $\{L_i\}$.
        \item For each $L\in \{L_i\}$
        \begin{enumerate}
            \item Choose an arbitrary edge $f\in L$ and compute the \rvmps\ of $G\setminus f$, $\{M_i\}$.
            \item Compare the intersection of $L$ with each $M$ to get its position relative to the other leaves.
        \end{enumerate}
        \item Compute nodes on the path from $G$ to $e$.
    \end{enumerate}
\end{frame}

\begin{frame}{Finding an Entire Pseudosequential DR-Plan}
    Building the DR-plan (of $G$):
    \begin{enumerate}
        \item Preprocessing: Compute the \rvmps\ of $G\setminus e$, for all $e$.
        \item Start with $G$ as the single node in the DR-plan.
        \item Recursively compute a branch for each leaf in the DR-plan.
    \end{enumerate}
\end{frame}

% \begin{frame}{The Algorithm}
%     We can find the pseudosequential DR-plan by computing the \rvmps\ of $G\setminus e$, for all $e$ and doing a quadratic amount of work for each leaf.

%     \n

%     More specifically we:
%     \begin{enumerate}
%         \item Basis step: Start with $G$ as the single node in the DR-plan.
%         \item Recursive step: Compute a branch for each leaf in the DR-plan (that isn't a single edge.)
%         \begin{enumerate}
%             \item Choose arbitrary leaf $L$ and edge $f\in L$.
%             \item Compute the \rvmps\ of $L\setminus f$.
%             \item For each \rvmps, choose an arbitrary edge $g$ and compute the \rvmps\ of $L\setminus g$.
%             \item Compute the branch from $L$ to $f$ (uses a linear number of graph intersections to position each leaf.)
%         \end{enumerate}
%         % \begin{enumerate}
%         %     \item For each leaf $L$ and some edge $f\in L$, compute \rvmps\ of $G\setminus f$.
%         %      % (\rvmps\ of $L\setminus f$ are a subset.)
%         %     \item Compute the branch from $L$ to $f$ (uses a linear number of graph intersections.)
%         % \end{enumerate}
%     \end{enumerate}
% \end{frame}

\begin{frame}{Algorithm Demonstration}
    \begin{figure}
        \includegraphics<1>[width=0.08\linewidth]{../../img/svg/algo_illustration_a}
        \includegraphics<2>[width=0.75\linewidth]{../../img/svg/algo_illustration_b}
        \includegraphics<3>[width=0.75\linewidth]{../../img/svg/algo_illustration_c}
        \includegraphics<4>[width=0.75\linewidth]{../../img/svg/algo_illustration_d}
        \includegraphics<5>[width=0.75\linewidth]{../../img/svg/algo_illustration_e}
        \includegraphics<6>[width=0.75\linewidth]{../../img/svg/algo_illustration_f}
    \end{figure}
\end{frame}

\begin{frame}{Finding an Entire Pseudosequential DR-Plan}
    Building the DR-plan (of $G$):
    \begin{enumerate}
        \item Preprocessing: Compute the \rvmps\ of $G\setminus e$, for all $e$.
        \item Start with $G$ as the single node in the DR-plan.
        \item Recursively compute a branch for each leaf in the DR-plan.
    \end{enumerate}
\end{frame}

\begin{frame}{Open Source Software}
    Under development.
    % \footfullcite{baker2015arxiv}

    Current version available at: \url{cise.ufl.edu/~tbaker/drp}
    \begin{figure}\centering
        \includegraphics[width=0.65\linewidth]{../../img/screenshots/node_in_drp}
    \end{figure}
\end{frame}

\section{Main Result: OMD}
\begin{frame}{What Next?}
    What comes after optimal DR-planning? We've decomposed to the extent possible.

    \n

    Na{\"i}vely, we would, bottom-up, recombine the solved children into parents.

    % \n

    % Recombining each node is equivalent to solving an indecomposable system.

    % Subsystems of any given node have solutions. We are trying to recombine these solved subsystems to get the parent.

    % \n

    % This is equivalent to solving an indecomposable system. Na{\"i}vely we would just start solving.
\end{frame}

\begin{frame}{OMD}
\small
    \begin{definition}[Optimal modification for decomposition (OMD)]
        % OMD$(G,k,s)$ asks if there exist $k$ edges to drop from $G$ and some set of edges to add to $G$, such that the new graph has a DR-plan of max fan-in $s$.
        Informally, can we drop some edges and add some others to make an easily realizable system (i.e.\ small max fan-in DR-plan)?
    \end{definition}

    % Added edges parameterize the realization space, which we search for a realization that satisfies the dropped constraints.

    % \n

    Dropped edges: Chosen so that the realization space has a convex parameterization.
    % \todo{citation}
    % We know how to find dropped edges so that the realization space has a convex parameterization, which can be efficiently searched.

    \n

    Added edges: Parameterize the realization space. Chosen so that the realization can be efficiently computed (e.g., triangle-decomposable.)

    \n

    We search the realization space for those that satisfy the dropped constraints.

    % We can efficiently compute the realization because of the edges we chose to add.

    \begin{figure}
        \includegraphics[height=0.23\linewidth]{../../img/svg/k33_omd_a}
        \hspace{3em}
        \includegraphics[height=0.23\linewidth]{../../img/svg/k33_omd_b}
        \hspace{3em}
        \includegraphics[height=0.23\linewidth]{../../img/svg/k33_omd_c}
    \end{figure}
\end{frame}

\begin{frame}{Open Problems}
    \begin{openproblem}
        Is there a more efficient algorithm than $O(|V|^3)$ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
    \end{openproblem}

    \begin{conjecture}
    \label{conj:mfaisoptimal:rephrase}
        For independent graphs, \term{cluster-minimal} DR-plans are optimal. In fact, for independent graphs, cluster-minimality and canonical are equivalent properties of a DR-plan.
    \end{conjecture}

    Additional problems can be found in the paper.
\end{frame}

% --------------------------------------------------------------
%                        End of Body
% --------------------------------------------------------------

\begin{frame}[allowframebreaks]{Bibliography}
    \bibliographystyle{plain}
    \bibliography{references}
    % \printbibliography
\end{frame}

% \bibliographystyle{plain}
% \bibliography{references}

\end{document}

% --------------------------------------------------------------
%                       End of Document
% --------------------------------------------------------------
